### 2021-09-17 main.tex
Initial Overleaf Import

\documentclass{article}
\usepackage[]{graphicx}
\usepackage[]{xcolor}
\usepackage{alltt}
\usepackage[left=2.3cm,right=2.8cm, top = 2.2cm, bottom = 3cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\PassOptionsToPackage{hyphens}{url}
\usepackage{url}
\usepackage[disable]{todonotes}
\usepackage{multicol}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage[colorlinks=false]{hyperref} 
\urlstyle{same}
\usepackage{lineno}
\linenumbers

% to handle authorship footnotes as numbers:
\makeatletter
\let\@fnsymbol\@arabic
\makeatother

\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\changed}[1]{#1}

\begin{document}

\title{To log or not to log}
  \author{Anonymous Alpaca\thanks{All Alpaca friends} $^{,}$\thanks{The Zoo} $^{ , *}$}

\maketitle

\begin{abstract}
In the abstract, this article is a very good one. 
\end{abstract}

\bigskip

{\footnotesize $^*$ Correspondence to: Anonymous Alpaca (\url{anonymous@alpaca.com}))}

\newpage

% General thought: maybe easier to do with WIS rather than CRPS?

\section{Introduction}

\paragraph{Context about scoring}

\begin{itemize}
    \item what is a probabilistic forecast and a proper scoring rule, what are scores used for
    \item CRPS generalisation of absolute error
    \item idea of scores on logged data: interpretation as a measure of relative improvement, just as when you log the absolute error
\end{itemize}

\paragraph{What's there in the literature}

\paragraph{Current problems / questions about scoring an epidemiological setting}
%probably narrow down to the ones we really want to discuss%
There are limitations with what we can do with scores on a natural scale and open questions whether we can do these things on a log-scale and also whether a log-scale might be inherently more appropriate
\begin{itemize}
    \item can't really compare quantities on different orders of magnitudes
    \item Can't really model score on a natural scale as errors are heavily skewed. Is there a way to model scores to get insight about different factors that systematically influence scores
    \item Maybe epidemiological forecasts are inherently better suited to be scored on an absolute scale due to the multiplicative nature of processes (related to over-prediction > under-prediction)? Does the log-scale imply we are scoring the growth rate? 
    \item unclear what trade-offs of logging / not logging are
    \item what score (log, not log) matches closest what our intuition (or policymakers) think is good? 
    \item are we better at forecasting deaths than cases? --> relative measure would help
    \item There exists confusion about what can be done to a score in general (e.g. people want to take the median, which they shouldn't) %maybe don't add as an extra point
\end{itemize}

\paragraph{Outline of what we do}

\begin{enumerate}
    \item Effects / Interpretation of scoring on a log scale
    \item Appropriate scales for epidemiological (and possibly other) processes
    \item Modelling log scores
\end{enumerate}

\section{Effects / Interpretation of scoring on a log scale}

\paragraph{How to log and what happens}
\begin{itemize}
    \item take log of (observation or truth + 1) 
    \item discuss how equations change for CRPS, WIS
    \item Optional: What happens to the decomposition of the WIS if we log?
    \item Optional: Discussion of parallels to point forecasts \\
\end{itemize}

\paragraph{What happens to the WIS / CRPS when you log it?}
Discuss Johannes' Figure from the WIS paper and what changes when you go from natural to log \\

\paragraph{Interpretation of scores on log-scale}

\begin{itemize}
    \item We can interpret as a relative measure
    \item scale-free
\end{itemize}

\paragraph{Practical considerations / limitations}

\begin{itemize}
    \item Before: lot of weight to scores of targets that are large
    \item Now: Potentially a lot of weight to things that are small
    \item Relative weight depends on the relation between mean and variance of the thing we analyse. If variance grows quicker / slower than the mean, than we give more / less relative weight to large things
\end{itemize}

\paragraph{Option: empirical analysis} 
Score different things from the European / US Forecast Hub and plot the relationship between mean and variance. Plot log-scale scores for different things and see which of these influence average scores most \\
What happens to the decomposition of the WIS if we log?

\section{Appropriate scales for epidemiological processes}
%Could also make this more general and compare any additive, multiplicative or even other processes and state which scale would be appropriate.

\paragraph{Scoring epidemiological processes}
Epidemiological processes are generally multiplicative, others may be additive or even something else. We want to explore whether epidemiological processes should in principle be scored on a log scale. Also: are there any processes which shouldn't be scored on a log scale?

\paragraph{What is the interpretation of logging in an epidemiological setting and is it appropriate?}

\paragraph{Over- and under-prediction}
If we want to keep this in, we could: 
\begin{itemize}
    \item Check whether there is actually a problem with over-prediction and under-prediction in the Hub. This could be the case because we are most interested in certain scenarios in which this might arise. 
    \item Discuss this in light of Johannes' analysis of how the decomposition of WIS values differs if the data are skewed
    \item compare this to the PIT-value-like relative bias scores we used for the German / Polish paper, which capture a relative tendency to over- or under-predict, rather than absolute penalties. 
\end{itemize}

\paragraph{Toy example: simulate an epidemic and apply 3 different modelse} Compare scores for these three models on the natural scale and the log scale

\paragraph{Optional: Empirical differences when going from natural to log-scale} 
We could score a lot of forecasts of the European Hub and show plots about how scores differs when going from natural to log-scale

\section{Modelling scores}
\paragraph{Motivation} Alternative is to use summary measures or pairwise comparison, which becomes cumbersome for many dimensions.

\paragraph{caveats, practical limitations}
\begin{itemize}
    \item What kind of transformations can we do? Propose random effect models
    \item is there anything we can't do? 
    \item How does logging change the error distribution? On the natural scale, you have huge outliers. What's the appropriate error distribution to use? confidence intervals. Maybe we can only say something about the effects
\end{itemize}

\paragraph{Optional: Application to data from the Euro-Hub}
Also: what, empirically is the distribution of errors?

\section{Other ideas}
\begin{itemize}
    \item Compare different predictive distributions and the score as a function of an outcome.
    \item ...
\end{itemize}

\section{Discussion}

\paragraph{summary of what we did}

\paragraph{context, implications, limitations}

\paragraph{outlook, future work}

\newpage

\end{document}

### 2021-09-24 _targets.Rmd
add Rmd targets workflow

---
title: "Target Markdown"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Target Markdown is a powerful R Markdown interface for reproducible analysis pipelines, and the chapter at https://books.ropensci.org/targets/markdown.html walks through it in detail. This R Markdown report the example from the chapter. Try it out in both interactive and non-interactive modes, either by running the code chunks in different ways or setting the `tar_interactive` chunk option.

# Packages

The example requires several R packages, and `targets` must be version 0.5.0.9000 or above.

```{r, eval = FALSE}
install.packages(c("biglm", "dplyr", "ggplot2", "readr", "targets", "tidyr"))
```

# Setup

If you are using old versions of `targets` (<= 0.7.0) and/or `knitr` (<= 1.33), you will need to load the `targets` package in the R Markdown document in order for Target Markdown code chunks to work.

```{r}
library(targets)
```

Near the top of the document, you may also wish to remove the `_targets_r` directory previously written by non-interactive runs of the report. Otherwise, your pipeline may contain superfluous targets.

```{r}
library(targets)
tar_unscript()
```

# Globals

We first define some global options/functions common to all targets. The function below plots a histogram of ozone concentrations, and our histogram target will need it.

```{targets example-globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c("biglm", "dplyr", "ggplot2", "readr", "tidyr"))
create_plot <- function(data) {
  ggplot(data) +
    geom_histogram(aes(x = Ozone), bins = 12) +
    theme_gray(24)
}
```

# Targets

Our first target borrows the `airquality` dataset built into base R.

```{targets raw-data}
tar_target(raw_data, airquality)
```

Our next targets preprocess the data, make a histogram, and fit a model.

```{targets downstream-targets}
list(
  tar_target(data, raw_data %>% filter(!is.na(Ozone))),
  tar_target(hist, create_plot(data))
)
```

Set the `tar_simple` chunk option to `TRUE` to define a single target with the command in the code chunk. The chunk below only contains `biglm(Ozone ~ Wind + Temp, data)` in the source, but because `tar_simple` is `TRUE`, it is shorthand for `tar_target(name = fit, command = biglm(Ozone ~ Wind + Temp, data))`. All other arguments to `tar_target()` are set to their default values (configurable with `tar_option_set()`).

# Pipeline

If you ran all the `{targets}` chunks in non-interactive mode, then your R scripts are set up to run the pipeline.

```{r}
tar_make()
```

# Output

You can retrieve results from the `_targets/` data store using `tar_read()` or `tar_load()`.

```{r}
tar_read(hist)
```

The `targets` dependency graph helps your readers understand the steps of your pipeline at a high level.

```{r}
tar_visnetwork()
```

At this point, you can go back and run `{targets}` chunks in interactive mode without interfering with the code or data of the non-interactive pipeline.

### 2021-09-24 _targets_r/globals/example-globals.R

@@ -1,0 +1,7 @@
+options(tidyverse.quiet = TRUE)
+tar_option_set(packages = c("biglm", "dplyr", "ggplot2", "readr", "tidyr"))
+create_plot <- function(data) {
+  ggplot(data) +
+    geom_histogram(aes(x = Ozone), bins = 12) +
+    theme_gray(24)
+}

### 2021-09-27 _targets.Rmd
add toy scoring workflow

@@
 ---
-title: "Target Markdown"
+title: "Supplementary information"
 output: html_document
@@
 
-Target Markdown is a powerful R Markdown interface for reproducible analysis pipelines, and the chapter at https://books.ropensci.org/targets/markdown.html walks through it in detail.
-This R Markdown report the example from the chapter.
-Try it out in both interactive and non-interactive modes, either by running the code chunks in different ways or setting the `tar_interactive` chunk option.
-
-# Packages
-
-The example requires several R packages, and `targets` must be version 0.5.0.9000 or above.
-
-```{r, eval = FALSE}
-install.packages(c("biglm", "dplyr", "ggplot2", "readr", "targets", "tidyr"))
-```
-
 # Setup
 
-If you are using old versions of `targets` (<= 0.7.0) and/or `knitr` (<= 1.33), you will need to load the `targets` package in the R Markdown document in order for Target Markdown code chunks to work.
+Set up the workflow pipeline and options. We first load the `targets` package and remove the potentially outdated workflow.
 
@@
 library(targets)
-```
-
-Near the top of the document, you may also wish to remove the `_targets_r` directory previously written by non-interactive runs of the report. Otherwise, your pipeline may contain superfluous targets.
-
-```{r}
-library(targets)
+library(tibble)
+library(tidyr)
+library(dplyr)
+library(scoringutils)
+library(ggplot2)
+library(purrr)
+library(here)
 tar_unscript()
@@
 
-# Globals
+We now define shared global options across our workflow and load R functions from the `R` folder.
 
-We first define some global options/functions common to all targets. The function below plots a histogram of ozone concentrations, and our histogram target will need it.
-
-```{targets example-globals, tar_globals = TRUE}
+```{targets globals, tar_globals = TRUE}
 options(tidyverse.quiet = TRUE)
-tar_option_set(packages = c("biglm", "dplyr", "ggplot2", "readr", "tidyr"))
-create_plot <- function(data) {
-  ggplot(data) +
-    geom_histogram(aes(x = Ozone), bins = 12) +
-    theme_gray(24)
-}
-```
-
-# Targets
-
-Our first target borrows the `airquality` dataset built into base R.
-
-```{targets raw-data}
-tar_target(raw_data, airquality)
-```
-
-Our next targets preprocess the data, make a histogram, and fit a model.
-
-```{targets downstream-targets}
-list(
-  tar_target(data, raw_data %>% filter(!is.na(Ozone))),
-  tar_target(hist, create_plot(data))
+library(purrr)
+library(here)
+functions <- list.files(here("R"), full.names = TRUE)
+walk(functions, source)
+tar_option_set(
+  packages = c("tibble", "tidyr", "dplyr", "scoringutils", "ggplot2", "purrr")
 )
@@
 
-Set the `tar_simple` chunk option to `TRUE` to define a single target with the command in the code chunk.
-The chunk below only contains `biglm(Ozone ~ Wind + Temp, data)` in the source, but because `tar_simple` is `TRUE`, it is shorthand for `tar_target(name = fit, command = biglm(Ozone ~ Wind + Temp, data))`.
-All other arguments to `tar_target()` are set to their default values (configurable with `tar_option_set()`).
+# Toy example
+
+An exponential process with a certain rate r; and 3 forecasting models: one that thinks it’s exponential with rate r-a , one that thinks it’s exponential with rate r+a and one that thinks it’s linear; and compare scoring on linear and log scale for increasing horizons.
+
+```{targets toy_scenarios, tar_simple = TRUE}
+expand_grid(
+  init = c(100),
+  r = c(0, 0.01, 0.1, -0.01, -0.1),
+  r_error = c(0, 0.01, 0.05, 0.1, -0.01, -0.05, -0.1),
+  additive_error = c(0, 1, 10, -1, -10)
+) |> 
+  mutate(id = 1:n())
+```
+
+```{targets toy_simulations, tar_simple = TRUE}
+toy_scenarios |>
+  rowwise() |>
+  mutate(y = list(
+    simulate_exp_poisson(init = init, time = 10, growth = r + r_error,
+                         additive_error = additive_error, sims = 1000)
+    )
+  ) |>
+  unnest(cols = c(y))
+```
+
+```{targets toy_truth, tar_simple = TRUE}
+toy_simulations |>
+  filter(r_error == 0, additive_error == 0, sim == 1) |>
+  select(init, r, time, true_value = obs)
+```
+
+```{targets toy_forecasts, tar_simple = TRUE}
+toy_simulations |>
+  filter(!(r_error == 0 & additive_error == 0)) |>
+  select(init, r, r_error, additive_error, time, sample = sim,
+         prediction = obs) |>
+  left_join(toy_truth, by = c("init", "r", "time"))
+```
+
+```{targets, toy_scores, tar_simple = TRUE}
+toy_forecasts |>
+  eval_forecasts(
+    summarise_by = c("init", "r", "r_error", "additive_error"),
+    metrics = c("crps")
+  )
+```
+
+```{targets, toy_log_scores, tar_simple = TRUE}
+toy_forecasts |>
+  mutate(across(.cols = c("true_value", "prediction"), ~ log(.x))) |>
+  eval_forecasts(
+    summarise_by = c("init", "r", "r_error", "additive_error"),
+    metrics = c("crps")
+  )
+```
 
@@
 
-If you ran all the `{targets}` chunks in non-interactive mode, then your R scripts are set up to run the pipeline.
+If all`{targets}` were run in non-interactive mode, then the pipeline is ready to run using the following,
 
-```{r}
+```{r, eval = FALSE}
 tar_make()
@@
 
-# Output
+The complete pipeline can be visualised using,
 
-You can retrieve results from the `_targets/` data store using `tar_read()` or `tar_load()`.
-
-```{r}
-tar_read(hist)
-```
-
-The `targets` dependency graph helps your readers understand the steps of your pipeline at a high level.
-
-```{r}
+```{r, eval = FALSE}
 tar_visnetwork()
 ```
-
-At this point, you can go back and run `{targets}` chunks in interactive mode without interfering with the code or data of the non-interactive pipeline.

### 2021-09-27 R/toy-example.R

@@ -1,0 +1,19 @@
+#' Simulate Count data using an underlying exponential process and a Poisson
+#' Observation model.
+#' @importFrom tibble tibble
+#' @importFrom purrr map
+#' @importFrom tidyr unnest
+simulate_exp_poisson <- function(init = 100, time = 10, growth = 0,
+                                 additive_error = 0, sims = 1) {
+  y <- init * exp(growth * 1:time)
+  y <- y + additive_error
+  y <- max(0, y)
+  obs <- tibble::tibble(
+    sim = list(1:sims),
+    time = 1:time,
+    y = y,
+    obs = purrr::map(y, ~ rpois(sims, .))
+  )
+  obs <- tidyr::unnest(obs, cols = c(obs, sim))
+  return(obs)
+}

### 2021-09-27 _targets_r/globals/example-globals.R

@@ -1,7 +1,0 @@
-options(tidyverse.quiet = TRUE)
-tar_option_set(packages = c("biglm", "dplyr", "ggplot2", "readr", "tidyr"))
-create_plot <- function(data) {
-  ggplot(data) +
-    geom_histogram(aes(x = Ozone), bins = 12) +
-    theme_gray(24)
-}

### 2021-09-27 _targets_r/globals/globals.R

@@ -1,0 +1,8 @@
+options(tidyverse.quiet = TRUE)
+library(purrr)
+library(here)
+functions <- list.files(here("R"), full.names = TRUE)
+walk(functions, source)
+tar_option_set(
+  packages = c("tibble", "tidyr", "dplyr", "scoringutils", "ggplot2", "purrr")
+)

### 2021-09-27 _targets_r/targets/toy_forecasts.R

@@ -1,0 +1,7 @@
+tar_target(toy_forecasts, {
+  toy_simulations |>
+    filter(!(r_error == 0 & additive_error == 0)) |>
+    select(init, r, r_error, additive_error, time, sample = sim,
+           prediction = obs) |>
+    left_join(toy_truth, by = c("init", "r", "time"))
+})

### 2021-09-27 _targets_r/targets/toy_log_scores.R

@@ -1,0 +1,8 @@
+tar_target(toy_log_scores, {
+  toy_forecasts |>
+    mutate(across(.cols = c("true_value", "prediction"), ~ log(.x))) |>
+    eval_forecasts(
+      summarise_by = c("init", "r", "r_error", "additive_error"),
+      metrics = c("crps")
+    )
+})

### 2021-09-27 _targets_r/targets/toy_scenarios.R

@@ -1,0 +1,9 @@
+tar_target(toy_scenarios, {
+  expand_grid(
+    init = c(100),
+    r = c(0, 0.01, 0.1, -0.01, -0.1),
+    r_error = c(0, 0.01, 0.05, 0.1, -0.01, -0.05, -0.1),
+    additive_error = c(0, 1, 10, -1, -10)
+  ) |> 
+    mutate(id = 1:n())
+})

### 2021-09-27 _targets_r/targets/toy_scores.R

@@ -1,0 +1,7 @@
+tar_target(toy_scores, {
+  toy_forecasts |>
+    eval_forecasts(
+      summarise_by = c("init", "r", "r_error", "additive_error"),
+      metrics = c("crps")
+    )
+})

### 2021-09-27 _targets_r/targets/toy_simulations.R

@@ -1,0 +1,10 @@
+tar_target(toy_simulations, {
+  toy_scenarios |>
+    rowwise() |>
+    mutate(y = list(
+      simulate_exp_poisson(init = init, time = 10, growth = r + r_error,
+                           additive_error = additive_error, sims = 1000)
+      )
+    ) |>
+    unnest(cols = c(y))
+})

### 2021-09-27 _targets_r/targets/toy_truth.R

@@ -1,0 +1,5 @@
+tar_target(toy_truth, {
+  toy_simulations |>
+    filter(r_error == 0, additive_error == 0, sim == 1) |>
+    select(init, r, time, true_value = obs)
+})

### 2021-09-27 _targets.Rmd
add toy example sketch

@@
 title: "Supplementary information"
-output: html_document
+output: github_document
 ---
@@
 
-An exponential process with a certain rate r; and 3 forecasting models: one that thinks it’s exponential with rate r-a , one that thinks it’s exponential with rate r+a and one that thinks it’s linear; and compare scoring on linear and log scale for increasing horizons.
+As a toy example we simulate a grid of data with varying exponential growth rates and a Poisson observation model.
+As example forecasts we use the same simulation model but add error to either the growth rate or the mean of observations.
+We do this using the following steps,
+
+- Define a grid of scenarios including the true values (zero growth rate or cumulative linear error).
 
@@
   r = c(0, 0.01, 0.1, -0.01, -0.1),
-  r_error = c(0, 0.01, 0.05, 0.1, -0.01, -0.05, -0.1),
+  r_error = c(0, 0.01, 0.025, 0.05, 0.1, -0.01, -0.025, -0.05, -0.1),
   additive_error = c(0, 1, 10, -1, -10)
@@
 
+- Simulate each scenario a 1000 times.
+
 ```{targets toy_simulations, tar_simple = TRUE}
@@
   mutate(y = list(
-    simulate_exp_poisson(init = init, time = 10, growth = r + r_error,
+    simulate_exp_poisson(init = init, time = 4, growth = r + r_error,
                          additive_error = additive_error, sims = 1000)
@@
 
+- Filter for the "truth" scenario and keep only the first simulation.
+
 ```{targets toy_truth, tar_simple = TRUE}
@@
 ```
+
+- Filter for all other scenarios.
 
@@
 
+- Score example forecasts and summarise by growth rate, growth rate error, additive error, and forecast horizon.
+
 ```{targets, toy_scores, tar_simple = TRUE}
@@
   eval_forecasts(
-    summarise_by = c("init", "r", "r_error", "additive_error"),
+    summarise_by = c("init", "r", "r_error", "additive_error", "time"),
     metrics = c("crps")
@@
 ```
+
+- Repeat the scoring process but on the log scale.
 
@@
   eval_forecasts(
-    summarise_by = c("init", "r", "r_error", "additive_error"),
+    summarise_by = c("init", "r", "r_error", "additive_error", "time"),
     metrics = c("crps")
+  )
+```
+
+- Join log and natural scale scores.
+
+```{targets, toy_all_scores, tar_simple = TRUE}
+toy_log_scores |>
+  mutate(scale = "log") |>
+  bind_rows(
+    toy_scores |>
+      mutate(scale = "natural")
+  )
+```
+
+- Plot natural scale scores by horizon.
+
+```{targets plot_toy_scores_by_horizon, tar_simple = TRUE}
+toy_scores |>
+  filter(time <= 4) |>
+  plot_scores_by_horizon()
+```
+
+- Plot log scale scores by horizon.
+
+```{targets plot_toy_log_scores_by_horizon, tar_simple = TRUE}
+toy_log_scores |>
+  filter(time <= 4) |>
+  plot_scores_by_horizon()
+```
+
+- Plot log and natural scale scores for the 4 week horizon.
+
+```{targets plot_toy_scores_by_scale, tar_simple = TRUE}
+toy_all_scores |>
+  filter(time == 4) |>
+  plot_scores_by_scale()
+```
+
+- Save output for external use.
+
+```{targets save_toy_output}
+list(
+  tar_target(
+    toy_scores_csv, 
+    save_csv(toy_all_scores, here("output/data/toy_all_scores.csv")),
+    format = "file"
+  ),
+  tar_target(
+    toy_scores_by_scale_file, 
+    save_plot(plot_toy_scores_by_scale, 
+              here("output", "figures", "toy_scores_by_scale.png"),
+    format = "file")
+            ),
+  tar_target(
+    toy_scores_by_horizon_file, 
+    save_plot(plot_toy_scores_by_horizon, 
+              here("output", "figures", "toy_scores_by_horizon.png"),
+    format = "file")
+            ),
+  tar_target(
+    toy_log_scores_by_horizon_file, 
+    save_plot(plot_toy_log_scores_by_horizon, 
+              here("output", "figures", "toy_log_scores_by_horizon.png"),
+    format = "file")
+            )
 )

### 2021-09-27 R/toy-example.R

@@ -5,15 +5,56 @@
 #' @importFrom tidyr unnest
 simulate_exp_poisson <- function(init = 100, time = 10, growth = 0,
                                  additive_error = 0, sims = 1) {
-  y <- init * exp(growth * 1:time)
-  y <- y + additive_error
+  y <- init * exp(cumsum(rep(growth, time)))
+  y <- y + cumsum(rep(additive_error, time))
   y <- max(0, y)
   obs <- tibble::tibble(
     sim = list(1:sims),
     time = 1:time,
     y = y,
-    obs = purrr::map(y, ~ rpois(sims, .))
+    obs = purrr::map(y, function(lambda) {
+      set.seed(runif(1, 0, 100) + lambda)
+      rpois(sims, lambda)
+      }
+    )
   )
   obs <- tidyr::unnest(obs, cols = c(obs, sim))
   return(obs)
+}
+
+#' Plot scores by horizon
+#' @import ggplot2
+#' @importFrom dplyr group_by mutate
+plot_scores_by_horizon <- function(scores) {
+  scores |>
+    group_by(additive_error, r) |>
+    mutate(relative_crps = crps / max(crps)) |>
+    ungroup() |>
+    ggplot() +
+    aes(y = relative_crps, x = time, col = r_error, group = r_error) +
+    geom_point(alpha = 0.8) +
+    geom_line(alpha = 0.8) +
+    theme_bw() +
+    theme(legend.position = "bottom") +
+    facet_grid(vars(additive_error), vars(r))
+}
+
+#' Plot scores by scale
+#' @import ggplot2
+#' @importFrom dplyr group_by mutate
+#' @importFrom tidyr pivot_wider
+plot_scores_by_scale <- function(scores) {
+  scores |>
+    group_by(additive_error, r, scale) |>
+    mutate(relative_crps = crps / max(crps)) |>
+    ungroup() |>
+    select(-crps) |>
+    pivot_wider(names_from = c("scale"), values_from = "relative_crps") |>
+    ggplot() +
+    aes(y = log, x = natural, col = r_error, group = r_error) +
+    geom_abline(intercept = 0, slope = 1, linetype = 2) +
+    geom_point(alpha = 0.8) +
+    theme_bw() +
+    theme(legend.position = "bottom") +
+    facet_grid(vars(additive_error), vars(r))
 }

### 2021-09-27 R/utils.R

@@ -1,0 +1,12 @@
+#' Save a plot and return the ppath for targets
+#' @importFrom ggplot2 ggsave
+save_plot <- function(plot, filename, ...) {
+  ggplot2::ggsave(filename, plot, height = 8, width = 8)
+  return(filename)
+}
+
+#' Save a dataframe to a csv and return the path for targets
+save_csv <- function(dt, path) {
+  readr::write_csv(dt, path)
+  return(path)
+}

### 2021-09-27 _targets_r/targets/plot_toy_log_scores_by_horizon.R

@@ -1,0 +1,5 @@
+tar_target(plot_toy_log_scores_by_horizon, {
+  toy_log_scores |>
+    filter(time <= 4) |>
+    plot_scores_by_horizon()
+})

### 2021-09-27 _targets_r/targets/plot_toy_scores_by_horizon.R

@@ -1,0 +1,5 @@
+tar_target(plot_toy_scores_by_horizon, {
+  toy_scores |>
+    filter(time <= 4) |>
+    plot_scores_by_horizon()
+})

### 2021-09-27 _targets_r/targets/plot_toy_scores_by_scale.R

@@ -1,0 +1,5 @@
+tar_target(plot_toy_scores_by_scale, {
+  toy_all_scores |>
+    filter(time == 4) |>
+    plot_scores_by_scale()
+})

### 2021-09-27 _targets_r/targets/save_toy_output.R

@@ -1,0 +1,25 @@
+list(
+  tar_target(
+    toy_scores_csv, 
+    save_csv(toy_all_scores, here("output/data/toy_all_scores.csv")),
+    format = "file"
+  ),
+  tar_target(
+    toy_scores_by_scale_file, 
+    save_plot(plot_toy_scores_by_scale, 
+              here("output", "figures", "toy_scores_by_scale.png"),
+    format = "file")
+            ),
+  tar_target(
+    toy_scores_by_horizon_file, 
+    save_plot(plot_toy_scores_by_horizon, 
+              here("output", "figures", "toy_scores_by_horizon.png"),
+    format = "file")
+            ),
+  tar_target(
+    toy_log_scores_by_horizon_file, 
+    save_plot(plot_toy_log_scores_by_horizon, 
+              here("output", "figures", "toy_log_scores_by_horizon.png"),
+    format = "file")
+            )
+)

### 2021-09-27 _targets_r/targets/toy_all_scores.R

@@ -1,0 +1,8 @@
+tar_target(toy_all_scores, {
+  toy_log_scores |>
+    mutate(scale = "log") |>
+    bind_rows(
+      toy_scores |>
+        mutate(scale = "natural")
+    )
+})

### 2021-09-27 _targets_r/targets/toy_log_scores.R

@@ -2,7 +2,7 @@ tar_target(toy_log_scores, {
   toy_forecasts |>
     mutate(across(.cols = c("true_value", "prediction"), ~ log(.x))) |>
     eval_forecasts(
-      summarise_by = c("init", "r", "r_error", "additive_error"),
+      summarise_by = c("init", "r", "r_error", "additive_error", "time"),
       metrics = c("crps")
     )
 })

### 2021-09-27 _targets_r/targets/toy_scenarios.R

@@ -2,7 +2,7 @@ tar_target(toy_scenarios, {
   expand_grid(
     init = c(100),
     r = c(0, 0.01, 0.1, -0.01, -0.1),
-    r_error = c(0, 0.01, 0.05, 0.1, -0.01, -0.05, -0.1),
+    r_error = c(0, 0.01, 0.025, 0.05, 0.1, -0.01, -0.025, -0.05, -0.1),
     additive_error = c(0, 1, 10, -1, -10)
   ) |> 
     mutate(id = 1:n())

### 2021-09-27 _targets_r/targets/toy_scores.R

@@ -1,7 +1,7 @@
 tar_target(toy_scores, {
   toy_forecasts |>
     eval_forecasts(
-      summarise_by = c("init", "r", "r_error", "additive_error"),
+      summarise_by = c("init", "r", "r_error", "additive_error", "time"),
       metrics = c("crps")
     )
 })

### 2021-09-27 _targets_r/targets/toy_simulations.R

@@ -2,7 +2,7 @@ tar_target(toy_simulations, {
   toy_scenarios |>
     rowwise() |>
     mutate(y = list(
-      simulate_exp_poisson(init = init, time = 10, growth = r + r_error,
+      simulate_exp_poisson(init = init, time = 4, growth = r + r_error,
                            additive_error = additive_error, sims = 1000)
       )
     ) |>

### 2021-09-28 _targets.Rmd
add oforecast plotting

@@
 library(targets)
+library(tarchetypes)
 library(tibble)
@@
 library(here)
+library(tibble)
+library(tarchetypes)
 functions <- list.files(here("R"), full.names = TRUE)
@@
 tar_option_set(
-  packages = c("tibble", "tidyr", "dplyr", "scoringutils", "ggplot2", "purrr")
+  packages = c("tibble", "tidyr", "dplyr", "scoringutils", "ggplot2", "purrr",
+               "readr"),
+  error = "workspace"
 )
@@
 
-- Define a grid of scenarios including the true values (zero growth rate or cumulative linear error).
+- Define a set of growth rates to explore.
+
+```{targets toy_globals, tar_globals = TRUE}
+toy_growth_rates <- tibble(r = c(0, 0.01, 0.1, -0.01, -0.1))
+```
+
+- Define a grid of scenarios including the true values (zero growth rate and cumulative linear error) for this set of growth rates.
 
@@
   init = c(100),
-  r = c(0, 0.01, 0.1, -0.01, -0.1),
+  r = toy_growth_rates$r,
   r_error = c(0, 0.01, 0.025, 0.05, 0.1, -0.01, -0.025, -0.05, -0.1),
@@
 toy_simulations |>
-  filter(!(r_error == 0 & additive_error == 0)) |>
   select(init, r, r_error, additive_error, time, sample = sim,
@@
   left_join(toy_truth, by = c("init", "r", "time"))
+```
+
+- Plot simulated observations and forecasts for each growth rate stratified by assumed growth rate and additive error.
+
+```{targets plot_toy_forecasts, tar_interactive = FALSE}
+tar_map(
+  toy_growth_rates,
+  tar_target(
+    toy_forecasts_plot,
+    plot_toy_forecasts(toy_forecasts, growth = r[[1]],
+                       alpha = 0.2, samples = 10)
+  ),
+  tar_target(
+    plot_toy_forecasts_file,
+    save_plot(
+      toy_forecasts_plot, 
+      here("output", "figures", 
+           paste0("plot_toy_forecasts_", r[[1]], ".png"))
+    ),
+    format = "file")
+)
 ```
@@
 
-```{targets plot_toy_scores_by_horizon, tar_simple = TRUE}
+```{targets toy_scores_by_horizon_plot, tar_simple = TRUE}
 toy_scores |>
   filter(time <= 4) |>
-  plot_scores_by_horizon()
+  plot_toy_scores_by_horizon()
 ```
@@
 
-```{targets plot_toy_log_scores_by_horizon, tar_simple = TRUE}
+```{targets toy_log_scores_by_horizon_plot, tar_simple = TRUE}
 toy_log_scores |>
   filter(time <= 4) |>
-  plot_scores_by_horizon()
+  plot_toy_scores_by_horizon()
 ```
@@
 
-```{targets plot_toy_scores_by_scale, tar_simple = TRUE}
+```{targets toy_scores_by_scale_plot, tar_simple = TRUE}
 toy_all_scores |>
   filter(time == 4) |>
-  plot_scores_by_scale()
+  plot_toy_scores_by_scale()
 ```
@@
     toy_scores_by_scale_file, 
-    save_plot(plot_toy_scores_by_scale, 
+    save_plot(toy_scores_by_scale_plot, 
               here("output", "figures", "toy_scores_by_scale.png"),
@@
     toy_scores_by_horizon_file, 
-    save_plot(plot_toy_scores_by_horizon, 
+    save_plot(toy_scores_by_horizon_plot, 
               here("output", "figures", "toy_scores_by_horizon.png"),
@@
     toy_log_scores_by_horizon_file, 
-    save_plot(plot_toy_log_scores_by_horizon, 
+    save_plot(toy_log_scores_by_horizon_plot, 
               here("output", "figures", "toy_log_scores_by_horizon.png"),

### 2021-09-28 R/toy-example.R

@@ -22,10 +22,26 @@ simulate_exp_poisson <- function(init = 100, time = 10, growth = 0,
   return(obs)
 }
 
+#' Plot simulated observations and forecasts
+#' @import ggplot2
+#' @importFrom dplyr filter
+plot_toy_forecasts <- function(forecasts, growth = 0.01, samples = 10,
+                               alpha = 0.2) {
+  forecasts |>
+    filter(sample <= samples, r == growth) |>
+    ggplot() +
+    aes(x = time, y = prediction, group = sample) +
+    geom_point(aes(y = true_value), alpha = 0.9, size = 1.1) +
+    geom_line(alpha = alpha) +
+    theme_bw() +
+    theme(legend.position = "bottom") +
+    facet_grid(vars(additive_error), vars(r_error))
+}
+
 #' Plot scores by horizon
 #' @import ggplot2
 #' @importFrom dplyr group_by mutate
-plot_scores_by_horizon <- function(scores) {
+plot_toy_scores_by_horizon <- function(scores) {
   scores |>
     group_by(additive_error, r) |>
     mutate(relative_crps = crps / max(crps)) |>
@@ -43,7 +59,7 @@ plot_scores_by_horizon <- function(scores) {
 #' @import ggplot2
 #' @importFrom dplyr group_by mutate
 #' @importFrom tidyr pivot_wider
-plot_scores_by_scale <- function(scores) {
+plot_toy_scores_by_scale <- function(scores) {
   scores |>
     group_by(additive_error, r, scale) |>
     mutate(relative_crps = crps / max(crps)) |>

### 2021-09-28 _targets.sh

@@ -1,0 +1,7 @@
+#!bin/bash
+
+# render the workflow
+Rscript -e "rmarkdown::render('_targets.Rmd')" 
+
+# run the workflow
+Rscript -e "targets::tar_make()" 

### 2021-09-28 _targets_r/globals/globals.R

@@ -1,8 +1,12 @@
 options(tidyverse.quiet = TRUE)
 library(purrr)
 library(here)
+library(tibble)
+library(tarchetypes)
 functions <- list.files(here("R"), full.names = TRUE)
 walk(functions, source)
 tar_option_set(
-  packages = c("tibble", "tidyr", "dplyr", "scoringutils", "ggplot2", "purrr")
+  packages = c("tibble", "tidyr", "dplyr", "scoringutils", "ggplot2", "purrr",
+               "readr"),
+  error = "workspace"
 )

### 2021-09-28 _targets_r/globals/toy_globals.R

@@ -1,0 +1,1 @@
+toy_growth_rates <- tibble(r = c(0, 0.01, 0.1, -0.01, -0.1))

### 2021-09-28 _targets_r/targets/plot_toy_forecasts.R

@@ -1,0 +1,16 @@
+tar_map(
+  toy_growth_rates,
+  tar_target(
+    toy_forecasts_plot,
+    plot_toy_forecasts(toy_forecasts, growth = r[[1]],
+                       alpha = 0.2, samples = 10)
+  ),
+  tar_target(
+    plot_toy_forecasts_file,
+    save_plot(
+      toy_forecasts_plot, 
+      here("output", "figures", 
+           paste0("plot_toy_forecasts_", r[[1]], ".png"))
+    ),
+    format = "file")
+)

### 2021-09-28 _targets_r/targets/plot_toy_log_scores_by_horizon.R

@@ -1,5 +1,0 @@
-tar_target(plot_toy_log_scores_by_horizon, {
-  toy_log_scores |>
-    filter(time <= 4) |>
-    plot_scores_by_horizon()
-})

### 2021-09-28 _targets_r/targets/plot_toy_scores_by_horizon.R

@@ -1,5 +1,0 @@
-tar_target(plot_toy_scores_by_horizon, {
-  toy_scores |>
-    filter(time <= 4) |>
-    plot_scores_by_horizon()
-})

### 2021-09-28 _targets_r/targets/plot_toy_scores_by_scale.R

@@ -1,5 +1,0 @@
-tar_target(plot_toy_scores_by_scale, {
-  toy_all_scores |>
-    filter(time == 4) |>
-    plot_scores_by_scale()
-})

### 2021-09-28 _targets_r/targets/save_toy_output.R

@@ -6,19 +6,19 @@ list(
   ),
   tar_target(
     toy_scores_by_scale_file, 
-    save_plot(plot_toy_scores_by_scale, 
+    save_plot(toy_scores_by_scale_plot, 
               here("output", "figures", "toy_scores_by_scale.png"),
     format = "file")
             ),
   tar_target(
     toy_scores_by_horizon_file, 
-    save_plot(plot_toy_scores_by_horizon, 
+    save_plot(toy_scores_by_horizon_plot, 
               here("output", "figures", "toy_scores_by_horizon.png"),
     format = "file")
             ),
   tar_target(
     toy_log_scores_by_horizon_file, 
-    save_plot(plot_toy_log_scores_by_horizon, 
+    save_plot(toy_log_scores_by_horizon_plot, 
               here("output", "figures", "toy_log_scores_by_horizon.png"),
     format = "file")
             )

### 2021-09-28 _targets_r/targets/toy_forecasts.R

@@ -1,6 +1,5 @@
 tar_target(toy_forecasts, {
   toy_simulations |>
-    filter(!(r_error == 0 & additive_error == 0)) |>
     select(init, r, r_error, additive_error, time, sample = sim,
            prediction = obs) |>
     left_join(toy_truth, by = c("init", "r", "time"))

### 2021-09-28 _targets_r/targets/toy_log_scores_by_horizon_plot.R

@@ -1,0 +1,5 @@
+tar_target(toy_log_scores_by_horizon_plot, {
+  toy_log_scores |>
+    filter(time <= 4) |>
+    plot_toy_scores_by_horizon()
+})

### 2021-09-28 _targets_r/targets/toy_scenarios.R

@@ -1,7 +1,7 @@
 tar_target(toy_scenarios, {
   expand_grid(
     init = c(100),
-    r = c(0, 0.01, 0.1, -0.01, -0.1),
+    r = toy_growth_rates$r,
     r_error = c(0, 0.01, 0.025, 0.05, 0.1, -0.01, -0.025, -0.05, -0.1),
     additive_error = c(0, 1, 10, -1, -10)
   ) |> 

### 2021-09-28 _targets_r/targets/toy_scores_by_horizon_plot.R

@@ -1,0 +1,5 @@
+tar_target(toy_scores_by_horizon_plot, {
+  toy_scores |>
+    filter(time <= 4) |>
+    plot_toy_scores_by_horizon()
+})

### 2021-09-28 _targets_r/targets/toy_scores_by_scale_plot.R

@@ -1,0 +1,5 @@
+tar_target(toy_scores_by_scale_plot, {
+  toy_all_scores |>
+    filter(time == 4) |>
+    plot_toy_scores_by_scale()
+})

### 2021-09-28 _targets.Rmd
update repro and options

@@
                "readr"),
-  error = "workspace"
+  error = "continue"
 )
@@
 ```{targets toy_globals, tar_globals = TRUE}
-toy_growth_rates <- tibble(r = c(0, 0.01, 0.1, -0.01, -0.1))
+toy_growth_rates <- tibble(r = c(0, 0.1, 0.5, -0.1, -0.5))
 ```
@@
   r = toy_growth_rates$r,
-  r_error = c(0, 0.01, 0.025, 0.05, 0.1, -0.01, -0.025, -0.05, -0.1),
-  additive_error = c(0, 1, 10, -1, -10)
+  r_error = c(0, 0.01, 0.05, 0.1, 0.25, -0.01, -0.05, -0.1, -0.25),
+  additive_error = c(0, 5, 10, -5, -10)
 ) |> 
@@
 
-- Repeat the scoring process but on the log scale.
+- Repeat the scoring process but on the log scale. Here we have to pad by a small number as cannot take the log of zero.
 
@@
 toy_forecasts |>
-  mutate(across(.cols = c("true_value", "prediction"), ~ log(.x))) |>
+  mutate(across(.cols = c("true_value", "prediction"), ~ log(.x + 0.001))) |>
   eval_forecasts(
@@
 
-If all`{targets}` were run in non-interactive mode, then the pipeline is ready to run using the following,
+The pipeline can be regenerated by rendering this file,
+
+```{r, eval = FALSE}
+rmarkdown::render("_targets.Rmd")
+```
+
+Thee pipeline can then be run using,
 
@@
 ```
+
+The pipeline can be regenerated and run using the following single step
+
+```{bash, eval = FALSE}
+. _targets.sh
+```

### 2021-09-28 R/toy-example.R

@@ -1,24 +1,23 @@
 #' Simulate Count data using an underlying exponential process and a Poisson
 #' Observation model.
 #' @importFrom tibble tibble
-#' @importFrom purrr map
+#' @importFrom purrr map map_dbl
 #' @importFrom tidyr unnest
 simulate_exp_poisson <- function(init = 100, time = 10, growth = 0,
                                  additive_error = 0, sims = 1) {
   y <- init * exp(cumsum(rep(growth, time)))
   y <- y + cumsum(rep(additive_error, time))
-  y <- max(0, y)
+  y <- purrr::map_dbl(y, ~ max(., 0))
   obs <- tibble::tibble(
     sim = list(1:sims),
     time = 1:time,
     y = y,
     obs = purrr::map(y, function(lambda) {
-      set.seed(runif(1, 0, 100) + lambda)
-      rpois(sims, lambda)
+        rpois(sims, lambda)
       }
     )
   )
-  obs <- tidyr::unnest(obs, cols = c(obs, sim))
+  obs <- tidyr::unnest(obs, c(obs, sim))
   return(obs)
 }
 

### 2021-09-28 R/utils.R

@@ -1,7 +1,7 @@
 #' Save a plot and return the ppath for targets
 #' @importFrom ggplot2 ggsave
 save_plot <- function(plot, filename, ...) {
-  ggplot2::ggsave(filename, plot, height = 8, width = 8)
+  ggplot2::ggsave(filename, plot, height = 9, width = 9)
   return(filename)
 }
 

### 2021-09-28 _targets_r/globals/globals.R

@@ -8,5 +8,5 @@ walk(functions, source)
 tar_option_set(
   packages = c("tibble", "tidyr", "dplyr", "scoringutils", "ggplot2", "purrr",
                "readr"),
-  error = "workspace"
+  error = "continue"
 )

### 2021-09-28 _targets_r/globals/toy_globals.R

@@ -1,1 +1,1 @@
-toy_growth_rates <- tibble(r = c(0, 0.01, 0.1, -0.01, -0.1))
+toy_growth_rates <- tibble(r = c(0, 0.1, 0.5, -0.1, -0.5))

### 2021-09-28 _targets_r/targets/toy_log_scores.R

@@ -1,6 +1,6 @@
 tar_target(toy_log_scores, {
   toy_forecasts |>
-    mutate(across(.cols = c("true_value", "prediction"), ~ log(.x))) |>
+    mutate(across(.cols = c("true_value", "prediction"), ~ log(.x + 0.001))) |>
     eval_forecasts(
       summarise_by = c("init", "r", "r_error", "additive_error", "time"),
       metrics = c("crps")

### 2021-09-28 _targets_r/targets/toy_scenarios.R

@@ -2,8 +2,8 @@ tar_target(toy_scenarios, {
   expand_grid(
     init = c(100),
     r = toy_growth_rates$r,
-    r_error = c(0, 0.01, 0.025, 0.05, 0.1, -0.01, -0.025, -0.05, -0.1),
-    additive_error = c(0, 1, 10, -1, -10)
+    r_error = c(0, 0.01, 0.05, 0.1, 0.25, -0.01, -0.05, -0.1, -0.25),
+    additive_error = c(0, 5, 10, -5, -10)
   ) |> 
     mutate(id = 1:n())
 })

### 2021-09-28 _targets.Rmd
clean up

@@
 ```{targets toy_globals, tar_globals = TRUE}
-toy_growth_rates <- tibble(r = c(0, 0.1, 0.5, -0.1, -0.5))
+toy_growth_rates <- tibble(r = c(0, 0.1, 0.25, -0.1, -0.25))
 ```
@@
 
-Thee pipeline can then be run using,
+The pipeline can then be run using,
 
@@
 ```
+
+Alternatively the pipeline can be explored interactively using this notebook.

### 2021-09-28 R/toy-example.R

@@ -34,6 +34,10 @@ plot_toy_forecasts <- function(forecasts, growth = 0.01, samples = 10,
     geom_line(alpha = alpha) +
     theme_bw() +
     theme(legend.position = "bottom") +
+    labs(x = "Forecast horizon", y = "Simulated value",
+         caption = "Points represent simulated true values whilst each line
+                    represents a sample forecast with some error. The plot is
+                    stratified by growth rate and additive error.") +
     facet_grid(vars(additive_error), vars(r_error))
 }
 
@@ -47,10 +51,15 @@ plot_toy_scores_by_horizon <- function(scores) {
     ungroup() |>
     ggplot() +
     aes(y = relative_crps, x = time, col = r_error, group = r_error) +
-    geom_point(alpha = 0.8) +
-    geom_line(alpha = 0.8) +
+    geom_point(alpha = 0.6) +
+    geom_line(alpha = 0.4) +
+    scale_color_viridis_c() +
     theme_bw() +
     theme(legend.position = "bottom") +
+    labs(x = "Forecast horizon", y = "CRPS relative to maximum panel value",
+         col = "Growth rate error",
+         caption = "The plot is stratified by true growth rate and additive
+                    error with growth rate error included in each panel.") +
     facet_grid(vars(additive_error), vars(r))
 }
 
@@ -69,7 +78,13 @@ plot_toy_scores_by_scale <- function(scores) {
     aes(y = log, x = natural, col = r_error, group = r_error) +
     geom_abline(intercept = 0, slope = 1, linetype = 2) +
     geom_point(alpha = 0.8) +
+    scale_color_viridis_c() +
     theme_bw() +
     theme(legend.position = "bottom") +
+    labs(x = "Natural scale CRPS relative to maximum panel value",
+         y = "Log scale CRPS relative to maximum panel value",
+         col = "Growth rate error",
+         caption = "The plot is stratified by true growth rate and additive
+                    error with growth rate error included in each panel.") +
     facet_grid(vars(additive_error), vars(r))
 }

### 2021-09-28 _targets_r/globals/toy_globals.R

@@ -1,1 +1,1 @@
-toy_growth_rates <- tibble(r = c(0, 0.1, 0.5, -0.1, -0.5))
+toy_growth_rates <- tibble(r = c(0, 0.1, 0.25, -0.1, -0.25))

### 2021-09-28 _targets.Rmd
add forecast csv

@@
   tar_target(
+    toy_forecasts_csv, 
+    save_csv(toy_forecasts, here("output/data/toy_forecasts.csv")),
+    format = "file"
+  ),
+  tar_target(
     toy_scores_csv, 

### 2021-09-28 _targets_r/targets/save_toy_output.R

@@ -1,5 +1,10 @@
 list(
   tar_target(
+    toy_forecasts_csv, 
+    save_csv(toy_forecasts, here("output/data/toy_forecasts.csv")),
+    format = "file"
+  ),
+  tar_target(
     toy_scores_csv, 
     save_csv(toy_all_scores, here("output/data/toy_all_scores.csv")),
     format = "file"

### 2021-10-23 analysis/Analysis.Rmd
create script with scenario grid

---
title: "Log or not"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE, 
                      warning = FALSE)

library(scoringutils)
library(purrr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(here)

n <- 100
quantiles <- c(0.01, 0.025, seq(0.05, 0.95, by = 0.05), 0.975, 0.99)

```

Define a scenario grid with two data-generating distributions and two corresponding predictive distributions. For the predictive distributions one parameter is set equal to the true parameter, while the other is varied either in a multiplicative or an additive way.

```{r}
truth_mean_equal <- c(FALSE, TRUE)
truth_var_equal <- c(FALSE, TRUE)
error_var <- c("mu", "sigma")
error_type <- c("multiplicative", "additive")

scenarios <- expand_grid(truth_mean_equal, 
                         truth_var_equal, 
                         error_var, 
                         error_type)

grid_sigma_mult <- seq(0.5, 2.5, by = 0.1)
grid_mu_mult <- seq(0.5, 2.5, by = 0.1)
grid_sigma_add <- seq(0, 100, by = 5)
grid_mu_add <- seq(0, 1000, by = 50)

grids <- tibble(error_var = 
                  c("sigma", "mu", "sigma", "mu"),  
                error_type = 
                  c("multiplicative", "multiplicative", 
                    "additive", "additive"), 
                grid = list(grid_sigma_mult, 
                            grid_mu_mult, 
                            grid_sigma_add, 
                            grid_mu_add))

sigma <- 10
mu <- 1000

scenarios <- scenarios |>
  inner_join(grids) |>
  mutate(name = paste0(error_type, "_error_", error_var,  
                       ifelse(equal_mean, "_equal_mean_", "_unequal_mean_"), 
                       ifelse(equal_var, "equal_var", "unequal_var")))

```

```{r}
# helper functions
# ----------------

# function to generate a quantile forecast with a given error
create_forecast <- function(grid, sigma_true, mu_true,
                            var = "sigma", type = "multiplicative") {

  mu <- function(.x, v = var, t = type) {
    if (v == "mu") {
      if (t == "multiplicative") {
        return(mu_true * .x)
      } else if (t == "additive") {
        return(mu_true + .x)
      }
    } else if (var == "sigma") {
      return(mu_true)
    }
  }

  sigma <- function(.x, v = var, t = type) {
    if (var == "sigma") {
      if (t == "multiplicative") {
        return(sigma_true * .x)
      } else if (t == "additive") {
        return(sigma_true + .x)
      }
    } else if (var == "mu") {
      return(sigma_true)
    }
  }

  map_dfr(.x = grid, 
      .f = function(.x) {
        data.frame(prediction = 
                     qnorm(p = quantiles, 
                           mean = mu(.x), 
                           sd = sigma(.x)), 
                   quantile = quantiles, 
                   sigma_hat = sigma(.x), 
                   sigma_true = sigma_true,
                   mu_hat = mu(.x), 
                   mu_true = mu_true,
                   error = .x, 
                   error_type = type, 
                   error_var = var)
      }
  ) |>
  rowwise() |>
  mutate(n = list(1:n)) |>
  ungroup() |>
  unnest(cols = c(n))
}

# helper function to log predictions and true values
create_log_data <- function(data) {
  data |>
  mutate(prediction = log(prediction + 1), 
         true_value = log(true_value + 1))
}

# helper function to score forecasts
score_forecasts <- function(data) {
  eval_forecasts(data = data, 
                 metrics = c("interval_score"),
                 summarise_by = c("mu_true", "sigma_true", "error", 
                                  "error_var", "error_type")) 
}

# helper function to join the score data 
join_scores <- function(scores_1, scores_2, scale = "natural") {
  full_join(scores_1 |>
             select(error, sigma_true, mu_true, error_var, error_type, interval_score), 
           scores_2 |>
             select(error, sigma_true, mu_true, error_var, error_type, interval_score)
           ) |>
    mutate(scale = scale)
}

# helper function to plot scores
plot_scores <- function(scores) {
  scores |>
    mutate("True parameters" = 
             paste0("mean: ", mu_true, ", sd: ", sigma_true)) |>
    ggplot(aes(x = error, y = interval_score, 
               colour = `True parameters`, group = `True parameters`)) + 
    geom_line() + 
    scale_color_brewer(palette = "Set1", name = "True parameters: ") + 
    facet_wrap(~ scale, scale = "free") + 
    theme_minimal() + 
    theme(legend.position = "bottom") 
}

```

```{r}
i = 1
equal_mean <- scenarios[i, ]$truth_mean_equal
equal_var <- scenarios[i, ]$truth_var_equal
grid <- unlist(scenarios[i, ]$grid)
error_var <- scenarios[i, ]$error_var
error_type <- scenarios[i, ]$error_type
scenario_name <- scenarios[i, ]$name

sigma_true <- c(sigma, sigma + (as.numeric(!equal_var) * 10))
mu_true <- c(mu, mu + as.numeric(!equal_mean) * 1000)

true_values_1 <- 
  data.frame(
    true_value = rnorm(n = n, mean = mu_true[1], sd = sigma_true[1]),
    n = 1:n,
    sigma_true = sigma_true[1]
  )

true_values_2 <- 
  data.frame(
    true_value = rnorm(n = n, mean = mu_true[2], sd = sigma_true[2]),
    n = 1:n,
    sigma_true = sigma_true[2]
  )

forecasts_1 <- create_forecast(grid, 
                              sigma_true = sigma_true[1],
                              mu_true = mu_true[1], 
                              var = error_var, 
                              type = error_type)

forecasts_2 <- create_forecast(grid, 
                              sigma_true = sigma_true[2],
                              mu_true = mu_true[2], 
                              var = error_var, 
                              type = error_type)

data_1 <- inner_join(forecasts_1, true_values_1)
data_2 <- inner_join(forecasts_2, true_values_2)
log_data_1 <- create_log_data(data_1)
log_data_2 <- create_log_data(data_2)

scores_1 <- score_forecasts(data_1)
scores_2 <- score_forecasts(data_2)
log_scores_1 <- score_forecasts(log_data_1)
log_scores_2 <- score_forecasts(log_data_2)

joined_scores <- join_scores(scores_1, scores_2)
joined_log_scores <- join_scores(log_scores_1, log_scores_2, scale = "log")
combined_scores <- rbind(joined_scores, joined_log_scores)

title <- paste(error_type, "error of", error_var, "for two variables with", 
               ifelse(equal_mean, "equal", "unequal"), 
               "mean and", 
               ifelse(equal_var, "equal", "unequal"), "variance") |>
    str_to_sentence()

xlab <- paste(error_type, "error on", error_var) |> str_to_sentence()

plot_scores(combined_scores) + 
  ggtitle(title) + 
  labs(y = "Interval Score", 
       x = xlab)

ggsave(here("output", "figures", paste0(scenario_name, ".png")))
```

Simulate data from two normal distributions with $\mu_{1,2} = 1000$ and $\sigma_1 = 0.1$ and $\sigma_2 = 1$.

```{r}
sigma <- c(1, 100)
mu_true <- c(1000, 2000)

true_values_1 <- 
  data.frame(
    true_value = rnorm(n = n, mean = mu_true[1], sd = sigma[1]),
    n = 1:n,
    sigma_true = sigma[1]
  )

true_values_2 <- 
  data.frame(
    true_value = rnorm(n = n, mean = mu_true[1], sd = sigma[2]),
    n = 1:n,
    sigma_true = sigma[2]
  )

```

Define different predictive distributions with mean 1000 and standard deviations between 0 and 10.

```{r}
mu_hat <- 1000

forecasts_1 <- create_forecast(grid_sigma_mult, sigma_true = sigma[1]) 
forecasts_2 <- create_forecast(grid_sigma_mult, sigma_true = sigma[2])

forecasts_mu_1 <- create_forecast(grid_mu_mult, sigma_true = sigma[1], 
                                 mu_true = 1000, var = "mu") 
forecasts_mu_2 <- create_forecast(grid_mu_mult, sigma_true = sigma[2], 
                                 mu_true = 1000, var = "mu") 
```

Join these and create a version with predictions and true values logged. 
```{r}
data_1 <- inner_join(forecasts_1, true_values_1)
data_2 <- inner_join(forecasts_2, true_values_2)

data_mu_1 <- inner_join(forecasts_mu_1, true_values_1)
data_mu_2 <- inner_join(forecasts_mu_2, true_values_2)

log_data_1 <- create_log_data(data_1)
log_data_2 <- create_log_data(data_2)

log_data_mu_1 <- create_log_data(data_mu_1)
log_data_mu_2 <- create_log_data(data_mu_2)
```

Score forecasts.

```{r}
scores_1 <- score_forecasts(data_1)
scores_2 <- score_forecasts(data_2)

log_scores_1 <- score_forecasts(log_data_1)
log_scores_2 <- score_forecasts(log_data_2)

scores_mu_1 <- score_forecasts(data_mu_1)
scores_mu_2 <- score_forecasts(data_mu_2)

log_scores_mu_1 <- score_forecasts(log_data_mu_1)
log_scores_mu_2 <- score_forecasts(log_data_mu_2)

joined_scores <- join_scores(scores_1, scores_2)
joined_log_scores <- join_scores(log_scores_1, log_scores_2, scale = "log")
combined_scores <- rbind(joined_scores, joined_log_scores)

joined_scores_mu <- join_scores(scores_mu_1, scores_mu_2)
joined_log_scores_mu <- join_scores(log_scores_mu_1, log_scores_mu_2, scale = "log")
combined_scores_mu <- rbind(joined_scores_mu, joined_log_scores_mu)

```

Create a plot for data on log scale and natural scale.

```{r}

```

## Scenario: Multiplicative error on the standard deviation - identical means

```{r}
plot_scores(combined_scores) +
    labs(y = "WIS", x = "sd of predictive distribution / sd of data distribution") 
```

*Plot shows forecasts where the mean of the forecast distribution is equal to the mean of the data-generating distribution. The variance of the forecast distribution is multiplied by an error factor between 0.5 and 10.*

Same plot with y axis on a log scale
```{r}
plot_scores(combined_scores) +
    labs(y = "WIS", x = "sd of predictive distribution / sd of data distribution") + 
  ggplot2::scale_y_continuous(trans = "log10")
```

What we learn from these plots:

- A perfect forecast that gets the mean as well as the variance right will incur higher scores if the true variance is higher. 
- getting the variance wrong (in multiplicative terms) is equally costly regardless of what the true variance is (curves look identical on a log scale plot)
- Those two effects are exactly the same on the linear as on the log scale? Not sure I understand this

## Multiplicative error on the mean - identical means

```{r}
plot_scores(combined_scores_mu) +
    labs(y = "WIS", x = "mu of predictive distribution / mu of data distribution")
```

*Plot shows forecasts where the standard deviation of the forecast distribution is equal to the standard deviation of the data-generating distribution. The mean of the forecast distribution is multiplied by an error factor between 0.5 and 10.*

What we learn from this plot:

- with a multiplicative error in the mean and the correct standard deviation, the standard deviation of the underlying data makes some difference, but not that much
- multiplicative errors in the mean get punished much more harshly when scoring on the natural scale compared to scoring on the log scale

```{r}
combined_scores_mu |>
  group_by(error, scale) |>
  summarise(mean_score = mean(interval_score))
```

### 2021-10-23 analysis/Analysis.Rmd
add plots and clean analysis script

@@
 quantiles <- c(0.01, 0.025, seq(0.05, 0.95, by = 0.05), 0.975, 0.99)
+rerun_simulation <- FALSE
 
@@
   mutate(name = paste0(error_type, "_error_", error_var,  
-                       ifelse(equal_mean, "_equal_mean_", "_unequal_mean_"), 
-                       ifelse(equal_var, "equal_var", "unequal_var")))
+                       ifelse(truth_mean_equal, "_equal_mean_", "_unequal_mean_"), 
+                       ifelse(truth_var_equal, "equal_var", "unequal_var")), 
+         title = paste(error_type, "error of", error_var, "for two variables with", 
+                       ifelse(truth_mean_equal, "equal", "unequal"), 
+                       "mean and", 
+                       ifelse(truth_var_equal, "equal", "unequal"), "variance") |>
+           str_to_sentence())
+
+scenarios |>
+  select(truth_mean_equal, truth_var_equal, error_var, error_type)
 
@@
 
-```{r}
+```{r, eval = rerun_simulation}
 # helper functions
@@
 
+# helper function to wrap a long title
+wrap_title <- function(x, ...) {
+  paste(strwrap(x, ...), collapse = "\n")
+}
+
 # helper function to plot scores
@@
              paste0("mean: ", mu_true, ", sd: ", sigma_true)) |>
+    group_by(`True parameters`, error, scale) |>
+    summarise(interval_score = mean(interval_score)) |>
     ggplot(aes(x = error, y = interval_score, 
@@
 
-```{r}
-i = 1
+```{r eval = rerun_simulation}
+for (i in 1:nrow(scenarios)) {
   equal_mean <- scenarios[i, ]$truth_mean_equal
@@
   scenario_name <- scenarios[i, ]$name
+  scenario_title <- scenarios[i, ]$title
 
-sigma_true <- c(sigma, sigma + (as.numeric(!equal_var) * 10))
+  sigma_true <- c(sigma, sigma + (as.numeric(!equal_var) * 90))
   mu_true <- c(mu, mu + as.numeric(!equal_mean) * 1000)
@@
 
-title <- paste(error_type, "error of", error_var, "for two variables with", 
-               ifelse(equal_mean, "equal", "unequal"), 
-               "mean and", 
-               ifelse(equal_var, "equal", "unequal"), "variance") |>
-    str_to_sentence()
-
-xlab <- paste(error_type, "error on", error_var) |> str_to_sentence()
-
   plot_scores(combined_scores) + 
-  ggtitle(title) + 
+    ggtitle(wrap_title(scenario_title, width = 60)) + 
     labs(y = "Interval Score", 
-       x = xlab)
+         x = paste(error_type, "error on", error_var) |> 
+           str_to_sentence())
 
   ggsave(here("output", "figures", paste0(scenario_name, ".png")))
-```
-
-Simulate data from two normal distributions with $\mu_{1,2} = 1000$ and $\sigma_1 = 0.1$ and $\sigma_2 = 1$.
-
-```{r}
-sigma <- c(1, 100)
-mu_true <- c(1000, 2000)
-
-true_values_1 <- 
-  data.frame(
-    true_value = rnorm(n = n, mean = mu_true[1], sd = sigma[1]),
-    n = 1:n,
-    sigma_true = sigma[1]
-  )
-
-true_values_2 <- 
-  data.frame(
-    true_value = rnorm(n = n, mean = mu_true[1], sd = sigma[2]),
-    n = 1:n,
-    sigma_true = sigma[2]
-  )
+}
 
@@
 
-Define different predictive distributions with mean 1000 and standard deviations between 0 and 10.
+---
 
 ```{r}
-mu_hat <- 1000
-
-forecasts_1 <- create_forecast(grid_sigma_mult, sigma_true = sigma[1]) 
-forecasts_2 <- create_forecast(grid_sigma_mult, sigma_true = sigma[2])
-
-forecasts_mu_1 <- create_forecast(grid_mu_mult, sigma_true = sigma[1], 
-                                 mu_true = 1000, var = "mu") 
-forecasts_mu_2 <- create_forecast(grid_mu_mult, sigma_true = sigma[2], 
-                                 mu_true = 1000, var = "mu") 
+i = 1
+file_name <- paste0(scenarios$name[i], ".png")
+knitr::include_graphics(here("output", "figures", file_name))
 ```
 
-Join these and create a version with predictions and true values logged. 
+---
+
 ```{r}
-data_1 <- inner_join(forecasts_1, true_values_1)
-data_2 <- inner_join(forecasts_2, true_values_2)
-
-data_mu_1 <- inner_join(forecasts_mu_1, true_values_1)
-data_mu_2 <- inner_join(forecasts_mu_2, true_values_2)
-
-log_data_1 <- create_log_data(data_1)
-log_data_2 <- create_log_data(data_2)
-
-log_data_mu_1 <- create_log_data(data_mu_1)
-log_data_mu_2 <- create_log_data(data_mu_2)
+i = 2
+file_name <- paste0(scenarios$name[i], ".png")
+knitr::include_graphics(here("output", "figures", file_name))
 ```
 
-Score forecasts.
+---
 
 ```{r}
-scores_1 <- score_forecasts(data_1)
-scores_2 <- score_forecasts(data_2)
+i = 3
+file_name <- paste0(scenarios$name[i], ".png")
+knitr::include_graphics(here("output", "figures", file_name))
+```
 
-log_scores_1 <- score_forecasts(log_data_1)
-log_scores_2 <- score_forecasts(log_data_2)
+---
 
-scores_mu_1 <- score_forecasts(data_mu_1)
-scores_mu_2 <- score_forecasts(data_mu_2)
+```{r}
+i = 4
+file_name <- paste0(scenarios$name[i], ".png")
+knitr::include_graphics(here("output", "figures", file_name))
+```
 
-log_scores_mu_1 <- score_forecasts(log_data_mu_1)
-log_scores_mu_2 <- score_forecasts(log_data_mu_2)
+---
 
-joined_scores <- join_scores(scores_1, scores_2)
-joined_log_scores <- join_scores(log_scores_1, log_scores_2, scale = "log")
-combined_scores <- rbind(joined_scores, joined_log_scores)
+```{r}
+i = 5
+file_name <- paste0(scenarios$name[i], ".png")
+knitr::include_graphics(here("output", "figures", file_name))
+```
 
-joined_scores_mu <- join_scores(scores_mu_1, scores_mu_2)
-joined_log_scores_mu <- join_scores(log_scores_mu_1, log_scores_mu_2, scale = "log")
-combined_scores_mu <- rbind(joined_scores_mu, joined_log_scores_mu)
+---
+
+```{r}
+i = 6
+file_name <- paste0(scenarios$name[i], ".png")
+knitr::include_graphics(here("output", "figures", file_name))
+```
+
+---
+
+```{r}
+i = 7
+file_name <- paste0(scenarios$name[i], ".png")
+knitr::include_graphics(here("output", "figures", file_name))
+```
+
+---
+
+```{r}
+i = 8
+file_name <- paste0(scenarios$name[i], ".png")
+knitr::include_graphics(here("output", "figures", file_name))
+```
+
+---
+
+```{r}
+i = 9
+file_name <- paste0(scenarios$name[i], ".png")
+knitr::include_graphics(here("output", "figures", file_name))
 
 ```
-
-Create a plot for data on log scale and natural scale.
-
-```{r}
-
-```
-
-## Scenario: Multiplicative error on the standard deviation - identical means
-
-```{r}
-plot_scores(combined_scores) +
-    labs(y = "WIS", x = "sd of predictive distribution / sd of data distribution") 
-```
-
-*Plot shows forecasts where the mean of the forecast distribution is equal to the mean of the data-generating distribution.
-The variance of the forecast distribution is multiplied by an error factor between 0.5 and 10.*
-
-Same plot with y axis on a log scale
-```{r}
-plot_scores(combined_scores) +
-    labs(y = "WIS", x = "sd of predictive distribution / sd of data distribution") + 
-  ggplot2::scale_y_continuous(trans = "log10")
-```
-
-What we learn from these plots:
-
-- A perfect forecast that gets the mean as well as the variance right will incur higher scores if the true variance is higher. 
-- getting the variance wrong (in multiplicative terms) is equally costly regardless of what the true variance is (curves look identical on a log scale plot)
-- Those two effects are exactly the same on the linear as on the log scale? Not sure I understand this
-
-## Multiplicative error on the mean - identical means
-
-```{r}
-plot_scores(combined_scores_mu) +
-    labs(y = "WIS", x = "mu of predictive distribution / mu of data distribution")
-```
-
-*Plot shows forecasts where the standard deviation of the forecast distribution is equal to the standard deviation of the data-generating distribution.
-The mean of the forecast distribution is multiplied by an error factor between 0.5 and 10.*
 
@@
 
+---
+
 ```{r}
-combined_scores_mu |>
-  group_by(error, scale) |>
-  summarise(mean_score = mean(interval_score))
+i = 10
+file_name <- paste0(scenarios$name[i], ".png")
+knitr::include_graphics(here("output", "figures", file_name))
 ```
+
+---
+
+```{r}
+i = 11
+file_name <- paste0(scenarios$name[i], ".png")
+knitr::include_graphics(here("output", "figures", file_name))
+```
+
+What we learn from this plot:
+
+- A perfect forecast that gets the mean as well as the variance right will incur higher scores if the true variance is higher. 
+- getting the variance wrong (in multiplicative terms) is equally costly regardless of what the true variance is (curves look identical on a log scale plot, I checked that previously)
+- Those two effects are exactly the same on the linear as on the log scale? Not sure I understand this
+
+---
+
+```{r}
+i = 12
+file_name <- paste0(scenarios$name[i], ".png")
+knitr::include_graphics(here("output", "figures", file_name))
+```
+
+---
+
+```{r}
+i = 13
+file_name <- paste0(scenarios$name[i], ".png")
+knitr::include_graphics(here("output", "figures", file_name))
+```
+
+---
+
+```{r}
+i = 14
+file_name <- paste0(scenarios$name[i], ".png")
+knitr::include_graphics(here("output", "figures", file_name))
+```
+
+---
+
+```{r}
+i = 15
+file_name <- paste0(scenarios$name[i], ".png")
+knitr::include_graphics(here("output", "figures", file_name))
+```
+
+---
+
+```{r}
+i = 16
+file_name <- paste0(scenarios$name[i], ".png")
+knitr::include_graphics(here("output", "figures", file_name))
+```

### 2021-10-23 main.tex
Updates from Overleaf

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Formulas for the scores}
+\textbf{Move to introduction probably}
+
+\subsection{CRPS}
+The CRPS is a proper scoring rule that generalises the absolute error to probabilistic forecasts.
+It measures the 'distance' of the predictive distribution to the observed data-generating distribution.
+The CRPS is given as
+$$\text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,$$
+where y is the true observed value and F the CDF of predictive distribution. Often An alternative representation is used:
+$$ \text{CRPS}(F, y) = \frac{1}{2} \mathbb{E}_{F} |X - X'| - \mathbb{E}_P |X - y|,$$ where $X$ and $X'$ are independent realisations from the predictive distributions $F$ with finite first moment and $y$ is the true value.
+In this represenation we can simply replace $X$ and $X'$ by samples sum over all possible combinations to obtain the CRPS.
+For integer-valued forecasts, the RPS is given as 
+$$ \text{RPS}(F, y) = \sum_{x = 0}^\infty (F(x) - 1(x \geq y))^2. $$
+
+\subsection{WIS}
+\textbf{maybe replace by formulas from WIS paper, maybe use quantile version as well}
+
+The (weighted) interval score is a proper scoring rule for quantile forecasts that converges to the crps for an increasing number of intervals.
+The score can be decomposed into a dispersion component and penalties for over- and under-prediction.
+For a single interval, the score is computed as
+$$IS_\alpha(F,y) = (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u), $$ 
+where $1()$ is the indicator function, $y$ is the true value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution $F$, i.e.
+the lower and upper bound of a single prediction interval.
+For a set of $K$ prediction intervals and the median $m$, the score is computed as a weighted sum,
+$$WIS = \frac{1}{K + 0.5} \cdot (w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y)),$$ 
+where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Effects / Interpretation of scoring on a log scale}
@@
 \begin{itemize}
-    \item take log of (observation or truth + 1) 
+    \item why you can't just log CRPS / WIS (simulation that Johannes did)
+    \item what to do instead: take log of (observation or truth + 1) 
+    \item What about happens when either the observation or the forecast is 0?
+\end{itemize}
+
+\paragraph{what does logging mean}
+\begin{itemize}
     \item discuss how equations change for CRPS, WIS
@@
 
-\paragraph{Toy example: simulate an epidemic and apply 3 different modelse} Compare scores for these three models on the natural scale and the log scale
+\paragraph{Toy example: simulate an epidemic and apply 3 different models} Compare scores for these three models on the natural scale and the log scale
 

### 2021-10-25 analysis/Analysis.Rmd
restructure plots

@@
 
-grid_sigma_mult <- seq(0.5, 2.5, by = 0.1)
-grid_mu_mult <- seq(0.5, 2.5, by = 0.1)
-grid_sigma_add <- seq(0, 100, by = 5)
-grid_mu_add <- seq(0, 1000, by = 50)
+grid_sigma_mult <- seq(0.5, 2, by = 0.1)
+grid_mu_mult <- seq(0.5, 2, by = 0.1)
+grid_sigma_add <- seq(-5, 10, by = 1)
+grid_mu_add <- seq(-500, 1000, by = 100)
 
@@
     summarise(interval_score = mean(interval_score)) |>
-    ggplot(aes(x = error, y = interval_score, 
+    group_by(scale) |>
+    mutate(multplicative_score = interval_score / min(interval_score)) |>
+    ggplot(aes(x = error, y = multplicative_score, 
                colour = `True parameters`, group = `True parameters`)) + 
@@
     scale_color_brewer(palette = "Set1", name = "True parameters: ") + 
+    scale_y_continuous(limits = c(1, NA), expand = c(1, NA)) + 
+    expand_limits(y = c(1, NA)) + 
     facet_wrap(~ scale, scale = "free") + 
@@
 ```{r eval = rerun_simulation}
+
+results <- list()
+
 for (i in 1:nrow(scenarios)) {
@@
   joined_log_scores <- join_scores(log_scores_1, log_scores_2, scale = "log")
-  combined_scores <- rbind(joined_scores, joined_log_scores)
+  combined_scores <- rbind(joined_scores, joined_log_scores) |>
+    mutate(equal_mean = equal_mean, 
+           equal_var = equal_var)
+
+  # store all results
+  results[[scenario_name]] <- combined_scores
 
@@
     ggtitle(wrap_title(scenario_title, width = 60)) + 
-    labs(y = "Interval Score", 
+    labs(y = "Multiplicative change in Interval Score", 
          x = paste(error_type, "error on", error_var) |> 
@@
 
+```
+
+```{r}
+results
+
+res <- bind_rows(results)
+
+plot_multiple <- function(res, equal_means, equal_vars) {
+  res <- res |>
+    filter(equal_mean == equal_means, 
+           equal_var == equal_vars) |>
+    mutate("True parameters" = 
+             paste0("mean: ", mu_true, ", sd: ", sigma_true)) |>
+    group_by(`True parameters`, error, error_type, error_var, scale) |>
+    summarise(interval_score = mean(interval_score)) |>
+    group_by(scale, error_var, error_type) |>
+    mutate(multplicative_score = interval_score / min(interval_score), 
+           error_title = str_to_title(
+             paste(error_type, "on", error_var)))
+
+  res |> 
+    ggplot(aes(x = error, y = multplicative_score, 
+               colour = `True parameters`, group = `True parameters`)) + 
+    geom_line() + 
+    scale_color_brewer(palette = "Set1", name = "True parameters: ") + 
+    # scale_y_continuous(limits = c(1, NA), expand = c(0, NA)) + 
+    expand_limits(y = c(1, NA)) + 
+    facet_wrap( ~ error_title + scale, scale = "free", nrow = 2) + 
+    theme_minimal() + 
+    theme(legend.position = "bottom") +
+    labs(y = "Multiplicative change in Interval Score", 
+         x = "Error") + 
+    geom_rect(data = subset(res, error_var == "sigma"), 
+              fill = NA, colour = "grey80", xmin = -Inf,xmax = Inf,
+              ymin = -Inf,ymax = Inf) + 
+    geom_rect(data = subset(res, error_var == "mu"), 
+              fill = NA, colour = "grey70", xmin = -Inf,xmax = Inf,
+              ymin = -Inf,ymax = Inf) 
+}
+```
+
+---
+
+```{r}
+p1 <- plot_multiple(res, equal_means = TRUE, equal_vars = TRUE) 
+ggsave(p1, filename = here("output", "effect_equal_mean_equal_var.png"))
+
+p1 + 
+  ggtitle("Effect of scoring on a log scale")
+```
+
+---
+
+```{r}
+p2 <- plot_multiple(res, equal_means = FALSE, equal_vars = TRUE) 
+ggsave(p2, filename = here("output", "effect_unequal_mean_equal_var.png"))
+
+p2 + 
+  ggtitle("Effect of scoring on a log scale for two variables with unequal means")
+```
+
+---
+
+```{r}
+p3 <- plot_multiple(res, equal_means = TRUE, equal_vars = FALSE) 
+ggsave(p3, filename = here("output", "effect_equal_mean_unequal_var.png"))
+
+p3 + 
+  ggtitle("Effect of scoring on a log scale for two variables with unequal variances")
+```
+
+```{r}
+p4 <- plot_multiple(res, equal_means = FALSE, equal_vars = FALSE) 
+ggsave(p4, filename = here("output", "effect_unequal_mean_unequal_var.png"))
+
+title <- "Effect of scoring on a log scale for two variables with unequal means and unequal variances"
+p4 + 
+  ggtitle(wrap_title(title, width = 80))
 ```

### 2021-10-25 analysis/Analysis.Rmd
update plots

@@
 
-```{r}
+```{r eval=rerun_simulation}
 results
@@
 }
+
+p1 <- plot_multiple(res, equal_means = TRUE, equal_vars = TRUE) 
+ggsave(p1, filename = here("output", "effect_equal_mean_equal_var.png"))
+
+p2 <- plot_multiple(res, equal_means = FALSE, equal_vars = TRUE) 
+ggsave(p2, filename = here("output", "effect_unequal_mean_equal_var.png"))
+
+p3 <- plot_multiple(res, equal_means = TRUE, equal_vars = FALSE) 
+ggsave(p3, filename = here("output", "effect_equal_mean_unequal_var.png"))
+
+p4 <- plot_multiple(res, equal_means = FALSE, equal_vars = FALSE) 
+ggsave(p4, filename = here("output", "effect_unequal_mean_unequal_var.png"))
+
 ```
@@
 
+Effect of scoring on a log scale
+
 ```{r}
-p1 <- plot_multiple(res, equal_means = TRUE, equal_vars = TRUE) 
-ggsave(p1, filename = here("output", "effect_equal_mean_equal_var.png"))
-
-p1 + 
-  ggtitle("Effect of scoring on a log scale")
+knitr::include_graphics(here("output", "effect_equal_mean_equal_var.png"))
 ```
@@
 
+Effect of scoring on a log scale for two variables with unequal means
+
 ```{r}
-p2 <- plot_multiple(res, equal_means = FALSE, equal_vars = TRUE) 
-ggsave(p2, filename = here("output", "effect_unequal_mean_equal_var.png"))
-
-p2 + 
-  ggtitle("Effect of scoring on a log scale for two variables with unequal means")
+knitr::include_graphics(here("output", "effect_unequal_mean_equal_var.png"))
 ```
@@
 
+Effect of scoring on a log scale for two variables with unequal variances
+
 ```{r}
-p3 <- plot_multiple(res, equal_means = TRUE, equal_vars = FALSE) 
-ggsave(p3, filename = here("output", "effect_equal_mean_unequal_var.png"))
-
-p3 + 
-  ggtitle("Effect of scoring on a log scale for two variables with unequal variances")
+knitr::include_graphics( here("output", "effect_equal_mean_unequal_var.png"))
 ```
 
+Effect of scoring on a log scale for two variables with unequal means and unequal variances
+
 ```{r}
-p4 <- plot_multiple(res, equal_means = FALSE, equal_vars = FALSE) 
-ggsave(p4, filename = here("output", "effect_unequal_mean_unequal_var.png"))
-
-title <- "Effect of scoring on a log scale for two variables with unequal means and unequal variances"
-p4 + 
-  ggtitle(wrap_title(title, width = 80))
+knitr::include_graphics(here("output", "effect_unequal_mean_unequal_var.png"))
 ```

### 2021-10-25 analysis/Analysis.Rmd
rerun plots with higher sample number

@@
 
-n <- 100
+n <- 10000
 quantiles <- c(0.01, 0.025, seq(0.05, 0.95, by = 0.05), 0.975, 0.99)
@@
 p1 <- plot_multiple(res, equal_means = TRUE, equal_vars = TRUE) 
-ggsave(p1, filename = here("output", "effect_equal_mean_equal_var.png"))
+ggsave(p1, filename = here("output", "figures", "effect_equal_mean_equal_var.png"))
 
 p2 <- plot_multiple(res, equal_means = FALSE, equal_vars = TRUE) 
-ggsave(p2, filename = here("output", "effect_unequal_mean_equal_var.png"))
+ggsave(p2, filename = here("output", "figures", "effect_unequal_mean_equal_var.png"))
 
 p3 <- plot_multiple(res, equal_means = TRUE, equal_vars = FALSE) 
-ggsave(p3, filename = here("output", "effect_equal_mean_unequal_var.png"))
+ggsave(p3, filename = here("output", "figures", "effect_equal_mean_unequal_var.png"))
 
 p4 <- plot_multiple(res, equal_means = FALSE, equal_vars = FALSE) 
-ggsave(p4, filename = here("output", "effect_unequal_mean_unequal_var.png"))
+ggsave(p4, filename = here("output", "figures", "effect_unequal_mean_unequal_var.png"))
 

### 2021-10-25 main.tex
write introduction

--- 2021-09-17 main.tex
@@
 % General thought: maybe easier to do with WIS rather than CRPS? 
+% Could do WIS in the main paper and CRPS in the SI?
 
@@
 
-\paragraph{Context about scoring}
+\paragraph{Role of forecasts and problem in comparing them}
 
-\begin{itemize}
-    \item what is a probabilistic forecast and a proper scoring rule, what are scores used for
-    \item CRPS generalisation of absolute error
-    \item idea of scores on logged data: interpretation as a measure of relative improvement, just as when you log the absolute error
-\end{itemize}
+Forecasts play an important role in decision-making in many areas such as for example economics [Citations], agriculture [Citations] and epidemiology [Citations].
+Forecasts are most useful if they are probabilistic in nature [Citation e.g.
+Held et al.], meaning that in addition to a point prediction, forecasters also provide a predictive distribution to quantify their uncertainty.
+Through the COVID-19 pandemic forecasts of case numbers, hospitalisations and deaths have had a strong influence on public policy and there merits and limitations have been widely discussed by politicians, experts and laypeople alike.
+Probabilistic COVID-19 Forecasts from different research institutions have been systematically collected and aggregated by the US Forecast Hub [Citation], the German and Polish Forecast Hub [Citation] and the European Forecast Hub [Citation].
+One major challenge in comparing and evaluating these forecasts is that it is not immediately how to deal with targets on different orders of magnitudes (e.g.
+case numbers and hospitalisations) or how to take the exponential nature of epidemiological processes into account.
+In this paper we will argue that evaluating forecasts on a logarithmic, rather than a natural scale, may help alleviate some of these concerns and may be better suited to epidemiological forecasts.
 
-\paragraph{What's there in the literature}
+%Forecasts should be well calibrated (i.e. should not systematically deviate from the observations) and subject to calibration should be as precise as possible (Cite Gneiting).
 
-\paragraph{Current problems / questions about scoring an epidemiological setting}
-%probably narrow down to the ones we really want to discuss%
-There are limitations with what we can do with scores on a natural scale and open questions whether we can do these things on a log-scale and also whether a log-scale might be inherently more appropriate
-\begin{itemize}
-    \item can't really compare quantities on different orders of magnitudes
-    \item Can't really model score on a natural scale as errors are heavily skewed. Is there a way to model scores to get insight about different factors that systematically influence scores
-    \item Maybe epidemiological forecasts are inherently better suited to be scored on an absolute scale due to the multiplicative nature of processes (related to over-prediction > under-prediction)?
-Does the log-scale imply we are scoring the growth rate?
-    \item unclear what trade-offs of logging / not logging are
-    \item what score (log, not log) matches closest what our intuition (or policymakers) think is good? 
-    \item are we better at forecasting deaths than cases? --> relative measure would help
-    \item There exists confusion about what can be done to a score in general (e.g. people want to take the median, which they shouldn't) %maybe don't add as an extra point
-\end{itemize}
+\paragraph{Context / explanation of scoring rules}
+Generally, probabilistic forecasts like those submitted to the Forecast Hubs are usually evaluated using proper scoring rules [Cite Gneiting et al.
+2007].
+A proper scoring rule incentivises any forecaster to state their true best belief and on average always gives the best score to the forecaster who reports a predictive distribution that is equal to the data-generating distribution.
+Forecasts submitted to the Forecast Hubs were required to follow a quantile-based format where forecasters would report a set of 23 quantiles (11 symmetric prediction intervals as well as a median prediction).
+A natural choice for quantile-based forecasts is the (weighted) interval score (WIS) [Cite Bracher et al.].
+
+The WIS (lower values are better) can be decomposed into a dispersion component and penalties for over- and under-prediction. For a single prediction interval, the interval score is computed as 
+$$IS_\alpha(F,y) = (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u), $$ 
+where $1()$ is the indicator function, $y$ is the true value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution $F$, i.e.
+the lower and upper bound of a single prediction interval.
+For a set of $K$ prediction intervals and the median $m$, the WIS is computed as a weighted sum,
+$$WIS = \frac{1}{K + 0.5} \cdot (w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y)),$$ 
+where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$
+
+The WIS is closely related to the continuous ranked probability score (CRPS, lower values are better) [Cite Gneiting 2007], which itself can be understood as a generalisation of the absolute error to probabilistic forecasts.
+For an increasing set of equally-spaced prediction intervals the WIS converges to the CRPS and shares many of its properties.
+The CRPS measures the 'distance' of the predictive distribution to the observed data-generating distribution as
+$$\text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,$$
+where y is the true observed value and F the CDF of predictive distribution. Often the following alternative representation is used
+$$ \text{CRPS}(F, y) = \frac{1}{2} \mathbb{E}_{F} |X - X'| - \mathbb{E}_P |X - y|,$$ 
+where $X$ and $X'$ are independent realisations from the predictive distributions $F$ with finite first moment and $y$ is the true value.
+In this representation we can simply replace $X$ and $X'$ by samples sum over all possible combinations to obtain the CRPS.
+
+Another proper scoring rule commonly used is the log score [CITATION], which we will not discuss further in this paper.
+
+\paragraph{Problems with the WIS and CRPS in an epidemiological setting}
+Their relation to the absolute error means that both the CRPS and the WIS scale with the prediction target.
+Forecasts of COVID-19 cases, for example typically have much higher scores than forecasts of hospitalisations of death.
+Similarly, when looking at performance across different locations or over time, average scores will be dominated by locations and times with high incidences.
+Single outliers also can have a disproportionate effect on aggregate scores.
+One can argue that this is meaningful and that we should care most about places and periods when incidences are high (maybe cite Bracher et al.).
+However, this may not always be true and is clearly not desirable when comparing performance of a model on different prediction targets: case numbers are not necessarily more important than hospitalisations, just because observed values tend be an order of magnitude higher.
+This makes forecasts often hard or impossible to compare.
+Distributions of forecast scores are usually highly skewed making it hard to model them using standard linear models.
+Researchers instead often resorted to evaluating different prediction targets or locations completely independently, a process that is often cumbersome and makes it hard to identify factors that systematically affect scores.
+In any epidemiological setting there is also a noticeable tension when processes (such as the spread of a disease) that are usually modelled multiplicatively are scored in terms of absolute (additive) errors.
 
 \paragraph{Outline of what we do} 
+A natural proposition is therefore to score forecasts on a logarithmic scale, making it easier to model and compare scores across targets by alleviating concerns about skewed forecast scores and also taking into account the multiplicative nature of epidemiological processes.
+In this paper we will ...
 
-\begin{enumerate}
-    \item Effects / Interpretation of scoring on a log scale
-    \item Appropriate scales for epidemiological (and possibly other) processes
-    \item Modelling log scores
-\end{enumerate}
+%\begin{enumerate}
+%    \item Effects / Interpretation of scoring on a log scale
+%    \item Appropriate scales for epidemiological (and possibly other) processes
+%    \item Modelling log scores
+%\end{enumerate}
 
-\section{Effects / Interpretation of scoring on a log scale}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Scoring on a log scale}
 
-\paragraph{How to log and what happens}
-\begin{itemize}
-    \item take log of (observation or truth + 1) 
-    \item discuss how equations change for CRPS, WIS
-    \item Optional: What happens to the decomposition of the WIS if we log?
-    \item Optional: Discussion of parallels to point forecasts \\
-\end{itemize}
+\paragraph{What we propose}
+Proper scoring rules yield a score as a function of a forecast $x$ and a corresponding observation $y$.
+For epidemiological forecasts of positive quantities we propose to replace WIS and CRPS scores on the natural scale computed as
+%
+$$S(x, y), $$
+%
+by scores computed on the logarithmic scale as 
+%
+$$S(\log{(x + 1)}, \log{(y + 1)}), $$
+where $S()$ is either the WIS or the CRPS and one is added to deal with zeros in the data.
+The so computed score is still proper, as for a single forecast the order between different forecasters is not affected by the monotone transformation.
+Differences in the ranking between forecasters only appear when averaging over multiple forecasts, as deviations from the perfect forecast get penalised differently depending on whether forecasts are scored on the log or the natural scale.
+Note that simply taking the log of a score computed on the natural scale, i.e.
+computing $\log S(x, y)$ results in an improper score and is therefore not advised.
+We illustrate this point in Figure \ref{fig:log-improper} in the SI.
+
+\paragraph{Scoring the log means scoring the growth rate}
+If the forecast target is an incidence (e.g.
+incident cases or hospitalisations), then computing a score based on the logarithm of the forecast and the observation is approximately equivalent to scoring a forecast of a multiplicative growth rate.
+The growth rate is defined as
+
+$$ g_{t, t+1} = \frac{I_{t+1} - I_t}{I_t},$$
+
+where $I_{t+1}$ is the incidence we want to forecast, $I_t$ is the last known observation and $g_{t, t+1}$ is the growth rate between now ($t$) and $t+1$. This can be rewritten as
+
+$$I_{t+1} = (1 + g_{t, t+1}) I_t.$$
+
+If we score $\log I_{t_+1}$ rather than $I_{t+1}$, this changes to
+
+$$\log I_{t+1} = \log (1 + g_{t, t+1}) + \log I_t.$$
+
+The logarithm of today's incident value, $\log I_t$, is a known quantity that we do not need to forecast. This additive shift does not change the forecast score. For small values of g, we have 
+$$\log (1+ g_{t, t+1}) \approx g_{t, t+1}.$$
+
+When we score the logarithm of the forecast and the observations, the interval score for a single prediction interval changes as follows
+
+$$IS_\alpha(\log F, \log y) = (\log u - \log l) + \frac{2}{\alpha} \cdot (\log l - \log y) \cdot 1(\log y \leq \log l) + \frac{2}{\alpha} \cdot (\log y - \log u) \cdot 1(\log y \geq \log u), $$
+
+where $y = I_{t+1}$ (pick consistent notation) and the part inside the indicator functions stays the same as the logarithm is a monotonic transformation and the inequality is preserved.
+Per above, we can rewrite this as
+
+$$IS_\alpha(\log F, \log y) \approx (\log u - \log l) + \frac{2}{\alpha} \cdot (\log l - g_{t, t+1}) \cdot 1(g_{t, t+1} \leq \log l) + \frac{2}{\alpha} \cdot (g_{t, t+1} - \log u) \cdot 1(g_{t, t+1} \geq \log u).$$
+
+% At the moment we're just dropping the g in here.
+Maybe need to rewrite this to explain that u and l really are lower und upper intevals for g * It and that therefore the "+ log It" really cancels out.
+From the formula as is it's not immmediately obvious why the additive constant disappears
+% maybe something like $u(\hat{I}_{t+1)}, where \hat{I}_{t+1} = \hat{g}_{t, t+1} I_{t}$
+
+Conveniently, the decomposition of the WIS into dispersion, over-prediction and under-prediction is preserved.
+The interpretation of the components has now changed in that they now approximately represent uncertainty, over-prediction, and under-prediction with respect to the growth rate.
+
+Note that the actual forecast and the task of the forecaster has not changed.
+We are merely making use of the fact that a forecast already contains several pieces of information.
+Any forecast makes an implicit statement about the growth rate of a process as well as about the expected absolute order of magnitude of the target.
+By scoring based on the log transformed forecasts and observations, we choose to evaluate the forecaster based on their ability to predict the growth rate, rather than absolute numbers.
+Applying the log transformation preservers propriety of the scoring rule.
+In both scenarios, forecasters are therefore incentivised to report their best possible forecast.
+Depending on the way we score the forecasts, we do however decide which types of errors (additive or multiplicative) we would like to penalise most.
+
+%\begin{itemize}
+%    \item discuss how equations change for CRPS
+%    \item Optional: Discussion of parallels to point forecasts and the point that you're still incentivised to report the median\\
+%    \item discuss that even if both scoring rules are proper, they still penalise different things %differently. How does that change incentives for the forecaster? 
+%\end{itemize}
 
 \paragraph{What happens to the WIS / CRPS when you log it?}
-Discuss Johannes' Figure from the WIS paper and what changes when you go from natural to log \\
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/crps log viz.png}
+    \caption{How CRPS changes when you log the observation}
+    \label{fig:log-crps-viz}
+\end{figure}
 
@@
     \item Relative weight depends on the relation between mean and variance of the thing we analyse.
-If variance grows quicker / slower than the mean, than we give more / less relative weight to large things
+If variance grows quicker / slower than the mean, than we give more / less relative weight to large things.
+It would be nice to have an analysis for how the average of a score changes for targets with different variances.
 \end{itemize}
@@
 
+\appendix
+\section{Supplementary information}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/placeholder-image.png}
+    \caption{Illustration of the fact that taking the log of WIS or CRPS results in an improper score}
+    \label{fig:log-improper}
+\end{figure}
+
 \end{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Idea parking lot 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\begin{itemize}
+%    \item idea of scores on logged data: interpretation as a measure of relative improvement, just as when you log the absolute error
+%\end{itemize}
+
+%\paragraph{What's there in the literature}
+
+%\paragraph{Current problems / questions about scoring an epidemiological setting}
+%probably narrow down to the ones we really want to discuss%
+%There are limitations with what we can do with scores on a natural scale and open questions whether we can do these things on a log-scale and also whether a log-scale might be inherently more appropriate
+%\begin{itemize}
+    %\item Can't really model score on a natural scale as errors are heavily skewed. Is there a way to model scores to get insight about different factors that systematically influence scores
+ %   \item make a plot somewhere that shows the distribution of scores
+  %  \item Maybe epidemiological forecasts are inherently better suited to be scored on an absolute scale due to the multiplicative nature of processes (related to over-prediction > under-prediction)?
+Does the log-scale imply we are scoring the growth rate?
+   % \item unclear what trade-offs of logging / not logging are
+    %\item what score (log, not log) matches closest what our intuition (or policymakers) think is good? 
+    %\item are we better at forecasting deaths than cases? --> relative measure would help
+    %\item There exists confusion about what can be done to a score in general (e.g. people want to take the median, which they shouldn't) %maybe don't add as an extra point
+%\end{itemize}

### 2021-10-26 analysis/Analysis ECDC forecasts.Rmd
add data and analysis script for european hub data

---
title: "Log or not - Analysis of ECDC forecasts"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE, 
                      warning = FALSE)

library(scoringutils)
library(purrr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(here)
library(data.table)
library(kableExtra)
library(performance)

recompute_scores <- FALSE
```

```{r}
# helper functions
# ------------------------------------------------------------------------------

make_table <- function(x, caption = "") {
  x %>%
    rename_all(gsub, pattern = "_", replacement = " ") %>%
    rename_all(str_to_sentence) %>%
    mutate_if(is.numeric, round, 2) %>%
    kable(caption = caption) %>%
    kable_styling()
}

# format plot scales
scale_fn <- function(x) {
  ifelse(x >= 1000000, 
         paste0(x / 1000000, "m"), 
         ifelse(x >= 5000,
                paste0(x / 1000, "k"),
                x))
}

```

```{r}
# load data 
# ------------------------------------------------------------------------------

hub_data <- rbindlist(list(
  fread(here("data", "full-data-european-forecast-hub-1.csv")), 
  fread(here("data", "full-data-european-forecast-hub-2.csv"))
)) |>
  unique()

```

```{r eval = recompute_scores}
# score forecasts
# ------------------------------------------------------------------------------

# helper function
score_forecasts <- function(data,
                            summarise_by = c("model", "location", 
                                             "target_end_date", "forecast_date",
                                             "horizon", "target_type"), 
                            scale = c("natural", "log")) {
  if (scale[1] == "log") {
    data <- data |>
      mutate(prediction = log(pmax(prediction, 0) + 1), 
             true_value = log(pmax(true_value, 0) + 1))
  }

  data |>
    eval_forecasts(summarise_by = summarise_by) |>
    mutate(scale = scale[1])
}

scores_natural <- score_forecasts(hub_data)
scores_log <- score_forecasts(hub_data, scale = "log")

scores <- bind_rows(scores_natural, scores_log)
fwrite(scores, here("output", "data", "all-scores-european-hub.csv"))
```

```{r}
scores <- fread(here("output", "data", "all-scores-european-hub.csv"))
```

General thoughts?

- Focus on Hub-ensemble or average across all models? 
- Focus on two weeks ahead?

## Scores across forecast targets

```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  ggplot(aes(y = interval_score, x = target_type)) + 
  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
  geom_boxplot(alpha = 0.5) + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ scale, scale = "free") + 
  theme_minimal() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn, trans = "log10") + 
  labs(y = "Interval score", x = "Target type")
```

*Scores are more comparable across targets*

Mean scores: 
```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(scale, target_type) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") |>
  make_table()
```

## Scores across locations

--> could also do this plot with multiplicative scores, i.e. how much higher is a score than the lowest score observed

```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(target_type, location, scale) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") |>
  ggplot(aes(y = interval_score, x = reorder(location, -interval_score))) + 
  geom_bar(stat = "identity") + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ target_type + scale, scale = "free") + 
  theme_minimal() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn) + 
  labs(y = "Interval score", x = "Target type")
```

*We see that the scores are much more evenly distributed across locations if we log the data*

## Scores over time

restrict to a few locations, e.g. GB, FR, ES (highest average scores there)

UK:

```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble",
         location == "GB",
         horizon == 2) |>
  group_by(target_type, scale) |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale, scale = "free_y") + 
  theme_minimal() + 
  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
```

France: Looks like a data revision

```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble",
         location == "FR",
         horizon == 2) |>
  group_by(target_type, scale) |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale, scale = "free_y") + 
  theme_minimal() + 
  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
```

Spain:

```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble",
         location == "ES",
         horizon == 2) |>
  group_by(target_type, scale) |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale, scale = "free_y") + 
  theme_minimal() + 
  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
```

*Overall it looks like there can still be substantial fluctuation in log scores, but maybe slightly less than when scored on an absolute scale. Interestingly, it seems like the smoothing effect is stronger for deaths than for cases.*

## Change of average scores for increasing forecast horizons

```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble") |>
  group_by(scale, target_type, horizon) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = horizon)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale) + 
  theme_minimal() + 
  labs(x = "Forecast horizon in weeks", y = "Multplicative change in interval score")
```

*Slightly less increase on the log scale, but overall not a lot of difference*

## Distribution of scores

--> not sure that is useful

```{r}
scores |>
  ggplot(aes(x = interval_score)) + 
  geom_density() + 
  facet_wrap(~ target_type + scale, scale = "free") + 
  scale_x_continuous(labels = scale_fn)
```

## Fit a model to data

```{r include=TRUE}
reg_log <- scores |>
  filter(scale == "log") |>
  glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 
      family = gaussian)

reg <- scores |>
  filter(scale == "natural") |>
  glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 
      family = gaussian)
```

```{r}
# check distribution
performance::check_distribution(reg_log) |>
  plot()

# check normality of residuals
check_normality(reg_log) |>
  plot()

# heteroskedasticity of error terms - the plot takes extremly long to run
performance::check_heteroskedasticity(reg_log)
# performance::check_heteroskedasticity(reg_log) |>
#   plot()

# check influence of outliers - plot also takes very long to run
performance::check_outliers(reg_log)
# performance::check_outliers(reg_log) |>
#   plot()

```

## Model rankings

Still to do

- restrict to full set of observations
- check how rankings change

```{r}
# find out which models have submitted forecasts for all locations and time points

```

### 2021-10-26 analysis/Analysis ECDC forecasts.Rmd
show regresssion code in markdown file

@@
 
-```{r include=TRUE}
+Check fit of the log scale model
+
+```{r echo=TRUE}
 reg_log <- scores |>
   filter(scale == "log") |>
-  glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 
-      family = gaussian)
-
-reg <- scores |>
-  filter(scale == "natural") |>
   glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 

### 2022-01-26 analysis/Analysis ECDC forecasts.Rmd
generate some plots for Hub evaluation

@@
   unique()
+```
 
+```{r}
+# create version of data where the true value is the true growth rate
+# interesting question: does it make a difference, whether we transform it 
+# to be the weekly growth rate, i.e. take the n-th root, where n is the horizon?
+
+latest_truth <- fread(here("data", "weekly-truth-Europe.csv")) |>
+  rename(last_known_true_value = true_value) |>
+  mutate(forecast_date = target_end_date + 2) |>
+  select(location, target_type, forecast_date, last_known_true_value) |>
+  unique()
+
+# unsure whether the pmax hack is the right way to do it
+growth_data <- hub_data |>
+  inner_join(latest_truth) |>
+  mutate(true_value = true_value / pmax(last_known_true_value + 1),
+         prediction = prediction / pmax(last_known_true_value + 1)) |>
+  select(-last_known_true_value)
 ```
@@
 
+scores_gr_natural <- score_forecasts(growth_data)
+scores_gr_log <- score_forecasts(growth_data, scale = "log")
+
 scores <- bind_rows(scores_natural, scores_log)
 fwrite(scores, here("output", "data", "all-scores-european-hub.csv"))
+
+scores_gr <- bind_rows(scores_gr_natural, scores_gr_log)
+fwrite(scores_gr, here("output", "data", "all-scores-growth-rate-european-hub.csv"))
 ```
@@
 scores <- fread(here("output", "data", "all-scores-european-hub.csv"))
+scores[, scale := factor(scale, levels = c("natural", "log"))]
+
+scores_gr <- fread(here("output", "data", "all-scores-growth-rate-european-hub.csv"))
+scores_gr[, scale := factor(scale, levels = c("natural", "log"))]
 ```
@@
 ## Scores across forecast targets
+
+This uses the Hub ensemble
 
@@
   labs(y = "Interval score", x = "Target type")
+
+ggsave("output/figures/HUB-average-scores.png", width = 7, height = 5)
 ```
@@
 ```{r}
+# alternatively, we could think about just summarising across all models for these plots? 
 scores |> 
@@
   labs(y = "Interval score", x = "Target type")
+
+ggsave("output/figures/HUB-scores-locations.png", width = 7, height = 5)
 ```
@@
 *We see that the scores are much more evenly distributed across locations if we log the data*
+
+## Variance scores across locations
+
+```{r}
+scores |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  group_by(target_type, location, scale) |>
+  summarise(mean_wis = mean(interval_score), 
+            sd_wis = sd(interval_score), 
+            .groups = "drop_last") |>
+  ggplot(aes(y = sd_wis, x = mean_wis)) + 
+  geom_abline(slope = 1, intercept = 0, linetype = "dashed", size = 0.5) + 
+  geom_point() + 
+  facet_wrap(~ target_type + scale, scale = "free") + 
+  theme_minimal() + 
+  theme(legend.position = "none") + 
+  # scale_x_continuous(labels = scale_fn) + 
+  # scale_y_continuous(labels = scale_fn) + 
+  scale_x_continuous(labels = scale_fn, trans = "log10") +
+  scale_y_continuous(labels = scale_fn, trans = "log10") +
+  labs(x = "Mean WIS", y = "Sd of WIS")
+
+ggsave("output/figures/HUB-sd-vs-mean-scores.png", width = 7, height = 5)
+
+```
 
@@
 
-UK:
-
 ```{r}
@@
   filter(model == "EuroCOVIDhub-ensemble",
-         location == "GB",
+         location %in% c("GB", "ES", "DE"),
          horizon == 2) |>
-  group_by(target_type, scale) |>
+  rename(Location = location) |>
+  group_by(target_type, scale, Location) |>
   mutate(multiplicative_score = interval_score / min(interval_score)) |>
-  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
+  ggplot(aes(y = multiplicative_score, x = forecast_date, colour = Location)) + 
   geom_line() + 
@@
   labs(x = "Forecast date", y = "WIS relative to smallest observed value")
-```
 
-France: Looks like a data revision
-
-```{r}
-scores |> 
-  filter(model == "EuroCOVIDhub-ensemble",
-         location == "FR",
-         horizon == 2) |>
-  group_by(target_type, scale) |>
-  mutate(multiplicative_score = interval_score / min(interval_score)) |>
-  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
-  geom_line() + 
-  geom_point() + 
-  facet_grid(target_type ~ scale, scale = "free_y") + 
-  theme_minimal() + 
-  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
-```
-
-Spain:
-
-```{r}
-scores |> 
-  filter(model == "EuroCOVIDhub-ensemble",
-         location == "ES",
-         horizon == 2) |>
-  group_by(target_type, scale) |>
-  mutate(multiplicative_score = interval_score / min(interval_score)) |>
-  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
-  geom_line() + 
-  geom_point() + 
-  facet_grid(target_type ~ scale, scale = "free_y") + 
-  theme_minimal() + 
-  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
+ggsave("output/figures/HUB-scores-over-time.png", width = 7, height = 5)
 ```
@@
   labs(x = "Forecast horizon in weeks", y = "Multplicative change in interval score")
+
+ggsave("output/figures/HUB-scores-over-horizon.png", width = 7, height = 5)
+
 ```
@@
 ## Distribution of scores
-
---> not sure that is useful
 
@@
 check_normality(reg_log) |>
-  plot()
+  plot(type = "qq")
 
@@
 ```
+
+---
+
+# Looking at growth rates directly
+
+## Scores of the growth rate across targets
+
+This uses the Hub ensemble
+
+```{r}
+scores_gr |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  ggplot(aes(y = interval_score, x = target_type)) + 
+  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
+  geom_boxplot(alpha = 0.5) + 
+  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
+  facet_wrap(~ scale, scale = "free") + 
+  theme_minimal() + 
+  theme(legend.position = "none") + 
+  scale_y_continuous(labels = scale_fn, trans = "log10") + 
+  labs(y = "Interval score", x = "Target type")
+```
+
+*Scoring the growth rate makes things more comparable*
+
+Mean scores: 
+```{r}
+scores_gr |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  group_by(scale, target_type) |>
+  summarise(interval_score = mean(interval_score), 
+            .groups = "drop_last") |>
+  make_table()
+```
+
+## Scores for the growth rate across locations
+
+```{r}
+scores_gr |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  group_by(target_type, location, scale) |>
+  summarise(interval_score = mean(interval_score), 
+            .groups = "drop_last") |>
+  ggplot(aes(y = interval_score, x = reorder(location, -interval_score))) + 
+  geom_bar(stat = "identity") + 
+  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
+  facet_wrap(~ target_type + scale, scale = "free") + 
+  theme_minimal() + 
+  theme(legend.position = "none") + 
+  scale_y_continuous(labels = scale_fn) + 
+  labs(y = "Interval score", x = "Target type")
+```
+
+## Scores for growth rate over time
+
+restrict to a few locations, e.g. GB, FR, ES (highest average scores there)
+
+UK:
+
+```{r}
+scores_gr |> 
+  filter(model == "EuroCOVIDhub-ensemble",
+         location == "GB",
+         horizon == 2) |>
+  group_by(target_type, scale) |>
+  mutate(multiplicative_score = interval_score / min(interval_score)) |>
+  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
+  geom_line() + 
+  geom_point() + 
+  facet_grid(target_type ~ scale, scale = "free_y") + 
+  theme_minimal() + 
+  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
+```
+
+France: Looks like a data revision
+
+```{r}
+scores_gr |> 
+  filter(model == "EuroCOVIDhub-ensemble",
+         location == "FR",
+         horizon == 2) |>
+  group_by(target_type, scale) |>
+  mutate(multiplicative_score = interval_score / min(interval_score)) |>
+  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
+  geom_line() + 
+  geom_point() + 
+  facet_grid(target_type ~ scale, scale = "free_y") + 
+  theme_minimal() + 
+  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
+```
+
+Spain:
+
+```{r}
+scores_gr |> 
+  filter(model == "EuroCOVIDhub-ensemble",
+         location == "ES",
+         horizon == 2) |>
+  group_by(target_type, scale) |>
+  mutate(multiplicative_score = interval_score / min(interval_score)) |>
+  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
+  geom_line() + 
+  geom_point() + 
+  facet_grid(target_type ~ scale, scale = "free_y") + 
+  theme_minimal() + 
+  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
+```
+
+## Change of average scores for the growth rate for increasing forecast horizons
+
+```{r}
+scores_gr |> 
+  filter(model == "EuroCOVIDhub-ensemble") |>
+  group_by(scale, target_type, horizon) |>
+  summarise(interval_score = mean(interval_score), 
+            .groups = "drop_last") |>
+  mutate(multiplicative_score = interval_score / min(interval_score)) |>
+  ggplot(aes(y = multiplicative_score, x = horizon)) + 
+  geom_line() + 
+  geom_point() + 
+  facet_grid(target_type ~ scale) + 
+  theme_minimal() + 
+  labs(x = "Forecast horizon in weeks", y = "Multplicative change in interval score")
+```
+
+## Distribution of scores growth rate
+
+```{r}
+scores_gr |>
+  ggplot(aes(x = interval_score)) + 
+  geom_density() + 
+  facet_wrap(~ target_type + scale, scale = "free") + 
+  scale_x_continuous(labels = scale_fn)
+```
+
+## Fit a model to growth rate data
+
+Check fit of the log scale model
+
+```{r echo=TRUE}
+reg_gr <- scores_gr |>
+  filter(scale == "natural") |>
+  glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 
+      family = gaussian)
+
+reg_gr_log <- scores_gr |>
+  filter(scale == "log") |>
+  glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 
+      family = gaussian)
+```
+
+```{r}
+# check distribution
+performance::check_distribution(reg_gr_log) |>
+  plot()
+
+# check normality of residuals
+check_normality(reg_gr_log) |>
+  plot(type = "qq")
+
+# heteroskedasticity of error terms - the plot takes extremly long to run
+performance::check_heteroskedasticity(reg_gr_log)
+# performance::check_heteroskedasticity(reg_log) |>
+#   plot()
+
+# check influence of outliers - plot also takes very long to run
+performance::check_outliers(reg_gr_log)
+# performance::check_outliers(reg_log) |>
+#   plot()
+
+```

### 2022-01-26 main.tex

@@
 
+% We could look at the log growth rate
+% with any linear model, an additive effect on the log growth rate would mean a multiplicative effect on the growth rate
+
 \section{Introduction}
@@
 The WIS (lower values are better) can be decomposed into a dispersion component and penalties for over- and under-prediction. For a single prediction interval, the interval score is computed as 
-$$IS_\alpha(F,y) = (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u), $$ 
-where $1()$ is the indicator function, $y$ is the true value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution $F$, i.e.
-the lower and upper bound of a single prediction interval.
+\begin{equation}
+ IS_\alpha(F,y) = (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u),    
+\end{equation}
+
+where $1()$ is the indicator function, $y$ is the observed value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution $F$, i.e.
+the lower and upper bound of a single central prediction interval.
 For a set of $K$ prediction intervals and the median $m$, the WIS is computed as a weighted sum,
-$$WIS = \frac{1}{K + 0.5} \cdot (w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y)),$$ 
+\begin{equation}
+WIS = \frac{1}{K + 0.5} \cdot (w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y)),    
+\end{equation} 
 where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$
@@
 The CRPS measures the 'distance' of the predictive distribution to the observed data-generating distribution as
-$$\text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,$$
+
+\begin{equation}
+    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,
+\end{equation}
+
 where y is the true observed value and F the CDF of predictive distribution. Often the following alternative representation is used
-$$ \text{CRPS}(F, y) = \frac{1}{2} \mathbb{E}_{F} |X - X'| - \mathbb{E}_P |X - y|,$$ 
+\begin{equation}
+    \text{CRPS}(F, y) = \frac{1}{2} \mathbb{E}_{F} |X - X'| - \mathbb{E}_P |X - y|,
+\end{equation}
+
 where $X$ and $X'$ are independent realisations from the predictive distributions $F$ with finite first moment and $y$ is the true value.
@@
 %
-$$S(x, y), $$
+\begin{equation}
+S(x, y),     
+\end{equation}
 %
@@
 %
-$$S(\log{(x + 1)}, \log{(y + 1)}), $$
+\begin{equation}
+S(\log{(x + 1)}, \log{(y + 1)}),    
+\end{equation}
+%
 where $S()$ is either the WIS or the CRPS and one is added to deal with zeros in the data.
@@
 
-\paragraph{Scoring the log means scoring the growth rate}
-If the forecast target is an incidence (e.g.
-incident cases or hospitalisations), then computing a score based on the logarithm of the forecast and the observation is approximately equivalent to scoring a forecast of a multiplicative growth rate.
+\paragraph{Scoring the log means scoring the growth rate} Computing a score based on the logarithm of the forecast and the observation is approximately equivalent to scoring a forecast of a multiplicative growth rate of the forecast target.
 The growth rate is defined as
+%
+\begin{equation}
+    g_{t, t+1} = \frac{y_{t+1} - y_t}{y_t},
+\end{equation}
+%
+where $y_{t+1}$ is the forecast target (e.g. reported cases of COVID-19), $y_t$ is the last known observation and $g_{t, t+1}$ is the growth rate between now ($t$) and $t+1$. This can be rewritten as 
+%
+\begin{equation}
+y_{t+1} = (1 + g_{t, t+1}) \cdot y_t.
+\end{equation}
+%
+If we score $\log y_{t_+1}$ rather than $y_{t+1}$, this changes to 
+%
+\begin{equation}
+\log y_{t+1} = \log (1 + g_{t, t+1}) + \log y_t.    
+\end{equation}
+%
+Note that $\log y_t$ is a known quantity that can be removed from both sides without changing the score.
+What is uncertain is $\log y_{t+1} - \log y_t = \log (1 + g_{t, t+1})$.
+For small values of $g$, we lastly have
+\begin{equation}
+    \log (1+ g_{t, t+1}) \approx g_{t, t+1}.
+\end{equation}
+%
+When we score a forecast for $y_{t+1}$ on the log scale, the interval score for a single prediction interval changes as follows
+%
+\begin{align}
+    \text{IS}_\alpha(\log F, \log y) = &(\log u - \log l) \\ 
+    &+ \frac{2}{\alpha} \cdot (\log l - \log y) \cdot 1(\log y \leq \log l) \\
+    &+ \frac{2}{\alpha} \cdot (\log y - \log u) \cdot 1(\log y \geq \log u).
+\end{align}
+%
+Note that the values of the indicator functions stay the same as the logarithm is a monotonic transformation and the inequality is preserved. 
+On the natural scale, $u$, $l$ are the lower and upper bounds of a central prediction interval for a forecast of $y_{t+1} = (1 + g_{t, t+1}) \cdot y_t$. 
+We can rewrite these as 
+\begin{equation}
+   l = l^* \cdot y_t 
+   \;\; \text{and} \;\; 
+   u = u^* \cdot y_t,  
+\end{equation}
+where $u^*$ and $l^*$ are the lower and upper bounds of a prediction interval for $(1 + g_{t, t+1})$. Consequently, 
+\begin{equation}
+  \log l = \log l^* + \log y_t
+  \;\; \text{and} \;\;
+  \log u = \log u^* + \log y_t. 
+\end{equation}
+%
+We can then rewrite the score for a forecast of $\log y_{t+1} = \log (1 + g_{t,t+1}) + \log y_t$ as
+%
+\begin{equation}
+\begin{aligned}
+IS_\alpha&(\log F_{t+1}, \log y_{t+a}) = \\
+&((\log u^* + \log y_t) - (\log l^* + \log y_t)) \\
+&+ \frac{2}{\alpha} \cdot ((\log l^* + \log y_t) - (\log (1 + g_{t, t+1}) + \log y_t)) 
+     \cdot 1((\log (1 + g_{t, t+1}) + \log y_t) \leq (\log l^* + \log y_t) \\
+&+ \frac{2}{\alpha} \cdot ((\log (1 + g_{t, t+1}) + \log y_t) - (\log u^* + \log y_t)) 
+      \cdot 1((\log (1 + g_{t, t+1}) + \log y_t) \geq (\log u^* + \log y_t),
+\end{aligned}
+\end{equation}
+%
+which simplifies to 
+%
+\begin{equation}
+\begin{aligned}
+\label{eqn:is-log}
+\text{IS}_\alpha(\log F_{t+1}, \log y_{t+a}) = &(\log u^* - \log l^*) \\
+&+ \frac{2}{\alpha} \cdot (\log l^* - \log (1 + g_{t, t+1}) 
+     \cdot 1(\log (1 + g_{t, t+1}) \leq \log l^*) \\
+&+ \frac{2}{\alpha} \cdot (\log (1 + g_{t, t+1}) - \log u^*) 
+      \cdot 1(\log (1 + g_{t, t+1}) \geq \log u^*).
+\end{aligned}
+\end{equation}
 
-$$ g_{t, t+1} = \frac{I_{t+1} - I_t}{I_t},$$
+The left-hand side of equation \ref{eqn:is-log} was left unchanged.
+From the right-hand side we can see that for a single prediction interval, scoring the log of a forecast ($\log F_{t+1}$) against the log observed value ($\log y_{t+1}$) is exactly equivalent to evaluating a forecast for $\log (1 + g_{t, t+1})$.
+With $\log (1 + g_{t, t+1} \approx g_{t, t+1}$, this is approximately a forecast for the growth rate for small values of $g_{t, t+1}$.
+Conveniently, the decomposition of the WIS into dispersion, over-prediction and under-prediction is preserved.
+The interpretation of the components has now changed in that they now approximately represent uncertainty, over-prediction, and under-prediction with the growth rate.
 
-where $I_{t+1}$ is the incidence we want to forecast, $I_t$ is the last known observation and $g_{t, t+1}$ is the growth rate between now ($t$) and $t+1$. This can be rewritten as
+\paragraph{Additive vs. multiplicative errors}
+Of course the growth rate (and more generally the term $\log (1 + g_{t, t+1})$) would be linked multiplicatively to the observed value on the natural scale.
+Recall that the WIS is closely linked to the absolute error.
+On the natural scale, we were conceptually evaluating an absolute error of our forecast with regards to the true observed value.
+Simplified to the case of a point forecast $\hat{y}_{t+1}$ we could express this as measuring an absolute error $\varepsilon_{t+1}$ such that
+%
+\begin{equation}
+    \hat{y}_{t+1} = y_{t+1} + \varepsilon_{t+1}. 
+¸\end{equation}
+In the case of scoring the log forecast, where we evaluate an absolute error on $\log (1 + g_{t, t+1})$, it makes more sense to understand the error $y_{t+1}$ as being a multiplicative error $\varepsilon^*_{t+1}$ so that
+\begin{align}
+    \hat{y}_{t+1} &= y_{t+1} \cdot \varepsilon^*_{t+1} \\
+    &= y_{t} \cdot (1 + g_{t, t+1}) \cdot \varepsilon^*_{t+1},
+\end{align}
+%
+which on the log scale becomes to 
+\begin{align}
+    \hat{y}_{t+1} = \log y_{t} + \log (1 + g_{t, t+1}) + \varepsilon^*_{t+1}. 
+\end{align}
 
-$$I_{t+1} = (1 + g_{t, t+1}) I_t.$$
+The log transformation therefore changes which types of errors (additive or multiplicative) are penalised.
+The WIS and CRPS computed on the natural scale penalise large absolute errors, regardless of whether the error is large in relative terms.
+Predicting 101,000 cases instead of the true 100,000 will be treated equally to predicting 2,000 hospitalisations, rather than the 1,000 observed.
+The WIS and CRPS based on log-transformed values however, penalise relative errors.
+Forecasting 101,000 rather than 100,000 cases will be treated equally to forecasting 2,020, rather than 2,000 hospitalisations.
+The score is relative and therefore less dependent on the scale of the forecast quantity, making it easier to compare (relative) performance across different forecast targets, periods and locations.
 
-If we score $\log I_{t_+1}$ rather than $I_{t+1}$, this changes to
+\paragraph{Note on propriety and shifting the focus of the evaluation}
+Applying the log transformation preservers propriety of the scoring rule.
+The task of the forecaster has not changed and in both scoring scenarios, forecasters are incentivised to report their best possible forecast.
+Any forecast that is made already implicitly contains a statement about the growth rate of a process as well as about the expected absolute order of magnitude of the target.
+By scoring based on the log transformed forecasts and observations, we merely choose to shift the attention towards evaluating the forecaster based on relative errors, rather than absolute numbers.
 
-$$\log I_{t+1} = \log (1 + g_{t, t+1}) + \log I_t.$$
+\section{What happens when you average many scores}
 
-The logarithm of today's incident value, $\log I_t$, is a known quantity that we do not need to forecast. This additive shift does not change the forecast score. For small values of g, we have 
-$$\log (1+ g_{t, t+1}) \approx g_{t, t+1}.$$
+THIS COULD INCLUDE JOHANNES EXAMPLE
 
-When we score the logarithm of the forecast and the observations, the interval score for a single prediction interval changes as follows
+COULD ALSO JUST MAKE A SIMULATION. 
+YOU TAKE 1000 little counties, combine them to 20 states with different sizes. Then you plot
 
-$$IS_\alpha(\log F, \log y) = (\log u - \log l) + \frac{2}{\alpha} \cdot (\log l - \log y) \cdot 1(\log y \leq \log l) + \frac{2}{\alpha} \cdot (\log y - \log u) \cdot 1(\log y \geq \log u), $$
+This means the WIS and CRPS on the natural scale are dependent on the scale of the forecast target, because forecasting errors will typically be large for large quantities. 
+However, note that large relative errors will likely more often occur with small quantities (e.g.
+predicting 9, rather than 3 deaths), potentially giving undue weight to small quantities.
+Whether or not more weight is given to smaller or larger forecast quantities depends on the exact relationship between the mean and the variance of the quantity of interest.
+If the variance grows slower than the mean, then scores will on average be larger for small quantities than for larger ones.
+In practice, this can be partly mitigated by restricting the analysis to forecast targets with with higher incidences.
+% the phrase "scores give a higher weight to a or b" could be made clearer: This is about a scenario where you average across different targets and the average score is influenced by outliers
 
-where $y = I_{t+1}$ (pick consistent notation) and the part inside the indicator functions stays the same as the logarithm is a monotonic transformation and the inequality is preserved.
-Per above, we can rewrite this as
+%COMPARISON TO SCORING THE ACTUAL GROWTH RATE - just harder to do and achieves the same thing? No objections if someone wants to? Might actually be better if there are negative values involved?
 
-$$IS_\alpha(\log F, \log y) \approx (\log u - \log l) + \frac{2}{\alpha} \cdot (\log l - g_{t, t+1}) \cdot 1(g_{t, t+1} \leq \log l) + \frac{2}{\alpha} \cdot (g_{t, t+1} - \log u) \cdot 1(g_{t, t+1} \geq \log u).$$
-
-% At the moment we're just dropping the g in here.
-Maybe need to rewrite this to explain that u and l really are lower und upper intevals for g * It and that therefore the "+ log It" really cancels out.
-From the formula as is it's not immmediately obvious why the additive constant disappears
-% maybe something like $u(\hat{I}_{t+1)}, where \hat{I}_{t+1} = \hat{g}_{t, t+1} I_{t}$
-
-Conveniently, the decomposition of the WIS into dispersion, over-prediction and under-prediction is preserved.
-The interpretation of the components has now changed in that they now approximately represent uncertainty, over-prediction, and under-prediction with respect to the growth rate.
-
-Note that the actual forecast and the task of the forecaster has not changed.
-We are merely making use of the fact that a forecast already contains several pieces of information.
-Any forecast makes an implicit statement about the growth rate of a process as well as about the expected absolute order of magnitude of the target.
-By scoring based on the log transformed forecasts and observations, we choose to evaluate the forecaster based on their ability to predict the growth rate, rather than absolute numbers.
-Applying the log transformation preservers propriety of the scoring rule.
-In both scenarios, forecasters are therefore incentivised to report their best possible forecast.
-Depending on the way we score the forecasts, we do however decide which types of errors (additive or multiplicative) we would like to penalise most.
+%What about other things like scoring the square root of the forecast?
 
@@
     \includegraphics[width=0.9\textwidth]{output/crps log viz.png}
-    \caption{How CRPS changes when you log the observation}
+    \caption{Figure from Johannes: How CRPS changes when you log the observation}
     \label{fig:log-crps-viz}
@@
 
-\paragraph{Interpretation of scores on log-scale}
+% could make an analysis re outliers: how do scores change if we remove the 1 or 2 worst forecasts? The one or two best forecasts?
 
-\begin{itemize}
-    \item We can interpret as a relative measure
-    \item scale-free
-\end{itemize}
+\paragraph{detailed interpretation of plots and when you give larger scores to what}
 
-\paragraph{Practical considerations / limitations}
-
-\begin{itemize}
-    \item Before: lot of weight to scores of targets that are large
-    \item Now: Potentially a lot of weight to things that are small
-    \item Relative weight depends on the relation between mean and variance of the thing we analyse.
-If variance grows quicker / slower than the mean, than we give more / less relative weight to large things.
-It would be nice to have an analysis for how the average of a score changes for targets with different variances.
-\end{itemize}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/placeholder-image.png}
+    \caption{Plot that shows the scores for log and non log and how they change}
+    \label{fig:change-in-scores}
+\end{figure}
 

### 2022-01-26 analysis/Analysis ECDC forecasts.Rmd
split data files into smaller files
Former-commit-id: 78e173bad2cc7709054fe238f7f35717d59fbed1

@@
 # ------------------------------------------------------------------------------
-
 hub_data <- rbindlist(list(
   fread(here("data", "full-data-european-forecast-hub-1.csv")), 
-  fread(here("data", "full-data-european-forecast-hub-2.csv"))
+  fread(here("data", "full-data-european-forecast-hub-2.csv")),
+  fread(here("data", "full-data-european-forecast-hub-3.csv")),
 )) |>

### 2022-01-27 analysis/illustration-effect-log-score.R
add illustration example for effect logging scores

@@ -1,0 +1,79 @@
+library(data.table)
+library(scoringutils)
+library(ggplot2)
+library(scoringRules)
+library(dplyr)
+library(patchwork)
+
+# define simulation parameters
+n_steps = 500
+n_rep <- 5000
+true_mean = 0
+true_sd = 5
+true_values <- rnorm(n = n_rep, mean = true_mean, sd = true_sd)
+sd <- 10^(seq(-1, 1.6, length.out = n_steps))
+mu <- seq(0, 100, length.out = n_steps)
+
+
+true_values <- seq(0.2, 5, length.out = 1000)
+
+vals <- expand.grid(
+  true_value = true_values, 
+  quantile = c(0.01, 0.025, seq(0.05, 0.95, 0.05), 0.975, 0.99)
+) |>
+  mutate(id = true_value) |>
+  arrange(true_value, quantile) |>
+  mutate(prediction = qnorm(p = quantile, mean = 1, sd = 0.4), 
+         model = "Model") 
+
+scores <- vals |>
+  mutate(scale = "natural") |>
+  eval_forecasts() |> 
+  rbind(
+    vals |>
+      mutate(true_value = log(true_value), 
+             prediction = log(prediction), 
+             scale = "log") |>
+      eval_forecasts()
+  ) |>
+  group_by(scale) |>
+  mutate(score = interval_score / min(interval_score), 
+         scale = factor(scale, levels = c("natural", "log"))) 
+
+scale_factor <- 3 # scale factor for the density
+
+p1 <- scores |>
+  ggplot(aes(x = id, y = interval_score/scale_factor)) + 
+  geom_area(stat = "function", 
+            fun = function(x) {dnorm(x, mean = 1, sd = 0.4)}, 
+            color = "grey", fill = "grey", alpha = 0.5) +
+  geom_line() + 
+  labs(y = "WIS", x = "Observed value") + 
+  scale_y_continuous(label = function(x) {paste(scale_factor * x)}) + 
+  facet_wrap(~ scale) + 
+  theme_minimal()
+
+scale_factor2 <- 3
+p2_log <- scores |>
+  ggplot(aes(x = id, y = interval_score/scale_factor2)) + 
+  geom_area(stat = "function", 
+            fun = function(x) {dnorm(x, mean = 1, sd = 0.4)}, 
+            color = "grey", fill = "grey", alpha = 0.5) +
+  geom_line() + 
+  labs(y = "WIS", x = "Observed value") + 
[…]

### 2022-01-28 analysis/Simulation-poisson-mean-sd.R
add simulation poisson mean and sd of WIS

@@ -1,0 +1,158 @@
+library(data.table)
+library(dplyr)
+library(scoringutils)
+library(ggplot2)
+library(tidyr)
+
+mean_county <- 100
+time_points <- 1000
+state_sizes <- 10^seq(0, 2, 0.01)
+q <- c(0.01, 0.025, seq(0.05, 0.95, 0.05), 0.975, 0.99)
+
+df <- data.table(
+  state = 1:length(state_sizes),
+  state_size = state_sizes,
+  quantile = list(q)
+) |>
+  unnest(quantile) |>
+  mutate(date = list(1:time_points)) |>
+  unnest(cols = date) |>
+  group_by(state, state_size, date) |>
+  mutate(true_value = rpois(1, mean_county * state_size), 
+         prediction = qpois(p = quantile, lambda = mean_county * state_size)) 
+
+
+scores_state <- df |>
+  eval_forecasts()
+
+scores_state_log <- df |>
+  mutate(true_value = log(true_value), 
+         prediction = log(prediction)) |>
+  eval_forecasts()
+
+make_plot <- function(scores, summary_fct = mean) {
+  scores |>
+    group_by(state_size) |>
+    summarise(interval_score = summary_fct(interval_score)) |>
+    ggplot(aes(y = interval_score, x = state_size)) +
+    geom_point() +  
+    labs(y = "WIS", x = "Size of state") + 
+    theme_minimal() 
+}
+
+mean_state <- scores_state |> 
+  make_plot() 
+
+mean_state2 <- scores_state |> 
+  make_plot() +
+  scale_x_continuous(trans = "log10")
+
+mean_state_log <- scores_state_log |> 
+  make_plot() 
+
+mean_state_log2 <- scores_state_log |> 
+  make_plot() +
+  scale_x_continuous(trans = "log10")
+
+(mean_state + mean_state_log) /
+  (mean_state2 + mean_state_log2)
+
+ggsave("output/figures/SIM-mean-sd-state-size.png", width = 7, height = 4)
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+# # know that sum of independent poisson variables is poisson with lambda = lambda1 + lambda2 etc. 
+# 
+# n_counties <- 10000
+# mean_county <- 100
+# dates <- 1:15
+# n_states <- 300
+# q <- c(0.01, 0.025, seq(0.05, 0.95, 0.05), 0.975, 0.99)
+# 
[…]

### 2022-01-28 analysis/illustration-effect-log-score.R

@@ -74,6 +74,6 @@ p2_log <- scores |>
 
 p1 / p2_log
 
-ggsave("output/figures/SIM-effect-log-score.png", width = 7, 
+ggsave("output/figures/SIM-effect-log-score.png", width = 6, 
        height = 4)
 

### 2022-01-28 main.tex

@@
 
+\tableofcontents
+
 \begin{abstract}
@@
 
+% Introduction
+%    - Role of forecasts and problems in comparing them
+%    - Introduction to scoring rules, WIS, CRPS
+%    - outline of the paper
+%Scoring on a log scale
+%    - what we propose
+%    - we approximately score the growth rate
+%    - focus is shifted from additive to multiplicative error
+%    - Interpretation: 
+%        - How do forecast weights change: what gets high and %what gets low scores? 
+%        - - Figure from Johannes: How CRPS changes when you log %the observation
+%
+%Appropriate scales for epidemiological processes
+%    - general point about how all processes are modelled %multiplicatively and therefore it might be odd to score them %based on absolute errors
+%    - Simulations Sam? 
+%    - Adding up several districts - how do scores change? %(Example Johannes) --> relationship between mean and variance %in scores
+%
+%Example: European Forecast Hub: 
+%    - paragraph introduction to the Forecast Hub
+%    - How does the distribution of scores change
+%        - target types, locations, over time
+%    - relationship between mean and sd of WIS
+
+% ===========================================================
 \section{Introduction}
@@
 The so computed score is still proper, as for a single forecast the order between different forecasters is not affected by the monotone transformation.
+% last sentence not very clear
 Differences in the ranking between forecasters only appear when averaging over multiple forecasts, as deviations from the perfect forecast get penalised differently depending on whether forecasts are scored on the log or the natural scale.
@@
 We illustrate this point in Figure \ref{fig:log-improper} in the SI.
+% ADD A COMMENT ABOUT LOG X + 1
 
@@
 
+\section{Empirical example: the European Forecast Hub}
+
+\paragraph{Introduction to the Forecast Hub}
+
+NEED A TABLE WITH SCORES
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-average-scores.png}
+    \caption{CAPTION}
+    \label{fig:HUB-average-scores}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-scores-locations.png}
+    \caption{CAPTION}
+    \label{fig:HUB-scores-location}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-scores-over-horizon.png}
+    \caption{CAPTION}
+    \label{fig:HUB-scores-horizon}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-scores-over-time.png}
+    \caption{CAPTION}
+    \label{fig:HUB-scores-time}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-sd-vs-mean-scores.png}
+    \caption{CAPTION}
+    \label{fig:HUB-mean-sd-scores}
+\end{figure}
+
 \section{Modelling scores}

### 2022-01-28 main.tex

@@
     \centering
-    \includegraphics[width=0.9\textwidth]{output/crps log viz.png}
-    \caption{Figure from Johannes: How CRPS changes when you log the observation}
+    \includegraphics[width=0.9\textwidth]{output/figures/SIM-effect-log-score.png}
+    \caption{Weighted interval score for a forecast distribution that is N(1, 0.2) and different observed values.
+Top: difference in absolute terms.
+Bottom: difference in relative terms.
+Interesting panels are top left and bottom right.}.
     \label{fig:log-crps-viz}

### 2022-01-28 analysis/Analysis ECDC forecasts.Rmd
update plots

--- 2022-01-26 analysis/Analysis ECDC forecasts.Rmd
@@
 # ------------------------------------------------------------------------------
-
 hub_data <- rbindlist(list(
   fread(here("data", "full-data-european-forecast-hub-1.csv")), 
-  fread(here("data", "full-data-european-forecast-hub-2.csv"))
+  fread(here("data", "full-data-european-forecast-hub-2.csv")),
+  fread(here("data", "full-data-european-forecast-hub-3.csv"))
 )) |>
@@
 ggsave("output/figures/HUB-sd-vs-mean-scores.png", width = 7, height = 5)
+
+wis_vs_total <- function(scores, summary_fct = mean, fulldata = hub_data) {
+
+  total_values <- 
+    fulldata |> 
+    filter(!is.na(prediction)) |>
+    group_by(location, target_type, target_end_date) |>
+    summarise(true_value = unique(true_value)) |>
+    group_by(location, target_type) |>
+    summarise(total = sum(true_value))
+
+  scores |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  group_by(target_type, location, scale) |>
+  summarise(wis = summary_fct(interval_score), 
+            .groups = "drop_last") |> 
+  inner_join(total_values) |>
+  ggplot(aes(y = wis, x = total)) + 
+  geom_point() + 
+  facet_wrap(~ target_type + scale, scale = "free") + 
+  theme_minimal() + 
+  theme(legend.position = "none") + 
+  # scale_x_continuous(labels = scale_fn) + 
+  # scale_y_continuous(labels = scale_fn) + 
+  scale_x_continuous(labels = scale_fn, trans = "log10") +
+  scale_y_continuous(labels = scale_fn, trans = "log10") +
+  labs(y = "REPLACE", x = "Total number of observed values")
+}
+
+mean_wis_vs_total <- wis_vs_total(scores)
+ggsave("output/figures/HUB-mean-scores-vs-total.png", width = 7, height = 5)
+
+sd_wis_vs_total <- wis_vs_total(scores, summary_fct = stats::sd)
+ggsave("output/figures/HUB-sd-scores-vs-total.png", width = 7, height = 5)
 

### 2022-01-28 analysis/Simulation-poisson-mean-sd.R

@@ -60,8 +60,19 @@ mean_state_log2 <- scores_state_log |>
 ggsave("output/figures/SIM-mean-sd-state-size.png", width = 7, height = 4)
 
 
+# plots mean vs. sd of wis -----------------------------------------------------
+mean_wis_vs_sd <- function(scores) {
+  scores |>
+    group_by(state_size) |>
+    summarise(mean_wis = mean(interval_score), 
+              sd_wis = sd(interval_score)) |>
+    ggplot(aes(y = mean_wis, x = sd_wis)) +
+    geom_point() +  
+    labs(y = "Mean WIS", x = "Sd WIS") + 
+    theme_minimal() 
+}
 
-
+mean_wis_vs_sd(scores_state_log)
 
 
 

### 2022-01-28 analysis/illustration-effect-log-score.R

@@ -62,7 +62,7 @@ p2_log <- scores |>
   geom_line() + 
   labs(y = "WIS", x = "Observed value") + 
   scale_y_continuous(label = function(x) {paste(scale_factor2 * x)}) + 
-  scale_x_continuous(trans = "log10", 
+  scale_x_continuous(trans = "log", 
                      label = function(x) {
                        ifelse(x < 1, 
                               paste0("1/", (1 / x)), 

### 2022-01-30 analysis/Analysis ECDC forecasts.Rmd
add plot with mean vs. var of observed values

@@
 - Focus on two weeks ahead?
+
+## Mean vs. sd of observed values
+
+```{r}
+
+label_fn <- function(x) {
+  ifelse(x < 1000, 
+         paste(x), 
+         ifelse(x < 1e6, 
+                paste0(x / 1000, "k"),
+                ifelse(x < 1e9, 
+                       paste0(x / 1e6, "m"), 
+                       paste0(x / 1e9, "b"))
+                )
+         )
+}
+
+hub_data |>
+  select(location, target_type, target_end_date, true_value) |>
+  unique() |>
+  group_by(location, target_type) |>
+  summarise(mean_value = mean(true_value), 
+            sd_value = sd(true_value), 
+            var_value = var(true_value)) |>
+  ggplot(aes(y = var_value, x = mean_value)) +
+  geom_point() + 
+  facet_wrap(~ target_type, scales = "free") + 
+  scale_x_continuous(trans = "log10", labels = label_fn) +
+  scale_y_continuous(trans = "log10", labels = label_fn) +
+  # scale_x_continuous(labels = label_fn) +
+  # scale_y_continuous(labels = label_fn) +
+  labs(y = "Variance observed value", x = "Mean observed value") + 
+  geom_function(fun = function(x) x^2, linetype = "dashed", size = 0.5) + 
+  theme_minimal()
+
+ggsave("output/figures/HUB-mean-var-obs.png", width = 7, height = 4)
+
+```
 

### 2022-01-30 analysis/Simulation-poisson-mean-sd.R

@@ -4,75 +4,97 @@ library(scoringutils)
 library(ggplot2)
 library(tidyr)
 
-mean_county <- 100
-time_points <- 1000
-state_sizes <- 10^seq(0, 2, 0.01)
-q <- c(0.01, 0.025, seq(0.05, 0.95, 0.05), 0.975, 0.99)
+setup_df <- function(state_sizes = 10^seq(0, 2, 0.01), 
+                     q = c(0.01, 0.025, seq(0.05, 0.95, 0.05), 0.975, 0.99), 
+                     time_points = 1000) {
+  df <- data.table(
+    state = 1:length(state_sizes),
+    state_size = state_sizes,
+    quantile = list(q)
+  ) |>
+    unnest(quantile) |>
+    mutate(date = list(1:time_points)) |>
+    unnest(cols = date) |>
+    group_by(state, state_size, date)  
+}
 
-df <- data.table(
-  state = 1:length(state_sizes),
-  state_size = state_sizes,
-  quantile = list(q)
-) |>
-  unnest(quantile) |>
-  mutate(date = list(1:time_points)) |>
-  unnest(cols = date) |>
-  group_by(state, state_size, date) |>
+score_states <- function(df) {
+  scores_state <- df |>
+    eval_forecasts() |>
+    mutate(scale = "natural")
+  
+  scores_state_log <- df |>
+    mutate(true_value = log(true_value + 1), 
+           prediction = log(prediction + 1)) |>
+    eval_forecasts() |>
+    mutate(scale = "log")
+  
+  scores <- 
+    rbind(scores_state, 
+          scores_state_log) |>
+    mutate(scale = factor(scale, levels = c("natural", "log")))
+  
+  return(scores)
+}
+
+make_plot <- function(scores, summary_fct = mean) {
+  p1 <- scores |>
+    group_by(state_size, scale) |>
+    summarise(interval_score = summary_fct(interval_score)) |>
+    ggplot(aes(y = interval_score, x = state_size)) +
+    geom_point(size = 0.4) +  
+    labs(y = "WIS", x = "Size of state") + 
+    theme_minimal() + 
+    facet_wrap(~ scale, scales = "free_y")
+  
+  p2 <- p1 + 
+    scale_x_continuous(trans = "log10")
+  
+  p1 / p2
+}
+
+
+
+# example poisson --------------------------------------------------------------
+mean_county <- 100
+df <- setup_df() |>
   mutate(true_value = rpois(1, mean_county * state_size), 
[…]

### 2022-01-30 main.tex

@@
 In this paper we will ...
-
 %\begin{enumerate}
@@
 The log transformation therefore changes which types of errors (additive or multiplicative) are penalised.
-The WIS and CRPS computed on the natural scale penalise large absolute errors, regardless of whether the error is large in relative terms.
+The WIS and CRPS computed on the natural scale increase linearly with increasing absolute errors (see Figure \ref{fig:change-in-scores}A, regardless of whether the error is large in relative terms.
 Predicting 101,000 cases instead of the true 100,000 will be treated equally to predicting 2,000 hospitalisations, rather than the 1,000 observed.
-The WIS and CRPS based on log-transformed values however, penalise relative errors.
+The WIS and CRPS based on log-transformed values however, penalise relative errors and increase linearly with an increase in relative errors (see Figure \ref{fig:change-in-scores}D).
 Forecasting 101,000 rather than 100,000 cases will be treated equally to forecasting 2,020, rather than 2,000 hospitalisations.
-The score is relative and therefore less dependent on the scale of the forecast quantity, making it easier to compare (relative) performance across different forecast targets, periods and locations.
+
+\paragraph{What happens to the WIS / CRPS when you log it?}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/SIM-effect-log-score.png}
+    \caption{Weighted interval score for a forecast distribution that is N(1, 0.2) and different observed values.
+Top: difference in absolute terms.
+Bottom: difference in relative terms.
+Interesting panels are top left and bottom right.}.
+    \label{fig:log-crps-viz}
+\end{figure}
 
@@
 
-\section{What happens when you average many scores}
+\section{Apropriate scales for epidemiological processes}
 
-THIS COULD INCLUDE JOHANNES EXAMPLE
+\paragraph{Modelling on the log scale}
 
-COULD ALSO JUST MAKE A SIMULATION. 
-YOU TAKE 1000 little counties, combine them to 20 states with different sizes. Then you plot
+General point: Scores are for a) improving models, b) choosing a model for decision making. 
+At least for a): models are multiplicative, so if a model is making error, it's making a multiplicative error. We therefore should score it based on that.
+
+%    - general point about how all processes are modelled %multiplicatively and therefore it might be odd to score them %based on absolute errors
+%    - Simulations Sam? 
+%    - Adding up several districts - how do scores change? %(Example Johannes) --> relationship between mean and variance %in scores
+
+%THIS COULD INCLUDE JOHANNES EXAMPLE
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/SIM-mean-sd-state-size.png}
+    \caption{Simulation of the effect of population size with ideal forecasts of a poisson-distributed variable.
+For each simulated state, we drew 1,000 observations from a poisson distribution with $\lambda = 100 \cdot \text{state size}$ and computed the WIS for an ideal forecast (i.e.
+the predictive distribution was poisson with $\lambda$ equal to the true $\lambda$ for every state).
+Left: Mean WIS depending on state size, right: Mean WIS depending on state sizes when scored on a log scale.
+Plots for the standard deviation, rather than the mean of WIS values look almost identical.}.
+    \label{fig:log-crps-viz}
+\end{figure}
+
+The score is relative and therefore less dependent on the scale of the forecast quantity, making it easier to compare (relative) performance across different forecast targets, periods and locations.
 
@@
 %\end{itemize}
-
-\paragraph{What happens to the WIS / CRPS when you log it?}
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/SIM-effect-log-score.png}
-    \caption{Weighted interval score for a forecast distribution that is N(1, 0.2) and different observed values.
-Top: difference in absolute terms.
-Bottom: difference in relative terms.
-Interesting panels are top left and bottom right.}.
-    \label{fig:log-crps-viz}
-\end{figure}
 

### 2022-01-31 analysis/Analysis ECDC forecasts.Rmd
add plot with mean observed values across locations

--- 2022-01-28 analysis/Analysis ECDC forecasts.Rmd
@@
 library(performance)
+library(tidytext)
 
 recompute_scores <- FALSE
+
+# situations with negative values: France 2021-05-22, CZ 2021-08-07, ES 2021-03-06
 ```
@@
 - Focus on two weeks ahead?
+
+## Mean vs. sd of observed values
+
+```{r}
+
+label_fn <- function(x) {
+  ifelse(x < 1000, 
+         paste(x), 
+         ifelse(x < 1e6, 
+                paste0(x / 1000, "k"),
+                ifelse(x < 1e9, 
+                       paste0(x / 1e6, "m"), 
+                       paste0(x / 1e9, "b"))
+                )
+         )
+}
+
+# average number of cases and deaths
+hub_data |>
+  dplyr::select(location, target_type, target_end_date, true_value) |>
+  unique() |>
+  group_by(target_type) |>
+  summarise(mean_value = mean(true_value), 
+            sd_value = sd(true_value), 
+            var_value = var(true_value))
+
+# negative binomial
+# MASS::glm.nb(true_value ~ target_type, 
+#        data = hub_data %>%
+#          mutate(true_value = pmax(true_value, 0)), 
+#        link = "identity")
+
+# plot with means across locations ---------------------------------------------
+label_fn_within <- function(x) {
+  x <- gsub(pattern = "___Deaths", "", x)
+  x <- gsub(pattern = "___Cases", "", x)
+  paste(x)
+}
+
+hub_data |>
+  dplyr::select(location, target_type, target_end_date, true_value) |>
+  unique() |>
+  group_by(target_type, location) |>
+  summarise(mean_value = mean(true_value), 
+            sd_value = sd(true_value), 
+            var_value = var(true_value)) |>
+  ggplot(aes(y = mean_value, x = reorder_within(location, -mean_value, target_type))) + 
+  geom_bar(stat = "identity") + 
+  facet_wrap(~ target_type, scales = "free") + 
+  theme_minimal() + 
+  scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) + 
+  labs(y = "Mean observed value", x = "Location")
+
+ggsave("output/figures/HUB-mean-obs-location.png", width = 9, height = 4)
+
+# plot with mean vs. variance --------------------------------------------------
+hub_data |>
+  dplyr::select(location, target_type, target_end_date, true_value) |>
+  unique() |>
+  group_by(location, target_type) |>
+  summarise(mean_value = mean(true_value), 
+            sd_value = sd(true_value), 
+            var_value = var(true_value)) |>
+  ggplot(aes(y = var_value, x = mean_value)) +
+  geom_point() + 
+  facet_wrap(~ target_type, scales = "free") + 
+  scale_x_continuous(trans = "log10", labels = label_fn) +
+  scale_y_continuous(trans = "log10", labels = label_fn) +
+  # scale_x_continuous(labels = label_fn) +
+  # scale_y_continuous(labels = label_fn) +
+  labs(y = "Variance observed value", x = "Mean observed value") + 
+  geom_function(fun = function(x) x^2, linetype = "dashed", size = 0.5) + 
+  theme_minimal()
+
+ggsave("output/figures/HUB-mean-var-obs.png", width = 7, height = 4)
+
+```
 

### 2022-01-31 main.tex
Updates from Overleaf

@@
 
+%CHECK WHY FOR LOG SCALE THE VALUES FOR 0.2 and 5 ARE NOT EQUAL
+
 \paragraph{Note on propriety and shifting the focus of the evaluation}
@@
 
-\section{Apropriate scales for epidemiological processes}
+\section{Appropriate scales for epidemiological processes}
 
-\paragraph{Modelling on the log scale}
+In this section we explore the rationale for and implications of scoring forecasts on the log scale in an epidemiological context.
+Forecast evaluation serves several purposes.
+It provides feedback for forecasters and researchers necessary to improve their models, and it helps decision makers with distinguishing good from bad predictions and choosing forecasts that inform future decisions.
 
-General point: Scores are for a) improving models, b) choosing a model for decision making. 
-At least for a): models are multiplicative, so if a model is making error, it's making a multiplicative error. We therefore should score it based on that.
+Generally, the way forecasts are evaluated should correspond to the error structure of the model and the type of errors models will usually make.
+In epidemiology, the spread of infectious diseases is usually modelled as a multiplicative process.
+COMPARTMENT MODELS.
+RENEWAL EQUATION.
+The number of infections usually grows with some exponential rate etc.
+The most important errors that models commonly make will therefore be multiplicative errors.
+It therefore makes sense to focus evaluation on these multiplicative errors.
+often, if the model is performing badly in multiplicative terms, this means that something with the model is wrong and needs fixing, regardless of performance in absolute terms.
+When selecting a model for decision-making, similar considerations apply.
+When a model is structured in a multiplicative way, we should base our trust in it on past relative performance, rather than absolute performance.
 
-%    - general point about how all processes are modelled %multiplicatively and therefore it might be odd to score them %based on absolute errors
-%    - Simulations Sam? 
-%    - Adding up several districts - how do scores change? %(Example Johannes) --> relationship between mean and variance %in scores
+\paragraph{What happens when you average things}
+In order to compare model, we usually need to compare aggregate performance.
+In many epidemiological settings, whether we decide to evaluate relative or absolute performance has implications for how we weigh performance of models across different dimensions when we aggregate and compare scores.
+For example, different forecast targets like reported case numbers, hospitalisations and deaths have very different orders of magnitude and scores (even for an ideal forecaster) usually also differ by several orders of magnitude.
+If we aggregated scores on the natural scale across e.g.
+cases and deaths, in order to rank models by overall score, an equal weight on absolute performance would implicitly result in a much higher weight on performance on cases.
+On the log scale, we would put equal weight on the relative performance on cases and on deaths, making performance usually more easily comparable (as scores of an ideal forecaster would not differ as much as on the natural scale).
+IS THAT TRUE? SHOULD PROBABLY PROVE THIS POINT BY SIMULATION.
+
+While it is intuitive that performance on cases is not orders of magnitude more important than performance on deaths, the question becomes trickier if we compare across locations or across time for the same forecast target.
+Indeed, predictive performance of a model may be more important in locations or at times when incidence values are high, as argued by \citep{Bracher}.
+Whether or not we should think this way depends on certain characteristics of the models.
+
+Let us assume that we care in particular about situations where incidence numbers are high.
+It could be useful to give a higher weight to these situations (by scoring on the natural scale), if we believe that certain models may indeed be inherently better in such situations where incidence numbers are high.
+For example, models may be better around epidemic peaks when numbers are highest, or they may be especially good at forecasting large-scale population dynamics.
+%(whatever that may be...)
+Then, scoring on the natural scale will help you identify identify the models that tend to be most situations in the situations we most care about.
+For example, when looking at one forecast target across time, it may be reasonable to assume that the model which performed best around the peak of one wave of COVID-19 may also be the one to take most seriously in the second wave of COVID-19.
+If, on the other hand, we believe that good models should do consistently well and that situations with high incidences do not provide orders of magnitude more information about which models perform well or badly in the circumstances we most care about, then scoring on the log scale may be more appropriate.
+For example, it may be reasonable not to think that predictive performance in a country like Germany would have a hundred times more importance than performance in a country like Luxembourg when determining the best model for future decision making.
+Similarly, we may not necessarily believe that predictive performance during the January 2022 wave of COVID-19 should carry three times as much weight than predictive performance during the January 2021 wave, just because numbers where three times as high.
+% Maybe make a plot with one country and scores for that one country in different waves.
+We could then argue that within one wave, performance of the peak is meaningful, but across different waves (with different numbers) it is unclear, whether that difference conveys meaningful information.
+
+When scoring forecasts on a log scale, one may fall in the opposite trap of giving to much weight to situations where incidences are low, as small errors (e.g.
+predicting 9, rather than 3 deaths) may have a much larger effect in relative terms than in absolute terms if the quantity to forecast is very small.
+How much weight is given to smaller or larger forecast quantities depends on the exact relationship between the mean and the variance of the quantity of interest.
+If the variance grows slower than the mean, then scores will on average be larger for small quantities than for larger ones.
+EXPLAIN FIGURE AND ALSO HUB EXAMPLE. 
+In practice, this can be partly mitigated by restricting the analysis to forecast targets with with higher incidences. 
+% the phrase "scores give a higher weight to a or b" could be made clearer: This is about a scenario where you average across different targets and the average score is influenced by outliers
+
+OVERALL, IT TENDS TO BE THE CASE THAT SCORING SOMETHING ON THE LOG SCALE EVENS THINGS OUT QUITE A BIT.
+
+%MAYBE A PARAGRAPH ABOUT OUTLIERS?
+%COULD ALSO MAKE A NICE PLOT ABOUT OUTLIERS, where we look at the effect of outliers. but maybe that is already included in the plot in Figure 1. 
+% could make an analysis re outliers: how do scores change if we remove the 1 or 2 worst forecasts? The one or two best forecasts?
 
@@
 
-The score is relative and therefore less dependent on the scale of the forecast quantity, making it easier to compare (relative) performance across different forecast targets, periods and locations.
-
-This means the WIS and CRPS on the natural scale are dependent on the scale of the forecast target, because forecasting errors will typically be large for large quantities. 
-However, note that large relative errors will likely more often occur with small quantities (e.g.
-predicting 9, rather than 3 deaths), potentially giving undue weight to small quantities.
-Whether or not more weight is given to smaller or larger forecast quantities depends on the exact relationship between the mean and the variance of the quantity of interest.
-If the variance grows slower than the mean, then scores will on average be larger for small quantities than for larger ones.
-In practice, this can be partly mitigated by restricting the analysis to forecast targets with with higher incidences.
-% the phrase "scores give a higher weight to a or b" could be made clearer: This is about a scenario where you average across different targets and the average score is influenced by outliers
+%Score different things from the European / US Forecast Hub and plot the relationship between mean and variance.
+Plot log-scale scores for different things and see which of these influence average scores most \\
+% What happens to the decomposition of the WIS if we log?
 
@@
 
-% could make an analysis re outliers: how do scores change if we remove the 1 or 2 worst forecasts? The one or two best forecasts?
-
-\paragraph{detailed interpretation of plots and when you give larger scores to what}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/placeholder-image.png}
-    \caption{Plot that shows the scores for log and non log and how they change}
-    \label{fig:change-in-scores}
-\end{figure}
-
-\paragraph{Option: empirical analysis} 
-Score different things from the European / US Forecast Hub and plot the relationship between mean and variance.
-Plot log-scale scores for different things and see which of these influence average scores most \\
-What happens to the decomposition of the WIS if we log?
-
-\section{Appropriate scales for epidemiological processes}
-%Could also make this more general and compare any additive, multiplicative or even other processes and state which scale would be appropriate.
-
-\paragraph{Scoring epidemiological processes}
-Epidemiological processes are generally multiplicative, others may be additive or even something else.
-We want to explore whether epidemiological processes should in principle be scored on a log scale.
-Also: are there any processes which shouldn't be scored on a log scale?
-
-\paragraph{What is the interpretation of logging in an epidemiological setting and is it appropriate?}
-
-\paragraph{Over- and under-prediction}
-If we want to keep this in, we could: 
-\begin{itemize}
-    \item Check whether there is actually a problem with over-prediction and under-prediction in the Hub.
-This could be the case because we are most interested in certain scenarios in which this might arise.
-    \item Discuss this in light of Johannes' analysis of how the decomposition of WIS values differs if the data are skewed
-    \item compare this to the PIT-value-like relative bias scores we used for the German / Polish paper, which capture a relative tendency to over- or under-predict, rather than absolute penalties. 
-\end{itemize}
-
 \paragraph{Toy example: simulate an epidemic and apply 3 different models} Compare scores for these three models on the natural scale and the log scale
-
-\paragraph{Optional: Empirical differences when going from natural to log-scale} 
-We could score a lot of forecasts of the European Hub and show plots about how scores differs when going from natural to log-scale
 
@@
 
-\paragraph{Introduction to the Forecast Hub}
+\paragraph{Introduction to the Forecast Hub} As an empirical example for evaluating forecasts on the natural and on the log scale we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021}.
+Every week the European Forecast Hub collates and aggregates forecasts for different COVID-19 related targets from teams around the world.
+Forecasts are made one to four weeks ahead into the future and follow a quantile-based format with a set of 22 quantiles plus the median ($0.01, 0.025, 0.05, ..., 0.5, ...
+0.95, 0.975, 0.99$).
+The forecasts for the purpose of this illustrations are forecasts made between XXX and XXX for reported cases and deaths from COVID-19 in XXX countries.
+
+\paragraph{Describing features of the data}
+Across all time points and locations, the mean number of cases (deaths) observed was 22811 (368) with a standard deviation of 47135 (831). 
+FIT A NEGATIVE BINOMIAL TO THESE OBSERVATIONS
 
@@
 \end{figure}
-
-\section{Modelling scores}
-\paragraph{Motivation} Alternative is to use summary measures or pairwise comparison, which becomes cumbersome for many dimensions.
-
-\paragraph{caveats, practical limitations}
-\begin{itemize}
-    \item What kind of transformations can we do? Propose random effect models
-    \item is there anything we can't do? 
-    \item How does logging change the error distribution?
-On the natural scale, you have huge outliers.
-What's the appropriate error distribution to use?
-confidence intervals.
-Maybe we can only say something about the effects
-\end{itemize}
-
-\paragraph{Optional: Application to data from the Euro-Hub}
-Also: what, empirically is the distribution of errors?
 
@@
 %\end{itemize}
+
+%\paragraph{Over- and under-prediction}
+%If we want to keep this in, we could: 
+%\begin{itemize}
+%    \item Check whether there is actually a problem with over-prediction and under-prediction in the Hub.
+This could be the case because we are most interested in certain scenarios in which this might arise.
+%    \item Discuss this in light of Johannes' analysis of how the decomposition of WIS values differs if the data are skewed
+%    \item compare this to the PIT-value-like relative bias scores we used for the German / Polish paper, which capture a relative tendency to over- or under-predict, rather than absolute penalties. 
+%\end{itemize}
+
+%\section{Modelling scores}
+%\paragraph{Motivation} Alternative is to use summary measures or pairwise comparison, which becomes cumbersome for many dimensions.
+
+%\paragraph{caveats, practical limitations}
+%\begin{itemize}
+%    \item What kind of transformations can we do? Propose random effect models
+%    \item is there anything we can't do? 
+%    \item How does logging change the error distribution?
+On the natural scale, you have huge outliers.
+What's the appropriate error distribution to use?
+confidence intervals.
+Maybe we can only say something about the effects
+%\end{itemize}
+
+%\paragraph{Optional: Application to data from the Euro-Hub}
+%Also: what, empirically is the distribution of errors?

### 2022-01-31 analysis/Analysis ECDC forecasts.Rmd
update plots

@@
   theme_minimal() + 
+  scale_y_continuous(labels = label_fn) + 
   scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) + 
@@
 # alternatively, we could think about just summarising across all models for these plots?
-scores |> 
+
+# make two separate plots to keep ordering. I.e. order locations according 
+# to log scale to illustrate that the ordering changes for cases
+
+get_order <- function(df, type = "Cases") {
+  df |> 
+    filter(scale == "natural", target_type == type) |>
+    arrange(-interval_score) |>
+    pull(location)
+}
+
+plot_df <- scores |> 
   filter(model == "EuroCOVIDhub-ensemble", 
@@
   summarise(interval_score = mean(interval_score), 
-            .groups = "drop_last") |>
-  ggplot(aes(y = interval_score, x = reorder(location, -interval_score))) + 
+            .groups = "drop_last")
+
+order_cases <- get_order(plot_df)
+order_deaths <- get_order(plot_df, type = "Deaths")
+
+plot <- function(plot_df, order) {
+  plot_df |>
+    mutate(location = factor(location, levels = order)) |>
+    ggplot(aes(y = interval_score, x = location)) + 
     geom_bar(stat = "identity") + 
@@
     scale_y_continuous(labels = scale_fn) + 
-  labs(y = "Interval score", x = "Target type")
+    scale_x_discrete(guide = guide_axis(n.dodge=2)) +
+    labs(y = "Interval score", x = "Location")
+}
+
+cases <- plot_df |>
+  filter(target_type == "Cases") |>
+  plot(order = order_cases)
+
+deaths <- plot_df |>
+  filter(target_type == "Deaths") |>
+  plot(order = order_deaths)
+
+cases / 
+  deaths
 
@@
   theme(legend.position = "none") + 
-  # scale_x_continuous(labels = scale_fn) + 
-  # scale_y_continuous(labels = scale_fn) + 
-  scale_x_continuous(labels = scale_fn, trans = "log10") +
-  scale_y_continuous(labels = scale_fn, trans = "log10") +
+  scale_x_continuous(labels = scale_fn) +
+  scale_y_continuous(labels = scale_fn) +
   labs(y = "REPLACE", x = "Total number of observed values")
@@
 
-mean_wis_vs_total <- wis_vs_total(scores)
+wis_vs_total(scores) + 
+  labs(y = "Mean WIS")
 ggsave("output/figures/HUB-mean-scores-vs-total.png", width = 7, height = 5)
 
-sd_wis_vs_total <- wis_vs_total(scores, summary_fct = stats::sd)
+wis_vs_total(scores, summary_fct = stats::sd) + 
+  labs(y = "Sd WIS")
 ggsave("output/figures/HUB-sd-scores-vs-total.png", width = 7, height = 5)
+
+# log-log versions of that plot
+wis_vs_total(scores) + 
+  labs(y = "Mean WIS") + 
+  scale_x_continuous(labels = scale_fn, trans = "log10") +
+  scale_y_continuous(labels = scale_fn, trans = "log10") 
+ggsave("output/figures/HUB-mean-scores-vs-total-log-log.png", width = 7, height = 5)
+
+wis_vs_total(scores, summary_fct = stats::sd) + 
+  labs(y = "Sd WIS") + 
+  scale_x_continuous(labels = scale_fn, trans = "log10") +
+  scale_y_continuous(labels = scale_fn, trans = "log10") 
+ggsave("output/figures/HUB-sd-scores-vs-total-log-log.png", width = 7, height = 5)
 

### 2022-01-31 main.tex

@@
 \paragraph{Describing features of the data}
-Across all time points and locations, the mean number of cases (deaths) observed was 22811 (368) with a standard deviation of 47135 (831). 
-FIT A NEGATIVE BINOMIAL TO THESE OBSERVATIONS
+Across all time points and locations, the mean number of cases (deaths) observed was 22811 (368) with an overall standard deviation of 47135 (831).
+The number of observed value does not only differ by forecast target, but also shows considerable heterogeneity by location (see Figure \ref{fig:HUB-mean-locations}.
+Across the different locations, the variance in observed values is roughly equal to the square of the mean, suggesting that, for a given location, both cases and deaths are not poisson-distributed, but rather exhibit over-dispersion.
+MAYBE FIT A NEGATIVE BINOMIAL TO THESE OBSERVATIONS
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-obs-location.png}
+    \caption{CAPTION}
+    \label{fig:HUB-mean-locations}
+\end{figure}
+
+\paragraph{Features of the scores}
 

### 2022-01-31 analysis/Analysis ECDC forecasts.Rmd
update WIS-size simulation plot

@@
 scores |>
+  dplyr::select(location, target_type, scale, target_end_date, interval_score) |>
+  unique() |>
+  group_by(target_type, scale) |>
+  summarise(mean_value = mean(interval_score), 
+            sd_value = sd(interval_score), 
+            var_value = var(interval_score))
+
+scores |> 
   filter(model == "EuroCOVIDhub-ensemble", 
@@
 
-ggsave("output/figures/HUB-average-scores.png", width = 7, height = 5)
+ggsave("output/figures/HUB-average-scores.png", width = 7, height = 3.5)
 ```
@@
 
+## Correlation between scores
+
+```{r}
+scores |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  group_by(target_type, location, scale) |>
+  summarise(interval_score = mean(interval_score), 
+            .groups = "drop_last") %>%
+  pivot_wider(values_from = interval_score, names_from = scale) %>%
+  group_by(target_type) %>%
+  summarise(correlation = cor(natural, log))
+```
+
 ## Variance scores across locations
@@
 
-  scores |> 
+  plot_df <- scores |> 
     filter(model == "EuroCOVIDhub-ensemble", 
@@
               .groups = "drop_last") |> 
-  inner_join(total_values) |>
+    inner_join(total_values)
+
+  plot_df |>
+    group_by(target_type, scale) |>
+    summarise(correlation = cor(wis, total)) |> 
+    print()
+
+  plot_df |>
     ggplot(aes(y = wis, x = total)) + 
@@
 ggsave("output/figures/HUB-sd-scores-vs-total-log-log.png", width = 7, height = 5)
+
+# correlation
 

### 2022-01-31 analysis/Simulation-poisson-mean-sd.R

@@ -3,6 +3,7 @@ library(dplyr)
 library(scoringutils)
 library(ggplot2)
 library(tidyr)
+library(patchwork)
 
 setup_df <- function(state_sizes = 10^seq(0, 2, 0.01), 
                      q = c(0.01, 0.025, seq(0.05, 0.95, 0.05), 0.975, 0.99), 
@@ -39,19 +40,72 @@ score_states <- function(df) {
 
 make_plot <- function(scores, summary_fct = mean) {
   p1 <- scores |>
-    group_by(state_size, scale) |>
+    group_by(state_size, scale, Theta) |>
     summarise(interval_score = summary_fct(interval_score)) |>
-    ggplot(aes(y = interval_score, x = state_size)) +
+    ggplot(aes(y = interval_score, x = state_size, colour = Theta)) +
     geom_point(size = 0.4) +  
     labs(y = "WIS", x = "Size of state") + 
     theme_minimal() + 
     facet_wrap(~ scale, scales = "free_y")
   
   p2 <- p1 + 
-    scale_x_continuous(trans = "log10")
+    scale_x_continuous(trans = "log10") + 
+    scale_y_continuous(trans = "log10")
   
   p1 / p2
 }
+
+
+
+
+# sample Negative Binomial -----------------------------------------------------
+# var = mu + mu^2/size --> increases a lot with mu
+# theta = size very high --> poisson distribution
+mean_county <- 100
+
+sizes_nbinom <- c(0.1, 1, 1e9)
+
+res <- list()
+for (size_nbinom in sizes_nbinom) {
+  df <- setup_df(time_points = 1000) |>
+    mutate(true_value = rnbinom(n = 1, size = size_nbinom, 
+                                mu = mean_county * state_size), 
+           prediction = qnbinom(p = quantile, size = size_nbinom, 
+                                mu = mean_county * state_size)) 
+  
+  scores <- score_states(df) |>
+    mutate(Theta = as.character(size_nbinom))
+  
+  res[[paste(size_nbinom)]] <- scores
+}
+
+saveRDS(res, file = "output/data/simulation-negative-binom.Rda")
+
+out <- rbindlist(res) |>
+  group_by(Theta, scale) |>
+  mutate(interval_score = interval_score / mean(interval_score), 
+         Theta = ifelse(Theta == "1e+09", "1b", Theta))
+
+
+make_plot(out, summary_fct = mean) +
[…]

### 2022-01-31 main.tex
Updates from Overleaf

@@
 0.95, 0.975, 0.99$).
-The forecasts for the purpose of this illustrations are forecasts made between XXX and XXX for reported cases and deaths from COVID-19 in XXX countries.
+The forecasts for the purpose of this illustrations are two-week-ahead forecasts made between XXX and XXX for reported cases and deaths from COVID-19 in XXX countries.
 
@@
 \paragraph{Features of the scores}
-
-NEED A TABLE WITH SCORES
+Across all time points and locations, the mean number weighted interval score on the natural scale for cases (deaths) was 8855 (59.3) with an overall standard deviation of 51274 (156).
+On the log scale, scores were much closer to each other (see Figure \ref{fig:HUB-average-scores}), with a mean wis for cases (deaths) of 0.507 (0.362) with an overall standard deviation of 0.745 (0.407).
+% NEED A TABLE WITH SCORES for the Appendix
+On the natural scale, scores vary by several orders of magnitudes across locations (see \ref{fig:HUB-scores-location}), whereas variation across locations is much smaller for scores on the log scale and wis values are more evenly distributed.
+The ordering of scores across locations is not preserved when going from the natural to the log scale.
+% Rather, there seems to be a slight inverse relationship that locations which previously had low scores now tend to receive higher scores. 
+The correlation between scores on the log and natural scale for different locations is XX for cases and XX for deaths. 
+%THERE SEEMS TO BE A TENDENCY FOR SMALLER
 
@@
     \includegraphics[width=0.9\textwidth]{output/figures/HUB-average-scores.png}
-    \caption{CAPTION}
+    \caption{Distribution of interval scores for two week ahead forecasts of COVID-19 cases and deaths evaluated on the natural scale (left) and on the log scale (right). }
     \label{fig:HUB-average-scores}
@@
     \includegraphics[width=0.9\textwidth]{output/figures/HUB-scores-locations.png}
-    \caption{CAPTION}
+    \caption{Weighted interval scores across different locations for two week ahead forecasts of COVID-19 cases and deaths evaluated on the natural scale (left) and on the log scale (right). }
     \label{fig:HUB-scores-location}
 \end{figure}
+
+When evaluated on the natural scale, there is  strong relationship between the mean of the weighted interval score and the total number of observed cases (correlation: 0.913) or deaths (cor: 0.849) in a location (see Figure \ref{fig:HUB-mean-scores-total-loglog} and Figure \ref{fig:HUB-mean-scores-total} in the SI).
+The relationship is less pronounced for scores on the log scale (cor: -0.019 for cases, -0.395 for deaths) and negative, meaning that locations with fewer observed cases or deaths tend to receive higher scores.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-scores-vs-total-log-log.png}
+    \caption{Plot with Weighted interval scores against the mean number of observed cases or deaths.}
+    \label{fig:HUB-mean-scores-total-loglog}
+\end{figure}
+
+Both on the natural scale as well as on the log scale, scores increase considerably with increasing forecast horizon (see Figure \ref{fig:HUB-scores-horizon}).
+The increase is more pronounced for cases than for deaths and arguably represents an increase in the relative difficulty to forecast values further into the future.
+A similar increase both on the natural and the log scale indicates that we are able to observe the increase in difficulty regardless of how we evaluate the scores.
 
@@
     \includegraphics[width=0.9\textwidth]{output/figures/HUB-scores-over-horizon.png}
-    \caption{CAPTION}
+    \caption{Weighted interval scores across forecast horizons}
     \label{fig:HUB-scores-horizon}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-scores-over-time.png}
-    \caption{CAPTION}
-    \label{fig:HUB-scores-time}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-sd-vs-mean-scores.png}
-    \caption{CAPTION}
-    \label{fig:HUB-mean-sd-scores}
 \end{figure}
@@
 \paragraph{context, implications, limitations}
+
+SOME DIFFERENCES BETWEEN SCORES ARE MEANINGFUL. E.G. WE WOULDN'T WANT SCORES FOR CASES AND DEATHS TO BE COMPLETELY EQUAL, BECAUSE DEATHS ARE LIKELY EASIER TO FORECAST
 
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-scores-vs-total.png}
+    \caption{Plot with Weighted interval scores against the mean number of observed cases or deaths.}
+    \label{fig:HUB-mean-scores-total}
+\end{figure}
+
 \end{document}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% not sure this plot is really meaningful
+%\begin{figure}[h!]
+%    \centering
+%    \includegraphics[width=0.9\textwidth]{output/figures/HUB-sd-vs-mean-scores.png}
+%    \caption{CAPTION}
+%    \label{fig:HUB-mean-sd-scores}
+%\end{figure}
 

### 2022-01-31 analysis/Analysis ECDC forecasts.Rmd
update plots

@@
   geom_function(fun = function(x) x^2, linetype = "dashed", size = 0.5) + 
-  theme_minimal()
+  theme_minimal() + 
+  geom_smooth(alpha = 0.2, size = 0.2)
 

### 2022-01-31 analysis/Simulation-poisson-mean-sd.R

@@ -42,6 +42,9 @@ make_plot <- function(scores, summary_fct = mean) {
   p1 <- scores |>
     group_by(state_size, scale, Theta) |>
     summarise(interval_score = summary_fct(interval_score)) |>
+    group_by(Theta, scale) |>
+    mutate(interval_score = interval_score / mean(interval_score), 
+           Theta = ifelse(Theta == "1e+09", "1b", Theta)) |>
     ggplot(aes(y = interval_score, x = state_size, colour = Theta)) +
     geom_point(size = 0.4) +  
     labs(y = "WIS", x = "Size of state") + 
@@ -80,25 +83,23 @@ for (size_nbinom in sizes_nbinom) {
 }
 
 saveRDS(res, file = "output/data/simulation-negative-binom.Rda")
+res <- readRDS(file = "output/data/simulation-negative-binom.Rda")
 
-out <- rbindlist(res) |>
-  group_by(Theta, scale) |>
-  mutate(interval_score = interval_score / mean(interval_score), 
-         Theta = ifelse(Theta == "1e+09", "1b", Theta))
-
+out <- rbindlist(res) 
 
 make_plot(out, summary_fct = mean) +
   plot_layout(guides = "collect") & 
   theme(legend.position = "bottom") &
-  labs(y = "WIS relative to mean")
+  labs(y = "Relative WIS")
 
-ggsave("output/figures/SIM-mean-sd-state-size.png", width = 7, height = 4)
+ggsave("output/figures/SIM-mean-state-size.png", width = 7, height = 4)
 
+make_plot(out, summary_fct = stats::sd) +
+  plot_layout(guides = "collect") &
+  theme(legend.position = "bottom") &
+  labs(y = "Realtive WIS sd")
 
-
-
-
-
+ggsave("output/figures/SIM-sd-state-size.png", width = 7, height = 4)
 
 
 

### 2022-02-07 main.tex
Updates from Overleaf

@@
 
-% General thought: maybe easier to do with WIS rather than CRPS? 
-% Could do WIS in the main paper and CRPS in the SI?
-
-% We could look at the log growth rate
-% with any linear model, an additive effect on the log growth rate would mean a multiplicative effect on the growth rate
-
-% Introduction
-%    - Role of forecasts and problems in comparing them
-%    - Introduction to scoring rules, WIS, CRPS
-%    - outline of the paper
-%Scoring on a log scale
-%    - what we propose
-%    - we approximately score the growth rate
-%    - focus is shifted from additive to multiplicative error
-%    - Interpretation: 
-%        - How do forecast weights change: what gets high and %what gets low scores? 
-%        - - Figure from Johannes: How CRPS changes when you log %the observation
-%
-%Appropriate scales for epidemiological processes
-%    - general point about how all processes are modelled %multiplicatively and therefore it might be odd to score them %based on absolute errors
-%    - Simulations Sam? 
-%    - Adding up several districts - how do scores change? %(Example Johannes) --> relationship between mean and variance %in scores
-%
-%Example: European Forecast Hub: 
-%    - paragraph introduction to the Forecast Hub
-%    - How does the distribution of scores change
-%        - target types, locations, over time
-%    - relationship between mean and sd of WIS
-
 % ===========================================================
@@
 One major challenge in comparing and evaluating these forecasts is that it is not immediately how to deal with targets on different orders of magnitudes (e.g.
-case numbers and hospitalisations) or how to take the exponential nature of epidemiological processes into account.
-In this paper we will argue that evaluating forecasts on a logarithmic, rather than a natural scale, may help alleviate some of these concerns and may be better suited to epidemiological forecasts.
+case numbers and hospitalisations) or how to take the potentially non-stationary nature of epidemiological processes into account.
+In this paper we will argue that evaluating forecasts on a logarithmic scale may be useful alongside scoring on the natural scale, and in some instances may be better suited to epidemiological forecasting contexts.
 
@@
 \begin{equation}
-WIS = \frac{1}{K + 0.5} \cdot (w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y)),    
+\text{WIS} = \frac{1}{K + 0.5} \cdot \left(w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y)\right),    
 \end{equation} 
@@
 Their relation to the absolute error means that both the CRPS and the WIS scale with the prediction target.
-Forecasts of COVID-19 cases, for example typically have much higher scores than forecasts of hospitalisations of death.
+Forecasts of COVID-19 cases, for example typically have higher scores than forecasts of hospitalisations of death.
 Similarly, when looking at performance across different locations or over time, average scores will be dominated by locations and times with high incidences.
-Single outliers also can have a disproportionate effect on aggregate scores.
-One can argue that this is meaningful and that we should care most about places and periods when incidences are high (maybe cite Bracher et al.).
+Outliers also can have a disproportionate effect on aggregate scores.
+One can argue that this is meaningful and that we should care most about places and periods when incidence is high (maybe cite Bracher et al.).
 However, this may not always be true and is clearly not desirable when comparing performance of a model on different prediction targets: case numbers are not necessarily more important than hospitalisations, just because observed values tend be an order of magnitude higher.
 This makes forecasts often hard or impossible to compare.
-Distributions of forecast scores are usually highly skewed making it hard to model them using standard linear models.
-Researchers instead often resorted to evaluating different prediction targets or locations completely independently, a process that is often cumbersome and makes it hard to identify factors that systematically affect scores.
-In any epidemiological setting there is also a noticeable tension when processes (such as the spread of a disease) that are usually modelled multiplicatively are scored in terms of absolute (additive) errors.
+Distributions of forecast scores, are also usually skewed making it hard to model them using standard linear models.
+Researchers instead often resorted to evaluating different prediction targets or locations independently, a process that is often cumbersome and makes it hard to identify factors that systematically affect scores.
+In epidemiological settings there is also a noticeable tension as processes (such as the spread of a disease) are typically considered by theory to evolve as exponential processes but using traditional proper scoring rules are evaluate in terms of absolute (additive on the linear scale) errors.
 
 \paragraph{Outline of what we do} 
-A natural proposition is therefore to score forecasts on a logarithmic scale, making it easier to model and compare scores across targets by alleviating concerns about skewed forecast scores and also taking into account the multiplicative nature of epidemiological processes.
-In this paper we will ...
+A natural proposition is therefore to score forecasts on a logarithmic scale, bringing the scoring process closer to the underlying theory of how disease transmission takes place, and making it easier to model and compare scores across targets by alleviating concerns about skewed forecast score.
+In this paper we will be amazing again and again and again...
 %\begin{enumerate}
@@
 %
-where $S()$ is either the WIS or the CRPS and one is added to deal with zeros in the data.
-The so computed score is still proper, as for a single forecast the order between different forecasters is not affected by the monotone transformation.
+where $S()$ is the WIS or the CRPS and one is added to deal with zeros in the data.
+The computed score is still proper, as for a single forecast the order between different forecasters is not affected by the monotone transformation (citation).
 % last sentence not very clear
@@
 
-\paragraph{Scoring the log means scoring the growth rate} Computing a score based on the logarithm of the forecast and the observation is approximately equivalent to scoring a forecast of a multiplicative growth rate of the forecast target.
+\paragraph{Scoring the log means scoring the approximate growth rate} Computing a score based on the logarithm of the forecast and the observation is approximately equivalent to scoring a forecast of a multiplicative growth rate of the forecast target.
 The growth rate is defined as
@@
 What is uncertain is $\log y_{t+1} - \log y_t = \log (1 + g_{t, t+1})$.
-For small values of $g$, we lastly have
+For small values of $g$, we can make the following approximation
 \begin{equation}
@@
 %
+
+This connection with the underlying epidemic growth rate is an attractive property in and of itself as it is the conceptual target many forecasters and forecast consumers make use of.
+
 When we score a forecast for $y_{t+1}$ on the log scale, the interval score for a single prediction interval changes as follows
@@
 Conveniently, the decomposition of the WIS into dispersion, over-prediction and under-prediction is preserved.
-The interpretation of the components has now changed in that they now approximately represent uncertainty, over-prediction, and under-prediction with the growth rate.
+The interpretation of the components has now changed in that they now approximately represent uncertainty, over-prediction, and under-prediction with the growth rate and so are independent of current incidence.
 
@@
 Generally, the way forecasts are evaluated should correspond to the error structure of the model and the type of errors models will usually make.
-In epidemiology, the spread of infectious diseases is usually modelled as a multiplicative process.
+In epidemiology, the spread of infectious diseases is usually modelled as an exponential process.
 COMPARTMENT MODELS.
@@
 The number of infections usually grows with some exponential rate etc.
-The most important errors that models commonly make will therefore be multiplicative errors.
-It therefore makes sense to focus evaluation on these multiplicative errors.
-often, if the model is performing badly in multiplicative terms, this means that something with the model is wrong and needs fixing, regardless of performance in absolute terms.
+Error from these models will then in general be dominated by multiplicative process errors.
+Under the assumption that these models are correctly capturing transmission dynamics inherent to the system it therefore makes sense to focus evaluation on these multiplicative errors.
+Often, if the model is performing badly in multiplicative terms, this means that there is a flaw with its assumptions that needs to be resolved, regardless of performance in absolute terms.
 When selecting a model for decision-making, similar considerations apply.
 When a model is structured in a multiplicative way, we should base our trust in it on past relative performance, rather than absolute performance.
+Though ultimately this decision is dependent on the users of the forecast who may prefer models that perform better on an absolute scale for a range of reasons (for example not over-estimating the peak size of an epidemic).
 
@@
 How much weight is given to smaller or larger forecast quantities depends on the exact relationship between the mean and the variance of the quantity of interest.
-If the variance grows slower than the mean, then scores will on average be larger for small quantities than for larger ones.
-EXPLAIN FIGURE AND ALSO HUB EXAMPLE. 
-In practice, this can be partly mitigated by restricting the analysis to forecast targets with with higher incidences. 
-% the phrase "scores give a higher weight to a or b" could be made clearer: This is about a scenario where you average across different targets and the average score is influenced by outliers
-
-OVERALL, IT TENDS TO BE THE CASE THAT SCORING SOMETHING ON THE LOG SCALE EVENS THINGS OUT QUITE A BIT.
+To investigate this we sampled forecasts from different negative binomial distributions.
+The negative binomial distribution has mean $\mu$ and variance $\sigma^2 = \mu + \mu ^2 / \theta$ and for $\lim_{\theta \to \infty}$ converges to the Poisson distribution.
+For large values of $\theta$, resulting in a variance approximately equal to the mean, forecasts for lower quantities on average received higher scores on the log scale (see Figure \ref{fig:SIM-wis-state-size-mean}.
+For $\theta = 1$ (and correspondingly, $\sigma^2 = \mu + \mu^2$, we found that scores on the log scale remained approximately constant regardless of the size of the quantity to forecast.
+For $\theta = 0.1$ (and $\sigma^2 = \mu + 10 \cdot \mu^2$), scores on the log scale increased with the quantity to forecast.
+When scored on the natural scale, a higher quantity to forecast always lead to higher scores regardless of the chosen distribution (as long as the variance grows together with the mean).
 
@@
     \includegraphics[width=0.9\textwidth]{output/figures/SIM-mean-sd-state-size.png}
-    \caption{Simulation of the effect of population size with ideal forecasts of a poisson-distributed variable.
-For each simulated state, we drew 1,000 observations from a poisson distribution with $\lambda = 100 \cdot \text{state size}$ and computed the WIS for an ideal forecast (i.e.
-the predictive distribution was poisson with $\lambda$ equal to the true $\lambda$ for every state).
+    \caption{Simulation of the effect of population size with ideal forecasts of a negative-binomially-distributed variable.
+For each simulated state, we drew 1,000 observations from a negative binomial distribution with $\mu = 100 \cdot \text{state size}$ and with values of $\theta$ equal to 0.1, 1, and 1 billion.
+The variance of the negative binomial is given as $\sigma^2 = \mu + \mu^2 / \theta$, meaning that for large theta the negative binomial distribution is equal to the poisson distribution.
+For these simulated values we computed the WIS for an ideal forecast (i.e.
+the predictive distribution was negative binomial with $\mu$ and $\theta$ equal to the true $\mu$ and $\theta$ for every state).
 Left: Mean WIS depending on state size, right: Mean WIS depending on state sizes when scored on a log scale.
-Plots for the standard deviation, rather than the mean of WIS values look almost identical.}.
-    \label{fig:log-crps-viz}
+Plots for the standard deviation, rather than the mean of WIS values look are given in Figure \ref{fig:SIM-wis-state-size-sd} in the SI.}.
+    \label{fig:SIM-wis-state-size-mean}
 \end{figure}
+
+%Other transformations: 
+%Box-Cox, square root
 
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/SIM-sd-state-size.png}
+    \caption{Simulation of the effect of population size with ideal forecasts of a negative-binomially-distributed variable.
+For each simulated state, we drew 1,000 observations from a negative binomial distribution with $\mu = 100 \cdot \text{state size}$ and with values of $\theta$ equal to 0.1, 1, and 1 billion.
+The variance of the negative binomial is given as $\sigma^2 = \mu + \mu^2 / \theta$, meaning that for large theta the negative binomial distribution is equal to the poisson distribution.
+For these simulated values we computed the WIS for an ideal forecast (i.e.
+the predictive distribution was negative binomial with $\mu$ and $\theta$ equal to the true $\mu$ and $\theta$ for every state).
+Left: Mean WIS depending on state size, right: Mean WIS depending on state sizes when scored on a log scale.
+Plots for the standard deviation, rather than the mean of WIS values look are given in Figure \ref{fig:SIM-wis-state-size-sd} in the SI.}.
+    \label{fig:SIM-wis-state-size-mean}
+\end{figure}
+
 \end{document}

### 2022-05-19 main.tex
update current version

@@
 In the abstract, this article is a very good one.
+We did a very good job of the research and got all the best results.
+People told us we were mad and we said no we are effective.
+Now gather around and listen to our thoughts you will be, shocked, scared, amazed, and astounded.
 \end{abstract}
@@
 
-\paragraph{Role of forecasts and problem in comparing them}
+Forecasts play an important role in decision-making in epidemiology and public health [Citations], as well as other areas as diverse as for example economics [Citations] or agriculture [Citations].
+Epidemiological modelling was an important tool to inform public policy throughout the COVID-19 pandemic and scenario modelling, as well as forecasts of cases, hospitalisations and deaths from COVID-19 have received widespread attention.
+Beginning in 2020, forecasts from different research institutions have been systematically collected, aggregated and evaluated by COVID-19 Forecast hubs in the US [Citation], Germany and Poland, and Europe [Citation].
 
-Forecasts play an important role in decision-making in many areas such as for example economics [Citations], agriculture [Citations] and epidemiology [Citations].
-Forecasts are most useful if they are probabilistic in nature [Citation e.g.
-Held et al.], meaning that in addition to a point prediction, forecasters also provide a predictive distribution to quantify their uncertainty.
-Through the COVID-19 pandemic forecasts of case numbers, hospitalisations and deaths have had a strong influence on public policy and there merits and limitations have been widely discussed by politicians, experts and laypeople alike.
-Probabilistic COVID-19 Forecasts from different research institutions have been systematically collected and aggregated by the US Forecast Hub [Citation], the German and Polish Forecast Hub [Citation] and the European Forecast Hub [Citation].
-One major challenge in comparing and evaluating these forecasts is that it is not immediately how to deal with targets on different orders of magnitudes (e.g.
-case numbers and hospitalisations) or how to take the potentially non-stationary nature of epidemiological processes into account.
-In this paper we will argue that evaluating forecasts on a logarithmic scale may be useful alongside scoring on the natural scale, and in some instances may be better suited to epidemiological forecasting contexts.
+We usually understand epidemiological processes to be exponential in nature and therefore most commonly describe and model them in terms of concepts like growth rates [CITATION?] or the effective reproduction number (the average number of secondary cases produced by each infected person) [CITATION].
+Consequently, errors from such models will generally be dominated by multiplicative process errors.
+Under the assumption that these models are correctly capturing the underlying inherent transmission dynamics, it would make sense to focus evaluation on these multiplicative errors.
+In the past, however, evaluation of forecasts submitted to the Forecast Hubs was mostly centred around metrics like the weighted interval score (WIS), a proper scoring rule [CITATION] which, applied to raw forecasts, captures additive, rather than multiplicative errors.
+This makes it harder to diagnose and quantify model mis-specification, but also makes it more difficult to deal with the potential non-stationarity of epidemiological processes, or to compare performance across targets that are on very different orders of magnitudes (like for example cases and hospitalisations).
+In this paper, we propose to transform forecasts and observed values prior to evaluation using the WIS in a way that shifts the focus of the evaluation towards relative, rather than absolute errors.
+We illustrate the effect of this transformation using forecasts submitted to the European COVID-19 Forecast Hub.
 
-%Forecasts should be well calibrated (i.e. should not systematically deviate from the observations) and subject to calibration should be as precise as possible (Cite Gneiting).
+\paragraph{Forecast Hub format and WIS}
+Forecasts submitted to the European (as well as other) COVID-19 Forecast Hubs are probabilistic in nature [Citation e.g.
+Held et al.], meaning that forecasters provide a full predictive distribution which quantifies their uncertainty.
+These predictive distributions where reported in the form of a set of 23 quantiles (11 symmetric prediction intervals as well as a median prediction).
+The quality of the forecasts was mainly assessed based on two grounds: first, by how well they were calibrated (judged primarily by whether the 50\% and 90\% central prediction intervals on average really covered 50\% and 90\% of observed values, respectively) and second, by how well they scored in terms of the weighted interval score [CITATION bracher].
+The WIS is a strictly proper scoring rule, meaning that it incentivises any forecaster to state their true best belief and in expectation gives the best score to the forecaster who reports a predictive distribution that is equal to the data-generating distribution.
 
-\paragraph{Context / explanation of scoring rules}
-Generally, probabilistic forecasts like those submitted to the Forecast Hubs are usually evaluated using proper scoring rules [Cite Gneiting et al.
-2007].
-A proper scoring rule incentivises any forecaster to state their true best belief and on average always gives the best score to the forecaster who reports a predictive distribution that is equal to the data-generating distribution.
-Forecasts submitted to the Forecast Hubs were required to follow a quantile-based format where forecasters would report a set of 23 quantiles (11 symmetric prediction intervals as well as a median prediction).
-A natural choice for quantile-based forecasts is the (weighted) interval score (WIS) [Cite Bracher et al.].
-
-The WIS (lower values are better) can be decomposed into a dispersion component and penalties for over- and under-prediction. For a single prediction interval, the interval score is computed as 
-\begin{equation}
- IS_\alpha(F,y) = (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u),    
-\end{equation}
+WIS values are always larger or equal than zero and lower values imply better performance.
+The WIS can be decomposed into a dispersion component and penalties for over- and under-prediction.
+For a single prediction interval, the interval score is computed as
+\begin{align}
+ IS_\alpha(F,y) &= (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u) \\
+ &= \text{dispersion} + \text{underprediction} + \text{overprediction},    
+\end{align}
 
@@
 the lower and upper bound of a single central prediction interval.
+For a single forecast interval, the interval score is the width of the prediction interval, if it covers the observed value, and the absolute error between the forecast and the lower (or upper, respectively) interval boundary.
 For a set of $K$ prediction intervals and the median $m$, the WIS is computed as a weighted sum,
@@
 
-The WIS is closely related to the continuous ranked probability score (CRPS, lower values are better) [Cite Gneiting 2007], which itself can be understood as a generalisation of the absolute error to probabilistic forecasts.
+The WIS is closely related to the continuous ranked probability score (CRPS) [Cite Gneiting 2007], which itself can be understood as a generalisation of the absolute error to probabilistic forecasts.
 For an increasing set of equally-spaced prediction intervals the WIS converges to the CRPS and shares many of its properties.
-The CRPS measures the 'distance' of the predictive distribution to the observed data-generating distribution as
-
-\begin{equation}
-    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,
-\end{equation}
-
-where y is the true observed value and F the CDF of predictive distribution. Often the following alternative representation is used
-\begin{equation}
-    \text{CRPS}(F, y) = \frac{1}{2} \mathbb{E}_{F} |X - X'| - \mathbb{E}_P |X - y|,
-\end{equation}
-
-where $X$ and $X'$ are independent realisations from the predictive distributions $F$ with finite first moment and $y$ is the true value.
-In this representation we can simply replace $X$ and $X'$ by samples sum over all possible combinations to obtain the CRPS.
-
+More information on the CRPS is provided in section \ref{crps} in the SI.
+In this paper we focus on the WIS due to its simplicity and widespread use by the COVID-19 Forecast Hubs, but arguments apply analogously to the CRPS.
 Another proper scoring rule commonly used is the log score [CITATION], which we will not discuss further in this paper.
 
-\paragraph{Problems with the WIS and CRPS in an epidemiological setting}
-Their relation to the absolute error means that both the CRPS and the WIS scale with the prediction target.
-Forecasts of COVID-19 cases, for example typically have higher scores than forecasts of hospitalisations of death.
-Similarly, when looking at performance across different locations or over time, average scores will be dominated by locations and times with high incidences.
-Outliers also can have a disproportionate effect on aggregate scores.
-One can argue that this is meaningful and that we should care most about places and periods when incidence is high (maybe cite Bracher et al.).
-However, this may not always be true and is clearly not desirable when comparing performance of a model on different prediction targets: case numbers are not necessarily more important than hospitalisations, just because observed values tend be an order of magnitude higher.
-This makes forecasts often hard or impossible to compare.
-Distributions of forecast scores, are also usually skewed making it hard to model them using standard linear models.
-Researchers instead often resorted to evaluating different prediction targets or locations independently, a process that is often cumbersome and makes it hard to identify factors that systematically affect scores.
-In epidemiological settings there is also a noticeable tension as processes (such as the spread of a disease) are typically considered by theory to evolve as exponential processes but using traditional proper scoring rules are evaluate in terms of absolute (additive on the linear scale) errors.
-
-\paragraph{Outline of what we do} 
-A natural proposition is therefore to score forecasts on a logarithmic scale, bringing the scoring process closer to the underlying theory of how disease transmission takes place, and making it easier to model and compare scores across targets by alleviating concerns about skewed forecast score.
-In this paper we will be amazing again and again and again...
-%\begin{enumerate}
-%    \item Effects / Interpretation of scoring on a log scale
-%    \item Appropriate scales for epidemiological (and possibly other) processes
-%    \item Modelling log scores
-%\end{enumerate}
+%Simplified to the case of a point prediction, we could express the future observed value $y_t$ we as the point prediction $\hat{y}_{t+1}$ plus an additive error $\varepsilon_{t+1}$:
+%
+%\begin{equation}
+%    y_{t+1} = \hat{y}_{t+1} + \varepsilon_{t+1}. 
+%¸\end{equation} 
+%Conceptually, the WIS and the CRPS evaluate the forecast based on this additive error $\varepsilon_{t+1}$.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Scoring on a log scale}
+\section{Transforming forecasts to shift the focus of the evaluation}
 
-\paragraph{What we propose}
-Proper scoring rules yield a score as a function of a forecast $x$ and a corresponding observation $y$.
-For epidemiological forecasts of positive quantities we propose to replace WIS and CRPS scores on the natural scale computed as
-%
-\begin{equation}
-S(x, y),     
-\end{equation}
-%
-by scores computed on the logarithmic scale as 
-%
-\begin{equation}
-S(\log{(x + 1)}, \log{(y + 1)}),    
-\end{equation}
-%
-where $S()$ is the WIS or the CRPS and one is added to deal with zeros in the data.
-The computed score is still proper, as for a single forecast the order between different forecasters is not affected by the monotone transformation (citation).
-% last sentence not very clear
-Differences in the ranking between forecasters only appear when averaging over multiple forecasts, as deviations from the perfect forecast get penalised differently depending on whether forecasts are scored on the log or the natural scale.
-Note that simply taking the log of a score computed on the natural scale, i.e.
-computing $\log S(x, y)$ results in an improper score and is therefore not advised.
-We illustrate this point in Figure \ref{fig:log-improper} in the SI.
-% ADD A COMMENT ABOUT LOG X + 1
-
-\paragraph{Scoring the log means scoring the approximate growth rate} Computing a score based on the logarithm of the forecast and the observation is approximately equivalent to scoring a forecast of a multiplicative growth rate of the forecast target.
+Applied to the 'raw' forecasts, the WIS measures the absolute distance between the predictive distribution and the forecast.
+We can, however, shift the focus of the evaluation by transforming the forecasts.
+Any forecast about the absolute value of a forecast target in the future implicitly also contains a statement about the growth rate of the underlying process (assuming that we have accurate knowledge of the target value today).
 The growth rate is defined as
@@
 %
-where $y_{t+1}$ is the forecast target (e.g. reported cases of COVID-19), $y_t$ is the last known observation and $g_{t, t+1}$ is the growth rate between now ($t$) and $t+1$. This can be rewritten as 
+where $y_{t+1}$ is the forecast target (e.g. reported cases of COVID-19) in the future, $y_t$ is the last known observation and $g_{t, t+1}$ is the growth rate between now ($t$) and $t+1$. 
+Using the growth rate, we can express the value of the forecast target in the future as such: 
 %
@@
 %
-If we score $\log y_{t_+1}$ rather than $y_{t+1}$, this changes to 
-%
+
+We discuss two different transformations in order to shift the focus towards evaluating the forecaster based on how well they predicted the growth relative to the present value, rather than how well they predicted the absolute value that was observed.
+The first one is to divide all observed values, and the forecasts by the last observed value.
+The second one is to take the logarithm of all forecasts and observed values.
+The computed score is still proper, as for a single forecast the order between different forecasters is not affected by the monotone transformation (citation).
+
+\paragraph{Scoring the growth rate directly}
+By dividing every observed value and every forecast by the last observed value prior to evaluation, we replace forecasting $y_{t+1}$ with forecasting  $y_{t+1} / y_t = 1 + g_{t, t+1}$.
+Just as we were previously penalised an additive error with respect the absolute value, we are now penalising an absolute error with respect to the growth rate.
+
+We can illustrate what's happening by thinking of an analogous point prediction $\hat{g_{t, t+1}}$ for the growth rate. We could then write
 \begin{equation}
-\log y_{t+1} = \log (1 + g_{t, t+1}) + \log y_t.    
+    1 + g_{t, t+1} = 1 + \hat{g}_{t, t+1} + \varepsilon^*_{t+1}. 
 \end{equation}
 %
-Note that $\log y_t$ is a known quantity that can be removed from both sides without changing the score.
-What is uncertain is $\log y_{t+1} - \log y_t = \log (1 + g_{t, t+1})$.
+The WIS and CRPS would evaluate the forecast based on the error $\varepsilon^*_{t+1}$.
+This translates to an absolute error on the actual forecast target that is multiplicatively connected to the last observed value:
+\begin{align}
+    y_{t+1} &= (1 + \hat{g}_{t, t+1} + \varepsilon^*_{t+1}) \cdot y_t \\
+            &= (1 + \hat{g}_{t, t+1}) \cdot y_t + \varepsilon^*_{t+1} \cdot y_t \\
+            &= \hat{y}_{t+1} + \varepsilon^*_{t+1} \cdot y_t.
+\end{align}
+
+One disadvantage of this is that we can't really do anything when an observation is zero. 
+%is this helpful?%
+%Maybe add some math on how we actually transform the distribution. E.g. divide every quantile or sample by the last observed value, divide the distribution?%
+
+\paragraph{Taking the logarithm of forecasts}
+Taking the logarithm of both the forecasts and the observed values means that we are now scoring an additive error on $\log y_{t_+1}$ rather than on $y_{t+1}$ (see a detailed derivation in section \ref{wis-log-derivation} in the SI).
+
+Using again the analogy of a point forecast it is easy to see that this corresponds to evaluating a multiplicative error on the original $y_{t+1}$:
+%
+\begin{align}
+\log y_{t+1} &= \log \hat{y}_{t+1} + \varepsilon_{t+1} \Leftrightarrow \\    
+y_{t+1} &= (1 + \hat{y}_{t+1}) \cdot \exp{\varepsilon_{t+1}}.    
+\end{align}
+%
+Computing a score based on the logarithm of the forecast and the observation is approximately equivalent to scoring a forecast of a multiplicative growth rate of the forecast target.
+
+To get an intuition, using again the analogy of point forecasts, note that we can decompose the error into a known and an unknown part: 
+%
+\begin{align}
+\log y_{t+1} &= \log (1 + \hat{g}_{t, t+1}) + \log y_t + \varepsilon_{t+1} \Leftrightarrow \\
+\log y_{t+1} &= \log (1 + \hat{g}_{t, t+1}) + \log y_t - \log y_t + \varepsilon'_{t+1} \Leftrightarrow \\    
+\log y_{t+1} &= \log (1 + \hat{g}_{t, t+1}) + \varepsilon'_{t+1}.
+\end{align}
+%
+% NOT SURE THIS ARGUMENT WORKS. 
+%
+What affects the score is only the unknown part, as shifting both the forecast and the observation by a known amount does not make a difference for evaluating the distance between forecast and observation.
+The WIS and CRPS therefore provide a measurement of an additive error on the forecast of $ \log (1 + \hat{g}_{t, t+1})$.
 For small values of $g$, we can make the following approximation
@@
 
-This connection with the underlying epidemic growth rate is an attractive property in and of itself as it is the conceptual target many forecasters and forecast consumers make use of.
+Conveniently, the decomposition of the WIS into dispersion, overprediction and underprediction is preserved for both transformations.
+The interpretation of the components has now changed in that they now (approximately) represent uncertainty, over-prediction, and under-prediction with the growth rate and so are independent of current incidence.
 
-When we score a forecast for $y_{t+1}$ on the log scale, the interval score for a single prediction interval changes as follows
-%
-\begin{align}
-    \text{IS}_\alpha(\log F, \log y) = &(\log u - \log l) \\ 
-    &+ \frac{2}{\alpha} \cdot (\log l - \log y) \cdot 1(\log y \leq \log l) \\
-    &+ \frac{2}{\alpha} \cdot (\log y - \log u) \cdot 1(\log y \geq \log u).
-\end{align}
-%
-Note that the values of the indicator functions stay the same as the logarithm is a monotonic transformation and the inequality is preserved. 
-On the natural scale, $u$, $l$ are the lower and upper bounds of a central prediction interval for a forecast of $y_{t+1} = (1 + g_{t, t+1}) \cdot y_t$. 
-We can rewrite these as 
-\begin{equation}
-   l = l^* \cdot y_t 
-   \;\; \text{and} \;\; 
-   u = u^* \cdot y_t,  
-\end{equation}
-where $u^*$ and $l^*$ are the lower and upper bounds of a prediction interval for $(1 + g_{t, t+1})$. Consequently, 
-\begin{equation}
-  \log l = \log l^* + \log y_t
-  \;\; \text{and} \;\;
-  \log u = \log u^* + \log y_t. 
-\end{equation}
-%
-We can then rewrite the score for a forecast of $\log y_{t+1} = \log (1 + g_{t,t+1}) + \log y_t$ as
-%
-\begin{equation}
-\begin{aligned}
-IS_\alpha&(\log F_{t+1}, \log y_{t+a}) = \\
-&((\log u^* + \log y_t) - (\log l^* + \log y_t)) \\
-&+ \frac{2}{\alpha} \cdot ((\log l^* + \log y_t) - (\log (1 + g_{t, t+1}) + \log y_t)) 
-     \cdot 1((\log (1 + g_{t, t+1}) + \log y_t) \leq (\log l^* + \log y_t) \\
-&+ \frac{2}{\alpha} \cdot ((\log (1 + g_{t, t+1}) + \log y_t) - (\log u^* + \log y_t)) 
-      \cdot 1((\log (1 + g_{t, t+1}) + \log y_t) \geq (\log u^* + \log y_t),
-\end{aligned}
-\end{equation}
-%
-which simplifies to 
-%
-\begin{equation}
-\begin{aligned}
-\label{eqn:is-log}
-\text{IS}_\alpha(\log F_{t+1}, \log y_{t+a}) = &(\log u^* - \log l^*) \\
-&+ \frac{2}{\alpha} \cdot (\log l^* - \log (1 + g_{t, t+1}) 
-     \cdot 1(\log (1 + g_{t, t+1}) \leq \log l^*) \\
-&+ \frac{2}{\alpha} \cdot (\log (1 + g_{t, t+1}) - \log u^*) 
-      \cdot 1(\log (1 + g_{t, t+1}) \geq \log u^*).
-\end{aligned}
-\end{equation}
+\paragraph{First transform, then score, rather than the other way round}
+Note that simply taking the log of a score computed on the natural scale, i.e.
+computing $\log S(x, y)$ results in an improper score and is therefore not advised.
+We illustrate this point in Figure \ref{fig:log-improper} in the SI.
 
-The left-hand side of equation \ref{eqn:is-log} was left unchanged.
-From the right-hand side we can see that for a single prediction interval, scoring the log of a forecast ($\log F_{t+1}$) against the log observed value ($\log y_{t+1}$) is exactly equivalent to evaluating a forecast for $\log (1 + g_{t, t+1})$.
-With $\log (1 + g_{t, t+1} \approx g_{t, t+1}$, this is approximately a forecast for the growth rate for small values of $g_{t, t+1}$.
-Conveniently, the decomposition of the WIS into dispersion, over-prediction and under-prediction is preserved.
-The interpretation of the components has now changed in that they now approximately represent uncertainty, over-prediction, and under-prediction with the growth rate and so are independent of current incidence.
-
-\paragraph{Additive vs. multiplicative errors}
-Of course the growth rate (and more generally the term $\log (1 + g_{t, t+1})$) would be linked multiplicatively to the observed value on the natural scale.
-Recall that the WIS is closely linked to the absolute error.
-On the natural scale, we were conceptually evaluating an absolute error of our forecast with regards to the true observed value.
-Simplified to the case of a point forecast $\hat{y}_{t+1}$ we could express this as measuring an absolute error $\varepsilon_{t+1}$ such that
-%
-\begin{equation}
-    \hat{y}_{t+1} = y_{t+1} + \varepsilon_{t+1}. 
-¸\end{equation}
-In the case of scoring the log forecast, where we evaluate an absolute error on $\log (1 + g_{t, t+1})$, it makes more sense to understand the error $y_{t+1}$ as being a multiplicative error $\varepsilon^*_{t+1}$ so that
-\begin{align}
-    \hat{y}_{t+1} &= y_{t+1} \cdot \varepsilon^*_{t+1} \\
-    &= y_{t} \cdot (1 + g_{t, t+1}) \cdot \varepsilon^*_{t+1},
-\end{align}
-%
-which on the log scale becomes to 
-\begin{align}
-    \hat{y}_{t+1} = \log y_{t} + \log (1 + g_{t, t+1}) + \varepsilon^*_{t+1}. 
-\end{align}
+\paragraph{How scores change}
+For a single forecast, the ranking between forecasters would be preserved and the scores remain proper [Proof?].
+Differences in the ranking between forecasters only appear when averaging over multiple forecasts, as deviations from the perfect forecast get penalised differently depending on whether forecasts are transformed or not.
 
@@
 
-%CHECK WHY FOR LOG SCALE THE VALUES FOR 0.2 and 5 ARE NOT EQUAL
-
-\paragraph{Note on propriety and shifting the focus of the evaluation}
-Applying the log transformation preservers propriety of the scoring rule.
-The task of the forecaster has not changed and in both scoring scenarios, forecasters are incentivised to report their best possible forecast.
-Any forecast that is made already implicitly contains a statement about the growth rate of a process as well as about the expected absolute order of magnitude of the target.
-By scoring based on the log transformed forecasts and observations, we merely choose to shift the attention towards evaluating the forecaster based on relative errors, rather than absolute numbers.
-
 \section{Appropriate scales for epidemiological processes}
 
-In this section we explore the rationale for and implications of scoring forecasts on the log scale in an epidemiological context.
+THE NEXT PARAGRAPH COULD BE MOVED TO THE INTRODUCTION?
+
+In this section we explore the rationale for and implications of scoring transformed forecasts in an epidemiological context in greater depth.
 Forecast evaluation serves several purposes.
@@
 
+SHOULD INSTEAD TALK ABOUT WHAT THE DIFFERNCE IS BETWEEN LOG AND DIVIDING?
+
 \paragraph{What happens when you average things}
-In order to compare model, we usually need to compare aggregate performance.
+In order to compare models, we usually need to compare aggregate performance.
 In many epidemiological settings, whether we decide to evaluate relative or absolute performance has implications for how we weigh performance of models across different dimensions when we aggregate and compare scores.
@@
 
+\subsection{Additional information on the CRPS} \label{crps}
+
+The CRPS measures the 'distance' of the predictive distribution to the observed data-generating distribution as
+
+\begin{equation}
+    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,
+\end{equation}
+
+where y is the true observed value and F the CDF of predictive distribution. Often the following alternative representation is used
+\begin{equation}
+    \text{CRPS}(F, y) = \frac{1}{2} \mathbb{E}_{F} |X - X'| - \mathbb{E}_P |X - y|,
+\end{equation}
+
+where $X$ and $X'$ are independent realisations from the predictive distributions $F$ with finite first moment and $y$ is the true value.
+In this representation we can simply replace $X$ and $X'$ by samples sum over all possible combinations to obtain the CRPS.
+
+\subsection{Changes in the WIS for forecasts on the log scale} \label{wis-log-derivation}
+
+This connection with the underlying epidemic growth rate is an attractive property in and of itself as it is the conceptual target many forecasters and forecast consumers make use of.
+
+When we score a forecast for $y_{t+1}$ on the log scale, the interval score for a single prediction interval changes as follows
+%
+\begin{align}
+    \text{IS}_\alpha(\log F, \log y) = &(\log u - \log l) \\ 
+    &+ \frac{2}{\alpha} \cdot (\log l - \log y) \cdot 1(\log y \leq \log l) \\
+    &+ \frac{2}{\alpha} \cdot (\log y - \log u) \cdot 1(\log y \geq \log u).
+\end{align}
+%
+Note that the values of the indicator functions stay the same as the logarithm is a monotonic transformation and the inequality is preserved. 
+On the natural scale, $u$, $l$ are the lower and upper bounds of a central prediction interval for a forecast of $y_{t+1} = (1 + g_{t, t+1}) \cdot y_t$. 
+We can rewrite these as 
+\begin{equation}
+   l = l^* \cdot y_t 
+   \;\; \text{and} \;\; 
+   u = u^* \cdot y_t,  
+\end{equation}
+where $u^*$ and $l^*$ are the lower and upper bounds of a prediction interval for $(1 + g_{t, t+1})$. Consequently, 
+\begin{equation}
+  \log l = \log l^* + \log y_t
+  \;\; \text{and} \;\;
+  \log u = \log u^* + \log y_t. 
+\end{equation}
+%
+We can then rewrite the score for a forecast of $\log y_{t+1} = \log (1 + g_{t,t+1}) + \log y_t$ as
+%
+\begin{equation}
+\begin{aligned}
+IS_\alpha&(\log F_{t+1}, \log y_{t+a}) = \\
+&((\log u^* + \log y_t) - (\log l^* + \log y_t)) \\
+&+ \frac{2}{\alpha} \cdot ((\log l^* + \log y_t) - (\log (1 + g_{t, t+1}) + \log y_t)) 
+     \cdot 1((\log (1 + g_{t, t+1}) + \log y_t) \leq (\log l^* + \log y_t) \\
+&+ \frac{2}{\alpha} \cdot ((\log (1 + g_{t, t+1}) + \log y_t) - (\log u^* + \log y_t)) 
+      \cdot 1((\log (1 + g_{t, t+1}) + \log y_t) \geq (\log u^* + \log y_t),
+\end{aligned}
+\end{equation}
+%
+which simplifies to 
+%
+\begin{equation}
+\begin{aligned}
+\label{eqn:is-log}
+\text{IS}_\alpha(\log F_{t+1}, \log y_{t+a}) = &(\log u^* - \log l^*) \\
+&+ \frac{2}{\alpha} \cdot (\log l^* - \log (1 + g_{t, t+1}) 
+     \cdot 1(\log (1 + g_{t, t+1}) \leq \log l^*) \\
+&+ \frac{2}{\alpha} \cdot (\log (1 + g_{t, t+1}) - \log u^*) 
+      \cdot 1(\log (1 + g_{t, t+1}) \geq \log u^*).
+\end{aligned}
+\end{equation}
+
+The left-hand side of equation \ref{eqn:is-log} was left unchanged.
+From the right-hand side we can see that for a single prediction interval, scoring the log of a forecast ($\log F_{t+1}$) against the log observed value ($\log y_{t+1}$) is exactly equivalent to evaluating a forecast for $\log (1 + g_{t, t+1})$.
+With $\log (1 + g_{t, t+1} \approx g_{t, t+1}$, this is approximately a forecast for the growth rate for small values of $g_{t, t+1}$.
+
+\subsection{Additional figures}
+
 \begin{figure}[h!]
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \paragraph{Problems with the WIS and CRPS in an epidemiological setting}
+% Their relation to the absolute error means that both the WIS and the CRPS usually scale with the prediction target.
+Forecasts of COVID-19 cases, for example typically have higher scores than forecasts of hospitalisations of death.
+Similarly, when looking at performance across different locations or over time, average scores will be dominated by locations and times with high incidences.
+Outliers also can have a disproportionate effect on aggregate scores.
+One can argue that this is meaningful and that we should care most about places and periods when incidence is high (maybe cite Bracher et al.).
+However, this may not always be true and is clearly not desirable when comparing performance of a model on different prediction targets: case numbers are not necessarily more important than hospitalisations, just because observed values tend be an order of magnitude higher.
+This makes forecasts often hard or impossible to compare.
 

### 2022-06-04 main.tex
latest update

@@
 Forecasts play an important role in decision-making in epidemiology and public health [Citations], as well as other areas as diverse as for example economics [Citations] or agriculture [Citations].
-Epidemiological modelling was an important tool to inform public policy throughout the COVID-19 pandemic and scenario modelling, as well as forecasts of cases, hospitalisations and deaths from COVID-19 have received widespread attention.
+Epidemiological modelling has been an important tool to inform public policy throughout the COVID-19 pandemic and scenario modelling, as well as forecasts of cases, hospitalisations and deaths from COVID-19 have received widespread attention.
 Beginning in 2020, forecasts from different research institutions have been systematically collected, aggregated and evaluated by COVID-19 Forecast hubs in the US [Citation], Germany and Poland, and Europe [Citation].

### 2022-06-05 main.tex
rewrite manuscript

--- 2022-05-19 main.tex
@@
 \usepackage[colorlinks=false]{hyperref} 
+\setlength{\parskip}{\baselineskip}%
+\setlength{\parindent}{0pt}%
 \urlstyle{same}
@@
 
-Forecasts play an important role in decision-making in epidemiology and public health [Citations], as well as other areas as diverse as for example economics [Citations] or agriculture [Citations].
-Epidemiological modelling was an important tool to inform public policy throughout the COVID-19 pandemic and scenario modelling, as well as forecasts of cases, hospitalisations and deaths from COVID-19 have received widespread attention.
+Probabilistic forecasts play an important role in decision-making in epidemiology and public health [Citations], as well as other areas as diverse as for example economics [Citations] or agriculture [Citations].
+Epidemiological modelling in particular has received widespread attention during the COVID-19 pandemic and has been an important tool to inform public policy.
+Evaluating forecasts is important, as the evaluation provides feedback for forecasters and researchers necessary to improve their models, and it helps decision makers with distinguishing good from bad predictions and choosing forecasts that inform future decisions.
+
+Proabbilistic forecasts are usually evaluated using so-called proper scoring rules, which return a numerical score as a function of the forecast and the later observed data.
+Forecasters, in expectaction, optimise their score by providing a predictive distribution that is equal to the data-generating distribution, and will be penalised if they fail to do so.
+Two proper scoring rules commonly used in epidemiology and other fields are the Continuous Ranked Probability Score (CRPS) [Gneiting 2007] and the Weighted Interval Score (WIS) [CITATION].
+Due to their widespread use we focus on these two in this paper, ignoring other proper scoring rules.
+Both the WIS and CRPS represent a measure of the absolute distance between forecast and observation and can be thought of as a generalisation of the absolute error to full predictive distributions (see details in the SI).
+
+Evaluating forecasts based on absolute errors (for simplicity we use the term interchangeably with absolute distance, even though it technically only applies to point forecasts) may not always be desirable, especially in an epidemiological context.
+% The reason for this is that generally, the way forecasts are evaluated should correspond to the error structure of the model and the type of errors models will usually make.
+Epidemiological processes are usually understood to be exponential in nature and therefore are most commonly described and modeled in multiplicative terms.
+Under the assumption that these models are correctly capturing the underlying inherent transmission dynamics, it would therefore make sense to focus evaluation on multiplicative, rather than absolute errors.
+Often, if the model is performing badly in multiplicative terms, this means that there is a flaw with its assumptions that needs to be resolved, regardless of performance in absolute terms.
+Evaluating epidemiological forecasts based on a measure of absolute error (as is common practice) can make it harder to diagnose and quantify model mis-specification.
+It also may make it more difficult to deal with the potential non-stationarity of epidemiological processes, or to compare performance across targets that are on very different orders of magnitudes (like for example cases and hospitalisations).
+
+In this paper we argue that it may therefore often be helpful to shift the focus of the evaluation on multiplicative, rather than absolute errors.
+This can be achieved by transforming the forecasts as well as the observations using the logarithm before applying the WIS or CRPS.
+While we will mostly focus on the log transformation in this paper, the point we like to make is a broader one: it is possible to flexible change the focus of the evaluation by applying a multitude of different transformations before scoring forecasts.
+Instead of scoring multiplicative errors, one may for example like to apply a transformation that allows to score forecasters based on the shape of the trajectory they predicted.
+Furthermore, it is possible to capture different aspects in a single number by constructing composite scores as a weighted sum of scores based on different transformations.
+This allows forecast consumers to assign explicit weights to different qualities of the forecasts they deem to be important when assessing predictive performance.
+
+In section [REF] we discuss which kinds of transformations can be applied.
+In section [REF] we give a detailed theoretical explanation of the rationale for applying the log transformation and discuss its interpretation.
+In section [REF] we present an applied example using forecasts submitted to the European COVID-19 Forecast Hub.
+The European COVID-19 Forecast Hub is one of several Forecast hubs that systematically collected, aggregated and evaluated forecasts of several COVID-19 targets created by different teams.
+
+REWRITE HERE. 
 Beginning in 2020, forecasts from different research institutions have been systematically collected, aggregated and evaluated by COVID-19 Forecast hubs in the US [Citation], Germany and Poland, and Europe [Citation].
+Forecasts submitted to the COVID-19 Forecast Hubs are probabilistic in nature [Citation e.g.
+Held et al.], meaning that forecasters provide a full predictive distribution which quantifies their uncertainty.
+These predictive distributions are reported in the form of a set of 23 quantiles (11 symmetric prediction intervals as well as a median prediction).
 
-We usually understand epidemiological processes to be exponential in nature and therefore most commonly describe and model them in terms of concepts like growth rates [CITATION?] or the effective reproduction number (the average number of secondary cases produced by each infected person) [CITATION].
-Consequently, errors from such models will generally be dominated by multiplicative process errors.
-Under the assumption that these models are correctly capturing the underlying inherent transmission dynamics, it would make sense to focus evaluation on these multiplicative errors.
-In the past, however, evaluation of forecasts submitted to the Forecast Hubs was mostly centred around metrics like the weighted interval score (WIS), a proper scoring rule [CITATION] which, applied to raw forecasts, captures additive, rather than multiplicative errors.
-This makes it harder to diagnose and quantify model mis-specification, but also makes it more difficult to deal with the potential non-stationarity of epidemiological processes, or to compare performance across targets that are on very different orders of magnitudes (like for example cases and hospitalisations).
-In this paper, we propose to transform forecasts and observed values prior to evaluation using the WIS in a way that shifts the focus of the evaluation towards relative, rather than absolute errors.
-We illustrate the effect of this transformation using forecasts submitted to the European COVID-19 Forecast Hub.
+% The WIS can be thought of as an approximation of the CRPS for forecasts where the predictive distribution is represented through a set of quantiles.
+The CRPS, and by extension the WIS, represents a generalisation of the absolute error to continuous distributions, meaning that forecasters are evaluated on the absolute distance of their predictive distribution to the observed value (i.e.
+some form of absolute error).
+For simplicity, we will use absolute error and absolute distance between forecast and observation interchangeably, even though the former applies to point predictions and the latter is used for probabilistic forecasts.
 
-\paragraph{Forecast Hub format and WIS}
-Forecasts submitted to the European (as well as other) COVID-19 Forecast Hubs are probabilistic in nature [Citation e.g.
-Held et al.], meaning that forecasters provide a full predictive distribution which quantifies their uncertainty.
-These predictive distributions where reported in the form of a set of 23 quantiles (11 symmetric prediction intervals as well as a median prediction).
-The quality of the forecasts was mainly assessed based on two grounds: first, by how well they were calibrated (judged primarily by whether the 50\% and 90\% central prediction intervals on average really covered 50\% and 90\% of observed values, respectively) and second, by how well they scored in terms of the weighted interval score [CITATION bracher].
-The WIS is a strictly proper scoring rule, meaning that it incentivises any forecaster to state their true best belief and in expectation gives the best score to the forecaster who reports a predictive distribution that is equal to the data-generating distribution.
+% 
+% 
+% 
+% We argue that when dealing with epidemiological processes it makes sense to shift the evaluation towards multiplicative errors by either taking the logarithm of both the forecast and the observations or by dividing them by the last known observation before scoring.
 
-WIS values are always larger or equal than zero and lower values imply better performance.
-The WIS can be decomposed into a dispersion component and penalties for over- and under-prediction.
-For a single prediction interval, the interval score is computed as
-\begin{align}
- IS_\alpha(F,y) &= (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u) \\
- &= \text{dispersion} + \text{underprediction} + \text{overprediction},    
-\end{align}
-
-where $1()$ is the indicator function, $y$ is the observed value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution $F$, i.e.
-the lower and upper bound of a single central prediction interval.
-For a single forecast interval, the interval score is the width of the prediction interval, if it covers the observed value, and the absolute error between the forecast and the lower (or upper, respectively) interval boundary.
-For a set of $K$ prediction intervals and the median $m$, the WIS is computed as a weighted sum,
-\begin{equation}
-\text{WIS} = \frac{1}{K + 0.5} \cdot \left(w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y)\right),    
-\end{equation} 
-where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$
-
-The WIS is closely related to the continuous ranked probability score (CRPS) [Cite Gneiting 2007], which itself can be understood as a generalisation of the absolute error to probabilistic forecasts.
-For an increasing set of equally-spaced prediction intervals the WIS converges to the CRPS and shares many of its properties.
-More information on the CRPS is provided in section \ref{crps} in the SI.
-In this paper we focus on the WIS due to its simplicity and widespread use by the COVID-19 Forecast Hubs, but arguments apply analogously to the CRPS.
+% What kinds of transformations can be considered meaningful depends on the context.
+We will illustrate the idea of transforming forecasts by focussing on the WIS as it is currently used in epidemiological settings such as the COVID-19 Forecast hubs.
+% In this paper we focus on the WIS due to its simplicity and widespread use by the COVID-19 Forecast Hubs, but arguments apply analogously to the CRPS.
 Another proper scoring rule commonly used is the log score [CITATION], which we will not discuss further in this paper.
-
-%Simplified to the case of a point prediction, we could express the future observed value $y_t$ we as the point prediction $\hat{y}_{t+1}$ plus an additive error $\varepsilon_{t+1}$:
-%
-%\begin{equation}
-%    y_{t+1} = \hat{y}_{t+1} + \varepsilon_{t+1}. 
-%¸\end{equation} 
-%Conceptually, the WIS and the CRPS evaluate the forecast based on this additive error $\varepsilon_{t+1}$.
 
@@
 
-Applied to the 'raw' forecasts, the WIS measures the absolute distance between the predictive distribution and the forecast.
-We can, however, shift the focus of the evaluation by transforming the forecasts.
-Any forecast about the absolute value of a forecast target in the future implicitly also contains a statement about the growth rate of the underlying process (assuming that we have accurate knowledge of the target value today).
-The growth rate is defined as
-%
-\begin{equation}
-    g_{t, t+1} = \frac{y_{t+1} - y_t}{y_t},
-\end{equation}
-%
-where $y_{t+1}$ is the forecast target (e.g. reported cases of COVID-19) in the future, $y_t$ is the last known observation and $g_{t, t+1}$ is the growth rate between now ($t$) and $t+1$. 
-Using the growth rate, we can express the value of the forecast target in the future as such: 
-%
-\begin{equation}
-y_{t+1} = (1 + g_{t, t+1}) \cdot y_t.
-\end{equation}
-%
+\subsection{Permissible transformations / propriety} - JOHANNES
 
-We discuss two different transformations in order to shift the focus towards evaluating the forecaster based on how well they predicted the growth relative to the present value, rather than how well they predicted the absolute value that was observed.
-The first one is to divide all observed values, and the forecasts by the last observed value.
-The second one is to take the logarithm of all forecasts and observed values.
-The computed score is still proper, as for a single forecast the order between different forecasters is not affected by the monotone transformation (citation).
+In order to preserve the propriety of the score, transformations must be strictly monotonic.
+This makes sure that a forecast that is further away from the observed value will also be further away after the transformation.
 
-\paragraph{Scoring the growth rate directly}
-By dividing every observed value and every forecast by the last observed value prior to evaluation, we replace forecasting $y_{t+1}$ with forecasting  $y_{t+1} / y_t = 1 + g_{t, t+1}$.
-Just as we were previously penalised an additive error with respect the absolute value, we are now penalising an absolute error with respect to the growth rate.
-
-We can illustrate what's happening by thinking of an analogous point prediction $\hat{g_{t, t+1}}$ for the growth rate. We could then write
-\begin{equation}
-    1 + g_{t, t+1} = 1 + \hat{g}_{t, t+1} + \varepsilon^*_{t+1}. 
-\end{equation}
-%
-The WIS and CRPS would evaluate the forecast based on the error $\varepsilon^*_{t+1}$.
-This translates to an absolute error on the actual forecast target that is multiplicatively connected to the last observed value:
-\begin{align}
-    y_{t+1} &= (1 + \hat{g}_{t, t+1} + \varepsilon^*_{t+1}) \cdot y_t \\
-            &= (1 + \hat{g}_{t, t+1}) \cdot y_t + \varepsilon^*_{t+1} \cdot y_t \\
-            &= \hat{y}_{t+1} + \varepsilon^*_{t+1} \cdot y_t.
-\end{align}
-
-One disadvantage of this is that we can't really do anything when an observation is zero. 
-%is this helpful?%
-%Maybe add some math on how we actually transform the distribution. E.g. divide every quantile or sample by the last observed value, divide the distribution?%
-
-\paragraph{Taking the logarithm of forecasts}
-Taking the logarithm of both the forecasts and the observed values means that we are now scoring an additive error on $\log y_{t_+1}$ rather than on $y_{t+1}$ (see a detailed derivation in section \ref{wis-log-derivation} in the SI).
-
-Using again the analogy of a point forecast it is easy to see that this corresponds to evaluating a multiplicative error on the original $y_{t+1}$:
-%
-\begin{align}
-\log y_{t+1} &= \log \hat{y}_{t+1} + \varepsilon_{t+1} \Leftrightarrow \\    
-y_{t+1} &= (1 + \hat{y}_{t+1}) \cdot \exp{\varepsilon_{t+1}}.    
-\end{align}
-%
-Computing a score based on the logarithm of the forecast and the observation is approximately equivalent to scoring a forecast of a multiplicative growth rate of the forecast target.
-
-To get an intuition, using again the analogy of point forecasts, note that we can decompose the error into a known and an unknown part: 
-%
-\begin{align}
-\log y_{t+1} &= \log (1 + \hat{g}_{t, t+1}) + \log y_t + \varepsilon_{t+1} \Leftrightarrow \\
-\log y_{t+1} &= \log (1 + \hat{g}_{t, t+1}) + \log y_t - \log y_t + \varepsilon'_{t+1} \Leftrightarrow \\    
-\log y_{t+1} &= \log (1 + \hat{g}_{t, t+1}) + \varepsilon'_{t+1}.
-\end{align}
-%
-% NOT SURE THIS ARGUMENT WORKS. 
-%
-What affects the score is only the unknown part, as shifting both the forecast and the observation by a known amount does not make a difference for evaluating the distance between forecast and observation.
-The WIS and CRPS therefore provide a measurement of an additive error on the forecast of $ \log (1 + \hat{g}_{t, t+1})$.
-For small values of $g$, we can make the following approximation
-\begin{equation}
-    \log (1+ g_{t, t+1}) \approx g_{t, t+1}.
-\end{equation}
-%
-
-Conveniently, the decomposition of the WIS into dispersion, overprediction and underprediction is preserved for both transformations.
-The interpretation of the components has now changed in that they now (approximately) represent uncertainty, over-prediction, and under-prediction with the growth rate and so are independent of current incidence.
+- WHAT KINDS OF DATA CAN I CONDITION ON? 
+- WOULD DIVIDING A FORECAST BY ITS RESOLUTION WORK? 
+- CAN I SCORE THE TRAJECTORY / WEEK-TO-WEEK GROWTH RATE BY DIVIDING the 4 week ahead forecast by the three week ahead forecast?
 
@@
 
-\paragraph{How scores change}
-For a single forecast, the ranking between forecasters would be preserved and the scores remain proper [Proof?].
+% \paragraph{Changes only come when you average across multiple scores}
+% For a single forecast, the ranking between forecasters would be preserved and the scores remain proper [Proof?].
 Differences in the ranking between forecasters only appear when averaging over multiple forecasts, as deviations from the perfect forecast get penalised differently depending on whether forecasts are transformed or not.
 
-The log transformation therefore changes which types of errors (additive or multiplicative) are penalised.
+\paragraph{Example transformations to score the growth rate}
+
+Take the log
+
+Divide by the latest known value
+
+Divide by the previous for each time value, which gives you a weekly growth rate, i.e. a trajectory.
+
+Box-Cox
+
+square root
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Log transformation}
+
+As explained in the introduction, it may be useful to assess epidemiological forecasts based on multiplicative rather than absolute errors, i.e.
+based on how good they predict relative changes in the quantity of interest as apposed to how good they forecast absolute values.
+For count data, as are common in epidemiology, we can achieve this by taking the logarithm of both the forecasts and the observed values.
+
+To illustrate this intuitively, let us consider a point forecast $\hat{y}_{t+1}$ for a quantity of interest $y_{t+1}$, such that 
+\begin{equation}
+y_{t+1} = \hat{y}_{t+1} + \varepsilon_{t+1}.
+\end{equation}
+
+The score will be determined by the size of the forecast error $\varepsilon_{t+1}$.
+When taking the logarithm of the forecast and the observation first, than the score is determined based on the absolute error $\varepsilon^*_{t+1}$:
+\begin{equation}
+\log y_{t+1} = \log \hat{y}_{t+1} + \varepsilon^*_{t+1}.
+\end{equation}
+
+It is easy to see that this corresponds to evaluating a multiplicative error on the original $y_{t+1}$:
+%
+\begin{align}
+\log y_{t+1} &= \log \hat{y}_{t+1} + \varepsilon^*_{t+1} \Leftrightarrow \\    
+y_{t+1} &= \hat{y}_{t+1} \cdot \exp{\varepsilon^*_{t+1}}.    
+\end{align}
+%
+
+\paragraph{Dealing with negative counts and small values}
+In an epidemiological setting where we deal with count data, negative values should likely be omitted entirely.
+Values of zero should potentially also be removed from the analsys, as it does not make sense to interpret relative errors on zero counts.
+Alternatively, if one is not willing to omit them, a small quantity (such as 1) could be added to all observations before taking the logarithm.
+Note that predictive distribtions have to be truncated in a similar way.
+WOULD WE ALSO ADD ONE TO THE FORECAST?
+
+\subsection{Interpretation}
+The log transformation changes which types of errors (absolute or relative) are penalised.
 The WIS and CRPS computed on the natural scale increase linearly with increasing absolute errors (see Figure \ref{fig:change-in-scores}A, regardless of whether the error is large in relative terms.
@@
 Forecasting 101,000 rather than 100,000 cases will be treated equally to forecasting 2,020, rather than 2,000 hospitalisations.
+Note that for a single forecast the relative rankings of different forecasts would be preserved.
+The transformation only makes a difference when aggregating predictions avross different dimensions.
 
-\paragraph{What happens to the WIS / CRPS when you log it?}
 \begin{figure}[h!]
@@
 Bottom: difference in relative terms.
-Interesting panels are top left and bottom right.}.
-    \label{fig:log-crps-viz}
+Interesting panels are top left and bottom right.}
+    \label{fig:change-in-scores}
 \end{figure}
 
-\section{Appropriate scales for epidemiological processes}
+Evaluating forecasts based on relative errors may be desirable, if different forecast targets, for example reported case numbers, hospitalisations and deaths, have very different orders of magnitude.
+In these instances scores (even for an ideal forecaster) usually also differ by several orders of magnitude.
+When scores are aggregated on the natural scale without any prior transformation, overall performance would be dominated by the performance on the target which has the highest absolute numbers.
+After the log transformation, equal weight would be put on the relative performance, making performance across different targets more easily comparable (as scores of an ideal forecaster would not differ as much as on the natural scale).
 
-THE NEXT PARAGRAPH COULD BE MOVED TO THE INTRODUCTION?
-
-In this section we explore the rationale for and implications of scoring transformed forecasts in an epidemiological context in greater depth.
-Forecast evaluation serves several purposes.
-It provides feedback for forecasters and researchers necessary to improve their models, and it helps decision makers with distinguishing good from bad predictions and choosing forecasts that inform future decisions.
-
-Generally, the way forecasts are evaluated should correspond to the error structure of the model and the type of errors models will usually make.
-In epidemiology, the spread of infectious diseases is usually modelled as an exponential process.
-COMPARTMENT MODELS.
-RENEWAL EQUATION.
-The number of infections usually grows with some exponential rate etc.
-Error from these models will then in general be dominated by multiplicative process errors.
-Under the assumption that these models are correctly capturing transmission dynamics inherent to the system it therefore makes sense to focus evaluation on these multiplicative errors.
-Often, if the model is performing badly in multiplicative terms, this means that there is a flaw with its assumptions that needs to be resolved, regardless of performance in absolute terms.
-When selecting a model for decision-making, similar considerations apply.
-When a model is structured in a multiplicative way, we should base our trust in it on past relative performance, rather than absolute performance.
-Though ultimately this decision is dependent on the users of the forecast who may prefer models that perform better on an absolute scale for a range of reasons (for example not over-estimating the peak size of an epidemic).
-
-SHOULD INSTEAD TALK ABOUT WHAT THE DIFFERNCE IS BETWEEN LOG AND DIVIDING?
-
-\paragraph{What happens when you average things}
-In order to compare models, we usually need to compare aggregate performance.
-In many epidemiological settings, whether we decide to evaluate relative or absolute performance has implications for how we weigh performance of models across different dimensions when we aggregate and compare scores.
-For example, different forecast targets like reported case numbers, hospitalisations and deaths have very different orders of magnitude and scores (even for an ideal forecaster) usually also differ by several orders of magnitude.
-If we aggregated scores on the natural scale across e.g.
-cases and deaths, in order to rank models by overall score, an equal weight on absolute performance would implicitly result in a much higher weight on performance on cases.
-On the log scale, we would put equal weight on the relative performance on cases and on deaths, making performance usually more easily comparable (as scores of an ideal forecaster would not differ as much as on the natural scale).
-IS THAT TRUE? SHOULD PROBABLY PROVE THIS POINT BY SIMULATION.
-
-While it is intuitive that performance on cases is not orders of magnitude more important than performance on deaths, the question becomes trickier if we compare across locations or across time for the same forecast target.
-Indeed, predictive performance of a model may be more important in locations or at times when incidence values are high, as argued by \citep{Bracher}.
-Whether or not we should think this way depends on certain characteristics of the models.
-
-Let us assume that we care in particular about situations where incidence numbers are high.
-It could be useful to give a higher weight to these situations (by scoring on the natural scale), if we believe that certain models may indeed be inherently better in such situations where incidence numbers are high.
-For example, models may be better around epidemic peaks when numbers are highest, or they may be especially good at forecasting large-scale population dynamics.
-%(whatever that may be...)
-Then, scoring on the natural scale will help you identify identify the models that tend to be most situations in the situations we most care about.
-For example, when looking at one forecast target across time, it may be reasonable to assume that the model which performed best around the peak of one wave of COVID-19 may also be the one to take most seriously in the second wave of COVID-19.
-If, on the other hand, we believe that good models should do consistently well and that situations with high incidences do not provide orders of magnitude more information about which models perform well or badly in the circumstances we most care about, then scoring on the log scale may be more appropriate.
+Using a log transformation, however, may not be appropriate if one believes that differences in absolute values are meaningful.
+This may be the case for example when comparing forecasts across locations or time points, rather than different forecasts targets.
+\citep{Bracher} for example argue that in epidemiological settings it is a desirable property of the CRPS / WIS that it assigns large scores to forecasts where counts are high, as this naturally reflects the situations we should most care about and helps us identify the models that perform best in similar situations.
+If, on the other hand, we believe that good models should do consistently well and that situations with high incidences do not provide significantly more information about which models perform well or badly, then scoring on the log scale may be more appropriate.
 For example, it may be reasonable not to think that predictive performance in a country like Germany would have a hundred times more importance than performance in a country like Luxembourg when determining the best model for future decision making.
@@
 
-When scoring forecasts on a log scale, one may fall in the opposite trap of giving to much weight to situations where incidences are low, as small errors (e.g.
-predicting 9, rather than 3 deaths) may have a much larger effect in relative terms than in absolute terms if the quantity to forecast is very small.
-How much weight is given to smaller or larger forecast quantities depends on the exact relationship between the mean and the variance of the quantity of interest.
+An aggregate of forecasts scored on the natural scale will be dominated by predictions of quantities with high absolute values.
+For scoring log transformed forecasts, the opposite may be true, as small absolut errors (e.g.
+predicting 9, rather than 3 deaths) may be large in relative terms if the quantity to forecast is very small.
+How much weight is given to smaller or larger forecast quantities for raw and transformed forecasts depends on the exact relationship between the mean and the variance of the quantity of interest.
+
 To investigate this we sampled forecasts from different negative binomial distributions.
@@
 
-%MAYBE A PARAGRAPH ABOUT OUTLIERS?
+In order to deal with small counts, we can extend the argument made above to small values in general (rather than just zeros) and conclude that it is hard to reliably evaluate relative errors when dealing with small discrete counts.
+One could therefore decide to remove forecasts for small quantities below a certain threshold from the analysis.
+Alternatively, one could add a number $n$ (e.g.
+$n = 100$) to all observations and forecasts before taking the logarithm.
+This constitutes a strictly monotonic transformation which preserves propriety and would reduce all scores.
+Scores for forecasts on small quantities would be particularly affected, resulting in a more equal relative importance of scores for different quantities.
+Researchers can choose to add eitehr a small or large number, depending on how much they want to equalise relative errors when aggregating scores.
+
 %COULD ALSO MAKE A NICE PLOT ABOUT OUTLIERS, where we look at the effect of outliers. but maybe that is already included in the plot in Figure 1. 
@@
 
-%Other transformations: 
-%Box-Cox, square root
+\subsection{Approximately scoring the growth rate}
+NOT SO CONVINCED ANYMORE THAT THIS WORKS. MY INTUITION IS IF YOU WANT THE GROWTH RATE YOU SHOULD ACTUALLY SCORE THE GROWTH RATE AND DIVIDE BY THE LAST OBSERVED VALUE.
+
+Computing a score based on the logarithm of the forecast and the observation is also approximately equivalent to scoring a forecast of a multiplicative growth rate of the forecast target.
+The growth rate is defined as
+%
+\begin{equation}
+    g_{t, t+1} = \frac{y_{t+1} - y_t}{y_t},
+\end{equation}
+%
+where $y_{t+1}$ is the forecast target (e.g. reported cases of COVID-19) in the future, $y_t$ is the last known observation and $g_{t, t+1}$ is the growth rate between now ($t$) and $t+1$. 
+Using the growth rate, we can express the value of the forecast target in the future as such: 
+%
+\begin{equation}
+y_{t+1} = (1 + g_{t, t+1}) \cdot y_t.
+\end{equation}
+%
+
+Now consider a forecast on $\log y_{t+1}$. Note that we can decompose the error into a known and an unknown part: 
+%
+\begin{align}
+\log y_{t+1}  &= \log \hat{y}_{t+1} + \varepsilon^*_{t+1} \\
+              &= \log ((1 + \hat{g}_{t+1}) \cdot y_{t}) + \varepsilon^*_{t+1} \\
+              &= \log (1 + \hat{g}_{t, t+1}) + \log y_t + \varepsilon_{t+1} \Leftrightarrow \\
+              &= \log (1 + \hat{g}_{t, t+1}) + \log y_t - \log y_t + \varepsilon'_{t+1} \Leftrightarrow \\    
+              &= \log (1 + \hat{g}_{t, t+1}) + \varepsilon'_{t+1}.
+\end{align}
+
+NOT SO SURE ANYMORE THAT THIS WORKS.
+CAN WE JUST SUBTRACT Y\_t?
+BECAUSE Y\_t STILL HAS INFORMATION ON THE SIZE OF THE QUANTITY YOU'RE FORECASTING AND ALSO WE'RE REALLY JUST SUBTRACTING IT FROM THE RIGHT HAND SIDE.
+
+What affects the score is only the unknown part, as shifting both the forecast and the observation by a known amount does not make a difference for evaluating the distance between forecast and observation.
+The WIS and CRPS therefore provide a measurement of an additive error on the forecast of$ \log (1 + \hat{g}_{t, t+1})$.
+For small values of $g$, we can make the following approximation
+\begin{equation}
+    \log (1+ g_{t, t+1}) \approx g_{t, t+1}.
+\end{equation}
+%
+
+Conveniently, the decomposition of the WIS into dispersion, overprediction and underprediction is preserved for both transformations.
+The interpretation of the components has now changed in that they now (approximately) represent uncertainty, over-prediction, and under-prediction with the growth rate and so are independent of current incidence.
 
@@
 
-\section{Other ideas}
-\begin{itemize}
-    \item Compare different predictive distributions and the score as a function of an outcome.
-    \item ...
-\end{itemize}
+% \section{Other ideas}
+% \begin{itemize}
+%     \item Compare different predictive distributions and the score as a function of an outcome.
+%     \item ...
+% \end{itemize}
 
@@
 
+\subsection{Additional information on the WIS} \label{wis}
+\paragraph{WIS}
+
+WIS values are always larger or equal than zero and lower values imply better performance.
+The WIS can be decomposed into a dispersion component and penalties for over- and under-prediction.
+For a single prediction interval, the interval score is computed as
+\begin{align}
+ IS_\alpha(F,y) &= (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u) \\
+ &= \text{dispersion} + \text{underprediction} + \text{overprediction},    
+\end{align}
+
+where $1()$ is the indicator function, $y$ is the observed value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution $F$, i.e.
+the lower and upper bound of a single central prediction interval.
+For a single forecast interval, the interval score is the width of the prediction interval, if it covers the observed value, and the absolute error between the forecast and the lower (or upper, respectively) interval boundary.
+For a set of $K$ prediction intervals and the median $m$, the WIS is computed as a weighted sum,
+\begin{equation}
+\text{WIS} = \frac{1}{K + 0.5} \cdot \left(w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y)\right),    
+\end{equation} 
+where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$.
+
+The WIS is closely related to the continuous ranked probability score (CRPS) [Cite Gneiting 2007], which itself can be understood as a generalisation of the absolute error to probabilistic forecasts.
+For an increasing set of equally-spaced prediction intervals the WIS converges to the CRPS and shares many of its properties.
+More information on the CRPS is provided in section \ref{crps} in the SI.
+
 \subsection{Additional information on the CRPS} \label{crps}
@@
 In this representation we can simply replace $X$ and $X'$ by samples sum over all possible combinations to obtain the CRPS.
+
+\subsection{Scoring the growth rate directly}
+By dividing every observed value and every forecast by the last observed value prior to evaluation, we replace forecasting $y_{t+1}$ with forecasting  $y_{t+1} / y_t = 1 + g_{t, t+1}$.
+Just as we were previously penalised an additive error with respect the absolute value, we are now penalising an absolute error with respect to the growth rate.
+
+We can illustrate what's happening by thinking of an analogous point prediction $\hat{g_{t, t+1}}$ for the growth rate. We could then write
+\begin{equation}
+    1 + g_{t, t+1} = 1 + \hat{g}_{t, t+1} + \varepsilon^*_{t+1}. 
+\end{equation}
+%
+The WIS and CRPS would evaluate the forecast based on the error $\varepsilon^*_{t+1}$.
+This translates to an absolute error on the actual forecast target that is multiplicatively connected to the last observed value:
+\begin{align}
+    y_{t+1} &= (1 + \hat{g}_{t, t+1} + \varepsilon^*_{t+1}) \cdot y_t \\
+            &= (1 + \hat{g}_{t, t+1}) \cdot y_t + \varepsilon^*_{t+1} \cdot y_t \\
+            &= \hat{y}_{t+1} + \varepsilon^*_{t+1} \cdot y_t.
+\end{align}
+
+One disadvantage of this is that we can't really do anything when an observation is zero. 
+%is this helpful?%
+%Maybe add some math on how we actually transform the distribution. E.g. divide every quantile or sample by the last observed value, divide the distribution?%
 

### 2022-06-06 main.tex
update manuscript

@@
 Due to their widespread use we focus on these two in this paper, ignoring other proper scoring rules.
-Both the WIS and CRPS represent a measure of the absolute distance between forecast and observation and can be thought of as a generalisation of the absolute error to full predictive distributions (see details in the SI).
+Both the WIS and CRPS represent a measure of the absolute distance between the predictive distribution and observation and can be thought of as a generalisation of the absolute error to full predictive distributions (see details in the SI).
 
@@
 In section [REF] we present an applied example using forecasts submitted to the European COVID-19 Forecast Hub.
-The European COVID-19 Forecast Hub is one of several Forecast hubs that systematically collected, aggregated and evaluated forecasts of several COVID-19 targets created by different teams.
-
-REWRITE HERE. 
-Beginning in 2020, forecasts from different research institutions have been systematically collected, aggregated and evaluated by COVID-19 Forecast hubs in the US [Citation], Germany and Poland, and Europe [Citation].
+The European COVID-19 Forecast Hub is one of several Forecast hubs (together with a Hub in the US [CITATION] and one in Germany and Poland [CITATION]) that systematically collected, aggregated and evaluated forecasts of several COVID-19 targets created by different teams.
 Forecasts submitted to the COVID-19 Forecast Hubs are probabilistic in nature [Citation e.g.
@@
 These predictive distributions are reported in the form of a set of 23 quantiles (11 symmetric prediction intervals as well as a median prediction).
-
-% The WIS can be thought of as an approximation of the CRPS for forecasts where the predictive distribution is represented through a set of quantiles.
-The CRPS, and by extension the WIS, represents a generalisation of the absolute error to continuous distributions, meaning that forecasters are evaluated on the absolute distance of their predictive distribution to the observed value (i.e.
-some form of absolute error).
-For simplicity, we will use absolute error and absolute distance between forecast and observation interchangeably, even though the former applies to point predictions and the latter is used for probabilistic forecasts.
-
-% 
-% 
-% 
-% We argue that when dealing with epidemiological processes it makes sense to shift the evaluation towards multiplicative errors by either taking the logarithm of both the forecast and the observations or by dividing them by the last known observation before scoring.
-
-% What kinds of transformations can be considered meaningful depends on the context.
-We will illustrate the idea of transforming forecasts by focussing on the WIS as it is currently used in epidemiological settings such as the COVID-19 Forecast hubs.
-% In this paper we focus on the WIS due to its simplicity and widespread use by the COVID-19 Forecast Hubs, but arguments apply analogously to the CRPS.
-Another proper scoring rule commonly used is the log score [CITATION], which we will not discuss further in this paper.
+%Currently, the US Flu Forecasting Hub and COVID-19 Scenario Hubs in the US and in Europe use the same format.
 
@@
 
-\paragraph{Example transformations to score the growth rate}
+\paragraph{Example transformations}
 
@@
 
-Divide by the latest known value
+Divide by the latest known value = score forecast on the growth rate
+
+Log that = score log growth rate
 
@@
 
-square root
+square root WAS THAT USED SOMEWHERE?
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Log transformation}
+\section{Log transforming forecasts and observations}
 
-As explained in the introduction, it may be useful to assess epidemiological forecasts based on multiplicative rather than absolute errors, i.e.
-based on how good they predict relative changes in the quantity of interest as apposed to how good they forecast absolute values.
-For count data, as are common in epidemiology, we can achieve this by taking the logarithm of both the forecasts and the observed values.
-
-To illustrate this intuitively, let us consider a point forecast $\hat{y}_{t+1}$ for a quantity of interest $y_{t+1}$, such that 
+In this section (and those that follow) we focus on one particular transformation, taking the logarithm of forecasts and observations, which makes it possible to change the focus of the evaluation from absolute to relative errors.
+To illustrate the effects of applying the logarithm intuitively, let us consider a point forecast $\hat{y}_{t+1}$ for a quantity of interest $y_{t+1}$, such that
 \begin{equation}
@@
 
-The score will be determined by the size of the forecast error $\varepsilon_{t+1}$.
-When taking the logarithm of the forecast and the observation first, than the score is determined based on the absolute error $\varepsilon^*_{t+1}$:
+The analogon of the WIS / CRPS for point forecasts is the absolute error, and so the score will be determined by the size of the forecast error $\varepsilon_{t+1}$.
+When taking the logarithm of the forecast and the observation first, than the score is determined based on the size of $\varepsilon^*_{t+1}$:
 \begin{equation}
@@
 \end{equation}
-
+%
 It is easy to see that this corresponds to evaluating a multiplicative error on the original $y_{t+1}$:
@@
 
-\paragraph{Dealing with negative counts and small values}
-In an epidemiological setting where we deal with count data, negative values should likely be omitted entirely.
-Values of zero should potentially also be removed from the analsys, as it does not make sense to interpret relative errors on zero counts.
-Alternatively, if one is not willing to omit them, a small quantity (such as 1) could be added to all observations before taking the logarithm.
-Note that predictive distribtions have to be truncated in a similar way.
-WOULD WE ALSO ADD ONE TO THE FORECAST?
-
-\subsection{Interpretation}
-The log transformation changes which types of errors (absolute or relative) are penalised.
-The WIS and CRPS computed on the natural scale increase linearly with increasing absolute errors (see Figure \ref{fig:change-in-scores}A, regardless of whether the error is large in relative terms.
+On the natural scale, WIS and CRPS increase linearly with increasing absolute errors (see Figure \ref{fig:change-in-scores}A, regardless of whether the error is large in relative terms.
 Predicting 101,000 cases instead of the true 100,000 will be treated equally to predicting 2,000 hospitalisations, rather than the 1,000 observed.
-The WIS and CRPS based on log-transformed values however, penalise relative errors and increase linearly with an increase in relative errors (see Figure \ref{fig:change-in-scores}D).
+On the log scale, WIS and CRPS penalise relative errors and increase linearly with an increase in relative errors (see Figure \ref{fig:change-in-scores}D).
 Forecasting 101,000 rather than 100,000 cases will be treated equally to forecasting 2,020, rather than 2,000 hospitalisations.
-Note that for a single forecast the relative rankings of different forecasts would be preserved.
-The transformation only makes a difference when aggregating predictions avross different dimensions.
+% Note that for a single forecast the relative rankings of different forecasts would be preserved. The transformation only makes a difference when aggregating predictions across different dimensions.
 
@@
 
-Evaluating forecasts based on relative errors may be desirable, if different forecast targets, for example reported case numbers, hospitalisations and deaths, have very different orders of magnitude.
-In these instances scores (even for an ideal forecaster) usually also differ by several orders of magnitude.
+\subsection{Approximately scoring the growth rate}
+Computing a score based on the logarithm of the forecast and the observation is also approximately equivalent to scoring a forecast of a multiplicative growth rate of the forecast target.
+The growth rate is defined as
+%
+\begin{equation}
+    g_{t, t+1} = \frac{y_{t+1} - y_t}{y_t},
+\end{equation}
+%
+where $y_{t+1}$ is the forecast target (e.g. reported cases of COVID-19) in the future, $y_t$ is the last known observation and $g_{t, t+1}$ is the growth rate between now ($t$) and $t+1$. 
+Using the growth rate, we can express the value of the forecast target in the future as such: 
+%
+\begin{equation}
+y_{t+1} = (1 + g_{t, t+1}) \cdot y_t.
+\end{equation}
+%
+
+Now consider a forecast on $\log y_{t+1}$.
+Using the fact that for small values of $x$, $\log (1+ x) \approx x$, we can see that scoring forecasts on the log scale is approximately equal to scoring an additive error on the growth rate (on the natural scale), as long as the growth rate is small.
+%
+\begin{align}
+\log y_{t+1}        &= \log \hat{y}_{t+1} + \varepsilon^*_{t+1} \\
+                    &= \log ((1 + \hat{g}_{t+1}) \cdot y_{t}) + \varepsilon^*_{t+1} \\
+                    &= \log (1 + \hat{g}_{t, t+1}) + \log y_t + \varepsilon_{t+1} \Leftrightarrow \\
+\varepsilon'_{t+1}  &= \log y_{t+1} -  \log y_t - \log (1 + \hat{g}_{t, t+1}) \Leftrightarrow \\    
+                    &= \log (\frac{y_{t+1}}{y_t}) - \log (1 + \hat{g}_{t, t+1}) \Leftrightarrow \\    
+                    &= \log (1 + g_{t, t+1}) - \log (1 + \hat{g}_{t, t+1}) \Leftrightarrow \\   
+                    &\approx g_{t, t+1} - \hat{g}_{t, t+1} \quad \text{for small } \hat{g}_{t, t+1}, g_{t, t+1}. 
+\end{align}
+
+This approximation breaks down when the growth rate becomes too large.
+If scoring the growth rate is the focus of the evaluation, then it may be better to transform forecasts and observations by the last known values instead in order to score the growth rate explicitly.
+
+\subsection{Dealing with negative counts and small values}
+Unfortunately, the log transformation cannot readily be applied when dealing with negative or zero values.
+In an epidemiological setting that involves count data, negative values should likely be omitted entirely.
+Values of zero should potentially also be removed from the analsys, as it does not make sense to interpret relative errors on zero counts.
+Alternatively, if one is not willing to omit them, a small quantity (such as 1) could be added to all observations before taking the logarithm.
+Note that predictive distribtions have to be truncated in a similar way.
+When dealing with count data we could extend this argument to all small values (rather than just zeros), as it is hard to reliably evaluate relative errors when dealing with small discrete values.
+One could therefore decide to remove forecasts for small quantities below a certain threshold from the analysis.
+Alternatively, one could add a number $n$ (e.g.
+$n = 100$) to all observations and forecasts before taking the logarithm.
+This constitutes a strictly monotonic transformation which preserves propriety and would reduce all scores (through reducing relative errors).
+Scores for forecasts on small quantities would be particularly affected, resulting in a more equal distribution of scores for forecasts of differing orders of magntitudes.
+Researchers can choose to add either a small or large number, depending on how much they want to equalise relative errors when aggregating scores.
+
+\subsection{Relative importance when summarising scores}
+For a single transformation, the ranking between models is preserved, even though the numeric value of the score changes.
+However, when averaging across forecasts, the relative influence of individual forecasts on the aggregate score may differ dramatically, leading to changes in the overall model rankings.
+While the main rationale for applying the log transformation is that this better reflects the types of errors common in epidemiological modelling, it may also be helpful when scores based on the natural scale are very unequal.
+This usually happens, for example, when different forecast targets, like reported case numbers, hospitalisations and deaths, or the same targets across time and space, have very different orders of magnitude.
+In these instances, scores (even for an ideal forecaster) usually also differ by several orders of magnitude.
 When scores are aggregated on the natural scale without any prior transformation, overall performance would be dominated by the performance on the target which has the highest absolute numbers.
@@
 
-Using a log transformation, however, may not be appropriate if one believes that differences in absolute values are meaningful.
+While on the natural scale, aggregate forecasts will usually be dominated by predictions of quantities with high absolute values, the opposite may be true for scores based on log transformed predictions.
+Small absolute errors (e.g.
+predicting 9, rather than 3 deaths) may be large in relative terms if the quantity to forecast is very small.
+The strength of this effect depends on the exact relationship between the mean and the variance of the quantity of interest.
+
+ADD TO FIGURE CAPTION? 
+To illustrate this with count data we sampled forecasts from different negative binomial distributions.
+The negative binomial distribution has mean $\mu$ and variance $\sigma^2 = \mu + \mu ^2 / \theta$ and for $\lim_{\theta \to \infty}$ converges to the Poisson distribution.
+For large values of $\theta$, resulting in a variance approximately equal to the mean, forecasts for lower quantities on average received higher scores on the log scale (see Figure \ref{fig:SIM-wis-state-size-mean}.
+For $\theta = 1$ (and correspondingly, $\sigma^2 = \mu + \mu^2$, we found that scores on the log scale remained approximately constant regardless of the size of the quantity to forecast.
+For $\theta = 0.1$ (and $\sigma^2 = \mu + 10 \cdot \mu^2$), scores on the log scale increased with the quantity to forecast.
+When scored on the natural scale, a higher quantity to forecast always lead to higher scores regardless of the chosen distribution (as long as the variance grows together with the mean).
+
+One way to handle situations where scores are very unequal, as mentioned above, is to add a number $n$ to all forecasts and observations in order to reduce the relative weight of forecast targets with small quantities.
+
+Whether or not a log transformation is apprpriate depends on whether one believes that differences in absolute values are meaningful.
 This may be the case for example when comparing forecasts across locations or time points, rather than different forecasts targets.
@@
 We could then argue that within one wave, performance of the peak is meaningful, but across different waves (with different numbers) it is unclear, whether that difference conveys meaningful information.
-
-An aggregate of forecasts scored on the natural scale will be dominated by predictions of quantities with high absolute values.
-For scoring log transformed forecasts, the opposite may be true, as small absolut errors (e.g.
-predicting 9, rather than 3 deaths) may be large in relative terms if the quantity to forecast is very small.
-How much weight is given to smaller or larger forecast quantities for raw and transformed forecasts depends on the exact relationship between the mean and the variance of the quantity of interest.
-
-To investigate this we sampled forecasts from different negative binomial distributions.
-The negative binomial distribution has mean $\mu$ and variance $\sigma^2 = \mu + \mu ^2 / \theta$ and for $\lim_{\theta \to \infty}$ converges to the Poisson distribution.
-For large values of $\theta$, resulting in a variance approximately equal to the mean, forecasts for lower quantities on average received higher scores on the log scale (see Figure \ref{fig:SIM-wis-state-size-mean}.
-For $\theta = 1$ (and correspondingly, $\sigma^2 = \mu + \mu^2$, we found that scores on the log scale remained approximately constant regardless of the size of the quantity to forecast.
-For $\theta = 0.1$ (and $\sigma^2 = \mu + 10 \cdot \mu^2$), scores on the log scale increased with the quantity to forecast.
-When scored on the natural scale, a higher quantity to forecast always lead to higher scores regardless of the chosen distribution (as long as the variance grows together with the mean).
-
-In order to deal with small counts, we can extend the argument made above to small values in general (rather than just zeros) and conclude that it is hard to reliably evaluate relative errors when dealing with small discrete counts.
-One could therefore decide to remove forecasts for small quantities below a certain threshold from the analysis.
-Alternatively, one could add a number $n$ (e.g.
-$n = 100$) to all observations and forecasts before taking the logarithm.
-This constitutes a strictly monotonic transformation which preserves propriety and would reduce all scores.
-Scores for forecasts on small quantities would be particularly affected, resulting in a more equal relative importance of scores for different quantities.
-Researchers can choose to add eitehr a small or large number, depending on how much they want to equalise relative errors when aggregating scores.
 
@@
 
-\subsection{Approximately scoring the growth rate}
-NOT SO CONVINCED ANYMORE THAT THIS WORKS. MY INTUITION IS IF YOU WANT THE GROWTH RATE YOU SHOULD ACTUALLY SCORE THE GROWTH RATE AND DIVIDE BY THE LAST OBSERVED VALUE.
-
-Computing a score based on the logarithm of the forecast and the observation is also approximately equivalent to scoring a forecast of a multiplicative growth rate of the forecast target.
-The growth rate is defined as
-%
-\begin{equation}
-    g_{t, t+1} = \frac{y_{t+1} - y_t}{y_t},
-\end{equation}
-%
-where $y_{t+1}$ is the forecast target (e.g. reported cases of COVID-19) in the future, $y_t$ is the last known observation and $g_{t, t+1}$ is the growth rate between now ($t$) and $t+1$. 
-Using the growth rate, we can express the value of the forecast target in the future as such: 
-%
-\begin{equation}
-y_{t+1} = (1 + g_{t, t+1}) \cdot y_t.
-\end{equation}
-%
-
-Now consider a forecast on $\log y_{t+1}$. Note that we can decompose the error into a known and an unknown part: 
-%
-\begin{align}
-\log y_{t+1}  &= \log \hat{y}_{t+1} + \varepsilon^*_{t+1} \\
-              &= \log ((1 + \hat{g}_{t+1}) \cdot y_{t}) + \varepsilon^*_{t+1} \\
-              &= \log (1 + \hat{g}_{t, t+1}) + \log y_t + \varepsilon_{t+1} \Leftrightarrow \\
-              &= \log (1 + \hat{g}_{t, t+1}) + \log y_t - \log y_t + \varepsilon'_{t+1} \Leftrightarrow \\    
-              &= \log (1 + \hat{g}_{t, t+1}) + \varepsilon'_{t+1}.
-\end{align}
-
-NOT SO SURE ANYMORE THAT THIS WORKS.
-CAN WE JUST SUBTRACT Y\_t?
-BECAUSE Y\_t STILL HAS INFORMATION ON THE SIZE OF THE QUANTITY YOU'RE FORECASTING AND ALSO WE'RE REALLY JUST SUBTRACTING IT FROM THE RIGHT HAND SIDE.
-
-What affects the score is only the unknown part, as shifting both the forecast and the observation by a known amount does not make a difference for evaluating the distance between forecast and observation.
-The WIS and CRPS therefore provide a measurement of an additive error on the forecast of$ \log (1 + \hat{g}_{t, t+1})$.
-For small values of $g$, we can make the following approximation
-\begin{equation}
-    \log (1+ g_{t, t+1}) \approx g_{t, t+1}.
-\end{equation}
-%
-
-Conveniently, the decomposition of the WIS into dispersion, overprediction and underprediction is preserved for both transformations.
+Conveniently, the decomposition of the WIS into dispersion, overprediction and underprediction is preserved.
 The interpretation of the components has now changed in that they now (approximately) represent uncertainty, over-prediction, and under-prediction with the growth rate and so are independent of current incidence.
@@
 Across all time points and locations, the mean number weighted interval score on the natural scale for cases (deaths) was 8855 (59.3) with an overall standard deviation of 51274 (156).
-On the log scale, scores were much closer to each other (see Figure \ref{fig:HUB-average-scores}), with a mean wis for cases (deaths) of 0.507 (0.362) with an overall standard deviation of 0.745 (0.407).
+On the log scale, scores were much closer to each other (see Figure \ref{fig:HUB-average-scores}), with a mean WIS for cases (deaths) of 0.507 (0.362) with an overall standard deviation of 0.745 (0.407).
 % NEED A TABLE WITH SCORES for the Appendix
@@
 
+% NOT SO CONVINCED ANYMORE THAT THIS WORKS. MY INTUITION IS IF YOU WANT THE GROWTH RATE YOU SHOULD ACTUALLY SCORE THE GROWTH RATE AND DIVIDE BY THE LAST OBSERVED VALUE. 
+% 
+% WELL MAYBE IT WORKS AFTER ALL AND IT WOULD BE THE EQUIVALENT OF LOGGING FIRST AND THEN SUBTRACTING THE LAST OBSERVED VALUE WAIT A MINUTE THAT'S A DIFFERENT TRANSFORMATION AND ALSO THAT's SCORING THE LOG GROWTH RATE.
+% 
+% BUT THEN AGAIN I'M STILL CONFUSED, BECAUSE ANY RELATIVE ERROR ON A QUANTITY Y\_t+1 IS AUTOMATICALLY AN ABSOLUTE ERROR ON THE GROWTH RATE IF WE KNOW Y\_t.
+AH BUT YOU DON'T KNOW WHAT THE ABSOLUTE ERROR ON THE GROWTH RATE IS (E.G.
+FOR Y\_t+1 of 10 IF YOU DON'T SPECIFICALLY LOOK AT THE LAST VALUE.
+BUT WE ARE ASSUMING THAT THIS IS WHAT THE MODEL DOES WHEN IT MAKES A FORECAST INTERNALLY
+
+% The WIS can be thought of as an approximation of the CRPS for forecasts where the predictive distribution is represented through a set of quantiles.
+The CRPS, and by extension the WIS, represents a generalisation of the absolute error to continuous distributions, meaning that forecasters are evaluated on the absolute distance of their predictive distribution to the observed value (i.e.
+some form of absolute error).
+For simplicity, we will use absolute error and absolute distance between forecast and observation interchangeably, even though the former applies to point predictions and the latter is used for probabilistic forecasts.
+
+% 
+% 
+% 
+% We argue that when dealing with epidemiological processes it makes sense to shift the evaluation towards multiplicative errors by either taking the logarithm of both the forecast and the observations or by dividing them by the last known observation before scoring.
+
+% What kinds of transformations can be considered meaningful depends on the context.
+We will illustrate the idea of transforming forecasts by focussing on the WIS as it is currently used in epidemiological settings such as the COVID-19 Forecast hubs.
+% In this paper we focus on the WIS due to its simplicity and widespread use by the COVID-19 Forecast Hubs, but arguments apply analogously to the CRPS.
+Another proper scoring rule commonly used is the log score [CITATION], which we will not discuss further in this paper.
+
 % \paragraph{Problems with the WIS and CRPS in an epidemiological setting}

### 2022-06-08 main.tex
update manuscript

@@
 In these instances, scores (even for an ideal forecaster) usually also differ by several orders of magnitude.
-When scores are aggregated on the natural scale without any prior transformation, overall performance would be dominated by the performance on the target which has the highest absolute numbers.
-After the log transformation, equal weight would be put on the relative performance, making performance across different targets more easily comparable (as scores of an ideal forecaster would not differ as much as on the natural scale).
+Evaluating forecasts based on relative errors often makes it easier to compare performance across different targets more easily (as scores of an ideal forecaster would not differ as much as on the natural scale).
 
-While on the natural scale, aggregate forecasts will usually be dominated by predictions of quantities with high absolute values, the opposite may be true for scores based on log transformed predictions.
+While on the natural scale, aggregate scores will usually be dominated by predictions of quantities with high absolute values, the opposite may be true for scores based on log transformed predictions.
 Small absolute errors (e.g.
@@
 The strength of this effect depends on the exact relationship between the mean and the variance of the quantity of interest.
+For simulated negative binomial count data, log predictions for small quantities receive on average higher scores than forecasts for large quantities if the mean and the variance grow at the same rate $\sigma^2 = \mu$, about equal scores when the variance grows at a rate of $\sigma^2 = \mu + \mu^2$ (Figure \ref{fig:SIM-wis-state-size-mean}), and smaller scores when the variance grows faster than that.
+As mentioned above, one way to handle situations where scores are very unequal, is to add a number $n$ to all forecasts and observations in order to reduce the relative weight of forecast targets with small quantities.
 
-ADD TO FIGURE CAPTION? 
-To illustrate this with count data we sampled forecasts from different negative binomial distributions.
+% To illustrate this with count data we sampled forecasts from different negative binomial distributions.
 The negative binomial distribution has mean $\mu$ and variance $\sigma^2 = \mu + \mu ^2 / \theta$ and for $\lim_{\theta \to \infty}$ converges to the Poisson distribution.
@@
 
-One way to handle situations where scores are very unequal, as mentioned above, is to add a number $n$ to all forecasts and observations in order to reduce the relative weight of forecast targets with small quantities.
-
+COULD ALSO BE DISCUSSION
 Whether or not a log transformation is apprpriate depends on whether one believes that differences in absolute values are meaningful.

### 2022-07-08 analysis/Analysis ECDC forecasts.Rmd
Update ECDC analasys and manuscript text

--- 2022-01-31 analysis/Analysis ECDC forecasts.Rmd
@@
 
+label_fn <- function(x) {
+  ifelse(x < 1000, 
+         paste(x), 
+         ifelse(x < 1e6, 
+                paste0(x / 1000, "k"),
+                ifelse(x < 1e9, 
+                       paste0(x / 1e6, "m"), 
+                       paste0(x / 1e9, "b"))
+                )
+         )
+}
+
 ```
@@
   unique()
-```
-
-```{r}
-# create version of data where the true value is the true growth rate
-# interesting question: does it make a difference, whether we transform it 
-# to be the weekly growth rate, i.e. take the n-th root, where n is the horizon?
-
-latest_truth <- fread(here("data", "weekly-truth-Europe.csv")) |>
-  rename(last_known_true_value = true_value) |>
-  mutate(forecast_date = target_end_date + 2) |>
-  select(location, target_type, forecast_date, last_known_true_value) |>
-  unique()
-
-# unsure whether the pmax hack is the right way to do it
-growth_data <- hub_data |>
-  inner_join(latest_truth) |>
-  mutate(true_value = true_value / pmax(last_known_true_value + 1),
-         prediction = prediction / pmax(last_known_true_value + 1)) |>
-  select(-last_known_true_value)
 ```
@@
 scores[, scale := factor(scale, levels = c("natural", "log"))]
+data.table::setnames(scores, old = c("aem", "sharpness"), new = c("ae_median", "dispersion"))
 
@@
 
+## =============================================================================
+## FIGURE 4
 ## Mean vs. sd of observed values
@@
 ```{r}
-
-label_fn <- function(x) {
-  ifelse(x < 1000, 
-         paste(x), 
-         ifelse(x < 1e6, 
-                paste0(x / 1000, "k"),
-                ifelse(x < 1e9, 
-                       paste0(x / 1e6, "m"), 
-                       paste0(x / 1e9, "b"))
-                )
-         )
-}
-
 # average number of cases and deaths
@@
 
-# negative binomial
-# MASS::glm.nb(true_value ~ target_type, 
-#        data = hub_data %>%
-#          mutate(true_value = pmax(true_value, 0)), 
-#        link = "identity")
-
 # plot with means across locations ---------------------------------------------
@@
 
-hub_data |>
+mean_observations <- hub_data |>
   dplyr::select(location, target_type, target_end_date, true_value) |>
@@
             sd_value = sd(true_value), 
-            var_value = var(true_value)) |>
+            var_value = var(true_value))
+
+plot_means_obs <- mean_observations |>
   ggplot(aes(y = mean_value, x = reorder_within(location, -mean_value, target_type))) + 
-  geom_bar(stat = "identity") + 
+  geom_bar(stat = "identity", fill = "grey50") + 
   facet_wrap(~ target_type, scales = "free") + 
-  theme_minimal() + 
+  theme_scoringutils() + 
   scale_y_continuous(labels = label_fn) + 
@@
 
-ggsave("output/figures/HUB-mean-obs-location.png", width = 9, height = 4)
+plot_df <- scores |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  group_by(target_type, location, scale) |>
+  summarise(interval_score = mean(interval_score), 
+            .groups = "drop_last") |> 
+  full_join(mean_observations)
 
+plot_mean_scores <- plot_df |>
+  ggplot(aes(y = interval_score,
+             x = reorder_within(location, -mean_value, target_type))) + 
+  geom_bar(stat = "identity") + 
+  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
+  facet_wrap(scale ~ target_type, scale = "free") + 
+  theme_scoringutils() + 
+  theme(legend.position = "none") + 
+  scale_y_continuous(labels = scale_fn) + 
+  scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) +
+  labs(y = "Interval score", x = "Location")
+
+plot_means_obs / 
+  plot_mean_scores + 
+  plot_layout(heights = c(1, 2))
+
+ggsave("output/figures/HUB-mean-obs-location.png", width = 9, height = 8)
+
+# This uses the ensemble, but alternatively, we could think about just summarising across all models for these plots? 
+```
+
+```{r}
 # plot with mean vs. variance --------------------------------------------------
@@
   geom_function(fun = function(x) x^2, linetype = "dashed", size = 0.5) + 
-  theme_minimal()
+  theme_scoringutils() + 
+  geom_smooth(alpha = 0.2, size = 0.2)
 
@@
   facet_wrap(~ scale, scale = "free") + 
-  theme_minimal() + 
+  theme_scoringutils() + 
   theme(legend.position = "none") + 
@@
 
-ggsave("output/figures/HUB-average-scores.png", width = 7, height = 3.5)
+ggsave("output/figures/HUB-average-scores.png", width = 7, height = 2.5)
 ```
@@
 
-## Scores across locations
-
---> could also do this plot with multiplicative scores, i.e. how much higher is a score than the lowest score observed
-
-```{r}
-# alternatively, we could think about just summarising across all models for these plots?
-
-# make two separate plots to keep ordering. I.e. order locations according 
-# to log scale to illustrate that the ordering changes for cases
-
-get_order <- function(df, type = "Cases") {
-  df |> 
-    filter(scale == "natural", target_type == type) |>
-    arrange(-interval_score) |>
-    pull(location)
-}
-
-plot_df <- scores |> 
-  filter(model == "EuroCOVIDhub-ensemble", 
-         horizon == 2) |>
-  group_by(target_type, location, scale) |>
-  summarise(interval_score = mean(interval_score), 
-            .groups = "drop_last")
-
-order_cases <- get_order(plot_df)
-order_deaths <- get_order(plot_df, type = "Deaths")
-
-plot <- function(plot_df, order) {
-  plot_df |>
-    mutate(location = factor(location, levels = order)) |>
-    ggplot(aes(y = interval_score, x = location)) + 
-    geom_bar(stat = "identity") + 
-    scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
-    facet_wrap(~ target_type + scale, scale = "free") + 
-    theme_minimal() + 
-    theme(legend.position = "none") + 
-    scale_y_continuous(labels = scale_fn) + 
-    scale_x_discrete(guide = guide_axis(n.dodge=2)) +
-    labs(y = "Interval score", x = "Location")
-}
-
-cases <- plot_df |>
-  filter(target_type == "Cases") |>
-  plot(order = order_cases)
-
-deaths <- plot_df |>
-  filter(target_type == "Deaths") |>
-  plot(order = order_deaths)
-
-cases / 
-  deaths
-
-ggsave("output/figures/HUB-scores-locations.png", width = 7, height = 5)
-```
-
-*We see that the scores are much more evenly distributed across locations if we log the data*
-
 ## Correlation between scores
@@
 
-## Variance scores across locations
-
-```{r}
-scores |> 
-  filter(model == "EuroCOVIDhub-ensemble", 
-         horizon == 2) |>
-  group_by(target_type, location, scale) |>
-  summarise(mean_wis = mean(interval_score), 
-            sd_wis = sd(interval_score), 
-            .groups = "drop_last") |>
-  ggplot(aes(y = sd_wis, x = mean_wis)) + 
-  geom_abline(slope = 1, intercept = 0, linetype = "dashed", size = 0.5) + 
-  geom_point() + 
-  facet_wrap(~ target_type + scale, scale = "free") + 
-  theme_minimal() + 
-  theme(legend.position = "none") + 
-  # scale_x_continuous(labels = scale_fn) + 
-  # scale_y_continuous(labels = scale_fn) + 
-  scale_x_continuous(labels = scale_fn, trans = "log10") +
-  scale_y_continuous(labels = scale_fn, trans = "log10") +
-  labs(x = "Mean WIS", y = "Sd of WIS")
-
-ggsave("output/figures/HUB-sd-vs-mean-scores.png", width = 7, height = 5)
-
-wis_vs_total <- function(scores, summary_fct = mean, fulldata = hub_data) {
-
-  total_values <- 
-    fulldata |> 
-    filter(!is.na(prediction)) |>
-    group_by(location, target_type, target_end_date) |>
-    summarise(true_value = unique(true_value)) |>
-    group_by(location, target_type) |>
-    summarise(total = sum(true_value))
-
-  plot_df <- scores |> 
-    filter(model == "EuroCOVIDhub-ensemble", 
-           horizon == 2) |>
-    group_by(target_type, location, scale) |>
-    summarise(wis = summary_fct(interval_score), 
-              .groups = "drop_last") |> 
-    inner_join(total_values)
-
-  plot_df |>
-    group_by(target_type, scale) |>
-    summarise(correlation = cor(wis, total)) |> 
-    print()
-
-  plot_df |>
-    ggplot(aes(y = wis, x = total)) + 
-    geom_point() + 
-    facet_wrap(~ target_type + scale, scale = "free") + 
-    theme_minimal() + 
-    theme(legend.position = "none") + 
-    scale_x_continuous(labels = scale_fn) +
-    scale_y_continuous(labels = scale_fn) +
-    labs(y = "REPLACE", x = "Total number of observed values")
-}
-
-wis_vs_total(scores) + 
-  labs(y = "Mean WIS")
-ggsave("output/figures/HUB-mean-scores-vs-total.png", width = 7, height = 5)
-
-wis_vs_total(scores, summary_fct = stats::sd) + 
-  labs(y = "Sd WIS")
-ggsave("output/figures/HUB-sd-scores-vs-total.png", width = 7, height = 5)
-
-# log-log versions of that plot
-wis_vs_total(scores) + 
-  labs(y = "Mean WIS") + 
-  scale_x_continuous(labels = scale_fn, trans = "log10") +
-  scale_y_continuous(labels = scale_fn, trans = "log10") 
-ggsave("output/figures/HUB-mean-scores-vs-total-log-log.png", width = 7, height = 5)
-
-wis_vs_total(scores, summary_fct = stats::sd) + 
-  labs(y = "Sd WIS") + 
-  scale_x_continuous(labels = scale_fn, trans = "log10") +
-  scale_y_continuous(labels = scale_fn, trans = "log10") 
-ggsave("output/figures/HUB-sd-scores-vs-total-log-log.png", width = 7, height = 5)
-
-# correlation
-
-```
-
 ## Scores over time
-
 restrict to a few locations, e.g. GB, FR, ES (highest average scores there)
@@
   facet_grid(target_type ~ scale, scale = "free_y") + 
-  theme_minimal() + 
+  theme_scoringutils() + 
   labs(x = "Forecast date", y = "WIS relative to smallest observed value")
@@
   facet_grid(target_type ~ scale) + 
-  theme_minimal() + 
+  theme_scoringutils() + 
   labs(x = "Forecast horizon in weeks", y = "Multplicative change in interval score")
@@
 *Slightly less increase on the log scale, but overall not a lot of difference*
+
+## Actual model rankings for single forecasts
+
+```{r}
+# we see that rankings change even for a single forecast. So we need to check everything is still correct. 
+# take one example
+test_data <- hub_data |>
+  filter(model %in% c("MUNI-ARIMA", "epiforecasts-EpiNow2"), 
+         forecast_date == "2021-10-18", 
+         target_type == "Cases", 
+         location == "AT")
+
+test_data |>
+  score(metrics = c("interval_score")) |>
+  summarise_scores()
+
+test_data |>
+  mutate(prediction = log(prediction), 
+         true_value = log(true_value)) |>
+  score(metrics = c("interval_score")) |>
+  summarise_scores()
+## indeed everything seems to be correct, but rankings still change FOR A SINGLE FORECAST. 
+# It seems this is because the trade-off between calibration and sharpness changes? 
+# need to check in what way
+
+rankings <- scores |>
+  select(-c(coverage_deviation, bias, aem, sharpness, underprediction, overprediction)) |>
+  group_by(location, forecast_date, horizon, target_type, scale) |>
+  mutate(ranking = rank(interval_score) / n()) |>
+   filter(forecast_date == "2021-10-18", 
+         target_type == "Cases", 
+         location == "AT")
+
+factor_levels <-  rankings |>
+  filter(horizon == 2, scale == "natural") |>
+  group_by(model) |>
+  summarise(ranking = mean(ranking)) |>
+  arrange(-ranking) |>
+  pull(model)
+
+p_rankings <- rankings |>
+  filter(horizon == 2) |>
+  mutate(model = factor(model, levels = factor_levels)) |>
+  ggplot(aes(y = model, x = ranking)) +
+  ggdist::stat_halfeye() +
+  facet_wrap(~ scale) +
+  theme_scoringutils() +
+  labs(y = "Model", x = "Normalised ranking")
+
+rankings_diff <- rankings |>
+  select(model, location, target_end_date, forecast_date, interval_score, scale, ranking) |>
+  pivot_wider(names_from = scale, values_from = ranking, id_cols = -interval_score)
+
+p_diff_ranking <- rankings_diff |>
+  filter(horizon == 2) |>
+  mutate(diff_ranking = natural - log) |>
+  mutate(model = factor(model, levels = factor_levels)) |>
+  ggplot(aes(y = model, x = diff_ranking)) +
+  ggdist::stat_halfeye() +
+  theme_scoringutils() +
+  labs(y = "", x = "Differences in ranking") +
+  theme(axis.text.y = element_blank(), 
+        axis.ticks.y = element_blank(), 
+        axis.line.y = element_blank())
+
+p_rankings + p_diff_ranking + 
+  plot_layout(widths = c(2, 1))
+
+ggsave("output/figures/HUB-ranking-diffs.png", width = 9, height = 5)
+
+rankings_diff |>
+  group_by(horizon, target_type) |>
+  summarise(cor = cor(natural, log))
+```
+
+## Look at pairwise comparisons
+
+```{r}
+summarised_pairwise <- scores |>
+  filter(horizon == 2, 
+         target_type == "Cases") |>
+  summarise_scores(by = c("model", "scale"), relative_skill = TRUE)
+
+ranking_natural <- 
+  summarised_pairwise |>
+  filter(scale == "natural") |>
+  arrange(-relative_skill) |>
+  pull(model)
+
+plot_rel_skill <- summarised_pairwise |>
+  mutate(model = factor(model, levels = ranking_natural)) |>
+  ggplot(aes(y = model, x = relative_skill)) +
+  geom_bar(stat = "identity") +
+  theme_scoringutils() +
+  facet_wrap(~scale) +
+  labs(y = "Model", x = "Relative skill")
+
+diffs <- summarised_pairwise |>
+  mutate(model = factor(model, levels = ranking_natural)) |>
+  select(model, relative_skill, scale) |>
+  pivot_wider(names_from = scale, values_from = relative_skill) |>
+  mutate(model = factor(model, levels = ranking_natural), 
+         difference = natural - log)
+
+plot_diffs <- diffs |>
+  mutate(title = "difference") |>
+  ggplot(aes(y = model, x = difference)) +
+  geom_bar(stat = "identity") +
+  theme_scoringutils() +
+  labs(y = "Model", x = "Difference in relative skill") +
+  theme(axis.title.y = element_blank(), 
+        axis.text.y = element_blank(), 
+        axis.ticks.y = element_blank(), 
+        axis.line.y = element_blank()) +
+  facet_wrap(~title)
+
+plot_distr_scores <- scores |>
+  filter(horizon == 2) |>
+  mutate(model = factor(model, levels = ranking_natural)) |>
+  ggplot(aes(y = model, x = interval_score)) +
+  ggdist::stat_pointinterval() +
+  facet_wrap(~ scale, scale = "free_x") +
+  theme_scoringutils() +
+  scale_x_log10(labels = label_fn) +
+  theme(axis.title.y = element_blank(), 
+        axis.text.y = element_blank(), 
+        axis.ticks.y = element_blank(), 
+        axis.line.y = element_blank()) +
+  labs(y = "Model", x = "WIS")
+
+plot_rel_wis <- scores |>
+  filter(horizon == 2, 
+         target_type == "Cases") |>
+  mutate(model = factor(model, levels = ranking_natural)) |>
+  summarise_scores(by = c("model", "scale")) |>
+  plot_wis(relative_contributions = FALSE) +
+    theme(axis.title.y = element_blank(), 
+        axis.text.y = element_blank(), 
+        axis.ticks.y = element_blank(), 
+        axis.line.y = element_blank()) +
+  scale_x_continuous(labels = label_fn) +
+  facet_wrap(~scale, scales = "free_x")
+
+plot_rel_skill + plot_diffs + plot_rel_wis + #plot_distr_scores +
+  plot_layout(widths = c(2, 1, 2)) &
+  plot_layout(guides = "collect") &
+  theme(legend.position = "bottom")
+
+ggsave("output/figures/HUB-pairwise-comparisons.png", width = 9, height = 5)
+
+#Issues: this currently ignores that some models have only made forecasts for one location.
+They should usually be excluded, but maybe not necessary here as we're really more comparing the results for natural and log rather than the models against each other.
+
+cor(diffs$log, diffs$natural)
+
+# "CovidMetrics-epiBATS" only has forecasts in DE
+# "ICM-agentModel" only in Poland
+# "UVA is everywhere
+hub_data |>
+  filter(model == "UVA-Ensemble") |> check_forecasts()
+
+plot_pred_score <- function(hub_data, scores, model,
+                            horizon = 2, type = "Cases") {
+  forecast_model <- model
+  h <- horizon
+  filtered_hub <- hub_data |>
+    filter(model == forecast_model, 
+           horizon %in% h, 
+           target_type == type)
+
+  locations <- unique(filtered_hub$location)
+
+  filtered_scores <- scores |>
+    filter(model == forecast_model, 
+           horizon %in% h, 
+           location %in% locations,
+           target_type == type)
+
+  plot_natural  <- filtered_hub |>  
+    plot_predictions(x = "target_end_date") +
+    scale_y_continuous(labels = label_fn)  +
+    theme(legend.position = "bottom")
+
+  plot_log <- filtered_hub |>
+    mutate(prediction = log(prediction + 1), 
+           true_value = log(true_value + 1)) |>
+    plot_predictions(x = "target_end_date") +
+    scale_y_continuous(labels = label_fn)  +
+    theme(legend.position = "bottom")
+
+  scores_natural <- filtered_scores |>
+    filter(scale == "natural") |>
+    summarise_scores("target_end_date") |>
+    mutate(target_end_date = factor(target_end_date)) |>
+    plot_wis(x = "target_end_date", flip = TRUE) +
+    scale_x_continuous(labels = label_fn)
+
+  scores_log <- filtered_scores |>
+    filter(scale == "log") |>
+    summarise_scores("target_end_date") |>
+    mutate(target_end_date = factor(target_end_date)) |>
+    plot_wis(x = "target_end_date", flip = TRUE)
+
+  (plot_natural + plot_log) /
+    (scores_natural + scores_log) +
+    plot_layout(guides = "collect") &
+    theme(legend.position = "bottom")
+}
+
+ICM <- plot_pred_score(hub_data, scores, "ICM-agentModel")
+
+epiBATS <- plot_pred_score(hub_data, scores, "CovidMetrics-epiBATS")
+
+BSLCoV <- plot_pred_score(hub_data, scores, "UB-BSLCoV")
+
+ggsave(filename = "output/figures/HUB-model-ICM.png", width = 7, height = 4.5, plot = ICM)
+ggsave(filename = "output/figures/HUB-model-epiBATS.png", width = 7, height = 4.5, plot = epiBATS)
+ggsave(filename = "output/figures/HUB-model-BSLCoV.png", width = 7, height = 4.5, plot = BSLCoV)
+
+```
 
@@
   facet_wrap(~ target_type + scale, scale = "free") + 
-  theme_minimal() + 
+  theme_scoringutils() + 
   theme(legend.position = "none") + 
@@
 ```
+
+## Code parking lot
+
+```{r}
+# create version of data where the true value is the true growth rate
+# interesting question: does it make a difference, whether we transform it 
+# to be the weekly growth rate, i.e. take the n-th root, where n is the horizon?
+
+latest_truth <- fread(here("data", "weekly-truth-Europe.csv")) |>
+  rename(last_known_true_value = true_value) |>
+  mutate(forecast_date = target_end_date + 2) |>
+  select(location, target_type, forecast_date, last_known_true_value) |>
+  unique()
+
+# unsure whether the pmax hack is the right way to do it
+growth_data <- hub_data |>
+  inner_join(latest_truth) |>
+  mutate(true_value = true_value / pmax(last_known_true_value + 1),
+         prediction = prediction / pmax(last_known_true_value + 1)) |>
+  select(-last_known_true_value)
+```
+
+## Variance of scores across locations
+
+```{r}
+scores |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  group_by(target_type, location, scale) |>
+  summarise(mean_wis = mean(interval_score), 
+            sd_wis = sd(interval_score), 
+            .groups = "drop_last") |>
+  ggplot(aes(y = sd_wis, x = mean_wis)) + 
+  geom_abline(slope = 1, intercept = 0, linetype = "dashed", size = 0.5) + 
+  geom_point() + 
+  facet_wrap(~ target_type + scale, scale = "free") + 
+  theme_scoringutils() + 
+  theme(legend.position = "none") + 
+  # scale_x_continuous(labels = scale_fn) + 
+  # scale_y_continuous(labels = scale_fn) + 
+  scale_x_continuous(labels = scale_fn, trans = "log10") +
+  scale_y_continuous(labels = scale_fn, trans = "log10") +
+  labs(x = "Mean WIS", y = "Sd of WIS")
+
+ggsave("output/figures/HUB-sd-vs-mean-scores.png", width = 7, height = 5)
+```
+
+```{r}
+# negative binomial
+# MASS::glm.nb(true_value ~ target_type, 
+#        data = hub_data %>%
+#          mutate(true_value = pmax(true_value, 0)), 
+#        link = "identity")
+```
+
+## WIS vs. the totsl number of observed values
+probably not needed anymore, as we have similar things
+
+```{r}
+wis_vs_total <- function(scores, summary_fct = mean, fulldata = hub_data) {
+
+  total_values <- 
+    fulldata |> 
+    filter(!is.na(prediction)) |>
+    group_by(location, target_type, target_end_date) |>
+    summarise(true_value = unique(true_value)) |>
+    group_by(location, target_type) |>
+    summarise(total = sum(true_value))
+
+  plot_df <- scores |> 
+    filter(model == "EuroCOVIDhub-ensemble", 
+           horizon == 2) |>
+    group_by(target_type, location, scale) |>
+    summarise(wis = summary_fct(interval_score), 
+              .groups = "drop_last") |> 
+    inner_join(total_values)
+
+  plot_df |>
+    group_by(target_type, scale) |>
+    summarise(correlation = cor(wis, total)) |> 
+    print()
+
+  plot_df |>
+    ggplot(aes(y = wis, x = total)) + 
+    geom_point() + 
+    facet_wrap(~ target_type + scale, scale = "free") + 
+    theme_scoringutils() + 
+    theme(legend.position = "none") + 
+    scale_x_continuous(labels = scale_fn) +
+    scale_y_continuous(labels = scale_fn) +
+    labs(y = "REPLACE", x = "Total number of observed values")
+}
+
+wis_vs_total(scores) + 
+  labs(y = "Mean WIS")
+ggsave("output/figures/HUB-mean-scores-vs-total.png", width = 7, height = 5)
+
+wis_vs_total(scores, summary_fct = stats::sd) + 
+  labs(y = "Sd WIS")
+ggsave("output/figures/HUB-sd-scores-vs-total.png", width = 7, height = 5)
+
+# log-log versions of that plot
+wis_vs_total(scores) + 
+  labs(y = "Mean WIS") + 
+  scale_x_continuous(labels = scale_fn, trans = "log10") +
+  scale_y_continuous(labels = scale_fn, trans = "log10") 
+ggsave("output/figures/HUB-mean-scores-vs-total-log-log.png", width = 7, height = 5)
+
+wis_vs_total(scores, summary_fct = stats::sd) + 
+  labs(y = "Sd WIS") + 
+  scale_x_continuous(labels = scale_fn, trans = "log10") +
+  scale_y_continuous(labels = scale_fn, trans = "log10") 
+ggsave("output/figures/HUB-sd-scores-vs-total-log-log.png", width = 7, height = 5)
+
+# correlation
+
+```

### 2022-07-08 main.tex
Update ECDC analasys and manuscript text

@@
 \newcommand{\changed}[1]{#1}
+
+% to write the simulated as iid symbol: 
+\newcommand{\distas}[1]{\mathbin{\overset{#1}{\kern\z@\sim}}}%
 
@@
 Two proper scoring rules commonly used in epidemiology and other fields are the Continuous Ranked Probability Score (CRPS) [Gneiting 2007] and the Weighted Interval Score (WIS) [CITATION].
-Due to their widespread use we focus on these two in this paper, ignoring other proper scoring rules.
+Due to their widespread use in epidemiology we focus on these two in this paper, ignoring other proper scoring rules.
 Both the WIS and CRPS represent a measure of the absolute distance between the predictive distribution and observation and can be thought of as a generalisation of the absolute error to full predictive distributions (see details in the SI).
 
-Evaluating forecasts based on absolute errors (for simplicity we use the term interchangeably with absolute distance, even though it technically only applies to point forecasts) may not always be desirable, especially in an epidemiological context.
+Evaluating forecasts based on absolute errors (for simplicity we use the term interchangeably with absolute distance, even though the former technically only applies to point forecasts) may not always be desirable, especially in an epidemiological context.
 % The reason for this is that generally, the way forecasts are evaluated should correspond to the error structure of the model and the type of errors models will usually make.
@@
 In this paper we argue that it may therefore often be helpful to shift the focus of the evaluation on multiplicative, rather than absolute errors.
-This can be achieved by transforming the forecasts as well as the observations using the logarithm before applying the WIS or CRPS.
-While we will mostly focus on the log transformation in this paper, the point we like to make is a broader one: it is possible to flexible change the focus of the evaluation by applying a multitude of different transformations before scoring forecasts.
-Instead of scoring multiplicative errors, one may for example like to apply a transformation that allows to score forecasters based on the shape of the trajectory they predicted.
-Furthermore, it is possible to capture different aspects in a single number by constructing composite scores as a weighted sum of scores based on different transformations.
-This allows forecast consumers to assign explicit weights to different qualities of the forecasts they deem to be important when assessing predictive performance.
+This is approximately equal to scoring the growth rate of a forecast and can be achieved by transforming the forecasts as well as the observations using the logarithm before applying the WIS or CRPS.
+We focus on the log transformation, as it represents a very convenient way of achieving our goal of evaluating how well a forecast captures the growth dynamic inherent to epidemiological process.
+However, we note that the same logic we use can also applied to other transformations which could shift the focus of evaluation in different ways.
+For example, given an appropriate transformation, it could also be possible to evaluate forecasters based on the shape of the trajectory they predicted.
+It would even be possible to construct composite scores as a weighted sum of scores based on different transformations, allowing forecast consumers to assign explicit weights to different qualities of the forecasts.
 
-In section [REF] we discuss which kinds of transformations can be applied.
-In section [REF] we give a detailed theoretical explanation of the rationale for applying the log transformation and discuss its interpretation.
+In section [REF] we discuss the broader idea of transforming forecasts, and in section [REF] give a detailed theoretical of the log transformation and its interpretation.
 In section [REF] we present an applied example using forecasts submitted to the European COVID-19 Forecast Hub.
@@
 
-\subsection{Permissible transformations / propriety} - JOHANNES
-
-In order to preserve the propriety of the score, transformations must be strictly monotonic.
-This makes sure that a forecast that is further away from the observed value will also be further away after the transformation.
-
-- WHAT KINDS OF DATA CAN I CONDITION ON? 
-- WOULD DIVIDING A FORECAST BY ITS RESOLUTION WORK? 
-- CAN I SCORE THE TRAJECTORY / WEEK-TO-WEEK GROWTH RATE BY DIVIDING the 4 week ahead forecast by the three week ahead forecast?
+The log transformation, among other transformations, is possible because it is a strictly monotonic transformation.
+A forecast that is closer to the target on the natural scale will still be closer to the target after applying a strictly monotonic transformation like the logarithm.
+For a single forecast, the ranking between forecasters would be preserved and the scores remain proper, meaning that forecasters still receive the best score in expectation if the predictive distribution equals the data-generating distribution.
+Differences in the ranking between forecasters only appear when averaging over multiple forecasts, as deviations from the perfect forecast get penalised differently depending on whether forecasts are transformed or not.
 
 \paragraph{First transform, then score, rather than the other way round}
-Note that simply taking the log of a score computed on the natural scale, i.e.
+Note that it is important to first apply the transformation and then score the transformed forecasts and observations.
+Simply taking the log of the WIS or the CRPS computed on the natural scale, i.e.
 computing $\log S(x, y)$ results in an improper score and is therefore not advised.
-We illustrate this point in Figure \ref{fig:log-improper} in the SI.
+This is because the logarithm of the WIS / CRPS does not penalise outliers enough and therefore incentivises overconfident predictions.
+We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can easily be seen that overconfident models perform best in terms of the log WIS.
 
-% \paragraph{Changes only come when you average across multiple scores}
-% For a single forecast, the ranking between forecasters would be preserved and the scores remain proper [Proof?].
-Differences in the ranking between forecasters only appear when averaging over multiple forecasts, as deviations from the perfect forecast get penalised differently depending on whether forecasts are transformed or not.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/example-log-first.png}
+    \caption{Scores for different forecasts evaluated the WIS and the log WIS.
+We simulated 1000 observations $Y_i = {\rm e}^{x_i}$, with $x_i \text{iid} \sim \mathcal{N}(0, 1)$.
+We then simulated 20 forecasters who would issue a predictive distribution $F = {\rm e}^{x_i}$, with $x \sim \mathcal{N}(0, \sigma)$, with values of $\sigma$ ranging from 0.1 to 2.}
+    \label{fig:log-improper}
+\end{figure}
 
-\paragraph{Example transformations}
+\paragraph{Other transformations}
+While we focus on the log transformation in this paper, there are other transformations that may be helpful in an epidemiological context.
+One could for example divide all forecasts by the population in a location, to obtain forecasts standardised incidences per e.g.
+100,000 inhabitants.
+Or one could apply a Box-Cox transformation MISSING. 
+Or a square-root transformation MISSING. 
+Or one could divide each forecast by the forecast of the previous week (and observations by the observation of the previous week), to transform preditions to week-to-week growth rates.
+This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory.
+Note that this is not well defined in the context of forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate.
 
-Take the log
-
-Divide by the latest known value = score forecast on the growth rate
-
-Log that = score log growth rate
-
-Divide by the previous for each time value, which gives you a weekly growth rate, i.e. a trajectory.
-
-Box-Cox
-
-square root WAS THAT USED SOMEWHERE?
+% Divide by the latest known value = score forecast on the growth rate
+% - WHAT KINDS OF DATA CAN I CONDITION ON? 
+% - WOULD DIVIDING A FORECAST BY ITS RESOLUTION WORK? 
+% - CAN I SCORE THE TRAJECTORY / WEEK-TO-WEEK GROWTH RATE BY DIVIDING the 4 week ahead forecast by the three week ahead forecast?
 
@@
 
-In this section (and those that follow) we focus on one particular transformation, taking the logarithm of forecasts and observations, which makes it possible to change the focus of the evaluation from absolute to relative errors.
+% In this section (and those that follow) we focus on one particular transformation, taking the logarithm of forecasts and observations, which makes it possible to change the focus of the evaluation from absolute to relative errors.
 To illustrate the effects of applying the logarithm intuitively, let us consider a point forecast $\hat{y}_{t+1}$ for a quantity of interest $y_{t+1}$, such that 
@@
 The analogon of the WIS / CRPS for point forecasts is the absolute error, and so the score will be determined by the size of the forecast error $\varepsilon_{t+1}$.
-When taking the logarithm of the forecast and the observation first, than the score is determined based on the size of $\varepsilon^*_{t+1}$:
+When taking the logarithm of the forecast and the observation first, then the score is determined based on the size of $\varepsilon^*_{t+1}$:
 \begin{equation}
@@
 Forecasting 101,000 rather than 100,000 cases will be treated equally to forecasting 2,020, rather than 2,000 hospitalisations.
-% Note that for a single forecast the relative rankings of different forecasts would be preserved. The transformation only makes a difference when aggregating predictions across different dimensions.
+While the absolute error on the log scale is exactly symmetric (e.g., predicting 2 instead of 1 would give the same error as predicting 0.5), this is not entirely true for the WIS.
+Due to the fact that upper and lower bounds of individual prediction intervals are affected differently, there may be slight deviations (compare e.g.
+in Figure \ref{fig:change-in-scores}D the scores for a relative error of $\frac{1}{5}$ and 5).
 
@@
                     &= \log (1 + g_{t, t+1}) - \log (1 + \hat{g}_{t, t+1}) \Leftrightarrow \\   
-                    &\approx g_{t, t+1} - \hat{g}_{t, t+1} \quad \text{for small } \hat{g}_{t, t+1}, g_{t, t+1}. 
+                    &\approx g_{t, t+1} - \hat{g}_{t, t+1} \quad \text{for small differences between} \hat{g}_{t, t+1} and g_{t, t+1}. 
 \end{align}
 
-This approximation breaks down when the growth rate becomes too large.
-If scoring the growth rate is the focus of the evaluation, then it may be better to transform forecasts and observations by the last known values instead in order to score the growth rate explicitly.
+This approximation breaks down when the difference between the predicted and the actual growth rate becomes too large (albeit forecasts are still evaluated in terms of relative errors).
+If this is of great concern one may want to use a different transformation like dividing forecasts and observations by the last known values instead in order to score the growth rate explicitly.
 
@@
 $n = 100$) to all observations and forecasts before taking the logarithm.
-This constitutes a strictly monotonic transformation which preserves propriety and would reduce all scores (through reducing relative errors).
+This also constitutes a strictly monotonic transformation which preserves propriety and would reduce all scores (through reducing relative errors).
 Scores for forecasts on small quantities would be particularly affected, resulting in a more equal distribution of scores for forecasts of differing orders of magntitudes.
@@
 \subsection{Relative importance when summarising scores}
-For a single transformation, the ranking between models is preserved, even though the numeric value of the score changes.
-However, when averaging across forecasts, the relative influence of individual forecasts on the aggregate score may differ dramatically, leading to changes in the overall model rankings.
 While the main rationale for applying the log transformation is that this better reflects the types of errors common in epidemiological modelling, it may also be helpful when scores based on the natural scale are very unequal.
@@
 
-While on the natural scale, aggregate scores will usually be dominated by predictions of quantities with high absolute values, the opposite may be true for scores based on log transformed predictions.
+On the natural scale, aggregate scores will usually be dominated by predictions of quantities with high absolute values.
+The opposite may be true for scores based on log transformed predictions.
 Small absolute errors (e.g.
@@
 For simulated negative binomial count data, log predictions for small quantities receive on average higher scores than forecasts for large quantities if the mean and the variance grow at the same rate $\sigma^2 = \mu$, about equal scores when the variance grows at a rate of $\sigma^2 = \mu + \mu^2$ (Figure \ref{fig:SIM-wis-state-size-mean}), and smaller scores when the variance grows faster than that.
-As mentioned above, one way to handle situations where scores are very unequal, is to add a number $n$ to all forecasts and observations in order to reduce the relative weight of forecast targets with small quantities.
+%As mentioned above, one way to handle situations where scores are very unequal, is to add a number $n$ to all forecasts and observations in order to reduce the relative weight of forecast targets with small quantities.
 
@@
 
-%COMPARISON TO SCORING THE ACTUAL GROWTH RATE - just harder to do and achieves the same thing? No objections if someone wants to? Might actually be better if there are negative values involved?
-
-%What about other things like scoring the square root of the forecast?
-
 %\begin{itemize}
-%    \item discuss how equations change for CRPS
 %    \item Optional: Discussion of parallels to point forecasts and the point that you're still incentivised to report the median\\
-%    \item discuss that even if both scoring rules are proper, they still penalise different things %differently. How does that change incentives for the forecaster? 
 %\end{itemize}
-
-\paragraph{Toy example: simulate an epidemic and apply 3 different models} Compare scores for these three models on the natural scale and the log scale
 
@@
 Across the different locations, the variance in observed values is roughly equal to the square of the mean, suggesting that, for a given location, both cases and deaths are not poisson-distributed, but rather exhibit over-dispersion.
-MAYBE FIT A NEGATIVE BINOMIAL TO THESE OBSERVATIONS
+% MAYBE FIT A NEGATIVE BINOMIAL TO THESE OBSERVATIONS
 
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-var-obs.png}
+    \caption{Observed mean against observed variance of observations.
+The slope of the dashed line is 2, which would represent a relationhip of Variance observed value = square of mean observed value.
+}
+    \label{fig:HUB-variance-mean}
+\end{figure}
+
 \paragraph{Features of the scores}
@@
 On the natural scale, scores vary by several orders of magnitudes across locations (see \ref{fig:HUB-scores-location}), whereas variation across locations is much smaller for scores on the log scale and wis values are more evenly distributed.
-The ordering of scores across locations is not preserved when going from the natural to the log scale.
-% Rather, there seems to be a slight inverse relationship that locations which previously had low scores now tend to receive higher scores. 
+The ordering of scores across locations is not preserved when going from the natural to the log scale (see Figure \ref{fig:HUB-mean-location}.
+Rather, there seems to be a slight inverse relationship that locations which previously had low scores now tend to receive higher scores.
 The correlation between scores on the log and natural scale for different locations is XX for cases and XX for deaths. 
-%THERE SEEMS TO BE A TENDENCY FOR SMALLER
+BUT: CORRELATION FOR RANKINGS IS STILL LAGRGE
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-ranking-diffs.png}
+    \caption{Distribution of rankings of the different Hub models.
+Rankings are computed as relative rankings, i.e.
+the ranking divided by the total number of models available for a given forecast target.}
+    \label{fig:HUB-rankings}
+\end{figure}
 
@@
     \label{fig:HUB-average-scores}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-scores-locations.png}
-    \caption{Weighted interval scores across different locations for two week ahead forecasts of COVID-19 cases and deaths evaluated on the natural scale (left) and on the log scale (right). }
-    \label{fig:HUB-scores-location}
 \end{figure}
@@
 
+\paragraph{Rankings}
+Rankings do change a little bit, but not massively. More difference comes when we look at greater forecast horizons.
+Really this is more a point about how rankings change even for one single forecast (which I originally thought they shouldn't).
+Would be good to tease out what happens here.
+Somewhow we do change our judegement on which model is better.
+Maybe the preservation of the order really only holds for one interval?
+Or maybe not even that?
+Example: True value = 1
+Forecaster 1: 0.8, 3.     Natural scale: 3.2 - 0.8 = 2.4.     Log scale: log 3 - log 0.8 = 1.386
+Forecaster 2: 0.5, 2.8    Natural scale: 2.8 - 0.5 = 2.3    Log scale: log (2.8/0.5) = 1.722767
+So one effect is that for a given dispersion on the natural scale, it's better if that interval is skewed, as that implies that at least you only off a lot in one direction in negative terms, and not in the other direction.
+Maybe looking at over-and underprediction is the key here.
+%ADD SAME PLOT TO SI FOR HORIZON 4.
+
+\paragraph{pairwise comparisons}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-pairwise-comparisons.png}
+    \caption{Changes in relative skill and the distribution of scores}
+    \label{fig:HUB-pairwise}
+\end{figure}
+
+Correlation between pairwise comparisons is 0.4414717.
+
 % \section{Other ideas}
@@
 \paragraph{summary of what we did}
+In this paper we suggested that epidemiological forecasts which are today evalaluated using the WIS or CRPS on a natural scale should instead be log transformed first before scoring.
+This shifts the focus of the evaluation from absolute to relative errors and is a convenient way to score forecasts based on how well they are able to predict the growth rate.
+This is useful as it makes sure that the evaluation is consistent with the kinds of errors we expect forecasting models to make in an epidemiological context.
+
+When applied to data from the Europeam Forecasting Hub, we found that
 

### 2022-07-08 analysis/example-log-first-Johannes.R

@@ -1,0 +1,42 @@
+library(dplyr)
+library(tidyr)
+library(scoringutils)
+library(ggplot2)
+
+n_sim <- 1000
+epsilon <- rnorm(n_sim)
+Y <- exp(epsilon)
+
+forecasts <- expand.grid(
+  sigma = 1:20/10, 
+  quantile = c(0.01, 0.025, 1:19/20, 0.975, 0.99)
+)
+
+forecasts <- forecasts |>
+  as_tibble() |>
+  mutate(model = 10 * sigma, 
+         prediction = exp(qnorm(quantile, sd = sigma)), 
+         true_value = list(Y), 
+         sample_id = list(1:length(Y))) |>
+  unnest(c(true_value, sample_id))
+
+
+scores <- score(forecasts)
+
+summary <- scores |>
+  mutate(log_wis = log(interval_score +1)) |>
+  group_by(sigma) |>
+  summarise(wis = mean(interval_score), 
+            log_wis = mean(log_wis))
+
+summary |> 
+  pivot_longer(cols = c(wis, log_wis), values_to = "score", names_to = "type") |>
+  mutate(type = factor(type, levels = c("wis", "log_wis"), labels = c("WIS", "log(WIS)"))) |>
+  ggplot(aes(x = sigma, y = score)) +
+  geom_vline(xintercept = 1, linetype = "dashed", size = 0.3, color = "grey80") +
+  geom_point() + 
+  facet_wrap(~type, scales = "free_y") + 
+  theme_scoringutils() +
+  labs(y = "Score", x = "Standard deviation of predictive distribution")
+
+ggsave("output/figures/example-log-first.png", width = 7, height = 5)

### 2022-07-15 analysis/Analysis ECDC forecasts.Rmd
update figures, manuscript, add citations

@@
 
-ggsave("output/figures/HUB-mean-obs-location.png", width = 9, height = 8)
+ggsave("output/figures/HUB-mean-obs-location.png", width = 9, height = 5.5)
 
@@
 
+Figure that shows Variance vs. mean of the observations. Makes the point that the
+data looks neg. binomial. 
 ```{r}
@@
   labs(y = "Variance observed value", x = "Mean observed value") + 
-  geom_function(fun = function(x) x^2, linetype = "dashed", size = 0.5) + 
+  geom_function(fun = function(x) {x + x^2}, linetype = "dashed", size = 0.5) + 
   theme_scoringutils() + 
@@
 
-ggsave("output/figures/HUB-mean-var-obs.png", width = 7, height = 4)
+ggsave("output/figures/HUB-mean-var-obs.png", width = 7, height = 2.5)
 
@@
 
-*Scores are more comparable across targets*
-
 Mean scores: 
@@
 
-## Scores over time
-restrict to a few locations, e.g. GB, FR, ES (highest average scores there)
-
-```{r}
-scores |> 
-  filter(model == "EuroCOVIDhub-ensemble",
-         location %in% c("GB", "ES", "DE"),
-         horizon == 2) |>
-  rename(Location = location) |>
-  group_by(target_type, scale, Location) |>
-  mutate(multiplicative_score = interval_score / min(interval_score)) |>
-  ggplot(aes(y = multiplicative_score, x = forecast_date, colour = Location)) + 
-  geom_line() + 
-  geom_point() + 
-  facet_grid(target_type ~ scale, scale = "free_y") + 
-  theme_scoringutils() + 
-  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
-
-ggsave("output/figures/HUB-scores-over-time.png", width = 7, height = 5)
-```
-
-*Overall it looks like there can still be substantial fluctuation in log scores, but maybe slightly less than when scored on an absolute scale.
-Interestingly, it seems like the smoothing effect is stronger for deaths than for cases.*
-
 ## Change of average scores for increasing forecast horizons
@@
   geom_point() + 
-  facet_grid(target_type ~ scale) + 
+  facet_grid(scale ~ target_type) + 
   theme_scoringutils() + 
@@
 
-ggsave("output/figures/HUB-scores-over-horizon.png", width = 7, height = 5)
+ggsave("output/figures/HUB-scores-over-horizon.png", width = 7, height = 4)
 
@@
 
-*Slightly less increase on the log scale, but overall not a lot of difference*
-
-## Actual model rankings for single forecasts
+## Comparison of the distribution of model rankings
 
 ```{r}
-# we see that rankings change even for a single forecast. So we need to check everything is still correct. 
-# take one example
-test_data <- hub_data |>
-  filter(model %in% c("MUNI-ARIMA", "epiforecasts-EpiNow2"), 
-         forecast_date == "2021-10-18", 
-         target_type == "Cases", 
-         location == "AT")
-
-test_data |>
-  score(metrics = c("interval_score")) |>
-  summarise_scores()
-
-test_data |>
-  mutate(prediction = log(prediction), 
-         true_value = log(true_value)) |>
-  score(metrics = c("interval_score")) |>
-  summarise_scores()
-## indeed everything seems to be correct, but rankings still change FOR A SINGLE FORECAST. 
-# It seems this is because the trade-off between calibration and sharpness changes? 
-# need to check in what way
 
 rankings <- scores |>
-  select(-c(coverage_deviation, bias, aem, sharpness, underprediction, overprediction)) |>
+  select(-c(coverage_deviation, bias, ae_median, dispersion, underprediction, overprediction)) |>
   group_by(location, forecast_date, horizon, target_type, scale) |>
   mutate(ranking = rank(interval_score) / n()) |>
-   filter(forecast_date == "2021-10-18", 
-         target_type == "Cases", 
-         location == "AT")
+  ungroup()
 
+rankings_diff <- rankings |>
+  select(model, location, target_end_date, forecast_date, horizon,
+         interval_score, scale, ranking, target_type) |>
+  pivot_wider(names_from = scale, values_from = ranking, 
+              id_cols = -interval_score)
+
+plot_ranking_diffs <- function(
+  rankings, 
+  rankings_diff,
+  target_type
+) {
+  type <- target_type
+  # maybe replace this by the same ranking for all plots
   factor_levels <- rankings |>
-  filter(horizon == 2, scale == "natural") |>
+    filter(horizon == 2, scale == "natural", target_type == type) |>
     group_by(model) |>
@@
 
-p_rankings <- rankings |>
-  filter(horizon == 2) |>
-  mutate(model = factor(model, levels = factor_levels)) |>
-  ggplot(aes(y = model, x = ranking)) +
-  ggdist::stat_halfeye() +
-  facet_wrap(~ scale) +
-  theme_scoringutils() +
-  labs(y = "Model", x = "Normalised ranking")
+  # p_rankings <- rankings |>
+  #   filter(horizon == 2) |>
+  #   mutate(model = factor(model, levels = factor_levels)) |>
+  #   ggplot(aes(y = model, x = ranking)) +
+  #   ggdist::stat_halfeye() +
+  #   facet_wrap(~ scale) +
+  #   theme_scoringutils() +
+  #   labs(y = "Model", x = "Normalised ranking")
 
-rankings_diff <- rankings |>
-  select(model, location, target_end_date, forecast_date, interval_score, scale, ranking) |>
-  pivot_wider(names_from = scale, values_from = ranking, id_cols = -interval_score)
-
-p_diff_ranking <- rankings_diff |>
-  filter(horizon == 2) |>
+  ranking_plot <- function(
+    hor
+  ) {
+    p <- rankings_diff |>
       mutate(diff_ranking = natural - log) |>
       mutate(model = factor(model, levels = factor_levels)) |>
+      filter(horizon == hor, 
+             !is.na(model)) |>
       ggplot(aes(y = model, x = diff_ranking)) +
-  ggdist::stat_halfeye() +
+      geom_vline(linetype = "dashed", xintercept = 0, 
+                 size = 0.6, color = "grey40") +
+      ggdist::stat_pointinterval() +
       theme_scoringutils() +
-  labs(y = "", x = "Differences in ranking") +
+      labs(y = "", x = "Differences in ranking")
+
+    if (hor > 1) {
+      p <- p +
         theme(axis.text.y = element_blank(), 
@@
               axis.line.y = element_blank())
+    }
+    return(p)
+  }
 
-p_rankings + p_diff_ranking + 
-  plot_layout(widths = c(2, 1))
+  ranking_plot(hor = 1) + ranking_plot(2) + ranking_plot(3) + ranking_plot(4) +
+    plot_layout(nrow = 1)
+}
 
-ggsave("output/figures/HUB-ranking-diffs.png", width = 9, height = 5)
+# should probably have one plot for both cases and deaths since this is rankings
+p_diff_cases <- plot_ranking_diffs(rankings, rankings_diff, 
+                                   target_type = "Cases")
+
+ggsave(filename = "output/figures/HUB-ranking-diffs.png", 
+       plot = p_diff_cases,
+       width = 9, height = 5)
 
@@
 
-## Distribution of scores
+--- 
+# Code parking lot
+
+## Scores over time
+restrict to a few locations, e.g. GB, FR, ES (highest average scores there)
 
@@
 scores |> 
-  ggplot(aes(x = interval_score)) + 
-  geom_density() + 
-  facet_wrap(~ target_type + scale, scale = "free") + 
-  scale_x_continuous(labels = scale_fn)
-```
-
-## Fit a model to data
-
-Check fit of the log scale model
-
-```{r echo=TRUE}
-reg_log <- scores |>
-  filter(scale == "log") |>
-  glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 
-      family = gaussian)
-```
-
-```{r}
-# check distribution
-performance::check_distribution(reg_log) |>
-  plot()
-
-# check normality of residuals
-check_normality(reg_log) |>
-  plot(type = "qq")
-
-# heteroskedasticity of error terms - the plot takes extremly long to run
-performance::check_heteroskedasticity(reg_log)
-# performance::check_heteroskedasticity(reg_log) |>
-#   plot()
-
-# check influence of outliers - plot also takes very long to run
-performance::check_outliers(reg_log)
-# performance::check_outliers(reg_log) |>
-#   plot()
-
-```
-
-## Model rankings
-
-Still to do
-
-- restrict to full set of observations
-- check how rankings change
-
-```{r}
-# find out which models have submitted forecasts for all locations and time points
-
-```
-
----
-
-# Looking at growth rates directly
-
-## Scores of the growth rate across targets
-
-This uses the Hub ensemble
-
-```{r}
-scores_gr |> 
   filter(model == "EuroCOVIDhub-ensemble",
+         location %in% c("GB", "ES", "DE"),
          horizon == 2) |>
-  ggplot(aes(y = interval_score, x = target_type)) + 
-  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
-  geom_boxplot(alpha = 0.5) + 
-  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
-  facet_wrap(~ scale, scale = "free") + 
-  theme_scoringutils() + 
-  theme(legend.position = "none") + 
-  scale_y_continuous(labels = scale_fn, trans = "log10") + 
-  labs(y = "Interval score", x = "Target type")
-```
-
-*Scoring the growth rate makes things more comparable*
-
-Mean scores: 
-```{r}
-scores_gr |> 
-  filter(model == "EuroCOVIDhub-ensemble", 
-         horizon == 2) |>
-  group_by(scale, target_type) |>
-  summarise(interval_score = mean(interval_score), 
-            .groups = "drop_last") |>
-  make_table()
-```
-
-## Scores for the growth rate across locations
-
-```{r}
-scores_gr |> 
-  filter(model == "EuroCOVIDhub-ensemble", 
-         horizon == 2) |>
-  group_by(target_type, location, scale) |>
-  summarise(interval_score = mean(interval_score), 
-            .groups = "drop_last") |>
-  ggplot(aes(y = interval_score, x = reorder(location, -interval_score))) + 
-  geom_bar(stat = "identity") + 
-  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
-  facet_wrap(~ target_type + scale, scale = "free") + 
-  theme_scoringutils() + 
-  theme(legend.position = "none") + 
-  scale_y_continuous(labels = scale_fn) + 
-  labs(y = "Interval score", x = "Target type")
-```
-
-## Scores for growth rate over time
-
-restrict to a few locations, e.g. GB, FR, ES (highest average scores there)
-
-UK:
-
-```{r}
-scores_gr |> 
-  filter(model == "EuroCOVIDhub-ensemble",
-         location == "GB",
-         horizon == 2) |>
-  group_by(target_type, scale) |>
+  rename(Location = location) |>
+  group_by(target_type, scale, Location) |>
   mutate(multiplicative_score = interval_score / min(interval_score)) |>
-  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
+  ggplot(aes(y = multiplicative_score, x = forecast_date, colour = Location)) + 
   geom_line() + 
@@
   labs(x = "Forecast date", y = "WIS relative to smallest observed value")
+
+ggsave("output/figures/HUB-scores-over-time.png", width = 7, height = 5)
 ```
 
-France: Looks like a data revision
-
-```{r}
-scores_gr |> 
-  filter(model == "EuroCOVIDhub-ensemble",
-         location == "FR",
-         horizon == 2) |>
-  group_by(target_type, scale) |>
-  mutate(multiplicative_score = interval_score / min(interval_score)) |>
-  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
-  geom_line() + 
-  geom_point() + 
-  facet_grid(target_type ~ scale, scale = "free_y") + 
-  theme_scoringutils() + 
-  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
-```
-
-Spain:
-
-```{r}
-scores_gr |> 
-  filter(model == "EuroCOVIDhub-ensemble",
-         location == "ES",
-         horizon == 2) |>
-  group_by(target_type, scale) |>
-  mutate(multiplicative_score = interval_score / min(interval_score)) |>
-  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
-  geom_line() + 
-  geom_point() + 
-  facet_grid(target_type ~ scale, scale = "free_y") + 
-  theme_scoringutils() + 
-  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
-```
-
-## Change of average scores for the growth rate for increasing forecast horizons
-
-```{r}
-scores_gr |> 
-  filter(model == "EuroCOVIDhub-ensemble") |>
-  group_by(scale, target_type, horizon) |>
-  summarise(interval_score = mean(interval_score), 
-            .groups = "drop_last") |>
-  mutate(multiplicative_score = interval_score / min(interval_score)) |>
-  ggplot(aes(y = multiplicative_score, x = horizon)) + 
-  geom_line() + 
-  geom_point() + 
-  facet_grid(target_type ~ scale) + 
-  theme_scoringutils() + 
-  labs(x = "Forecast horizon in weeks", y = "Multplicative change in interval score")
-```
-
-## Distribution of scores growth rate
-
-```{r}
-scores_gr |>
-  ggplot(aes(x = interval_score)) + 
-  geom_density() + 
-  facet_wrap(~ target_type + scale, scale = "free") + 
-  scale_x_continuous(labels = scale_fn)
-```
-
-## Fit a model to growth rate data
-
-Check fit of the log scale model
-
-```{r echo=TRUE}
-reg_gr <- scores_gr |>
-  filter(scale == "natural") |>
-  glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 
-      family = gaussian)
-
-reg_gr_log <- scores_gr |>
-  filter(scale == "log") |>
-  glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 
-      family = gaussian)
-```
-
-```{r}
-# check distribution
-performance::check_distribution(reg_gr_log) |>
-  plot()
-
-# check normality of residuals
-check_normality(reg_gr_log) |>
-  plot(type = "qq")
-
-# heteroskedasticity of error terms - the plot takes extremly long to run
-performance::check_heteroskedasticity(reg_gr_log)
-# performance::check_heteroskedasticity(reg_log) |>
-#   plot()
-
-# check influence of outliers - plot also takes very long to run
-performance::check_outliers(reg_gr_log)
-# performance::check_outliers(reg_log) |>
-#   plot()
-
-```
-
-## Code parking lot
+*Overall it looks like there can still be substantial fluctuation in log scores, but maybe slightly less than when scored on an absolute scale.
+Interestingly, it seems like the smoothing effect is stronger for deaths than for cases.*
 
@@
 ```
+
+Test for a single example forecast to verify that rankings change
+```{r}
+# we see that rankings change even for a single forecast. So we need to check everything is still correct. 
+# take one example
+test_data <- hub_data |>
+  filter(model %in% c("MUNI-ARIMA", "epiforecasts-EpiNow2"), 
+         forecast_date == "2021-10-18", 
+         target_type == "Cases", 
+         location == "AT")
+
+test_data |>
+  score(metrics = c("interval_score")) |>
+  summarise_scores()
+
+test_data |>
+  mutate(prediction = log(prediction), 
+         true_value = log(true_value)) |>
+  score(metrics = c("interval_score")) |>
+  summarise_scores()
+## indeed everything seems to be correct, but rankings still change FOR A SINGLE FORECAST. 
+# It seems this is because the trade-off between calibration and sharpness changes? 
+# need to check in what way
+```
+
+## Distribution of scores
+
+```{r}
+scores |>
+  ggplot(aes(x = interval_score)) + 
+  geom_density() + 
+  facet_wrap(~ target_type + scale, scale = "free") + 
+  scale_x_continuous(labels = scale_fn)
+```
+
+## Fit a model to data
+
+Check fit of the log scale model
+```{r echo=TRUE}
+reg_log <- scores |>
+  filter(scale == "log") |>
+  glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 
+      family = gaussian)
+```
+
+```{r}
+# check distribution
+performance::check_distribution(reg_log) |>
+  plot()
+
+# check normality of residuals
+check_normality(reg_log) |>
+  plot(type = "qq")
+
+# heteroskedasticity of error terms - the plot takes extremly long to run
+performance::check_heteroskedasticity(reg_log)
+# performance::check_heteroskedasticity(reg_log) |>
+#   plot()
+
+# check influence of outliers - plot also takes very long to run
+performance::check_outliers(reg_log)
+# performance::check_outliers(reg_log) |>
+#   plot()
+
+```
+
+---
+
+# Growth Rate stuff - Looking at growth rates directly
+
+## Scores of the growth rate across targets
+
+This uses the Hub ensemble
+```{r}
+scores_gr |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  ggplot(aes(y = interval_score, x = target_type)) + 
+  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
+  geom_boxplot(alpha = 0.5) + 
+  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
+  facet_wrap(~ scale, scale = "free") + 
+  theme_scoringutils() + 
+  theme(legend.position = "none") + 
+  scale_y_continuous(labels = scale_fn, trans = "log10") + 
+  labs(y = "Interval score", x = "Target type")
+```
+*Scoring the growth rate makes things more comparable*
+
+Mean scores: 
+```{r}
+scores_gr |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  group_by(scale, target_type) |>
+  summarise(interval_score = mean(interval_score), 
+            .groups = "drop_last") |>
+  make_table()
+```
+
+## Scores for the growth rate across locations
+```{r}
+scores_gr |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  group_by(target_type, location, scale) |>
+  summarise(interval_score = mean(interval_score), 
+            .groups = "drop_last") |>
+  ggplot(aes(y = interval_score, x = reorder(location, -interval_score))) + 
+  geom_bar(stat = "identity") + 
+  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
+  facet_wrap(~ target_type + scale, scale = "free") + 
+  theme_scoringutils() + 
+  theme(legend.position = "none") + 
+  scale_y_continuous(labels = scale_fn) + 
+  labs(y = "Interval score", x = "Target type")
+```
+
+## Scores for growth rate over time
+restrict to a few locations, e.g. GB, FR, ES (highest average scores there)
+
+UK: 
+```{r}
+scores_gr |> 
+  filter(model == "EuroCOVIDhub-ensemble",
+         location == "GB",
+         horizon == 2) |>
+  group_by(target_type, scale) |>
+  mutate(multiplicative_score = interval_score / min(interval_score)) |>
+  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
+  geom_line() + 
+  geom_point() + 
+  facet_grid(target_type ~ scale, scale = "free_y") + 
+  theme_scoringutils() + 
+  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
+```
+
+France: Looks like a data revision
+```{r}
+scores_gr |> 
+  filter(model == "EuroCOVIDhub-ensemble",
+         location == "FR",
+         horizon == 2) |>
+  group_by(target_type, scale) |>
+  mutate(multiplicative_score = interval_score / min(interval_score)) |>
+  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
+  geom_line() + 
+  geom_point() + 
+  facet_grid(target_type ~ scale, scale = "free_y") + 
+  theme_scoringutils() + 
+  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
+```
+
+Spain: 
+```{r}
+scores_gr |> 
+  filter(model == "EuroCOVIDhub-ensemble",
+         location == "ES",
+         horizon == 2) |>
+  group_by(target_type, scale) |>
+  mutate(multiplicative_score = interval_score / min(interval_score)) |>
+  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
+  geom_line() + 
+  geom_point() + 
+  facet_grid(target_type ~ scale, scale = "free_y") + 
+  theme_scoringutils() + 
+  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
+```
+
+## Change of average scores for the growth rate for increasing forecast horizons
+```{r}
+scores_gr |> 
+  filter(model == "EuroCOVIDhub-ensemble") |>
+  group_by(scale, target_type, horizon) |>
+  summarise(interval_score = mean(interval_score), 
+            .groups = "drop_last") |>
+  mutate(multiplicative_score = interval_score / min(interval_score)) |>
+  ggplot(aes(y = multiplicative_score, x = horizon)) + 
+  geom_line() + 
+  geom_point() + 
+  facet_grid(target_type ~ scale) + 
+  theme_scoringutils() + 
+  labs(x = "Forecast horizon in weeks", y = "Multplicative change in interval score")
+```
+
+## Distribution of scores growth rate
+
+```{r}
+scores_gr |>
+  ggplot(aes(x = interval_score)) + 
+  geom_density() + 
+  facet_wrap(~ target_type + scale, scale = "free") + 
+  scale_x_continuous(labels = scale_fn)
+```
+
+## Fit a model to growth rate data
+
+Check fit of the log scale model
+```{r echo=TRUE}
+reg_gr <- scores_gr |>
+  filter(scale == "natural") |>
+  glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 
+      family = gaussian)
+
+reg_gr_log <- scores_gr |>
+  filter(scale == "log") |>
+  glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 
+      family = gaussian)
+```
+
+Check distribution
+```{r}
+# check distribution
+performance::check_distribution(reg_gr_log) |>
+  plot()
+
+# check normality of residuals
+check_normality(reg_gr_log) |>
+  plot(type = "qq")
+
+# heteroskedasticity of error terms - the plot takes extremly long to run
+performance::check_heteroskedasticity(reg_gr_log)
+# performance::check_heteroskedasticity(reg_log) |>
+#   plot()
+
+# check influence of outliers - plot also takes very long to run
+performance::check_outliers(reg_gr_log)
+# performance::check_outliers(reg_log) |>
+#   plot()
+
+```

### 2022-07-15 main.tex
update figures, manuscript, add citations

@@
 \linenumbers
+\bibliographystyle{apalike}
 
@@
 
-Probabilistic forecasts play an important role in decision-making in epidemiology and public health [Citations], as well as other areas as diverse as for example economics [Citations] or agriculture [Citations].
+Probabilistic forecasts play an important role in decision-making in epidemiology and public health \citep{reichCollaborativeMultiyearMultimodel2019, funkShorttermForecastsInform2020, cramerEvaluationIndividualEnsemble2021, bracherShorttermForecastingCOVID192021, europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}, as well as other areas as diverse as for example economics \citep{timmermannForecastingMethodsFinance2018, elliottForecastingEconomicsFinance2016} or meteorology \citep{gneitingWeatherForecastingEnsemble2005, kukkonenReviewOperationalRegionalscale2012}.
 Epidemiological modelling in particular has received widespread attention during the COVID-19 pandemic and has been an important tool to inform public policy.
-Evaluating forecasts is important, as the evaluation provides feedback for forecasters and researchers necessary to improve their models, and it helps decision makers with distinguishing good from bad predictions and choosing forecasts that inform future decisions.
+Evaluating forecasts is important, as the evaluation provides feedback for forecasters and researchers necessary to improve their models, and it helps decision makers distinguish good from bad predictions and choosing forecasts that inform future decisions.
 
@@
 Forecasters, in expectaction, optimise their score by providing a predictive distribution that is equal to the data-generating distribution, and will be penalised if they fail to do so.
-Two proper scoring rules commonly used in epidemiology and other fields are the Continuous Ranked Probability Score (CRPS) [Gneiting 2007] and the Weighted Interval Score (WIS) [CITATION].
+Two proper scoring rules commonly used in epidemiology and other fields are the Continuous Ranked Probability Score (CRPS) \citep{gneitingStrictlyProperScoring2007} and the Weighted Interval Score (WIS) \citep{bracherEvaluatingEpidemicForecasts2021}.
 Due to their widespread use in epidemiology we focus on these two in this paper, ignoring other proper scoring rules.
@@
 Evaluating forecasts based on absolute errors (for simplicity we use the term interchangeably with absolute distance, even though the former technically only applies to point forecasts) may not always be desirable, especially in an epidemiological context.
-% The reason for this is that generally, the way forecasts are evaluated should correspond to the error structure of the model and the type of errors models will usually make.
+% The reason for this is that generally, the way for ecasts are evaluated should correspond to the error structure of the model and the type of errors models will usually make.
 Epidemiological processes are usually understood to be exponential in nature and therefore are most commonly described and modeled in multiplicative terms.
@@
 
-In this paper we argue that it may therefore often be helpful to shift the focus of the evaluation on multiplicative, rather than absolute errors.
+In this paper we argue that it may therefore be helpful to shift the focus of the evaluation on multiplicative, rather than absolute errors.
 This is approximately equal to scoring the growth rate of a forecast and can be achieved by transforming the forecasts as well as the observations using the logarithm before applying the WIS or CRPS.
@@
 In section [REF] we discuss the broader idea of transforming forecasts, and in section [REF] give a detailed theoretical of the log transformation and its interpretation.
-In section [REF] we present an applied example using forecasts submitted to the European COVID-19 Forecast Hub.
-The European COVID-19 Forecast Hub is one of several Forecast hubs (together with a Hub in the US [CITATION] and one in Germany and Poland [CITATION]) that systematically collected, aggregated and evaluated forecasts of several COVID-19 targets created by different teams.
+In section [REF] we present an applied example using forecasts submitted to the European COVID-19 Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
+The European COVID-19 Forecast Hub is one of several Forecast hubs (together with a Hub in the US \citep{cramerEvaluationIndividualEnsemble2021} and one in Germany and Poland \citep{bracherShorttermForecastingCOVID192021} that systematically collected, aggregated and evaluated forecasts of several COVID-19 targets created by different teams.
 Forecasts submitted to the COVID-19 Forecast Hubs are probabilistic in nature [Citation e.g.
@@
 
-The log transformation, among other transformations, is possible because it is a strictly monotonic transformation.
-A forecast that is closer to the target on the natural scale will still be closer to the target after applying a strictly monotonic transformation like the logarithm.
-For a single forecast, the ranking between forecasters would be preserved and the scores remain proper, meaning that forecasters still receive the best score in expectation if the predictive distribution equals the data-generating distribution.
-Differences in the ranking between forecasters only appear when averaging over multiple forecasts, as deviations from the perfect forecast get penalised differently depending on whether forecasts are transformed or not.
+Applying the log transformation, among other transformations, is permissible because it is a strictly monotonic transformation.
+While rankings between different models may indeed change, even for a single forecast, the relevant condition for a scoring rule to be strictly proper is still satisfied: In expectation a forecaster will receive the best score if she reports a predictive distribution that is equal to the data-generating distribution, even after transforming both the forecasts and the observations (representing draws from the data-generating distribution).
 
-\paragraph{First transform, then score, rather than the other way round}
-Note that it is important to first apply the transformation and then score the transformed forecasts and observations.
+It is important to first apply the transformation and then score the transformed forecasts and observations.
 Simply taking the log of the WIS or the CRPS computed on the natural scale, i.e.
@@
 
-\paragraph{Other transformations}
+\paragraph{Other transformations - maybe move to Discussion}
 While we focus on the log transformation in this paper, there are other transformations that may be helpful in an epidemiological context.
@@
 This may be the case for example when comparing forecasts across locations or time points, rather than different forecasts targets.
-\citep{Bracher} for example argue that in epidemiological settings it is a desirable property of the CRPS / WIS that it assigns large scores to forecasts where counts are high, as this naturally reflects the situations we should most care about and helps us identify the models that perform best in similar situations.
+\cite{bracherEvaluatingEpidemicForecasts2021} for example argue that in epidemiological settings it is a desirable property of the CRPS / WIS that it assigns large scores to forecasts where counts are high, as this naturally reflects the situations we should most care about and helps us identify the models that perform best in similar situations.
 If, on the other hand, we believe that good models should do consistently well and that situations with high incidences do not provide significantly more information about which models perform well or badly, then scoring on the log scale may be more appropriate.
@@
 
-\paragraph{Introduction to the Forecast Hub} As an empirical example for evaluating forecasts on the natural and on the log scale we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021}.
+\paragraph{Introduction to the Forecast Hub} As an empirical example for evaluating forecasts on the natural and on the log scale we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
 Every week the European Forecast Hub collates and aggregates forecasts for different COVID-19 related targets from teams around the world.
@@
     \caption{Observed mean against observed variance of observations.
-The slope of the dashed line is 2, which would represent a relationhip of Variance observed value = square of mean observed value.
+The dashed line marks what would be expecetd from a negative binomially distributed variable with variance of the observed value equal to the mean + the square of mean observed value.
 }
@@
 
+\bibliography{log-or-not.bib}
+
 \end{document}

### 2022-07-15 analysis/example-log-first-Johannes.R

@@ -39,4 +39,4 @@ summary |>
   theme_scoringutils() +
   labs(y = "Score", x = "Standard deviation of predictive distribution")
 
-ggsave("output/figures/example-log-first.png", width = 7, height = 5)
+ggsave("output/figures/example-log-first.png", width = 7, height = 2.5)

### 2022-07-15 main.tex
update manuscript

@@
 Two proper scoring rules commonly used in epidemiology and other fields are the Continuous Ranked Probability Score (CRPS) \citep{gneitingStrictlyProperScoring2007} and the Weighted Interval Score (WIS) \citep{bracherEvaluatingEpidemicForecasts2021}.
-Due to their widespread use in epidemiology we focus on these two in this paper, ignoring other proper scoring rules.
-Both the WIS and CRPS represent a measure of the absolute distance between the predictive distribution and observation and can be thought of as a generalisation of the absolute error to full predictive distributions (see details in the SI).
+Both the WIS and CRPS represent a measure of the absolute distance between the predictive distribution and observation and represent a generalisation of the absolute error to full predictive distributions (see details in the SI).
+The WIS can be thought of as an approxmiation of the CRPS for forecasts in a quantile format and has very similar properties (the WIS converges towards the CRPS for an increasing set of equally-spaced prediction intervals).
+Due to its widespread use in epidemiology (and its similarity to the CRPS) we focus on the WIS in this paper, ignoring other proper scoring rules.
 
-Evaluating forecasts based on absolute errors (for simplicity we use the term interchangeably with absolute distance, even though the former technically only applies to point forecasts) may not always be desirable, especially in an epidemiological context.
+% Maybe insert WIS formula here?
+
+Evaluating forecasts based on absolute errors (for simplicity we use the term interchangeably with absolute distance, even though the former technically only applies to point forecasts), as is done by the WIS, may not always be desirable, especially in an epidemiological context.
 % The reason for this is that generally, the way for ecasts are evaluated should correspond to the error structure of the model and the type of errors models will usually make.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Transforming forecasts to shift the focus of the evaluation}
+\section{Log transforming forecasts and observations}
 
@@
 
-\paragraph{Other transformations - maybe move to Discussion}
-While we focus on the log transformation in this paper, there are other transformations that may be helpful in an epidemiological context.
-One could for example divide all forecasts by the population in a location, to obtain forecasts standardised incidences per e.g.
-100,000 inhabitants.
-Or one could apply a Box-Cox transformation MISSING. 
-Or a square-root transformation MISSING. 
-Or one could divide each forecast by the forecast of the previous week (and observations by the observation of the previous week), to transform preditions to week-to-week growth rates.
-This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory.
-Note that this is not well defined in the context of forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate.
+Unfortunately, the log transformation cannot readily be applied when dealing with negative or zero values.
+In an epidemiological setting that involves count data, negative values should likely be omitted entirely.
+Values of zero should potentially also be removed from the analsys, as it does not make sense to interpret relative errors on zero counts.
+Alternatively, if one is not willing to omit them, a small quantity (such as 1) could be added to all observations before taking the logarithm.
+Note that predictive distribtions have to be truncated in a similar way.
+When dealing with count data we could extend this argument to all small values (rather than just zeros), as it is hard to reliably evaluate relative errors when dealing with small discrete values.
+One could therefore decide to remove forecasts for small quantities below a certain threshold from the analysis.
 
-% Divide by the latest known value = score forecast on the growth rate
-% - WHAT KINDS OF DATA CAN I CONDITION ON? 
-% - WOULD DIVIDING A FORECAST BY ITS RESOLUTION WORK? 
-% - CAN I SCORE THE TRAJECTORY / WEEK-TO-WEEK GROWTH RATE BY DIVIDING the 4 week ahead forecast by the three week ahead forecast?
+\subsection{Scoring multplicative errors}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Log transforming forecasts and observations}
-
-% In this section (and those that follow) we focus on one particular transformation, taking the logarithm of forecasts and observations, which makes it possible to change the focus of the evaluation from absolute to relative errors.
 To illustrate the effects of applying the logarithm intuitively, let us consider a point forecast $\hat{y}_{t+1}$ for a quantity of interest $y_{t+1}$, such that 
@@
 \end{align}
-%
+% %>%
 
@@
 
-\subsection{Dealing with negative counts and small values}
-Unfortunately, the log transformation cannot readily be applied when dealing with negative or zero values.
-In an epidemiological setting that involves count data, negative values should likely be omitted entirely.
-Values of zero should potentially also be removed from the analsys, as it does not make sense to interpret relative errors on zero counts.
-Alternatively, if one is not willing to omit them, a small quantity (such as 1) could be added to all observations before taking the logarithm.
-Note that predictive distribtions have to be truncated in a similar way.
-When dealing with count data we could extend this argument to all small values (rather than just zeros), as it is hard to reliably evaluate relative errors when dealing with small discrete values.
-One could therefore decide to remove forecasts for small quantities below a certain threshold from the analysis.
-Alternatively, one could add a number $n$ (e.g.
-$n = 100$) to all observations and forecasts before taking the logarithm.
-This also constitutes a strictly monotonic transformation which preserves propriety and would reduce all scores (through reducing relative errors).
-Scores for forecasts on small quantities would be particularly affected, resulting in a more equal distribution of scores for forecasts of differing orders of magntitudes.
-Researchers can choose to add either a small or large number, depending on how much they want to equalise relative errors when aggregating scores.
+\subsection{Effects on model rankings}
+Rankings between different forecasts may be affected by applying the log transformation even when looking at a single forecast.
+As an example, consider two forecasters, A and B.
+Both issue a single 50\%-prediction interval for a target which later resolves as 1.
+Forecaster A issues the interval [0.5, 2], while forecaster B issues [0.9.
+2.5].
+On the natural scale, the score would equal the dispersion of the forecast, i.e.
+the length of this interval (since the observed value falls within the prediction interval).
+Forecaster A would receive a score of $2 - 0.5$ = 1.5, beating forecaster B who would receive a score of $2.5 - 0.9 = 1.6$.
+When we transform the forecast, forecaster A receives a score of $\log (\frac{2}{0.5}) = 1.39$, while forecaster B obtains a score of $\log (\frac{2.6}{0.9} = 1.06$ and now wins over forecaster A.
+The main difference between forecaster A and B is that forecaster A has issued a prediction interval ([0.5, 2]) that was more symmetric around the true value than the interval issued by forecaster B ([0.9.
+2.6], which was more skewed).
+This means that in relative terms, Forecaster B was noticeably closer to the true value with the lower bound of her prediction interval, while only being slightly further away with the upper interval.
+This is likely to lead to a more lenient treatment of (one-sided) outlier forecasts.
 
-\subsection{Relative importance when summarising scores}
-While the main rationale for applying the log transformation is that this better reflects the types of errors common in epidemiological modelling, it may also be helpful when scores based on the natural scale are very unequal.
-This usually happens, for example, when different forecast targets, like reported case numbers, hospitalisations and deaths, or the same targets across time and space, have very different orders of magnitude.
-In these instances, scores (even for an ideal forecaster) usually also differ by several orders of magnitude.
-Evaluating forecasts based on relative errors often makes it easier to compare performance across different targets more easily (as scores of an ideal forecaster would not differ as much as on the natural scale).
-
+Overall model rankings usually differ even more when performance is averaged across multiple forecasts.
+This effect is not mainly driven by the kinds of changes in model rankings for single forecasts we just discussed.
+Rather, aggregate scores across multiple forecasts are affected by the fact some scores will in general be much larger than others.
 On the natural scale, aggregate scores will usually be dominated by predictions of quantities with high absolute values.
-The opposite may be true for scores based on log transformed predictions.
-Small absolute errors (e.g.
+Even an ideal forecaster will make much larger errors when predicting a quantity such as reported case numbers, than when predicting hospitalisations.
+Evaluating forecasts based on relative errors may help with comparing performance across different targets.
+However, log transforming forecasts and observations may also result in the opposite extreme, since small absolute errors (e.g.
 predicting 9, rather than 3 deaths) may be large in relative terms if the quantity to forecast is very small.
-The strength of this effect depends on the exact relationship between the mean and the variance of the quantity of interest.
+This is especially true when forecasts are restricted to discrete values.
+Whether or not forecast targets with smaller quantities receive higher scores than those for targets with larger quantities, depends on the exact relationship between the mean and the variance of the quantity of interest.
 For simulated negative binomial count data, log predictions for small quantities receive on average higher scores than forecasts for large quantities if the mean and the variance grow at the same rate $\sigma^2 = \mu$, about equal scores when the variance grows at a rate of $\sigma^2 = \mu + \mu^2$ (Figure \ref{fig:SIM-wis-state-size-mean}), and smaller scores when the variance grows faster than that.
-%As mentioned above, one way to handle situations where scores are very unequal, is to add a number $n$ to all forecasts and observations in order to reduce the relative weight of forecast targets with small quantities.
-
 % To illustrate this with count data we sampled forecasts from different negative binomial distributions.
@@
 
-COULD ALSO BE DISCUSSION
-Whether or not a log transformation is apprpriate depends on whether one believes that differences in absolute values are meaningful.
-This may be the case for example when comparing forecasts across locations or time points, rather than different forecasts targets.
-\cite{bracherEvaluatingEpidemicForecasts2021} for example argue that in epidemiological settings it is a desirable property of the CRPS / WIS that it assigns large scores to forecasts where counts are high, as this naturally reflects the situations we should most care about and helps us identify the models that perform best in similar situations.
-If, on the other hand, we believe that good models should do consistently well and that situations with high incidences do not provide significantly more information about which models perform well or badly, then scoring on the log scale may be more appropriate.
-For example, it may be reasonable not to think that predictive performance in a country like Germany would have a hundred times more importance than performance in a country like Luxembourg when determining the best model for future decision making.
-Similarly, we may not necessarily believe that predictive performance during the January 2022 wave of COVID-19 should carry three times as much weight than predictive performance during the January 2021 wave, just because numbers where three times as high.
-% Maybe make a plot with one country and scores for that one country in different waves.
-We could then argue that within one wave, performance of the peak is meaningful, but across different waves (with different numbers) it is unclear, whether that difference conveys meaningful information.
+If one finds that average scores are dominated by forecasts for small quantities one could omit these forecasts entirely, arguing that relative errors are difficult to meaningfully interpret for small quantities.
+One might also take a more experimental route, namely adding a number $n$ (e.g.
+$n = 100$) to all observations and forecasts before taking the logarithm.
+This also constitutes a strictly monotonic transformation which preserves propriety and would reduce all scores (through reducing relative errors).
+Scores for forecasts on small quantities would be particularly affected, resulting in a more equal distribution of scores for forecasts of differing orders of magntitudes.
+Researchers could choose to add either a small or large number, depending on how much they want to equalise relative errors when aggregating scores.
+I KNOW YOU ALL HATE THIS, BUT IT IS SOOOO CLEVER! :-(
 
@@
 % could make an analysis re outliers: how do scores change if we remove the 1 or 2 worst forecasts? The one or two best forecasts?
-
-%THIS COULD INCLUDE JOHANNES EXAMPLE
 
@@
 
+PUT SOMEWHERE ELSE
 Conveniently, the decomposition of the WIS into dispersion, overprediction and underprediction is preserved.
@@
 
+\paragraph{Other transformations - maybe move to Discussion}
+While we focus on the log transformation in this paper, there are other transformations that may be helpful in an epidemiological context.
+One could for example divide all forecasts by the population in a location, to obtain forecasts standardised incidences per e.g.
+100,000 inhabitants.
+Or one could apply a Box-Cox transformation MISSING. 
+Or a square-root transformation MISSING. 
+Or one could divide each forecast by the forecast of the previous week (and observations by the observation of the previous week), to transform preditions to week-to-week growth rates.
+This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory.
+Note that this is not well defined in the context of forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate.
+
+COULD ALSO BE DISCUSSION
+Whether or not a log transformation is apprpriate depends on whether one believes that differences in absolute values are meaningful.
+This may be the case for example when comparing forecasts across locations or time points, rather than different forecasts targets.
+\cite{bracherEvaluatingEpidemicForecasts2021} for example argue that in epidemiological settings it is a desirable property of the CRPS / WIS that it assigns large scores to forecasts where counts are high, as this naturally reflects the situations we should most care about and helps us identify the models that perform best in similar situations.
+If, on the other hand, we believe that good models should do consistently well and that situations with high incidences do not provide significantly more information about which models perform well or badly, then scoring on the log scale may be more appropriate.
+For example, it may be reasonable not to think that predictive performance in a country like Germany would have a hundred times more importance than performance in a country like Luxembourg when determining the best model for future decision making.
+Similarly, we may not necessarily believe that predictive performance during the January 2022 wave of COVID-19 should carry three times as much weight than predictive performance during the January 2021 wave, just because numbers where three times as high.
+
+% Divide by the latest known value = score forecast on the growth rate
+% - WHAT KINDS OF DATA CAN I CONDITION ON? 
+% - WOULD DIVIDING A FORECAST BY ITS RESOLUTION WORK? 
+% - CAN I SCORE THE TRAJECTORY / WEEK-TO-WEEK GROWTH RATE BY DIVIDING the 4 week ahead forecast by the three week ahead forecast?
+
 \newpage
@@
     \includegraphics[width=0.9\textwidth]{output/placeholder-image.png}
-    \caption{Illustration of the fact that taking the log of WIS or CRPS results in an improper score}
-    \label{fig:log-improper}
+    \caption{Placeholder}
+    \label{fig:placeholder}
 \end{figure}

### 2022-07-21 analysis/Analysis ECDC forecasts.Rmd
update manuscript

@@
 # average number of cases and deaths
-hub_data |>
-  dplyr::select(location, target_type, target_end_date, true_value) |>
-  unique() |>
-  group_by(target_type) |>
-  summarise(mean_value = mean(true_value), 
-            sd_value = sd(true_value), 
-            var_value = var(true_value))
 
@@
 
-ggsave("output/figures/HUB-mean-obs-location.png", width = 9, height = 5.5)
+ggsave("output/figures/HUB-mean-obs-location.png", width = 9, height = 7)
 
@@
 ```{r}
-
 rankings <- scores |>
@@
 
-rankings_diff <- rankings |>
+rankings_wide <- rankings |>
   select(model, location, target_end_date, forecast_date, horizon,
@@
   rankings, 
-  rankings_diff,
+  rankings_wide,
   target_type
@@
   ) {
-    p <- rankings_diff |>
+    p <- rankings_wide |>
       mutate(diff_ranking = natural - log) |>
@@
 # should probably have one plot for both cases and deaths since this is rankings
-p_diff_cases <- plot_ranking_diffs(rankings, rankings_diff, 
+p_diff_cases <- plot_ranking_diffs(rankings, rankings_wide, 
                                    target_type = "Cases")
@@
        width = 9, height = 5)
-
-rankings_diff |>
-  group_by(horizon, target_type) |>
-  summarise(cor = cor(natural, log))
 ```
@@
 summarised_pairwise <- scores |>
-  filter(horizon == 2, 
-         target_type == "Cases") |>
+  filter(target_type == "Cases") |>
   summarise_scores(by = c("model", "scale"), relative_skill = TRUE)
@@
 They should usually be excluded, but maybe not necessary here as we're really more comparing the results for natural and log rather than the models against each other.
+```
 
-cor(diffs$log, diffs$natural)
+# Correlations for ranks, scores and pairwise comparisons
 
+```{r}
+# correlation between values
+hub_data |>
+  select(location, target_end_date, target_type, true_value) |>
+  unique() |>
+  mutate(log_obs = log(pmax(true_value, 0) + 1)) |>
+  group_by(target_type) |>
+  summarise(cor = cor(true_value, log_obs))
+
+# correlation between scores
+cor_scores <- scores |>
+  filter(horizon <=4) |>
+  select(interval_score, scale, forecast_date, model, horizon, location, target_type) |>
+  pivot_wider(names_from = scale, values_from = interval_score) |>
+  group_by(horizon, target_type) |>
+  summarise(cor = cor(natural, log), 
+            spearman = cor(natural, log))
+
+cor_scores |> 
+  mutate(`Target type` = target_type) |>
+  ggplot(aes(x = horizon, y = cor, 
+             fill = `Target type`, color = `Target type`)) + 
+  geom_histogram(stat = "identity", position = "dodge") +
+  theme_scoringutils() + 
+  coord_cartesian(ylim = c(0, 1)) +
+  labs(y = "Correlation (scores natural vs. log)", x = "Forecast horizon")
+
+# Correlation between ranks when going from natural to log
+correlation_ranks <- rankings_wide |>
+  filter(horizon < 5) |>
+  group_by(horizon, target_type) |>
+  summarise(cor = cor(natural, log)) |>
+  rename(`Target type` = target_type)
+
+plot_cor_ranks <- correlation_ranks |>
+  ggplot(aes(x = horizon, y = cor, 
+             fill = `Target type`, color = `Target type`)) + 
+  geom_histogram(stat = "identity", position = "dodge") +
+  theme_scoringutils() + 
+  coord_cartesian(ylim = c(0, 1)) +
+  labs(y = "Correlation (rank natural vs. log)", x = "Forecast horizon")
+
+ggsave(filename = "output/figures/HUB-cor-ranks.png", width = 7, height = 4.5, 
+       plot = plot_cor_ranks)
+
+# Correlation between pairwise comparisons
+correlation_rel_skill <- scores |>
+  filter(horizon < 5) |>
+  summarise_scores(
+    by = c("model", "scale", "horizon", "target_type"), 
+    relative_skill = TRUE) |>
+  select(model, horizon, target_type, relative_skill, scale) |>
+  pivot_wider(values_from = relative_skill, names_from = scale) |>
+  group_by(horizon, target_type) |>
+  summarise(cor = cor(natural, log))
+
+plot_cor_skill <- correlation_rel_skill |>
+  rename(`Target type` = target_type) |>
+  ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
+  # geom_line() +
+  # geom_point() +
+  geom_histogram(stat = "identity", position = "dodge") +
+  theme_scoringutils() + 
+  coord_cartesian(ylim = c(0, 1)) +
+  labs(y = "Correlation (rel. skill natural vs. log)", x = "Forecast horizon")
+
+ggsave(filename = "output/figures/HUB-cor-rel-skill.png", width = 7, height = 4.5, 
+       plot = plot_cor_skill)
+
+plot_cors <- plot_cor_ranks + plot_cor_skill +
+  plot_layout(guides = "collect") & 
+  theme(legend.position = "bottom")
+
+ggsave(filename = "output/figures/HUB-cors-skill-rank.png", width = 7, height = 3, 
+       plot = plot_cors)
+
+```
+
+Example comparisons of forecasts on the log and the natural scale
+
+```{r}
 # "CovidMetrics-epiBATS" only has forecasts in DE
@@
 # "UVA is everywhere
-hub_data |>
-  filter(model == "UVA-Ensemble") |> check_forecasts()
 
@@
 
-ggsave(filename = "output/figures/HUB-model-ICM.png", width = 7, height = 4.5, plot = ICM)
-ggsave(filename = "output/figures/HUB-model-epiBATS.png", width = 7, height = 4.5, plot = epiBATS)
-ggsave(filename = "output/figures/HUB-model-BSLCoV.png", width = 7, height = 4.5, plot = BSLCoV)
+ggsave(filename = "output/figures/HUB-model-ICM.png", width = 9, height = 5, plot = ICM)
+ggsave(filename = "output/figures/HUB-model-epiBATS.png", width = 9, height = 5, plot = epiBATS)
+ggsave(filename = "output/figures/HUB-model-BSLCoV.png", width = 9, height = 5, plot = BSLCoV)
+
+```
+
+# Table with relevant values
+
+```{r}
+summary <- hub_data |>
+  filter(horizon == 2) |>
+  mutate(natural = true_value, 
+         log = log(pmax(0, true_value) +1)) |>
+  select(location, target_type, target_end_date, natural, log) |>
+  pivot_longer(cols = c(natural, log), values_to = "true_value", names_to = "scale") |>
+  unique() |>
+  group_by(target_type, scale) |>
+  summarise(mean = mean(true_value), 
+            sd = sd(true_value), 
+            var = var(true_value))
+
+table_obs <- summary |>
+  pivot_longer(cols = c(mean, sd, var), names_to = "measure") |>
+  pivot_wider(values_from = value, names_from = scale) |>
+  select(target_type, measure, natural, log) |>
+  mutate(quantity = "Observations", 
+         natural = round(natural), 
+         log = round(log, 2)) |>
+  select(target_type, quantity, measure, natural, log)
+
+table_scores <- scores |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  group_by(target_type, scale) |>
+  summarise(mean = mean(interval_score), 
+            sd = sd(interval_score), 
+            .groups = "drop_last") |>
+  pivot_longer(cols = c(mean, sd), names_to = "measure") |>
+  pivot_wider(values_from = value, names_from = scale) |>
+ mutate(quantity = "WIS", 
+         natural = round(natural), 
+         log = round(log, 2)) |>
+  select(target_type, quantity, measure, natural, log)
+
+table <- rbind(table_obs, 
+               table_scores) |>
+  arrange(quantity, target_type, measure)
+
+table |>
+  kable(format = "latex", booktabs = TRUE,
+        align = c("l", "l", "l", rep("c", 2)),
+        # col.names = c(" ", names(scores)[-(1:2)])) %>%
+  ) #|>
+  # collapse_rows(columns = 1) %>%
+  # kable_styling(latex_options = c("scale_down",
+  #                                 "hold_position")) %>%
+  # column_spec(1, background = "white") %>%
+  # pack_rows("Observations", 1, 6, indent = FALSE,
+  #           hline_after = TRUE) %>%
+  # pack_rows("WIS", 7, 10, indent = FALSE,
+  #           hline_after = TRUE)
 
@@
 
-## Variance of scores across locations
+## Variance of scores vs. mean of scores across locations
 

### 2022-07-21 main.tex
update manuscript

@@
 The WIS can be thought of as an approxmiation of the CRPS for forecasts in a quantile format and has very similar properties (the WIS converges towards the CRPS for an increasing set of equally-spaced prediction intervals).
-Due to its widespread use in epidemiology (and its similarity to the CRPS) we focus on the WIS in this paper, ignoring other proper scoring rules.
+Due to its widespread use in epidemiology (and its similarity to the CRPS) we mostly focus on the WIS in this paper, ignoring other proper scoring rules.
+Formulas for the WIS are given in the SI.
 
@@
                     &= \log (1 + g_{t, t+1}) - \log (1 + \hat{g}_{t, t+1}) \Leftrightarrow \\   
-                    &\approx g_{t, t+1} - \hat{g}_{t, t+1} \quad \text{for small differences between} \hat{g}_{t, t+1} and g_{t, t+1}. 
+                    &\approx g_{t, t+1} - \hat{g}_{t, t+1} \quad \text{for small differences between } \hat{g}_{t, t+1} \text{ and } g_{t, t+1}. 
 \end{align}
@@
 This approximation breaks down when the difference between the predicted and the actual growth rate becomes too large (albeit forecasts are still evaluated in terms of relative errors).
-If this is of great concern one may want to use a different transformation like dividing forecasts and observations by the last known values instead in order to score the growth rate explicitly.
+If this is of great concern one may want to use a different transformation like dividing forecasts and observations by the last known values instead in order to score the growth rate explicitly (although this lacks the elegance of having one single transformation per time point).
+
+Conveniently, the decomposition of the WIS into dispersion, overprediction and underprediction is preserved.
+The interpretation of the components has now changed in that they now (approximately) represent uncertainty, over-prediction, and under-prediction with the growth rate and so are independent of current incidence.
 
@@
 Whether or not forecast targets with smaller quantities receive higher scores than those for targets with larger quantities, depends on the exact relationship between the mean and the variance of the quantity of interest.
+
+MAYBE MOVE TO SI
 For simulated negative binomial count data, log predictions for small quantities receive on average higher scores than forecasts for large quantities if the mean and the variance grow at the same rate $\sigma^2 = \mu$, about equal scores when the variance grows at a rate of $\sigma^2 = \mu + \mu^2$ (Figure \ref{fig:SIM-wis-state-size-mean}), and smaller scores when the variance grows faster than that.
@@
 
+I KNOW YOU ALL ARE GOING TO HATE THIS, BUT IT IS SOOOO CLEVER! :-(
 If one finds that average scores are dominated by forecasts for small quantities one could omit these forecasts entirely, arguing that relative errors are difficult to meaningfully interpret for small quantities.
@@
 Researchers could choose to add either a small or large number, depending on how much they want to equalise relative errors when aggregating scores.
-I KNOW YOU ALL HATE THIS, BUT IT IS SOOOO CLEVER! :-(
-
-%COULD ALSO MAKE A NICE PLOT ABOUT OUTLIERS, where we look at the effect of outliers. but maybe that is already included in the plot in Figure 1. 
-% could make an analysis re outliers: how do scores change if we remove the 1 or 2 worst forecasts? The one or two best forecasts?
 
@@
 
-PUT SOMEWHERE ELSE
-Conveniently, the decomposition of the WIS into dispersion, overprediction and underprediction is preserved.
-The interpretation of the components has now changed in that they now (approximately) represent uncertainty, over-prediction, and under-prediction with the growth rate and so are independent of current incidence.
-
 %Score different things from the European / US Forecast Hub and plot the relationship between mean and variance.
@@
 
-\paragraph{Introduction to the Forecast Hub} As an empirical example for evaluating forecasts on the natural and on the log scale we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
+As an empirical example for evaluating forecasts on the natural and on the log scale we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
 Every week the European Forecast Hub collates and aggregates forecasts for different COVID-19 related targets from teams around the world.
@@
 The forecasts for the purpose of this illustrations are two-week-ahead forecasts made between XXX and XXX for reported cases and deaths from COVID-19 in XXX countries.
+Where not otherwise stated, we report results for a two-week-ahead forecast horizon.
 
-\paragraph{Describing features of the data}
-Across all time points and locations, the mean number of cases (deaths) observed was 22811 (368) with an overall standard deviation of 47135 (831).
-The number of observed value does not only differ by forecast target, but also shows considerable heterogeneity by location (see Figure \ref{fig:HUB-mean-locations}.
-Across the different locations, the variance in observed values is roughly equal to the square of the mean, suggesting that, for a given location, both cases and deaths are not poisson-distributed, but rather exhibit over-dispersion.
-% MAYBE FIT A NEGATIVE BINOMIAL TO THESE OBSERVATIONS
+Pooled across all time points and locations, the mean number of cases (deaths) observed was 22811 (368) with an overall standard deviation of 47135 (831) (see Table \ref{tab:HUB-summary}).
+The average number of observed cases and deaths varies considerably by location (see Figure \ref{fig:HUB-mean-locations}A and B).
+Observations are overdispersed, meaning that for a given location, the variance in observations exceeds the mean of all observed values (see Figure \ref{fig:HUB-variance-mean}).
+%maybe make a line of best fit of the form variance = alpha * mean^2 + beta * mean
+
+Pooled across all time points and locations, the mean WIS on the natural scale for cases (deaths) was 8855 (59.3) with an overall standard deviation of 51274 (156) (see Table \ref{tab:HUB-summary}).
+On the log scale, scores were much closer to each other (see Figure \ref{fig:HUB-average-scores}), with a mean WIS for cases (deaths) of 0.507 (0.362) with an overall standard deviation of 0.745 (0.407).
+
+Variation of scores across targets (see Figure XX) and locations (see Figure \ref{fig:HUB-mean-locations}A, B) is much larger on the natural than on the log scale, where scores are more evenly distributed.
+On the natural scale, average scores per location correlate strongly (XX) with the average number of observed cases or deaths (See Figure XX).
+On the log scale, this correlation is much weaker and slightly negative.
+Forecasts across different locations receive more similar scores, with a tendency for scores to be slightly higher in locations with few observed cases and deaths (see Figures XX and \ref{fig:HUB-mean-locations}.
+%This is in line with the observation that the association between variance is smaller than mu^2 + mu, meaning that we would expect this. 
+For individual forecasts, rankings between models are mostly preserved (correlation: X for cases, Y for deaths), with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors-scores}).
+
+The correlation between scores on the log and natural scale for different locations is XX for cases and XX for deaths. 
+The correlation between the actual observations is 0.628 for cases and 0.631 for deaths. IS THAT HELPFUL?
+
+\begin{table}[h!]]
+    \centering
+
+    \begin{tabular}{lllcc}
+    \toprule
+    target\_type & quantity & measure & natural & log\\
+    \midrule
+    Cases & Observations & mean & 20308 & 8.39\\
+    Cases & Observations & sd & 43405 & 2.11\\
+    Cases & Observations & var & 1884002308 & 4.46\\
+    Deaths & Observations & mean & 259 & 3.79\\
+    Deaths & Observations & sd & 532 & 2.11\\
+    \addlinespace
+    Deaths & Observations & var & 283492 & 4.46\\
+    Cases & WIS & mean & 4544 & 0.28\\
+    Cases & WIS & sd & 17585 & 0.49\\
+    Deaths & WIS & mean & 28 & 0.22\\
+    Deaths & WIS & sd & 67 & 0.26\\
+    \bottomrule
+    \end{tabular}
+
+    \caption{CAPTION}
+    \label{tab:HUB-summary}
+\end{table}
 
@@
     \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-obs-location.png}
-    \caption{CAPTION}
+    \caption{Scores and average observed values across 1ocations.
+A, B: Mean of observed cases and deaths across all time points for different locations.
+C, D: XXX WIS for forecasts made by the Hub ensemble.
+D, E: Average WIS for log-transformed forecasts.
+}
     \label{fig:HUB-mean-locations}
@@
     \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-var-obs.png}
-    \caption{Observed mean against observed variance of observations.
+    \caption{Observed mean against observed variance of observations in different locations across time.
 The dashed line marks what would be expecetd from a negative binomially distributed variable with variance of the observed value equal to the mean + the square of mean observed value.
@@
 
-\paragraph{Features of the scores}
-Across all time points and locations, the mean number weighted interval score on the natural scale for cases (deaths) was 8855 (59.3) with an overall standard deviation of 51274 (156).
-On the log scale, scores were much closer to each other (see Figure \ref{fig:HUB-average-scores}), with a mean WIS for cases (deaths) of 0.507 (0.362) with an overall standard deviation of 0.745 (0.407).
-% NEED A TABLE WITH SCORES for the Appendix
-On the natural scale, scores vary by several orders of magnitudes across locations (see \ref{fig:HUB-scores-location}), whereas variation across locations is much smaller for scores on the log scale and wis values are more evenly distributed.
-The ordering of scores across locations is not preserved when going from the natural to the log scale (see Figure \ref{fig:HUB-mean-location}.
-Rather, there seems to be a slight inverse relationship that locations which previously had low scores now tend to receive higher scores.
-The correlation between scores on the log and natural scale for different locations is XX for cases and XX for deaths. 
-BUT: CORRELATION FOR RANKINGS IS STILL LAGRGE
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-cors-skill-rank.png}
+    \caption{Correlation between scores on the natural and on the log scale.
+A: Correlation between relative ranks (the rank in terms of WIS divided by the number of models) for individual forecasts, separated by forecast horizon.
+B: Correlation between relative skill values achieved by models across locations and time points, separated by forecast horizon.}
+    \label{fig:HUB-cors-scores}
+\end{figure}
 
@@
 \paragraph{Rankings}
-Rankings do change a little bit, but not massively. More difference comes when we look at greater forecast horizons.
-Really this is more a point about how rankings change even for one single forecast (which I originally thought they shouldn't).
-Would be good to tease out what happens here.
-Somewhow we do change our judegement on which model is better.
-Maybe the preservation of the order really only holds for one interval?
-Or maybe not even that?
-Example: True value = 1
-Forecaster 1: 0.8, 3.     Natural scale: 3.2 - 0.8 = 2.4.     Log scale: log 3 - log 0.8 = 1.386
-Forecaster 2: 0.5, 2.8    Natural scale: 2.8 - 0.5 = 2.3    Log scale: log (2.8/0.5) = 1.722767
 So one effect is that for a given dispersion on the natural scale, it's better if that interval is skewed, as that implies that at least you only off a lot in one direction in negative terms, and not in the other direction.
@@
 % \end{itemize}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-model-ICM.png}
+    \caption{Top: 2-week-ahead predictions. Bottom: scores }
+    \label{fig:HUB-model-ICM}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-model-epiBATS.png}
+    \caption{Top: 2-week-ahead predictions. Bottom: scores }
+    \label{fig:HUB-model-epiBATS}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-model-BSLCoV.png}
+    \caption{Top: 2-week-ahead predictions. Bottom: scores }
+    \label{fig:HUB-model-BSLCoV}
+\end{figure}
 
@@
 % - CAN I SCORE THE TRAJECTORY / WEEK-TO-WEEK GROWTH RATE BY DIVIDING the 4 week ahead forecast by the three week ahead forecast?
+
+ %COULD ALSO MAKE A NICE PLOT ABOUT OUTLIERS, where we look at the effect of outliers. but maybe that is already included in the plot in Figure 1. 
+% could make an analysis re outliers: how do scores change if we remove the 1 or 2 worst forecasts? The one or two best forecasts?
 

### 2022-07-21 main.tex
Updates from Overleaf

--- 2022-07-15 main.tex
@@
 \linenumbers
-\bibliographystyle{apalike}
 
@@
 \newcommand{\changed}[1]{#1}
-
-% to write the simulated as iid symbol: 
-\newcommand{\distas}[1]{\mathbin{\overset{#1}{\kern\z@\sim}}}%
 
@@
 
-Probabilistic forecasts play an important role in decision-making in epidemiology and public health \citep{reichCollaborativeMultiyearMultimodel2019, funkShorttermForecastsInform2020, cramerEvaluationIndividualEnsemble2021, bracherShorttermForecastingCOVID192021, europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}, as well as other areas as diverse as for example economics \citep{timmermannForecastingMethodsFinance2018, elliottForecastingEconomicsFinance2016} or meteorology \citep{gneitingWeatherForecastingEnsemble2005, kukkonenReviewOperationalRegionalscale2012}.
+Probabilistic forecasts play an important role in decision-making in epidemiology and public health [Citations], as well as other areas as diverse as for example economics [Citations] or agriculture [Citations].
 Epidemiological modelling in particular has received widespread attention during the COVID-19 pandemic and has been an important tool to inform public policy.
-Evaluating forecasts is important, as the evaluation provides feedback for forecasters and researchers necessary to improve their models, and it helps decision makers distinguish good from bad predictions and choosing forecasts that inform future decisions.
+Evaluating forecasts is important, as the evaluation provides feedback for forecasters and researchers necessary to improve their models, and it helps decision makers with distinguishing good from bad predictions and choosing forecasts that inform future decisions.
 
@@
 Forecasters, in expectaction, optimise their score by providing a predictive distribution that is equal to the data-generating distribution, and will be penalised if they fail to do so.
-Two proper scoring rules commonly used in epidemiology and other fields are the Continuous Ranked Probability Score (CRPS) \citep{gneitingStrictlyProperScoring2007} and the Weighted Interval Score (WIS) \citep{bracherEvaluatingEpidemicForecasts2021}.
-Due to their widespread use in epidemiology we focus on these two in this paper, ignoring other proper scoring rules.
+Two proper scoring rules commonly used in epidemiology and other fields are the Continuous Ranked Probability Score (CRPS) [Gneiting 2007] and the Weighted Interval Score (WIS) [CITATION].
+Due to their widespread use we focus on these two in this paper, ignoring other proper scoring rules.
 Both the WIS and CRPS represent a measure of the absolute distance between the predictive distribution and observation and can be thought of as a generalisation of the absolute error to full predictive distributions (see details in the SI).
 
-Evaluating forecasts based on absolute errors (for simplicity we use the term interchangeably with absolute distance, even though the former technically only applies to point forecasts) may not always be desirable, especially in an epidemiological context.
-% The reason for this is that generally, the way for ecasts are evaluated should correspond to the error structure of the model and the type of errors models will usually make.
+Evaluating forecasts based on absolute errors (for simplicity we use the term interchangeably with absolute distance, even though it technically only applies to point forecasts) may not always be desirable, especially in an epidemiological context.
+% The reason for this is that generally, the way forecasts are evaluated should correspond to the error structure of the model and the type of errors models will usually make.
 Epidemiological processes are usually understood to be exponential in nature and therefore are most commonly described and modeled in multiplicative terms.
@@
 
-In this paper we argue that it may therefore be helpful to shift the focus of the evaluation on multiplicative, rather than absolute errors.
-This is approximately equal to scoring the growth rate of a forecast and can be achieved by transforming the forecasts as well as the observations using the logarithm before applying the WIS or CRPS.
-We focus on the log transformation, as it represents a very convenient way of achieving our goal of evaluating how well a forecast captures the growth dynamic inherent to epidemiological process.
-However, we note that the same logic we use can also applied to other transformations which could shift the focus of evaluation in different ways.
-For example, given an appropriate transformation, it could also be possible to evaluate forecasters based on the shape of the trajectory they predicted.
-It would even be possible to construct composite scores as a weighted sum of scores based on different transformations, allowing forecast consumers to assign explicit weights to different qualities of the forecasts.
+In this paper we argue that it may therefore often be helpful to shift the focus of the evaluation on multiplicative, rather than absolute errors.
+This can be achieved by transforming the forecasts as well as the observations using the logarithm before applying the WIS or CRPS.
+While we will mostly focus on the log transformation in this paper, the point we like to make is a broader one: it is possible to flexible change the focus of the evaluation by applying a multitude of different transformations before scoring forecasts.
+Instead of scoring multiplicative errors, one may for example like to apply a transformation that allows to score forecasters based on the shape of the trajectory they predicted.
+Furthermore, it is possible to capture different aspects in a single number by constructing composite scores as a weighted sum of scores based on different transformations.
+This allows forecast consumers to assign explicit weights to different qualities of the forecasts they deem to be important when assessing predictive performance.
 
-In section [REF] we discuss the broader idea of transforming forecasts, and in section [REF] give a detailed theoretical of the log transformation and its interpretation.
-In section [REF] we present an applied example using forecasts submitted to the European COVID-19 Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
-The European COVID-19 Forecast Hub is one of several Forecast hubs (together with a Hub in the US \citep{cramerEvaluationIndividualEnsemble2021} and one in Germany and Poland \citep{bracherShorttermForecastingCOVID192021} that systematically collected, aggregated and evaluated forecasts of several COVID-19 targets created by different teams.
+In section [REF] we discuss which kinds of transformations can be applied.
+In section [REF] we give a detailed theoretical explanation of the rationale for applying the log transformation and discuss its interpretation.
+In section [REF] we present an applied example using forecasts submitted to the European COVID-19 Forecast Hub.
+The European COVID-19 Forecast Hub is one of several Forecast hubs (together with a Hub in the US [CITATION] and one in Germany and Poland [CITATION]) that systematically collected, aggregated and evaluated forecasts of several COVID-19 targets created by different teams.
 Forecasts submitted to the COVID-19 Forecast Hubs are probabilistic in nature [Citation e.g.
@@
 
-Applying the log transformation, among other transformations, is permissible because it is a strictly monotonic transformation.
-While rankings between different models may indeed change, even for a single forecast, the relevant condition for a scoring rule to be strictly proper is still satisfied: In expectation a forecaster will receive the best score if she reports a predictive distribution that is equal to the data-generating distribution, even after transforming both the forecasts and the observations (representing draws from the data-generating distribution).
+\subsection{Permissible transformations / propriety} - JOHANNES
 
-It is important to first apply the transformation and then score the transformed forecasts and observations.
-Simply taking the log of the WIS or the CRPS computed on the natural scale, i.e.
+In order to preserve the propriety of the score, transformations must be strictly monotonic.
+This makes sure that a forecast that is further away from the observed value will also be further away after the transformation.
+
+- WHAT KINDS OF DATA CAN I CONDITION ON? 
+- WOULD DIVIDING A FORECAST BY ITS RESOLUTION WORK? 
+- CAN I SCORE THE TRAJECTORY / WEEK-TO-WEEK GROWTH RATE BY DIVIDING the 4 week ahead forecast by the three week ahead forecast? 
+Probably does not work. 
+Problem: Median von (X\_t+4 / X\_t+3) ist anders als Median
+
+\paragraph{First transform, then score, rather than the other way round}
+Note that simply taking the log of a score computed on the natural scale, i.e.
 computing $\log S(x, y)$ results in an improper score and is therefore not advised.
-This is because the logarithm of the WIS / CRPS does not penalise outliers enough and therefore incentivises overconfident predictions.
-We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can easily be seen that overconfident models perform best in terms of the log WIS.
+We illustrate this point in Figure \ref{fig:log-improper} in the SI.
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/example-log-first.png}
-    \caption{Scores for different forecasts evaluated the WIS and the log WIS.
-We simulated 1000 observations $Y_i = {\rm e}^{x_i}$, with $x_i \text{iid} \sim \mathcal{N}(0, 1)$.
-We then simulated 20 forecasters who would issue a predictive distribution $F = {\rm e}^{x_i}$, with $x \sim \mathcal{N}(0, \sigma)$, with values of $\sigma$ ranging from 0.1 to 2.}
-    \label{fig:log-improper}
-\end{figure}
+% \paragraph{Changes only come when you average across multiple scores}
+% For a single forecast, the ranking between forecasters would be preserved and the scores remain proper [Proof?].
+Differences in the ranking between forecasters only appear when averaging over multiple forecasts, as deviations from the perfect forecast get penalised differently depending on whether forecasts are transformed or not.
 
-\paragraph{Other transformations - maybe move to Discussion}
-While we focus on the log transformation in this paper, there are other transformations that may be helpful in an epidemiological context.
-One could for example divide all forecasts by the population in a location, to obtain forecasts standardised incidences per e.g.
-100,000 inhabitants.
-Or one could apply a Box-Cox transformation MISSING. 
-Or a square-root transformation MISSING. 
-Or one could divide each forecast by the forecast of the previous week (and observations by the observation of the previous week), to transform preditions to week-to-week growth rates.
-This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory.
-Note that this is not well defined in the context of forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate.
+\paragraph{Example transformations}
 
-% Divide by the latest known value = score forecast on the growth rate
-% - WHAT KINDS OF DATA CAN I CONDITION ON? 
-% - WOULD DIVIDING A FORECAST BY ITS RESOLUTION WORK? 
-% - CAN I SCORE THE TRAJECTORY / WEEK-TO-WEEK GROWTH RATE BY DIVIDING the 4 week ahead forecast by the three week ahead forecast?
+Take the log
+
+Divide by the latest known value = score forecast on the growth rate
+
+Log that = score log growth rate
+
+Divide by the previous for each time value, which gives you a weekly growth rate, i.e. a trajectory.
+
+Box-Cox
+
+square root WAS THAT USED SOMEWHERE? This is stabilising the variance. Kraus und Kraus. Idea: if X is poisson with some mean, then sqrt(x) has asymptotically a constant variance
+
+Standardising by population size
 
@@
 
-% In this section (and those that follow) we focus on one particular transformation, taking the logarithm of forecasts and observations, which makes it possible to change the focus of the evaluation from absolute to relative errors.
+In this section (and those that follow) we focus on one particular transformation, taking the logarithm of forecasts and observations, which makes it possible to change the focus of the evaluation from absolute to relative errors.
 To illustrate the effects of applying the logarithm intuitively, let us consider a point forecast $\hat{y}_{t+1}$ for a quantity of interest $y_{t+1}$, such that
@@
 The analogon of the WIS / CRPS for point forecasts is the absolute error, and so the score will be determined by the size of the forecast error $\varepsilon_{t+1}$.
-When taking the logarithm of the forecast and the observation first, then the score is determined based on the size of $\varepsilon^*_{t+1}$:
+When taking the logarithm of the forecast and the observation first, than the score is determined based on the size of $\varepsilon^*_{t+1}$:
 \begin{equation}
@@
 Forecasting 101,000 rather than 100,000 cases will be treated equally to forecasting 2,020, rather than 2,000 hospitalisations.
-While the absolute error on the log scale is exactly symmetric (e.g., predicting 2 instead of 1 would give the same error as predicting 0.5), this is not entirely true for the WIS.
-Due to the fact that upper and lower bounds of individual prediction intervals are affected differently, there may be slight deviations (compare e.g.
-in Figure \ref{fig:change-in-scores}D the scores for a relative error of $\frac{1}{5}$ and 5).
+% Note that for a single forecast the relative rankings of different forecasts would be preserved. The transformation only makes a difference when aggregating predictions across different dimensions.
 
@@
                     &= \log (1 + g_{t, t+1}) - \log (1 + \hat{g}_{t, t+1}) \Leftrightarrow \\   
-                    &\approx g_{t, t+1} - \hat{g}_{t, t+1} \quad \text{for small differences between} \hat{g}_{t, t+1} and g_{t, t+1}. 
+                    &\approx g_{t, t+1} - \hat{g}_{t, t+1} \quad \text{for small } \hat{g}_{t, t+1}, g_{t, t+1}. 
 \end{align}
 
-This approximation breaks down when the difference between the predicted and the actual growth rate becomes too large (albeit forecasts are still evaluated in terms of relative errors).
-If this is of great concern one may want to use a different transformation like dividing forecasts and observations by the last known values instead in order to score the growth rate explicitly.
+%maybe it suffices that the difference between them is small. You can write them as a ratio and then the ratio just needs to be similar to one.
+
+This approximation breaks down when the growth rate becomes too large.
+If scoring the growth rate is the focus of the evaluation, then it may be better to transform forecasts and observations by the last known values instead in order to score the growth rate explicitly.
 
@@
 $n = 100$) to all observations and forecasts before taking the logarithm.
-This also constitutes a strictly monotonic transformation which preserves propriety and would reduce all scores (through reducing relative errors).
+This constitutes a strictly monotonic transformation which preserves propriety and would reduce all scores (through reducing relative errors).
 Scores for forecasts on small quantities would be particularly affected, resulting in a more equal distribution of scores for forecasts of differing orders of magntitudes.
@@
 \subsection{Relative importance when summarising scores}
+% Stationarity of forecast timelines
+
+For a single transformation, the ranking between models is preserved, even though the numeric value of the score changes.
+However, when averaging across forecasts, the relative influence of individual forecasts on the aggregate score may differ dramatically, leading to changes in the overall model rankings.
 While the main rationale for applying the log transformation is that this better reflects the types of errors common in epidemiological modelling, it may also be helpful when scores based on the natural scale are very unequal.
@@
 
-On the natural scale, aggregate scores will usually be dominated by predictions of quantities with high absolute values.
-The opposite may be true for scores based on log transformed predictions.
+While on the natural scale, aggregate scores will usually be dominated by predictions of quantities with high absolute values, the opposite may be true for scores based on log transformed predictions.
 Small absolute errors (e.g.
@@
 For simulated negative binomial count data, log predictions for small quantities receive on average higher scores than forecasts for large quantities if the mean and the variance grow at the same rate $\sigma^2 = \mu$, about equal scores when the variance grows at a rate of $\sigma^2 = \mu + \mu^2$ (Figure \ref{fig:SIM-wis-state-size-mean}), and smaller scores when the variance grows faster than that.
-%As mentioned above, one way to handle situations where scores are very unequal, is to add a number $n$ to all forecasts and observations in order to reduce the relative weight of forecast targets with small quantities.
+As mentioned above, one way to handle situations where scores are very unequal, is to add a number $n$ to all forecasts and observations in order to reduce the relative weight of forecast targets with small quantities.
 
@@
 This may be the case for example when comparing forecasts across locations or time points, rather than different forecasts targets.
-\cite{bracherEvaluatingEpidemicForecasts2021} for example argue that in epidemiological settings it is a desirable property of the CRPS / WIS that it assigns large scores to forecasts where counts are high, as this naturally reflects the situations we should most care about and helps us identify the models that perform best in similar situations.
+\citep{Bracher} for example argue that in epidemiological settings it is a desirable property of the CRPS / WIS that it assigns large scores to forecasts where counts are high, as this naturally reflects the situations we should most care about and helps us identify the models that perform best in similar situations.
 If, on the other hand, we believe that good models should do consistently well and that situations with high incidences do not provide significantly more information about which models perform well or badly, then scoring on the log scale may be more appropriate.
@@
 
+%COMPARISON TO SCORING THE ACTUAL GROWTH RATE - just harder to do and achieves the same thing? No objections if someone wants to? Might actually be better if there are negative values involved?
+
+%What about other things like scoring the square root of the forecast?
+
 %\begin{itemize}
+%    \item discuss how equations change for CRPS
 %    \item Optional: Discussion of parallels to point forecasts and the point that you're still incentivised to report the median\\
+%    \item discuss that even if both scoring rules are proper, they still penalise different things %differently. How does that change incentives for the forecaster? 
 %\end{itemize}
+
+\paragraph{Toy example: simulate an epidemic and apply 3 different models} Compare scores for these three models on the natural scale and the log scale
 
@@
 
-\paragraph{Introduction to the Forecast Hub} As an empirical example for evaluating forecasts on the natural and on the log scale we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
+\paragraph{Introduction to the Forecast Hub} As an empirical example for evaluating forecasts on the natural and on the log scale we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021}.
 Every week the European Forecast Hub collates and aggregates forecasts for different COVID-19 related targets from teams around the world.
@@
 Across the different locations, the variance in observed values is roughly equal to the square of the mean, suggesting that, for a given location, both cases and deaths are not poisson-distributed, but rather exhibit over-dispersion.
-% MAYBE FIT A NEGATIVE BINOMIAL TO THESE OBSERVATIONS
+MAYBE FIT A NEGATIVE BINOMIAL TO THESE OBSERVATIONS
 
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-var-obs.png}
-    \caption{Observed mean against observed variance of observations.
-The dashed line marks what would be expecetd from a negative binomially distributed variable with variance of the observed value equal to the mean + the square of mean observed value.
-}
-    \label{fig:HUB-variance-mean}
-\end{figure}
-
 \paragraph{Features of the scores}
@@
 On the natural scale, scores vary by several orders of magnitudes across locations (see \ref{fig:HUB-scores-location}), whereas variation across locations is much smaller for scores on the log scale and wis values are more evenly distributed.
-The ordering of scores across locations is not preserved when going from the natural to the log scale (see Figure \ref{fig:HUB-mean-location}.
-Rather, there seems to be a slight inverse relationship that locations which previously had low scores now tend to receive higher scores.
+The ordering of scores across locations is not preserved when going from the natural to the log scale.
+% Rather, there seems to be a slight inverse relationship that locations which previously had low scores now tend to receive higher scores. 
 The correlation between scores on the log and natural scale for different locations is XX for cases and XX for deaths. 
-BUT: CORRELATION FOR RANKINGS IS STILL LAGRGE
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-ranking-diffs.png}
-    \caption{Distribution of rankings of the different Hub models.
-Rankings are computed as relative rankings, i.e.
-the ranking divided by the total number of models available for a given forecast target.}
-    \label{fig:HUB-rankings}
-\end{figure}
+%THERE SEEMS TO BE A TENDENCY FOR SMALLER
 
@@
 \end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-scores-locations.png}
+    \caption{Weighted interval scores across different locations for two week ahead forecasts of COVID-19 cases and deaths evaluated on the natural scale (left) and on the log scale (right). }
+    \label{fig:HUB-scores-location}
+\end{figure}
+% Could also sort this figure according to population size?
+%Make multi panel Figure with Figure 2 and sort according to mean outcome
 
@@
 
-\paragraph{Rankings}
-Rankings do change a little bit, but not massively. More difference comes when we look at greater forecast horizons.
-Really this is more a point about how rankings change even for one single forecast (which I originally thought they shouldn't).
-Would be good to tease out what happens here.
-Somewhow we do change our judegement on which model is better.
-Maybe the preservation of the order really only holds for one interval?
-Or maybe not even that?
-Example: True value = 1
-Forecaster 1: 0.8, 3.     Natural scale: 3.2 - 0.8 = 2.4.     Log scale: log 3 - log 0.8 = 1.386
-Forecaster 2: 0.5, 2.8    Natural scale: 2.8 - 0.5 = 2.3    Log scale: log (2.8/0.5) = 1.722767
-So one effect is that for a given dispersion on the natural scale, it's better if that interval is skewed, as that implies that at least you only off a lot in one direction in negative terms, and not in the other direction.
-Maybe looking at over-and underprediction is the key here.
-%ADD SAME PLOT TO SI FOR HORIZON 4.
-
-\paragraph{pairwise comparisons}
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-pairwise-comparisons.png}
-    \caption{Changes in relative skill and the distribution of scores}
-    \label{fig:HUB-pairwise}
-\end{figure}
-
-Correlation between pairwise comparisons is 0.4414717.
-
 % \section{Other ideas}
@@
 \paragraph{summary of what we did}
-In this paper we suggested that epidemiological forecasts which are today evalaluated using the WIS or CRPS on a natural scale should instead be log transformed first before scoring.
-This shifts the focus of the evaluation from absolute to relative errors and is a convenient way to score forecasts based on how well they are able to predict the growth rate.
-This is useful as it makes sure that the evaluation is consistent with the kinds of errors we expect forecasting models to make in an epidemiological context.
-
-When applied to data from the Europeam Forecasting Hub, we found that
 
@@
 
-\bibliography{log-or-not.bib}
-
 \end{document}

### 2022-08-02 analysis/Analysis ECDC forecasts.Rmd
update manuscript

@@
 
+```{r}
+
+check <- hub_data |> 
+  check_forecasts()
+
+filtermodels <- check$unique_values |>
+  filter(horizon >= 4, 
+         target_type == 2,
+         true_value > 900,
+         quantile == 23, 
+         forecast_date > 16,
+         location == 32) |>
+  pull(model)
+
+scores <- scores[model %in% filtermodels]
+scores[, type_and_scale := paste0(target_type, " - ", scale)]
+scores[, type_and_scale := factor(type_and_scale, 
+                                  levels = c("Cases - natural", "Deaths - natural",
+                                             "Cases - log", "Deaths - log"))]
+
+hub_data <- hub_data[model %in% filtermodels]
+hub_data[, true_value := pmax(true_value, 0)]
+
+```
+
 General thoughts?
@@
 
-Figure that shows Variance vs. mean of the observations. Makes the point that the
+Figure 5? 
+that shows Variance vs. mean of the observations. Makes the point that the
 data looks neg. binomial. 
@@
 # plot with mean vs. variance --------------------------------------------------
-hub_data |>
+plot_var_mean <- function(hub_data, log = FALSE) {
+  df <- hub_data |>
     dplyr::select(location, target_type, target_end_date, true_value) |>
     unique() |>
-  group_by(location, target_type) |>
+    group_by(location, target_type)
+
+  f <- function(x) {x + x^2}
+
+  if (log) {
+    df <- mutate(df, true_value = log(pmax(true_value, 0) + 1))
+
+    f <- function(x) {log(x + x^2 + 1)}
+  }
+
+  df |>
   summarise(mean_value = mean(true_value), 
@@
   scale_y_continuous(trans = "log10", labels = label_fn) +
-  # scale_x_continuous(labels = label_fn) +
-  # scale_y_continuous(labels = label_fn) +
   labs(y = "Variance observed value", x = "Mean observed value") + 
-  geom_function(fun = function(x) {x + x^2}, linetype = "dashed", size = 0.5) + 
+  geom_function(fun = f, linetype = "dashed", size = 0.5) + 
   theme_scoringutils() + 
   geom_smooth(alpha = 0.2, size = 0.2)
-
-ggsave("output/figures/HUB-mean-var-obs.png", width = 7, height = 2.5)
-
+}
+
+p1 <- plot_var_mean(hub_data)
+p2 <- plot_var_mean(hub_data, log = TRUE)
+
+wis_vs_total <- function(scores, summary_fct = mean, fulldata = hub_data) {
+
+  total_values <- 
+    fulldata |> 
+    filter(!is.na(prediction)) |>
+    group_by(location, target_type, target_end_date) |>
+    summarise(true_value = unique(true_value)) |>
+    group_by(location, target_type) |>
+    summarise(total = sum(true_value))
+
+  plot_df <- scores |> 
+    filter(model == "EuroCOVIDhub-ensemble", 
+           horizon == 2) |>
+    group_by(location, target_type, scale, type_and_scale) |>
+    summarise(wis = summary_fct(interval_score), 
+              .groups = "drop_last") |> 
+    inner_join(total_values)
+
+  plot_df |>
+    group_by(target_type, scale) |>
+    summarise(correlation = cor(wis, total)) |> 
+    print()
+
+  plot_df |>
+    ggplot(aes(y = wis, x = total)) + 
+    geom_point() + 
+    facet_wrap(~ type_and_scale, scale = "free") + 
+    theme_scoringutils() + 
+    theme(legend.position = "none") + 
+    scale_x_continuous(labels = scale_fn) +
+    scale_y_continuous(labels = scale_fn) +
+    labs(y = "REPLACE", x = "Total number of observed values")
+}
+
+# log-log versions of that plot
+
+p3 <- wis_vs_total(scores) + 
+  labs(y = "Mean WIS") + 
+  scale_x_continuous(labels = scale_fn, trans = "log10") +
+  scale_y_continuous(labels = scale_fn, trans = "log10")
+
+p1 / p2 / p3 + 
+  plot_layout(heights = c(1, 1, 2))
+
+ggsave("output/figures/HUB-mean-scores-vs-total-log-log.png", width = 7, height = 6)
+
+# ggsave("output/figures/HUB-mean-scores-vs-total-log-log.png", width = 7, height = 4)
+
+# ggsave("output/figures/HUB-mean-var-obs.png", width = 7, height = 2.5)
 ```
@@
 
+```{r}
+v1 <- c("a","b","c","d","e","f","g")
+v2 <- c("b","x","e","c","z","d","a")
+
+o <- 0.05
+DF <- data.frame(x = c(rep(1, length(v1)), rep(2, length(v2))),
+                 x1 = c(rep(1 + o, length(v1)), rep(2 - o, length(v2))),
+                 y = c(rev(seq_along(v1)), rev(seq_along(v2))),
+                 g = c(v1, v2))
+
+library(ggplot2)
+library(grid)
+ggplot(DF, aes(x=x, y=y, group=g, label=g)) +
+  geom_path(aes(x=x1), arrow = arrow(length = unit(0.02,"npc")), 
+            size=1, color="green") +
+  geom_text(size=10) +
+  theme_minimal() +
+  theme(axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        panel.grid = element_blank())
+```
+
 ## Look at pairwise comparisons
@@
   mutate(model = factor(model, levels = ranking_natural), 
-         difference = natural - log)
+         difference = log - natural)
 
@@
 ```{r}
+
+# 1. observations natural vs. log --> pearson
+# 2. correlation scores natural vs. log (single forecasts) --> spearman
+
+# Table
 # correlation between values
-hub_data |>
+cor_data <- hub_data |>
   select(location, target_end_date, target_type, true_value) |>
@@
 
-# correlation between scores
-cor_scores <- scores |>
-  filter(horizon <=4) |>
-  select(interval_score, scale, forecast_date, model, horizon, location, target_type) |>
-  pivot_wider(names_from = scale, values_from = interval_score) |>
-  group_by(horizon, target_type) |>
-  summarise(cor = cor(natural, log), 
-            spearman = cor(natural, log))
-
-cor_scores |> 
-  mutate(`Target type` = target_type) |>
-  ggplot(aes(x = horizon, y = cor, 
-             fill = `Target type`, color = `Target type`)) + 
-  geom_histogram(stat = "identity", position = "dodge") +
-  theme_scoringutils() + 
-  coord_cartesian(ylim = c(0, 1)) +
-  labs(y = "Correlation (scores natural vs. log)", x = "Forecast horizon")
+# # correlation between scores
+# # not really meaningful: if we do ranks it's just going to be ranks over everything which doesn't make sense
+# # if we do individual scores than it's a bit unclear what the "correct" correlation should be
+# cor_scores <- scores |>
+#   filter(horizon <=4) |>
+#   select(interval_score, scale, forecast_date, model, horizon, location, target_type) |>
+#   pivot_wider(names_from = scale, values_from = interval_score) |>
+#   group_by(horizon, target_type) |>
+#   summarise(cor = cor(natural, log), 
+#             spearman = cor(natural, log, method = "spearman"))
 
@@
   group_by(horizon, target_type) |>
-  summarise(cor = cor(natural, log)) |>
+  summarise(cor = cor(natural, log), 
+            spearman = cor(natural, log, method = "spearman")) |>
   rename(`Target type` = target_type)
-
-plot_cor_ranks <- correlation_ranks |>
-  ggplot(aes(x = horizon, y = cor, 
-             fill = `Target type`, color = `Target type`)) + 
-  geom_histogram(stat = "identity", position = "dodge") +
-  theme_scoringutils() + 
-  coord_cartesian(ylim = c(0, 1)) +
-  labs(y = "Correlation (rank natural vs. log)", x = "Forecast horizon")
-
-ggsave(filename = "output/figures/HUB-cor-ranks.png", width = 7, height = 4.5, 
-       plot = plot_cor_ranks)
 
@@
   group_by(horizon, target_type) |>
-  summarise(cor = cor(natural, log))
-
-plot_cor_skill <- correlation_rel_skill |>
+  summarise(cor = cor(natural, log), 
+            spearman = cor(natural, log, method = "spearman"))
+
+p_cor_ranks <- correlation_ranks |>
+  ggplot(aes(x = horizon, y = cor, 
+             fill = `Target type`, color = `Target type`)) + 
+  geom_histogram(stat = "identity", position = "dodge") +
+  theme_scoringutils() + 
+  coord_cartesian(ylim = c(0, 1)) +
+  labs(y = "Corr. ranks single forecasts", x = "Forecast horizon")
+
+# p_cor_scores <- cor_scores |> 
+#   mutate(`Target type` = target_type) |>
+#   ggplot(aes(x = horizon, y = cor, 
+#              fill = `Target type`, color = `Target type`)) + 
+#   geom_histogram(stat = "identity", position = "dodge") +
+#   theme_scoringutils() + 
+#   coord_cartesian(ylim = c(0, 1)) +
+#   labs(y = "Spearman corr. scores single forecasts", x = "Forecast horizon")
+
+p_cor_skill <- correlation_rel_skill |>
   rename(`Target type` = target_type) |>
   ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
-  # geom_line() +
-  # geom_point() +
   geom_histogram(stat = "identity", position = "dodge") +
@@
   coord_cartesian(ylim = c(0, 1)) +
-  labs(y = "Correlation (rel. skill natural vs. log)", x = "Forecast horizon")
-
-ggsave(filename = "output/figures/HUB-cor-rel-skill.png", width = 7, height = 4.5, 
-       plot = plot_cor_skill)
-
-plot_cors <- plot_cor_ranks + plot_cor_skill +
-  plot_layout(guides = "collect") & 
+  labs(y = "Correlation rel. skill", x = "Forecast horizon")
+
+plot_cors <- p_cor_ranks  + p_cor_skill +
+  plot_layout(guides = "collect") +
+  plot_annotation(tag_levels = "A") &
   theme(legend.position = "bottom")
 
-ggsave(filename = "output/figures/HUB-cors-skill-rank.png", width = 7, height = 3, 
+ggsave(filename = "output/figures/HUB-cors.png", width = 7, height = 3, 
        plot = plot_cors)
@@
 ```{r}
-wis_vs_total <- function(scores, summary_fct = mean, fulldata = hub_data) {
-
-  total_values <- 
-    fulldata |> 
-    filter(!is.na(prediction)) |>
-    group_by(location, target_type, target_end_date) |>
-    summarise(true_value = unique(true_value)) |>
-    group_by(location, target_type) |>
-    summarise(total = sum(true_value))
-
-  plot_df <- scores |> 
-    filter(model == "EuroCOVIDhub-ensemble", 
-           horizon == 2) |>
-    group_by(target_type, location, scale) |>
-    summarise(wis = summary_fct(interval_score), 
-              .groups = "drop_last") |> 
-    inner_join(total_values)
-
-  plot_df |>
-    group_by(target_type, scale) |>
-    summarise(correlation = cor(wis, total)) |> 
-    print()
-
-  plot_df |>
-    ggplot(aes(y = wis, x = total)) + 
-    geom_point() + 
-    facet_wrap(~ target_type + scale, scale = "free") + 
-    theme_scoringutils() + 
-    theme(legend.position = "none") + 
-    scale_x_continuous(labels = scale_fn) +
-    scale_y_continuous(labels = scale_fn) +
-    labs(y = "REPLACE", x = "Total number of observed values")
-}
-
-wis_vs_total(scores) + 
-  labs(y = "Mean WIS")
-ggsave("output/figures/HUB-mean-scores-vs-total.png", width = 7, height = 5)
-
-wis_vs_total(scores, summary_fct = stats::sd) + 
-  labs(y = "Sd WIS")
-ggsave("output/figures/HUB-sd-scores-vs-total.png", width = 7, height = 5)
-
-# log-log versions of that plot
-wis_vs_total(scores) + 
-  labs(y = "Mean WIS") + 
-  scale_x_continuous(labels = scale_fn, trans = "log10") +
-  scale_y_continuous(labels = scale_fn, trans = "log10") 
-ggsave("output/figures/HUB-mean-scores-vs-total-log-log.png", width = 7, height = 5)
-
-wis_vs_total(scores, summary_fct = stats::sd) + 
-  labs(y = "Sd WIS") + 
-  scale_x_continuous(labels = scale_fn, trans = "log10") +
-  scale_y_continuous(labels = scale_fn, trans = "log10") 
-ggsave("output/figures/HUB-sd-scores-vs-total-log-log.png", width = 7, height = 5)
-
-# correlation
+
+# wis_vs_total(scores) + 
+#   labs(y = "Mean WIS")
+# ggsave("output/figures/HUB-mean-scores-vs-total.png", width = 7, height = 5)
+# 
+# wis_vs_total(scores, summary_fct = stats::sd) + 
+#   labs(y = "Sd WIS")
+# ggsave("output/figures/HUB-sd-scores-vs-total.png", width = 7, height = 5)
+
+# wis_vs_total(scores, summary_fct = stats::sd) + 
+#   labs(y = "Sd WIS") + 
+#   scale_x_continuous(labels = scale_fn, trans = "log10") +
+#   scale_y_continuous(labels = scale_fn, trans = "log10") 
+# ggsave("output/figures/HUB-sd-scores-vs-total-log-log.png", width = 7, height = 5)
 

### 2022-08-02 main.tex
update manuscript

--- 2022-07-21 main.tex
@@
 The forecasts for the purpose of this illustrations are two-week-ahead forecasts made between XXX and XXX for reported cases and deaths from COVID-19 in XXX countries.
+We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 16 forecast dates.
 Where not otherwise stated, we report results for a two-week-ahead forecast horizon.
 
-Pooled across all time points and locations, the mean number of cases (deaths) observed was 22811 (368) with an overall standard deviation of 47135 (831) (see Table \ref{tab:HUB-summary}).
-The average number of observed cases and deaths varies considerably by location (see Figure \ref{fig:HUB-mean-locations}A and B).
-Observations are overdispersed, meaning that for a given location, the variance in observations exceeds the mean of all observed values (see Figure \ref{fig:HUB-variance-mean}).
-%maybe make a line of best fit of the form variance = alpha * mean^2 + beta * mean
-
-Pooled across all time points and locations, the mean WIS on the natural scale for cases (deaths) was 8855 (59.3) with an overall standard deviation of 51274 (156) (see Table \ref{tab:HUB-summary}).
-On the log scale, scores were much closer to each other (see Figure \ref{fig:HUB-average-scores}), with a mean WIS for cases (deaths) of 0.507 (0.362) with an overall standard deviation of 0.745 (0.407).
-
-Variation of scores across targets (see Figure XX) and locations (see Figure \ref{fig:HUB-mean-locations}A, B) is much larger on the natural than on the log scale, where scores are more evenly distributed.
-On the natural scale, average scores per location correlate strongly (XX) with the average number of observed cases or deaths (See Figure XX).
-On the log scale, this correlation is much weaker and slightly negative.
-Forecasts across different locations receive more similar scores, with a tendency for scores to be slightly higher in locations with few observed cases and deaths (see Figures XX and \ref{fig:HUB-mean-locations}.
-%This is in line with the observation that the association between variance is smaller than mu^2 + mu, meaning that we would expect this. 
-For individual forecasts, rankings between models are mostly preserved (correlation: X for cases, Y for deaths), with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors-scores}).
-
-The correlation between scores on the log and natural scale for different locations is XX for cases and XX for deaths. 
-The correlation between the actual observations is 0.628 for cases and 0.631 for deaths. IS THAT HELPFUL?
-
-\begin{table}[h!]]
+\begin{table}[h!]
     \centering
@@
     Cases & Observations & var & 1884002308 & 4.46\\
+    \addlinespace
     Deaths & Observations & mean & 259 & 3.79\\
     Deaths & Observations & sd & 532 & 2.11\\
+    Deaths & Observations & var & 283492 & 4.46\\
     \addlinespace
-    Deaths & Observations & var & 283492 & 4.46\\
+    \hline
+    \addlinespace
     Cases & WIS & mean & 4544 & 0.28\\
     Cases & WIS & sd & 17585 & 0.49\\
+    \addlinespace
     Deaths & WIS & mean & 28 & 0.22\\
@@
 
-    \caption{CAPTION}
+    \caption{Table with summary statistics for observations and scores for forecasts from the ECDC data set.}
     \label{tab:HUB-summary}
 \end{table}
+
+Pooled across all time points and locations, the mean number of cases (deaths) observed was 22811 (368) with an overall standard deviation of 47135 (831) (see Table \ref{tab:HUB-summary}).
+The average number of observed cases and deaths varies considerably by location (see Figure \ref{fig:HUB-mean-locations}A and B).
+Observations are overdispersed, meaning that for a given location, the variance in observations exceeds the mean of all observed values (see Figure \ref{fig:HUB-variance-mean}).
+%maybe make a line of best fit of the form variance = alpha * mean^2 + beta * mean
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-average-scores.png}
+    \caption{Distribution of interval scores for two week ahead forecasts of COVID-19 cases and deaths evaluated on the natural scale (left) and on the log scale (right). }
+    \label{fig:HUB-average-scores}
+\end{figure}
+
+Pooled across all time points and locations, the mean WIS on the natural scale for cases (deaths) was 8855 (59.3) with an overall standard deviation of 51274 (156) (see Table \ref{tab:HUB-summary}).
+On the log scale, scores were much closer to each other (see Figure \ref{fig:HUB-average-scores}), with a mean WIS for cases (deaths) of 0.507 (0.362) with an overall standard deviation of 0.745 (0.407).
+
+Variation of scores across targets (see Figure \ref{fig:HUBXXX} and locations (see Figure \ref{fig:HUB-mean-locations}A, B) is much larger on the natural than on the log scale, where scores are more evenly distributed.
+On the natural scale, average scores per location correlate strongly (XX) with the average number of observed cases or deaths (See Figure XX).
+On the log scale, this correlation is much weaker and slightly negative.
+Forecasts across different locations receive more similar scores, with a tendency for scores to be slightly higher in locations with few observed cases and deaths (see Figures XX and \ref{fig:HUB-mean-locations}.
+%This is in line with the observation that the association between variance is smaller than mu^2 + mu, meaning that we would expect this. 
+For individual forecasts, rankings between models are mostly preserved (correlation: X for cases, Y for deaths), with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}).
 
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-cors-skill-rank.png}
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-cors.png}
     \caption{Correlation between scores on the natural and on the log scale.
 A: Correlation between relative ranks (the rank in terms of WIS divided by the number of models) for individual forecasts, separated by forecast horizon.
-B: Correlation between relative skill values achieved by models across locations and time points, separated by forecast horizon.}
-    \label{fig:HUB-cors-scores}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-ranking-diffs.png}
-    \caption{Distribution of rankings of the different Hub models.
-Rankings are computed as relative rankings, i.e.
-the ranking divided by the total number of models available for a given forecast target.}
-    \label{fig:HUB-rankings}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-average-scores.png}
-    \caption{Distribution of interval scores for two week ahead forecasts of COVID-19 cases and deaths evaluated on the natural scale (left) and on the log scale (right). }
-    \label{fig:HUB-average-scores}
+B: Correlation between relative skill values achieved by models across locations and time points, separated by forecast horizon.
+SHOULD THIS BE SPEARMAN?}
+    \label{fig:HUB-cors}
 \end{figure}
@@
 The increase is more pronounced for cases than for deaths and arguably represents an increase in the relative difficulty to forecast values further into the future.
-A similar increase both on the natural and the log scale indicates that we are able to observe the increase in difficulty regardless of how we evaluate the scores.
+A similar increase both on the natural and the log scale suggests that we are able to observe the increase in difficulty regardless of how we evaluate the scores.
 

### 2022-08-02 analysis/example-log-first-Johannes.R

@@ -20,18 +20,32 @@ forecasts <- forecasts |>
          sample_id = list(1:length(Y))) |>
   unnest(c(true_value, sample_id))
 
+forecasts <- forecasts |>
+  mutate(scale = "natural") |>
+  rbind(forecasts |>
+          mutate(prediction = log(prediction + 1), 
+                 true_value = log(true_value + 1), 
+                 scale = "log"))
 
 scores <- score(forecasts)
 
 summary <- scores |>
   mutate(log_wis = log(interval_score +1)) |>
-  group_by(sigma) |>
+  group_by(sigma, scale) |>
   summarise(wis = mean(interval_score), 
-            log_wis = mean(log_wis))
+            log_wis = mean(log_wis)) |>
+  mutate(log_wis = ifelse(scale == "log", log_wis, NA)) |>
+  pivot_wider(values_from = wis, names_from = scale) |>
+  group_by(sigma) |>
+  summarise(log_wis = sum(log_wis, na.rm = TRUE), 
+            wis_log = sum(log, na.rm = TRUE), 
+            wis = sum(natural, na.rm = TRUE))
 
 summary |> 
-  pivot_longer(cols = c(wis, log_wis), values_to = "score", names_to = "type") |>
-  mutate(type = factor(type, levels = c("wis", "log_wis"), labels = c("WIS", "log(WIS)"))) |>
+  pivot_longer(cols = c(wis, log_wis, wis_log), values_to = "score", names_to = "type") |>
+  mutate(type = factor(type, 
+                       levels = c("wis", "wis_log", "log_wis"), 
+                       labels = c("WIS", "WIS (log scale)", "log(WIS)"))) |>
   ggplot(aes(x = sigma, y = score)) +
   geom_vline(xintercept = 1, linetype = "dashed", size = 0.3, color = "grey80") +
   geom_point() + 
@@ -39,4 +53,4 @@ summary |>
   theme_scoringutils() +
   labs(y = "Score", x = "Standard deviation of predictive distribution")
 
-ggsave("output/figures/example-log-first.png", width = 7, height = 2.5)
+ggsave("output/figures/example-log-first.png", width = 8, height = 2.5)

### 2022-08-02 analysis/illustration-effect-log-score.R

@@ -28,13 +28,15 @@ vals <- expand.grid(
 
 scores <- vals |>
   mutate(scale = "natural") |>
-  eval_forecasts() |> 
+  score() |> 
+  summarise_scores(by = c("scale", "id")) |>
   rbind(
     vals |>
       mutate(true_value = log(true_value), 
              prediction = log(prediction), 
              scale = "log") |>
-      eval_forecasts()
+      score() |>
+      summarise_scores(by = c("scale", "id"))
   ) |>
   group_by(scale) |>
   mutate(score = interval_score / min(interval_score), 
@@ -50,8 +52,8 @@ p1 <- scores |>
   geom_line() + 
   labs(y = "WIS", x = "Observed value") + 
   scale_y_continuous(label = function(x) {paste(scale_factor * x)}) + 
-  facet_wrap(~ scale) + 
-  theme_minimal()
+  facet_wrap(~ scale, ncol = 1, scales = "free") + 
+  theme_scoringutils()
 
 scale_factor2 <- 3
 p2_log <- scores |>
@@ -69,10 +71,10 @@ p2_log <- scores |>
                               paste(x))
                      }, 
                      breaks = c(0.2, 1/2, 1, 2, 5)) + 
-  facet_wrap(~ scale) + 
-  theme_minimal()
+  facet_wrap(~ scale, ncol = 1, scales = "free") + 
+  theme_scoringutils()
 
-p1 / p2_log
+p1 + p2_log
 
 ggsave("output/figures/SIM-effect-log-score.png", width = 6, 
        height = 4)

### 2022-08-21 main.tex
manuscript update

@@
 
-%Score different things from the European / US Forecast Hub and plot the relationship between mean and variance.
-Plot log-scale scores for different things and see which of these influence average scores most \\
-% What happens to the decomposition of the WIS if we log?
-
-%\begin{itemize}
-%    \item Optional: Discussion of parallels to point forecasts and the point that you're still incentivised to report the median\\
-%\end{itemize}
-
 \section{Empirical example: the European Forecast Hub}
 
+\paragraph{Intro Hub}
 As an empirical example for evaluating forecasts on the natural and on the log scale we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
@@
 
-\begin{table}[h!]
+In addition to the WIS we use pairwise comparisons (CITATION Kramer) to evaluate models against each other.
+In a first step, score ratios are computed for all pairs of models by taking the set of overlapping forecasts between the two models and dividing the score of one model over the score achieved by the other model.
+The relative skill for a given model is then obtained by taking the geometric mean of all score ratios which involve that model.
+Low values are better, and the average model receives a relative skill score of 1.
+
+\begin{figure}[h!]
     \centering
-
-    \begin{tabular}{lllcc}
-    \toprule
-    target\_type & quantity & measure & natural & log\\
-    \midrule
-    Cases & Observations & mean & 20308 & 8.39\\
-    Cases & Observations & sd & 43405 & 2.11\\
-    Cases & Observations & var & 1884002308 & 4.46\\
-    \addlinespace
-    Deaths & Observations & mean & 259 & 3.79\\
-    Deaths & Observations & sd & 532 & 2.11\\
-    Deaths & Observations & var & 283492 & 4.46\\
-    \addlinespace
-    \hline
-    \addlinespace
-    Cases & WIS & mean & 4544 & 0.28\\
-    Cases & WIS & sd & 17585 & 0.49\\
-    \addlinespace
-    Deaths & WIS & mean & 28 & 0.22\\
-    Deaths & WIS & sd & 67 & 0.26\\
-    \bottomrule
-    \end{tabular}
-
-    \caption{Table with summary statistics for observations and scores for forecasts from the ECDC data set.}
-    \label{tab:HUB-summary}
-\end{table}
-
-Pooled across all time points and locations, the mean number of cases (deaths) observed was 22811 (368) with an overall standard deviation of 47135 (831) (see Table \ref{tab:HUB-summary}).
-The average number of observed cases and deaths varies considerably by location (see Figure \ref{fig:HUB-mean-locations}A and B).
-Observations are overdispersed, meaning that for a given location, the variance in observations exceeds the mean of all observed values (see Figure \ref{fig:HUB-variance-mean}).
-%maybe make a line of best fit of the form variance = alpha * mean^2 + beta * mean
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-var-obs.png}
+    \caption{Observed mean against observed variance of observations in different locations across time.
+The dashed line marks what would be expecetd from a negative binomially distributed variable with variance of the observed value equal to the mean + the square of mean observed value.
+}
+    \label{fig:HUB-variance-mean}
+\end{figure}
 
@@
 
-Pooled across all time points and locations, the mean WIS on the natural scale for cases (deaths) was 8855 (59.3) with an overall standard deviation of 51274 (156) (see Table \ref{tab:HUB-summary}).
+SECTION STRUCUTRE / IMPORTANT THINGS
+\begin{itemize}
+\item X variation gets smaller / more even
+\item X relationship with country size changes
+\item X inndividual rankings stay similar
+\item relative skill scores change a lot
+\item deep dive into what happens for individual models
+\item scores increase across forecast horizons
+\end{itemize}
+
+Maybe look at correlation between bias and score?
+Correlation between relative WIS contributions.?
+
+\paragraph{Mean / sd of cases and WIS values}
+Pooled across all time points and locations, the mean number of cases (deaths) observed was 22811 (368) with an overall standard deviation of 47135 (831) (see Table \ref{tab:HUB-summary}).
+The average number of observed cases and deaths varies considerably by location (see Figure \ref{fig:HUB-mean-locations}A and B).
+Observations were overdispersed, meaning that for a given location, the variance in observations exceeds the mean of all observed values (see Figure \ref{fig:HUB-variance-mean}).
+%Could fit a model var = alpha * x2 + beta *x
+
+On the natural scale, scores vary considerably across targets (see Fig \ref{}) and locations (see Fig \ref{}) and correlate strongly (XX) with the average number of observed cases / deaths in a given country (see Fig \ref{fig:HUB-mean-locations}A, B).
+On the log scale, scores are more evenly distributed between targets (see Fig \ref{}) and locations (see Fig \ref{fig:HUB-mean-locations}.
+The correlation between average WIS and the mean number of observations in a country is much weaker and tends to be negative.
+Locations therefore tend to have a more equal influence on avearage scores, with smaller countries potentially being slightly overrepresented.
+%This is in line with the observation that the association between variance is smaller than mu^2 + mu, meaning that we would expect this.
+
+ % Pooled across all time points and locations, the mean WIS on the natural scale for cases (deaths) was 8855 (59.3) with an overall standard deviation of 51274 (156) (see Table \ref{tab:HUB-summary}).
 On the log scale, scores were much closer to each other (see Figure \ref{fig:HUB-average-scores}), with a mean WIS for cases (deaths) of 0.507 (0.362) with an overall standard deviation of 0.745 (0.407).
 
-Variation of scores across targets (see Figure \ref{fig:HUBXXX} and locations (see Figure \ref{fig:HUB-mean-locations}A, B) is much larger on the natural than on the log scale, where scores are more evenly distributed.
-On the natural scale, average scores per location correlate strongly (XX) with the average number of observed cases or deaths (See Figure XX).
-On the log scale, this correlation is much weaker and slightly negative.
-Forecasts across different locations receive more similar scores, with a tendency for scores to be slightly higher in locations with few observed cases and deaths (see Figures XX and \ref{fig:HUB-mean-locations}.
-%This is in line with the observation that the association between variance is smaller than mu^2 + mu, meaning that we would expect this. 
-For individual forecasts, rankings between models are mostly preserved (correlation: X for cases, Y for deaths), with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}).
+% When evaluated on the natural scale, there is  strong relationship between the mean of the weighted interval score and the total number of observed cases (correlation: 0.913) or deaths (cor: 0.849) in a location (see Figure \ref{fig:HUB-mean-scores-total-loglog} and Figure \ref{fig:HUB-mean-scores-total} in the SI).
+The relationship is less pronounced for scores on the log scale (cor: -0.019 for cases, -0.395 for deaths) and negative, meaning that locations with fewer observed cases or deaths tend to receive higher scores.
 
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-var-obs.png}
-    \caption{Observed mean against observed variance of observations in different locations across time.
-The dashed line marks what would be expecetd from a negative binomially distributed variable with variance of the observed value equal to the mean + the square of mean observed value.
-}
-    \label{fig:HUB-variance-mean}
-\end{figure}
+\paragraph{individual rankings}
+For individual forecasts, rankings between models for single forecasts are mostly preserved (correlation: X for cases, Y for deaths), with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}).
+When evaluating performance across different forecasts and forecast targets, relative skill scores of the models change considerably.
+At least for cases, the effect is stronger with increasing forecast horizon.
 
@@
 
-When evaluated on the natural scale, there is  strong relationship between the mean of the weighted interval score and the total number of observed cases (correlation: 0.913) or deaths (cor: 0.849) in a location (see Figure \ref{fig:HUB-mean-scores-total-loglog} and Figure \ref{fig:HUB-mean-scores-total} in the SI).
-The relationship is less pronounced for scores on the log scale (cor: -0.019 for cases, -0.395 for deaths) and negative, meaning that locations with fewer observed cases or deaths tend to receive higher scores.
-
 \begin{figure}[h!]
@@
     \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-scores-vs-total-log-log.png}
-    \caption{Plot with Weighted interval scores against the mean number of observed cases or deaths.}
+    \caption{My current thinking is that the top row is helpful, as it explalins what the mean / var relationship is which is relevant for how much weight small countries get.
+I think the 2nd row is just completely useless.
+The 3rd and 4th row may be helpful.
+In some sense it is duplicated information from above, but it gives some nice intuiton for the correlations.
+I could alternatively also turn it into a correlation plot and add it to the other one.
+}
     \label{fig:HUB-mean-scores-total-loglog}
@@
 
-\paragraph{Rankings}
-So one effect is that for a given dispersion on the natural scale, it's better if that interval is skewed, as that implies that at least you only off a lot in one direction in negative terms, and not in the other direction.
+% \paragraph{Rankings}
+% So one effect is that for a given dispersion on the natural scale, it's better if that interval is skewed, as that implies that at least you only off a lot in one direction in negative terms, and not in the other direction.
 Maybe looking at over-and underprediction is the key here.
@@
     \includegraphics[width=0.9\textwidth]{output/figures/HUB-pairwise-comparisons.png}
-    \caption{Changes in relative skill and the distribution of scores}
+    \caption{Changes in relative skill and the distribution of scores. MAYBE THIS SHOULD ONLY BE CASES?}
     \label{fig:HUB-pairwise}
@@
 
-Correlation between pairwise comparisons is 0.4414717.
-
-% \section{Other ideas}
-% \begin{itemize}
-%     \item Compare different predictive distributions and the score as a function of an outcome.
-%     \item ...
-% \end{itemize}
+% Correlation between pairwise comparisons is 0.4414717.
 
@@
 \section{Supplementary information}
+
+\begin{table}[h!]
+    \centering
+
+    \begin{tabular}{lllcc}
+    \toprule
+    target\_type & quantity & measure & natural & log\\
+    \midrule
+    Cases & Observations & mean & 20308 & 8.39\\
+    Cases & Observations & sd & 43405 & 2.11\\
+    Cases & Observations & var & 1884002308 & 4.46\\
+    \addlinespace
+    Deaths & Observations & mean & 259 & 3.79\\
+    Deaths & Observations & sd & 532 & 2.11\\
+    Deaths & Observations & var & 283492 & 4.46\\
+    \addlinespace
+    \hline
+    \addlinespace
+    Cases & WIS & mean & 4544 & 0.28\\
+    Cases & WIS & sd & 17585 & 0.49\\
+    \addlinespace
+    Deaths & WIS & mean & 28 & 0.22\\
+    Deaths & WIS & sd & 67 & 0.26\\
+    \bottomrule
+    \end{tabular}
+    \caption{Table with summary statistics for observations and scores for forecasts from the ECDC data set.}
+    \label{tab:HUB-summary}
+\end{table}
 
@@
 
+% \section{Other ideas}
+% \begin{itemize}
+%     \item Compare different predictive distributions and the score as a function of an outcome.
+%     \item ...
+% \end{itemize}
+
+%Score different things from the European / US Forecast Hub and plot the relationship between mean and variance.
+Plot log-scale scores for different things and see which of these influence average scores most \\
+% What happens to the decomposition of the WIS if we log?
+
+%\begin{itemize}
+%    \item Optional: Discussion of parallels to point forecasts and the point that you're still incentivised to report the median\\
+%\end{itemize}
+
 % NOT SO CONVINCED ANYMORE THAT THIS WORKS. MY INTUITION IS IF YOU WANT THE GROWTH RATE YOU SHOULD ACTUALLY SCORE THE GROWTH RATE AND DIVIDE BY THE LAST OBSERVED VALUE. 

### 2022-08-24 analysis/Analysis ECDC forecasts.Rmd
update plots

@@
 library(tidytext)
+library(grid)
 
@@
          paste0(x / 1000000, "m"), 
-         ifelse(x >= 5000,
+         ifelse(x >= 1000,
                 paste0(x / 1000, "k"),
@@
 ## FIGURE 4
-## Mean vs. sd of observed values
+## Mean vs. sd of observed values + boxplots for distribution across target type + location
 
@@
 plot_means_obs <- mean_observations |>
-  ggplot(aes(y = mean_value, x = reorder_within(location, -mean_value, target_type))) + 
-  geom_bar(stat = "identity", fill = "grey50") + 
+  ggplot(aes(y = mean_value, x = reorder_within(location, -mean_value, target_type))) +   geom_bar(stat = "identity", fill = "grey50") + 
   facet_wrap(~ target_type, scales = "free") + 
@@
   labs(y = "Mean observed value", x = "Location")
+
+box_plot_obs <- hub_data |>
+  mutate(true_value = pmax(true_value, 1)) |>
+  dplyr::select(location, target_type, target_end_date, true_value) |>
+  unique() |>
+  ggplot(aes(y = true_value, x = target_type)) + 
+  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
+  geom_boxplot(alpha = 0.5) + 
+  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
+  theme_scoringutils() + 
+  theme(legend.position = "none") + 
+  scale_y_continuous(labels = scale_fn, trans = "log10") + 
+  labs(y = "Observations", x = "Target type")
 
@@
   scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) +
-  labs(y = "Interval score", x = "Location")
-
-plot_means_obs / 
-  plot_mean_scores + 
-  plot_layout(heights = c(1, 2))
-
-ggsave("output/figures/HUB-mean-obs-location.png", width = 9, height = 7)
+  labs(y = "Mean interval score", x = "Location")
+
+box_plot_scores <- scores |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  ggplot(aes(y = interval_score, x = target_type)) + 
+  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
+  geom_boxplot(alpha = 0.5) + 
+  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
+  facet_wrap(~ scale, scale = "free", nrow = 2) + 
+  theme_scoringutils() + 
+  theme(legend.position = "none") + 
+  scale_y_continuous(labels = scale_fn, trans = "log10") + 
+  labs(y = "Interval score", x = "Target type")
+
+box_plot_horizon <- scores |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon <= 4) |>
+  group_by(forecast_date, model, location, target_type, scale) |>
+  mutate(interval_score = interval_score / interval_score[horizon == 1]) |>
+  ggplot(aes(y = interval_score,
+             x = as.factor(horizon), fill = target_type)) + 
+  geom_hline(aes(yintercept = 1), linetype = "dashed", color = "grey40") +
+  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) +
+  geom_boxplot(alpha = 0.5, position = "dodge2") + 
+  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
+  facet_wrap(~ scale, nrow = 1) + 
+  theme_scoringutils() + 
+  theme(legend.position = "none") + 
+  scale_y_continuous(labels = scale_fn, trans = "log10") +
+  labs(y = "Rel. increase in WIS", x = "Forecast horizon (weeks)")
+
+layout <- "
+AB
+CD
+CD
+EE"
+
+plot_means_obs  + box_plot_obs + plot_mean_scores + box_plot_scores + box_plot_horizon +
+  plot_layout(heights = c(1, 2, 1), 
+              widths = c(3, 1)) +
+  plot_layout(design = layout) +
+  plot_annotation(tag_levels = "A")
+
+ggsave("output/figures/HUB-mean-obs-location.png", width = 10, height = 10)
 
@@
 
-Figure 5? 
+Figure
+Currently discontinuoued and just there to discuss it. 
 that shows Variance vs. mean of the observations. Makes the point that the
@@
   df <- hub_data |>
-    dplyr::select(location, target_type, target_end_date, true_value) |>
+    mutate("x" = true_value) |>
+    dplyr::select(location, target_type, target_end_date, true_value, x) |>
     unique() |>
@@
 
-    f <- function(x) {log(x + x^2 + 1)}
-  }
-
-  df |>
-  summarise(mean_value = mean(true_value), 
+    f <- function(x) {log(log(x) + log(x)^2 + 1)}
+  }
+
+  out <- df |>
+  summarise(mean_value = mean(x), 
             sd_value = sd(true_value), 
@@
   scale_y_continuous(trans = "log10", labels = label_fn) +
-  labs(y = "Variance observed value", x = "Mean observed value") + 
-  geom_function(fun = f, linetype = "dashed", size = 0.5) + 
+  labs(y = "Var observations", x = "Mean observations") + 
+  # geom_function(fun = f, linetype = "dashed", size = 0.5) +
   theme_scoringutils() + 
   geom_smooth(alpha = 0.2, size = 0.2)
+
+  if (log) {
+    out <- out +
+      labs(y = "Var log(observations)", x = "Mean observations")
+  }
+
+  return(out)
 }
@@
     group_by(location, target_type) |>
-    summarise(total = sum(true_value))
+    summarise(total = mean(true_value))
 
@@
 
-p3 <- wis_vs_total(scores) + 
+scatter_wis_obs <- wis_vs_total(scores) + 
   labs(y = "Mean WIS") + 
@@
 
-p1 / p2 / p3 + 
+p1 / p2 / scatter_wis_obs + 
   plot_layout(heights = c(1, 1, 2))
 
-ggsave("output/figures/HUB-mean-scores-vs-total-log-log.png", width = 7, height = 6)
-
-# ggsave("output/figures/HUB-mean-scores-vs-total-log-log.png", width = 7, height = 4)
-
-# ggsave("output/figures/HUB-mean-var-obs.png", width = 7, height = 2.5)
+ggsave("output/figures/HUB-mean-scores-vs-total-log-log.png", width = 7, height = 7)
+```
+
+## Figure 5 - Correlations between everything
+```{r}
+df <- scores |>
+  filter(horizon < 5) |>
+  select(-c(coverage_deviation, bias, ae_median, dispersion, underprediction, overprediction, type_and_scale)) |>
+  pivot_wider(names_from = scale, values_from = interval_score) |> 
+  group_by(horizon, target_type, forecast_date, target_end_date, location) |>
+  summarise(cor = cor(natural, log), 
+            spearman = cor(natural, log, method = "spearman")) |>
+  group_by(horizon, target_type) |>
+  summarise(cor = mean(cor, na.rm = TRUE), 
+            spearman = mean(cor, na.rm = TRUE)) |>
+  rename(`Target type` = target_type)
+
+p_cor_scores <- df |>
+  ggplot(aes(x = horizon, y = cor, 
+             fill = `Target type`, color = `Target type`)) + 
+  geom_histogram(stat = "identity", position = "dodge") +
+  theme_scoringutils() + 
+  coord_cartesian(ylim = c(0, 1)) +
+  labs(y = "Avg. Spearman cor(scores)", x = "Forecast horizon")
+
+# Correlation between pairwise comparisons
+correlation_rel_skill <- scores |>
+  filter(horizon < 5) |>
+  summarise_scores(
+    by = c("model", "scale", "horizon", "target_type"), 
+    relative_skill = TRUE) |>
+  select(model, horizon, target_type, relative_skill, scale) |>
+  pivot_wider(values_from = relative_skill, names_from = scale) |>
+  group_by(horizon, target_type) |>
+  summarise(cor = cor(natural, log), 
+            spearman = cor(natural, log, method = "spearman"))
+
+p_cor_skill <- correlation_rel_skill |>
+  rename(`Target type` = target_type) |>
+  ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
+  geom_histogram(stat = "identity", position = "dodge") +
+  theme_scoringutils() + 
+  coord_cartesian(ylim = c(0, 1)) +
+  labs(y = "Correlation rel. skill", x = "Forecast horizon")
+
+(scatter_wis_obs) / (p_cor_scores + p_cor_skill) +
+  plot_layout(guides = "collect") +
+  plot_annotation(tag_levels = "A") +
+  plot_layout(widths = c(1, 1), heights = c(2, 1)) &
+  theme(legend.position = "bottom")
+
+ggsave("output/figures/HUB-correlations.png", width = 7, height = 6)
+
 ```
@@
             var_value = var(interval_score))
-
-scores |> 
-  filter(model == "EuroCOVIDhub-ensemble", 
-         horizon == 2) |>
-  ggplot(aes(y = interval_score, x = target_type)) + 
-  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
-  geom_boxplot(alpha = 0.5) + 
-  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
-  facet_wrap(~ scale, scale = "free") + 
-  theme_scoringutils() + 
-  theme(legend.position = "none") + 
-  scale_y_continuous(labels = scale_fn, trans = "log10") + 
-  labs(y = "Interval score", x = "Target type")
 
@@
 ```
+
+```{r}
+
+```
+
+## Figure 6 - Look at pairwise comparisons and changes in rankings
+
+```{r}
+summarised_pairwise <- scores |>
+  filter(target_type == "Cases", 
+         horizon == 2) |>
+  summarise_scores(by = c("model", "scale"), relative_skill = TRUE)
+
+ranking_natural <- 
+  summarised_pairwise |>
+  filter(scale == "natural") |>
+  arrange(-relative_skill) |>
+  pull(model)
+
+plot_rel_skill <- summarised_pairwise |>
+  mutate(model = factor(model, levels = ranking_natural)) |>
+  ggplot(aes(y = model, x = relative_skill)) +
+  geom_bar(stat = "identity") +
+  theme_scoringutils() +
+  facet_wrap(~scale) +
+  labs(y = "Model", x = "Relative skill")
+
+diffs <- summarised_pairwise |>
+  mutate(model = factor(model, levels = ranking_natural)) |>
+  select(model, relative_skill, scale) |>
+  pivot_wider(names_from = scale, values_from = relative_skill) |>
+  mutate(model = factor(model, levels = ranking_natural), 
+         difference = log - natural)
+
+plot_diffs <- diffs |>
+  mutate(title = "difference") |>
+  ggplot(aes(y = model, x = difference)) +
+  geom_bar(stat = "identity") +
+  theme_scoringutils() +
+  labs(y = "Model", x = "Difference in relative skill") +
+  theme(axis.title.y = element_blank(), 
+        axis.text.y = element_blank(), 
+        axis.ticks.y = element_blank(), 
+        axis.line.y = element_blank()) +
+  facet_wrap(~title)
+
+# plot_distr_scores <- scores |>
+#   filter(horizon == 2) |>
+#   mutate(model = factor(model, levels = ranking_natural)) |>
+#   ggplot(aes(y = model, x = interval_score)) +
+#   ggdist::stat_pointinterval() +
+#   facet_wrap(~ scale, scale = "free_x") +
+#   theme_scoringutils() +
+#   scale_x_log10(labels = label_fn) +
+#   theme(axis.title.y = element_blank(), 
+#         axis.text.y = element_blank(), 
+#         axis.ticks.y = element_blank(), 
+#         axis.line.y = element_blank()) +
+#   labs(y = "Model", x = "WIS")
+
+plot_rel_wis <- scores |>
+  filter(horizon == 2, 
+         target_type == "Cases") |>
+  mutate(model = factor(model, levels = ranking_natural)) |>
+  summarise_scores(by = c("model", "scale")) |>
+  plot_wis(relative_contributions = FALSE) +
+    theme(axis.title.y = element_blank(), 
+        axis.text.y = element_blank(), 
+        axis.ticks.y = element_blank(), 
+        axis.line.y = element_blank()) +
+  scale_x_continuous(labels = label_fn) +
+  facet_wrap(~scale, scales = "free_x")
+
+plot_ranking_change <- summarised_pairwise |>
+  select(model, scale, relative_skill) |>
+  group_by(scale) |>
+  mutate(rank = rank(-relative_skill), 
+         x = ifelse(scale == "natural", 0.4, 2.8), 
+         x_arrow = ifelse(scale == "natural", 1.7, 2.7)) |>
+  group_by(model) |>
+  mutate(color = ifelse(diff(relative_skill) > 0, "deteriorated", "not deteriorated")) |>
+  ggplot(aes(y = rank, x = x, group = model, label = model)) +
+    geom_path(aes(x=x_arrow, color = color), arrow = arrow(length = unit(0.02,"npc")), 
+            size=1, show.legend = FALSE) +
+  geom_text(size=3.5, hjust = 0) +
+  coord_cartesian(xlim = c(0.5, 4)) +
+  theme_scoringutils() +
+  theme(axis.title = element_blank(),
+        axis.line.x = element_blank(),
+        axis.line.y = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        panel.grid = element_blank(), 
+        legend.position="none")
+
+layout <- "
+AAA
+BCD"
+
+plot_ranking_change + plot_rel_skill + plot_diffs + plot_rel_wis + #plot_distr_scores 
+  plot_layout(design = layout) +
+  plot_layout(heights = c(1, 2)) +
+  plot_layout(widths = c(2, 1, 2)) +
+  plot_layout(guides = "collect") &
+  theme(legend.position = "bottom")
+
+ggsave("output/figures/HUB-pairwise-comparisons.png", width = 9, height = 5)
+
+# summarised_pairwise |>
+#   select(model, scale, relative_skill) |>
+#   pivot_wider(names_from = scale, values_from = relative_skill) |>
+#   ggplot(aes(y = log, x = natural)) +
+#   geom_point() + 
+#   # geom_abline(slope = 1) +
+#   theme_scoringutils()
+
+```
+
+# Correlations for ranks, scores and pairwise comparisons
+
+```{r}
+
+# 1. observations natural vs. log --> pearson
+# 2. correlation scores natural vs. log (single forecasts) --> spearman
+
+# Table
+# correlation between values
+cor_data <- hub_data |>
+  select(location, target_end_date, target_type, true_value) |>
+  unique() |>
+  mutate(log_obs = log(pmax(true_value, 0) + 1)) |>
+  group_by(target_type) |>
+  summarise(cor = cor(true_value, log_obs))
+
+# # correlation between scores
+# # not really meaningful: if we do ranks it's just going to be ranks over everything which doesn't make sense
+# # if we do individual scores than it's a bit unclear what the "correct" correlation should be
+# cor_scores <- scores |>
+#   filter(horizon <=4) |>
+#   select(interval_score, scale, forecast_date, model, horizon, location, target_type) |>
+#   pivot_wider(names_from = scale, values_from = interval_score) |>
+#   group_by(horizon, target_type) |>
+#   summarise(cor = cor(natural, log), 
+#             spearman = cor(natural, log, method = "spearman"))
+
+# Correlation between ranks when going from natural to log
+correlation_ranks <- rankings_wide |>
+  filter(horizon < 5) |>
+  group_by(horizon, target_type) |>
+  summarise(cor = cor(natural, log), 
+            spearman = cor(natural, log, method = "spearman")) |>
+  rename(`Target type` = target_type)
+
+# we could also do spearman per target and then plot an average of the spearman correlation.
+
+# Correlation between pairwise comparisons
+correlation_rel_skill <- scores |>
+  filter(horizon < 5) |>
+  summarise_scores(
+    by = c("model", "scale", "horizon", "target_type"), 
+    relative_skill = TRUE) |>
+  select(model, horizon, target_type, relative_skill, scale) |>
+  pivot_wider(values_from = relative_skill, names_from = scale) |>
+  group_by(horizon, target_type) |>
+  summarise(cor = cor(natural, log), 
+            spearman = cor(natural, log, method = "spearman"))
+
+p_cor_ranks <- correlation_ranks |>
+  ggplot(aes(x = horizon, y = cor, 
+             fill = `Target type`, color = `Target type`)) + 
+  geom_histogram(stat = "identity", position = "dodge") +
+  theme_scoringutils() + 
+  coord_cartesian(ylim = c(0, 1)) +
+  labs(y = "Corr. ranks per forecast", x = "Forecast horizon")
+
+# p_cor_scores <- cor_scores |> 
+#   mutate(`Target type` = target_type) |>
+#   ggplot(aes(x = horizon, y = cor, 
+#              fill = `Target type`, color = `Target type`)) + 
+#   geom_histogram(stat = "identity", position = "dodge") +
+#   theme_scoringutils() + 
+#   coord_cartesian(ylim = c(0, 1)) +
+#   labs(y = "Spearman corr. scores single forecasts", x = "Forecast horizon")
+
+p_cor_skill <- correlation_rel_skill |>
+  rename(`Target type` = target_type) |>
+  ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
+  geom_histogram(stat = "identity", position = "dodge") +
+  theme_scoringutils() + 
+  coord_cartesian(ylim = c(0, 1)) +
+  labs(y = "Correlation rel. skill", x = "Forecast horizon")
+
+combined <- scores[, c("model", "location", "target_end_date", "scale", "target_type", "interval_score", "horizon")] |>
+  full_join(hub_data[, c("location", "target_end_date", "target_type", "model", "true_value")]) |>
+  mutate(true_value = ifelse(scale == "natural", true_value, log(true_value + 1)))
+
+combined |>
+  filter(horizon <= 4) |>
+  rename(`Target type` = target_type) |>
+  group_by(horizon, scale, `Target type`) |>
+  summarise(cor = cor(interval_score, true_value)) |>
+    ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
+  geom_histogram(stat = "identity", position = "dodge") +
+  theme_scoringutils() + 
+  facet_wrap(~ scale) +
+  expand_limits(ylim = c(0, 1)) +
+  labs(y = "Correlation score and observation", x = "Forecast horizon")
+
+plot_cors <- p_cor_ranks  + p_cor_skill +
+  plot_layout(guides = "collect") +
+  plot_annotation(tag_levels = "A") &
+  theme(legend.position = "bottom")
+
+ggsave(filename = "output/figures/HUB-cors.png", width = 7, height = 3, 
+       plot = plot_cors)
+
+```
+
+## Figure 7 - Example comparisons of forecasts on the log and the natural scale
+
+```{r}
+# "CovidMetrics-epiBATS" only has forecasts in DE
+# "ICM-agentModel" only in Poland
+# "UVA is everywhere
+
+plot_pred_score <- function(hub_data, scores, model,
+                            horizon = 2, type = "Cases", 
+                            remove_x_text = TRUE) {
+  forecast_model <- model
+  h <- horizon
+  filtered_hub <- hub_data |>
+    filter(model == forecast_model, 
+           horizon %in% h, 
+           target_type == type)
+
+  locations <- unique(filtered_hub$location)
+
+  filtered_scores <- scores |>
+    filter(model == forecast_model, 
+           horizon %in% h, 
+           location %in% locations,
+           target_type == type)
+
+  plot_natural  <- filtered_hub |>  
+    mutate(id = paste(type, "- natural")) |>
+    plot_predictions(x = "target_end_date") +
+    scale_y_continuous(labels = label_fn)  +
+    facet_wrap(~ id) +
+    theme(legend.position = "bottom", 
+          axis.text.x = element_blank(), 
+          axis.title.x = element_blank()) +
+    labs(y = "Predictions")
+
+  plot_log <- filtered_hub |>
+    mutate(prediction = log(prediction + 1), 
+           true_value = log(true_value + 1)) |>
+    mutate(id = paste(type, "- log")) |>
+    plot_predictions(x = "target_end_date") +
+    scale_y_continuous(labels = label_fn)  +
+    facet_wrap(~ id) +
+    theme(legend.position = "bottom", 
+          axis.text.x = element_blank(), 
+          axis.title.x = element_blank()) +
+    labs(y = "Predictions")
+
+  scores_natural <- filtered_scores |>
+    filter(scale == "natural") |>
+    summarise_scores("target_end_date") |>
+    mutate(target_end_date = factor(target_end_date)) |>
+    plot_wis(x = "target_end_date", flip = TRUE) +
+    scale_x_continuous(labels = label_fn) +
+    labs(y = "Target end date")
+
+  scores_log <- filtered_scores |>
+    filter(scale == "log") |>
+    summarise_scores("target_end_date") |>
+    mutate(target_end_date = factor(target_end_date)) |>
+    plot_wis(x = "target_end_date", flip = TRUE) +
+    labs(y = "Target end date")
+
+  if (remove_x_text) {
+    scores_natural <- scores_natural +
+      theme(axis.text.x = element_blank(), 
+            axis.title.x = element_blank())
+    # scores_log <- scores_log +
+    #   theme(axis.text.x = element_blank(), 
+    #         axis.title.x = element_blank())
+  }
+
+  list(plot_natural, scores_natural, plot_log, scores_log )
+}
+
+# ICM <- plot_pred_score(hub_data, scores, "ICM-agentModel")
+# epiBATS <- plot_pred_score(hub_data, scores, "CovidMetrics-epiBATS")
+# BSLCoV <- plot_pred_score(hub_data, scores, "UB-BSLCoV")
+# ggsave(filename = "output/figures/HUB-model-ICM.png", width = 9, height = 5, plot = ICM)
+# ggsave(filename = "output/figures/HUB-model-epiBATS.png", width = 9, height = 5, plot = epiBATS)
+
+ensemble_cases <- plot_pred_score(hub_data |> filter(location == "DE"), 
+                                  scores, "EuroCOVIDhub-ensemble", 
+                                  remove_x_text = TRUE, 
+                                  type = "Cases")
+
+ensemble_deaths <- plot_pred_score(hub_data |> filter(location == "DE"), 
+                                   scores, "EuroCOVIDhub-ensemble", 
+                                   type = "Deaths")
+
+put_plot_together <- function(plot_cases, 
+                              plot_deaths) {
+  layout <- "
+AE
+BF
+CG
+DH"
+
+  plot_cases[[1]] + plot_cases[[2]] + plot_cases[[3]] +
+    plot_cases[[4]] + plot_deaths[[1]] +
+    plot_deaths[[2]] + plot_deaths[[3]] + plot_deaths[[4]] +
+    plot_annotation(tag_levels = "A") +
+    plot_layout(guides = "collect", 
+                design = layout) &
+    theme(legend.position = "bottom")
+}
+
+put_plot_together(ensemble_cases, ensemble_deaths)
+ggsave(filename = "output/figures/HUB-model-comparison-ensemble.png", width = 9, height = 7)
+
+# baseline <- plot_pred_score(hub_data |> filter(location == "DE"), scores, "EuroCOVIDhub-baseline") 
+# epinow <- plot_pred_score(hub_data |> filter(location == "DE"), scores, "epiforecasts-EpiNow2") 
+# 
+# ensemble / epinow +
+#   plot_layout(guides = "collect", 
+#               heights = c(1, 1, 2)) +
+#   theme(legend.position = "bottom")
+# 
+# ggsave(filename = "output/figures/HUB-model-comparison.png", width = 9, height = 7)
+
+```
+
+```{r}
+scores |>
+  summarise_scores(c("model", "target_type", "scale")) |>
+  summarise_scores(c("model", "target_type", "scale"), fun = signif, digits = 2) |>
+  plot_score_table(by = c("target_type", "scale")) +
+  facet_wrap(target_type ~ scale)
+```
+
+```{r}
+
+split_data <- split(scores, by = c("type_and_scale"))
+cors <- lapply(split_data, FUN = correlation, 
+               metrics = c("interval_score", 
+                            "bias", "overprediction", 
+                            "underprediction",
+                            "dispersion"))
+
+# scoringutils dev stuff
+# split_data <- split(scores, by = c("target_type", "scale"))
+# list_name_as_col <- function(list, by = NULL) {
+#   names <- names(list)
+#   print(names(list_element))
+# }
+
+plots <- lapply(cors, plot_correlation)
+
+(plots[[1]] + plots[[2]]) /
+  (plots[[3]] + plots[[4]])
+
+correlation
+```
+
+# Table with relevant values
+
+```{r}
+summary <- hub_data |>
+  filter(horizon == 2) |>
+  mutate(natural = true_value, 
+         log = log(pmax(0, true_value) +1)) |>
+  select(location, target_type, target_end_date, natural, log) |>
+  pivot_longer(cols = c(natural, log), values_to = "true_value", names_to = "scale") |>
+  unique() |>
+  group_by(target_type, scale) |>
+  summarise(mean = mean(true_value), 
+            sd = sd(true_value), 
+            var = var(true_value))
+
+table_obs <- summary |>
+  pivot_longer(cols = c(mean, sd, var), names_to = "measure") |>
+  pivot_wider(values_from = value, names_from = scale) |>
+  select(target_type, measure, natural, log) |>
+  mutate(quantity = "Observations", 
+         natural = round(natural), 
+         log = round(log, 2)) |>
+  select(target_type, quantity, measure, natural, log)
+
+table_scores <- scores |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  group_by(target_type, scale) |>
+  summarise(mean = mean(interval_score), 
+            sd = sd(interval_score), 
+            .groups = "drop_last") |>
+  pivot_longer(cols = c(mean, sd), names_to = "measure") |>
+  pivot_wider(values_from = value, names_from = scale) |>
+ mutate(quantity = "WIS", 
+         natural = round(natural), 
+         log = round(log, 2)) |>
+  select(target_type, quantity, measure, natural, log)
+
+table <- rbind(table_obs, 
+               table_scores) |>
+  arrange(quantity, target_type, measure)
+
+table |>
+  kable(format = "latex", booktabs = TRUE,
+        align = c("l", "l", "l", rep("c", 2)),
+        # col.names = c(" ", names(scores)[-(1:2)])) %>%
+  ) #|>
+  # collapse_rows(columns = 1) %>%
+  # kable_styling(latex_options = c("scale_down",
+  #                                 "hold_position")) %>%
+  # column_spec(1, background = "white") %>%
+  # pack_rows("Observations", 1, 6, indent = FALSE,
+  #           hline_after = TRUE) %>%
+  # pack_rows("WIS", 7, 10, indent = FALSE,
+  #           hline_after = TRUE)
+
+```
+
+--- 
+# Code parking lot
+
+## Scores over time
+restrict to a few locations, e.g. GB, FR, ES (highest average scores there)
+
+```{r}
+scores |> 
+  filter(model == "EuroCOVIDhub-ensemble",
+         location %in% c("GB", "ES", "DE"),
+         horizon == 2) |>
+  rename(Location = location) |>
+  group_by(target_type, scale, Location) |>
+  mutate(multiplicative_score = interval_score / min(interval_score)) |>
+  ggplot(aes(y = multiplicative_score, x = forecast_date, colour = Location)) + 
+  geom_line() + 
+  geom_point() + 
+  facet_grid(target_type ~ scale, scale = "free_y") + 
+  theme_scoringutils() + 
+  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
+
+ggsave("output/figures/HUB-scores-over-time.png", width = 7, height = 5)
+```
+
+*Overall it looks like there can still be substantial fluctuation in log scores, but maybe slightly less than when scored on an absolute scale.
+Interestingly, it seems like the smoothing effect is stronger for deaths than for cases.*
+
+```{r}
+# create version of data where the true value is the true growth rate
+# interesting question: does it make a difference, whether we transform it 
+# to be the weekly growth rate, i.e. take the n-th root, where n is the horizon?
+
+latest_truth <- fread(here("data", "weekly-truth-Europe.csv")) |>
+  rename(last_known_true_value = true_value) |>
+  mutate(forecast_date = target_end_date + 2) |>
+  select(location, target_type, forecast_date, last_known_true_value) |>
+  unique()
+
+# unsure whether the pmax hack is the right way to do it
+growth_data <- hub_data |>
+  inner_join(latest_truth) |>
+  mutate(true_value = true_value / pmax(last_known_true_value + 1),
+         prediction = prediction / pmax(last_known_true_value + 1)) |>
+  select(-last_known_true_value)
+```
+
+## Variance of scores vs. mean of scores across locations
+
+```{r}
+scores |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  group_by(target_type, location, scale) |>
+  summarise(mean_wis = mean(interval_score), 
+            sd_wis = sd(interval_score), 
+            .groups = "drop_last") |>
+  ggplot(aes(y = sd_wis, x = mean_wis)) + 
+  geom_abline(slope = 1, intercept = 0, linetype = "dashed", size = 0.5) + 
+  geom_point() + 
+  facet_wrap(~ target_type + scale, scale = "free") + 
+  theme_scoringutils() + 
+  theme(legend.position = "none") + 
+  # scale_x_continuous(labels = scale_fn) + 
+  # scale_y_continuous(labels = scale_fn) + 
+  scale_x_continuous(labels = scale_fn, trans = "log10") +
+  scale_y_continuous(labels = scale_fn, trans = "log10") +
+  labs(x = "Mean WIS", y = "Sd of WIS")
+
+ggsave("output/figures/HUB-sd-vs-mean-scores.png", width = 7, height = 5)
+```
+
+```{r}
+# negative binomial
+# MASS::glm.nb(true_value ~ target_type, 
+#        data = hub_data %>%
+#          mutate(true_value = pmax(true_value, 0)), 
+#        link = "identity")
+```
+
+## WIS vs. the totsl number of observed values
+probably not needed anymore, as we have similar things
+
+```{r}
+
+# wis_vs_total(scores) + 
+#   labs(y = "Mean WIS")
+# ggsave("output/figures/HUB-mean-scores-vs-total.png", width = 7, height = 5)
+# 
+# wis_vs_total(scores, summary_fct = stats::sd) + 
+#   labs(y = "Sd WIS")
+# ggsave("output/figures/HUB-sd-scores-vs-total.png", width = 7, height = 5)
+
+# wis_vs_total(scores, summary_fct = stats::sd) + 
+#   labs(y = "Sd WIS") + 
+#   scale_x_continuous(labels = scale_fn, trans = "log10") +
+#   scale_y_continuous(labels = scale_fn, trans = "log10") 
+# ggsave("output/figures/HUB-sd-scores-vs-total-log-log.png", width = 7, height = 5)
+
+```
+
+Test for a single example forecast to verify that rankings change
+```{r}
+# we see that rankings change even for a single forecast. So we need to check everything is still correct. 
+# take one example
+test_data <- hub_data |>
+  filter(model %in% c("MUNI-ARIMA", "epiforecasts-EpiNow2"), 
+         forecast_date == "2021-10-18", 
+         target_type == "Cases", 
+         location == "AT")
+
+test_data |>
+  score(metrics = c("interval_score")) |>
+  summarise_scores()
+
+test_data |>
+  mutate(prediction = log(prediction), 
+         true_value = log(true_value)) |>
+  score(metrics = c("interval_score")) |>
+  summarise_scores()
+## indeed everything seems to be correct, but rankings still change FOR A SINGLE FORECAST. 
+# It seems this is because the trade-off between calibration and sharpness changes? 
+# need to check in what way
+```
+
+## Distribution of scores
+
+```{r}
+scores |>
+  ggplot(aes(x = interval_score)) + 
+  geom_density() + 
+  facet_wrap(~ target_type + scale, scale = "free") + 
+  scale_x_continuous(labels = scale_fn)
+```
+
+## Fit a model to data
+
+Check fit of the log scale model
+```{r echo=TRUE}
+reg_log <- scores |>
+  filter(scale == "log") |>
+  glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 
+      family = gaussian)
+```
+
+```{r}
+# check distribution
+performance::check_distribution(reg_log) |>
+  plot()
+
+# check normality of residuals
+check_normality(reg_log) |>
+  plot(type = "qq")
+
+# heteroskedasticity of error terms - the plot takes extremly long to run
+performance::check_heteroskedasticity(reg_log)
+# performance::check_heteroskedasticity(reg_log) |>
+#   plot()
+
+# check influence of outliers - plot also takes very long to run
+performance::check_outliers(reg_log)
+# performance::check_outliers(reg_log) |>
+#   plot()
+
+```
+
+---
 
@@
 
-```{r}
-v1 <- c("a","b","c","d","e","f","g")
-v2 <- c("b","x","e","c","z","d","a")
-
-o <- 0.05
-DF <- data.frame(x = c(rep(1, length(v1)), rep(2, length(v2))),
-                 x1 = c(rep(1 + o, length(v1)), rep(2 - o, length(v2))),
-                 y = c(rev(seq_along(v1)), rev(seq_along(v2))),
-                 g = c(v1, v2))
-
-library(ggplot2)
-library(grid)
-ggplot(DF, aes(x=x, y=y, group=g, label=g)) +
-  geom_path(aes(x=x1), arrow = arrow(length = unit(0.02,"npc")), 
-            size=1, color="green") +
-  geom_text(size=10) +
-  theme_minimal() +
-  theme(axis.title = element_blank(),
-        axis.text = element_blank(),
-        axis.ticks = element_blank(),
-        panel.grid = element_blank())
-```
-
-## Look at pairwise comparisons
-
-```{r}
-summarised_pairwise <- scores |>
-  filter(target_type == "Cases") |>
-  summarise_scores(by = c("model", "scale"), relative_skill = TRUE)
-
-ranking_natural <- 
-  summarised_pairwise |>
-  filter(scale == "natural") |>
-  arrange(-relative_skill) |>
-  pull(model)
-
-plot_rel_skill <- summarised_pairwise |>
-  mutate(model = factor(model, levels = ranking_natural)) |>
-  ggplot(aes(y = model, x = relative_skill)) +
-  geom_bar(stat = "identity") +
-  theme_scoringutils() +
-  facet_wrap(~scale) +
-  labs(y = "Model", x = "Relative skill")
-
-diffs <- summarised_pairwise |>
-  mutate(model = factor(model, levels = ranking_natural)) |>
-  select(model, relative_skill, scale) |>
-  pivot_wider(names_from = scale, values_from = relative_skill) |>
-  mutate(model = factor(model, levels = ranking_natural), 
-         difference = log - natural)
-
-plot_diffs <- diffs |>
-  mutate(title = "difference") |>
-  ggplot(aes(y = model, x = difference)) +
-  geom_bar(stat = "identity") +
-  theme_scoringutils() +
-  labs(y = "Model", x = "Difference in relative skill") +
-  theme(axis.title.y = element_blank(), 
-        axis.text.y = element_blank(), 
-        axis.ticks.y = element_blank(), 
-        axis.line.y = element_blank()) +
-  facet_wrap(~title)
-
-plot_distr_scores <- scores |>
-  filter(horizon == 2) |>
-  mutate(model = factor(model, levels = ranking_natural)) |>
-  ggplot(aes(y = model, x = interval_score)) +
-  ggdist::stat_pointinterval() +
-  facet_wrap(~ scale, scale = "free_x") +
-  theme_scoringutils() +
-  scale_x_log10(labels = label_fn) +
-  theme(axis.title.y = element_blank(), 
-        axis.text.y = element_blank(), 
-        axis.ticks.y = element_blank(), 
-        axis.line.y = element_blank()) +
-  labs(y = "Model", x = "WIS")
-
-plot_rel_wis <- scores |>
-  filter(horizon == 2, 
-         target_type == "Cases") |>
-  mutate(model = factor(model, levels = ranking_natural)) |>
-  summarise_scores(by = c("model", "scale")) |>
-  plot_wis(relative_contributions = FALSE) +
-    theme(axis.title.y = element_blank(), 
-        axis.text.y = element_blank(), 
-        axis.ticks.y = element_blank(), 
-        axis.line.y = element_blank()) +
-  scale_x_continuous(labels = label_fn) +
-  facet_wrap(~scale, scales = "free_x")
-
-plot_rel_skill + plot_diffs + plot_rel_wis + #plot_distr_scores +
-  plot_layout(widths = c(2, 1, 2)) &
-  plot_layout(guides = "collect") &
-  theme(legend.position = "bottom")
-
-ggsave("output/figures/HUB-pairwise-comparisons.png", width = 9, height = 5)
-
-#Issues: this currently ignores that some models have only made forecasts for one location.
-They should usually be excluded, but maybe not necessary here as we're really more comparing the results for natural and log rather than the models against each other.
-```
-
-# Correlations for ranks, scores and pairwise comparisons
-
-```{r}
-
-# 1. observations natural vs. log --> pearson
-# 2. correlation scores natural vs. log (single forecasts) --> spearman
-
-# Table
-# correlation between values
-cor_data <- hub_data |>
-  select(location, target_end_date, target_type, true_value) |>
-  unique() |>
-  mutate(log_obs = log(pmax(true_value, 0) + 1)) |>
-  group_by(target_type) |>
-  summarise(cor = cor(true_value, log_obs))
-
-# # correlation between scores
-# # not really meaningful: if we do ranks it's just going to be ranks over everything which doesn't make sense
-# # if we do individual scores than it's a bit unclear what the "correct" correlation should be
-# cor_scores <- scores |>
-#   filter(horizon <=4) |>
-#   select(interval_score, scale, forecast_date, model, horizon, location, target_type) |>
-#   pivot_wider(names_from = scale, values_from = interval_score) |>
-#   group_by(horizon, target_type) |>
-#   summarise(cor = cor(natural, log), 
-#             spearman = cor(natural, log, method = "spearman"))
-
-# Correlation between ranks when going from natural to log
-correlation_ranks <- rankings_wide |>
-  filter(horizon < 5) |>
-  group_by(horizon, target_type) |>
-  summarise(cor = cor(natural, log), 
-            spearman = cor(natural, log, method = "spearman")) |>
-  rename(`Target type` = target_type)
-
-# Correlation between pairwise comparisons
-correlation_rel_skill <- scores |>
-  filter(horizon < 5) |>
-  summarise_scores(
-    by = c("model", "scale", "horizon", "target_type"), 
-    relative_skill = TRUE) |>
-  select(model, horizon, target_type, relative_skill, scale) |>
-  pivot_wider(values_from = relative_skill, names_from = scale) |>
-  group_by(horizon, target_type) |>
-  summarise(cor = cor(natural, log), 
-            spearman = cor(natural, log, method = "spearman"))
-
-p_cor_ranks <- correlation_ranks |>
-  ggplot(aes(x = horizon, y = cor, 
-             fill = `Target type`, color = `Target type`)) + 
-  geom_histogram(stat = "identity", position = "dodge") +
-  theme_scoringutils() + 
-  coord_cartesian(ylim = c(0, 1)) +
-  labs(y = "Corr. ranks single forecasts", x = "Forecast horizon")
-
-# p_cor_scores <- cor_scores |> 
-#   mutate(`Target type` = target_type) |>
-#   ggplot(aes(x = horizon, y = cor, 
-#              fill = `Target type`, color = `Target type`)) + 
-#   geom_histogram(stat = "identity", position = "dodge") +
-#   theme_scoringutils() + 
-#   coord_cartesian(ylim = c(0, 1)) +
-#   labs(y = "Spearman corr. scores single forecasts", x = "Forecast horizon")
-
-p_cor_skill <- correlation_rel_skill |>
-  rename(`Target type` = target_type) |>
-  ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
-  geom_histogram(stat = "identity", position = "dodge") +
-  theme_scoringutils() + 
-  coord_cartesian(ylim = c(0, 1)) +
-  labs(y = "Correlation rel. skill", x = "Forecast horizon")
-
-plot_cors <- p_cor_ranks  + p_cor_skill +
-  plot_layout(guides = "collect") +
-  plot_annotation(tag_levels = "A") &
-  theme(legend.position = "bottom")
-
-ggsave(filename = "output/figures/HUB-cors.png", width = 7, height = 3, 
-       plot = plot_cors)
-
-```
-
-Example comparisons of forecasts on the log and the natural scale
-
-```{r}
-# "CovidMetrics-epiBATS" only has forecasts in DE
-# "ICM-agentModel" only in Poland
-# "UVA is everywhere
-
-plot_pred_score <- function(hub_data, scores, model,
-                            horizon = 2, type = "Cases") {
-  forecast_model <- model
-  h <- horizon
-  filtered_hub <- hub_data |>
-    filter(model == forecast_model, 
-           horizon %in% h, 
-           target_type == type)
-
-  locations <- unique(filtered_hub$location)
-
-  filtered_scores <- scores |>
-    filter(model == forecast_model, 
-           horizon %in% h, 
-           location %in% locations,
-           target_type == type)
-
-  plot_natural  <- filtered_hub |>  
-    plot_predictions(x = "target_end_date") +
-    scale_y_continuous(labels = label_fn)  +
-    theme(legend.position = "bottom")
-
-  plot_log <- filtered_hub |>
-    mutate(prediction = log(prediction + 1), 
-           true_value = log(true_value + 1)) |>
-    plot_predictions(x = "target_end_date") +
-    scale_y_continuous(labels = label_fn)  +
-    theme(legend.position = "bottom")
-
-  scores_natural <- filtered_scores |>
-    filter(scale == "natural") |>
-    summarise_scores("target_end_date") |>
-    mutate(target_end_date = factor(target_end_date)) |>
-    plot_wis(x = "target_end_date", flip = TRUE) +
-    scale_x_continuous(labels = label_fn)
-
-  scores_log <- filtered_scores |>
-    filter(scale == "log") |>
-    summarise_scores("target_end_date") |>
-    mutate(target_end_date = factor(target_end_date)) |>
-    plot_wis(x = "target_end_date", flip = TRUE)
-
-  (plot_natural + plot_log) /
-    (scores_natural + scores_log) +
-    plot_layout(guides = "collect") &
-    theme(legend.position = "bottom")
-}
-
-ICM <- plot_pred_score(hub_data, scores, "ICM-agentModel")
-
-epiBATS <- plot_pred_score(hub_data, scores, "CovidMetrics-epiBATS")
-
-BSLCoV <- plot_pred_score(hub_data, scores, "UB-BSLCoV")
-
-ggsave(filename = "output/figures/HUB-model-ICM.png", width = 9, height = 5, plot = ICM)
-ggsave(filename = "output/figures/HUB-model-epiBATS.png", width = 9, height = 5, plot = epiBATS)
-ggsave(filename = "output/figures/HUB-model-BSLCoV.png", width = 9, height = 5, plot = BSLCoV)
-
-```
-
-# Table with relevant values
-
-```{r}
-summary <- hub_data |>
-  filter(horizon == 2) |>
-  mutate(natural = true_value, 
-         log = log(pmax(0, true_value) +1)) |>
-  select(location, target_type, target_end_date, natural, log) |>
-  pivot_longer(cols = c(natural, log), values_to = "true_value", names_to = "scale") |>
-  unique() |>
-  group_by(target_type, scale) |>
-  summarise(mean = mean(true_value), 
-            sd = sd(true_value), 
-            var = var(true_value))
-
-table_obs <- summary |>
-  pivot_longer(cols = c(mean, sd, var), names_to = "measure") |>
-  pivot_wider(values_from = value, names_from = scale) |>
-  select(target_type, measure, natural, log) |>
-  mutate(quantity = "Observations", 
-         natural = round(natural), 
-         log = round(log, 2)) |>
-  select(target_type, quantity, measure, natural, log)
-
-table_scores <- scores |> 
-  filter(model == "EuroCOVIDhub-ensemble", 
-         horizon == 2) |>
-  group_by(target_type, scale) |>
-  summarise(mean = mean(interval_score), 
-            sd = sd(interval_score), 
-            .groups = "drop_last") |>
-  pivot_longer(cols = c(mean, sd), names_to = "measure") |>
-  pivot_wider(values_from = value, names_from = scale) |>
- mutate(quantity = "WIS", 
-         natural = round(natural), 
-         log = round(log, 2)) |>
-  select(target_type, quantity, measure, natural, log)
-
-table <- rbind(table_obs, 
-               table_scores) |>
-  arrange(quantity, target_type, measure)
-
-table |>
-  kable(format = "latex", booktabs = TRUE,
-        align = c("l", "l", "l", rep("c", 2)),
-        # col.names = c(" ", names(scores)[-(1:2)])) %>%
-  ) #|>
-  # collapse_rows(columns = 1) %>%
-  # kable_styling(latex_options = c("scale_down",
-  #                                 "hold_position")) %>%
-  # column_spec(1, background = "white") %>%
-  # pack_rows("Observations", 1, 6, indent = FALSE,
-  #           hline_after = TRUE) %>%
-  # pack_rows("WIS", 7, 10, indent = FALSE,
-  #           hline_after = TRUE)
-
-```
-
---- 
-# Code parking lot
-
-## Scores over time
-restrict to a few locations, e.g. GB, FR, ES (highest average scores there)
-
-```{r}
-scores |> 
-  filter(model == "EuroCOVIDhub-ensemble",
-         location %in% c("GB", "ES", "DE"),
-         horizon == 2) |>
-  rename(Location = location) |>
-  group_by(target_type, scale, Location) |>
-  mutate(multiplicative_score = interval_score / min(interval_score)) |>
-  ggplot(aes(y = multiplicative_score, x = forecast_date, colour = Location)) + 
-  geom_line() + 
-  geom_point() + 
-  facet_grid(target_type ~ scale, scale = "free_y") + 
-  theme_scoringutils() + 
-  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
-
-ggsave("output/figures/HUB-scores-over-time.png", width = 7, height = 5)
-```
-
-*Overall it looks like there can still be substantial fluctuation in log scores, but maybe slightly less than when scored on an absolute scale.
-Interestingly, it seems like the smoothing effect is stronger for deaths than for cases.*
-
-```{r}
-# create version of data where the true value is the true growth rate
-# interesting question: does it make a difference, whether we transform it 
-# to be the weekly growth rate, i.e. take the n-th root, where n is the horizon?
-
-latest_truth <- fread(here("data", "weekly-truth-Europe.csv")) |>
-  rename(last_known_true_value = true_value) |>
-  mutate(forecast_date = target_end_date + 2) |>
-  select(location, target_type, forecast_date, last_known_true_value) |>
-  unique()
-
-# unsure whether the pmax hack is the right way to do it
-growth_data <- hub_data |>
-  inner_join(latest_truth) |>
-  mutate(true_value = true_value / pmax(last_known_true_value + 1),
-         prediction = prediction / pmax(last_known_true_value + 1)) |>
-  select(-last_known_true_value)
-```
-
-## Variance of scores vs. mean of scores across locations
-
-```{r}
-scores |> 
-  filter(model == "EuroCOVIDhub-ensemble", 
-         horizon == 2) |>
-  group_by(target_type, location, scale) |>
-  summarise(mean_wis = mean(interval_score), 
-            sd_wis = sd(interval_score), 
-            .groups = "drop_last") |>
-  ggplot(aes(y = sd_wis, x = mean_wis)) + 
-  geom_abline(slope = 1, intercept = 0, linetype = "dashed", size = 0.5) + 
-  geom_point() + 
-  facet_wrap(~ target_type + scale, scale = "free") + 
-  theme_scoringutils() + 
-  theme(legend.position = "none") + 
-  # scale_x_continuous(labels = scale_fn) + 
-  # scale_y_continuous(labels = scale_fn) + 
-  scale_x_continuous(labels = scale_fn, trans = "log10") +
-  scale_y_continuous(labels = scale_fn, trans = "log10") +
-  labs(x = "Mean WIS", y = "Sd of WIS")
-
-ggsave("output/figures/HUB-sd-vs-mean-scores.png", width = 7, height = 5)
-```
-
-```{r}
-# negative binomial
-# MASS::glm.nb(true_value ~ target_type, 
-#        data = hub_data %>%
-#          mutate(true_value = pmax(true_value, 0)), 
-#        link = "identity")
-```
-
-## WIS vs. the totsl number of observed values
-probably not needed anymore, as we have similar things
-
-```{r}
-
-# wis_vs_total(scores) + 
-#   labs(y = "Mean WIS")
-# ggsave("output/figures/HUB-mean-scores-vs-total.png", width = 7, height = 5)
-# 
-# wis_vs_total(scores, summary_fct = stats::sd) + 
-#   labs(y = "Sd WIS")
-# ggsave("output/figures/HUB-sd-scores-vs-total.png", width = 7, height = 5)
-
-# wis_vs_total(scores, summary_fct = stats::sd) + 
-#   labs(y = "Sd WIS") + 
-#   scale_x_continuous(labels = scale_fn, trans = "log10") +
-#   scale_y_continuous(labels = scale_fn, trans = "log10") 
-# ggsave("output/figures/HUB-sd-scores-vs-total-log-log.png", width = 7, height = 5)
-
-```
-
-Test for a single example forecast to verify that rankings change
-```{r}
-# we see that rankings change even for a single forecast. So we need to check everything is still correct. 
-# take one example
-test_data <- hub_data |>
-  filter(model %in% c("MUNI-ARIMA", "epiforecasts-EpiNow2"), 
-         forecast_date == "2021-10-18", 
-         target_type == "Cases", 
-         location == "AT")
-
-test_data |>
-  score(metrics = c("interval_score")) |>
-  summarise_scores()
-
-test_data |>
-  mutate(prediction = log(prediction), 
-         true_value = log(true_value)) |>
-  score(metrics = c("interval_score")) |>
-  summarise_scores()
-## indeed everything seems to be correct, but rankings still change FOR A SINGLE FORECAST. 
-# It seems this is because the trade-off between calibration and sharpness changes? 
-# need to check in what way
-```
-
-## Distribution of scores
-
-```{r}
-scores |>
-  ggplot(aes(x = interval_score)) + 
-  geom_density() + 
-  facet_wrap(~ target_type + scale, scale = "free") + 
-  scale_x_continuous(labels = scale_fn)
-```
-
-## Fit a model to data
-
-Check fit of the log scale model
-```{r echo=TRUE}
-reg_log <- scores |>
-  filter(scale == "log") |>
-  glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 
-      family = gaussian)
-```
-
-```{r}
-# check distribution
-performance::check_distribution(reg_log) |>
-  plot()
-
-# check normality of residuals
-check_normality(reg_log) |>
-  plot(type = "qq")
-
-# heteroskedasticity of error terms - the plot takes extremly long to run
-performance::check_heteroskedasticity(reg_log)
-# performance::check_heteroskedasticity(reg_log) |>
-#   plot()
-
-# check influence of outliers - plot also takes very long to run
-performance::check_outliers(reg_log)
-# performance::check_outliers(reg_log) |>
-#   plot()
-
-```
-
----
-
 # Growth Rate stuff - Looking at growth rates directly

### 2022-08-24 main.tex
update plots

@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-var-obs.png}
-    \caption{Observed mean against observed variance of observations in different locations across time.
-The dashed line marks what would be expecetd from a negative binomially distributed variable with variance of the observed value equal to the mean + the square of mean observed value.
-}
-    \label{fig:HUB-variance-mean}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-average-scores.png}
-    \caption{Distribution of interval scores for two week ahead forecasts of COVID-19 cases and deaths evaluated on the natural scale (left) and on the log scale (right). }
-    \label{fig:HUB-average-scores}
-\end{figure}
-
 SECTION STRUCUTRE / IMPORTANT THINGS
@@
 \item X inndividual rankings stay similar
-\item relative skill scores change a lot
+\item X relative skill scores change a lot
 \item deep dive into what happens for individual models
@@
 
-\paragraph{Mean / sd of cases and WIS values}
-Pooled across all time points and locations, the mean number of cases (deaths) observed was 22811 (368) with an overall standard deviation of 47135 (831) (see Table \ref{tab:HUB-summary}).
-The average number of observed cases and deaths varies considerably by location (see Figure \ref{fig:HUB-mean-locations}A and B).
-Observations were overdispersed, meaning that for a given location, the variance in observations exceeds the mean of all observed values (see Figure \ref{fig:HUB-variance-mean}).
-%Could fit a model var = alpha * x2 + beta *x
+\paragraph{Variation of scores gets smaller}
+The average number of observed cases and deaths varies considerably by location and target type (see Figure \ref{fig:HUB-mean-locations}A and B).
+On the natural scale, scores also vary considerably across targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}C).
+On the log scale, scores are more evenly distributed between targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}.
 
-On the natural scale, scores vary considerably across targets (see Fig \ref{}) and locations (see Fig \ref{}) and correlate strongly (XX) with the average number of observed cases / deaths in a given country (see Fig \ref{fig:HUB-mean-locations}A, B).
-On the log scale, scores are more evenly distributed between targets (see Fig \ref{}) and locations (see Fig \ref{fig:HUB-mean-locations}.
+Both on the natural scale as well as on the log scale, scores increase considerably with increasing forecast horizon (see Figure \ref{fig:HUB-scores-horizon}).
+The increase is more pronounced for cases than for deaths and arguably represents an increase in the relative difficulty to forecast values further into the future.
+A similar increase both on the natural and the log scale suggests that we are able to observe the increase in difficulty regardless of how we evaluate the scores.
+
+\paragraph{Correlation WIS - obs, location relationship}
+On the natural scale, scores correlate strongly (XX) with the average number of observed cases / deaths in a given country (see Fig \ref{fig:HUB-mean-locations}A, B).
 The correlation between average WIS and the mean number of observations in a country is much weaker and tends to be negative.
 Locations therefore tend to have a more equal influence on avearage scores, with smaller countries potentially being slightly overrepresented.
+
+For individual forecasts, rankings between models for single forecasts are mostly preserved (correlation: X for cases, Y for deaths), with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}).
+When evaluating performance across different forecasts and forecast targets, relative skill scores of the models change considerably.
+At least for cases, the effect is stronger with increasing forecast horizon.
+
+For relative skill scores, however, rankings do change a lot.
+The correlation between skill scores for two week ahead forecasts on the log and natural scale is XX.
+The correlation is stronger for deaths than for cases and tends to be lower for longer horizons.
+This might be because outliers are more likely for case forecasts than for death forecasts and also are more likely with increasing forecast horizon.
+
+\paragraph{Deep dive}
+Differences are likely driven by
+- on average, under-prediction plays a larger role (in relative terms), at least for cases. 
+- confidence intervals that go very low can produce artifacts --> baseline plot, not shown. 
+- natural scale punishes missing the peak very harshly. Log scale also punishes when you miss a new upswing. Arguably this is very good and potentially even closer to what we want.
+
+% Observations were overdispersed, meaning that for a given location, the variance in observations exceeds the mean of all observed values (see Figure \ref{fig:HUB-variance-mean}). 
+%Could fit a model var = alpha * x2 + beta *x
 %This is in line with the observation that the association between variance is smaller than mu^2 + mu, meaning that we would expect this.
@@
     \caption{Scores and average observed values across 1ocations.
-A, B: Mean of observed cases and deaths across all time points for different locations.
-C, D: XXX WIS for forecasts made by the Hub ensemble.
-D, E: Average WIS for log-transformed forecasts.
-}
+A, Mean of observed cases and deaths across all time points for different locations.
+B: XXX WIS for forecasts made by the Hub ensemble.
+C: Distribution of interval scores for two week ahead forecasts of COVID-19 cases and deaths evaluated on the natural scale (left) and on the log scale (right).}
     \label{fig:HUB-mean-locations}
@@
 
-\paragraph{individual rankings}
-For individual forecasts, rankings between models for single forecasts are mostly preserved (correlation: X for cases, Y for deaths), with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}).
-When evaluating performance across different forecasts and forecast targets, relative skill scores of the models change considerably.
-At least for cases, the effect is stronger with increasing forecast horizon.
-
 \begin{figure}[h!]
     \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-cors.png}
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-correlations.png}
     \caption{Correlation between scores on the natural and on the log scale.
@@
     \label{fig:HUB-cors}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-scores-vs-total-log-log.png}
-    \caption{My current thinking is that the top row is helpful, as it explalins what the mean / var relationship is which is relevant for how much weight small countries get.
-I think the 2nd row is just completely useless.
-The 3rd and 4th row may be helpful.
-In some sense it is duplicated information from above, but it gives some nice intuiton for the correlations.
-I could alternatively also turn it into a correlation plot and add it to the other one.
-}
-    \label{fig:HUB-mean-scores-total-loglog}
-\end{figure}
-
-Both on the natural scale as well as on the log scale, scores increase considerably with increasing forecast horizon (see Figure \ref{fig:HUB-scores-horizon}).
-The increase is more pronounced for cases than for deaths and arguably represents an increase in the relative difficulty to forecast values further into the future.
-A similar increase both on the natural and the log scale suggests that we are able to observe the increase in difficulty regardless of how we evaluate the scores.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-scores-over-horizon.png}
-    \caption{Weighted interval scores across forecast horizons}
-    \label{fig:HUB-scores-horizon}
 \end{figure}
@@
     \includegraphics[width=0.9\textwidth]{output/figures/HUB-pairwise-comparisons.png}
-    \caption{Changes in relative skill and the distribution of scores. MAYBE THIS SHOULD ONLY BE CASES?}
+    \caption{Changes in relative skill and the distribution of scores.Based on case forecasts only.}
     \label{fig:HUB-pairwise}
-\end{figure}
-
-% Correlation between pairwise comparisons is 0.4414717.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-model-ICM.png}
-    \caption{Top: 2-week-ahead predictions. Bottom: scores }
-    \label{fig:HUB-model-ICM}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-model-epiBATS.png}
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-model-comparison.png}
     \caption{Top: 2-week-ahead predictions. Bottom: scores }
-    \label{fig:HUB-model-epiBATS}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-model-BSLCoV.png}
-    \caption{Top: 2-week-ahead predictions. Bottom: scores }
-    \label{fig:HUB-model-BSLCoV}
+    \label{fig:HUB-model-comparison-ensemble}
 \end{figure}
@@
 \end{table}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-scores-vs-total-log-log.png}
+    \caption{My current thinking is that the top row is helpful, as it explalins what the mean / var relationship is which is relevant for how much weight small countries get.
+I think the 2nd row is just completely useless.
+The 3rd and 4th row may be helpful.
+In some sense it is duplicated information from above, but it gives some nice intuiton for the correlations.
+I could alternatively also turn it into a correlation plot and add it to the other one.
+}
+    \label{fig:HUB-mean-scores-total-loglog}
+\end{figure}
 

### 2022-08-30 analysis/Analysis ECDC forecasts.Rmd
update paper draft

@@
 # situations with negative values: France 2021-05-22, CZ 2021-08-07, ES 2021-03-06
-```
-
-```{r}
+
 # helper functions
@@
 
-```
-
-```{r}
 # load data 
@@
   unique()
+
+scores <- fread(here("output", "data", "all-scores-european-hub.csv"))
+scores[, scale := factor(scale, levels = c("natural", "log"))]
+data.table::setnames(scores, old = c("aem", "sharpness"), new = c("ae_median", "dispersion"))
+
+scores_gr <- fread(here("output", "data", "all-scores-growth-rate-european-hub.csv"))
+scores_gr[, scale := factor(scale, levels = c("natural", "log"))]
+
+check <- hub_data |> 
+  check_forecasts()
+
+filtermodels <- check$unique_values |>
+  filter(horizon >= 4, 
+         target_type == 2,
+         true_value > 900,
+         quantile == 23, 
+         forecast_date > 16,
+         location == 32) |>
+  pull(model)
+
+scores <- scores[model %in% filtermodels]
+scores[, type_and_scale := paste0(target_type, " - ", scale)]
+scores[, type_and_scale := factor(type_and_scale, 
+                                  levels = c("Cases - natural", "Deaths - natural",
+                                             "Cases - log", "Deaths - log"))]
+
+hub_data <- hub_data[model %in% filtermodels]
+hub_data[, true_value := pmax(true_value, 0)]
+
 ```
@@
 
-```{r}
-scores <- fread(here("output", "data", "all-scores-european-hub.csv"))
-scores[, scale := factor(scale, levels = c("natural", "log"))]
-data.table::setnames(scores, old = c("aem", "sharpness"), new = c("ae_median", "dispersion"))
-
-scores_gr <- fread(here("output", "data", "all-scores-growth-rate-european-hub.csv"))
-scores_gr[, scale := factor(scale, levels = c("natural", "log"))]
-```
-
-```{r}
-
-check <- hub_data |> 
-  check_forecasts()
-
-filtermodels <- check$unique_values |>
-  filter(horizon >= 4, 
-         target_type == 2,
-         true_value > 900,
-         quantile == 23, 
-         forecast_date > 16,
-         location == 32) |>
-  pull(model)
-
-scores <- scores[model %in% filtermodels]
-scores[, type_and_scale := paste0(target_type, " - ", scale)]
-scores[, type_and_scale := factor(type_and_scale, 
-                                  levels = c("Cases - natural", "Deaths - natural",
-                                             "Cases - log", "Deaths - log"))]
-
-hub_data <- hub_data[model %in% filtermodels]
-hub_data[, true_value := pmax(true_value, 0)]
-
-```
-
 General thoughts?
@@
   scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) + 
-  labs(y = "Mean observed value", x = "Location")
+  labs(y = "Mean observed value", x = "") +
+  theme(axis.title.x = element_blank())
 
@@
   scale_y_continuous(labels = scale_fn, trans = "log10") + 
-  labs(y = "Observations", x = "Target type")
+  labs(y = "Observations", x = "")  +
+  theme(axis.title.x = element_blank())
 
@@
          horizon == 2) |>
-  group_by(target_type, location, scale) |>
+  group_by(target_type, location, scale, type_and_scale) |>
   summarise(interval_score = mean(interval_score), 
@@
   scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
-  facet_wrap(scale ~ target_type, scale = "free") + 
+  facet_wrap(~ type_and_scale, scale = "free") + 
   theme_scoringutils() + 
@@
   scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) +
-  labs(y = "Mean interval score", x = "Location")
+  labs(y = "Mean interval score", x = "Loaction")
 
@@
   scale_y_continuous(labels = scale_fn, trans = "log10") +
-  labs(y = "Rel. increase in WIS", x = "Forecast horizon (weeks)")
+  labs(y = "Rel. change in WIS", x = "Forecast horizon (weeks)")
 
@@
 CD
-EE"
-
-plot_means_obs  + box_plot_obs + plot_mean_scores + box_plot_scores + box_plot_horizon +
+EF"
+
+plot_means_obs  + box_plot_obs + plot_mean_scores + box_plot_scores + box_plot_horizon + plot_spacer() +
   plot_layout(heights = c(1, 2, 1), 
@@
   coord_cartesian(ylim = c(0, 1)) +
-  labs(y = "Correlation rel. skill", x = "Forecast horizon")
+  labs(y = "Cor(rel. skill)", x = "Forecast horizon")
 
@@
 
-ggsave("output/figures/HUB-correlations.png", width = 7, height = 6)
+ggsave("output/figures/HUB-correlations.png", width = 7, height = 5.5)
 
@@
 ```{r}
+
+ranking_figure <- function(target = "Cases") {
   summarised_pairwise <- scores |>
-  filter(target_type == "Cases", 
+  filter(target_type == target, 
          horizon == 2) |>
@@
 
-ranking_natural <- 
+ranking_log <- 
   summarised_pairwise |>
-  filter(scale == "natural") |>
+  filter(scale == "log") |>
   arrange(-relative_skill) |>
@@
 
-plot_rel_skill <- summarised_pairwise |>
-  mutate(model = factor(model, levels = ranking_natural)) |>
-  ggplot(aes(y = model, x = relative_skill)) +
+plot_rel_skill_natural <- summarised_pairwise |>
+  filter(scale == "natural") |>
+  ggplot(aes(y = reorder(model, -relative_skill), x = relative_skill)) +
   geom_bar(stat = "identity") +
@@
 
+plot_rel_skill_log <- summarised_pairwise |>
+  filter(scale == "log") |>
+  ggplot(aes(y = reorder(model, -relative_skill), x = relative_skill)) +
+  geom_bar(stat = "identity") +
+  theme_scoringutils() +
+  facet_wrap(~scale) +
+  labs(y = "", x = "Relative skill") +
+  theme(axis.title.y = element_blank())
+
 diffs <- summarised_pairwise |>
-  mutate(model = factor(model, levels = ranking_natural)) |>
   select(model, relative_skill, scale) |>
   pivot_wider(names_from = scale, values_from = relative_skill) |>
-  mutate(model = factor(model, levels = ranking_natural), 
+  mutate(model = factor(model, levels = ranking_log), 
          difference = log - natural)
@@
   theme_scoringutils() +
-  labs(y = "Model", x = "Difference in relative skill") +
+  labs(y = "Model", x = "\u0394 Relative skill") +
   theme(axis.title.y = element_blank(), 
@@
 
-# plot_distr_scores <- scores |>
-#   filter(horizon == 2) |>
-#   mutate(model = factor(model, levels = ranking_natural)) |>
-#   ggplot(aes(y = model, x = interval_score)) +
-#   ggdist::stat_pointinterval() +
-#   facet_wrap(~ scale, scale = "free_x") +
-#   theme_scoringutils() +
-#   scale_x_log10(labels = label_fn) +
-#   theme(axis.title.y = element_blank(), 
-#         axis.text.y = element_blank(), 
-#         axis.ticks.y = element_blank(), 
-#         axis.line.y = element_blank()) +
-#   labs(y = "Model", x = "WIS")
-
 plot_rel_wis <- scores |>
@@
          target_type == "Cases") |>
-  mutate(model = factor(model, levels = ranking_natural)) |>
+  mutate(model = factor(model, levels = ranking_log)) |>
   summarise_scores(by = c("model", "scale")) |>
-  plot_wis(relative_contributions = FALSE) +
+  plot_wis(relative_contributions = TRUE) +
     theme(axis.title.y = element_blank(), 
@@
          x = ifelse(scale == "natural", 0.4, 2.8), 
-         x_arrow = ifelse(scale == "natural", 1.7, 2.7)) |>
+         x_arrow = ifelse(scale == "natural", 1.7, 2.7)
+         ) |>
   group_by(model) |>
@@
   ggplot(aes(y = rank, x = x, group = model, label = model)) +
-    geom_path(aes(x=x_arrow, color = color), arrow = arrow(length = unit(0.02,"npc")), 
+    geom_path(
+      aes(x=x_arrow, color = color), 
+      arrow = arrow(length = unit(0.09,"npc")), 
+      lineend = "round", linejoin = "mitre",
       size=1, show.legend = FALSE) +
-  geom_text(size=3.5, hjust = 0) +
-  coord_cartesian(xlim = c(0.5, 4)) +
+  # geom_text(size=3.3, hjust = 0) +
+  coord_cartesian(#xlim = c(0.5, 4), 
+                  ylim = c(0.8, 9.15)) +
   theme_scoringutils() +
@@
 
-layout <- "
-AAA
-BCD"
+# layout <- "
+# AAA
+# BCD"
+
+return(
+  list(
+    plot_ranking_change, plot_rel_skill_natural, plot_rel_skill_log, plot_diffs, plot_rel_wis
+  )
+)
 
@@
   theme(legend.position = "bottom")
-
-ggsave("output/figures/HUB-pairwise-comparisons.png", width = 9, height = 5)
+}
+
+ranking_cases <- ranking_figure(target = "Cases")
+ranking_deaths <- ranking_figure(target = "Deaths")
+
+layout <- "
+ABCDE
+FGHIJ"
+
+ranking_cases[[2]] + ranking_cases[[1]] + 
+ranking_cases[[3]] + ranking_cases[[4]] + ranking_cases[[5]] + 
+  ranking_deaths[[2]] + ranking_deaths[[1]] + 
+ranking_deaths[[3]] + ranking_deaths[[4]] + ranking_deaths[[5]] +
+  plot_annotation(tag_levels = "A") +
+  plot_layout(design = layout, 
+              guides = "collect") &
+  theme(legend.position = "bottom")
+
+ggsave("output/figures/HUB-pairwise-comparisons.png", width = 10.5, height = 6)
 
@@
 ggsave(filename = "output/figures/HUB-model-comparison-ensemble.png", width = 9, height = 7)
+
+ensemble_cases <- plot_pred_score(hub_data |> filter(location == "DE"), 
+                                  scores, "epiforecasts-EpiNow2", 
+                                  remove_x_text = TRUE, 
+                                  type = "Cases")
+
+ensemble_deaths <- plot_pred_score(hub_data |> filter(location == "DE"), 
+                                   scores, "EuroCOVIDhub-baseline", 
+                                   type = "Deaths")
 

### 2022-08-30 main.tex
update paper draft

@@
     \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/example-log-first.png}
+    \includegraphics[width=0.99\textwidth]{output/figures/example-log-first.png}
     \caption{Scores for different forecasts evaluated the WIS and the log WIS.
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/SIM-effect-log-score.png}
+    \includegraphics[width=0.99\textwidth]{output/figures/SIM-effect-log-score.png}
     \caption{Weighted interval score for a forecast distribution that is N(1, 0.2) and different observed values.
@@
 
-MAYBE MOVE TO SI
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/SIM-mean-state-size.png}
+    \caption{Simulation showing how the relationship between mean and variance of the forecast quantity affects scores.
+1,000 samples were drawn from different negative binomial distributions and scores were computed assuming an ideal forecaster (forecast distribution equal to data-generating distribution).
+Observations were simulated for different means of the negative-binomial distributions (representing, for example, the number of cases in differently sized states) as well as different relationships between mean and variance (representing, for example, three different infectious processes).
+The variance of the negative binomial is given as $\sigma^2 = \mu + \mu^2 / \theta$, meaning that for large theta the negative binomial distribution is equal to the poisson distribution.
+We used values of $\theta = 0.1$ (red), 1 (green) and 1b (blue).
+To make the scores for the different distributios comparable, scores were normalised to one, meaning that the mean score for every distribution (red, green, blue) is one.
+    A: Normalised WIS for ideal forecasts with increasing means of three distribution with different relationships between mean and variance.
+B: Same plot as A with both axes log-transformed.
+MAYBE CAN GET RID OF B.}
+    \label{fig:SIM-wis-state-size-mean}
+\end{figure}
+
+MAYBE MOVE TO SI. NEED TO ASK JOHANNES ABOUT IT. 
 For simulated negative binomial count data, log predictions for small quantities receive on average higher scores than forecasts for large quantities if the mean and the variance grow at the same rate $\sigma^2 = \mu$, about equal scores when the variance grows at a rate of $\sigma^2 = \mu + \mu^2$ (Figure \ref{fig:SIM-wis-state-size-mean}), and smaller scores when the variance grows faster than that.
@@
 
-I KNOW YOU ALL ARE GOING TO HATE THIS, BUT IT IS SOOOO CLEVER! :-(
-If one finds that average scores are dominated by forecasts for small quantities one could omit these forecasts entirely, arguing that relative errors are difficult to meaningfully interpret for small quantities.
-One might also take a more experimental route, namely adding a number $n$ (e.g.
-$n = 100$) to all observations and forecasts before taking the logarithm.
-This also constitutes a strictly monotonic transformation which preserves propriety and would reduce all scores (through reducing relative errors).
-Scores for forecasts on small quantities would be particularly affected, resulting in a more equal distribution of scores for forecasts of differing orders of magntitudes.
-Researchers could choose to add either a small or large number, depending on how much they want to equalise relative errors when aggregating scores.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/SIM-mean-sd-state-size.png}
-    \caption{Simulation of the effect of population size with ideal forecasts of a negative-binomially-distributed variable.
-For each simulated state, we drew 1,000 observations from a negative binomial distribution with $\mu = 100 \cdot \text{state size}$ and with values of $\theta$ equal to 0.1, 1, and 1 billion.
-The variance of the negative binomial is given as $\sigma^2 = \mu + \mu^2 / \theta$, meaning that for large theta the negative binomial distribution is equal to the poisson distribution.
-For these simulated values we computed the WIS for an ideal forecast (i.e.
-the predictive distribution was negative binomial with $\mu$ and $\theta$ equal to the true $\mu$ and $\theta$ for every state).
-Left: Mean WIS depending on state size, right: Mean WIS depending on state sizes when scored on a log scale.
-Plots for the standard deviation, rather than the mean of WIS values look are given in Figure \ref{fig:SIM-wis-state-size-sd} in the SI.}.
-    \label{fig:SIM-wis-state-size-mean}
-\end{figure}
+In the previous subsection we suggested to add a small number to all values in order to handle zeroes in the observations.
+While the choice of the number to add is somewhat arbitrary, it does have an influence on scores.
+By adding a number to forecasts as well as observations, measures of relative errors are slighly altered (while still preserving propriety of the score as we're applying a strictly monotonic transformation).
+Especially for forecasts of small values, the choice of the number that is added may affect scores quite strongly, as relative errors shrhink most for small quantities when adding a number.
+In principle, it is also possible to add a large number (e.g.
+100 or 10,000) to all observations and forecasts in order to equalise relative errors between forecasts of different quantities if one finds that average scores are dominated too strongly by forecasts for small quantities.
 
@@
 
-\paragraph{Intro Hub}
 As an empirical example for evaluating forecasts on the natural and on the log scale we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
@@
 
-SECTION STRUCUTRE / IMPORTANT THINGS
-\begin{itemize}
-\item X variation gets smaller / more even
-\item X relationship with country size changes
-\item X inndividual rankings stay similar
-\item X relative skill scores change a lot
-\item deep dive into what happens for individual models
-\item scores increase across forecast horizons
-\end{itemize}
+% SECTION STRUCUTRE / IMPORTANT THINGS
+% \begin{itemize}
+% \item X variation gets smaller / more even
+% \item X relationship with country size changes
+% \item X inndividual rankings stay similar
+% \item X relative skill scores change a lot
+% \item deep dive into what happens for individual models
+% \item scores increase across forecast horizons
+% \end{itemize}
+% 
+% Maybe look at correlation between bias and score?
+% Correlation between relative WIS contributions.?
 
-Maybe look at correlation between bias and score?
-Correlation between relative WIS contributions.?
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/HUB-mean-obs-location.png}
+    \caption{Observations and scores across locations and forecast horizons.
+A, Average (across all time points) of observed cases and deaths for different locations.
+B: Boxplot (y-axis on log-scale) of all cases and deaths.
+C: Scores for two-week-ahead forecasts from the EuroCOVIDhub-ensemble (averaged across all forecast dates) for different locations, evaluated on the natural as well as the logarithmic scale.
+D: Corresponding box-plots of all individual scores for two-week-ahead predictions.
+E: Boxplots for the relative change of scores for the EuroCOVIDhub-ensemble across forecast horizons.
+For any given forecast date and location, forecasts were made for four different forecast horizons, resulting in four scores.
+All scores were divided by the score for forecast horizon one.}
+    \label{fig:HUB-mean-locations}
+\end{figure}
 
-\paragraph{Variation of scores gets smaller}
-The average number of observed cases and deaths varies considerably by location and target type (see Figure \ref{fig:HUB-mean-locations}A and B).
-On the natural scale, scores also vary considerably across targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}C).
-On the log scale, scores are more evenly distributed between targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}.
+Across the dataset, the average number of observed cases and deaths varied considerably by location and target type (see Figure \ref{fig:HUB-mean-locations}A and B).
+On the natural scale, scores also varied considerably across targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}C).
+On the log scale, scores were more evenly distributed between targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}.
+Both on the natural scale as well as on the log scale, scores increased considerably with increasing forecast horizon (see Figure \ref{fig:HUB-scores-horizon}).
+The increase was more pronounced for cases than for deaths.
+We hypothesise that this increase in scores across horizons represents the relative difficulty to forecast values into the future, which was noticeable both on the natural as well as the logarithmic scale.
 
-Both on the natural scale as well as on the log scale, scores increase considerably with increasing forecast horizon (see Figure \ref{fig:HUB-scores-horizon}).
-The increase is more pronounced for cases than for deaths and arguably represents an increase in the relative difficulty to forecast values further into the future.
-A similar increase both on the natural and the log scale suggests that we are able to observe the increase in difficulty regardless of how we evaluate the scores.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/HUB-correlations.png}
+    \caption{Correlations between observations and scores.
+A: Mean WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against the average number of observations in a location.
+B: Average Spearman (rank-) correlation of scores for individual forecasts from \textit{all} models.
+For every individual target (defined by a combination of forecast date, target type, horizon, location), one score was obtained per model.
+Then, the rank correlation was computed between the scores for all models on the natural scale vs.
+on the log scale.
+All rank correlations were averaged and stratified by horizon and target type.
+C: Correlation between relative skill scores.
+For every forecast horizon and target type, a separate rel.
+skill score was computed per model using pairwise comparisons.
+The plot shows the correlation between the rel.
+skill scores on the natural vs.
+on the log scale.}
+    \label{fig:HUB-cors}
+\end{figure}
 
-\paragraph{Correlation WIS - obs, location relationship}
 On the natural scale, scores correlate strongly (XX) with the average number of observed cases / deaths in a given country (see Fig \ref{fig:HUB-mean-locations}A, B).
-The correlation between average WIS and the mean number of observations in a country is much weaker and tends to be negative.
-Locations therefore tend to have a more equal influence on avearage scores, with smaller countries potentially being slightly overrepresented.
+On the log scale, the correlation between average WIS and the mean number of observations in a country is much weaker and if anything tends to be slightly negative (giving higher weight to locations with few observations when averaging scores).
 
-For individual forecasts, rankings between models for single forecasts are mostly preserved (correlation: X for cases, Y for deaths), with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}).
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/HUB-model-comparison-ensemble.png}
+    \caption{
+    Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-ensemble made in Germany.
+A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale.
+B, F: Corresponding scores.
+C, G: Forecasts and observations on the log scale.
+D, H: Corresponding scores.
+    }
+    \label{fig:HUB-model-comparison-ensemble}
+\end{figure}
+
+For individual forecasts, rankings between models for single forecasts are mostly preserved (average rank correlation for single forecasts: X for cases, Y for deaths), with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}).
 When evaluating performance across different forecasts and forecast targets, relative skill scores of the models change considerably.
-At least for cases, the effect is stronger with increasing forecast horizon.
+The change is stronger for cases than for deaths (Figure \ref{fig:HUB-cors}XX) and tends to increase with increasing forecast horizon.
 
-For relative skill scores, however, rankings do change a lot.
-The correlation between skill scores for two week ahead forecasts on the log and natural scale is XX.
-The correlation is stronger for deaths than for cases and tends to be lower for longer horizons.
-This might be because outliers are more likely for case forecasts than for death forecasts and also are more likely with increasing forecast horizon.
+When comparing examples of forecasts on the natural scale with log-transformed ones (see Figures XX, XX, XX) a few interesting patterns emerge.
+Missing the peak, i.e.
+predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the log scale (see forecasts in May in Figure XXA, B).
+On the log scale, these have less of an influence, as errors are smaller in relative terms (see panels C and D).
+We hypothesise that the larger divergence in scores for cases (compared to scores for deaths) displayed in Figure \ref{fig:HUB-cors}C is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths.
+Conversely, failure to predict an upswing while numbers are still low, is not severely punished on the natural scale (see forecasts in July in Figure XX A, B), as overall absolute errors are low.
+On the log scale, missing lower inflection points tends to lead to more severe penalties (see panels C and D).
+One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure XXE, F).
+On the log scale, scores increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.
+Furthermore, including very low numbers in the confidence intervals may lead to very high scores on the log-scale.
+This may be an issue especially when forecasting small quantities.
+Figure XXX shows forecasts and scores for the baseline model, which had zero included in its 50 percent intervals for the forecasts in XX to XX, leading to excessive dispersion values on the log scale.
+Notwithstanding the general concern, one can argue that in the specific example including zero in the prediction intervals was unreasonable and rightly penalised given that case numbers were over 1000.
 
-\paragraph{Deep dive}
-Differences are likely driven by
-- on average, under-prediction plays a larger role (in relative terms), at least for cases. 
-- confidence intervals that go very low can produce artifacts --> baseline plot, not shown. 
-- natural scale punishes missing the peak very harshly. Log scale also punishes when you miss a new upswing. Arguably this is very good and potentially even closer to what we want.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/HUB-pairwise-comparisons.png}
+    \caption{Changes in model ratings as measured by relative skill for two-week-ahead predictions. Top row: cases, bottom row: deaths. NEED TO CHANGE THE ANNOTATION LEVELS.}
+    \label{fig:HUB-pairwise}
+\end{figure}
+
+\paragraph{SAM}
+
+COULD MAYBE ADD SOME ANALYSIS ABOUT WHAT KINDS OF MODELS BENEFITTED AND WHICH ONES GOT PENALISED MORE.
+AT THE MOMENT WE HAVE A VERY NICE FIGURE THAT ISN'T REALLY REFERENCED IN THE TEXT.
+SAM TO THE RESCUE!
 
@@
 %This is in line with the observation that the association between variance is smaller than mu^2 + mu, meaning that we would expect this.
-
- % Pooled across all time points and locations, the mean WIS on the natural scale for cases (deaths) was 8855 (59.3) with an overall standard deviation of 51274 (156) (see Table \ref{tab:HUB-summary}).
-On the log scale, scores were much closer to each other (see Figure \ref{fig:HUB-average-scores}), with a mean WIS for cases (deaths) of 0.507 (0.362) with an overall standard deviation of 0.745 (0.407).
-
-% When evaluated on the natural scale, there is  strong relationship between the mean of the weighted interval score and the total number of observed cases (correlation: 0.913) or deaths (cor: 0.849) in a location (see Figure \ref{fig:HUB-mean-scores-total-loglog} and Figure \ref{fig:HUB-mean-scores-total} in the SI).
-The relationship is less pronounced for scores on the log scale (cor: -0.019 for cases, -0.395 for deaths) and negative, meaning that locations with fewer observed cases or deaths tend to receive higher scores.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-obs-location.png}
-    \caption{Scores and average observed values across 1ocations.
-A, Mean of observed cases and deaths across all time points for different locations.
-B: XXX WIS for forecasts made by the Hub ensemble.
-C: Distribution of interval scores for two week ahead forecasts of COVID-19 cases and deaths evaluated on the natural scale (left) and on the log scale (right).}
-    \label{fig:HUB-mean-locations}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-correlations.png}
-    \caption{Correlation between scores on the natural and on the log scale.
-A: Correlation between relative ranks (the rank in terms of WIS divided by the number of models) for individual forecasts, separated by forecast horizon.
-B: Correlation between relative skill values achieved by models across locations and time points, separated by forecast horizon.
-SHOULD THIS BE SPEARMAN?}
-    \label{fig:HUB-cors}
-\end{figure}
-
-% \paragraph{Rankings}
-% So one effect is that for a given dispersion on the natural scale, it's better if that interval is skewed, as that implies that at least you only off a lot in one direction in negative terms, and not in the other direction.
-Maybe looking at over-and underprediction is the key here.
-%ADD SAME PLOT TO SI FOR HORIZON 4.
-
-\paragraph{pairwise comparisons}
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-pairwise-comparisons.png}
-    \caption{Changes in relative skill and the distribution of scores.Based on case forecasts only.}
-    \label{fig:HUB-pairwise}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-model-comparison.png}
-    \caption{Top: 2-week-ahead predictions. Bottom: scores }
-    \label{fig:HUB-model-comparison-ensemble}
-\end{figure}
 
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-scores-vs-total-log-log.png}
+    \includegraphics[width=0.99\textwidth]{output/figures/HUB-mean-scores-vs-total-log-log.png}
     \caption{My current thinking is that the top row is helpful, as it explalins what the mean / var relationship is which is relevant for how much weight small countries get.
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{output/placeholder-image.png}
+    \includegraphics[width=0.99\textwidth]{output/placeholder-image.png}
     \caption{Placeholder}
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-scores-vs-total.png}
+    \includegraphics[width=0.99\textwidth]{output/figures/HUB-mean-scores-vs-total.png}
     \caption{Plot with Weighted interval scores against the mean number of observed cases or deaths.}
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{output/figures/SIM-sd-state-size.png}
-    \caption{Simulation of the effect of population size with ideal forecasts of a negative-binomially-distributed variable.
-For each simulated state, we drew 1,000 observations from a negative binomial distribution with $\mu = 100 \cdot \text{state size}$ and with values of $\theta$ equal to 0.1, 1, and 1 billion.
-The variance of the negative binomial is given as $\sigma^2 = \mu + \mu^2 / \theta$, meaning that for large theta the negative binomial distribution is equal to the poisson distribution.
-For these simulated values we computed the WIS for an ideal forecast (i.e.
-the predictive distribution was negative binomial with $\mu$ and $\theta$ equal to the true $\mu$ and $\theta$ for every state).
-Left: Mean WIS depending on state size, right: Mean WIS depending on state sizes when scored on a log scale.
-Plots for the standard deviation, rather than the mean of WIS values look are given in Figure \ref{fig:SIM-wis-state-size-sd} in the SI.}.
-    \label{fig:SIM-wis-state-size-mean}
-\end{figure}
-
+\clearpage
 \bibliography{log-or-not.bib}
@@
 %    \centering
-%    \includegraphics[width=0.9\textwidth]{output/figures/HUB-sd-vs-mean-scores.png}
+%    \includegraphics[width=0.99\textwidth]{output/figures/HUB-sd-vs-mean-scores.png}
 %    \caption{CAPTION}
@@
 
-%\paragraph{Optional: Application to data from the Euro-Hub}
 %Also: what, empirically is the distribution of errors?
+
+ % Pooled across all time points and locations, the mean WIS on the natural scale for cases (deaths) was 8855 (59.3) with an overall standard deviation of 51274 (156) (see Table \ref{tab:HUB-summary}).
+On the log scale, scores were much closer to each other (see Figure \ref{fig:HUB-average-scores}), with a mean WIS for cases (deaths) of 0.507 (0.362) with an overall standard deviation of 0.745 (0.407).
+
+% When evaluated on the natural scale, there is  strong relationship between the mean of the weighted interval score and the total number of observed cases (correlation: 0.913) or deaths (cor: 0.849) in a location (see Figure \ref{fig:HUB-mean-scores-total-loglog} and Figure \ref{fig:HUB-mean-scores-total} in the SI).
+The relationship is less pronounced for scores on the log scale (cor: -0.019 for cases, -0.395 for deaths) and negative, meaning that locations with fewer observed cases or deaths tend to receive higher scores.

### 2022-08-30 analysis/Simulation-poisson-mean-sd.R

@@ -38,24 +38,35 @@ score_states <- function(df) {
   return(scores)
 }
 
+label_fn <- function(x) {
+  return(100*x)
+}
+
 make_plot <- function(scores, summary_fct = mean) {
   p1 <- scores |>
     group_by(state_size, scale, Theta) |>
     summarise(interval_score = summary_fct(interval_score)) |>
     group_by(Theta, scale) |>
     mutate(interval_score = interval_score / mean(interval_score), 
+           Variance = as.factor(as.character(Theta)),
+           Variance = recode_factor(Variance, 
+             "0.1" = "\u03C3 = \u03BC + 10 \u00B7 \u03BC^2", 
+             "1" = "\u03C3 = \u03BC + \u03BC^2", 
+             "1e+09" = "\u03C3 = \u03BC"),
            Theta = ifelse(Theta == "1e+09", "1b", Theta)) |>
-    ggplot(aes(y = interval_score, x = state_size, colour = Theta)) +
+    ggplot(aes(y = interval_score, x = state_size, colour = Variance)) +
     geom_point(size = 0.4) +  
-    labs(y = "WIS", x = "Size of state") + 
-    theme_minimal() + 
+    labs(y = "WIS", x = "Mean") +
+    scale_x_continuous(labels = label_fn) +
+    theme_scoringutils() + 
     facet_wrap(~ scale, scales = "free_y")
   
   p2 <- p1 + 
     scale_x_continuous(trans = "log10") + 
     scale_y_continuous(trans = "log10")
   
-  p1 / p2
+  p1 / p2 +
+    plot_annotation(tag_levels = "A")
 }
 
 
@@ -68,21 +79,21 @@ mean_county <- 100
 
 sizes_nbinom <- c(0.1, 1, 1e9)
 
-res <- list()
-for (size_nbinom in sizes_nbinom) {
-  df <- setup_df(time_points = 1000) |>
-    mutate(true_value = rnbinom(n = 1, size = size_nbinom, 
-                                mu = mean_county * state_size), 
-           prediction = qnbinom(p = quantile, size = size_nbinom, 
-                                mu = mean_county * state_size)) 
-  
-  scores <- score_states(df) |>
-    mutate(Theta = as.character(size_nbinom))
-  
-  res[[paste(size_nbinom)]] <- scores
-}
+# res <- list()
+# for (size_nbinom in sizes_nbinom) {
+#   df <- setup_df(time_points = 1000) |>
[…]

### 2022-08-30 analysis/example-log-first-Johannes.R

@@ -51,6 +51,8 @@ summary |>
   geom_point() + 
   facet_wrap(~type, scales = "free_y") + 
   theme_scoringutils() +
-  labs(y = "Score", x = "Standard deviation of predictive distribution")
+  labs(y = "Score", x = "Standard deviation of predictive distribution") + 
+  theme(panel.spacing = unit(1, "lines"))
 
-ggsave("output/figures/example-log-first.png", width = 8, height = 2.5)
+ggsave("output/figures/example-log-first.png", width = 7, height = 2.1)
+

### 2022-08-30 analysis/illustration-effect-log-score.R

@@ -53,7 +53,8 @@ p1 <- scores |>
   labs(y = "WIS", x = "Observed value") + 
   scale_y_continuous(label = function(x) {paste(scale_factor * x)}) + 
   facet_wrap(~ scale, ncol = 1, scales = "free") + 
-  theme_scoringutils()
+  theme_scoringutils() + 
+  theme(panel.spacing = unit(1, "lines"))
 
 scale_factor2 <- 3
 p2_log <- scores |>
@@ -72,10 +73,11 @@ p2_log <- scores |>
                      }, 
                      breaks = c(0.2, 1/2, 1, 2, 5)) + 
   facet_wrap(~ scale, ncol = 1, scales = "free") + 
-  theme_scoringutils()
+  theme_scoringutils() + 
+  theme(panel.spacing = unit(1, "lines"))
 
 p1 + p2_log
 
-ggsave("output/figures/SIM-effect-log-score.png", width = 6, 
-       height = 4)
+ggsave("output/figures/SIM-effect-log-score.png", width = 7, 
+       height = 3)
 

### 2022-08-30 analysis/Analysis ECDC forecasts.Rmd
update paper draft

@@
 ```{r}
-# "CovidMetrics-epiBATS" only has forecasts in DE
-# "ICM-agentModel" only in Poland
-# "UVA is everywhere
-
 plot_pred_score <- function(hub_data, scores, model,
@@
 
-# ICM <- plot_pred_score(hub_data, scores, "ICM-agentModel")
-# epiBATS <- plot_pred_score(hub_data, scores, "CovidMetrics-epiBATS")
-# BSLCoV <- plot_pred_score(hub_data, scores, "UB-BSLCoV")
-# ggsave(filename = "output/figures/HUB-model-ICM.png", width = 9, height = 5, plot = ICM)
-# ggsave(filename = "output/figures/HUB-model-epiBATS.png", width = 9, height = 5, plot = epiBATS)
-
-ensemble_cases <- plot_pred_score(hub_data |> filter(location == "DE"), 
-                                  scores, "EuroCOVIDhub-ensemble", 
+put_plot_together <- function(modelname = "EuroCOVIDhub-ensemble", 
+                              locationname = "DE") {
+
+  plot_cases <- plot_pred_score(hub_data |> filter(location == locationname), 
+                                  scores, modelname, 
                                   remove_x_text = TRUE, 
@@
 
-ensemble_deaths <- plot_pred_score(hub_data |> filter(location == "DE"), 
-                                   scores, "EuroCOVIDhub-ensemble", 
+  plot_deaths <- plot_pred_score(hub_data |> filter(location == locationname), 
+                                     scores, modelname, 
                                      type = "Deaths")
 
-put_plot_together <- function(plot_cases, 
-                              plot_deaths) {
   layout <- "
@@
 
-put_plot_together(ensemble_cases, ensemble_deaths)
+put_plot_together("EuroCOVIDhub-ensemble")
 ggsave(filename = "output/figures/HUB-model-comparison-ensemble.png", width = 9, height = 7)
 
-ensemble_cases <- plot_pred_score(hub_data |> filter(location == "DE"), 
-                                  scores, "epiforecasts-EpiNow2", 
-                                  remove_x_text = TRUE, 
-                                  type = "Cases")
-
-ensemble_deaths <- plot_pred_score(hub_data |> filter(location == "DE"), 
-                                   scores, "EuroCOVIDhub-baseline", 
-                                   type = "Deaths")
-
-# baseline <- plot_pred_score(hub_data |> filter(location == "DE"), scores, "EuroCOVIDhub-baseline") 
-# epinow <- plot_pred_score(hub_data |> filter(location == "DE"), scores, "epiforecasts-EpiNow2") 
-# 
-# ensemble / epinow +
-#   plot_layout(guides = "collect", 
-#               heights = c(1, 1, 2)) +
-#   theme(legend.position = "bottom")
-# 
-# ggsave(filename = "output/figures/HUB-model-comparison.png", width = 9, height = 7)
+put_plot_together("epiforecasts-EpiNow2")
+ggsave(filename = "output/figures/HUB-model-comparison-epinow.png", width = 9, height = 7)
+
+put_plot_together("EuroCOVIDhub-baseline", locationname = "DE")
+ggsave(filename = "output/figures/HUB-model-comparison-baseline.png", width = 9, height = 7)
 

### 2022-08-30 main.tex
update paper draft

@@
 % Maybe look at correlation between bias and score?
-% Correlation between relative WIS contributions.?
+% Correlation between relative WIS contributions on the log vs. the natural scale.?
 
@@
 This may be an issue especially when forecasting small quantities.
-Figure XXX shows forecasts and scores for the baseline model, which had zero included in its 50 percent intervals for the forecasts in XX to XX, leading to excessive dispersion values on the log scale.
-Notwithstanding the general concern, one can argue that in the specific example including zero in the prediction intervals was unreasonable and rightly penalised given that case numbers were over 1000.
+Figure \ref{fig:HUB-model-comparison-baseline} shows forecasts and scores for the baseline model, which had zero included in its 50 percent intervals for the forecasts in XX to XX, leading to excessive dispersion values on the log scale.
+Notwithstanding the general concern, one can argue that in the specific example (at least for cases) including zero in the prediction intervals was unreasonable and rightly penalised given that case numbers were in the thousands.
 
@@
 
-\subsection{Scoring the growth rate directly}
-By dividing every observed value and every forecast by the last observed value prior to evaluation, we replace forecasting $y_{t+1}$ with forecasting  $y_{t+1} / y_t = 1 + g_{t, t+1}$.
-Just as we were previously penalised an additive error with respect the absolute value, we are now penalising an absolute error with respect to the growth rate.
-
-We can illustrate what's happening by thinking of an analogous point prediction $\hat{g_{t, t+1}}$ for the growth rate. We could then write
-\begin{equation}
-    1 + g_{t, t+1} = 1 + \hat{g}_{t, t+1} + \varepsilon^*_{t+1}. 
-\end{equation}
-%
-The WIS and CRPS would evaluate the forecast based on the error $\varepsilon^*_{t+1}$.
-This translates to an absolute error on the actual forecast target that is multiplicatively connected to the last observed value:
-\begin{align}
-    y_{t+1} &= (1 + \hat{g}_{t, t+1} + \varepsilon^*_{t+1}) \cdot y_t \\
-            &= (1 + \hat{g}_{t, t+1}) \cdot y_t + \varepsilon^*_{t+1} \cdot y_t \\
-            &= \hat{y}_{t+1} + \varepsilon^*_{t+1} \cdot y_t.
-\end{align}
-
-One disadvantage of this is that we can't really do anything when an observation is zero. 
-%is this helpful?%
-%Maybe add some math on how we actually transform the distribution. E.g. divide every quantile or sample by the last observed value, divide the distribution?%
-
-\subsection{Changes in the WIS for forecasts on the log scale} \label{wis-log-derivation}
+\subsection{PROBABLY DELETE? Changes in the WIS for forecasts on the log scale} \label{wis-log-derivation}
 
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/placeholder-image.png}
-    \caption{Placeholder}
-    \label{fig:placeholder}
+    \includegraphics[width=0.99\textwidth]{output/figures/HUB-model-comparison-baseline.png}
+    \caption{
+    Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-ensemble made in Germany.
+A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale.
+B, F: Corresponding scores.
+C, G: Forecasts and observations on the log scale.
+D, H: Corresponding scores.
+    }
+    \label{fig:HUB-model-comparison-baseline}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-mean-scores-vs-total.png}
-    \caption{Plot with Weighted interval scores against the mean number of observed cases or deaths.}
-    \label{fig:HUB-mean-scores-total}
+    \includegraphics[width=0.99\textwidth]{output/figures/HUB-model-comparison-epinow.png}
+    \caption{
+    Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-ensemble made in Germany.
+A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale.
+B, F: Corresponding scores.
+C, G: Forecasts and observations on the log scale.
+D, H: Corresponding scores.
+    }
+    \label{fig:HUB-model-comparison-epinow}
 \end{figure}
@@
 
-%\section{Modelling scores}
-%\paragraph{Motivation} Alternative is to use summary measures or pairwise comparison, which becomes cumbersome for many dimensions.
-
-%\paragraph{caveats, practical limitations}
-%\begin{itemize}
-%    \item What kind of transformations can we do? Propose random effect models
-%    \item is there anything we can't do? 
-%    \item How does logging change the error distribution?
-On the natural scale, you have huge outliers.
-What's the appropriate error distribution to use?
-confidence intervals.
-Maybe we can only say something about the effects
-%\end{itemize}
-
-%Also: what, empirically is the distribution of errors?
-
  % Pooled across all time points and locations, the mean WIS on the natural scale for cases (deaths) was 8855 (59.3) with an overall standard deviation of 51274 (156) (see Table \ref{tab:HUB-summary}).

### 2022-11-15 main.tex
Addressing co-author comments

@@
 
-Probabilistic forecasts play an important role in decision-making in epidemiology and public health \citep{reichCollaborativeMultiyearMultimodel2019, funkShorttermForecastsInform2020, cramerEvaluationIndividualEnsemble2021, bracherShorttermForecastingCOVID192021, europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}, as well as other areas as diverse as for example economics \citep{timmermannForecastingMethodsFinance2018, elliottForecastingEconomicsFinance2016} or meteorology \citep{gneitingWeatherForecastingEnsemble2005, kukkonenReviewOperationalRegionalscale2012}.
-Epidemiological modelling in particular has received widespread attention during the COVID-19 pandemic and has been an important tool to inform public policy.
-Evaluating forecasts is important, as the evaluation provides feedback for forecasters and researchers necessary to improve their models, and it helps decision makers distinguish good from bad predictions and choosing forecasts that inform future decisions.
+Probabilistic forecasts play an important role in decision-making in epidemiology and public health \citep{reichCollaborativeMultiyearMultimodel2019, funkShorttermForecastsInform2020, cramerEvaluationIndividualEnsemble2021, bracherShorttermForecastingCOVID192021, europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}, as well as other areas as diverse as economics \citep{timmermannForecastingMethodsFinance2018, elliottForecastingEconomicsFinance2016} or meteorology \citep{gneitingWeatherForecastingEnsemble2005, kukkonenReviewOperationalRegionalscale2012}.
+Epidemiological modelling in particular has received widespread attention during the COVID-19 pandemic, during which it has demonstrated the critical role it can play in informing public policy.
+Forecast evaluation is important, as it provides feedback for forecasters and researchers to improve their models, and it helps decision makers distinguish good from bad predictions and choosing forecasts that will inform future decisions.
 
-Proabbilistic forecasts are usually evaluated using so-called proper scoring rules, which return a numerical score as a function of the forecast and the later observed data.
-Forecasters, in expectaction, optimise their score by providing a predictive distribution that is equal to the data-generating distribution, and will be penalised if they fail to do so.
-Two proper scoring rules commonly used in epidemiology and other fields are the Continuous Ranked Probability Score (CRPS) \citep{gneitingStrictlyProperScoring2007} and the Weighted Interval Score (WIS) \citep{bracherEvaluatingEpidemicForecasts2021}.
-Both the WIS and CRPS represent a measure of the absolute distance between the predictive distribution and observation and represent a generalisation of the absolute error to full predictive distributions (see details in the SI).
-The WIS can be thought of as an approxmiation of the CRPS for forecasts in a quantile format and has very similar properties (the WIS converges towards the CRPS for an increasing set of equally-spaced prediction intervals).
-Due to its widespread use in epidemiology (and its similarity to the CRPS) we mostly focus on the WIS in this paper, ignoring other proper scoring rules.
-Formulas for the WIS are given in the SI.
+Probabilistic forecasts are usually evaluated using so-called proper scoring rules, which return a numerical score as a function of the forecast and the later observed data.
+Forecasters, in expectation, optimise their score by providing a predictive distribution that is equal to the data-generating distribution, and will be penalised if they fail to do so.
+This paper focuses on one particular proper scoring rule, the Weighted Interval Score (WIS, formulas given in Section XXX in the SI) \citep{bracherEvaluatingEpidemicForecasts2021}, which is commonly used in epidemiology and other fields.
+The WIS can be understood as an approximation of the Continuous Ranked Probability Score (CRPS) \citep{gneitingStrictlyProperScoring2007} for predictive distributions represented by a set of predictive quantiles.
+The WIS and the CRPS measure the absolute distance between the predictive distribution and the observed value and can be understood as a generalisation of the absolute error to full predictive distributions (see details in section XX in the SI).
+For simplicity, we will use "absolute error" and "absolute distance" interchangeably, even though the former applies only to point forecasts and the latter would be more appropriate for predictive distributions.
+The WIS has been used by several Forecast Hubs for COVID-19 in the US [CITATION], Europe [CItation], Germany and Poland [CITATION], as well as the US Flu Forecasting Hub and the Scenario Hubs in US and Europe and XX.
 
-% Maybe insert WIS formula here?
+Evaluating forecasts based on the absolute distance between forecast and observed value may not always be optimal in an epidemiological context.
+Epidemiological processes are commonly understood and modeled to be inherently exponential in nature, as every infectious individual is assumed to lead to some number of additional infections [CITATION].
+It would therefore seem more appropriate to focus the evaluation on errors on the multiplicative growth rate, rather than on absolute errors on the observed values, to better represent the underlying disease dynamics.
+Evaluating forecasts based on absolute errors in practice also means that scores are strongly influenced by the overall order of magnitude of the quantity to forecast.
+This makes it difficult compare predictive performance across time, given that epidemiological processes may be non-stationary, or to compare performance across targets that are on very different orders of magnitudes (for example cases and hospitalisations) [CITE KATH ECDC, and others].
 
-Evaluating forecasts based on absolute errors (for simplicity we use the term interchangeably with absolute distance, even though the former technically only applies to point forecasts), as is done by the WIS, may not always be desirable, especially in an epidemiological context.
-% The reason for this is that generally, the way for ecasts are evaluated should correspond to the error structure of the model and the type of errors models will usually make.
-Epidemiological processes are usually understood to be exponential in nature and therefore are most commonly described and modeled in multiplicative terms.
-Under the assumption that these models are correctly capturing the underlying inherent transmission dynamics, it would therefore make sense to focus evaluation on multiplicative, rather than absolute errors.
-Often, if the model is performing badly in multiplicative terms, this means that there is a flaw with its assumptions that needs to be resolved, regardless of performance in absolute terms.
-Evaluating epidemiological forecasts based on a measure of absolute error (as is common practice) can make it harder to diagnose and quantify model mis-specification.
-It also may make it more difficult to deal with the potential non-stationarity of epidemiological processes, or to compare performance across targets that are on very different orders of magnitudes (like for example cases and hospitalisations).
-
-In this paper we argue that it may therefore be helpful to shift the focus of the evaluation on multiplicative, rather than absolute errors.
-This is approximately equal to scoring the growth rate of a forecast and can be achieved by transforming the forecasts as well as the observations using the logarithm before applying the WIS or CRPS.
-We focus on the log transformation, as it represents a very convenient way of achieving our goal of evaluating how well a forecast captures the growth dynamic inherent to epidemiological process.
-However, we note that the same logic we use can also applied to other transformations which could shift the focus of evaluation in different ways.
-For example, given an appropriate transformation, it could also be possible to evaluate forecasters based on the shape of the trajectory they predicted.
-It would even be possible to construct composite scores as a weighted sum of scores based on different transformations, allowing forecast consumers to assign explicit weights to different qualities of the forecasts.
-
-In section [REF] we discuss the broader idea of transforming forecasts, and in section [REF] give a detailed theoretical of the log transformation and its interpretation.
-In section [REF] we present an applied example using forecasts submitted to the European COVID-19 Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
-The European COVID-19 Forecast Hub is one of several Forecast hubs (together with a Hub in the US \citep{cramerEvaluationIndividualEnsemble2021} and one in Germany and Poland \citep{bracherShorttermForecastingCOVID192021} that systematically collected, aggregated and evaluated forecasts of several COVID-19 targets created by different teams.
-Forecasts submitted to the COVID-19 Forecast Hubs are probabilistic in nature [Citation e.g.
-Held et al.], meaning that forecasters provide a full predictive distribution which quantifies their uncertainty.
-These predictive distributions are reported in the form of a set of 23 quantiles (11 symmetric prediction intervals as well as a median prediction).
-%Currently, the US Flu Forecasting Hub and COVID-19 Scenario Hubs in the US and in Europe use the same format.
+In this paper, we discuss how log transforming the forecasts and observations prior to applying the WIS (or CRPS) may be an appropriate way to address the aforementioned issues with evaluating forecasts based on absolute errors.
+We first outline the motivation for transforming forecasts prior to scoring (section [REF]).
+We then provide a mathematical intuition of the impact of scoring on the log scale and its interpretation (section [REF]).
+We illustrate our findings on a series of examples using forecasts submitted to the European COVID-19 Forecast Hub  \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
+Finally, we give scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues on the use of transformed scores in the evaluation of infectious disease forecasts.
 
@@
 
+Log transforming forecasts and observations shifts the focus of the evaluation while preserving propriety of the score.
+Instead of scoring the magnitude of absolute errors this approach captures relative errors and is approximately equivalent to scoring a forecast for the growth rate of the observed quantity.
+
+We focus on the log transformation, as it represents a convenient way of achieving our goal of evaluating how well a forecast captures the growth dynamic inherent to epidemiological process.
+
 Applying the log transformation, among other transformations, is permissible because it is a strictly monotonic transformation.
+
 While rankings between different models may indeed change, even for a single forecast, the relevant condition for a scoring rule to be strictly proper is still satisfied: In expectation a forecaster will receive the best score if she reports a predictive distribution that is equal to the data-generating distribution, even after transforming both the forecasts and the observations (representing draws from the data-generating distribution).
@@
 In an epidemiological setting that involves count data, negative values should likely be omitted entirely.
-Values of zero should potentially also be removed from the analsys, as it does not make sense to interpret relative errors on zero counts.
-Alternatively, if one is not willing to omit them, a small quantity (such as 1) could be added to all observations before taking the logarithm.
-Note that predictive distribtions have to be truncated in a similar way.
+One potential way to deal values of zero is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm as is common practice when log transforming counts.
 When dealing with count data we could extend this argument to all small values (rather than just zeros), as it is hard to reliably evaluate relative errors when dealing with small discrete values.
-One could therefore decide to remove forecasts for small quantities below a certain threshold from the analysis.
+An alternative solution is to remove forcasts for small, relative, values from ongoing analysis.
+In the following sections we will explore the implications of both of these strategies.
 
-\subsection{Scoring multplicative errors}
+\subsection{Scoring multiplicative errors}
 
@@
 This approximation breaks down when the difference between the predicted and the actual growth rate becomes too large (albeit forecasts are still evaluated in terms of relative errors).
-If this is of great concern one may want to use a different transformation like dividing forecasts and observations by the last known values instead in order to score the growth rate explicitly (although this lacks the elegance of having one single transformation per time point).
+If this is of great concern one may want to use a different transformation such as dividing forecasts and observations by the last known values in order to score the growth rate explicitly (although this lacks the elegance of having one single transformation per time point).
 
@@
 Forecaster A would receive a score of $2 - 0.5$ = 1.5, beating forecaster B who would receive a score of $2.5 - 0.9 = 1.6$.
-When we transform the forecast, forecaster A receives a score of $\log (\frac{2}{0.5}) = 1.39$, while forecaster B obtains a score of $\log (\frac{2.6}{0.9} = 1.06$ and now wins over forecaster A.
+When we transform the forecast, forecaster A receives a score of $\log (\frac{2}{0.5}) = 1.39$, while forecaster B obtains a score of $\log (\frac{2.6}{0.9}) = 1.06$ and now wins over forecaster A.
 The main difference between forecaster A and B is that forecaster A has issued a prediction interval ([0.5, 2]) that was more symmetric around the true value than the interval issued by forecaster B ([0.9.
@@
 This effect is not mainly driven by the kinds of changes in model rankings for single forecasts we just discussed.
-Rather, aggregate scores across multiple forecasts are affected by the fact some scores will in general be much larger than others.
-On the natural scale, aggregate scores will usually be dominated by predictions of quantities with high absolute values.
-Even an ideal forecaster will make much larger errors when predicting a quantity such as reported case numbers, than when predicting hospitalisations.
+Rather, as observations from exponential processes are likely to be on very different absolute scales, large observations will account for the majority of absolute score but will not dominate relative scores in the same way.
+When aggregating scores across metrices with different scales, for example reported cases, and hospitalisations even an ideal forecaster will make much larger errors when predicting a quantity such as reported case numbers, than when predicting hospitalisations.
 Evaluating forecasts based on relative errors may help with comparing performance across different targets.
@@
 Observations were simulated for different means of the negative-binomial distributions (representing, for example, the number of cases in differently sized states) as well as different relationships between mean and variance (representing, for example, three different infectious processes).
-The variance of the negative binomial is given as $\sigma^2 = \mu + \mu^2 / \theta$, meaning that for large theta the negative binomial distribution is equal to the poisson distribution.
+The variance of the negative binomial is given as $\sigma^2 = \mu + \mu^2 / \theta$, meaning that for large theta the negative binomial distribution is equal to the Poisson distribution.
 We used values of $\theta = 0.1$ (red), 1 (green) and 1b (blue).
-To make the scores for the different distributios comparable, scores were normalised to one, meaning that the mean score for every distribution (red, green, blue) is one.
-    A: Normalised WIS for ideal forecasts with increasing means of three distribution with different relationships between mean and variance.
-B: Same plot as A with both axes log-transformed.
-MAYBE CAN GET RID OF B.}
+To make the scores for the different distributions comparable, scores were normalised to one, meaning that the mean score for every distribution (red, green, blue) is one.
+    A: Normalised WIS for ideal forecasts with increasing means of three distribution with different relationships between mean and variance. B: A but with log scale axis.}
     \label{fig:SIM-wis-state-size-mean}
@@
 
-MAYBE MOVE TO SI. NEED TO ASK JOHANNES ABOUT IT. 
 For simulated negative binomial count data, log predictions for small quantities receive on average higher scores than forecasts for large quantities if the mean and the variance grow at the same rate $\sigma^2 = \mu$, about equal scores when the variance grows at a rate of $\sigma^2 = \mu + \mu^2$ (Figure \ref{fig:SIM-wis-state-size-mean}), and smaller scores when the variance grows faster than that.
@@
 While the choice of the number to add is somewhat arbitrary, it does have an influence on scores.
-By adding a number to forecasts as well as observations, measures of relative errors are slighly altered (while still preserving propriety of the score as we're applying a strictly monotonic transformation).
-Especially for forecasts of small values, the choice of the number that is added may affect scores quite strongly, as relative errors shrhink most for small quantities when adding a number.
+By adding a number to forecasts as well as observations, measures of relative errors are slightly altered (while still preserving propriety of the score as we're applying a strictly monotonic transformation).
+Especially for forecasts of small values, the choice of the number that is added may affect scores quite strongly, as relative errors shrink most for small quantities when adding a number.
 In principle, it is also possible to add a large number (e.g.
@@
 \section{Empirical example: the European Forecast Hub}
+
+PUT SOMEWHERE
+The European COVID-19 Forecast Hub is one of several Forecast hubs (together with a Hub in the US \citep{cramerEvaluationIndividualEnsemble2021} and one in Germany and Poland \citep{bracherShorttermForecastingCOVID192021} that systematically collected, aggregated and evaluated forecasts of several COVID-19 targets created by different teams.
+Forecasts submitted to the COVID-19 Forecast Hubs are probabilistic in nature [Citation e.g.
+Held et al.], meaning that forecasters provide a full predictive distribution which quantifies their uncertainty.
+These predictive distributions are reported in the form of a set of 23 quantiles (11 symmetric prediction intervals as well as a median prediction).
 
@@
 On the natural scale, scores also varied considerably across targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}C).
-On the log scale, scores were more evenly distributed between targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}.
+On the log scale, scores were more evenly distributed between targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}).
 Both on the natural scale as well as on the log scale, scores increased considerably with increasing forecast horizon (see Figure \ref{fig:HUB-scores-horizon}).
@@
 
-On the natural scale, scores correlate strongly (XX) with the average number of observed cases / deaths in a given country (see Fig \ref{fig:HUB-mean-locations}A, B).
-On the log scale, the correlation between average WIS and the mean number of observations in a country is much weaker and if anything tends to be slightly negative (giving higher weight to locations with few observations when averaging scores).
+As expected on the natural scale, scores correlate strongly (XX) with the average number of observed cases / deaths in a given country (see Fig \ref{fig:HUB-mean-locations}A, B) and this relationship is not evident on the log scale.
+Interestingly, there is some evidence a negative relationship between the mean number of observations in a country and score when evaluated on a log scale implying that countries with lower mean incidence may be harder to forecast.
 
@@
 Missing the peak, i.e.
-predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the log scale (see forecasts in May in Figure XXA, B).
+predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts in May in Figure XXA, B).
 On the log scale, these have less of an influence, as errors are smaller in relative terms (see panels C and D).
@@
 
-\paragraph{SAM}
-
-COULD MAYBE ADD SOME ANALYSIS ABOUT WHAT KINDS OF MODELS BENEFITTED AND WHICH ONES GOT PENALISED MORE.
-AT THE MOMENT WE HAVE A VERY NICE FIGURE THAT ISN'T REALLY REFERENCED IN THE TEXT.
-SAM TO THE RESCUE!
+Evaluating the rank order of forecasts from the ECDC forecast hub we saw that scoring after a log transform altered the order of forecast models as expected based on our simulated case study.
+In particular, we note that models that are known to not make use of a exponential error model and to be biased towards under-prediction, such the the MUNI-ARMA model were ranked lower when using a log transform.
+A submission by the authors of this study, the epiforecasts-EpiNow2 forecast model, was the most impacted by the change to log scale scoring moving from last place for reported cases to 4th place.
+This makes sense given that it explicitly models errors on an exponential scale, and is relatively unbiased meaning it may be over-penalised due to over-prediction during periods of peak incidence.
+Interestingly, other models that are growth rate based, such as the LANL-GrowthRate forecaster, do less well after a log-transform.
+Encouragingly for the ECDC forecast Hub the Hub ensemble, which is the forecast the organisers suggest forecast consumers make use of, remains one of the top forecasts across scoring schemes and indeed ranks higher for reported cases when a log transform is used.
 
@@
 
-\paragraph{Other transformations - maybe move to Discussion}
+\paragraph{Other transformations}
 While we focus on the log transformation in this paper, there are other transformations that may be helpful in an epidemiological context.
@@
 
-COULD ALSO BE DISCUSSION
-Whether or not a log transformation is apprpriate depends on whether one believes that differences in absolute values are meaningful.
+Whether or not a log transformation is appropriate depends on whether one believes that differences in absolute values are meaningful.
 This may be the case for example when comparing forecasts across locations or time points, rather than different forecasts targets.
@@
 % could make an analysis re outliers: how do scores change if we remove the 1 or 2 worst forecasts? The one or two best forecasts?
+
+NEEDS TO GO SOMEWHERE
+However, we note that the same logic we use can also applied to other transformations which could shift the focus of evaluation in different ways.
+For example, given an appropriate transformation, it could also be possible to evaluate forecasters based on the shape of the trajectory [CITATION https://arxiv.org/abs/2209.04035] they predicted.
+It would even be possible to construct composite scores as a weighted sum of scores based on different transformations, allowing forecast consumers to assign explicit weights to different qualities of the forecasts.
 

### 2022-11-15 main.tex
Addressing comments before larger restructuring

@@
 
-Log transforming forecasts and observations shifts the focus of the evaluation while preserving propriety of the score.
-Instead of scoring the magnitude of absolute errors this approach captures relative errors and is approximately equivalent to scoring a forecast for the growth rate of the observed quantity.
+The goal of log transforming forecasts and observations is to shift the focus of the evaluation while preserving the propriety of the score.
+Instead of a score representing the magnitude of absolute errors this approach allows to obtain a score which measures relative error and provides an approximate measure for how well a forecast captures the multiplicative growth rate of the target quantity.
+In this paper we focus on the log transformation, but note that many other transformations may be useful and appropriate (see section XXXX) .
 
-We focus on the log transformation, as it represents a convenient way of achieving our goal of evaluating how well a forecast captures the growth dynamic inherent to epidemiological process.
-
-Applying the log transformation, among other transformations, is permissible because it is a strictly monotonic transformation.
-
-While rankings between different models may indeed change, even for a single forecast, the relevant condition for a scoring rule to be strictly proper is still satisfied: In expectation a forecaster will receive the best score if she reports a predictive distribution that is equal to the data-generating distribution, even after transforming both the forecasts and the observations (representing draws from the data-generating distribution).
-
-It is important to first apply the transformation and then score the transformed forecasts and observations.
-Simply taking the log of the WIS or the CRPS computed on the natural scale, i.e.
-computing $\log S(x, y)$ results in an improper score and is therefore not advised.
-This is because the logarithm of the WIS / CRPS does not penalise outliers enough and therefore incentivises overconfident predictions.
+The log transformation is permissible (in the sense that it preserves propriety of the score), because it is a strictly monotonic transformation.
+Even though actual rankings between models may change, a forecaster will in expectation still minimise her score if she reports a predictive distribution that is equal to the data-generating distribution.
+The order of the operations matters, and applying the log transformation after scores have been computed, i.e.
+computing $\log S(x, y)$ results in an improper score.
+This is because taking the logarithm of the WIS / CRPS results in a score that does not penalise outliers enough and therefore incentivises overconfident predictions.
 We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can easily be seen that overconfident models perform best in terms of the log WIS.
@@
 Rather, as observations from exponential processes are likely to be on very different absolute scales, large observations will account for the majority of absolute score but will not dominate relative scores in the same way.
-When aggregating scores across metrices with different scales, for example reported cases, and hospitalisations even an ideal forecaster will make much larger errors when predicting a quantity such as reported case numbers, than when predicting hospitalisations.
+When aggregating scores across metrics with different scales, for example reported cases, and hospitalisations, even an ideal (i.e.
+perfect) forecaster will make much larger errors when predicting case numbers than when predicting hospitalisations.
 Evaluating forecasts based on relative errors may help with comparing performance across different targets.
@@
 This is especially true when forecasts are restricted to discrete values.
-Whether or not forecast targets with smaller quantities receive higher scores than those for targets with larger quantities, depends on the exact relationship between the mean and the variance of the quantity of interest.
 
@@
 \end{figure}
+
+Whether or not forecast targets with smaller quantities receive higher scores than those for targets with larger quantities, depends on the exact relationship between the mean and the variance of the quantity of interest.
 
@@
 
-As expected on the natural scale, scores correlate strongly (XX) with the average number of observed cases / deaths in a given country (see Fig \ref{fig:HUB-mean-locations}A, B) and this relationship is not evident on the log scale.
-Interestingly, there is some evidence a negative relationship between the mean number of observations in a country and score when evaluated on a log scale implying that countries with lower mean incidence may be harder to forecast.
+As expected on the natural scale, scores correlate strongly (XX) with the average number of observed cases / deaths in a given country (see Fig \ref{fig:HUB-mean-locations}A, B).
+This relationship is not evident on the log scale, where correlation is much weaker and maybe slightly negative, implying larger scores for for countries with lower mean incidences.
 

### 2022-11-21 analysis/Analysis ECDC forecasts.Rmd
update correlation plot, add plot with number of available forecasts

@@
 library(grid)
+library(ggpubr)
 
@@
 
+remove <- hub_data |> 
+  group_by(model, forecast_date, location, target_type) |>
+  summarise(q = length(unique(quantile))) |>
+  filter(q < 23) |>
+  pull(model) |>
+  unique()
+
+filtermodels <- setdiff(filtermodels, remove)
+
 scores <- scores[model %in% filtermodels]
@@
 hub_data <- hub_data[model %in% filtermodels]
+hub_data <- hub_data[forecast_date >= "2021-03-05"]
 hub_data[, true_value := pmax(true_value, 0)]
@@
 
-General thoughts?
-
-- Focus on Hub-ensemble or average across all models? 
-- Focus on two weeks ahead?
+## Basic data summary
+```{r}
+hub_data |>
+  summarise(min_date = min(forecast_date), 
+            max_date = max(forecast_date)) |>
+  mutate(week_diff = (max_date - min_date - 1) / 7)
+
+hub_data$location |> 
+  unique() |>
+  length()
+
+hub_data |> 
+  filter(horizon == 2) |>
+  avail_forecasts(by = c("target_end_date", "target_type", "model")) |>
+  mutate(forecast_date = as.Date(target_end_date) - 12) |>
+  plot_avail_forecasts(show_numbers = FALSE) +
+  facet_grid(~ target_type) + 
+  labs(y = "Model", x = "Forecast date")
+
+ggsave(filename = "output/figures/number-avail-forecasts.png", height = 4, width = 10)
+```
 
@@
 
-Figure
-Currently discontinuoued and just there to discuss it. 
-that shows Variance vs. mean of the observations. Makes the point that the
-data looks neg. binomial. 
-```{r}
-# plot with mean vs. variance --------------------------------------------------
-plot_var_mean <- function(hub_data, log = FALSE) {
-  df <- hub_data |>
-    mutate("x" = true_value) |>
-    dplyr::select(location, target_type, target_end_date, true_value, x) |>
-    unique() |>
-    group_by(location, target_type)
-
-  f <- function(x) {x + x^2}
-
-  if (log) {
-    df <- mutate(df, true_value = log(pmax(true_value, 0) + 1))
-
-    f <- function(x) {log(log(x) + log(x)^2 + 1)}
-  }
-
-  out <- df |>
-  summarise(mean_value = mean(x), 
-            sd_value = sd(true_value), 
-            var_value = var(true_value)) |>
-  ggplot(aes(y = var_value, x = mean_value)) +
-  geom_point() + 
-  facet_wrap(~ target_type, scales = "free") + 
-  scale_x_continuous(trans = "log10", labels = label_fn) +
-  scale_y_continuous(trans = "log10", labels = label_fn) +
-  labs(y = "Var observations", x = "Mean observations") + 
-  # geom_function(fun = f, linetype = "dashed", size = 0.5) +
-  theme_scoringutils() + 
-  geom_smooth(alpha = 0.2, size = 0.2)
-
-  if (log) {
-    out <- out +
-      labs(y = "Var log(observations)", x = "Mean observations")
-  }
-
-  return(out)
-}
-
-p1 <- plot_var_mean(hub_data)
-p2 <- plot_var_mean(hub_data, log = TRUE)
-
+## Figure 5 - Correlations between everything
+```{r}
 wis_vs_total <- function(scores, summary_fct = mean, fulldata = hub_data) {
@@
 
-# log-log versions of that plot
-
 scatter_wis_obs <- wis_vs_total(scores) + 
@@
   scale_x_continuous(labels = scale_fn, trans = "log10") +
-  scale_y_continuous(labels = scale_fn, trans = "log10")
-
-p1 / p2 / scatter_wis_obs + 
-  plot_layout(heights = c(1, 1, 2))
-
-ggsave("output/figures/HUB-mean-scores-vs-total-log-log.png", width = 7, height = 7)
-```
-
-## Figure 5 - Correlations between everything
-```{r}
+  scale_y_continuous(labels = scale_fn, trans = "log10") + 
+  ggpubr::stat_cor(method = "spearman", p.accuracy = 0.001, 
+                   cor.coef.name = "rho", size = 3)
+
 df <- scores |>
@@
   plot_annotation(tag_levels = "A") +
-  plot_layout(widths = c(1, 1), heights = c(2, 1)) &
+  plot_layout(widths = c(1, 1), heights = c(3, 1)) &
   theme(legend.position = "bottom")
@@
 
+```{r}
+s <- hub_data |>
+  filter(model %in% c("EuroCOVIDhub-ensemble", "epiforecasts-EpiNow2")) |>
+  score()
+
+s_summarised <- s |>
+  summarise_scores(by = c("model", "range", "horizon"), metric = c("interval_score"))
+
+s_summarised |>
+  select(c(model, interval_score, range, horizon)) |>
+  pivot_wider(names_from = model, values_from = interval_score) |>
+  mutate(rel = `epiforecasts-EpiNow2` / `EuroCOVIDhub-ensemble`) |>
+  filter(horizon == 4)
+  ggplot(aes(y = rel, x = range, 
+             color = factor(horizon), group = factor(horizon))) +
+  geom_line()
+
+s_summarised |>
+  filter(horizon == 1) |>
+  select(c(model, interval_score, range, horizon)) |>
+  ggplot(aes(y = interval_score, x = range, 
+             color = model, group = model)) +
+  geom_line()
+
+```
+
 # Table with relevant values
@@
 ```
+
+Figure
+Currently discontinuoued and just there to discuss it. 
+that shows Variance vs. mean of the observations. Makes the point that the
+data looks neg. binomial. 
+```{r}
+# plot with mean vs. variance --------------------------------------------------
+plot_var_mean <- function(hub_data, log = FALSE) {
+  df <- hub_data |>
+    mutate("x" = true_value) |>
+    dplyr::select(location, target_type, target_end_date, true_value, x) |>
+    unique() |>
+    group_by(location, target_type)
+
+  f <- function(x) {x + x^2}
+
+  if (log) {
+    df <- mutate(df, true_value = log(pmax(true_value, 0) + 1))
+
+    f <- function(x) {log(log(x) + log(x)^2 + 1)}
+  }
+
+  out <- df |>
+  summarise(mean_value = mean(x), 
+            sd_value = sd(true_value), 
+            var_value = var(true_value)) |>
+  ggplot(aes(y = var_value, x = mean_value)) +
+  geom_point() + 
+  facet_wrap(~ target_type, scales = "free") + 
+  scale_x_continuous(trans = "log10", labels = label_fn) +
+  scale_y_continuous(trans = "log10", labels = label_fn) +
+  labs(y = "Var observations", x = "Mean observations") + 
+  # geom_function(fun = f, linetype = "dashed", size = 0.5) +
+  theme_scoringutils() + 
+  geom_smooth(alpha = 0.2, size = 0.2)
+
+  if (log) {
+    out <- out +
+      labs(y = "Var log(observations)", x = "Mean observations")
+  }
+
+  return(out)
+}
+
+p1 <- plot_var_mean(hub_data)
+p2 <- plot_var_mean(hub_data, log = TRUE)
+
+wis_vs_total <- function(scores, summary_fct = mean, fulldata = hub_data) {
+
+  total_values <- 
+    fulldata |> 
+    filter(!is.na(prediction)) |>
+    group_by(location, target_type, target_end_date) |>
+    summarise(true_value = unique(true_value)) |>
+    group_by(location, target_type) |>
+    summarise(total = mean(true_value))
+
+  plot_df <- scores |> 
+    filter(model == "EuroCOVIDhub-ensemble", 
+           horizon == 2) |>
+    group_by(location, target_type, scale, type_and_scale) |>
+    summarise(wis = summary_fct(interval_score), 
+              .groups = "drop_last") |> 
+    inner_join(total_values)
+
+  plot_df |>
+    group_by(target_type, scale) |>
+    summarise(correlation = cor(wis, total)) |> 
+    print()
+
+  plot_df |>
+    ggplot(aes(y = wis, x = total)) + 
+    geom_point() + 
+    facet_wrap(~ type_and_scale, scale = "free") + 
+    theme_scoringutils() + 
+    theme(legend.position = "none") + 
+    scale_x_continuous(labels = scale_fn) +
+    scale_y_continuous(labels = scale_fn) +
+    labs(y = "REPLACE", x = "Total number of observed values")
+}
+
+# log-log versions of that plot
+
+scatter_wis_obs <- wis_vs_total(scores) + 
+  labs(y = "Mean WIS") + 
+  scale_x_continuous(labels = scale_fn, trans = "log10") +
+  scale_y_continuous(labels = scale_fn, trans = "log10")
+
+p1 / p2 / scatter_wis_obs + 
+  plot_layout(heights = c(1, 1, 2))
+
+ggsave("output/figures/HUB-mean-scores-vs-total-log-log.png", width = 7, height = 7)
+```

### 2022-11-21 main.tex
Manuscript updates

@@
 
-\tableofcontents
+%\tableofcontents
 
@@
 Probabilistic forecasts play an important role in decision-making in epidemiology and public health \citep{reichCollaborativeMultiyearMultimodel2019, funkShorttermForecastsInform2020, cramerEvaluationIndividualEnsemble2021, bracherShorttermForecastingCOVID192021, europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}, as well as other areas as diverse as economics \citep{timmermannForecastingMethodsFinance2018, elliottForecastingEconomicsFinance2016} or meteorology \citep{gneitingWeatherForecastingEnsemble2005, kukkonenReviewOperationalRegionalscale2012}.
-Epidemiological modelling in particular has received widespread attention during the COVID-19 pandemic, during which it has demonstrated the critical role it can play in informing public policy.
-Forecast evaluation is important, as it provides feedback for forecasters and researchers to improve their models, and it helps decision makers distinguish good from bad predictions and choosing forecasts that will inform future decisions.
+Epidemiological modelling in particular has received widespread attention during the COVID-19 pandemic.
+Evaluations of forecasts can provide feedback for forecasters and researchers to improve their models, and help decision makers distinguish good from bad predictions and choosing forecasters and models that should inform future decisions.
 
-Probabilistic forecasts are usually evaluated using so-called proper scoring rules, which return a numerical score as a function of the forecast and the later observed data.
+Probabilistic forecasts \citep{heldProbabilisticForecastingInfectious2017} are usually evaluated using so-called proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data.
 Forecasters, in expectation, optimise their score by providing a predictive distribution that is equal to the data-generating distribution, and will be penalised if they fail to do so.
@@
 The WIS and the CRPS measure the absolute distance between the predictive distribution and the observed value and can be understood as a generalisation of the absolute error to full predictive distributions (see details in section XX in the SI).
-For simplicity, we will use "absolute error" and "absolute distance" interchangeably, even though the former applies only to point forecasts and the latter would be more appropriate for predictive distributions.
-The WIS has been used by several Forecast Hubs for COVID-19 in the US [CITATION], Europe [CItation], Germany and Poland [CITATION], as well as the US Flu Forecasting Hub and the Scenario Hubs in US and Europe and XX.
+For simplicity, we will use "absolute error" and "absolute distance" interchangeably, even though the former applies only to point forecasts and the latter is more appropriate for predictive distributions.
+The WIS has been and is currently used by several Forecast Hubs for COVID-19 in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Flu Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
 
-Evaluating forecasts based on the absolute distance between forecast and observed value may not always be optimal in an epidemiological context.
 Epidemiological processes are commonly understood and modeled to be inherently exponential in nature, as every infectious individual is assumed to lead to some number of additional infections [CITATION].
-It would therefore seem more appropriate to focus the evaluation on errors on the multiplicative growth rate, rather than on absolute errors on the observed values, to better represent the underlying disease dynamics.
-Evaluating forecasts based on absolute errors in practice also means that scores are strongly influenced by the overall order of magnitude of the quantity to forecast.
-This makes it difficult compare predictive performance across time, given that epidemiological processes may be non-stationary, or to compare performance across targets that are on very different orders of magnitudes (for example cases and hospitalisations) [CITE KATH ECDC, and others].
+Evaluating forecasts based on the absolute distance between forecast and observed value may therefore not be optimal in an epidemiological context, and it could be more appropriate to focus the evaluation on errors on the multiplicative growth rate, in order for the evaluation to better represent the underlying disease dynamics.
+Evaluating forecasts based on absolute, rather than relative errors, in practice also means that scores are strongly influenced by the overall order of magnitude of the quantity to forecast.
+This makes it difficult compare predictive performance across time, given that epidemiological processes may be non-stationary, or to compare performance across targets that are on very different orders of magnitudes (for example cases and hospitalisations) [CITE \citep{sherrattPredictivePerformanceMultimodel2022}, and others].
 
@@
 
-The goal of log transforming forecasts and observations is to shift the focus of the evaluation while preserving the propriety of the score.
-Instead of a score representing the magnitude of absolute errors this approach allows to obtain a score which measures relative error and provides an approximate measure for how well a forecast captures the multiplicative growth rate of the target quantity.
-In this paper we focus on the log transformation, but note that many other transformations may be useful and appropriate (see section XXXX) .
+The goal of log transforming forecasts and observations is to shift the focus of the evaluation in a way that may more appropriate for epidemiological forecasts, while preserving the propriety of the score.
+Instead of a score representing the magnitude of absolute errors this approach makes it possible to obtain a score which a) measures relative error and b) provides an approximate measure for how well a forecast captures the multiplicative growth rate of the target quantity.
+Focusing on relative errors mitigates issues with scores based on absolute errors that are heavily influenced by the order of magnitude of the quantity to forecast.
+The interpretation of the log transformed score as an approximate score for a forecast of the growth rate of a given quantity aligns the evaluation with our understanding of how epidemiological processes should be modeled and the kinds of errors models would typically make [CITATION?].
+In this paper we focus on the log transformation, but note that many other transformations may be useful and appropriate (see section \ref{sec:discussion}) .
 
-The log transformation is permissible (in the sense that it preserves propriety of the score), because it is a strictly monotonic transformation.
-Even though actual rankings between models may change, a forecaster will in expectation still minimise her score if she reports a predictive distribution that is equal to the data-generating distribution.
-The order of the operations matters, and applying the log transformation after scores have been computed, i.e.
-computing $\log S(x, y)$ results in an improper score.
-This is because taking the logarithm of the WIS / CRPS results in a score that does not penalise outliers enough and therefore incentivises overconfident predictions.
-We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can easily be seen that overconfident models perform best in terms of the log WIS.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/example-log-first.png}
-    \caption{Scores for different forecasts evaluated the WIS and the log WIS.
-We simulated 1000 observations $Y_i = {\rm e}^{x_i}$, with $x_i \text{iid} \sim \mathcal{N}(0, 1)$.
-We then simulated 20 forecasters who would issue a predictive distribution $F = {\rm e}^{x_i}$, with $x \sim \mathcal{N}(0, \sigma)$, with values of $\sigma$ ranging from 0.1 to 2.}
-    \label{fig:log-improper}
-\end{figure}
-
-Unfortunately, the log transformation cannot readily be applied when dealing with negative or zero values.
-In an epidemiological setting that involves count data, negative values should likely be omitted entirely.
-One potential way to deal values of zero is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm as is common practice when log transforming counts.
-When dealing with count data we could extend this argument to all small values (rather than just zeros), as it is hard to reliably evaluate relative errors when dealing with small discrete values.
-An alternative solution is to remove forcasts for small, relative, values from ongoing analysis.
-In the following sections we will explore the implications of both of these strategies.
-
-\subsection{Scoring multiplicative errors}
+\subsection{Interpretation as a score based on multiplicative (relative) errors}
 
@@
 
-On the natural scale, WIS and CRPS increase linearly with increasing absolute errors (see Figure \ref{fig:change-in-scores}A, regardless of whether the error is large in relative terms.
-Predicting 101,000 cases instead of the true 100,000 will be treated equally to predicting 2,000 hospitalisations, rather than the 1,000 observed.
-On the log scale, WIS and CRPS penalise relative errors and increase linearly with an increase in relative errors (see Figure \ref{fig:change-in-scores}D).
-Forecasting 101,000 rather than 100,000 cases will be treated equally to forecasting 2,020, rather than 2,000 hospitalisations.
-While the absolute error on the log scale is exactly symmetric (e.g., predicting 2 instead of 1 would give the same error as predicting 0.5), this is not entirely true for the WIS.
-Due to the fact that upper and lower bounds of individual prediction intervals are affected differently, there may be slight deviations (compare e.g.
-in Figure \ref{fig:change-in-scores}D the scores for a relative error of $\frac{1}{5}$ and 5).
+On the natural scale, WIS and CRPS increase linearly with increasing absolute distance between forecast and observation (see Figure \ref{fig:change-in-scores}A), regardless of whether the error is large in relative terms.
+On the log scale, WIS and CRPS conceptually penalise relative errors and increase linearly with an increase in relative errors (see Figure \ref{fig:change-in-scores}D).
+
+Note that the analogy between the WIS on the log scale and the absolute error of a point forecast on the log scale is slightly inaccurate: While the absolute error on the log scale is exactly symmetric (e.g., predicting 2 instead of 1 would give the same relative error as predicting 0.5), this is not entirely true for the WIS, which is computed for a set of prediction intervals.
+Due to the fact that upper and lower bounds of individual prediction intervals are affected slightly differently by the log transformations, small deviations may occur (compare e.g.
+in Figure \ref{fig:change-in-scores}D the small difference between scores for an observed outcome of $\frac{1}{5}$ and 5).
 
@@
     \includegraphics[width=0.99\textwidth]{output/figures/SIM-effect-log-score.png}
-    \caption{Weighted interval score for a forecast distribution that is N(1, 0.2) and different observed values.
-Top: difference in absolute terms.
-Bottom: difference in relative terms.
-Interesting panels are top left and bottom right.}
+    \caption{Illustration of the weighted interval score (black line) for a $\mathcal{N}(1, 0.2)$ predictive distribution (density represented by the grey area) as a function of the observed value.
+A: Score on the natural scale, linear x-axis.
+B: Score on the natural scale, logarithmic x-axis.
+C: Score on the log scale, linear x-axis.
+D: Score on the log scale, logarithmic x-axis.}
     \label{fig:change-in-scores}
@@
 
-\subsection{Approximately scoring the growth rate}
+\subsection{Interpretation as an approximate score of the growth rate}
+
 Computing a score based on the logarithm of the forecast and the observation is also approximately equivalent to scoring a forecast of a multiplicative growth rate of the forecast target.
@@
 
-Now consider a forecast on $\log y_{t+1}$.
+Now consider a forecast on $\log y_{t+1}$, where again, the score is determined by the size of $\varepsilon^*_{t+1}$.
 Using the fact that for small values of $x$, $\log (1+ x) \approx x$, we can see that scoring forecasts on the log scale is approximately equal to scoring an additive error on the growth rate (on the natural scale), as long as the growth rate is small.
@@
                     &= \log ((1 + \hat{g}_{t+1}) \cdot y_{t}) + \varepsilon^*_{t+1} \\
-                    &= \log (1 + \hat{g}_{t, t+1}) + \log y_t + \varepsilon_{t+1} \Leftrightarrow \\
-\varepsilon'_{t+1}  &= \log y_{t+1} -  \log y_t - \log (1 + \hat{g}_{t, t+1}) \Leftrightarrow \\    
-                    &= \log (\frac{y_{t+1}}{y_t}) - \log (1 + \hat{g}_{t, t+1}) \Leftrightarrow \\    
-                    &= \log (1 + g_{t, t+1}) - \log (1 + \hat{g}_{t, t+1}) \Leftrightarrow \\   
+                    &= \log (1 + \hat{g}_{t, t+1}) + \log y_t + \varepsilon^*_{t+1} \Leftrightarrow \\
+\\
+\varepsilon^*_{t+1}  &= \log y_{t+1} -  \log y_t - \log (1 + \hat{g}_{t, t+1}) \\    
+                    &= \log (\frac{y_{t+1}}{y_t}) - \log (1 + \hat{g}_{t, t+1}) \\    
+                    &= \log (1 + g_{t, t+1}) - \log (1 + \hat{g}_{t, t+1}) \\   
                     &\approx g_{t, t+1} - \hat{g}_{t, t+1} \quad \text{for small differences between } \hat{g}_{t, t+1} \text{ and } g_{t, t+1}. 
@@
 
-This approximation breaks down when the difference between the predicted and the actual growth rate becomes too large (albeit forecasts are still evaluated in terms of relative errors).
-If this is of great concern one may want to use a different transformation such as dividing forecasts and observations by the last known values in order to score the growth rate explicitly (although this lacks the elegance of having one single transformation per time point).
+The score is determined by the size of $\varepsilon^*_{t+1}$, which is approximately equal to the difference between the observed and the predicted growth rate.
+Note that this approximation becomes increasingly inaccurate when the difference between the predicted and the actual growth rate grows (although forecasts can still be interpreted in terms of relative errors as shown above).
+%If this is of great concern one may want to use a different transformation such as dividing forecasts and observations by the last known values in order to score the growth rate explicitly (although this lacks the elegance of having one single transformation per time point).
 
-Conveniently, the decomposition of the WIS into dispersion, overprediction and underprediction is preserved.
+% Conveniently, the decomposition of the WIS into dispersion, overprediction and underprediction is preserved.
 The interpretation of the components has now changed in that they now (approximately) represent uncertainty, over-prediction, and under-prediction with the growth rate and so are independent of current incidence.
@@
 \subsection{Effects on model rankings}
-Rankings between different forecasts may be affected by applying the log transformation even when looking at a single forecast.
-As an example, consider two forecasters, A and B.
+Rankings between different forecasters may change through the log transformation both when looking at aggregate scores as well as when looking at a single forecast.
+As an example for the effect on a single forecast, consider two forecasters, A and B.
 Both issue a single 50\%-prediction interval for a target which later resolves as 1.
@@
 This means that in relative terms, Forecaster B was noticeably closer to the true value with the lower bound of her prediction interval, while only being slightly further away with the upper interval.
-This is likely to lead to a more lenient treatment of (one-sided) outlier forecasts.
+This implies that scores on the log scale may treat (one-sided) outlier forecasts more leniently.
 
-Overall model rankings usually differ even more when performance is averaged across multiple forecasts.
-This effect is not mainly driven by the kinds of changes in model rankings for single forecasts we just discussed.
+% PARAGRAPH NEEDS SOME WORK
+Overall model rankings usually differ even more when scores are averaged across multiple forecasts.
+The change in rankings of aggregate scores is not mainly driven by the kinds of changes in model rankings for single forecasts just discussed.
 Rather, as observations from exponential processes are likely to be on very different absolute scales, large observations will account for the majority of absolute score but will not dominate relative scores in the same way.
@@
 Whether or not forecast targets with smaller quantities receive higher scores than those for targets with larger quantities, depends on the exact relationship between the mean and the variance of the quantity of interest.
-
-For simulated negative binomial count data, log predictions for small quantities receive on average higher scores than forecasts for large quantities if the mean and the variance grow at the same rate $\sigma^2 = \mu$, about equal scores when the variance grows at a rate of $\sigma^2 = \mu + \mu^2$ (Figure \ref{fig:SIM-wis-state-size-mean}), and smaller scores when the variance grows faster than that.
+For simulated count data following a negative binomial distribution, for example, log predictions for small quantities receive on average higher scores than forecasts for large quantities if the mean and the variance grow at the same rate $\sigma^2 = \mu$, about equal scores when the variance grows at a rate of $\sigma^2 = \mu + \mu^2$ (Figure \ref{fig:SIM-wis-state-size-mean}), and smaller scores when the variance grows faster than that.
 % To illustrate this with count data we sampled forecasts from different negative binomial distributions.
@@
 
-In the previous subsection we suggested to add a small number to all values in order to handle zeroes in the observations.
-While the choice of the number to add is somewhat arbitrary, it does have an influence on scores.
-By adding a number to forecasts as well as observations, measures of relative errors are slightly altered (while still preserving propriety of the score as we're applying a strictly monotonic transformation).
-Especially for forecasts of small values, the choice of the number that is added may affect scores quite strongly, as relative errors shrink most for small quantities when adding a number.
-In principle, it is also possible to add a large number (e.g.
-100 or 10,000) to all observations and forecasts in order to equalise relative errors between forecasts of different quantities if one finds that average scores are dominated too strongly by forecasts for small quantities.
+\subsection{Practical considerations}
+
+The log transformation is permissible (in the sense that it preserves propriety of the score), because it is a strictly monotonic transformation.
+Even though actual rankings between models may change, a forecaster will in expectation still minimise her score if she reports a predictive distribution that is equal to the data-generating distribution.
+The order of the operations matters, and applying the log transformation after scores have been computed, i.e.
+computing $\log S(F, y)$ results in an improper score.
+This is because taking the logarithm of the WIS / CRPS results in a score that does not penalise outliers enough and therefore incentivises overconfident predictions.
+We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can easily be seen that overconfident models perform best in terms of the log WIS.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/example-log-first.png}
+    \caption{Scores for different forecasts evaluated the WIS and the log WIS.
+We simulated 1000 observations $Y_i = {\rm e}^{x_i}$, with $x_i \text{iid} \sim \mathcal{N}(0, 1)$.
+We then simulated 20 forecasters who would issue a predictive distribution $F = {\rm e}^{x_i}$, with $x \sim \mathcal{N}(0, \sigma)$, with values of $\sigma$ ranging from 0.1 to 2.}
+    \label{fig:log-improper}
+\end{figure}
+
+In practice, one issue with the log transformation is that it cannot readily be applied to negative numbers or zero values.
+In addition, it may difficult to reliably evaluate relative errors when dealing with small discrete values.
+In an epidemiological setting that involves count data, negative values should likely be omitted entirely.
+One solution to deal with zeroes and small values is to also exclude those from ongoing analysis.
+An alternative is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm as is common practice when log transforming counts.
+This represents a strictly monotonic transformation and therefore preserves propriety of the resulting score.
+The choice of the number to add influences scores and rankings, as measures of relative errors shrink when adding a constant to the forecast and the observation.
+Adding something to the forecasts and observations changes scores for forecasts of small quantities most, as relative errors shrink more the smaller the original value before adding a constant.
+In principle, it would also be possible to exploit the difference in how scores are affected in order to equalise scores for forecasts of quantities across different scales by adding a larger number (e.g.
+100 or 10,000) if average scores are dominated too strongly by forecasts for small quantities.
 
@@
 
-PUT SOMEWHERE
-The European COVID-19 Forecast Hub is one of several Forecast hubs (together with a Hub in the US \citep{cramerEvaluationIndividualEnsemble2021} and one in Germany and Poland \citep{bracherShorttermForecastingCOVID192021} that systematically collected, aggregated and evaluated forecasts of several COVID-19 targets created by different teams.
-Forecasts submitted to the COVID-19 Forecast Hubs are probabilistic in nature [Citation e.g.
-Held et al.], meaning that forecasters provide a full predictive distribution which quantifies their uncertainty.
-These predictive distributions are reported in the form of a set of 23 quantiles (11 symmetric prediction intervals as well as a median prediction).
-
 As an empirical example for evaluating forecasts on the natural and on the log scale we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
-Every week the European Forecast Hub collates and aggregates forecasts for different COVID-19 related targets from teams around the world.
+The European COVID-19 Forecast Hub is one of several COVID-19 Forecast hubs (together with a Hub in the US \citep{cramerEvaluationIndividualEnsemble2021} and one in Germany and Poland \citep{bracherShorttermForecastingCOVID192021} that systematically collects, aggregated and evaluated forecasts of several COVID-19 targets created by different teams every week.
 Forecasts are made one to four weeks ahead into the future and follow a quantile-based format with a set of 22 quantiles plus the median ($0.01, 0.025, 0.05, ..., 0.5, ...
 0.95, 0.975, 0.99$).
-The forecasts for the purpose of this illustrations are two-week-ahead forecasts made between XXX and XXX for reported cases and deaths from COVID-19 in XXX countries.
+
+The forecasts used for the purpose of this illustrations are two-week-ahead forecasts submitted between February 8 2021 and October 18 2021 for reported cases and deaths from COVID-19.
 We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 16 forecast dates.
@@
 
-In addition to the WIS we use pairwise comparisons (CITATION Kramer) to evaluate models against each other.
+In addition to the WIS we use pairwise comparisons \citep{cramerEvaluationIndividualEnsemble2021} to evaluate models against each other.
 In a first step, score ratios are computed for all pairs of models by taking the set of overlapping forecasts between the two models and dividing the score of one model over the score achieved by the other model.
@@
 On the log scale, scores were more evenly distributed between targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}).
-Both on the natural scale as well as on the log scale, scores increased considerably with increasing forecast horizon (see Figure \ref{fig:HUB-scores-horizon}).
+Both on the natural scale as well as on the log scale, scores increased considerably with increasing forecast horizon (see Figure \ref{fig:HUB-mean-locations}E).
 The increase was more pronounced for cases than for deaths.
@@
 A: Mean WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against the average number of observations in a location.
-B: Average Spearman (rank-) correlation of scores for individual forecasts from \textit{all} models.
+B: Average Spearman CHECK AGAIN (rank-) correlation of scores for individual forecasts from \textit{all} models.
 For every individual target (defined by a combination of forecast date, target type, horizon, location), one score was obtained per model.
@@
 
-As expected on the natural scale, scores correlate strongly (XX) with the average number of observed cases / deaths in a given country (see Fig \ref{fig:HUB-mean-locations}A, B).
-This relationship is not evident on the log scale, where correlation is much weaker and maybe slightly negative, implying larger scores for for countries with lower mean incidences.
+As expected on the natural scale, scores correlate strongly with the average number of observed cases and deaths (Spearman correlation: XX and XX) in a given country (see Fig \ref{fig:HUB-mean-locations}A, B).
+This relationship is not evident on the log scale, where correlation is much weaker and maybe slightly negative (Spearman correlation: XX and XX), implying larger scores for for countries with lower mean incidences.
 
@@
 
-For individual forecasts, rankings between models for single forecasts are mostly preserved (average rank correlation for single forecasts: X for cases, Y for deaths), with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}).
-When evaluating performance across different forecasts and forecast targets, relative skill scores of the models change considerably.
+For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}).
+When evaluating performance \textit{across} different forecasts and forecast targets, relative skill scores of the models change considerably.
 The change is stronger for cases than for deaths (Figure \ref{fig:HUB-cors}XX) and tends to increase with increasing forecast horizon.
@@
 On the log scale, these have less of an influence, as errors are smaller in relative terms (see panels C and D).
-We hypothesise that the larger divergence in scores for cases (compared to scores for deaths) displayed in Figure \ref{fig:HUB-cors}C is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths.
+We hypothesise that the larger divergence in scores for cases (compared to scores for deaths) displayed in Figure \ref{fig:HUB-pairwise}C is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths.
 Conversely, failure to predict an upswing while numbers are still low, is not severely punished on the natural scale (see forecasts in July in Figure XX A, B), as overall absolute errors are low.
@@
 On the log scale, scores increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.
+
 Furthermore, including very low numbers in the confidence intervals may lead to very high scores on the log-scale.
@@
 
-Evaluating the rank order of forecasts from the ECDC forecast hub we saw that scoring after a log transform altered the order of forecast models as expected based on our simulated case study.
+Evaluating the rank order of forecasts from the ECDC forecast hub we saw that scoring after a log transform altered the order of forecast models.
 In particular, we note that models that are known to not make use of a exponential error model and to be biased towards under-prediction, such the the MUNI-ARMA model were ranked lower when using a log transform.
@@
 \section{Discussion}
+\label{sec:discussion}
 
-\paragraph{summary of what we did}
-In this paper we suggested that epidemiological forecasts which are today evalaluated using the WIS or CRPS on a natural scale should instead be log transformed first before scoring.
-This shifts the focus of the evaluation from absolute to relative errors and is a convenient way to score forecasts based on how well they are able to predict the growth rate.
-This is useful as it makes sure that the evaluation is consistent with the kinds of errors we expect forecasting models to make in an epidemiological context.
+In this paper we proposed a log transformation of forecasts and observed values prior to evaluation with the WIS (or CRPS) in order to address issues that arise when evaluating epidemiological forecasts based on a measure of absolute error.
+We gave a mathematical intuition for the transformation and showed that scores on the log scale can be interpreted as a) a measure of relative (multiplicative) prediction errors, as well as b) as an approximate score for a forecast of the multiplicative growth rate of the target quantity.
+When applying this approach to forecasts from the European COVID-19 Forecast Hub, we found overall scores to be more equal across, time, location and target type (cases, deaths).
+Scores on the log scale were much less influenced by the overall incidence level in a country, and showed a slight tendency to be higher in locations with a very incidences.
+We found that model rankings changed noticeably, favouring models like epiforecasts-EpiNow2 which explicitly has an exponential error structure, and disfavouring models like MUNI-ARIMA which is purely statistical model with a linear error structure.
+On the natural scale, missing the peak and overshooting was more severely penalised, while missing the nadir and a following upswing in numbers tended to be more severely penalised on the log scale than on the natural scale.
 
-When applied to data from the Europeam Forecasting Hub, we found that
+\paragraph{strengths and weaknesses}
 
-\paragraph{context, implications, limitations}
+Applying a log transformation prior to the WIS means that forecasts are evaluated in terms of relative errors and errors on the multiplicative growth rate, rather than absolute errors.
+The most important strength of this approach is that the evaluation better accommodates the exponential nature of epidemiological process and the types of errors forecasters who accurately model those processes are expected to make.
+The log transformation also helps avoid issues with scores being strongly influenced by the order of magnitude of the forecast quantity, which can be seen when evaluating forecasts on the natural scale.
+A potential downside is that forecast evaluation is unreliable in situations where observed values are zero or very small.
+Similarly, locations with lower incidences may get disproportionate weight (i.e.
+high scores) when evaluating forecasts on the log scale.
+\cite{bracherEvaluatingEpidemicForecasts2021} argue that the large weight given to forecasts for locations with high incidences is a desirable property, as it reflects performance on the targets we should care about most.
+On the other hand, scoring forecasts on the log scale may be less influenced by single outliers and better reflect consistent performance across time, space, and forecast targets.
+It also gives higher weight to another type of situation one may care about, namely one in which numbers start to rise from a previously low level.
+While scores on the log scale can always be interpreted as a measure of relative (multiplicative) predictive performance, the interpretation as a score for a forecast of the multiplicative growth rate is only approximate.
+If one is explicitly interested in scoring a forecast of the growth rate, it may be better to use a different transformation, e.g.
+dividing every forecast and corresponding observed value by the previously last observed value when the forecast was made.
 
-SOME DIFFERENCES BETWEEN SCORES ARE MEANINGFUL. E.G. WE WOULDN'T WANT SCORES FOR CASES AND DEATHS TO BE COMPLETELY EQUAL, BECAUSE DEATHS ARE LIKELY EASIER TO FORECAST
+It is important to note that the log transformation is only one of many transformation that may be useful and appropriate in an epidemiological context.
+One easy transformation is to divide all forecasts by the population in a location, in order to obtain forecasts standardised incidences per e.g.
+100,000 inhabitants.
+A second transformation was mentioned above, namely converting forecasts into forecasts for the multiplicative growth rate by dividing by the last observed value.
+One could also think about applying a Box-Cox or a square-root transformation in order to stabilise the variance of the forecasts.
+Another promising transformation would be to divide each forecast by the forecast of the previous week (and observations by the observation of the previous week), in order to obtain forecasts for week-to-week growth rates.
+This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory (for a different approach to evaluation the shape of a forecast, see https://arxiv.org/abs/2209.04035).
+This, unfortunately is not well defined in the context of forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate, but may be an interesting approach if predictive samples are available.
+What makes this approach of transforming forecasts before scoring particularly interesting is the possibility to construct composite scores as a weighted sum of scores based on different transformations.
+This would allow forecast consumers to assign explicit weights to different qualities of the forecasts they might care about.
 
-\paragraph{outlook, future work}
-
-\paragraph{Other transformations}
-While we focus on the log transformation in this paper, there are other transformations that may be helpful in an epidemiological context.
-One could for example divide all forecasts by the population in a location, to obtain forecasts standardised incidences per e.g.
-100,000 inhabitants.
-Or one could apply a Box-Cox transformation MISSING. 
-Or a square-root transformation MISSING. 
-Or one could divide each forecast by the forecast of the previous week (and observations by the observation of the previous week), to transform preditions to week-to-week growth rates.
-This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory.
-Note that this is not well defined in the context of forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate.
-
-Whether or not a log transformation is appropriate depends on whether one believes that differences in absolute values are meaningful.
-This may be the case for example when comparing forecasts across locations or time points, rather than different forecasts targets.
-\cite{bracherEvaluatingEpidemicForecasts2021} for example argue that in epidemiological settings it is a desirable property of the CRPS / WIS that it assigns large scores to forecasts where counts are high, as this naturally reflects the situations we should most care about and helps us identify the models that perform best in similar situations.
-If, on the other hand, we believe that good models should do consistently well and that situations with high incidences do not provide significantly more information about which models perform well or badly, then scoring on the log scale may be more appropriate.
-For example, it may be reasonable not to think that predictive performance in a country like Germany would have a hundred times more importance than performance in a country like Luxembourg when determining the best model for future decision making.
-Similarly, we may not necessarily believe that predictive performance during the January 2022 wave of COVID-19 should carry three times as much weight than predictive performance during the January 2021 wave, just because numbers where three times as high.
-
-% Divide by the latest known value = score forecast on the growth rate
-% - WHAT KINDS OF DATA CAN I CONDITION ON? 
-% - WOULD DIVIDING A FORECAST BY ITS RESOLUTION WORK? 
-% - CAN I SCORE THE TRAJECTORY / WEEK-TO-WEEK GROWTH RATE BY DIVIDING the 4 week ahead forecast by the three week ahead forecast?
+Exploring these different transformations is a promising avenue for future work that could help bridge the gap between modellers and policy makers by providing scoring rules that better reflect what forecast consumers actually care about.
+We have showed that the log transformation can lead to significant changes in the relative rankings of models against each other, with potentially important implications for decision makers who rely on the knowledge of past performance to make a decision about which forecasts should inform future decisions.
+We think it would therefore be a good idea to show different versions of the WIS in order to obtain a richer picture of model performance.
+While we have provided a small case study, more work needs to be done to better understand the effects of applying a log transformation in different contexts and how it might affect decision making.
 
@@
 % could make an analysis re outliers: how do scores change if we remove the 1 or 2 worst forecasts? The one or two best forecasts?
-
-NEEDS TO GO SOMEWHERE
-However, we note that the same logic we use can also applied to other transformations which could shift the focus of evaluation in different ways.
-For example, given an appropriate transformation, it could also be possible to evaluate forecasters based on the shape of the trajectory [CITATION https://arxiv.org/abs/2209.04035] they predicted.
-It would even be possible to construct composite scores as a weighted sum of scores based on different transformations, allowing forecast consumers to assign explicit weights to different qualities of the forecasts.
 
@@
 \subsection{Additional information on the WIS} \label{wis}
-\paragraph{WIS}
 
+The Weighted Interval Score is a proper scoring rule.
 WIS values are always larger or equal than zero and lower values imply better performance.
@@
 
-The CRPS measures the 'distance' of the predictive distribution to the observed data-generating distribution as
+The CRPS measures the distance of the predictive distribution to the observed data-generating distribution as
 
@@
 
-\subsection{PROBABLY DELETE? Changes in the WIS for forecasts on the log scale} \label{wis-log-derivation}
+\subsection{Additional figures}
 
-This connection with the underlying epidemic growth rate is an attractive property in and of itself as it is the conceptual target many forecasters and forecast consumers make use of.
-
-When we score a forecast for $y_{t+1}$ on the log scale, the interval score for a single prediction interval changes as follows
-%
-\begin{align}
-    \text{IS}_\alpha(\log F, \log y) = &(\log u - \log l) \\ 
-    &+ \frac{2}{\alpha} \cdot (\log l - \log y) \cdot 1(\log y \leq \log l) \\
-    &+ \frac{2}{\alpha} \cdot (\log y - \log u) \cdot 1(\log y \geq \log u).
-\end{align}
-%
-Note that the values of the indicator functions stay the same as the logarithm is a monotonic transformation and the inequality is preserved. 
-On the natural scale, $u$, $l$ are the lower and upper bounds of a central prediction interval for a forecast of $y_{t+1} = (1 + g_{t, t+1}) \cdot y_t$. 
-We can rewrite these as 
-\begin{equation}
-   l = l^* \cdot y_t 
-   \;\; \text{and} \;\; 
-   u = u^* \cdot y_t,  
-\end{equation}
-where $u^*$ and $l^*$ are the lower and upper bounds of a prediction interval for $(1 + g_{t, t+1})$. Consequently, 
-\begin{equation}
-  \log l = \log l^* + \log y_t
-  \;\; \text{and} \;\;
-  \log u = \log u^* + \log y_t. 
-\end{equation}
-%
-We can then rewrite the score for a forecast of $\log y_{t+1} = \log (1 + g_{t,t+1}) + \log y_t$ as
-%
-\begin{equation}
-\begin{aligned}
-IS_\alpha&(\log F_{t+1}, \log y_{t+a}) = \\
-&((\log u^* + \log y_t) - (\log l^* + \log y_t)) \\
-&+ \frac{2}{\alpha} \cdot ((\log l^* + \log y_t) - (\log (1 + g_{t, t+1}) + \log y_t)) 
-     \cdot 1((\log (1 + g_{t, t+1}) + \log y_t) \leq (\log l^* + \log y_t) \\
-&+ \frac{2}{\alpha} \cdot ((\log (1 + g_{t, t+1}) + \log y_t) - (\log u^* + \log y_t)) 
-      \cdot 1((\log (1 + g_{t, t+1}) + \log y_t) \geq (\log u^* + \log y_t),
-\end{aligned}
-\end{equation}
-%
-which simplifies to 
-%
-\begin{equation}
-\begin{aligned}
-\label{eqn:is-log}
-\text{IS}_\alpha(\log F_{t+1}, \log y_{t+a}) = &(\log u^* - \log l^*) \\
-&+ \frac{2}{\alpha} \cdot (\log l^* - \log (1 + g_{t, t+1}) 
-     \cdot 1(\log (1 + g_{t, t+1}) \leq \log l^*) \\
-&+ \frac{2}{\alpha} \cdot (\log (1 + g_{t, t+1}) - \log u^*) 
-      \cdot 1(\log (1 + g_{t, t+1}) \geq \log u^*).
-\end{aligned}
-\end{equation}
-
-The left-hand side of equation \ref{eqn:is-log} was left unchanged.
-From the right-hand side we can see that for a single prediction interval, scoring the log of a forecast ($\log F_{t+1}$) against the log observed value ($\log y_{t+1}$) is exactly equivalent to evaluating a forecast for $\log (1 + g_{t, t+1})$.
-With $\log (1 + g_{t, t+1} \approx g_{t, t+1}$, this is approximately a forecast for the growth rate for small values of $g_{t, t+1}$.
-
-\subsection{Additional figures}
+NUMBER OF AVAILABLE FORECASTS
 
@@
 The relationship is less pronounced for scores on the log scale (cor: -0.019 for cases, -0.395 for deaths) and negative, meaning that locations with fewer observed cases or deaths tend to receive higher scores.
+
+%Whether or not a log transformation is appropriate depends on whether one believes that differences in absolute values are meaningful.
+This may be the case for example when comparing forecasts across locations or time points, rather than different forecasts targets.
+\cite{bracherEvaluatingEpidemicForecasts2021} for example argue that in epidemiological settings it is a desirable property of the CRPS / WIS that it assigns large scores to forecasts where counts are high, as this naturally reflects the situations we should most care about and helps us identify the models that perform best in similar situations.
+If, on the other hand, we believe that good models should do consistently well and that situations with high incidences do not provide significantly more information about which models perform well or badly, then scoring on the log scale may be more appropriate.
+For example, it may be reasonable not to think that predictive performance in a country like Germany would have a hundred times more importance than performance in a country like Luxembourg when determining the best model for future decision making.
+Similarly, we may not necessarily believe that predictive performance during the January 2022 wave of COVID-19 should carry three times as much weight than predictive performance during the January 2021 wave, just because numbers where three times as high.
+
+% Divide by the latest known value = score forecast on the growth rate
+% - WHAT KINDS OF DATA CAN I CONDITION ON? 
+% - WOULD DIVIDING A FORECAST BY ITS RESOLUTION WORK? 
+% - CAN I SCORE THE TRAJECTORY / WEEK-TO-WEEK GROWTH RATE BY DIVIDING the 4 week ahead forecast by the three week ahead forecast?

### 2022-11-21 analysis/Analysis ECDC forecasts.Rmd
update plots

@@
 
-## Basic data summary
-```{r}
-hub_data |>
-  summarise(min_date = min(forecast_date), 
-            max_date = max(forecast_date)) |>
-  mutate(week_diff = (max_date - min_date - 1) / 7)
-
-hub_data$location |> 
-  unique() |>
-  length()
-
-hub_data |> 
-  filter(horizon == 2) |>
-  avail_forecasts(by = c("target_end_date", "target_type", "model")) |>
-  mutate(forecast_date = as.Date(target_end_date) - 12) |>
-  plot_avail_forecasts(show_numbers = FALSE) +
-  facet_grid(~ target_type) + 
-  labs(y = "Model", x = "Forecast date")
-
-ggsave(filename = "output/figures/number-avail-forecasts.png", height = 4, width = 10)
-```
-
 ## =============================================================================
@@
 
+## =============================================================================
 ## Figure 5 - Correlations between everything
@@
 
-## Scores across forecast targets
-
-This uses the Hub ensemble
-
-```{r}
-scores |>
-  dplyr::select(location, target_type, scale, target_end_date, interval_score) |>
-  unique() |>
-  group_by(target_type, scale) |>
-  summarise(mean_value = mean(interval_score), 
-            sd_value = sd(interval_score), 
-            var_value = var(interval_score))
-
-ggsave("output/figures/HUB-average-scores.png", width = 7, height = 2.5)
-```
-
-Mean scores: 
-```{r}
-scores |> 
-  filter(model == "EuroCOVIDhub-ensemble", 
-         horizon == 2) |>
-  group_by(scale, target_type) |>
-  summarise(interval_score = mean(interval_score), 
-            .groups = "drop_last") |>
-  make_table()
-```
-
-## Correlation between scores
-
-```{r}
-scores |> 
-  filter(model == "EuroCOVIDhub-ensemble", 
-         horizon == 2) |>
-  group_by(target_type, location, scale) |>
-  summarise(interval_score = mean(interval_score), 
-            .groups = "drop_last") %>%
-  pivot_wider(values_from = interval_score, names_from = scale) %>%
-  group_by(target_type) %>%
-  summarise(correlation = cor(natural, log))
-```
-
-## Change of average scores for increasing forecast horizons
-
-```{r}
-scores |> 
-  filter(model == "EuroCOVIDhub-ensemble") |>
-  group_by(scale, target_type, horizon) |>
-  summarise(interval_score = mean(interval_score), 
-            .groups = "drop_last") |>
-  mutate(multiplicative_score = interval_score / min(interval_score)) |>
-  ggplot(aes(y = multiplicative_score, x = horizon)) + 
-  geom_line() + 
-  geom_point() + 
-  facet_grid(scale ~ target_type) + 
-  theme_scoringutils() + 
-  labs(x = "Forecast horizon in weeks", y = "Multplicative change in interval score")
-
-ggsave("output/figures/HUB-scores-over-horizon.png", width = 7, height = 4)
-
-```
-
-```{r}
-
-```
-
-## Figure 6 - Look at pairwise comparisons and changes in rankings
+## =============================================================================
+## Figure 6 - Example comparisons of forecasts on the log and the natural scale
+
+```{r}
+plot_pred_score <- function(hub_data, scores, model,
+                            horizon = 2, type = "Cases", 
+                            remove_x_text = TRUE) {
+  forecast_model <- model
+  h <- horizon
+  filtered_hub <- hub_data |>
+    filter(model == forecast_model, 
+           horizon %in% h, 
+           target_type == type)
+
+  locations <- unique(filtered_hub$location)
+
+  filtered_scores <- scores |>
+    filter(model == forecast_model, 
+           horizon %in% h, 
+           location %in% locations,
+           target_type == type)
+
+  plot_natural  <- filtered_hub |>  
+    mutate(id = paste(type, "- natural")) |>
+    plot_predictions(x = "target_end_date") +
+    scale_y_continuous(labels = label_fn)  +
+    facet_wrap(~ id) +
+    theme(legend.position = "bottom", 
+          axis.text.x = element_blank(), 
+          axis.title.x = element_blank()) +
+    labs(y = "Predictions")
+
+  plot_log <- filtered_hub |>
+    mutate(prediction = log(prediction + 1), 
+           true_value = log(true_value + 1)) |>
+    mutate(id = paste(type, "- log")) |>
+    plot_predictions(x = "target_end_date") +
+    scale_y_continuous(labels = label_fn)  +
+    facet_wrap(~ id) +
+    theme(legend.position = "bottom", 
+          axis.text.x = element_blank(), 
+          axis.title.x = element_blank()) +
+    labs(y = "Predictions")
+
+  scores_natural <- filtered_scores |>
+    filter(scale == "natural") |>
+    summarise_scores("target_end_date") |>
+    mutate(target_end_date = factor(target_end_date)) |>
+    plot_wis(x = "target_end_date", flip = TRUE) +
+    scale_x_continuous(labels = label_fn) +
+    labs(y = "Target end date")
+
+  scores_log <- filtered_scores |>
+    filter(scale == "log") |>
+    summarise_scores("target_end_date") |>
+    mutate(target_end_date = factor(target_end_date)) |>
+    plot_wis(x = "target_end_date", flip = TRUE) +
+    labs(y = "Target end date")
+
+  if (remove_x_text) {
+    scores_natural <- scores_natural +
+      theme(axis.text.x = element_blank(), 
+            axis.title.x = element_blank())
+    # scores_log <- scores_log +
+    #   theme(axis.text.x = element_blank(), 
+    #         axis.title.x = element_blank())
+  }
+
+  list(plot_natural, scores_natural, plot_log, scores_log )
+}
+
+put_plot_together <- function(modelname = "EuroCOVIDhub-ensemble", 
+                              locationname = "DE") {
+
+  plot_cases <- plot_pred_score(hub_data |> filter(location == locationname), 
+                                  scores, modelname, 
+                                  remove_x_text = TRUE, 
+                                  type = "Cases")
+
+  plot_deaths <- plot_pred_score(hub_data |> filter(location == locationname), 
+                                     scores, modelname, 
+                                     type = "Deaths")
+
+  layout <- "
+AE
+BF
+CG
+DH"
+
+  plot_cases[[1]] + plot_cases[[2]] + plot_cases[[3]] +
+    plot_cases[[4]] + plot_deaths[[1]] +
+    plot_deaths[[2]] + plot_deaths[[3]] + plot_deaths[[4]] +
+    plot_annotation(tag_levels = "A") +
+    plot_layout(guides = "collect", 
+                design = layout) &
+    theme(legend.position = "bottom")
+}
+
+put_plot_together("EuroCOVIDhub-ensemble")
+ggsave(filename = "output/figures/HUB-model-comparison-ensemble.png", width = 10, height = 7)
+
+put_plot_together("epiforecasts-EpiNow2")
+ggsave(filename = "output/figures/HUB-model-comparison-epinow.png", width = 10, height = 7)
+
+put_plot_together("EuroCOVIDhub-baseline", locationname = "DE")
+ggsave(filename = "output/figures/HUB-model-comparison-baseline.png", width = 10, height = 7)
+
+```
+
+## =============================================================================
+## Figure 7 - Look at pairwise comparisons and changes in rankings (plus Appendix figures)
 
@@
   coord_cartesian(#xlim = c(0.5, 4), 
-                  ylim = c(0.8, 9.15)) +
+                  ylim = c(0.75, 8.25)) +
   theme_scoringutils() +
@@
 
+plot_rel_wis + plot_ranking_change
+
 # layout <- "
@@
 )
-
-plot_ranking_change + plot_rel_skill + plot_diffs + plot_rel_wis + #plot_distr_scores 
-  plot_layout(design = layout) +
-  plot_layout(heights = c(1, 2)) +
-  plot_layout(widths = c(2, 1, 2)) +
-  plot_layout(guides = "collect") &
-  theme(legend.position = "bottom")
 }
@@
 
-# Correlations for ranks, scores and pairwise comparisons
-
-```{r}
-
-# 1. observations natural vs. log --> pearson
-# 2. correlation scores natural vs. log (single forecasts) --> spearman
-
-# Table
-# correlation between values
-cor_data <- hub_data |>
-  select(location, target_end_date, target_type, true_value) |>
+## =============================================================================
+## Summary numbers for the text + Figure Available Forecasts fo Appendix
+
+```{r}
+hub_data |>
+  summarise(min_date = min(forecast_date), 
+            max_date = max(forecast_date)) |>
+  mutate(week_diff = (max_date - min_date - 1) / 7)
+
+hub_data$location |> 
   unique() |>
-  mutate(log_obs = log(pmax(true_value, 0) + 1)) |>
-  group_by(target_type) |>
-  summarise(cor = cor(true_value, log_obs))
-
-# # correlation between scores
-# # not really meaningful: if we do ranks it's just going to be ranks over everything which doesn't make sense
-# # if we do individual scores than it's a bit unclear what the "correct" correlation should be
-# cor_scores <- scores |>
-#   filter(horizon <=4) |>
-#   select(interval_score, scale, forecast_date, model, horizon, location, target_type) |>
-#   pivot_wider(names_from = scale, values_from = interval_score) |>
-#   group_by(horizon, target_type) |>
-#   summarise(cor = cor(natural, log), 
-#             spearman = cor(natural, log, method = "spearman"))
-
-# Correlation between ranks when going from natural to log
-correlation_ranks <- rankings_wide |>
-  filter(horizon < 5) |>
-  group_by(horizon, target_type) |>
-  summarise(cor = cor(natural, log), 
-            spearman = cor(natural, log, method = "spearman")) |>
-  rename(`Target type` = target_type)
-
-# we could also do spearman per target and then plot an average of the spearman correlation.
-
-# Correlation between pairwise comparisons
-correlation_rel_skill <- scores |>
-  filter(horizon < 5) |>
-  summarise_scores(
-    by = c("model", "scale", "horizon", "target_type"), 
-    relative_skill = TRUE) |>
-  select(model, horizon, target_type, relative_skill, scale) |>
-  pivot_wider(values_from = relative_skill, names_from = scale) |>
-  group_by(horizon, target_type) |>
-  summarise(cor = cor(natural, log), 
-            spearman = cor(natural, log, method = "spearman"))
-
-p_cor_ranks <- correlation_ranks |>
-  ggplot(aes(x = horizon, y = cor, 
-             fill = `Target type`, color = `Target type`)) + 
-  geom_histogram(stat = "identity", position = "dodge") +
-  theme_scoringutils() + 
-  coord_cartesian(ylim = c(0, 1)) +
-  labs(y = "Corr. ranks per forecast", x = "Forecast horizon")
-
-# p_cor_scores <- cor_scores |> 
-#   mutate(`Target type` = target_type) |>
-#   ggplot(aes(x = horizon, y = cor, 
-#              fill = `Target type`, color = `Target type`)) + 
-#   geom_histogram(stat = "identity", position = "dodge") +
-#   theme_scoringutils() + 
-#   coord_cartesian(ylim = c(0, 1)) +
-#   labs(y = "Spearman corr. scores single forecasts", x = "Forecast horizon")
-
-p_cor_skill <- correlation_rel_skill |>
-  rename(`Target type` = target_type) |>
-  ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
-  geom_histogram(stat = "identity", position = "dodge") +
-  theme_scoringutils() + 
-  coord_cartesian(ylim = c(0, 1)) +
-  labs(y = "Correlation rel. skill", x = "Forecast horizon")
-
-combined <- scores[, c("model", "location", "target_end_date", "scale", "target_type", "interval_score", "horizon")] |>
-  full_join(hub_data[, c("location", "target_end_date", "target_type", "model", "true_value")]) |>
-  mutate(true_value = ifelse(scale == "natural", true_value, log(true_value + 1)))
-
-combined |>
-  filter(horizon <= 4) |>
-  rename(`Target type` = target_type) |>
-  group_by(horizon, scale, `Target type`) |>
-  summarise(cor = cor(interval_score, true_value)) |>
-    ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
-  geom_histogram(stat = "identity", position = "dodge") +
-  theme_scoringutils() + 
-  facet_wrap(~ scale) +
-  expand_limits(ylim = c(0, 1)) +
-  labs(y = "Correlation score and observation", x = "Forecast horizon")
-
-plot_cors <- p_cor_ranks  + p_cor_skill +
-  plot_layout(guides = "collect") +
-  plot_annotation(tag_levels = "A") &
-  theme(legend.position = "bottom")
-
-ggsave(filename = "output/figures/HUB-cors.png", width = 7, height = 3, 
-       plot = plot_cors)
-
-```
-
-## Figure 7 - Example comparisons of forecasts on the log and the natural scale
-
-```{r}
-plot_pred_score <- function(hub_data, scores, model,
-                            horizon = 2, type = "Cases", 
-                            remove_x_text = TRUE) {
-  forecast_model <- model
-  h <- horizon
-  filtered_hub <- hub_data |>
-    filter(model == forecast_model, 
-           horizon %in% h, 
-           target_type == type)
-
-  locations <- unique(filtered_hub$location)
-
-  filtered_scores <- scores |>
-    filter(model == forecast_model, 
-           horizon %in% h, 
-           location %in% locations,
-           target_type == type)
-
-  plot_natural  <- filtered_hub |>  
-    mutate(id = paste(type, "- natural")) |>
-    plot_predictions(x = "target_end_date") +
-    scale_y_continuous(labels = label_fn)  +
-    facet_wrap(~ id) +
-    theme(legend.position = "bottom", 
-          axis.text.x = element_blank(), 
-          axis.title.x = element_blank()) +
-    labs(y = "Predictions")
-
-  plot_log <- filtered_hub |>
-    mutate(prediction = log(prediction + 1), 
-           true_value = log(true_value + 1)) |>
-    mutate(id = paste(type, "- log")) |>
-    plot_predictions(x = "target_end_date") +
-    scale_y_continuous(labels = label_fn)  +
-    facet_wrap(~ id) +
-    theme(legend.position = "bottom", 
-          axis.text.x = element_blank(), 
-          axis.title.x = element_blank()) +
-    labs(y = "Predictions")
-
-  scores_natural <- filtered_scores |>
-    filter(scale == "natural") |>
-    summarise_scores("target_end_date") |>
-    mutate(target_end_date = factor(target_end_date)) |>
-    plot_wis(x = "target_end_date", flip = TRUE) +
-    scale_x_continuous(labels = label_fn) +
-    labs(y = "Target end date")
-
-  scores_log <- filtered_scores |>
-    filter(scale == "log") |>
-    summarise_scores("target_end_date") |>
-    mutate(target_end_date = factor(target_end_date)) |>
-    plot_wis(x = "target_end_date", flip = TRUE) +
-    labs(y = "Target end date")
-
-  if (remove_x_text) {
-    scores_natural <- scores_natural +
-      theme(axis.text.x = element_blank(), 
-            axis.title.x = element_blank())
-    # scores_log <- scores_log +
-    #   theme(axis.text.x = element_blank(), 
-    #         axis.title.x = element_blank())
-  }
-
-  list(plot_natural, scores_natural, plot_log, scores_log )
-}
-
-put_plot_together <- function(modelname = "EuroCOVIDhub-ensemble", 
-                              locationname = "DE") {
-
-  plot_cases <- plot_pred_score(hub_data |> filter(location == locationname), 
-                                  scores, modelname, 
-                                  remove_x_text = TRUE, 
-                                  type = "Cases")
-
-  plot_deaths <- plot_pred_score(hub_data |> filter(location == locationname), 
-                                     scores, modelname, 
-                                     type = "Deaths")
-
-  layout <- "
-AE
-BF
-CG
-DH"
-
-  plot_cases[[1]] + plot_cases[[2]] + plot_cases[[3]] +
-    plot_cases[[4]] + plot_deaths[[1]] +
-    plot_deaths[[2]] + plot_deaths[[3]] + plot_deaths[[4]] +
-    plot_annotation(tag_levels = "A") +
-    plot_layout(guides = "collect", 
-                design = layout) &
-    theme(legend.position = "bottom")
-}
-
-put_plot_together("EuroCOVIDhub-ensemble")
-ggsave(filename = "output/figures/HUB-model-comparison-ensemble.png", width = 9, height = 7)
-
-put_plot_together("epiforecasts-EpiNow2")
-ggsave(filename = "output/figures/HUB-model-comparison-epinow.png", width = 9, height = 7)
-
-put_plot_together("EuroCOVIDhub-baseline", locationname = "DE")
-ggsave(filename = "output/figures/HUB-model-comparison-baseline.png", width = 9, height = 7)
-
-```
-
-```{r}
-scores |>
-  summarise_scores(c("model", "target_type", "scale")) |>
-  summarise_scores(c("model", "target_type", "scale"), fun = signif, digits = 2) |>
-  plot_score_table(by = c("target_type", "scale")) +
-  facet_wrap(target_type ~ scale)
-```
-
-```{r}
-
-split_data <- split(scores, by = c("type_and_scale"))
-cors <- lapply(split_data, FUN = correlation, 
-               metrics = c("interval_score", 
-                            "bias", "overprediction", 
-                            "underprediction",
-                            "dispersion"))
-
-# scoringutils dev stuff
-# split_data <- split(scores, by = c("target_type", "scale"))
-# list_name_as_col <- function(list, by = NULL) {
-#   names <- names(list)
-#   print(names(list_element))
-# }
-
-plots <- lapply(cors, plot_correlation)
-
-(plots[[1]] + plots[[2]]) /
-  (plots[[3]] + plots[[4]])
-
-correlation
-```
-
-```{r}
-s <- hub_data |>
-  filter(model %in% c("EuroCOVIDhub-ensemble", "epiforecasts-EpiNow2")) |>
-  score()
-
-s_summarised <- s |>
-  summarise_scores(by = c("model", "range", "horizon"), metric = c("interval_score"))
-
-s_summarised |>
-  select(c(model, interval_score, range, horizon)) |>
-  pivot_wider(names_from = model, values_from = interval_score) |>
-  mutate(rel = `epiforecasts-EpiNow2` / `EuroCOVIDhub-ensemble`) |>
-  filter(horizon == 4)
-  ggplot(aes(y = rel, x = range, 
-             color = factor(horizon), group = factor(horizon))) +
-  geom_line()
-
-s_summarised |>
-  filter(horizon == 1) |>
-  select(c(model, interval_score, range, horizon)) |>
-  ggplot(aes(y = interval_score, x = range, 
-             color = model, group = model)) +
-  geom_line()
-
-```
+  length()
+
+hub_data |> 
+  filter(horizon == 2) |>
+  avail_forecasts(by = c("target_end_date", "target_type", "model")) |>
+  mutate(forecast_date = as.Date(target_end_date) - 12) |>
+  plot_avail_forecasts(show_numbers = FALSE) +
+  facet_grid(~ target_type) + 
+  labs(y = "Model", x = "Forecast date")
+
+ggsave(filename = "output/figures/number-avail-forecasts.png", height = 4, width = 10)
+```
+
+## =============================================================================
+## Summary table for Appendix
 
@@
 
+## Scores across forecast targets
+
+This uses the Hub ensemble
+
+```{r}
+scores |>
+  dplyr::select(location, target_type, scale, target_end_date, interval_score) |>
+  unique() |>
+  group_by(target_type, scale) |>
+  summarise(mean_value = mean(interval_score), 
+            sd_value = sd(interval_score), 
+            var_value = var(interval_score))
+
+ggsave("output/figures/HUB-average-scores.png", width = 7, height = 2.5)
+```
+
+Mean scores: 
+```{r}
+scores |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  group_by(scale, target_type) |>
+  summarise(interval_score = mean(interval_score), 
+            .groups = "drop_last") |>
+  make_table()
+```
+
+## Correlation between scores
+
+```{r}
+scores |> 
+  filter(model == "EuroCOVIDhub-ensemble", 
+         horizon == 2) |>
+  group_by(target_type, location, scale) |>
+  summarise(interval_score = mean(interval_score), 
+            .groups = "drop_last") %>%
+  pivot_wider(values_from = interval_score, names_from = scale) %>%
+  group_by(target_type) %>%
+  summarise(correlation = cor(natural, log))
+```
+
+# Correlations for ranks, scores and pairwise comparisons
+
+```{r}
+
+# 1. observations natural vs. log --> pearson
+# 2. correlation scores natural vs. log (single forecasts) --> spearman
+
+# Table
+# correlation between values
+cor_data <- hub_data |>
+  select(location, target_end_date, target_type, true_value) |>
+  unique() |>
+  mutate(log_obs = log(pmax(true_value, 0) + 1)) |>
+  group_by(target_type) |>
+  summarise(cor = cor(true_value, log_obs))
+
+# # correlation between scores
+# # not really meaningful: if we do ranks it's just going to be ranks over everything which doesn't make sense
+# # if we do individual scores than it's a bit unclear what the "correct" correlation should be
+# cor_scores <- scores |>
+#   filter(horizon <=4) |>
+#   select(interval_score, scale, forecast_date, model, horizon, location, target_type) |>
+#   pivot_wider(names_from = scale, values_from = interval_score) |>
+#   group_by(horizon, target_type) |>
+#   summarise(cor = cor(natural, log), 
+#             spearman = cor(natural, log, method = "spearman"))
+
+# Correlation between ranks when going from natural to log
+correlation_ranks <- rankings_wide |>
+  filter(horizon < 5) |>
+  group_by(horizon, target_type) |>
+  summarise(cor = cor(natural, log), 
+            spearman = cor(natural, log, method = "spearman")) |>
+  rename(`Target type` = target_type)
+
+# we could also do spearman per target and then plot an average of the spearman correlation.
+
+# Correlation between pairwise comparisons
+correlation_rel_skill <- scores |>
+  filter(horizon < 5) |>
+  summarise_scores(
+    by = c("model", "scale", "horizon", "target_type"), 
+    relative_skill = TRUE) |>
+  select(model, horizon, target_type, relative_skill, scale) |>
+  pivot_wider(values_from = relative_skill, names_from = scale) |>
+  group_by(horizon, target_type) |>
+  summarise(cor = cor(natural, log), 
+            spearman = cor(natural, log, method = "spearman"))
+
+p_cor_ranks <- correlation_ranks |>
+  ggplot(aes(x = horizon, y = cor, 
+             fill = `Target type`, color = `Target type`)) + 
+  geom_histogram(stat = "identity", position = "dodge") +
+  theme_scoringutils() + 
+  coord_cartesian(ylim = c(0, 1)) +
+  labs(y = "Corr. ranks per forecast", x = "Forecast horizon")
+
+# p_cor_scores <- cor_scores |> 
+#   mutate(`Target type` = target_type) |>
+#   ggplot(aes(x = horizon, y = cor, 
+#              fill = `Target type`, color = `Target type`)) + 
+#   geom_histogram(stat = "identity", position = "dodge") +
+#   theme_scoringutils() + 
+#   coord_cartesian(ylim = c(0, 1)) +
+#   labs(y = "Spearman corr. scores single forecasts", x = "Forecast horizon")
+
+p_cor_skill <- correlation_rel_skill |>
+  rename(`Target type` = target_type) |>
+  ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
+  geom_histogram(stat = "identity", position = "dodge") +
+  theme_scoringutils() + 
+  coord_cartesian(ylim = c(0, 1)) +
+  labs(y = "Correlation rel. skill", x = "Forecast horizon")
+
+combined <- scores[, c("model", "location", "target_end_date", "scale", "target_type", "interval_score", "horizon")] |>
+  full_join(hub_data[, c("location", "target_end_date", "target_type", "model", "true_value")]) |>
+  mutate(true_value = ifelse(scale == "natural", true_value, log(true_value + 1)))
+
+combined |>
+  filter(horizon <= 4) |>
+  rename(`Target type` = target_type) |>
+  group_by(horizon, scale, `Target type`) |>
+  summarise(cor = cor(interval_score, true_value)) |>
+    ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
+  geom_histogram(stat = "identity", position = "dodge") +
+  theme_scoringutils() + 
+  facet_wrap(~ scale) +
+  expand_limits(ylim = c(0, 1)) +
+  labs(y = "Correlation score and observation", x = "Forecast horizon")
+
+plot_cors <- p_cor_ranks  + p_cor_skill +
+  plot_layout(guides = "collect") +
+  plot_annotation(tag_levels = "A") &
+  theme(legend.position = "bottom")
+
+ggsave(filename = "output/figures/HUB-cors.png", width = 7, height = 3, 
+       plot = plot_cors)
+
+```
+
+```{r}
+scores |>
+  summarise_scores(c("model", "target_type", "scale")) |>
+  summarise_scores(c("model", "target_type", "scale"), fun = signif, digits = 2) |>
+  plot_score_table(by = c("target_type", "scale")) +
+  facet_wrap(target_type ~ scale)
+```
+
+```{r}
+
+split_data <- split(scores, by = c("type_and_scale"))
+cors <- lapply(split_data, FUN = correlation, 
+               metrics = c("interval_score", 
+                            "bias", "overprediction", 
+                            "underprediction",
+                            "dispersion"))
+
+# scoringutils dev stuff
+# split_data <- split(scores, by = c("target_type", "scale"))
+# list_name_as_col <- function(list, by = NULL) {
+#   names <- names(list)
+#   print(names(list_element))
+# }
+
+plots <- lapply(cors, plot_correlation)
+
+(plots[[1]] + plots[[2]]) /
+  (plots[[3]] + plots[[4]])
+
+correlation
+```
+
+```{r}
+s <- hub_data |>
+  filter(model %in% c("EuroCOVIDhub-ensemble", "epiforecasts-EpiNow2")) |>
+  score()
+
+s_summarised <- s |>
+  summarise_scores(by = c("model", "range", "horizon"), metric = c("interval_score"))
+
+s_summarised |>
+  select(c(model, interval_score, range, horizon)) |>
+  pivot_wider(names_from = model, values_from = interval_score) |>
+  mutate(rel = `epiforecasts-EpiNow2` / `EuroCOVIDhub-ensemble`) |>
+  filter(horizon == 4)
+  ggplot(aes(y = rel, x = range, 
+             color = factor(horizon), group = factor(horizon))) +
+  geom_line()
+
+s_summarised |>
+  filter(horizon == 1) |>
+  select(c(model, interval_score, range, horizon)) |>
+  ggplot(aes(y = interval_score, x = range, 
+             color = model, group = model)) +
+  geom_line()
+
+```
+
 --- 

### 2022-11-21 main.tex

@@
 Forecasters, in expectation, optimise their score by providing a predictive distribution that is equal to the data-generating distribution, and will be penalised if they fail to do so.
-This paper focuses on one particular proper scoring rule, the Weighted Interval Score (WIS, formulas given in Section XXX in the SI) \citep{bracherEvaluatingEpidemicForecasts2021}, which is commonly used in epidemiology and other fields.
+This paper focuses on one particular proper scoring rule, the Weighted Interval Score (WIS, formulas given in Section \ref{wis} in the SI) \citep{bracherEvaluatingEpidemicForecasts2021}, which is commonly used in epidemiology and other fields.
 The WIS can be understood as an approximation of the Continuous Ranked Probability Score (CRPS) \citep{gneitingStrictlyProperScoring2007} for predictive distributions represented by a set of predictive quantiles.
-The WIS and the CRPS measure the absolute distance between the predictive distribution and the observed value and can be understood as a generalisation of the absolute error to full predictive distributions (see details in section XX in the SI).
+The WIS and the CRPS measure the absolute distance between the predictive distribution and the observed value and can be understood as a generalisation of the absolute error to full predictive distributions (see details in Sections \ref{wis} and \ref{crps} in the SI).
 For simplicity, we will use "absolute error" and "absolute distance" interchangeably, even though the former applies only to point forecasts and the latter is more appropriate for predictive distributions.
@@
 
-The forecasts used for the purpose of this illustrations are two-week-ahead forecasts submitted between February 8 2021 and October 18 2021 for reported cases and deaths from COVID-19.
-We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 16 forecast dates.
+ 2021 and October 18 2021 for reported cases and deaths from COVID-19.
+We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 16 forecast dates (see Figure \ref{fig:HUB-num-avail-models}.
 Where not otherwise stated, we report results for a two-week-ahead forecast horizon.
@@
 
-As expected on the natural scale, scores correlate strongly with the average number of observed cases and deaths (Spearman correlation: XX and XX) in a given country (see Fig \ref{fig:HUB-mean-locations}A, B).
-This relationship is not evident on the log scale, where correlation is much weaker and maybe slightly negative (Spearman correlation: XX and XX), implying larger scores for for countries with lower mean incidences.
+As expected on the natural scale, scores correlate strongly with the average number of observed cases and deaths (Spearman rank correlation: 0.96 and 0.97) in a given country (see Fig \ref{fig:HUB-mean-locations}A, B).
+On the log scale correlation is much weaker and slightly negative, at least in terms of rank correlation (Spearman correlation: -0.33 and -0.36), implying larger scores for for countries with lower mean incidences.
 
@@
 When evaluating performance \textit{across} different forecasts and forecast targets, relative skill scores of the models change considerably.
-The change is stronger for cases than for deaths (Figure \ref{fig:HUB-cors}XX) and tends to increase with increasing forecast horizon.
+The change is stronger for cases than for deaths (Figure \ref{fig:HUB-cors}C) and tends to increase with increasing forecast horizon.
 
-When comparing examples of forecasts on the natural scale with log-transformed ones (see Figures XX, XX, XX) a few interesting patterns emerge.
+When comparing examples of forecasts on the natural scale with log-transformed ones (see Figures \ref{fig:HUB-model-comparison-ensemble}, \ref{fig:HUB-model-comparison-baseline}, \ref{fig:HUB-model-comparison-epinow}) a few interesting patterns emerge.
 Missing the peak, i.e.
-predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts in May in Figure XXA, B).
+predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts in May in Figure \ref{fig:HUB-model-comparison-ensemble}A, B).
 On the log scale, these have less of an influence, as errors are smaller in relative terms (see panels C and D).
 We hypothesise that the larger divergence in scores for cases (compared to scores for deaths) displayed in Figure \ref{fig:HUB-pairwise}C is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths.
-Conversely, failure to predict an upswing while numbers are still low, is not severely punished on the natural scale (see forecasts in July in Figure XX A, B), as overall absolute errors are low.
+Conversely, failure to predict an upswing while numbers are still low, is less severely punished on the natural scale (see forecasts in July in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low.
 On the log scale, missing lower inflection points tends to lead to more severe penalties (see panels C and D).
-One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure XXE, F).
+One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure \ref{fig:HUB-model-comparison-ensemble}E, F).
 On the log scale, scores increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.
-
-Furthermore, including very low numbers in the confidence intervals may lead to very high scores on the log-scale.
-This may be an issue especially when forecasting small quantities.
-Figure \ref{fig:HUB-model-comparison-baseline} shows forecasts and scores for the baseline model, which had zero included in its 50 percent intervals for the forecasts in XX to XX, leading to excessive dispersion values on the log scale.
-Notwithstanding the general concern, one can argue that in the specific example (at least for cases) including zero in the prediction intervals was unreasonable and rightly penalised given that case numbers were in the thousands.
 
@@
 
-\paragraph{strengths and weaknesses}
-
 Applying a log transformation prior to the WIS means that forecasts are evaluated in terms of relative errors and errors on the multiplicative growth rate, rather than absolute errors.
@@
 A potential downside is that forecast evaluation is unreliable in situations where observed values are zero or very small.
+Including very small values in prediction intervals (see e.g.
+Figure \ref{fig:HUB-model-comparison-baseline}) can lead to excessive dispersion values on the log scale.
 Similarly, locations with lower incidences may get disproportionate weight (i.e.
@@
 It also gives higher weight to another type of situation one may care about, namely one in which numbers start to rise from a previously low level.
+
 While scores on the log scale can always be interpreted as a measure of relative (multiplicative) predictive performance, the interpretation as a score for a forecast of the multiplicative growth rate is only approximate.
@@
 Another promising transformation would be to divide each forecast by the forecast of the previous week (and observations by the observation of the previous week), in order to obtain forecasts for week-to-week growth rates.
-This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory (for a different approach to evaluation the shape of a forecast, see https://arxiv.org/abs/2209.04035).
+This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory (for a different approach to evaluation the shape of a forecast, see \citep{srivastavaShapebasedEvaluationEpidemic2022}).
 This, unfortunately is not well defined in the context of forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate, but may be an interesting approach if predictive samples are available.
@@
         \midrule
-    Cases & Observations & mean & 20308 & 8.39\\
-    Cases & Observations & sd & 43405 & 2.11\\
-    Cases & Observations & var & 1884002308 & 4.46\\
-    \addlinespace
-    Deaths & Observations & mean & 259 & 3.79\\
-    Deaths & Observations & sd & 532 & 2.11\\
-    Deaths & Observations & var & 283492 & 4.46\\
+        Cases & Observations & mean & 19941 & 8.33\\
+        Cases & Observations & sd & 43149 & 2.09\\
+        Cases & Observations & var & 1861850319 & 4.35\\
+        Deaths & Observations & mean & 222 & 3.65\\
+        Deaths & Observations & sd & 481 & 2.07\\
+        Deaths & Observations & var & 231025 & 4.29\\
         \addlinespace
@@
         Cases & WIS & sd & 17585 & 0.49\\
-    \addlinespace
         Deaths & WIS & mean & 28 & 0.22\\
@@
 \end{table}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-mean-scores-vs-total-log-log.png}
-    \caption{My current thinking is that the top row is helpful, as it explalins what the mean / var relationship is which is relevant for how much weight small countries get.
-I think the 2nd row is just completely useless.
-The 3rd and 4th row may be helpful.
-In some sense it is duplicated information from above, but it gives some nice intuiton for the correlations.
-I could alternatively also turn it into a correlation plot and add it to the other one.
-}
-    \label{fig:HUB-mean-scores-total-loglog}
-\end{figure}
 
@@
 
-NUMBER OF AVAILABLE FORECASTS
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/number-avail-forecasts.png}
+    \caption{
+    Number of forecasts available from different models for each forecast date. 
+    }
+    \label{fig:HUB-num-avail-models}
+\end{figure}
 
@@
     \caption{
-    Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-ensemble made in Germany.
+    Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-baseline made in Germany.
+The model had zero included in some of its 50 percent intervals (e.g.
+for case forecasts in July), leading to excessive dispersion values on the log scale.
+One could argue that including zero in the prediction intervals constituted an unreasonable forecast that was rightly penalised, but in general care has to be taken with small numbers.
 A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale.
@@
     \caption{
-    Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-ensemble made in Germany.
+    Forecasts and scores for two-week-ahead predictions from the epiforecasts-EpiNow2 model made in Germany.
 A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale.

### 2022-11-26 main.tex
Version before incorporating Seb's Feedback

@@
 
-Probabilistic forecasts play an important role in decision-making in epidemiology and public health \citep{reichCollaborativeMultiyearMultimodel2019, funkShorttermForecastsInform2020, cramerEvaluationIndividualEnsemble2021, bracherShorttermForecastingCOVID192021, europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}, as well as other areas as diverse as economics \citep{timmermannForecastingMethodsFinance2018, elliottForecastingEconomicsFinance2016} or meteorology \citep{gneitingWeatherForecastingEnsemble2005, kukkonenReviewOperationalRegionalscale2012}.
+Probabilistic forecasts \citep{heldProbabilisticForecastingInfectious2017} play an important role in decision-making in epidemiology and public health \citep{reichCollaborativeMultiyearMultimodel2019, funkShorttermForecastsInform2020, cramerEvaluationIndividualEnsemble2021, bracherShorttermForecastingCOVID192021, europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}, as well as other areas as diverse as economics \citep{timmermannForecastingMethodsFinance2018, elliottForecastingEconomicsFinance2016} or meteorology \citep{gneitingWeatherForecastingEnsemble2005, kukkonenReviewOperationalRegionalscale2012}.
 Epidemiological modelling in particular has received widespread attention during the COVID-19 pandemic.
-Evaluations of forecasts can provide feedback for forecasters and researchers to improve their models, and help decision makers distinguish good from bad predictions and choosing forecasters and models that should inform future decisions.
+Evaluations of forecasts can provide feedback for forecasters and researchers to improve their models and train ensembles, and help decision makers distinguish good from bad predictions and choosing forecasters and models that should inform future decisions.
 
-Probabilistic forecasts \citep{heldProbabilisticForecastingInfectious2017} are usually evaluated using so-called proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data.
-Forecasters, in expectation, optimise their score by providing a predictive distribution that is equal to the data-generating distribution, and will be penalised if they fail to do so.
-This paper focuses on one particular proper scoring rule, the Weighted Interval Score (WIS, formulas given in Section \ref{wis} in the SI) \citep{bracherEvaluatingEpidemicForecasts2021}, which is commonly used in epidemiology and other fields.
-The WIS can be understood as an approximation of the Continuous Ranked Probability Score (CRPS) \citep{gneitingStrictlyProperScoring2007} for predictive distributions represented by a set of predictive quantiles.
-The WIS and the CRPS measure the absolute distance between the predictive distribution and the observed value and can be understood as a generalisation of the absolute error to full predictive distributions (see details in Sections \ref{wis} and \ref{crps} in the SI).
+Probabilistic forecasts are usually evaluated using so-called proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data.
+Forecasters, in expectation, optimise their score by providing a predictive distribution that is equal to the data-generating distribution, and are penalised if they fail to do so.
+Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funk19_assessingperformancereal}}, and the Weighted Interval Score \citep[WIS,][formulas given in Section \ref{wis} in the SI]{bracherEvaluatingEpidemicForecasts2021} for predictive distributions represented by a set of predictive quantiles, used to assess forecasts in COVID-19 forecast hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
+The WIS and the (C)RPS measure the absolute distance between the predictive distribution and the observed value and can be understood as a generalisation of the absolute error to full predictive distributions (see details in Sections \ref{wis} and \ref{crps} in the SI).
 For simplicity, we will use "absolute error" and "absolute distance" interchangeably, even though the former applies only to point forecasts and the latter is more appropriate for predictive distributions.
-The WIS has been and is currently used by several Forecast Hubs for COVID-19 in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Flu Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
 
-Epidemiological processes are commonly understood and modeled to be inherently exponential in nature, as every infectious individual is assumed to lead to some number of additional infections [CITATION].
-Evaluating forecasts based on the absolute distance between forecast and observed value may therefore not be optimal in an epidemiological context, and it could be more appropriate to focus the evaluation on errors on the multiplicative growth rate, in order for the evaluation to better represent the underlying disease dynamics.
-Evaluating forecasts based on absolute, rather than relative errors, in practice also means that scores are strongly influenced by the overall order of magnitude of the quantity to forecast.
-This makes it difficult compare predictive performance across time, given that epidemiological processes may be non-stationary, or to compare performance across targets that are on very different orders of magnitudes (for example cases and hospitalisations) [CITE \citep{sherrattPredictivePerformanceMultimodel2022}, and others].
+Infectious processes are commonly modelled around the concept of the reproduction number, the average number of new cases generated by each case in the population.
+In the absence of changes in immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
+In that case, the number of new infections in the population grows exponentially in time~\citep{dushoff2021_speedstrengthepidemic}.
+This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellis21_challengescontrolcovid}.
+If case numbers are changing exponentially and the true modelling task revolves around estimating the reproduction number or the corresponding growth rate, evaluating forecasts based on the absolute distance between forecast and observed value rewards underprediction compared to overprediction by the same amount, as these errors grow exponentially with the error on the estimated reproduction number or growth rate.
+Instead, if one is to assess the ability of forecasts to assess and forecast the underlying infection dynamics, it may be more appropriate to evaluate errors on the growth rate directly.
 
-In this paper, we discuss how log transforming the forecasts and observations prior to applying the WIS (or CRPS) may be an appropriate way to address the aforementioned issues with evaluating forecasts based on absolute errors.
+Moreover, if forecasts based are evaluated on absolute rather than relative errors whereas in reality errors scale with the level of observations, scores are strongly influenced by the overall order of magnitude of the quantity to forecast.
+This makes it difficult compare predictive performance across time or to compare performance across targets that generally occur different orders of magnitudes.
+In the example of COVID-19 where case numbers underwent drastic changes over time following the initial emergence, forecast evaluation based on the absolute numbers would be dominated by performance around the peaks, and evaluation based on jointly evaluating the ability to predict numbers of cases, hospitalisation and deaths dominated by cases as they occur at largest numbers.
+
+In this paper, we discuss how log-transforming the forecasts and observations prior to applying the WIS (or CRPS) may be an appropriate way to address the aforementioned issues with evaluating forecasts based on absolute errors.
 We first outline the motivation for transforming forecasts prior to scoring (section [REF]).
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Log transforming forecasts and observations}
+\section{log-transforming forecasts and observations}
 
-The goal of log transforming forecasts and observations is to shift the focus of the evaluation in a way that may more appropriate for epidemiological forecasts, while preserving the propriety of the score.
-Instead of a score representing the magnitude of absolute errors this approach makes it possible to obtain a score which a) measures relative error and b) provides an approximate measure for how well a forecast captures the multiplicative growth rate of the target quantity.
+Our aim in log-transforming forecasts and observations is to shift the focus of the evaluation in a way that may more appropriate for epidemiological forecasts, while preserving the propriety of the score.
+Instead of a score representing the magnitude of absolute errors this approach makes it possible to obtain a score which a) measures relative error and at the same time b) provides an approximate measure for how well a forecast captures the multiplicative growth rate of the target quantity.
 Focusing on relative errors mitigates issues with scores based on absolute errors that are heavily influenced by the order of magnitude of the quantity to forecast.
-The interpretation of the log transformed score as an approximate score for a forecast of the growth rate of a given quantity aligns the evaluation with our understanding of how epidemiological processes should be modeled and the kinds of errors models would typically make [CITATION?].
-In this paper we focus on the log transformation, but note that many other transformations may be useful and appropriate (see section \ref{sec:discussion}) .
+The interpretation of the log-transformed score as an approximate score for a forecast of the growth rate of a given quantity aligns the evaluation with our understanding of how epidemiological processes should be modeled and the kinds of errors models would typically make [CITATION?].
+In this paper we focus on the log-transformation, but note that many other transformations may be useful and appropriate (see section \ref{sec:discussion}).
+We further focus on the WIS as evaluation metric, given its widespread use throughout the COVID-19 pandemic.
 
@@
 
-To illustrate the effects of applying the logarithm intuitively, let us consider a point forecast $\hat{y}_{t+1}$ for a quantity of interest $y_{t+1}$, such that 
+To illustrate the effects of applying the logarithm, let us consider a point forecast $\hat{y}_{t+1}$ for a quantity of interest $y_{t+1}$, such that 
 \begin{equation}
@@
 
-The analogon of the WIS / CRPS for point forecasts is the absolute error, and so the score will be determined by the size of the forecast error $\varepsilon_{t+1}$.
-When taking the logarithm of the forecast and the observation first, then the score is determined based on the size of $\varepsilon^*_{t+1}$:
+The analogue of the WIS / CRPS for point forecasts is the absolute error, and so the score will be determined by the size of the forecast error $\varepsilon_{t+1}$.
+When taking the logarithm of the forecast and the observation first, then the score is determined by $\varepsilon^*_{t+1}$:
 \begin{equation}
@@
 %
-It is easy to see that this corresponds to evaluating a multiplicative error on the original $y_{t+1}$:
+This corresponds to evaluating a multiplicative error on the original $y_{t+1}$:
 %
@@
 
-Note that the analogy between the WIS on the log scale and the absolute error of a point forecast on the log scale is slightly inaccurate: While the absolute error on the log scale is exactly symmetric (e.g., predicting 2 instead of 1 would give the same relative error as predicting 0.5), this is not entirely true for the WIS, which is computed for a set of prediction intervals.
-Due to the fact that upper and lower bounds of individual prediction intervals are affected slightly differently by the log transformations, small deviations may occur (compare e.g.
+The analogy between the WIS on the log scale and the absolute error of a point forecast on the log scale is slightly inaccurate: While the absolute error on the log scale is exactly symmetric (e.g., predicting 2 instead of 1 would give the same relative error as predicting 0.5), this is not entirely true for the WIS, which is computed for a set of prediction intervals.
+Due to the fact that upper and lower bounds of individual prediction intervals are affected slightly differently by the log-transformations, small deviations may occur (compare e.g.
 in Figure \ref{fig:change-in-scores}D the small difference between scores for an observed outcome of $\frac{1}{5}$ and 5).
@@
 
-Computing a score based on the logarithm of the forecast and the observation is also approximately equivalent to scoring a forecast of a multiplicative growth rate of the forecast target.
-The growth rate is defined as
+Computing a score based on the logarithm of the forecast and the observation is also approximately equivalent to scoring a forecast of a multiplicative growth rate of the forecast target, which approximates the more commonly used exponential growth rate.
+The multiplicative growth rate is defined as
 %
@@
 
-Now consider a forecast on $\log y_{t+1}$, where again, the score is determined by the size of $\varepsilon^*_{t+1}$.
+Now consider a forecast of $\log y_{t+1}$, where again, the score is determined by the size of $\varepsilon^*_{t+1}$.
 Using the fact that for small values of $x$, $\log (1+ x) \approx x$, we can see that scoring forecasts on the log scale is approximately equal to scoring an additive error on the growth rate (on the natural scale), as long as the growth rate is small.
@@
 \begin{align}
-\log y_{t+1}        &= \log \hat{y}_{t+1} + \varepsilon^*_{t+1} \\
+\log y_{t+1}        &= \log \hat{y}_{t+1} + \varepsilon^*_{t+1} \nonumber\\
                     &= \log ((1 + \hat{g}_{t+1}) \cdot y_{t}) + \varepsilon^*_{t+1} \\
-                    &= \log (1 + \hat{g}_{t, t+1}) + \log y_t + \varepsilon^*_{t+1} \Leftrightarrow \\
-\\
-\varepsilon^*_{t+1}  &= \log y_{t+1} -  \log y_t - \log (1 + \hat{g}_{t, t+1}) \\    
-                    &= \log (\frac{y_{t+1}}{y_t}) - \log (1 + \hat{g}_{t, t+1}) \\    
-                    &= \log (1 + g_{t, t+1}) - \log (1 + \hat{g}_{t, t+1}) \\   
-                    &\approx g_{t, t+1} - \hat{g}_{t, t+1} \quad \text{for small differences between } \hat{g}_{t, t+1} \text{ and } g_{t, t+1}. 
+                    &= \log (1 + \hat{g}_{t, t+1}) + \log y_t + \varepsilon^*_{t+1} \Leftrightarrow \nonumber \\
+\nonumber \\
+\varepsilon^*_{t+1}  &= \log y_{t+1} -  \log y_t - \log (1 + \hat{g}_{t, t+1}) \nonumber \\    
+                    &= \log \left(\frac{y_{t+1}}{y_t}\right) - \log (1 + \hat{g}_{t, t+1}) \\    
+                    &= \log (1 + g_{t, t+1}) - \log (1 + \hat{g}_{t, t+1})\nonumber  \\   
+                    &\approx g_{t, t+1} - \hat{g}_{t, t+1} \quad \text{for small differences between } \hat{g}_{t, t+1} \text{ and } g_{t, t+1}. \nonumber
 \end{align}
@@
 \subsection{Effects on model rankings}
-Rankings between different forecasters may change through the log transformation both when looking at aggregate scores as well as when looking at a single forecast.
+Rankings between different forecasters may change through the log-transformation both when looking at aggregate scores as well as when looking at a single forecast.
 As an example for the effect on a single forecast, consider two forecasters, A and B.
@@
 % PARAGRAPH NEEDS SOME WORK
-Overall model rankings usually differ even more when scores are averaged across multiple forecasts.
+Overall model rankings would be expected to differ even more when scores are averaged across multiple forecasts or targets.
 The change in rankings of aggregate scores is not mainly driven by the kinds of changes in model rankings for single forecasts just discussed.
@@
 Evaluating forecasts based on relative errors may help with comparing performance across different targets.
-However, log transforming forecasts and observations may also result in the opposite extreme, since small absolute errors (e.g.
+However, log-transforming forecasts and observations may also result in the opposite extreme because small absolute errors (e.g.
 predicting 9, rather than 3 deaths) may be large in relative terms if the quantity to forecast is very small.
-This is especially true when forecasts are restricted to discrete values.
+This can be exacerbated when forecasts are restricted to discrete values, where relative steps between possible values become greater closer to zero.
 
@@
 
-The log transformation is permissible (in the sense that it preserves propriety of the score), because it is a strictly monotonic transformation.
-Even though actual rankings between models may change, a forecaster will in expectation still minimise her score if she reports a predictive distribution that is equal to the data-generating distribution.
-The order of the operations matters, and applying the log transformation after scores have been computed, i.e.
+The log-transformation is permissible (in the sense that it preserves propriety of the score), because it is a strictly monotonic transformation.
+Even though actual rankings between models may change, forecasters will  still minimise her score if they report a predictive distribution that is equal to the data-generating distribution.
+The order of the operations matters, and applying the log-transformation after scores have been computed, i.e.
 computing $\log S(F, y)$ results in an improper score.
@@
 
-In practice, one issue with the log transformation is that it cannot readily be applied to negative numbers or zero values.
+In practice, one issue with the log-transformation is that it cannot readily be applied to negative numbers or zero values.
 In addition, it may difficult to reliably evaluate relative errors when dealing with small discrete values.
@@
 One solution to deal with zeroes and small values is to also exclude those from ongoing analysis.
-An alternative is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm as is common practice when log transforming counts.
+An alternative is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm as is common practice when log-transforming counts.
 This represents a strictly monotonic transformation and therefore preserves propriety of the resulting score.
@@
 As an empirical example for evaluating forecasts on the natural and on the log scale we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
-The European COVID-19 Forecast Hub is one of several COVID-19 Forecast hubs (together with a Hub in the US \citep{cramerEvaluationIndividualEnsemble2021} and one in Germany and Poland \citep{bracherShorttermForecastingCOVID192021} that systematically collects, aggregated and evaluated forecasts of several COVID-19 targets created by different teams every week.
+The European COVID-19 Forecast Hub is one of several COVID-19 Forecast hubs (together with a Hub in the US \citep{cramerEvaluationIndividualEnsemble2021} and one in Germany and Poland \citep{bracherShorttermForecastingCOVID192021} that systematically collects, aggregates and evaluates forecasts of several COVID-19 targets created by different teams every week.
 Forecasts are made one to four weeks ahead into the future and follow a quantile-based format with a set of 22 quantiles plus the median ($0.01, 0.025, 0.05, ..., 0.5, ...
@@
 
-In addition to the WIS we use pairwise comparisons \citep{cramerEvaluationIndividualEnsemble2021} to evaluate models against each other.
+In addition to the WIS we use pairwise comparisons \citep{cramerEvaluationIndividualEnsemble2021} to evaluate the relative performance of models across countries in the presence of missingness.
 In a first step, score ratios are computed for all pairs of models by taking the set of overlapping forecasts between the two models and dividing the score of one model over the score achieved by the other model.
@@
 Across the dataset, the average number of observed cases and deaths varied considerably by location and target type (see Figure \ref{fig:HUB-mean-locations}A and B).
-On the natural scale, scores also varied considerably across targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}C).
-On the log scale, scores were more evenly distributed between targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}).
-Both on the natural scale as well as on the log scale, scores increased considerably with increasing forecast horizon (see Figure \ref{fig:HUB-mean-locations}E).
-The increase was more pronounced for cases than for deaths.
+On the natural scale, scores also varied considerably across targets (see Figure\ref{fig:HUB-mean-locations}D) and locations (see Figure\ref{fig:HUB-mean-locations}C).
+On the log scale, scores were more evenly distributed between targets (see Figure\ref{fig:HUB-mean-locations}D) and locations (see Figure\ref{fig:HUB-mean-locations}C).
+Both on the natural scale as well to a slightly lesser extent on the log scale, scores increased considerably with increasing forecast horizon (see Figure \ref{fig:HUB-mean-locations}E).
 We hypothesise that this increase in scores across horizons represents the relative difficulty to forecast values into the future, which was noticeable both on the natural as well as the logarithmic scale.
@@
 
-As expected on the natural scale, scores correlate strongly with the average number of observed cases and deaths (Spearman rank correlation: 0.96 and 0.97) in a given country (see Fig \ref{fig:HUB-mean-locations}A, B).
-On the log scale correlation is much weaker and slightly negative, at least in terms of rank correlation (Spearman correlation: -0.33 and -0.36), implying larger scores for for countries with lower mean incidences.
+As expected on the natural scale, scores correlate strongly with the average number of observed cases and deaths (Spearman rank correlation: 0.96 and 0.97) in a given country (see Figure\ref{fig:HUB-mean-locations}A, B).
+On the log scale, the correlation is much weaker and slightly negative, at least in terms of rank correlation (Spearman correlation: -0.33 and -0.36), implying a tendency towards larger scores in countries with lower mean incidences.
 
@@
 
-For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}).
-When evaluating performance \textit{across} different forecasts and forecast targets, relative skill scores of the models change considerably.
-The change is stronger for cases than for deaths (Figure \ref{fig:HUB-cors}C) and tends to increase with increasing forecast horizon.
+For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}B).
+When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models change considerably.
+The change is stronger for cases than for deaths (Figure \ref{fig:HUB-cors}C) and possibly tends to increase with increasing forecast horizon.
 
-When comparing examples of forecasts on the natural scale with log-transformed ones (see Figures \ref{fig:HUB-model-comparison-ensemble}, \ref{fig:HUB-model-comparison-baseline}, \ref{fig:HUB-model-comparison-epinow}) a few interesting patterns emerge.
+When comparing examples of forecasts on the natural scale with those on the log scale (see Figures \ref{fig:HUB-model-comparison-ensemble}, \ref{fig:HUB-model-comparison-baseline}, \ref{fig:HUB-model-comparison-epinow}) a few interesting patterns emerge.
 Missing the peak, i.e.
 predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts in May in Figure \ref{fig:HUB-model-comparison-ensemble}A, B).
-On the log scale, these have less of an influence, as errors are smaller in relative terms (see panels C and D).
-We hypothesise that the larger divergence in scores for cases (compared to scores for deaths) displayed in Figure \ref{fig:HUB-pairwise}C is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths.
+On the log scale, these have less of an influence, as errors are smaller in relative terms (see \ref{fig:HUB-model-comparison-ensemble}C, D).
+We hypothesise that the lower correlation between skill scores for cases (compared to the correlation of scores for deaths) as displayed in Figure \ref{fig:HUB-cors}C is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths, and therefore relative skill scores are more strongly affected for cases.
 Conversely, failure to predict an upswing while numbers are still low, is less severely punished on the natural scale (see forecasts in July in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low.
-On the log scale, missing lower inflection points tends to lead to more severe penalties (see panels C and D).
+On the log scale, missing lower inflection points tends to lead to more severe penalties (see Figure \ref{fig:HUB-model-comparison-ensemble}C, D)).
 One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure \ref{fig:HUB-model-comparison-ensemble}E, F).
-On the log scale, scores increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.
+On the log scale, scores do not exhibit this behaviour and rather increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.
 
@@
     \caption{Changes in model ratings as measured by relative skill for two-week-ahead predictions. Top row: cases, bottom row: deaths. NEED TO CHANGE THE ANNOTATION LEVELS.}
-    \label{fig:HUB-pairwise}
+    \label{fig:HUB-rank-order}
 \end{figure}
 
-Evaluating the rank order of forecasts from the ECDC forecast hub we saw that scoring after a log transform altered the order of forecast models.
-In particular, we note that models that are known to not make use of a exponential error model and to be biased towards under-prediction, such the the MUNI-ARMA model were ranked lower when using a log transform.
-A submission by the authors of this study, the epiforecasts-EpiNow2 forecast model, was the most impacted by the change to log scale scoring moving from last place for reported cases to 4th place.
-This makes sense given that it explicitly models errors on an exponential scale, and is relatively unbiased meaning it may be over-penalised due to over-prediction during periods of peak incidence.
-Interestingly, other models that are growth rate based, such as the LANL-GrowthRate forecaster, do less well after a log-transform.
-Encouragingly for the ECDC forecast Hub the Hub ensemble, which is the forecast the organisers suggest forecast consumers make use of, remains one of the top forecasts across scoring schemes and indeed ranks higher for reported cases when a log transform is used.
+The rank order of different forecasting models (see Figure \ref{fig:HUB-rank-order} indicates that models that do not make use of a exponential error model and thus would be biased towards under-prediction during periods of exponential growth, such as the MUNI-ARMA model, were ranked lower when using a log-transform.
+The epiforecasts-EpiNow2 forecast model, on the other hand~(created by co-authors of this study), which makes forecasts based on the behaviour of the reproduction number and thus implicitly uses an exponential error model, was the most impacted by the change to log-scale scoring moving from last place for reported cases to 4th place.
+Other models that are based on modelling the growth rate, such as the LANL-GrowthRate forecaster, do less well after a log-transform, indicating that this potential association between model structure and score under different approximations does not hold strictly.
+Encouragingly for the European Forecast Hub, the Hub ensemble, which is the forecast the organisers suggest forecast consumers make use of, remains one of the top forecasts across scoring schemes and indeed ranks higher for reported cases when a log-transform is used.
 
@@
 
-In this paper we proposed a log transformation of forecasts and observed values prior to evaluation with the WIS (or CRPS) in order to address issues that arise when evaluating epidemiological forecasts based on a measure of absolute error.
-We gave a mathematical intuition for the transformation and showed that scores on the log scale can be interpreted as a) a measure of relative (multiplicative) prediction errors, as well as b) as an approximate score for a forecast of the multiplicative growth rate of the target quantity.
-When applying this approach to forecasts from the European COVID-19 Forecast Hub, we found overall scores to be more equal across, time, location and target type (cases, deaths).
-Scores on the log scale were much less influenced by the overall incidence level in a country, and showed a slight tendency to be higher in locations with a very incidences.
-We found that model rankings changed noticeably, favouring models like epiforecasts-EpiNow2 which explicitly has an exponential error structure, and disfavouring models like MUNI-ARIMA which is purely statistical model with a linear error structure.
+In this paper we proposed a log-transformation of forecasts and observed values prior to evaluation with the WIS (or CRPS) in order to address issues that arise when evaluating epidemiological forecasts based on a measure of absolute error.
+We gave a mathematical intuition for the transformation and showed that scores on the log scale can be interpreted as a) a measure of relative (multiplicative) prediction errors, as well as b) an approximate score for a forecast of the multiplicative growth rate of the target quantity.
+When applying this approach to forecasts from the European COVID-19 Forecast Hub, we found overall scores on the log scale to be more equal across, time, location and target type (cases, deaths) than scores on the natural scale.
+Scores on the log scale were much less influenced by the overall incidence level in a country, and showed a slight tendency to be higher in locations with very low incidences.
+We found that model rankings changed noticeably, favouring models like epiforecasts-EpiNow2 which explicitly has an exponential error structure, and disfavouring models like MUNI-ARIMA which is a purely statistical model with a linear error structure.
 On the natural scale, missing the peak and overshooting was more severely penalised, while missing the nadir and a following upswing in numbers tended to be more severely penalised on the log scale than on the natural scale.
 
-Applying a log transformation prior to the WIS means that forecasts are evaluated in terms of relative errors and errors on the multiplicative growth rate, rather than absolute errors.
+Applying a log-transformation prior to the WIS means that forecasts are evaluated in terms of relative errors and errors on the multiplicative growth rate, rather than absolute errors.
 The most important strength of this approach is that the evaluation better accommodates the exponential nature of epidemiological process and the types of errors forecasters who accurately model those processes are expected to make.
-The log transformation also helps avoid issues with scores being strongly influenced by the order of magnitude of the forecast quantity, which can be seen when evaluating forecasts on the natural scale.
+The log-transformation also helps avoid issues with scores being strongly influenced by the order of magnitude of the forecast quantity, which can be an issue when evaluating forecasts on the natural scale.
 A potential downside is that forecast evaluation is unreliable in situations where observed values are zero or very small.
@@
 \cite{bracherEvaluatingEpidemicForecasts2021} argue that the large weight given to forecasts for locations with high incidences is a desirable property, as it reflects performance on the targets we should care about most.
-On the other hand, scoring forecasts on the log scale may be less influenced by single outliers and better reflect consistent performance across time, space, and forecast targets.
+On the other hand, scoring forecasts on the log scale may be less influenced by outliers and better reflect consistent performance across time, space, and forecast targets.
 It also gives higher weight to another type of situation one may care about, namely one in which numbers start to rise from a previously low level.
-
 While scores on the log scale can always be interpreted as a measure of relative (multiplicative) predictive performance, the interpretation as a score for a forecast of the multiplicative growth rate is only approximate.
@@
 
-It is important to note that the log transformation is only one of many transformation that may be useful and appropriate in an epidemiological context.
-One easy transformation is to divide all forecasts by the population in a location, in order to obtain forecasts standardised incidences per e.g.
+The log-transformation is only one of many transformation that may be useful and appropriate in an epidemiological context.
+One easy transformation is to divide all forecasts by the population in a location, in order to obtain forecasts of standardised incidences per e.g.
 100,000 inhabitants.
-A second transformation was mentioned above, namely converting forecasts into forecasts for the multiplicative growth rate by dividing by the last observed value.
+A second transformation was mentioned above, namely converting forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last observed value.
 One could also think about applying a Box-Cox or a square-root transformation in order to stabilise the variance of the forecasts.
-Another promising transformation would be to divide each forecast by the forecast of the previous week (and observations by the observation of the previous week), in order to obtain forecasts for week-to-week growth rates.
-This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory (for a different approach to evaluation the shape of a forecast, see \citep{srivastavaShapebasedEvaluationEpidemic2022}).
+Another promising transformation would be to divide each forecast by the forecast of the previous week (and observations by the observation of the previous week), in order to obtain forecasts for week-to-week multiplicative growth rates, or alternatively to ask forecasters to provide estimates of the weekly relative changed applied to the latest data and subsequent forecast points directly.
+This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluation the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}.
 This, unfortunately is not well defined in the context of forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate, but may be an interesting approach if predictive samples are available.
@@
 Exploring these different transformations is a promising avenue for future work that could help bridge the gap between modellers and policy makers by providing scoring rules that better reflect what forecast consumers actually care about.
-We have showed that the log transformation can lead to significant changes in the relative rankings of models against each other, with potentially important implications for decision makers who rely on the knowledge of past performance to make a decision about which forecasts should inform future decisions.
-We think it would therefore be a good idea to show different versions of the WIS in order to obtain a richer picture of model performance.
-While we have provided a small case study, more work needs to be done to better understand the effects of applying a log transformation in different contexts and how it might affect decision making.
+We have showed that the log-transformation can lead to significant changes in the relative rankings of models against each other, with potentially important implications for decision makers who rely on the knowledge of past performance to make a decision about which forecasts should inform future decisions.
+While it is commonly accepted that multiple proper scoring rules should usually be considered when comparing forecasts, we think this should be supplemented by considering different transformations of the data to obtain a richer picture of model performance.
+While we have produced this work as an initial case study, more work needs to be done to better understand the effects of applying a logarithmic or other transformations more comprehensively and in different contexts, and how it might affect decision making.
 
@@
 
-%Whether or not a log transformation is appropriate depends on whether one believes that differences in absolute values are meaningful.
+%Whether or not a log-transformation is appropriate depends on whether one believes that differences in absolute values are meaningful.
 This may be the case for example when comparing forecasts across locations or time points, rather than different forecasts targets.

### 2022-11-29 main.tex
add abstract and authors, update paper after feedback

@@
 
-\title{To log or not to log}
-  \author{Anonymous Alpaca\thanks{All Alpaca friends} $^{,}$\thanks{The Zoo} $^{ , *}$}
+\title{Transformations of forecasts for evaluation in an epidemiological context}
+  \author{Nikos I.
+Bosse\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene & Tropical Medicine, London, United Kingdom} $^{,}$\thanks{Centre for the Mathematical Modelling of Infectious Diseases, London, United Kingdom} $^{,*}$,
+
+  Sam Abbott\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$,
+
+  Anne Cori\thanks{MRC Centre for Outbreak Analysis and Modelling, Department of Infectious Disease Epidemiology, School of Public Health, Imperial College London, London, United Kingdom} $^{}$, \\
+
+  Edwin van Leeuwen\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene & Tropical Medicine, London, United Kingdom} $^{,}$\thanks{UK Health Security Agency, London, United Kingdom} $^{}$,
+
+  Johannes Bracher\thanks{Institute of Economic Theory and Statistics, Karlsruhe Institute of Technology, Karlsruhe, Germany} $^{}$,
+
+  Sebastian Funk\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$
+  }
 
 \maketitle
-
 %\tableofcontents
-
 \begin{abstract}
-In the abstract, this article is a very good one.
-We did a very good job of the research and got all the best results.
-People told us we were mad and we said no we are effective.
-Now gather around and listen to our thoughts you will be, shocked, scared, amazed, and astounded.
+Forecast evaluation plays an essential role in improving forecasting models can help inform future decision making.
+In epidemiology, forecasts are often scored using the Weighted Interval Score (WIS) or the Continuous Ranked Probability Score (CRPS), which are a measure of absolute distance between forecast and observation.
+Infectious disease processes are commonly described in terms of the growth rate or the effective reproduction number.
+Scoring based on the absolute distance between forecast and observation may not reflect the true underlying modelling task well (estimating the growth rate or reproduction number) and makes it difficult to compare forecasts across time, space, and different targets.
+As a solution we propose transforming the data before scoring the forecasts.
+Depending on the context, many transformations may be appropriate.
+We describe the log-transformation as one specific example in detail and discuss its interpretation as a measure of relative error and an approximate score for a forecast of the growth rate.
+Applying the log-transformation to data from the European COVID-19 Forecast Hub, we found that it changed model rankings, evened scores across time, locations and target types, and emphasised situations in which models missed lower inflection points and consequent rises in numbers, while penalising missed peaks less severely.
 \end{abstract}
@@
 
-{\footnotesize $^*$ Correspondence to: Anonymous Alpaca (\url{anonymous@alpaca.com}))}
+{\footnotesize $^*$ Correspondence to \url{nikos.bosse@lshtm.ac.uk})}
 
@@
 Probabilistic forecasts are usually evaluated using so-called proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data.
-Forecasters, in expectation, optimise their score by providing a predictive distribution that is equal to the data-generating distribution, and are penalised if they fail to do so.
-Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funk19_assessingperformancereal}}, and the Weighted Interval Score \citep[WIS,][formulas given in Section \ref{wis} in the SI]{bracherEvaluatingEpidemicForecasts2021} for predictive distributions represented by a set of predictive quantiles, used to assess forecasts in COVID-19 forecast hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
-The WIS and the (C)RPS measure the absolute distance between the predictive distribution and the observed value and can be understood as a generalisation of the absolute error to full predictive distributions (see details in Sections \ref{wis} and \ref{crps} in the SI).
-For simplicity, we will use "absolute error" and "absolute distance" interchangeably, even though the former applies only to point forecasts and the latter is more appropriate for predictive distributions.
+Proper scoring rules are constructed such that forecasters are incentivised to report their true belief about the future.
+Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funk19_assessingperformancereal}, and the Weighted Interval Score \citep[WIS,][formulas given in Section \ref{wis} in the SI]{bracherEvaluatingEpidemicForecasts2021}.
+The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in COVID-19 forecast hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
+The WIS and the CRPS measure the absolute distance between the predictive distribution and the observed value and can be understood as a generalisation of the absolute error to predictive distributions (see details in Sections \ref{wis} and \ref{crps} in the SI).
+For simplicity, we will use "absolute error" and "absolute distance" interchangeably, even though the former is more commonly used for point forecasts and the latter is more appropriate for predictive distributions.
 
-Infectious processes are commonly modelled around the concept of the reproduction number, the average number of new cases generated by each case in the population.
+Infectious processes are commonly modelled around the concept of the reproduction number \citep{gosticPracticalConsiderationsMeasuring2020}, the average number of new cases generated by each case in the population.
 In the absence of changes in immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
@@
 This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellis21_challengescontrolcovid}.
-If case numbers are changing exponentially and the true modelling task revolves around estimating the reproduction number or the corresponding growth rate, evaluating forecasts based on the absolute distance between forecast and observed value rewards underprediction compared to overprediction by the same amount, as these errors grow exponentially with the error on the estimated reproduction number or growth rate.
-Instead, if one is to assess the ability of forecasts to assess and forecast the underlying infection dynamics, it may be more appropriate to evaluate errors on the growth rate directly.
+If case numbers are growing exponentially (for an effective reproduction number greater than 1) and the true modelling task revolves around estimating the reproduction number or the corresponding growth rate, evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number) less than overprediction by the same amount, as errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
+If one is to assess the ability of forecasts to assess and forecast the underlying infection dynamics, it may be more appropriate to evaluate errors on the growth rate directly.
 
-Moreover, if forecasts based are evaluated on absolute rather than relative errors whereas in reality errors scale with the level of observations, scores are strongly influenced by the overall order of magnitude of the quantity to forecast.
-This makes it difficult compare predictive performance across time or to compare performance across targets that generally occur different orders of magnitudes.
+Moreover, if forecasts are evaluated based on absolute rather than relative errors, whereas in reality errors scale with the level of observations, scores are strongly influenced by the overall order of magnitude of the quantity to forecast.
+This makes it difficult to compare predictive performance across time or to compare performance across targets that generally occur at different orders of magnitudes.
 In the example of COVID-19 where case numbers underwent drastic changes over time following the initial emergence, forecast evaluation based on the absolute numbers would be dominated by performance around the peaks, and evaluation based on jointly evaluating the ability to predict numbers of cases, hospitalisation and deaths dominated by cases as they occur at largest numbers.
 
-In this paper, we discuss how log-transforming the forecasts and observations prior to applying the WIS (or CRPS) may be an appropriate way to address the aforementioned issues with evaluating forecasts based on absolute errors.
-We first outline the motivation for transforming forecasts prior to scoring (section [REF]).
-We then provide a mathematical intuition of the impact of scoring on the log scale and its interpretation (section [REF]).
-We illustrate our findings on a series of examples using forecasts submitted to the European COVID-19 Forecast Hub  \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
-Finally, we give scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues on the use of transformed scores in the evaluation of infectious disease forecasts.
+In this paper, we discuss how transforming the forecasts and observations prior to applying the WIS (or CRPS) may be an appropriate way to address the aforementioned issues.
+Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects of the forecasts that forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).
+For conceptual clarity and to allow for a more in-depth discussion, we focus most of our attention on the log-transformation as one example of many possible transformations, chosen for its simplicity and usefulness in an epidemiological context.
+We further focus on the WIS as evaluation metric, given its widespread use throughout the COVID-19 pandemic.
+
+We first outline the motivation for transforming forecasts prior to scoring and specifically the motivation for applying a log-transformation (Section \ref{sec:methods}).
+We provide a mathematical intuition of the impact of the log-transformation, its interpretation (Section \ref{sec:methods:relative} and \ref{sec:methods:growthrate}) and effects on forecast rankings (section \ref{sec:methods:rankings} and provide some discuss some practical considerations relevant for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}).
+We illustrate our the effect of the log-transformation using forecasts submitted to the European COVID-19 Forecast Hub  \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022} (Section \ref{sec:results}).
+Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues on the use of transformed scores in the evaluation of infectious disease forecasts (Section \ref{sec:discussion}).
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{log-transforming forecasts and observations}
+\section{Transforming forecasts and observations}
+\label{sec:methods}
 
-Our aim in log-transforming forecasts and observations is to shift the focus of the evaluation in a way that may more appropriate for epidemiological forecasts, while preserving the propriety of the score.
-Instead of a score representing the magnitude of absolute errors this approach makes it possible to obtain a score which a) measures relative error and at the same time b) provides an approximate measure for how well a forecast captures the multiplicative growth rate of the target quantity.
-Focusing on relative errors mitigates issues with scores based on absolute errors that are heavily influenced by the order of magnitude of the quantity to forecast.
-The interpretation of the log-transformed score as an approximate score for a forecast of the growth rate of a given quantity aligns the evaluation with our understanding of how epidemiological processes should be modeled and the kinds of errors models would typically make [CITATION?].
-In this paper we focus on the log-transformation, but note that many other transformations may be useful and appropriate (see section \ref{sec:discussion}).
-We further focus on the WIS as evaluation metric, given its widespread use throughout the COVID-19 pandemic.
+Our aim in transforming forecasts and observations is to shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score.
+Instead of a score representing the magnitude of absolute errors this approach makes it possible to obtain a score which a) measures relative error and at the same time b) provides an approximate measure for how well a forecast captures the multiplicative growth rate of the target quantity (see Section \ref{sec:methods:growthrate}.
+The interpretation of the log-transformed score as an approximate score for a forecast of the growth rate of a given quantity aligns the evaluation with our understanding of how epidemiological processes should be modeled and the kinds of errors models would typically make.
 
 \subsection{Interpretation as a score based on multiplicative (relative) errors}
+\label{sec:methods:relative}
 
@@
 
-The analogy between the WIS on the log scale and the absolute error of a point forecast on the log scale is slightly inaccurate: While the absolute error on the log scale is exactly symmetric (e.g., predicting 2 instead of 1 would give the same relative error as predicting 0.5), this is not entirely true for the WIS, which is computed for a set of prediction intervals.
+The analogy between the WIS on the log scale and the absolute error of a point forecast on the log scale is slightly inaccurate: while the absolute error on the log scale is exactly symmetric (e.g., predicting 2 instead of 1 would give the same relative error as predicting 0.5), this is not entirely true for the WIS, which is computed for a set of prediction intervals.
 Due to the fact that upper and lower bounds of individual prediction intervals are affected slightly differently by the log-transformations, small deviations may occur (compare e.g.
@@
 \subsection{Interpretation as an approximate score of the growth rate}
+\label{sec:methods:growthrate}
 
 Computing a score based on the logarithm of the forecast and the observation is also approximately equivalent to scoring a forecast of a multiplicative growth rate of the forecast target, which approximates the more commonly used exponential growth rate.
-The multiplicative growth rate is defined as
+The multiplicative growth rate can be defined as
 %
@@
                     &= \log (1 + g_{t, t+1}) - \log (1 + \hat{g}_{t, t+1})\nonumber  \\   
-                    &\approx g_{t, t+1} - \hat{g}_{t, t+1} \quad \text{for small differences between } \hat{g}_{t, t+1} \text{ and } g_{t, t+1}. \nonumber
+                    &\approx g_{t, t+1} - \hat{g}_{t, t+1} \quad \text{for small } \hat{g}_{t, t+1} \text{ and } g_{t, t+1}. \nonumber
 \end{align}
@@
 \subsection{Effects on model rankings}
-Rankings between different forecasters may change through the log-transformation both when looking at aggregate scores as well as when looking at a single forecast.
+\label{sec:methods:rankings}
+Rankings between different forecasters based on their scores may change through the log-transformation both when looking at aggregate scores as well as when looking at a single forecast.
 As an example for the effect on a single forecast, consider two forecasters, A and B.
@@
 Forecaster A would receive a score of $2 - 0.5$ = 1.5, beating forecaster B who would receive a score of $2.5 - 0.9 = 1.6$.
-When we transform the forecast, forecaster A receives a score of $\log (\frac{2}{0.5}) = 1.39$, while forecaster B obtains a score of $\log (\frac{2.6}{0.9}) = 1.06$ and now wins over forecaster A.
+When we transform the forecast, forecaster A receives a score of $\log (\frac{2}{0.5}) = 1.39$, while forecaster B obtains a score of $\log (\frac{2.5}{0.9}) = 1.02$ and now wins over forecaster A.
 The main difference between forecaster A and B is that forecaster A has issued a prediction interval ([0.5, 2]) that was more symmetric around the true value than the interval issued by forecaster B ([0.9.
 2.6], which was more skewed).
-This means that in relative terms, Forecaster B was noticeably closer to the true value with the lower bound of her prediction interval, while only being slightly further away with the upper interval.
+This means that in relative terms, Forecaster B was noticeably closer to the true value with the lower bound of their prediction interval, while only being slightly further away with the upper interval.
 This implies that scores on the log scale may treat (one-sided) outlier forecasts more leniently.
@@
 Rather, as observations from exponential processes are likely to be on very different absolute scales, large observations will account for the majority of absolute score but will not dominate relative scores in the same way.
-When aggregating scores across metrics with different scales, for example reported cases, and hospitalisations, even an ideal (i.e.
-perfect) forecaster will make much larger errors when predicting case numbers than when predicting hospitalisations.
 Evaluating forecasts based on relative errors may help with comparing performance across different targets.
@@
 predicting 9, rather than 3 deaths) may be large in relative terms if the quantity to forecast is very small.
-This can be exacerbated when forecasts are restricted to discrete values, where relative steps between possible values become greater closer to zero.
+This can be exacerbated when forecasts are restricted to discrete values, where relative steps between possible values become greater for values closer to zero.
 
@@
 Whether or not forecast targets with smaller quantities receive higher scores than those for targets with larger quantities, depends on the exact relationship between the mean and the variance of the quantity of interest.
-For simulated count data following a negative binomial distribution, for example, log predictions for small quantities receive on average higher scores than forecasts for large quantities if the mean and the variance grow at the same rate $\sigma^2 = \mu$, about equal scores when the variance grows at a rate of $\sigma^2 = \mu + \mu^2$ (Figure \ref{fig:SIM-wis-state-size-mean}), and smaller scores when the variance grows faster than that.
+To provide a better intuition, we simulated count data following a negative binomial distribution.
+For these simulated data, log predictions for small quantities received on average higher scores than forecasts for large quantities if the mean and the variance grew at the same rate $\sigma^2 = \mu$, about equal scores when the variance grew at a rate of $\sigma^2 = \mu + \mu^2$, and smaller scores when the variance grew faster than that (Figure \ref{fig:SIM-wis-state-size-mean}).
 % To illustrate this with count data we sampled forecasts from different negative binomial distributions.
@@
 \subsection{Practical considerations}
+\label{sec:methods:considerations}
 
-The log-transformation is permissible (in the sense that it preserves propriety of the score), because it is a strictly monotonic transformation.
-Even though actual rankings between models may change, forecasters will  still minimise her score if they report a predictive distribution that is equal to the data-generating distribution.
-The order of the operations matters, and applying the log-transformation after scores have been computed, i.e.
-computing $\log S(F, y)$ results in an improper score.
+The log-transformation (among many other transformations) is permissible (in the sense that it preserves propriety of the score), because it is a strictly monotonic transformation.
+Even though actual rankings between models may change, forecasters will in expectation still minimise their score if they report a predictive distribution that is equal to the data-generating distribution.
+The order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee propriety.
+In the specific example, taking the logarithm of the scores, rather than scoring the log-transformed forecasts and data, results in an improper score.
 This is because taking the logarithm of the WIS / CRPS results in a score that does not penalise outliers enough and therefore incentivises overconfident predictions.
@@
     \includegraphics[width=0.99\textwidth]{output/figures/example-log-first.png}
-    \caption{Scores for different forecasts evaluated the WIS and the log WIS.
-We simulated 1000 observations $Y_i = {\rm e}^{x_i}$, with $x_i \text{iid} \sim \mathcal{N}(0, 1)$.
+    \caption{Scores for different forecasts evaluated using the WIS and the log WIS.
+We simulated 1000 observations $Y_i = {\rm e}^{x_i}$, with $x_i \text{ iid} \sim \mathcal{N}(0, 1)$.
 We then simulated 20 forecasters who would issue a predictive distribution $F = {\rm e}^{x_i}$, with $x \sim \mathcal{N}(0, \sigma)$, with values of $\sigma$ ranging from 0.1 to 2.}
@@
 In practice, one issue with the log-transformation is that it cannot readily be applied to negative numbers or zero values.
-In addition, it may difficult to reliably evaluate relative errors when dealing with small discrete values.
+In addition, it may be difficult to reliably evaluate relative errors when dealing with small discrete values.
 In an epidemiological setting that involves count data, negative values should likely be omitted entirely.
@@
 \section{Empirical example: the European Forecast Hub}
+\label{sec:results}
 
 As an empirical example for evaluating forecasts on the natural and on the log scale we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
-The European COVID-19 Forecast Hub is one of several COVID-19 Forecast hubs (together with a Hub in the US \citep{cramerEvaluationIndividualEnsemble2021} and one in Germany and Poland \citep{bracherShorttermForecastingCOVID192021} that systematically collects, aggregates and evaluates forecasts of several COVID-19 targets created by different teams every week.
+The European COVID-19 Forecast Hub is one of several COVID-19 Forecast hubs (together with a Hub in the US \citep{cramerEvaluationIndividualEnsemble2021} and one in Germany and Poland \citep{bracherShorttermForecastingCOVID192021}) which have been systematically collecting, aggregating and evaluating forecasts of several COVID-19 targets created by different teams every week.
 Forecasts are made one to four weeks ahead into the future and follow a quantile-based format with a set of 22 quantiles plus the median ($0.01, 0.025, 0.05, ..., 0.5, ...
@@
 
- 2021 and October 18 2021 for reported cases and deaths from COVID-19.
-We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 16 forecast dates (see Figure \ref{fig:HUB-num-avail-models}.
+The forecasts used for the purpose of this illustration are forecasts submitted between March 8 2021 and October 18 2021 for reported cases and deaths from COVID-19.
+We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 16 forecast dates (see Figure \ref{fig:HUB-num-avail-models}).
 Where not otherwise stated, we report results for a two-week-ahead forecast horizon.
@@
 In a first step, score ratios are computed for all pairs of models by taking the set of overlapping forecasts between the two models and dividing the score of one model over the score achieved by the other model.
-The relative skill for a given model is then obtained by taking the geometric mean of all score ratios which involve that model.
+The relative skill for a given model compared to others is then obtained by taking the geometric mean of all score ratios which involve that model.
 Low values are better, and the average model receives a relative skill score of 1.
@@
     \caption{Observations and scores across locations and forecast horizons.
-A, Average (across all time points) of observed cases and deaths for different locations.
-B: Boxplot (y-axis on log-scale) of all cases and deaths.
+A: Average (across all time points) of observed cases and deaths for different locations.
+B: Corresponding boxplot (y-axis on log-scale) of all cases and deaths.
 C: Scores for two-week-ahead forecasts from the EuroCOVIDhub-ensemble (averaged across all forecast dates) for different locations, evaluated on the natural as well as the logarithmic scale.
-D: Corresponding box-plots of all individual scores for two-week-ahead predictions.
+D: Corresponding boxplots of all individual scores for two-week-ahead predictions.
 E: Boxplots for the relative change of scores for the EuroCOVIDhub-ensemble across forecast horizons.
@@
 As expected on the natural scale, scores correlate strongly with the average number of observed cases and deaths (Spearman rank correlation: 0.96 and 0.97) in a given country (see Figure\ref{fig:HUB-mean-locations}A, B).
+We report the Spearman rank correlation here, rather than the Pearson correlation, so that correlations remain the same regardless of whether we look at observed values on the log or natural scale.
 On the log scale, the correlation is much weaker and slightly negative, at least in terms of rank correlation (Spearman correlation: -0.33 and -0.36), implying a tendency towards larger scores in countries with lower mean incidences.
@@
 We found that model rankings changed noticeably, favouring models like epiforecasts-EpiNow2 which explicitly has an exponential error structure, and disfavouring models like MUNI-ARIMA which is a purely statistical model with a linear error structure.
-On the natural scale, missing the peak and overshooting was more severely penalised, while missing the nadir and a following upswing in numbers tended to be more severely penalised on the log scale than on the natural scale.
+On the natural scale, missing the peak and overshooting was more severely penalised than missing the nadir and a following upswing in numbers.
+Both failure modes tended to be more equally penalised on the log scale (with undershooting receiving slightly higher penalties in our example).
 
@@
 One could also think about applying a Box-Cox or a square-root transformation in order to stabilise the variance of the forecasts.
-Another promising transformation would be to divide each forecast by the forecast of the previous week (and observations by the observation of the previous week), in order to obtain forecasts for week-to-week multiplicative growth rates, or alternatively to ask forecasters to provide estimates of the weekly relative changed applied to the latest data and subsequent forecast points directly.
+Another promising transformation would be to divide each forecast by the forecast of the previous week (and observations by the observation of the previous week), in order to obtain forecasts for week-to-week multiplicative growth rates, or alternatively, to ask forecasters to provide estimates of the weekly relative change applied to the latest data and subsequent forecast points directly.
 This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluation the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}.

### 2022-11-29 analysis/Analysis ECDC forecasts.Rmd
update plots

@@
 ## =============================================================================
+## FIGURE 1
+# illustration-effect-log-score.R
+
+## =============================================================================
+## FIGURE 2
+# Simulation poisson-mean-sd.R
+
+## =============================================================================
+## FIGURE 3
+# example-log-first-Johannes.R
+
+## =============================================================================
 ## FIGURE 4
@@
 
-  total_values <- 
+  mean_values <- 
     fulldata |> 
@@
     group_by(location, target_type) |>
-    summarise(total = mean(true_value))
+    summarise(mean_obs = mean(true_value))
 
@@
               .groups = "drop_last") |> 
-    inner_join(total_values)
+    inner_join(mean_values)
 
@@
     group_by(target_type, scale) |>
-    summarise(correlation = cor(wis, total)) |> 
+    summarise(correlation = cor(wis, mean_obs)) |> 
     print()
@@
   plot_df |>
-    ggplot(aes(y = wis, x = total)) + 
+    ggplot(aes(y = wis, x = mean_obs)) + 
     geom_point() + 
@@
     scale_y_continuous(labels = scale_fn) +
-    labs(y = "REPLACE", x = "Total number of observed values")
+    labs(y = "REPLACE", x = "Mean observed value")
 }
@@
 p_cor_scores <- df |>
-  ggplot(aes(x = horizon, y = cor, 
-             fill = `Target type`, color = `Target type`)) + 
-  geom_histogram(stat = "identity", position = "dodge") +
+  ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
+  scale_fill_brewer(palette = "Set1", name = "Target type") + 
+  scale_color_brewer(palette = "Set1", name = "Target type") +
+  geom_histogram(stat = "identity", position = "dodge", alpha = 0.4) +
   theme_scoringutils() + 
   coord_cartesian(ylim = c(0, 1)) +
-  labs(y = "Avg. Spearman cor(scores)", x = "Forecast horizon")
+  labs(y = "Avg. rank cor(scores)", x = "Forecast horizon") + 
+  theme(axis.title.y = element_text(size=8.5))
 
@@
   ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
-  geom_histogram(stat = "identity", position = "dodge") +
+  scale_fill_brewer(palette = "Set1", name = "Target type") + 
+  scale_color_brewer(palette = "Set1", name = "Target type") + 
+  geom_histogram(stat = "identity", position = "dodge", alpha = 0.4) +
   theme_scoringutils() + 
@@
 
-ggsave("output/figures/HUB-correlations.png", width = 7, height = 5.5)
+ggsave("output/figures/HUB-correlations.png", width = 7, height = 7)
 
@@
 put_plot_together("EuroCOVIDhub-ensemble")
-ggsave(filename = "output/figures/HUB-model-comparison-ensemble.png", width = 10, height = 7)
+ggsave(filename = "output/figures/HUB-model-comparison-ensemble.png", width = 10, height = 8.5)
 
 put_plot_together("epiforecasts-EpiNow2")
-ggsave(filename = "output/figures/HUB-model-comparison-epinow.png", width = 10, height = 7)
+ggsave(filename = "output/figures/HUB-model-comparison-epinow.png", width = 10, height = 8.5)
 
 put_plot_together("EuroCOVIDhub-baseline", locationname = "DE")
-ggsave(filename = "output/figures/HUB-model-comparison-baseline.png", width = 10, height = 7)
+ggsave(filename = "output/figures/HUB-model-comparison-baseline.png", width = 10, height = 8.5)
 
@@
   plot_layout(design = layout, 
-              guides = "collect") &
+              guides = "collect", 
+              widths = c(1, 1, 1, 1, 2.5)) &
   theme(legend.position = "bottom")

### 2022-11-29 analysis/example-log-first-Johannes.R

@@ -2,6 +2,7 @@ library(dplyr)
 library(tidyr)
 library(scoringutils)
 library(ggplot2)
+library(patchwork)
 
 n_sim <- 1000
 epsilon <- rnorm(n_sim)
@@ -39,20 +40,25 @@ summary <- scores |>
   group_by(sigma) |>
   summarise(log_wis = sum(log_wis, na.rm = TRUE), 
             wis_log = sum(log, na.rm = TRUE), 
-            wis = sum(natural, na.rm = TRUE))
-
-summary |> 
+            wis = sum(natural, na.rm = TRUE)) |>
   pivot_longer(cols = c(wis, log_wis, wis_log), values_to = "score", names_to = "type") |>
   mutate(type = factor(type, 
                        levels = c("wis", "wis_log", "log_wis"), 
-                       labels = c("WIS", "WIS (log scale)", "log(WIS)"))) |>
-  ggplot(aes(x = sigma, y = score)) +
-  geom_vline(xintercept = 1, linetype = "dashed", size = 0.3, color = "grey80") +
-  geom_point() + 
-  facet_wrap(~type, scales = "free_y") + 
-  theme_scoringutils() +
-  labs(y = "Score", x = "Standard deviation of predictive distribution") + 
-  theme(panel.spacing = unit(1, "lines"))
+                       labels = c("WIS", "WIS (log scale)", "log(WIS)")))
+
+
+score_plot <- function(summary) {
+  summary |>
+    ggplot(aes(x = sigma, y = score)) +
+    geom_vline(xintercept = 1, linetype = "dashed", size = 0.3, color = "grey80") +
+    geom_point() + 
+    facet_wrap(~type, scales = "free_y") + 
+    theme_scoringutils() +
+    labs(y = "Score", x = "Standard deviation of predictive distribution") + 
+    theme(panel.spacing = unit(1, "lines"))
+}
+
+score_plot(summary)
 
 ggsave("output/figures/example-log-first.png", width = 7, height = 2.1)
 

### 2022-11-29 analysis/illustration-effect-log-score.R

@@ -44,39 +44,60 @@ scores <- vals |>
 
 scale_factor <- 3 # scale factor for the density
 
-p1 <- scores |>
-  ggplot(aes(x = id, y = interval_score/scale_factor)) + 
-  geom_area(stat = "function", 
-            fun = function(x) {dnorm(x, mean = 1, sd = 0.4)}, 
-            color = "grey", fill = "grey", alpha = 0.5) +
-  geom_line() + 
-  labs(y = "WIS", x = "Observed value") + 
-  scale_y_continuous(label = function(x) {paste(scale_factor * x)}) + 
-  facet_wrap(~ scale, ncol = 1, scales = "free") + 
-  theme_scoringutils() + 
-  theme(panel.spacing = unit(1, "lines"))
+plot_linear_x <- function(scores) {
+  scores |>
+    ggplot(aes(x = id, y = interval_score/scale_factor)) + 
+    geom_area(stat = "function", 
+              fun = function(x) {dnorm(x, mean = 1, sd = 0.4)}, 
+              color = "grey", fill = "grey", alpha = 0.5) +
+    geom_line() + 
+    labs(y = "WIS", x = "Observed value") + 
+    scale_y_continuous(label = function(x) {paste(scale_factor * x)}) + 
+    facet_wrap(~ scale, ncol = 1, scales = "free") + 
+    theme_scoringutils() + 
+    theme(panel.spacing = unit(1, "lines"))
+}
+
+p1 <- filter(scores, scale == "natural") |>
+  plot_linear_x()
+
+p2 <- filter(scores, scale == "log") |>
+  plot_linear_x()
+
+
+
+
 
 scale_factor2 <- 3
-p2_log <- scores |>
-  ggplot(aes(x = id, y = interval_score/scale_factor2)) + 
-  geom_area(stat = "function", 
-            fun = function(x) {dnorm(x, mean = 1, sd = 0.4)}, 
-            color = "grey", fill = "grey", alpha = 0.5) +
-  geom_line() + 
-  labs(y = "WIS", x = "Observed value") + 
-  scale_y_continuous(label = function(x) {paste(scale_factor2 * x)}) + 
-  scale_x_continuous(trans = "log", 
-                     label = function(x) {
-                       ifelse(x < 1, 
-                              paste0("1/", (1 / x)), 
-                              paste(x))
-                     }, 
-                     breaks = c(0.2, 1/2, 1, 2, 5)) + 
[…]

### 2022-11-29 main.tex
Updates from Overleaf

@@
 Our aim in transforming forecasts and observations is to shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score.
+A particularly relevant transformation in the context of epidemic phenomena is the natural logarithm.
 Instead of a score representing the magnitude of absolute errors this approach makes it possible to obtain a score which a) measures relative error and at the same time b) provides an approximate measure for how well a forecast captures the multiplicative growth rate of the target quantity (see Section \ref{sec:methods:growthrate}.
@@
     \caption{Illustration of the weighted interval score (black line) for a $\mathcal{N}(1, 0.2)$ predictive distribution (density represented by the grey area) as a function of the observed value.
-A: Score on the natural scale, linear x-axis.
-B: Score on the natural scale, logarithmic x-axis.
-C: Score on the log scale, linear x-axis.
-D: Score on the log scale, logarithmic x-axis.}
+
+    A: Scores computed on the natural scale (without prior transformation of the predictive distribution and the observed values). 
+    Scores as a function of observed values (x-axis) on a linear scale.
+
+    Score on the natural scale, linear x-axis. B: Score on the natural scale, logarithmic x-axis. C: Score on the log scale, linear x-axis. D: Score on the log scale, logarithmic x-axis.} 
     \label{fig:change-in-scores}

### 2022-11-29 main.tex
update Figure legends. Clean manuscript from comments and old ideas

@@
 A particularly relevant transformation in the context of epidemic phenomena is the natural logarithm.
-Instead of a score representing the magnitude of absolute errors this approach makes it possible to obtain a score which a) measures relative error and at the same time b) provides an approximate measure for how well a forecast captures the multiplicative growth rate of the target quantity (see Section \ref{sec:methods:growthrate}.
+Instead of a score representing the magnitude of absolute errors this approach makes it possible to obtain a score which a) measures relative error and atx the same time b) provides an approximate measure for how well a forecast captures the multiplicative growth rate of the target quantity (see Section \ref{sec:methods:growthrate}.
 The interpretation of the log-transformed score as an approximate score for a forecast of the growth rate of a given quantity aligns the evaluation with our understanding of how epidemiological processes should be modeled and the kinds of errors models would typically make.
@@
     \includegraphics[width=0.99\textwidth]{output/figures/SIM-effect-log-score.png}
-    \caption{Illustration of the weighted interval score (black line) for a $\mathcal{N}(1, 0.2)$ predictive distribution (density represented by the grey area) as a function of the observed value.
-
+    \caption{Illustration of the weighted interval score on the natural and the log scale. The WIS
+    (black line) is shown for a $\mathcal{N}(1, 0.2)$ predictive distribution (truncated at 0, density represented by the grey area) as a function of the observed value. 
     A: Scores computed on the natural scale (without prior transformation of the predictive distribution and the observed values).
-    Scores as a function of observed values (x-axis) on a linear scale.
-
-    Score on the natural scale, linear x-axis. B: Score on the natural scale, logarithmic x-axis. C: Score on the log scale, linear x-axis. D: Score on the log scale, logarithmic x-axis.} 
+The x-axis is shown on a linear scale.
+C: Scores computed on the logarithmic scale (log-transforming the predictive distribution and the observed values prior to scoring).
+The x-axis is shown on a linear scale.
+B, D: Like A, C, but the x-axis is shown on a relative scale.}
     \label{fig:change-in-scores}
@@
 Note that this approximation becomes increasingly inaccurate when the difference between the predicted and the actual growth rate grows (although forecasts can still be interpreted in terms of relative errors as shown above).
-%If this is of great concern one may want to use a different transformation such as dividing forecasts and observations by the last known values in order to score the growth rate explicitly (although this lacks the elegance of having one single transformation per time point).
-
-% Conveniently, the decomposition of the WIS into dispersion, overprediction and underprediction is preserved.
-The interpretation of the components has now changed in that they now (approximately) represent uncertainty, over-prediction, and under-prediction with the growth rate and so are independent of current incidence.
 
@@
 
-% PARAGRAPH NEEDS SOME WORK
 Overall model rankings would be expected to differ even more when scores are averaged across multiple forecasts or targets.
@@
 For these simulated data, log predictions for small quantities received on average higher scores than forecasts for large quantities if the mean and the variance grew at the same rate $\sigma^2 = \mu$, about equal scores when the variance grew at a rate of $\sigma^2 = \mu + \mu^2$, and smaller scores when the variance grew faster than that (Figure \ref{fig:SIM-wis-state-size-mean}).
-% To illustrate this with count data we sampled forecasts from different negative binomial distributions.
-The negative binomial distribution has mean $\mu$ and variance $\sigma^2 = \mu + \mu ^2 / \theta$ and for $\lim_{\theta \to \infty}$ converges to the Poisson distribution.
-For large values of $\theta$, resulting in a variance approximately equal to the mean, forecasts for lower quantities on average received higher scores on the log scale (see Figure \ref{fig:SIM-wis-state-size-mean}.
-For $\theta = 1$ (and correspondingly, $\sigma^2 = \mu + \mu^2$, we found that scores on the log scale remained approximately constant regardless of the size of the quantity to forecast.
-For $\theta = 0.1$ (and $\sigma^2 = \mu + 10 \cdot \mu^2$), scores on the log scale increased with the quantity to forecast.
-When scored on the natural scale, a higher quantity to forecast always lead to higher scores regardless of the chosen distribution (as long as the variance grows together with the mean).
 
@@
     \includegraphics[width=0.99\textwidth]{output/figures/example-log-first.png}
-    \caption{Scores for different forecasts evaluated using the WIS and the log WIS.
+    \caption{Scores for different forecasts evaluated using the WIS, the WIS on the log-scale and the logarithm of the WIS.
 We simulated 1000 observations $Y_i = {\rm e}^{x_i}$, with $x_i \text{ iid} \sim \mathcal{N}(0, 1)$.
-We then simulated 20 forecasters who would issue a predictive distribution $F = {\rm e}^{x_i}$, with $x \sim \mathcal{N}(0, \sigma)$, with values of $\sigma$ ranging from 0.1 to 2.}
+We then simulated 20 forecasters who would issue a predictive distribution $F = {\rm e}^{x_i}$, with $x \sim \mathcal{N}(0, \sigma)$, with values of $\sigma$ ranging from 0.1 to 2.
+For forecasts evaluated by the WIS and the WIS on log-transformed values, the score is minimised when the predictive distribution equals the data-generating distribution.
+When taking the logarithm of the WIS, scores are minimised for a predictive distribution that is too narrow compared to the data-generating distribution.}
     \label{fig:log-improper}
@@
 
-% SECTION STRUCUTRE / IMPORTANT THINGS
-% \begin{itemize}
-% \item X variation gets smaller / more even
-% \item X relationship with country size changes
-% \item X inndividual rankings stay similar
-% \item X relative skill scores change a lot
-% \item deep dive into what happens for individual models
-% \item scores increase across forecast horizons
-% \end{itemize}
-% 
-% Maybe look at correlation between bias and score?
-% Correlation between relative WIS contributions on the log vs. the natural scale.?
-
 \begin{figure}[h!]
@@
     \includegraphics[width=0.99\textwidth]{output/figures/HUB-mean-obs-location.png}
-    \caption{Observations and scores across locations and forecast horizons.
+    \caption{Observations and scores across locations and forecast horizons for the European COVID-19 Forecast Hub data. Locations are sorted according to the mean observed value in that location. 
     A: Average (across all time points) of observed cases and deaths for different locations.
@@
 A: Mean WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against the average number of observations in a location.
-B: Average Spearman CHECK AGAIN (rank-) correlation of scores for individual forecasts from \textit{all} models.
+B: Average Spearman rank correlation of scores for individual forecasts from all models.
 For every individual target (defined by a combination of forecast date, target type, horizon, location), one score was obtained per model.
@@
     \includegraphics[width=0.99\textwidth]{output/figures/HUB-pairwise-comparisons.png}
-    \caption{Changes in model ratings as measured by relative skill for two-week-ahead predictions. Top row: cases, bottom row: deaths. NEED TO CHANGE THE ANNOTATION LEVELS.}
+    \caption{Changes in model ratings as measured by relative skill for two-week-ahead predictions for cases (top row) and deaths (bottom row).
+A: Relative skill scores for case forecasts from different models submitted to the European COVID-19 Forecast Hub computed on the natural scale.
+B: Change in rankings as determined by relative skill scores when moving from an evaluation on the natural scale to one on the logarithmic scale.
+C: Relative skill scores based on scores on the log scale.
+D: Difference in relative skill scores computed on the natural and on the logarithmic scale.
+E: Relative contributions of the different WIS components (overprediction, underprediction, and dispersion) to overall model scores on the natural and the logarithmic scale.
+F, G, H, I, J, K: Analogously for deaths.}
     \label{fig:HUB-rank-order}
@@
 Encouragingly for the European Forecast Hub, the Hub ensemble, which is the forecast the organisers suggest forecast consumers make use of, remains one of the top forecasts across scoring schemes and indeed ranks higher for reported cases when a log-transform is used.
-
-% Observations were overdispersed, meaning that for a given location, the variance in observations exceeds the mean of all observed values (see Figure \ref{fig:HUB-variance-mean}). 
-%Could fit a model var = alpha * x2 + beta *x
-%This is in line with the observation that the association between variance is smaller than mu^2 + mu, meaning that we would expect this.
 
@@
 
- %COULD ALSO MAKE A NICE PLOT ABOUT OUTLIERS, where we look at the effect of outliers. but maybe that is already included in the plot in Figure 1. 
-% could make an analysis re outliers: how do scores change if we remove the 1 or 2 worst forecasts? The one or two best forecasts?
-
 \newpage
@@
         \end{tabular}
-    \caption{Table with summary statistics for observations and scores for forecasts from the ECDC data set.}
+    \caption{Summary statistics for observations and scores for forecasts from the ECDC data set.}
     \label{tab:HUB-summary}
@@
 \end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Idea parking lot 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \section{Other ideas}
-% \begin{itemize}
-%     \item Compare different predictive distributions and the score as a function of an outcome.
-%     \item ...
-% \end{itemize}
-
-%Score different things from the European / US Forecast Hub and plot the relationship between mean and variance.
-Plot log-scale scores for different things and see which of these influence average scores most \\
-% What happens to the decomposition of the WIS if we log?
-
-%\begin{itemize}
-%    \item Optional: Discussion of parallels to point forecasts and the point that you're still incentivised to report the median\\
-%\end{itemize}
-
-% NOT SO CONVINCED ANYMORE THAT THIS WORKS. MY INTUITION IS IF YOU WANT THE GROWTH RATE YOU SHOULD ACTUALLY SCORE THE GROWTH RATE AND DIVIDE BY THE LAST OBSERVED VALUE. 
-% 
-% WELL MAYBE IT WORKS AFTER ALL AND IT WOULD BE THE EQUIVALENT OF LOGGING FIRST AND THEN SUBTRACTING THE LAST OBSERVED VALUE WAIT A MINUTE THAT'S A DIFFERENT TRANSFORMATION AND ALSO THAT's SCORING THE LOG GROWTH RATE.
-% 
-% BUT THEN AGAIN I'M STILL CONFUSED, BECAUSE ANY RELATIVE ERROR ON A QUANTITY Y\_t+1 IS AUTOMATICALLY AN ABSOLUTE ERROR ON THE GROWTH RATE IF WE KNOW Y\_t.
-AH BUT YOU DON'T KNOW WHAT THE ABSOLUTE ERROR ON THE GROWTH RATE IS (E.G.
-FOR Y\_t+1 of 10 IF YOU DON'T SPECIFICALLY LOOK AT THE LAST VALUE.
-BUT WE ARE ASSUMING THAT THIS IS WHAT THE MODEL DOES WHEN IT MAKES A FORECAST INTERNALLY
-
-% The WIS can be thought of as an approximation of the CRPS for forecasts where the predictive distribution is represented through a set of quantiles.
-The CRPS, and by extension the WIS, represents a generalisation of the absolute error to continuous distributions, meaning that forecasters are evaluated on the absolute distance of their predictive distribution to the observed value (i.e.
-some form of absolute error).
-For simplicity, we will use absolute error and absolute distance between forecast and observation interchangeably, even though the former applies to point predictions and the latter is used for probabilistic forecasts.
-
-% 
-% 
-% 
-% We argue that when dealing with epidemiological processes it makes sense to shift the evaluation towards multiplicative errors by either taking the logarithm of both the forecast and the observations or by dividing them by the last known observation before scoring.
-
-% What kinds of transformations can be considered meaningful depends on the context.
-We will illustrate the idea of transforming forecasts by focussing on the WIS as it is currently used in epidemiological settings such as the COVID-19 Forecast hubs.
-% In this paper we focus on the WIS due to its simplicity and widespread use by the COVID-19 Forecast Hubs, but arguments apply analogously to the CRPS.
-Another proper scoring rule commonly used is the log score [CITATION], which we will not discuss further in this paper.
-
-% \paragraph{Problems with the WIS and CRPS in an epidemiological setting}
-% Their relation to the absolute error means that both the WIS and the CRPS usually scale with the prediction target.
-Forecasts of COVID-19 cases, for example typically have higher scores than forecasts of hospitalisations of death.
-Similarly, when looking at performance across different locations or over time, average scores will be dominated by locations and times with high incidences.
-Outliers also can have a disproportionate effect on aggregate scores.
-One can argue that this is meaningful and that we should care most about places and periods when incidence is high (maybe cite Bracher et al.).
-However, this may not always be true and is clearly not desirable when comparing performance of a model on different prediction targets: case numbers are not necessarily more important than hospitalisations, just because observed values tend be an order of magnitude higher.
-This makes forecasts often hard or impossible to compare.
-
-% not sure this plot is really meaningful
-%\begin{figure}[h!]
-%    \centering
-%    \includegraphics[width=0.99\textwidth]{output/figures/HUB-sd-vs-mean-scores.png}
-%    \caption{CAPTION}
-%    \label{fig:HUB-mean-sd-scores}
-%\end{figure}
-
-%\begin{itemize}
-%    \item idea of scores on logged data: interpretation as a measure of relative improvement, just as when you log the absolute error
-%\end{itemize}
-
-%\paragraph{What's there in the literature}
-
-%\paragraph{Current problems / questions about scoring an epidemiological setting}
-%probably narrow down to the ones we really want to discuss%
-%There are limitations with what we can do with scores on a natural scale and open questions whether we can do these things on a log-scale and also whether a log-scale might be inherently more appropriate
-%\begin{itemize}
-    %\item Can't really model score on a natural scale as errors are heavily skewed. Is there a way to model scores to get insight about different factors that systematically influence scores
- %   \item make a plot somewhere that shows the distribution of scores
-  %  \item Maybe epidemiological forecasts are inherently better suited to be scored on an absolute scale due to the multiplicative nature of processes (related to over-prediction > under-prediction)?
-Does the log-scale imply we are scoring the growth rate?
-   % \item unclear what trade-offs of logging / not logging are
-    %\item what score (log, not log) matches closest what our intuition (or policymakers) think is good? 
-    %\item are we better at forecasting deaths than cases? --> relative measure would help
-    %\item There exists confusion about what can be done to a score in general (e.g. people want to take the median, which they shouldn't) %maybe don't add as an extra point
-%\end{itemize}
-
-%\paragraph{Over- and under-prediction}
-%If we want to keep this in, we could: 
-%\begin{itemize}
-%    \item Check whether there is actually a problem with over-prediction and under-prediction in the Hub.
-This could be the case because we are most interested in certain scenarios in which this might arise.
-%    \item Discuss this in light of Johannes' analysis of how the decomposition of WIS values differs if the data are skewed
-%    \item compare this to the PIT-value-like relative bias scores we used for the German / Polish paper, which capture a relative tendency to over- or under-predict, rather than absolute penalties. 
-%\end{itemize}
-
- % Pooled across all time points and locations, the mean WIS on the natural scale for cases (deaths) was 8855 (59.3) with an overall standard deviation of 51274 (156) (see Table \ref{tab:HUB-summary}).
-On the log scale, scores were much closer to each other (see Figure \ref{fig:HUB-average-scores}), with a mean WIS for cases (deaths) of 0.507 (0.362) with an overall standard deviation of 0.745 (0.407).
-
-% When evaluated on the natural scale, there is  strong relationship between the mean of the weighted interval score and the total number of observed cases (correlation: 0.913) or deaths (cor: 0.849) in a location (see Figure \ref{fig:HUB-mean-scores-total-loglog} and Figure \ref{fig:HUB-mean-scores-total} in the SI).
-The relationship is less pronounced for scores on the log scale (cor: -0.019 for cases, -0.395 for deaths) and negative, meaning that locations with fewer observed cases or deaths tend to receive higher scores.
-
-%Whether or not a log-transformation is appropriate depends on whether one believes that differences in absolute values are meaningful.
-This may be the case for example when comparing forecasts across locations or time points, rather than different forecasts targets.
-\cite{bracherEvaluatingEpidemicForecasts2021} for example argue that in epidemiological settings it is a desirable property of the CRPS / WIS that it assigns large scores to forecasts where counts are high, as this naturally reflects the situations we should most care about and helps us identify the models that perform best in similar situations.
-If, on the other hand, we believe that good models should do consistently well and that situations with high incidences do not provide significantly more information about which models perform well or badly, then scoring on the log scale may be more appropriate.
-For example, it may be reasonable not to think that predictive performance in a country like Germany would have a hundred times more importance than performance in a country like Luxembourg when determining the best model for future decision making.
-Similarly, we may not necessarily believe that predictive performance during the January 2022 wave of COVID-19 should carry three times as much weight than predictive performance during the January 2021 wave, just because numbers where three times as high.
-
-% Divide by the latest known value = score forecast on the growth rate
-% - WHAT KINDS OF DATA CAN I CONDITION ON? 
-% - WOULD DIVIDING A FORECAST BY ITS RESOLUTION WORK? 
-% - CAN I SCORE THE TRAJECTORY / WEEK-TO-WEEK GROWTH RATE BY DIVIDING the 4 week ahead forecast by the three week ahead forecast?

### 2022-11-29 analysis/illustration-effect-log-score.R
add citation, update plot

@@ -100,5 +100,5 @@ p4 <- filter(scores, scale == "log") |>
   plot_annotation(tag_levels = "A")
 
 ggsave("output/figures/SIM-effect-log-score.png", width = 7, 
-       height = 3)
+       height = 4)
 

### 2022-11-29 main.tex
add citations

@@
   \author{Nikos I.
-Bosse\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene & Tropical Medicine, London, United Kingdom} $^{,}$\thanks{Centre for the Mathematical Modelling of Infectious Diseases, London, United Kingdom} $^{,*}$,
-
+Bosse\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom} $^{,}$\thanks{Centre for the Mathematical Modelling of Infectious Diseases, London, United Kingdom} $^{,*}$,
   Sam Abbott\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$, 
-
   Anne Cori\thanks{MRC Centre for Outbreak Analysis and Modelling, Department of Infectious Disease Epidemiology, School of Public Health, Imperial College London, London, United Kingdom} $^{}$, \\
-
-  Edwin van Leeuwen\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene & Tropical Medicine, London, United Kingdom} $^{,}$\thanks{UK Health Security Agency, London, United Kingdom} $^{}$,
-
+  Edwin van Leeuwen\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom} $^{,}$\thanks{UK Health Security Agency, London, United Kingdom} $^{}$,
   Johannes Bracher\thanks{Institute of Economic Theory and Statistics, Karlsruhe Institute of Technology, Karlsruhe, Germany} $^{}$, 
-
-  Sebastian Funk\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$
-  }
+  Sebastian Funk\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$}
 
@@
 Proper scoring rules are constructed such that forecasters are incentivised to report their true belief about the future.
-Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funk19_assessingperformancereal}, and the Weighted Interval Score \citep[WIS,][formulas given in Section \ref{wis} in the SI]{bracherEvaluatingEpidemicForecasts2021}.
+Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the Weighted Interval Score \citep[WIS,][formulas given in Section \ref{wis} in the SI]{bracherEvaluatingEpidemicForecasts2021}.
 The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in COVID-19 forecast hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
@@
 In the absence of changes in immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
-In that case, the number of new infections in the population grows exponentially in time~\citep{dushoff2021_speedstrengthepidemic}.
-This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellis21_challengescontrolcovid}.
+In that case, the number of new infections in the population grows exponentially in time~\citep{dushoffSpeedStrengthEpidemic2021}.
+This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellisChallengesControlCOVID192021}.
 If case numbers are growing exponentially (for an effective reproduction number greater than 1) and the true modelling task revolves around estimating the reproduction number or the corresponding growth rate, evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number) less than overprediction by the same amount, as errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.

### 2022-11-29 main.tex
Change formulation of multiplicative to exponential error, add WIS formulas back in

@@
 Depending on the context, many transformations may be appropriate.
-We describe the log-transformation as one specific example in detail and discuss its interpretation as a measure of relative error and an approximate score for a forecast of the growth rate.
+We describe the log-transformation as one specific example in detail and discuss its interpretation as a measure of relative error and as score for an implicit forecast of the exponential growth rate.
 Applying the log-transformation to data from the European COVID-19 Forecast Hub, we found that it changed model rankings, evened scores across time, locations and target types, and emphasised situations in which models missed lower inflection points and consequent rises in numbers, while penalising missed peaks less severely.
@@
 Proper scoring rules are constructed such that forecasters are incentivised to report their true belief about the future.
-Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the Weighted Interval Score \citep[WIS,][formulas given in Section \ref{wis} in the SI]{bracherEvaluatingEpidemicForecasts2021}.
+Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][see formulas in Section \ref{crps}]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the Weighted Interval Score \citep{bracherEvaluatingEpidemicForecasts2021}.
 The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in COVID-19 forecast hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
-The WIS and the CRPS measure the absolute distance between the predictive distribution and the observed value and can be understood as a generalisation of the absolute error to predictive distributions (see details in Sections \ref{wis} and \ref{crps} in the SI).
-For simplicity, we will use "absolute error" and "absolute distance" interchangeably, even though the former is more commonly used for point forecasts and the latter is more appropriate for predictive distributions.
+The WIS and the CRPS measure the absolute distance between the predictive distribution and the observed value and can be understood as a generalisation of the absolute error to predictive distributions.
+For a single prediction interval, the interval score is computed as the sum of three penalty components, dispersion (width of the prediction interval), underprediction and overprediction,  
+%
+\begin{linenomath*}
+\begin{align*}
+ IS_\alpha(F,y) &= (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u) \\
+ &= \text{dispersion} + \text{underprediction} + \text{overprediction},    
+\end{align*}
+\end{linenomath*}
+%
+where $1()$ is the indicator function, $y$ is the observed value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution, i.e.
+the lower and upper bound of a single central prediction interval.
+For a set of $K$ prediction intervals and the median $m$, the WIS is computed as a weighted sum,
+\begin{linenomath*}
+\begin{equation*}
+\text{WIS} = \frac{1}{K + 0.5} \cdot \left(w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y)\right),    
+\end{equation*} 
+\end{linenomath*}
+where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$.
 
@@
 A particularly relevant transformation in the context of epidemic phenomena is the natural logarithm.
-Instead of a score representing the magnitude of absolute errors this approach makes it possible to obtain a score which a) measures relative error and atx the same time b) provides an approximate measure for how well a forecast captures the multiplicative growth rate of the target quantity (see Section \ref{sec:methods:growthrate}.
-The interpretation of the log-transformed score as an approximate score for a forecast of the growth rate of a given quantity aligns the evaluation with our understanding of how epidemiological processes should be modeled and the kinds of errors models would typically make.
+Instead of a score representing the magnitude of absolute errors this approach makes it possible to obtain a score which a) measures relative error and at the same time b) provides an a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate}.
+The interpretation of the log-transformed score as a score for a forecast of the exponential growth rate of a given quantity aligns the evaluation with our understanding of how epidemiological processes should be modeled and the kinds of errors models would typically make.
 
@@
 To illustrate the effects of applying the logarithm, let us consider a point forecast $\hat{y}_{t+1}$ for a quantity of interest $y_{t+1}$, such that 
-\begin{equation}
+%
+\begin{linenomath*}
+\begin{equation*}
 y_{t+1} = \hat{y}_{t+1} + \varepsilon_{t+1}.
-\end{equation}
-
-The analogue of the WIS / CRPS for point forecasts is the absolute error, and so the score will be determined by the size of the forecast error $\varepsilon_{t+1}$.
-When taking the logarithm of the forecast and the observation first, then the score is determined by $\varepsilon^*_{t+1}$:
-\begin{equation}
+\end{equation*}
+\end{linenomath*}
+%
+The analogue of the WIS / CRPS for point forecasts is the absolute error, and so the score will be determined by the size of the forecast error $\left|\varepsilon_{t+1}\right|$.
+When taking the logarithm of the forecast and the observation first, then the score is determined by $\left|\varepsilon^*_{t+1}\right|$:
+\begin{linenomath*}
+\begin{equation*}
 \log y_{t+1} = \log \hat{y}_{t+1} + \varepsilon^*_{t+1}.
-\end{equation}
+\end{equation*}
+\end{linenomath*}
 %
@@
 %
-\begin{align}
+\begin{linenomath*}
+\begin{align*}
 \log y_{t+1} &= \log \hat{y}_{t+1} + \varepsilon^*_{t+1} \Leftrightarrow \\    
 y_{t+1} &= \hat{y}_{t+1} \cdot \exp{\varepsilon^*_{t+1}}.    
-\end{align}
+\end{align*}
+\end{linenomath*}
 % %>%
@@
 
-\subsection{Interpretation as an approximate score of the growth rate}
+\section{Interpretation as scoring the exponential growth rate}
 \label{sec:methods:growthrate}
 
-Computing a score based on the logarithm of the forecast and the observation is also approximately equivalent to scoring a forecast of a multiplicative growth rate of the forecast target, which approximates the more commonly used exponential growth rate.
-The multiplicative growth rate can be defined as
+Computing a score based on the logarithm of the forecast and the observation is also equivalent to scoring a forecast of the exponential growth rate of the forecast target.
+The exponential growth rate $r$ governs epidemic growth via (see e.g.
+\cite{wallingaHowGenerationIntervals2007})
 %
-\begin{equation}
-    g_{t, t+1} = \frac{y_{t+1} - y_t}{y_t},
-\end{equation}
+\begin{linenomath*}
+\begin{equation*}
+y_{t + \Delta t} = y_t \cdot \exp(r \cdot \Delta t).
+\end{equation*}
+\end{linenomath*}
+For small $\Delta t$ and $r$ this is approximately equivalent to
+\begin{linenomath*}
+\begin{equation*}
+ y_{t + \Delta t} = y_t \cdot (1 + r \cdot \Delta t),
+%y(t) \cdot \exp(r \cdot \Delta t) \approx y(t) \cdot (1 + r \cdot \Delta t),
+\end{equation*}
+\end{linenomath*}
 %
-where $y_{t+1}$ is the forecast target (e.g. reported cases of COVID-19) in the future, $y_t$ is the last known observation and $g_{t, t+1}$ is the growth rate between now ($t$) and $t+1$. 
-Using the growth rate, we can express the value of the forecast target in the future as such: 
+which is sometimes used as an alternative definition ($r$ in this formulation is then called the multiplicative growth rate).
+Now assume that we predict one step time step ahead, i.e., at time $t$ we predict $y_{t + 1} = \hat{y}_{t + 1} + \varepsilon^*_{t+1}$.
+Moreover assume that $r$ is time-constant between $t$ and $t + 1$.
+The score is determined by the size of $\left|\varepsilon^*_{t+1}\right|$, the absolute error of the log-transformed incidence.
 %
-\begin{equation}
-y_{t+1} = (1 + g_{t, t+1}) \cdot y_t.
-\end{equation}
+\begin{linenomath*}
+\begin{align*}
+%\varepsilon^*_{t+1} & = \left|\log \hat{y}_{t + 1} - \log y_{t + 1} \right|\\
+\varepsilon^*_{t+1} & = \log \hat{y}_{t + 1} - \log y_{t + 1} \\
+& = \log \left[\exp(\hat{r}) \cdot y_t\right] - \log \left[\exp(r) \cdot y_t\right] \\
+& = \hat{r} - r,
+\end{align*}
+\end{linenomath*}
+where $\hat{r}$ is the predicted exponential growth rate as implied by $\hat{y}(t + 1)$.
+By log-transforming both the predicted value and the observation prior to applying the absolute error we thus just assess the error in the predicted exponential growth rate.
+If we apply the same reasoning to an $n$-week-ahead forecast and assume that the growth rate is piece-wise constant with value $r_t$ on $[t, t + 1)$, we get
 %
-
-Now consider a forecast of $\log y_{t+1}$, where again, the score is determined by the size of $\varepsilon^*_{t+1}$.
-Using the fact that for small values of $x$, $\log (1+ x) \approx x$, we can see that scoring forecasts on the log scale is approximately equal to scoring an additive error on the growth rate (on the natural scale), as long as the growth rate is small.
-%
-\begin{align}
-\log y_{t+1}        &= \log \hat{y}_{t+1} + \varepsilon^*_{t+1} \nonumber\\
-                    &= \log ((1 + \hat{g}_{t+1}) \cdot y_{t}) + \varepsilon^*_{t+1} \\
-                    &= \log (1 + \hat{g}_{t, t+1}) + \log y_t + \varepsilon^*_{t+1} \Leftrightarrow \nonumber \\
-\nonumber \\
-\varepsilon^*_{t+1}  &= \log y_{t+1} -  \log y_t - \log (1 + \hat{g}_{t, t+1}) \nonumber \\    
-                    &= \log \left(\frac{y_{t+1}}{y_t}\right) - \log (1 + \hat{g}_{t, t+1}) \\    
-                    &= \log (1 + g_{t, t+1}) - \log (1 + \hat{g}_{t, t+1})\nonumber  \\   
-                    &\approx g_{t, t+1} - \hat{g}_{t, t+1} \quad \text{for small } \hat{g}_{t, t+1} \text{ and } g_{t, t+1}. \nonumber
-\end{align}
-
-The score is determined by the size of $\varepsilon^*_{t+1}$, which is approximately equal to the difference between the observed and the predicted growth rate.
-Note that this approximation becomes increasingly inaccurate when the difference between the predicted and the actual growth rate grows (although forecasts can still be interpreted in terms of relative errors as shown above).
+\begin{linenomath*}
+\begin{align*}
+\epsilon^* & = \log \hat{y}_{t + n} - \log y_{t + n}\\
+& = \log\left[\exp\left(\prod_{i = 0}^{n - 1}\hat{r}_{t + i}\right) \cdot y_t\right] - \log\left[\exp\left(\prod_{i = 0}^{n - 1}r_{t + i} \right) \cdot y_t\right]\\
+& =   \prod_{i = 0}^{n - 1}\hat{r}_{t + i} \ - \ \prod_{i = 0}^{n - 1}r_{t + i}.
+\end{align*}
+\end{linenomath*}
+We thus evaluate the error in the cumulative growth over the considered $n$-week period.
 
@@
 In this paper we proposed a log-transformation of forecasts and observed values prior to evaluation with the WIS (or CRPS) in order to address issues that arise when evaluating epidemiological forecasts based on a measure of absolute error.
-We gave a mathematical intuition for the transformation and showed that scores on the log scale can be interpreted as a) a measure of relative (multiplicative) prediction errors, as well as b) an approximate score for a forecast of the multiplicative growth rate of the target quantity.
+We gave a mathematical intuition for the transformation and showed that scores on the log scale can be interpreted as a) a measure of relative (multiplicative) prediction errors, as well as b) a score for a forecast of the exponential growth rate of the target quantity.
 When applying this approach to forecasts from the European COVID-19 Forecast Hub, we found overall scores on the log scale to be more equal across, time, location and target type (cases, deaths) than scores on the natural scale.
@@
 It also gives higher weight to another type of situation one may care about, namely one in which numbers start to rise from a previously low level.
-While scores on the log scale can always be interpreted as a measure of relative (multiplicative) predictive performance, the interpretation as a score for a forecast of the multiplicative growth rate is only approximate.
-If one is explicitly interested in scoring a forecast of the growth rate, it may be better to use a different transformation, e.g.
+If one is interested in scoring a forecast of the multiplicative, rather than the exponential  growth rate, it may be better to use a different transformation, e.g.
 dividing every forecast and corresponding observed value by the previously last observed value when the forecast was made.
@@
 
-\subsection{Additional information on the WIS} \label{wis}
-
-The Weighted Interval Score is a proper scoring rule.
-WIS values are always larger or equal than zero and lower values imply better performance.
-The WIS can be decomposed into a dispersion component and penalties for over- and under-prediction.
-For a single prediction interval, the interval score is computed as
-\begin{align}
- IS_\alpha(F,y) &= (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u) \\
- &= \text{dispersion} + \text{underprediction} + \text{overprediction},    
-\end{align}
-
-where $1()$ is the indicator function, $y$ is the observed value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution $F$, i.e.
-the lower and upper bound of a single central prediction interval.
-For a single forecast interval, the interval score is the width of the prediction interval, if it covers the observed value, and the absolute error between the forecast and the lower (or upper, respectively) interval boundary.
-For a set of $K$ prediction intervals and the median $m$, the WIS is computed as a weighted sum,
-\begin{equation}
-\text{WIS} = \frac{1}{K + 0.5} \cdot \left(w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y)\right),    
-\end{equation} 
-where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$.
-
-The WIS is closely related to the continuous ranked probability score (CRPS) [Cite Gneiting 2007], which itself can be understood as a generalisation of the absolute error to probabilistic forecasts.
-For an increasing set of equally-spaced prediction intervals the WIS converges to the CRPS and shares many of its properties.
-More information on the CRPS is provided in section \ref{crps} in the SI.
-
 \subsection{Additional information on the CRPS} \label{crps}
@@
 
-\begin{equation}
+\begin{equation*}
     \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,
-\end{equation}
+\end{equation*}
 
 where y is the true observed value and F the CDF of predictive distribution. Often the following alternative representation is used
-\begin{equation}
+\begin{equation*}
     \text{CRPS}(F, y) = \frac{1}{2} \mathbb{E}_{F} |X - X'| - \mathbb{E}_P |X - y|,
-\end{equation}
+\end{equation*}
 

### 2022-11-29 main.tex
include crps formula in main text

@@
 Proper scoring rules are constructed such that forecasters are incentivised to report their true belief about the future.
-Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][see formulas in Section \ref{crps}]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the Weighted Interval Score \citep{bracherEvaluatingEpidemicForecasts2021}.
-The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in COVID-19 forecast hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
-The WIS and the CRPS measure the absolute distance between the predictive distribution and the observed value and can be understood as a generalisation of the absolute error to predictive distributions.
+Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the Weighted Interval Score \citep{bracherEvaluatingEpidemicForecasts2021}.
+The CRPS measures the distance of the predictive distribution to the observed data as 
+\begin{linenomath*}
+\begin{equation*}
+    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,
+\end{equation*}    
+\end{linenomath*}
+where y is the true observed value and F the cumulative distribution function (CDF) of predictive distribution.
+Both the WIS and the CRPS can be understood as a generalisation of the absolute error to predictive distributions.
+The WIS is itself an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in COVID-19 forecast hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
 For a single prediction interval, the interval score is computed as the sum of three penalty components, dispersion (width of the prediction interval), underprediction and overprediction,  
@@
 
-\subsection{Additional information on the CRPS} \label{crps}
-
-The CRPS measures the distance of the predictive distribution to the observed data-generating distribution as
-
-\begin{equation*}
-    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,
-\end{equation*}
-
-where y is the true observed value and F the CDF of predictive distribution. Often the following alternative representation is used
-\begin{equation*}
-    \text{CRPS}(F, y) = \frac{1}{2} \mathbb{E}_{F} |X - X'| - \mathbb{E}_P |X - y|,
-\end{equation*}
-
-where $X$ and $X'$ are independent realisations from the predictive distributions $F$ with finite first moment and $y$ is the true value.
-In this representation we can simply replace $X$ and $X'$ by samples sum over all possible combinations to obtain the CRPS.
-
-\subsection{Additional figures}
-
 \begin{figure}[h!]

### 2022-11-30 R/toy-example.R
clean files

@@ -1,90 +1,0 @@
-#' Simulate Count data using an underlying exponential process and a Poisson
-#' Observation model.
-#' @importFrom tibble tibble
-#' @importFrom purrr map map_dbl
-#' @importFrom tidyr unnest
-simulate_exp_poisson <- function(init = 100, time = 10, growth = 0,
-                                 additive_error = 0, sims = 1) {
-  y <- init * exp(cumsum(rep(growth, time)))
-  y <- y + cumsum(rep(additive_error, time))
-  y <- purrr::map_dbl(y, ~ max(., 0))
-  obs <- tibble::tibble(
-    sim = list(1:sims),
-    time = 1:time,
-    y = y,
-    obs = purrr::map(y, function(lambda) {
-        rpois(sims, lambda)
-      }
-    )
-  )
-  obs <- tidyr::unnest(obs, c(obs, sim))
-  return(obs)
-}
-
-#' Plot simulated observations and forecasts
-#' @import ggplot2
-#' @importFrom dplyr filter
-plot_toy_forecasts <- function(forecasts, growth = 0.01, samples = 10,
-                               alpha = 0.2) {
-  forecasts |>
-    filter(sample <= samples, r == growth) |>
-    ggplot() +
-    aes(x = time, y = prediction, group = sample) +
-    geom_point(aes(y = true_value), alpha = 0.9, size = 1.1) +
-    geom_line(alpha = alpha) +
-    theme_bw() +
-    theme(legend.position = "bottom") +
-    labs(x = "Forecast horizon", y = "Simulated value",
-         caption = "Points represent simulated true values whilst each line
-                    represents a sample forecast with some error. The plot is
-                    stratified by growth rate and additive error.") +
-    facet_grid(vars(additive_error), vars(r_error))
-}
-
-#' Plot scores by horizon
-#' @import ggplot2
-#' @importFrom dplyr group_by mutate
-plot_toy_scores_by_horizon <- function(scores) {
-  scores |>
-    group_by(additive_error, r) |>
-    mutate(relative_crps = crps / max(crps)) |>
-    ungroup() |>
-    ggplot() +
-    aes(y = relative_crps, x = time, col = r_error, group = r_error) +
-    geom_point(alpha = 0.6) +
-    geom_line(alpha = 0.4) +
-    scale_color_viridis_c() +
-    theme_bw() +
-    theme(legend.position = "bottom") +
-    labs(x = "Forecast horizon", y = "CRPS relative to maximum panel value",
-         col = "Growth rate error",
[…]

### 2022-11-30 _targets_r/targets/toy_log_scores_by_horizon_plot.R

@@ -1,5 +1,0 @@
-tar_target(toy_log_scores_by_horizon_plot, {
-  toy_log_scores |>
-    filter(time <= 4) |>
-    plot_toy_scores_by_horizon()
-})

### 2022-11-30 analysis/00 full paper code.R

@@ -1,0 +1,1728 @@
+library(scoringutils) # version 1.0.1
+library(scoringRules)
+library(purrr)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(patchwork)
+library(stringr)
+library(here)
+library(data.table)
+library(kableExtra)
+library(performance)
+library(tidytext)
+library(grid)
+library(ggpubr)
+
+set.seed(1234)
+
+recompute_scores <- FALSE
+
+# situations with negative values: France 2021-05-22, CZ 2021-08-07, ES 2021-03-06
+
+
+
+## ========================================================================== ##
+## FIGURE 1
+## ========================================================================== ##
+figure_1 <- function() {
+  true_values <- seq(0.2, 5, length.out = 1000)
+  quantiles <- c(0.01, 0.025, seq(0.05, 0.95, 0.05), 0.975, 0.99)
+  
+  vals <- expand.grid(
+    true_value = true_values, 
+    quantile = quantiles
+  ) |>
+    mutate(id = true_value) |>
+    arrange(true_value, quantile) |>
+    mutate(prediction = qnorm(p = quantile, mean = 1, sd = 0.4), 
+           model = "Model")
+  
+  scores <- vals |>
+    mutate(scale = "natural") |>
+    score() |> 
+    summarise_scores(by = c("scale", "id")) 
+  
+  scores_log <- vals |>
+    mutate(true_value = log(true_value), 
+           prediction = log(prediction), 
+           scale = "log") |>
+    score() |>
+    summarise_scores(by = c("scale", "id"))
+  
+  scores <- scores |>
+    rbind(scores_log) |>
+    group_by(scale) |>
+    mutate(score = interval_score / min(interval_score), 
+           scale = factor(scale, levels = c("natural", "log"))) 
+  
+  # scale factor for the density in the plot to make it look nicer
+  scale_factor <- 3
+  
+  plot_fct <- function(scores, filter = "natural", relative_x = FALSE) {
+    p <- scores |>
+      filter(scale == filter) |>
+      ggplot(aes(x = id, y = interval_score/scale_factor)) + 
+      geom_area(stat = "function", 
+                fun = function(x) {dnorm(x, mean = 1, sd = 0.4)}, 
+                color = "grey", fill = "grey", alpha = 0.5) +
+      geom_line() + 
+      labs(y = "WIS", x = "Observed value") + 
+      scale_y_continuous(label = function(x) {paste(scale_factor * x)}) + 
[…]

### 2022-12-01 analysis/00 full paper code.R
add new figure example model ranking, clean code

@@ -20,86 +20,54 @@ recompute_scores <- FALSE
 
 # situations with negative values: France 2021-05-22, CZ 2021-08-07, ES 2021-03-06
 
-
-
 ## ========================================================================== ##
 ## FIGURE 1
 ## ========================================================================== ##
-figure_1 <- function() {
-  true_values <- seq(0.2, 5, length.out = 1000)
-  quantiles <- c(0.01, 0.025, seq(0.05, 0.95, 0.05), 0.975, 0.99)
-  
-  vals <- expand.grid(
-    true_value = true_values, 
-    quantile = quantiles
-  ) |>
-    mutate(id = true_value) |>
-    arrange(true_value, quantile) |>
-    mutate(prediction = qnorm(p = quantile, mean = 1, sd = 0.4), 
-           model = "Model")
-  
-  scores <- vals |>
-    mutate(scale = "natural") |>
-    score() |> 
-    summarise_scores(by = c("scale", "id")) 
-  
-  scores_log <- vals |>
-    mutate(true_value = log(true_value), 
-           prediction = log(prediction), 
-           scale = "log") |>
-    score() |>
-    summarise_scores(by = c("scale", "id"))
-  
-  scores <- scores |>
-    rbind(scores_log) |>
-    group_by(scale) |>
-    mutate(score = interval_score / min(interval_score), 
-           scale = factor(scale, levels = c("natural", "log"))) 
-  
-  # scale factor for the density in the plot to make it look nicer
-  scale_factor <- 3
-  
-  plot_fct <- function(scores, filter = "natural", relative_x = FALSE) {
-    p <- scores |>
-      filter(scale == filter) |>
-      ggplot(aes(x = id, y = interval_score/scale_factor)) + 
-      geom_area(stat = "function", 
-                fun = function(x) {dnorm(x, mean = 1, sd = 0.4)}, 
-                color = "grey", fill = "grey", alpha = 0.5) +
-      geom_line() + 
-      labs(y = "WIS", x = "Observed value") + 
-      scale_y_continuous(label = function(x) {paste(scale_factor * x)}) + 
-      facet_wrap(~ scale, ncol = 1, scales = "free") + 
-      theme_scoringutils() + 
-      theme(panel.spacing = unit(1, "lines"))
-    
-    if (relative_x) {
-      p <- p +
-        scale_x_continuous(
[…]

### 2022-12-01 main.tex

@@
 Proper scoring rules are constructed such that forecasters are incentivised to report their true belief about the future.
-Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the Weighted Interval Score \citep{bracherEvaluatingEpidemicForecasts2021}.
+Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the Weighted Interval Score \citep{bracherEvaluatingEpidemicForecasts2021}.
 The CRPS measures the distance of the predictive distribution to the observed data as 
@@
 \end{linenomath*}
-where y is the true observed value and F the cumulative distribution function (CDF) of predictive distribution.
+where $y$ is the true observed value and $F$ the cumulative distribution function (CDF) of predictive distribution.
 Both the WIS and the CRPS can be understood as a generalisation of the absolute error to predictive distributions.
 The WIS is itself an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in COVID-19 forecast hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
-For a single prediction interval, the interval score is computed as the sum of three penalty components, dispersion (width of the prediction interval), underprediction and overprediction,  
-%
-\begin{linenomath*}
-\begin{align*}
- IS_\alpha(F,y) &= (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u) \\
- &= \text{dispersion} + \text{underprediction} + \text{overprediction},    
-\end{align*}
-\end{linenomath*}
-%
-where $1()$ is the indicator function, $y$ is the observed value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution, i.e.
-the lower and upper bound of a single central prediction interval.
-For a set of $K$ prediction intervals and the median $m$, the WIS is computed as a weighted sum,
+The WIS is commonly defined as
 \begin{linenomath*}
 \begin{equation*}
-\text{WIS} = \frac{1}{K + 0.5} \cdot \left(w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y)\right),    
+    \text{WIS}(F, y) = \frac{1}{2K + 1} \times \sum_{k = 1}^{2K + 1} \times 2 \left[ 1(y \leq q_{\tau_k}) - \tau_k \right] \times ( q_{\tau_k} - y), 
 \end{equation*}
 \end{linenomath*}
-where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$.
+where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, $K$ is the number of central prediction intervals provided, and $m$ is the median forecast.
 
-Infectious processes are commonly modelled around the concept of the reproduction number \citep{gosticPracticalConsiderationsMeasuring2020}, the average number of new cases generated by each case in the population.
+The dynamics of infectious processes are often described by the complementary concepts of the reproduction number R \citep{gosticPracticalConsiderationsMeasuring2020} and growth rate r \citep{wallingaHowGenerationIntervals2007}, where R describes the strength and r the speed of epidemic growth \citep{dushoffSpeedStrengthEpidemic2021}.
 In the absence of changes in immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
-In that case, the number of new infections in the population grows exponentially in time~\citep{dushoffSpeedStrengthEpidemic2021}.
+In that case, the number of new infections in the population grows exponentially in time.
 This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellisChallengesControlCOVID192021}.
-If case numbers are growing exponentially (for an effective reproduction number greater than 1) and the true modelling task revolves around estimating the reproduction number or the corresponding growth rate, evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number) less than overprediction by the same amount, as errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
+If case numbers are growing exponentially and the true modelling task revolves around estimating the reproduction number or the corresponding growth rate, evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount, as errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
 If one is to assess the ability of forecasts to assess and forecast the underlying infection dynamics, it may be more appropriate to evaluate errors on the growth rate directly.
@@
 
-\subsection{Interpretation as a score based on multiplicative (relative) errors}
+\subsection*{Interpretation as a relative error}
 \label{sec:methods:relative}
 
-To illustrate the effects of applying the logarithm, let us consider a point forecast $\hat{y}_{t+1}$ for a quantity of interest $y_{t+1}$, such that 
+To illustrate the effect of applying the logarithm prior to evaluating forecasts we consider the absolute error, which the CRPS and WIS generalize to probabilistic forecasts.
+When considering a point forecast $\hat{y}$ for a quantity of interest $y$, such that
 %
@@
 \begin{equation*}
-y_{t+1} = \hat{y}_{t+1} + \varepsilon_{t+1}.
+y = \hat{y} + \varepsilon,
+\end{equation*}
+\end{linenomath*}
+the absolute error is given by $|\varepsilon|$. When taking the logarithm of the forecast and the observation first (assuming that both are strictly positive), thus considering 
+\begin{linenomath*}
+\begin{equation*}
+\log y = \log \hat{y} + \varepsilon^*,
+\end{equation*}
+\end{linenomath*}
+the resulting absolute error $\left|\varepsilon^*\right|$ can be interpreted as an approximation of various common relative error measures.
+Using that $\log(a) \approx a - 1$ if $a$ is close to 1, we get
+$$
+|\varepsilon^*| = |\log \hat{y} - \log y| = \left|\log\left(\frac{\hat{y}}{y}\right) \right| \ \ \stackrel{\text{if } \hat{y} \ \approx \ y}{\approx} \ \ \left| \frac{\hat{y}}{y} - 1 \right| \ \ = \ \ \left| \frac{\hat{y} - y}{y} \right|.
+$$
+The absolute error after log transformation is thus an approximation of the \textit{absolute percentage error} (APE) \cite{Gneiting2011JASA}.
+As in the approximation we assumed that $\hat{y} \approx y$, we can just as well interpret it as an approximation of the \textit{relative error} (RE)
+$$
+\left| \frac{\hat{y} - y}{\hat{y}} \right|
+$$
+and the \textit{symmetric absolute percentage error} (SAPE)
+$$
+\left| \ \frac{\hat{y} - y}{y/2 + \hat{y}/2} \ \right|.
+$$
+As Figure \ref{fig:SAPE} shows, the alignment with the SAPE is in fact the closest and holds quite well even if predicted and observed value differ by a factor of two or three.
+Moving back to probabilistic forecasts, the CRPS or WIS applied to log-transformed forecasts and outcomes can thus be seen as a probabilistic counterpart to the symmetric absolute percentage error, which offers an appealing intuitive interpretation.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width = 0.8\textwidth]{output/figures/SAPE.png}
+\caption{Numerical comparison of different measures of relative error/ We consider a prediction of $\hat{y} = 1$ and for different observed values show the  absolute percentage error (APE), relative error (RE), symmetric absolute percentage error (SAPE) and the absolute error applied to log-transformed predictions and observations.
+$\hat{y}$ denotes the predicted value.
+A: Error measures as a function of the observed value with the x-axis shown on a linear scale.
+B: Same, but x-axis shown on a logarithmic scale.}
+\label{fig:SAPE}
+\end{figure}
+
+%The analogy between the WIS on the log scale and the absolute error of a point forecast on the log scale is slightly inaccurate: while the absolute error on the log scale is exactly symmetric (e.g., predicting 2 instead of 1 would give the same relative error as predicting 0.5), this is not entirely true for the WIS, which is computed for a set of prediction intervals.
+Due to the fact that upper and lower bounds of individual prediction intervals are affected slightly differently by the log-transformations, small deviations may occur (compare e.g.
+in Figure \ref{fig:change-in-scores}D the small difference between scores for an observed outcome of $\frac{1}{5}$ and 5).
+
+\subsection{Interpretation as scoring the exponential growth rate}
+\label{sec:methods:growthrate}
+
+Another interpretation can be based around viewing the process to be forecast as exponential with time-varying growth rate $r(t)$ (see e.g. \cite{wallingaHowGenerationIntervals2007}), i.e.
+\begin{linenomath*}
+\begin{equation*}
+\frac{d}{dt}y(t) = r(t)y(t)
 \end{equation*}
@@
 %
-The analogue of the WIS / CRPS for point forecasts is the absolute error, and so the score will be determined by the size of the forecast error $\left|\varepsilon_{t+1}\right|$.
-When taking the logarithm of the forecast and the observation first, then the score is determined by $\left|\varepsilon^*_{t+1}\right|$:
+which is solved by
+%
 \begin{linenomath*}
 \begin{equation*}
-\log y_{t+1} = \log \hat{y}_{t+1} + \varepsilon^*_{t+1}.
+y(t) = y_0 \exp \left( \int_0^t r(t') dt' \right) = y_0 \exp (\bar{r}t)
 \end{equation*}
 \end{linenomath*}
-%
-This corresponds to evaluating a multiplicative error on the original $y_{t+1}$:
+where $y_0$ is an initial data point and $\bar{r}$ is the mean of the growth rate between the initial time point $0$ and time $t$.
+
+If a forecast $y$ of the time series at time $t$ starts at time 0 with the data point $y_0$ then the absolute error after log transformation at time $t$ is
 %
@@
 \begin{align*}
-\log y_{t+1} &= \log \hat{y}_{t+1} + \varepsilon^*_{t+1} \Leftrightarrow \\    
-y_{t+1} &= \hat{y}_{t+1} \cdot \exp{\varepsilon^*_{t+1}}.    
+\epsilon^* &= \left| \log \left[ \hat{y}( t ) \right] - \log \left[ y ( t ) \right] \right|\\
+           &= \left| \log \left[ y_0 \exp (\bar{\hat{r}} t ) \right] - \log \left[ y_0 \exp (\bar{r}t) \right] \right|\\
+           &= t \left| \bar{\hat{r}} - \bar{r} \right|.
 \end{align*}
 \end{linenomath*}
-% %>%
+%
+We thus evaluate the error in the mean exponential growth rate, scaled by the length of the time period considered.
 
-On the natural scale, WIS and CRPS increase linearly with increasing absolute distance between forecast and observation (see Figure \ref{fig:change-in-scores}A), regardless of whether the error is large in relative terms.
-On the log scale, WIS and CRPS conceptually penalise relative errors and increase linearly with an increase in relative errors (see Figure \ref{fig:change-in-scores}D).
-
-The analogy between the WIS on the log scale and the absolute error of a point forecast on the log scale is slightly inaccurate: while the absolute error on the log scale is exactly symmetric (e.g., predicting 2 instead of 1 would give the same relative error as predicting 0.5), this is not entirely true for the WIS, which is computed for a set of prediction intervals.
-Due to the fact that upper and lower bounds of individual prediction intervals are affected slightly differently by the log-transformations, small deviations may occur (compare e.g.
-in Figure \ref{fig:change-in-scores}D the small difference between scores for an observed outcome of $\frac{1}{5}$ and 5).
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/SIM-effect-log-score.png}
-    \caption{Illustration of the weighted interval score on the natural and the log scale. The WIS
-    (black line) is shown for a $\mathcal{N}(1, 0.2)$ predictive distribution (truncated at 0, density represented by the grey area) as a function of the observed value. 
-    A: Scores computed on the natural scale (without prior transformation of the predictive distribution and the observed values).
-The x-axis is shown on a linear scale.
-C: Scores computed on the logarithmic scale (log-transforming the predictive distribution and the observed values prior to scoring).
-The x-axis is shown on a linear scale.
-B, D: Like A, C, but the x-axis is shown on a relative scale.}
-    \label{fig:change-in-scores}
-\end{figure}
-
-\section{Interpretation as scoring the exponential growth rate}
-\label{sec:methods:growthrate}
-
-Computing a score based on the logarithm of the forecast and the observation is also equivalent to scoring a forecast of the exponential growth rate of the forecast target.
-The exponential growth rate $r$ governs epidemic growth via (see e.g.
-\cite{wallingaHowGenerationIntervals2007})
-%
-\begin{linenomath*}
-\begin{equation*}
-y_{t + \Delta t} = y_t \cdot \exp(r \cdot \Delta t).
-\end{equation*}
-\end{linenomath*}
-For small $\Delta t$ and $r$ this is approximately equivalent to
-\begin{linenomath*}
-\begin{equation*}
- y_{t + \Delta t} = y_t \cdot (1 + r \cdot \Delta t),
-%y(t) \cdot \exp(r \cdot \Delta t) \approx y(t) \cdot (1 + r \cdot \Delta t),
-\end{equation*}
-\end{linenomath*}
-%
-which is sometimes used as an alternative definition ($r$ in this formulation is then called the multiplicative growth rate).
-Now assume that we predict one step time step ahead, i.e., at time $t$ we predict $y_{t + 1} = \hat{y}_{t + 1} + \varepsilon^*_{t+1}$.
-Moreover assume that $r$ is time-constant between $t$ and $t + 1$.
-The score is determined by the size of $\left|\varepsilon^*_{t+1}\right|$, the absolute error of the log-transformed incidence.
-%
-\begin{linenomath*}
-\begin{align*}
-%\varepsilon^*_{t+1} & = \left|\log \hat{y}_{t + 1} - \log y_{t + 1} \right|\\
-\varepsilon^*_{t+1} & = \log \hat{y}_{t + 1} - \log y_{t + 1} \\
-& = \log \left[\exp(\hat{r}) \cdot y_t\right] - \log \left[\exp(r) \cdot y_t\right] \\
-& = \hat{r} - r,
-\end{align*}
-\end{linenomath*}
-where $\hat{r}$ is the predicted exponential growth rate as implied by $\hat{y}(t + 1)$.
-By log-transforming both the predicted value and the observation prior to applying the absolute error we thus just assess the error in the predicted exponential growth rate.
-If we apply the same reasoning to an $n$-week-ahead forecast and assume that the growth rate is piece-wise constant with value $r_t$ on $[t, t + 1)$, we get
-%
-\begin{linenomath*}
-\begin{align*}
-\epsilon^* & = \log \hat{y}_{t + n} - \log y_{t + n}\\
-& = \log\left[\exp\left(\prod_{i = 0}^{n - 1}\hat{r}_{t + i}\right) \cdot y_t\right] - \log\left[\exp\left(\prod_{i = 0}^{n - 1}r_{t + i} \right) \cdot y_t\right]\\
-& =   \prod_{i = 0}^{n - 1}\hat{r}_{t + i} \ - \ \prod_{i = 0}^{n - 1}r_{t + i}.
-\end{align*}
-\end{linenomath*}
-We thus evaluate the error in the cumulative growth over the considered $n$-week period.
+For small $t$ and $\bar{r}$, $ y_0 \exp (\bar{r}t)$ is approximately equivalent to $ y_0 (1 + \bar{r}t)$, which is sometimes used as an alternative definition (r in this second formulation is then called the multiplicative growth rate).
 
@@
 2.5].
-On the natural scale, the score would equal the dispersion of the forecast, i.e.
+On the natural scale, the score would be determined  the dispersion of the forecast, i.e.
 the length of this interval (since the observed value falls within the prediction interval).
@@
 
+\subsection{WIS}
+For a single prediction interval, the interval score is computed as the sum of three penalty components, dispersion (width of the prediction interval), underprediction and overprediction,  
+%
+\begin{linenomath*}
+\begin{align*}
+ IS_\alpha(F,y) &= (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u) \\
+ &= \text{dispersion} + \text{underprediction} + \text{overprediction},    
+\end{align*}
+\end{linenomath*}
+%
+where $1()$ is the indicator function, $y$ is the observed value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution, i.e.
+the lower and upper bound of a single central prediction interval.
+For a set of $K$ prediction intervals and the median $m$, the WIS is computed as a weighted sum,
+\begin{linenomath*}
+\begin{equation*}
+\text{WIS} = \frac{1}{K + 0.5} \cdot \left(w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y)\right),    
+\end{equation*} 
+\end{linenomath*}
+where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$.
+
+Computing a score based on the logarithm of the forecast and the observation is also equivalent to scoring a forecast of the exponential growth rate of the forecast target.
+The exponential growth rate $r$ governs epidemic growth via (see e.g.
+\cite{wallingaHowGenerationIntervals2007})
+%
+\begin{linenomath*}
+\begin{equation*}
+y_{t + \Delta t} = y_t \cdot \exp(r \cdot \Delta t).
+\end{equation*}
+\end{linenomath*}
+For small $\Delta t$ and $r$ this is approximately equivalent to
+\begin{linenomath*}
+\begin{equation*}
+ y_{t + \Delta t} = y_t \cdot (1 + r \cdot \Delta t),
+%y(t) \cdot \exp(r \cdot \Delta t) \approx y(t) \cdot (1 + r \cdot \Delta t),
+\end{equation*}
+\end{linenomath*}
+%
+which is sometimes used as an alternative definition ($r$ in this formulation is then called the multiplicative growth rate).
+Now assume that we predict one step time step ahead, i.e., at time $t$ we predict $y_{t + 1} = \hat{y}_{t + 1} + \varepsilon^*_{t+1}$.
+Moreover assume that $r$ is time-constant between $t$ and $t + 1$.
+The score is determined by the size of $\left|\varepsilon^*_{t+1}\right|$, the absolute error of the log-transformed incidence.
+%
+\begin{linenomath*}
+\begin{align*}
+%\varepsilon^*_{t+1} & = \left|\log \hat{y}_{t + 1} - \log y_{t + 1} \right|\\
+\varepsilon^*_{t+1} & = \log \hat{y}_{t + 1} - \log y_{t + 1} \\
+& = \log \left[\exp(\hat{r}) \cdot y_t\right] - \log \left[\exp(r) \cdot y_t\right] \\
+& = \hat{r} - r,
+\end{align*}
+\end{linenomath*}
+where $\hat{r}$ is the predicted exponential growth rate as implied by $\hat{y}(t + 1)$.
+By log-transforming both the predicted value and the observation prior to applying the absolute error we thus just assess the error in the predicted exponential growth rate.
+If we apply the same reasoning to an $n$-week-ahead forecast and assume that the growth rate is piece-wise constant with value $r_t$ on $[t, t + 1)$, we get
+%
+\begin{linenomath*}
+\begin{align*}
+\epsilon^* & = \log \hat{y}_{t + n} - \log y_{t + n}\\
+& = \log\left[\exp\left(\sum_{i = 0}^{n - 1}\hat{r}_{t + i}\right) \cdot y_t\right] - \log\left[\exp\left(\sum_{i = 0}^{n - 1}r_{t + i} \right) \cdot y_t\right]\\
+& =   \sum_{i = 0}^{n - 1}\hat{r}_{t + i} \ - \ \sum_{i = 0}^{n - 1}r_{t + i}.
+\end{align*}
+\end{linenomath*}
+We thus evaluate the error in the cumulative growth over the considered $n$-week period.
+
 \clearpage

### 2022-12-01 analysis/00 full paper code.R
update plot

@@ -67,7 +67,7 @@ p2 <- plot_fct(data) +
   plot_layout(guides = "collect") &
   theme(legend.position = "bottom")
 
-ggsave("output/figures/different-relative-errors.png", width = 10, height = 3)
+ggsave("output/figures/different-relative-errors.png", width = 7, height = 3)
 
 
 
@@ -164,8 +164,65 @@ ggsave("output/figures/SIM-mean-state-size.png", width = 7, height = 4.1)
 
 ## ========================================================================== ##
 ## FIGURE 3
-# example-log-first-Johannes.R
 ## ========================================================================== ##
+
+n_sim <- 1000
+epsilon <- rnorm(n_sim)
+Y <- exp(epsilon)
+
+forecasts <- expand.grid(
+  sigma = 1:20/10, 
+  quantile = c(0.01, 0.025, 1:19/20, 0.975, 0.99)
+)
+
+forecasts <- forecasts |>
+  as_tibble() |>
+  mutate(model = 10 * sigma, 
+         prediction = exp(qnorm(quantile, sd = sigma)), 
+         true_value = list(Y), 
+         sample_id = list(1:length(Y))) |>
+  unnest(c(true_value, sample_id))
+
+forecasts <- forecasts |>
+  mutate(scale = "natural") |>
+  rbind(forecasts |>
+          mutate(prediction = log(prediction + 1), 
+                 true_value = log(true_value + 1), 
+                 scale = "log"))
+
+scores <- score(forecasts)
+
+summary <- scores |>
+  mutate(log_wis = log(interval_score +1)) |>
+  group_by(sigma, scale) |>
+  summarise(wis = mean(interval_score), 
+            log_wis = mean(log_wis)) |>
+  mutate(log_wis = ifelse(scale == "log", log_wis, NA)) |>
+  pivot_wider(values_from = wis, names_from = scale) |>
+  group_by(sigma) |>
+  summarise(log_wis = sum(log_wis, na.rm = TRUE), 
+            wis_log = sum(log, na.rm = TRUE), 
+            wis = sum(natural, na.rm = TRUE)) |>
+  pivot_longer(cols = c(wis, log_wis, wis_log), values_to = "score", names_to = "type") |>
+  mutate(type = factor(type, 
+                       levels = c("wis", "wis_log", "log_wis"), 
+                       labels = c("WIS", "WIS (log scale)", "log(WIS)")))
+
+
+score_plot <- function(summary) {
[…]

### 2022-12-01 main.tex

@@
 \centering
-\includegraphics[width = 0.8\textwidth]{output/figures/SAPE.png}
+\includegraphics[width = 1\textwidth]{output/figures/different-relative-errors.png}
 \caption{Numerical comparison of different measures of relative error/ We consider a prediction of $\hat{y} = 1$ and for different observed values show the  absolute percentage error (APE), relative error (RE), symmetric absolute percentage error (SAPE) and the absolute error applied to log-transformed predictions and observations.
@@
 
-\subsection{Interpretation as scoring the exponential growth rate}
+\subsection*{Interpretation as scoring the exponential growth rate}
 \label{sec:methods:growthrate}
 
-Another interpretation can be based around viewing the process to be forecast as exponential with time-varying growth rate $r(t)$ (see e.g. \cite{wallingaHowGenerationIntervals2007}), i.e.
+Another interpretation can be based around viewing the process to be forecast as exponential with time-varying growth rate $r(t)$~\citep[see, e.g.,][]{wallingaHowGenerationIntervals2007}, i.e.
 \begin{linenomath*}
@@
 
-For small $t$ and $\bar{r}$, $ y_0 \exp (\bar{r}t)$ is approximately equivalent to $ y_0 (1 + \bar{r}t)$, which is sometimes used as an alternative definition (r in this second formulation is then called the multiplicative growth rate).
+For small $\bar{r}t$, $ y_0 \exp (\bar{r}t)$ is approximately equivalent to $ y_0 (1 + \bar{r}t)$, which is sometimes used as an alternative definition (r in this second formulation is then called the multiplicative growth rate).
 
@@
 Rankings between different forecasters based on their scores may change through the log-transformation both when looking at aggregate scores as well as when looking at a single forecast.
+We illustrate this in Figure \ref{fig:illustration-ranking} with two forecasters issuing a negative binomial distribution with different dispersion.
+When showing the obtained score as a function of the observed value, it can be seen that the ranking between the two forecasters may change when scoring the forecast on the logarithmic, rather than the natural scale.
+
 As an example for the effect on a single forecast, consider two forecasters, A and B.
@@
 This implies that scores on the log scale may treat (one-sided) outlier forecasts more leniently.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width = 1\textwidth]{output/figures/illustration-effect-log-ranking-crps.png}
+\caption{Illustration of the effect of the log-transformation of the ranking for a single forecast.
+Shown are CRPS (or WIS, respectively) values as a function of the observed value for two forecasters.
+Model A issues a negative binomial forecast with mean = 10 and $\theta = 0.5$ (resulting in a variance of 210), while Model B issues a negative binomial forecast with mean = 10 and $\theta = 10$ (resulting in a variance of 20).}
+\label{fig:illustration-ranking}
+\end{figure}
 

### 2022-12-01 analysis/00 full paper code.R
update plots

@@ -71,11 +71,105 @@ ggsave("output/figures/different-relative-errors.png", width = 7, height = 3)
 
 
 
+## ========================================================================== ##
+## FIGURE 2
+## ========================================================================== ##
+
+quantile <- seq(0.005, 0.995, 0.005)
+mu <- 10
+theta_1 <- 0.5
+theta_2 <- 10
+
+data <- data.table(
+  quantile = quantile,
+  model_A = qnbinom(p = quantile, size = theta_1, mu = mu), 
+  model_B = qnbinom(p = quantile, size = theta_2, mu = mu)
+) |>
+  pivot_longer(cols = c(model_A, model_B), names_to = "model", values_to = "prediction") |>
+  mutate(true_value = list(1:50), 
+         id = list(1:50)) |>
+  unnest(cols = c(true_value, id))
+
+scores_natural <- data |>
+  score(metrics = "interval_score") |>
+  summarise_scores(by = c("model", "id")) |>
+  mutate(scale = "natural")
+
+scores_log <- data |>
+  mutate(true_value = log(true_value), 
+         prediction = log(prediction)) |>
+  score(metrics = "interval_score") |>
+  summarise_scores(by = c("model", "id"), na.rm = TRUE) |>
+  mutate(scale = "log")
+
+scores <- rbind(scores_natural, scores_log) |>
+  mutate(scale = factor(scale, levels = c("natural", "log")), 
+         Forecaster = ifelse(model == "model_A", "A", "B"))
+
+nbinom_natural <- data.table(
+  A = rnbinom(100000, mu = 10, size = theta_1), 
+  B = rnbinom(100000, mu = 10, size = theta_2)
+) |>
+  pivot_longer(cols = c(A, B), names_to = "Forecaster") |>
+  mutate(scale = "natural")
+
+nbinom_log <- nbinom |>
+  mutate(value = log(value), 
+         scale = "log")
+
+nbinom <- rbind(nbinom_natural, nbinom_log)
+
+nbinom <- filter(nbinom, scale == "natural") |> 
+  select(-scale)
+
+plot_fct <- function(scores, scale_factor, nbinom, filter_scale = "natural") {
+  
+  scores <- filter(scores, scale == filter_scale)
+  # nbinom <- filter(nbinom, scale == filter_scale)
+  
+  scores |>
+    ggplot(aes(x = id, y = interval_score / scale_factor, color = Forecaster)) + 
+    geom_histogram(data = filter(nbinom, Forecaster == "A"), inherit.aes = FALSE,
[…]

### 2022-12-01 main.tex

@@
 Rankings between different forecasters based on their scores may change through the log-transformation both when looking at aggregate scores as well as when looking at a single forecast.
-We illustrate this in Figure \ref{fig:illustration-ranking} with two forecasters issuing a negative binomial distribution with different dispersion.
+We illustrate this in Figure \ref{fig:illustration-ranking} with two forecasters, A and B, issuing a negative binomial distribution with different dispersion.
 When showing the obtained score as a function of the observed value, it can be seen that the ranking between the two forecasters may change when scoring the forecast on the logarithmic, rather than the natural scale.
-
-As an example for the effect on a single forecast, consider two forecasters, A and B.
-Both issue a single 50\%-prediction interval for a target which later resolves as 1.
-Forecaster A issues the interval [0.5, 2], while forecaster B issues [0.9.
-2.5].
-On the natural scale, the score would be determined  the dispersion of the forecast, i.e.
-the length of this interval (since the observed value falls within the prediction interval).
-Forecaster A would receive a score of $2 - 0.5$ = 1.5, beating forecaster B who would receive a score of $2.5 - 0.9 = 1.6$.
-When we transform the forecast, forecaster A receives a score of $\log (\frac{2}{0.5}) = 1.39$, while forecaster B obtains a score of $\log (\frac{2.5}{0.9}) = 1.02$ and now wins over forecaster A.
-The main difference between forecaster A and B is that forecaster A has issued a prediction interval ([0.5, 2]) that was more symmetric around the true value than the interval issued by forecaster B ([0.9.
-2.6], which was more skewed).
-This means that in relative terms, Forecaster B was noticeably closer to the true value with the lower bound of their prediction interval, while only being slightly further away with the upper interval.
-This implies that scores on the log scale may treat (one-sided) outlier forecasts more leniently.
+In particular, on the natural scale, forecaster A, who issues a more uncertain distribution, receives a better score for observed values far away from the centre of the respective predictive distribution than forecaster B.
+On the log scale, however, forecaster A receives a lower score for larger observed values, being more heavily penalised for assigning large probability to small values (which, in relative terms, are far away from the actual observation).
+Similarly, we would expect assigning large probability to very large values to be less severely penalised on the log scale than on the natural scale, as usually the distance between the observation and forecast that overpredicts tends to be large in absolute terms, but small in relative terms.
 
@@
 
-\subsection{WIS}
+\subsection{Alternative Formulation of the WIS}
 For a single prediction interval, the interval score is computed as the sum of three penalty components, dispersion (width of the prediction interval), underprediction and overprediction,  
@@
 
-Computing a score based on the logarithm of the forecast and the observation is also equivalent to scoring a forecast of the exponential growth rate of the forecast target.
-The exponential growth rate $r$ governs epidemic growth via (see e.g.
-\cite{wallingaHowGenerationIntervals2007})
-%
-\begin{linenomath*}
-\begin{equation*}
-y_{t + \Delta t} = y_t \cdot \exp(r \cdot \Delta t).
-\end{equation*}
-\end{linenomath*}
-For small $\Delta t$ and $r$ this is approximately equivalent to
-\begin{linenomath*}
-\begin{equation*}
- y_{t + \Delta t} = y_t \cdot (1 + r \cdot \Delta t),
-%y(t) \cdot \exp(r \cdot \Delta t) \approx y(t) \cdot (1 + r \cdot \Delta t),
-\end{equation*}
-\end{linenomath*}
-%
-which is sometimes used as an alternative definition ($r$ in this formulation is then called the multiplicative growth rate).
-Now assume that we predict one step time step ahead, i.e., at time $t$ we predict $y_{t + 1} = \hat{y}_{t + 1} + \varepsilon^*_{t+1}$.
-Moreover assume that $r$ is time-constant between $t$ and $t + 1$.
-The score is determined by the size of $\left|\varepsilon^*_{t+1}\right|$, the absolute error of the log-transformed incidence.
-%
-\begin{linenomath*}
-\begin{align*}
-%\varepsilon^*_{t+1} & = \left|\log \hat{y}_{t + 1} - \log y_{t + 1} \right|\\
-\varepsilon^*_{t+1} & = \log \hat{y}_{t + 1} - \log y_{t + 1} \\
-& = \log \left[\exp(\hat{r}) \cdot y_t\right] - \log \left[\exp(r) \cdot y_t\right] \\
-& = \hat{r} - r,
-\end{align*}
-\end{linenomath*}
-where $\hat{r}$ is the predicted exponential growth rate as implied by $\hat{y}(t + 1)$.
-By log-transforming both the predicted value and the observation prior to applying the absolute error we thus just assess the error in the predicted exponential growth rate.
-If we apply the same reasoning to an $n$-week-ahead forecast and assume that the growth rate is piece-wise constant with value $r_t$ on $[t, t + 1)$, we get
-%
-\begin{linenomath*}
-\begin{align*}
-\epsilon^* & = \log \hat{y}_{t + n} - \log y_{t + n}\\
-& = \log\left[\exp\left(\sum_{i = 0}^{n - 1}\hat{r}_{t + i}\right) \cdot y_t\right] - \log\left[\exp\left(\sum_{i = 0}^{n - 1}r_{t + i} \right) \cdot y_t\right]\\
-& =   \sum_{i = 0}^{n - 1}\hat{r}_{t + i} \ - \ \sum_{i = 0}^{n - 1}r_{t + i}.
-\end{align*}
-\end{linenomath*}
-We thus evaluate the error in the cumulative growth over the considered $n$-week period.
-
 \clearpage

### 2022-12-06 analysis/00 full paper code.R
update plots

@@ -13,6 +13,7 @@ library(performance)
 library(tidytext)
 library(grid)
 library(ggpubr)
+library(latex2exp)
 
 set.seed(1234)
 
@@ -171,53 +172,106 @@ ggsave(filename = "output/figures/illustration-effect-log-ranking-crps.png",
 ## ========================================================================== ##
 ## FIGURE 3
 ## ========================================================================== ##
-
-
-time_points <- 5000
-q = c(0.01, 0.025, seq(0.05, 0.95, 0.05), 0.975, 0.99)
-n_means <- 100
-means = 100 * (10^seq(0, 2, length.out = n_means))
-thetas <- c(0.1, 1, 1e9)
-model_names <- c("theta_0.1", "theta_1", "theta_1e9")
-
 if (file.exists("output/data/simulation-negative-binom.Rda")) {
   scores <- readRDS(file = "output/data/simulation-negative-binom.Rda")
 } else {
+  
+  time_points <- 1e5
+  n_means <- 60
+  means = 100 * (10^seq(0, 1.2, length.out = n_means))
+  model_names <- c("theta_1000_sqrt_mu", "theta_1", "theta_1e9")
+  
   df <- data.table(
     state = 1:length(mean),
     mean = means
   ) |>
     mutate(date = list(1:time_points)) |>
     unnest(cols = date) |>
-    mutate(theta = list(thetas), 
+    rowwise() |>
+    mutate(theta = list(c(1000/sqrt(mean), 10, 1e9)), 
            model = list(model_names)) |>
+    ungroup() |>
     unnest(cols = c("theta", "model")) |>
-    mutate(true_value = rnbinom(n = time_points * n_means * length(thetas), 
-                                size = theta,
-                                mu = mean)) |>
-    mutate(quantile = list(q)) |>
-    unnest(quantile) |>
-    mutate(prediction = qnbinom(p = quantile, size = theta, mu = mean))
+    mutate(predictive_sample = rnbinom(n = time_points * n_means * length(model_names), 
+                                        size = theta,
+                                        mu = mean)) 
   
-  scores_natural <- df |>
-    score(metrics = c("interval_score")) |>
-    summarise_scores(c("model", "mean", "theta")) |>
-    mutate(scale = "natural")
-  
-  scores_log <- df |>
[…]

### 2022-12-06 analysis/Analysis ECDC forecasts.Rmd

@@
 library(ggpubr)
+library(gh)
 
@@
   unique()
+
+## load anomalies (retrospecrtively revised data)
+anomalies_file <- here::here("data", "anomalies.csv")
+if (file.exists(anomalies_file)) {
+  anomalies <- data.table::fread(file = anomalies_file)
+} else {
+  ## get anomalies file at date of last data made
+  owner <- "covid19-forecast-hub-europe"
+  repo <- "covid19-forecast-hub-europe"
+  path <- "data-truth/anomalies/anomalies.csv"
+  commit <- gh::gh(
+    "/repos/{owner}/{repo}/commits?path={path}&until={date}",
+    owner = owner,
+    repo = repo,
+    path = path,
+    until = max(hub_data$target_end_date),
+    .limit = 1
+  )
+  anomalies <- data.table::fread(input = URLencode(URL = paste(
+    "https://raw.githubusercontent.com", owner, repo, commit[[1]]$sha, path, sep = "/"
+  )))
+  data.table::fwrite(x = anomalies, file = anomalies_file)
+}
+
+## prepare for mergeing into hub_data
+anomalies <- anomalies[, list(
+  target_end_date,
+  location,
+  target_type = paste0(stringr::str_to_title(stringr::str_remove(
+    string = target_variable, pattern = "^inc ")), "s")
+)]
+
+hub_data <- hub_data[!anomalies, on = .(target_end_date, location, target_type)]
 

### 2022-12-06 analysis/00 full paper code.R
update plots and code

@@ -14,6 +14,8 @@ library(tidytext)
 library(grid)
 library(ggpubr)
 library(latex2exp)
+library(gh)
+library(covidHubUtils) # remotes::install_github("https://github.com/reichlab/covidHubUtils")
 
 set.seed(1234)
 
@@ -224,7 +226,7 @@ p1 <- scores |>
   labs(y = "WIS", x = "Mean") +
   theme_scoringutils() + 
   scale_colour_discrete(
-    labels=c(TeX(r'($\sigma^2 = \mu + \mu^2$)'), 
+    labels=c(TeX(r'($\sigma^2 = \mu + 0.1 \cdot \mu^2$)'), 
              TeX(r'($\sigma^2 = \mu + \frac{\mu^{2.5}}{1000}$)'), 
              TeX(r'($\sigma^2 = \mu$)'))
   ) +
@@ -250,7 +252,7 @@ scores |>
   ungroup() |>
   mutate(
     model = factor(model, 
-                   labels = c(TeX(r'($\sigma^2 = \mu + \mu^2$)'), 
+                   labels = c(TeX(r'($\sigma^2 = \mu + 0.1 \cdot \mu^2$)'), 
                               TeX(r'($\sigma^2 = \mu + \mu^{2.5}/1000$)'), 
                               TeX(r'($\sigma^2 = \mu$)')))
   ) |>
@@ -260,53 +262,13 @@ scores |>
   facet_wrap(scale ~ model, scales = "free", labeller = label_parsed) + 
   theme_scoringutils() + 
   scale_colour_discrete(
-    labels=c(TeX(r'($\sigma^2 = \mu + \mu^2$)'), 
+    labels=c(TeX(r'($\sigma^2 = \mu + 0.1 \cdot \mu^2$)'), 
              TeX(r'($\sigma^2 = \mu + \frac{\mu^{2.5}}{1000}$)'), 
              TeX(r'($\sigma^2 = \mu$)'))
   ) + 
   scale_x_continuous(trans = "log10")
 
 ggsave("output/figures/SIM-score-approximation.png", width = 7, height = 4.1)
-
-
-
-
-
-
-
-
-p1 <- scores |>
-  rename(Theta = theta) |>
-  group_by(mean, scale, Theta) |>
-  summarise(interval_score = mean(interval_score)) |>
-  group_by(Theta, scale) |>
-  mutate(interval_score = interval_score / mean(interval_score), 
-         Theta = as.character(Theta),
[…]

### 2022-12-06 main.tex
Updates from Overleaf

@@
 
-\title{Transformations of forecasts for evaluation in an epidemiological context}
+\title{Transformation of forecasts for evaluating predictive performance in an epidemiological context}
   \author{Nikos I.
@@
 \end{linenomath*}
-where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, $K$ is the number of central prediction intervals provided, and $m$ is the median forecast.
+where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, $K$ is the number of predictive quantiles provided, and $m$ is the median forecast.
 
@@
 We provide a mathematical intuition of the impact of the log-transformation, its interpretation (Section \ref{sec:methods:relative} and \ref{sec:methods:growthrate}) and effects on forecast rankings (section \ref{sec:methods:rankings} and provide some discuss some practical considerations relevant for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}).
-We illustrate our the effect of the log-transformation using forecasts submitted to the European COVID-19 Forecast Hub  \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022} (Section \ref{sec:results}).
+We illustrate the effect of the log-transformation using forecasts submitted to the European COVID-19 Forecast Hub  \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022} (Section \ref{sec:results}).
 Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues on the use of transformed scores in the evaluation of infectious disease forecasts (Section \ref{sec:discussion}).
@@
 
-\subsection*{Interpretation as a relative error}
+\subsection{Interpretation as a relative error}
 \label{sec:methods:relative}
@@
 
-\subsection*{Interpretation as scoring the exponential growth rate}
+\subsection{Interpretation as scoring the exponential growth rate}
 \label{sec:methods:growthrate}
@@
 
-If a forecast $y$ of the time series at time $t$ starts at time 0 with the data point $y_0$ then the absolute error after log transformation at time $t$ is
+If a forecast $\hat{y}(t)$ for the value of the time series at time $t$ is issued at time 0 based on the data point $y_0$ then the absolute error after log transformation is
 %
@@
 We thus evaluate the error in the mean exponential growth rate, scaled by the length of the time period considered.
+For small $\bar{r}t$, $ y_0 \exp (\bar{r}t)$ is approximately equivalent to $ y_0 (1 + \bar{r}t)$, which is sometimes used as an alternative definition of the growth rate (then called the multiplicative growth rate).
 
-For small $\bar{r}t$, $ y_0 \exp (\bar{r}t)$ is approximately equivalent to $ y_0 (1 + \bar{r}t)$, which is sometimes used as an alternative definition (r in this second formulation is then called the multiplicative growth rate).
+\subsection{Interpretation as a variance-stabilising transformation}
+
+In epidemiological contexts, the variance of the variable of interest usually scales with its mean.
+For a poisson-distributed variable, for example, the variance equals its mean, and for a negative-binomially distributed variable the relationship can be described by $\sigma^2 = \mu = \frac{\mu^2}{\theta}$.
+A variance transformation (VST, CITATION) is a transformation that serves to make the variance of a variable independent of its mean.
+
+If the predictive distribution for the quantity $Y$ equals the true data-generating process $F$, the expected CRPS is given by \citep{gneitingStrictlyProperScoring2007}
+$$
+\mathbb{E}[\text{CRPS}(F, y)] = 0.5\times\mathbb{E}|Y - Y'|,
+$$
+where $Y$ and $Y'$ are independent samples from $F$.
+This corresponds to half the \textit{mean absolute difference}, which is a measure of dispersion.
+If $F$ is well-approximated by a normal distribution $\text{N}(\mu, \sigma^2)$, the approximation
+$$
+\mathbb{E}_F[\text{CRPS}(F, y)] \approx \frac{\sigma}{\sqrt{\pi}}
+$$
+can be used, meaning that the expected CRPS scales roughly with the standard deviation. Denoting by $F^*$ the predictive distribution for $\log(Y)$, the delta method implies that
+$$
+\mathbb{E}_F[\text{CRPS}\{F^*, \log(y)\}] \approx \frac{\sigma/\mu}{\sqrt{\pi}}.
+$$
+The expected CRPS for the log-transformed outcome thus scales with the coefficient of variation.
+In instances where the variance scales with the square of the mean, the log-transformation serves as a variance-stabilising transformation, making the expected CRPS independent of the mean of the forecast quantity.
+If the observations follow a poisson-distribution where the variance equals the mean, then a square-root-transformation can serve as VST.
+
+RUN REGRESSIONS
 
@@
 When showing the obtained score as a function of the observed value, it can be seen that the ranking between the two forecasters may change when scoring the forecast on the logarithmic, rather than the natural scale.
-In particular, on the natural scale, forecaster A, who issues a more uncertain distribution, receives a better score for observed values far away from the centre of the respective predictive distribution than forecaster B.
-On the log scale, however, forecaster A receives a lower score for larger observed values, being more heavily penalised for assigning large probability to small values (which, in relative terms, are far away from the actual observation).
-Similarly, we would expect assigning large probability to very large values to be less severely penalised on the log scale than on the natural scale, as usually the distance between the observation and forecast that overpredicts tends to be large in absolute terms, but small in relative terms.
+In particular, on the natural scale, forecaster A, who issues a more uncertain distribution, receives a better score than forecaster B for observed values far away from the centre of the respective predictive distribution.
+On the log scale, however, forecaster A receives a lower score for large observed values, being more heavily penalised for assigning large probability to small values (which, in relative terms, are far away from the actual observation).
 
@@
 The variance of the negative binomial is given as $\sigma^2 = \mu + \mu^2 / \theta$, meaning that for large theta the negative binomial distribution is equal to the Poisson distribution.
-We used values of $\theta = 0.1$ (red), 1 (green) and 1b (blue).
+We used values of $\theta = 0.1$ (red), 1 (green) and 1 billion (blue).
 To make the scores for the different distributions comparable, scores were normalised to one, meaning that the mean score for every distribution (red, green, blue) is one.
@@
 
-Whether or not forecast targets with smaller quantities receive higher scores than those for targets with larger quantities, depends on the exact relationship between the mean and the variance of the quantity of interest.
+Whether forecast targets of small or high order of magnitude receive higher scores depends on the relationship between the mean and the variability of the quantity of interest.
+
+MENTION EXAMPLE
+
 To provide a better intuition, we simulated count data following a negative binomial distribution.
@@
 When taking the logarithm of the WIS, scores are minimised for a predictive distribution that is too narrow compared to the data-generating distribution.}
+    \caption{\textbf{Alternative suggestions for legend:} Illustration of impropriety of log-transformed CRPS.
+We assume $Y \sim \text{LogNormal}(0, 1)$ and evaluate the expected CRPS for predictive distributions $\text{LogNormal}(0, \sigma)$ with varying values of $\sigma \in [0.1, 2]$.
+For the regular CRPS (left) and CRPS applied to log-transformed outcomes (middle), the lowest expectation is achieved for the true value $\sigma = 1$.
+For the log-transformed CRPS, the optimal value is \textbf{XXX}, i.e.
+there is an incentive to report a too sharp forecast.}
     \label{fig:log-improper}
@@
 In this paper we proposed a log-transformation of forecasts and observed values prior to evaluation with the WIS (or CRPS) in order to address issues that arise when evaluating epidemiological forecasts based on a measure of absolute error.
-We gave a mathematical intuition for the transformation and showed that scores on the log scale can be interpreted as a) a measure of relative (multiplicative) prediction errors, as well as b) a score for a forecast of the exponential growth rate of the target quantity.
+We gave a mathematical intuition for the transformation and showed that scores on the log scale can be interpreted as a) a measure of relative prediction errors, as well as b) a score for a forecast of the exponential growth rate of the target quantity.
 When applying this approach to forecasts from the European COVID-19 Forecast Hub, we found overall scores on the log scale to be more equal across, time, location and target type (cases, deaths) than scores on the natural scale.
@@
 
-Applying a log-transformation prior to the WIS means that forecasts are evaluated in terms of relative errors and errors on the multiplicative growth rate, rather than absolute errors.
+Applying a log-transformation prior to the WIS means that forecasts are evaluated in terms of relative errors and errors on the exponential growth rate, rather than absolute errors.
 The most important strength of this approach is that the evaluation better accommodates the exponential nature of epidemiological process and the types of errors forecasters who accurately model those processes are expected to make.

### 2022-12-09 analysis/00 full paper code.R
update plots and code

@@ -19,10 +19,6 @@ library(covidHubUtils) # remotes::install_github("https://github.com/reichlab/co
 
 set.seed(1234)
 
-recompute_scores <- FALSE
-
-# situations with negative values: France 2021-05-22, CZ 2021-08-07, ES 2021-03-06
-
 ## ========================================================================== ##
 ## FIGURE 1
 ## ========================================================================== ##
@@ -50,19 +46,20 @@ plot_fct <- function(data) {
          x = "Observed value")
 }
 
-label_fn <- function(x) {
-  ifelse(x%%1 == 0, x, 
-         paste0("1/", 1/x))
+label_fn_hat <- function(x) {
+  ifelse(x%%1 == 0, 
+         TeX(paste0(x, "\\hat{y}")), 
+         TeX(paste0("1/", 1/x, "\\hat{y}")))
 }
 
 p1 <- plot_fct(data) +
-  scale_x_continuous(breaks = c(1/4, 0.5, 1, 2, 4), 
-                     labels = label_fn) + 
+  scale_x_continuous(breaks = c(1/4, 1, 2, 4), 
+                     labels = label_fn_hat) + 
   scale_color_brewer(palette = "Set1")
 
 p2 <- plot_fct(data) +
   scale_x_continuous(trans = "log10", breaks = c(1/4, 0.5, 1, 2, 4), 
-                     labels = label_fn) + 
+                     labels = label_fn_hat) + 
   scale_color_brewer(palette = "Set1")
 
 (p1 + p2) +
@@ -80,8 +77,8 @@ ggsave("output/figures/different-relative-errors.png", width = 7, height = 3)
 
 quantile <- seq(0.005, 0.995, 0.005)
 mu <- 10
-theta_1 <- 0.5
-theta_2 <- 10
+theta_1 <- 1
+theta_2 <- 1e9
 
 data <- data.table(
   quantile = quantile,
@@ -116,7 +113,7 @@ nbinom_natural <- data.table(
   pivot_longer(cols = c(A, B), names_to = "Forecaster") |>
   mutate(scale = "natural")
 
-nbinom_log <- nbinom |>
+nbinom_log <- nbinom_natural |>
   mutate(value = log(value), 
          scale = "log")
 
@@ -149,18 +146,26 @@ plot_fct <- function(scores, scale_factor, nbinom, filter_scale = "natural") {
     theme_scoringutils() + 
[…]

### 2022-12-09 analysis/Refetch data from Forecast Hub.R

@@ -1,0 +1,106 @@
+# Code to refetch data from the Forecast Hub
+  
+# download folder using SVN
+# svn checkout https://github.com/epiforecasts/covid19-forecast-hub-europe/trunk/data-processed
+
+# load truth data using the covidHubutils package ------------------------------
+truth <- covidHubUtils::load_truth(hub = "ECDC") |>
+  filter(target_variable %in% c("inc case", "inc death")) |>
+  mutate(target_variable = ifelse(target_variable == "inc case", 
+                                  "Cases", "Deaths")) |>
+  rename(target_type = target_variable, 
+         true_value = value) |>
+  select(-model)
+
+fwrite(truth, "data/weekly-truth-Europe.csv")
+
+
+# get the correct file paths to all forecasts ----------------------------------
+folders <- here("data-processed", list.files("data-processed"))
+folders <- folders[
+  !(grepl("\\.R", folders) | grepl(".sh", folders) | grepl(".csv", folders))
+]
+
+file_paths <- purrr::map(folders, 
+                         .f = function(folder) {
+                           files <- list.files(folder)
+                           out <- here::here(folder, files)
+                           return(out)}) %>%
+  unlist()
+file_paths <- file_paths[grepl(".csv", file_paths)]
+
+# load all past forecasts ------------------------------------------------------
+# ceate a helper function to get model name from a file path
+get_model_name <- function(file_path) {
+  split <- str_split(file_path, pattern = "/")[[1]]
+  model <- split[length(split) - 1]
+  return(model)
+}
+
+# load forecasts
+prediction_data <- map_dfr(file_paths, 
+                           .f = function(file_path) {
+                             data <- fread(file_path)
+                             data[, `:=`(
+                               target_end_date = as.Date(target_end_date),
+                               quantile = as.numeric(quantile),
+                               forecast_date = as.Date(forecast_date), 
+                               model = get_model_name(file_path)
+                             )]
+                             return(data)
+                           }) %>%
+  filter(grepl("case", target) | grepl("death", target)) %>%
+  mutate(target_type = ifelse(grepl("death", target), 
+                              "Deaths", "Cases"), 
+         horizon = as.numeric(substr(target, 1, 1))) %>%
+  rename(prediction = value) %>%
+  filter(type == "quantile", 
+         grepl("inc", target)) %>%
[…]

### 2022-12-09 main.tex

@@
 \begin{equation*}
-    \text{WIS}(F, y) = \frac{1}{2K + 1} \times \sum_{k = 1}^{2K + 1} \times 2 \left[ 1(y \leq q_{\tau_k}) - \tau_k \right] \times ( q_{\tau_k} - y), 
+    \text{WIS}(F, y) = \frac{1}{K} \times \sum_{k = 1}^{K} \times 2 \left[ 1(y \leq q_{\tau_k}) - \tau_k \right] \times ( q_{\tau_k} - y), 
 \end{equation*}
 \end{linenomath*}
-where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, $K$ is the number of predictive quantiles provided, and $m$ is the median forecast.
+where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, and $K$ is the number of predictive quantiles provided.
 
@@
 Moreover, if forecasts are evaluated based on absolute rather than relative errors, whereas in reality errors scale with the level of observations, scores are strongly influenced by the overall order of magnitude of the quantity to forecast.
-This makes it difficult to compare predictive performance across time or to compare performance across targets that generally occur at different orders of magnitudes.
-In the example of COVID-19 where case numbers underwent drastic changes over time following the initial emergence, forecast evaluation based on the absolute numbers would be dominated by performance around the peaks, and evaluation based on jointly evaluating the ability to predict numbers of cases, hospitalisation and deaths dominated by cases as they occur at largest numbers.
+There exist other scores that do not depend on the scale of the forecast target, such as the log score \citep{goodRationalDecisions1952}, which in addition is also invariant under monotonic transformations of the target, or a scaled version of the CRPS \citep{bolinLocalScaleInvariance2021}.
+These, however, may not have all of the desired properties of the CRPS and in particular cannot easily be transferred to a setting with predictive quantiles.
+With the regular CRPS and WIS the dependence on the scale of the forecast target makes it difficult to compare predictive performance across time and in different phases of an epidemic or to compare performance across targets that generally occur at different orders of magnitudes.
+In the example of COVID-19 where case numbers underwent drastic changes over time following the initial emergence, forecast evaluation based on the absolute numbers would be dominated by performance around the peaks.
+The ability to correctly predict future upswings while case numbers are still low would count little towards overall model scores.
+Scores based on jointly evaluating the ability to predict numbers of cases, hospitalisation and deaths would be dominated by cases as they occur at largest numbers.
 
-In this paper, we discuss how transforming the forecasts and observations prior to applying the WIS (or CRPS) may be an appropriate way to address the aforementioned issues.
+In many fields it is common practice to forecast transformed quantities (see e.g.
+\cite{taylorEvaluatingVolatilityInterval1999} in finance, \cite{mayrLogLevelVAR2015} in macroeconomics, \cite{loweStochasticRainfallrunoffForecasting2014} in hydrology or \cite{fuglstadDoesNonstationarySpatial2015} in meteorology).
+The goal of the transformations is usually to improve accuracy of the predictions.
+In this paper, we discuss how transforming the forecasts and observations prior to applying the WIS (or CRPS) may be an appropriate way to improve the evaluation process by addressing the aforementioned issues.
 Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects of the forecasts that forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).
 For conceptual clarity and to allow for a more in-depth discussion, we focus most of our attention on the log-transformation as one example of many possible transformations, chosen for its simplicity and usefulness in an epidemiological context.
-We further focus on the WIS as evaluation metric, given its widespread use throughout the COVID-19 pandemic.
+We further focus on the CRPS/WIS as evaluation metric, given its widespread use throughout the COVID-19 pandemic.
 
@@
 A particularly relevant transformation in the context of epidemic phenomena is the natural logarithm.
-Instead of a score representing the magnitude of absolute errors this approach makes it possible to obtain a score which a) measures relative error and at the same time b) provides an a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate}.
+Instead of a score representing the magnitude of absolute errors this approach makes it possible to obtain a score which a) measures relative error and at the same time b) provides an a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate} and is c) independent from the mean of the observed quantity (see Section \ref{sec:methods:vst}).
 The interpretation of the log-transformed score as a score for a forecast of the exponential growth rate of a given quantity aligns the evaluation with our understanding of how epidemiological processes should be modeled and the kinds of errors models would typically make.
@@
 $$
-The absolute error after log transformation is thus an approximation of the \textit{absolute percentage error} (APE) \cite{Gneiting2011JASA}.
+The absolute error after log transformation is thus an approximation of the \textit{absolute percentage error} \citep[APE,][]{gneitingMakingEvaluatingPoint2011a}.
 As in the approximation we assumed that $\hat{y} \approx y$, we can just as well interpret it as an approximation of the \textit{relative error} (RE)
@@
 \subsection{Interpretation as a variance-stabilising transformation}
+\label{sec:methods:vst}
 
 In epidemiological contexts, the variance of the variable of interest usually scales with its mean.
-For a poisson-distributed variable, for example, the variance equals its mean, and for a negative-binomially distributed variable the relationship can be described by $\sigma^2 = \mu = \frac{\mu^2}{\theta}$.
-A variance transformation (VST, CITATION) is a transformation that serves to make the variance of a variable independent of its mean.
+A variance-stabilising transformation \citep[VST][]{bartlettSquareRootTransformation1936} is a transformation that serves to make the variance of a variable independent of its mean.
 
-If the predictive distribution for the quantity $Y$ equals the true data-generating process $F$, the expected CRPS is given by \citep{gneitingStrictlyProperScoring2007}
+If the predictive distribution for the quantity $Y$ equals the true data-generating process $F$ (an ideal forecast), the expected CRPS is given by \citep{gneitingStrictlyProperScoring2007}
 $$
@@
 The expected CRPS for the log-transformed outcome thus scales with the coefficient of variation.
-In instances where the variance scales with the square of the mean, the log-transformation serves as a variance-stabilising transformation, making the expected CRPS independent of the mean of the forecast quantity.
-If the observations follow a poisson-distribution where the variance equals the mean, then a square-root-transformation can serve as VST.
-
-RUN REGRESSIONS
+In instances where the variance scales with the square of the mean, the log-transformation serves as a variance-stabilising transformation \citep{guerreroTimeseriesAnalysisSupported1993}, making the expected CRPS independent of the mean of the forecast quantity.
+We illustrate this in Figure \ref{fig:SIM-wis-state-size-mean}, where we show CRPS scores for different ideal forecasters.
+For a negative-binomial distribution with $\sigma^2 = \mu + \mu^2$ and sufficiently large $\mu$ we see that on the log scale scores are independent of the mean of the forecast target.
+For a Poisson-distribution with
+$\sigma = \mu$, scores on the log scale decrease as the mean increases.
+The appropriate variance-stabilising transformation here to make scores independent of the mean would be a square-root-transformation \citep{bartlettSquareRootTransformation1936}.
 
@@
 Rankings between different forecasters based on their scores may change through the log-transformation both when looking at aggregate scores as well as when looking at a single forecast.
-We illustrate this in Figure \ref{fig:illustration-ranking} with two forecasters, A and B, issuing a negative binomial distribution with different dispersion.
+We illustrate this in Figure \ref{fig:illustration-ranking} with two forecasters, A and B, issuing two different distributions with different dispersion.
 When showing the obtained score as a function of the observed value, it can be seen that the ranking between the two forecasters may change when scoring the forecast on the logarithmic, rather than the natural scale.
@@
 Shown are CRPS (or WIS, respectively) values as a function of the observed value for two forecasters.
-Model A issues a negative binomial forecast with mean = 10 and $\theta = 0.5$ (resulting in a variance of 210), while Model B issues a negative binomial forecast with mean = 10 and $\theta = 10$ (resulting in a variance of 20).}
+Model A issues a negative binomial forecast with mean = 10 and $\theta = 1$ (resulting in a variance of 110), while Model B issues a Poisson distribution with mean = 10 (resulting in a variance of 10).}
 \label{fig:illustration-ranking}
@@
     \includegraphics[width=0.99\textwidth]{output/figures/SIM-mean-state-size.png}
-    \caption{Simulation showing how the relationship between mean and variance of the forecast quantity affects scores.
-1,000 samples were drawn from different negative binomial distributions and scores were computed assuming an ideal forecaster (forecast distribution equal to data-generating distribution).
-Observations were simulated for different means of the negative-binomial distributions (representing, for example, the number of cases in differently sized states) as well as different relationships between mean and variance (representing, for example, three different infectious processes).
+    \caption{Expected CRPS scores as a function of the mean and variance of the forecast quantity.
+We generated 10,000 samples for three distributions with differing relationships between mean and variance.
+We then computed expected CRPS values assuming an ideal forecaster with predictive distribution equal to the data-generating distribution.
 The variance of the negative binomial is given as $\sigma^2 = \mu + \mu^2 / \theta$, meaning that for large theta the negative binomial distribution is equal to the Poisson distribution.
-We used values of $\theta = 0.1$ (red), 1 (green) and 1 billion (blue).
+We used values of $\theta = 1$ (red), $\theta = \frac{1000}{\sqrt{\mu}}$ (green) and $\theta = 10^9$ (blue, approximately corresponding to a Poisson distribution).
 To make the scores for the different distributions comparable, scores were normalised to one, meaning that the mean score for every distribution (red, green, blue) is one.
-    A: Normalised WIS for ideal forecasts with increasing means of three distribution with different relationships between mean and variance. B: A but with log scale axis.}
+    A: Normalised expected CRPS for ideal forecasts with increasing means for three distribution with different relationships between mean and variance. B: A but with x axis on the log scale.}
     \label{fig:SIM-wis-state-size-mean}
@@
 
-Whether forecast targets of small or high order of magnitude receive higher scores depends on the relationship between the mean and the variability of the quantity of interest.
-
-MENTION EXAMPLE
-
-To provide a better intuition, we simulated count data following a negative binomial distribution.
-For these simulated data, log predictions for small quantities received on average higher scores than forecasts for large quantities if the mean and the variance grew at the same rate $\sigma^2 = \mu$, about equal scores when the variance grew at a rate of $\sigma^2 = \mu + \mu^2$, and smaller scores when the variance grew faster than that (Figure \ref{fig:SIM-wis-state-size-mean}).
+Whether forecasts of small or large targets receive higher scores than those for targets with larger quantities, depends on the relationship between the mean and the variability of the quantity of interest (see Section \ref{sec:methods:vst}.
+For the simulated count data, log predictions for small quantities received on average higher scores than forecasts for large quantities if the mean and the variance grew at the same rate $\sigma^2 = \mu$, about equal scores when the variance grew at a rate of $\sigma^2 = \mu + \mu^2$ (which for large $\mu$ is dominated by the quadratic term), and smaller scores when the variance grew faster than that (Figure \ref{fig:SIM-wis-state-size-mean}).
 
@@
     \includegraphics[width=0.99\textwidth]{output/figures/example-log-first.png}
-    \caption{Scores for different forecasts evaluated using the WIS, the WIS on the log-scale and the logarithm of the WIS.
-We simulated 1000 observations $Y_i = {\rm e}^{x_i}$, with $x_i \text{ iid} \sim \mathcal{N}(0, 1)$.
-We then simulated 20 forecasters who would issue a predictive distribution $F = {\rm e}^{x_i}$, with $x \sim \mathcal{N}(0, \sigma)$, with values of $\sigma$ ranging from 0.1 to 2.
-For forecasts evaluated by the WIS and the WIS on log-transformed values, the score is minimised when the predictive distribution equals the data-generating distribution.
-When taking the logarithm of the WIS, scores are minimised for a predictive distribution that is too narrow compared to the data-generating distribution.}
-    \caption{\textbf{Alternative suggestions for legend:} Illustration of impropriety of log-transformed CRPS.
+    \caption{Illustration of impropriety of log-transformed CRPS.
 We assume $Y \sim \text{LogNormal}(0, 1)$ and evaluate the expected CRPS for predictive distributions $\text{LogNormal}(0, \sigma)$ with varying values of $\sigma \in [0.1, 2]$.
 For the regular CRPS (left) and CRPS applied to log-transformed outcomes (middle), the lowest expectation is achieved for the true value $\sigma = 1$.
-For the log-transformed CRPS, the optimal value is \textbf{XXX}, i.e.
+For the log-transformed CRPS, the optimal value is  0.9, i.e.
 there is an incentive to report a too sharp forecast.}
@@
 We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 16 forecast dates (see Figure \ref{fig:HUB-num-avail-models}).
+We removed all observations marked as data anomalies by the European Forecast Hub as well as all remaining negative observed values.
+All predictive quantiles were truncated at 0.
+We applied the log-transformation after adding 1 to all predictions and observed values.
 Where not otherwise stated, we report results for a two-week-ahead forecast horizon.
@@
 Both on the natural scale as well to a slightly lesser extent on the log scale, scores increased considerably with increasing forecast horizon (see Figure \ref{fig:HUB-mean-locations}E).
-We hypothesise that this increase in scores across horizons represents the relative difficulty to forecast values into the future, which was noticeable both on the natural as well as the logarithmic scale.
+This reflects the increasing difficulty of forecasts further into the future.
 
@@
     \includegraphics[width=0.99\textwidth]{output/figures/HUB-correlations.png}
-    \caption{Correlations between observations and scores.
-A: Mean WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against the average number of observations in a location.
+    \caption{Correlations for median forecasts, scores, and rankings on the natural and logarithmic scale.
+A: WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against median predicted values.
 B: Average Spearman rank correlation of scores for individual forecasts from all models.
@@
 on the log scale.
-All rank correlations were averaged and stratified by horizon and target type.
+All rank correlations were averaged across locations, and target types and stratified by horizon and target type.
 C: Correlation between relative skill scores.
@@
 We report the Spearman rank correlation here, rather than the Pearson correlation, so that correlations remain the same regardless of whether we look at observed values on the log or natural scale.
-On the log scale, the correlation is much weaker and slightly negative, at least in terms of rank correlation (Spearman correlation: -0.33 and -0.36), implying a tendency towards larger scores in countries with lower mean incidences.
+On the log scale, the correlation is much weaker and slightly negative, at least in terms of rank correlation (Spearman correlation: CHECK AGAIN -0.33 and -0.36), implying a tendency towards larger scores in countries with lower mean incidences.
+
+In order to test the relationship between scores and the order of magnitude of the forecast, we ran the following regression on scores on the natural scale
+
+\begin{equation*}
+    \text{CRPS}(F, y) = \exp(\alpha) \cdot \text{median}(F)^{\beta}, 
+\end{equation*}
+where $F$ is the predictive distribution and $y$ the observed value. In addition, we ran the following regression on scores on the log scale
+
+\begin{equation*}
+    \text{CRPS}(\log F, \log y) = \alpha^* + \beta^* \cdot \log{(\text{median}(F))}. 
+\end{equation*}
+
+We found that on the natural scale, the order of magnitude of the forecast (represented by the median, rather than the mean, as forecasts were stored as quantiles and the mean was not available) strongly influenced scores, with estimated $\beta$-values close to one.
+On the natural, this influence was small, with estimated $\beta^*$-values close to zero.
+Both on the natural and on the log scale we found scores to increase over time (estimated $\alpha$- and $\alpha^*$-values increased over horizons).
+Estimates are summarised in Table \ref{tab:HUB-regression}.
 
@@
 \toprule
-        target\_type & quantity & measure & natural & log\\
+Target & Quantity & Measure & natural & log\\
 \midrule
-        Cases & Observations & mean & 19941 & 8.33\\
-        Cases & Observations & sd & 43149 & 2.09\\
-        Cases & Observations & var & 1861850319 & 4.35\\
-        Deaths & Observations & mean & 222 & 3.65\\
-        Deaths & Observations & sd & 481 & 2.07\\
-        Deaths & Observations & var & 231025 & 4.29\\
+Cases & Observations & mean & 20013 & 8.33\\
+Cases & Observations & sd & 43256 & 2.07\\
+Cases & Observations & var & 1871039511 & 4.30\\
+\addlinespace
+Deaths & Observations & mean & 244 & 4.01\\
+Deaths & Observations & sd & 498 & 1.81\\
+Deaths & Observations & var & 248306 & 3.28\\
 \addlinespace
@@
 \addlinespace
-        Cases & WIS & mean & 4544 & 0.28\\
-        Cases & WIS & sd & 17585 & 0.49\\
-        Deaths & WIS & mean & 28 & 0.22\\
-        Deaths & WIS & sd & 67 & 0.26\\
+Cases & WIS & mean & 4220 & 0.26\\
+Cases & WIS & sd & 13978 & 0.35\\
+\addlinespace
+Deaths & WIS & mean & 30 & 0.21\\
+Deaths & WIS & sd & 70 & 0.19\\
 \bottomrule
@@
 \subsection{Alternative Formulation of the WIS}
-For a single prediction interval, the interval score is computed as the sum of three penalty components, dispersion (width of the prediction interval), underprediction and overprediction,  
+
+Instead of defining the WIS as an average of scores for individual quantiles, we can define it using an average of scores for symmetric predictive intervals.
+For a single prediction interval, the interval scoren (IS) is computed as the sum of three penalty components, dispersion (width of the prediction interval), underprediction and overprediction,
 %
@@
 the lower and upper bound of a single central prediction interval.
-For a set of $K$ prediction intervals and the median $m$, the WIS is computed as a weighted sum,
+For a set of $K^*$ prediction intervals and the median $m$, the WIS is computed as a weighted sum,
 \begin{linenomath*}
 \begin{equation*}
-\text{WIS} = \frac{1}{K + 0.5} \cdot \left(w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y)\right),    
+\text{WIS} = \frac{1}{K^* + 0.5} \cdot \left(w_0 \cdot |y - m| + \sum_{k = 1}^{K^*} w_k \cdot IS_{\alpha_{k}}(F, y)\right),    
 \end{equation*} 
@@
 where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$.
+
+\begin{table}
+\centering
+\begin{tabular}{ccrrrr}
+\toprule
+Horizon & Target & $\alpha$ & $\beta$ & $\alpha^*$ & $\beta^*$\\
+\midrule
+all & all & -0.556 & 0.877 & 0.579 & -0.028\\
+\addlinespace
+all & Cases & 1.112 & 0.710 & 1.280 & -0.101\\
+all & Deaths & -0.433 & 0.762 & 0.523 & -0.044\\
+\addlinespace
+1 & all & -0.999 & 0.855 & 0.419 & -0.031\\
+2 & all & -0.756 & 0.890 & 0.517 & -0.028\\
+3 & all & -0.430 & 0.895 & 0.628 & -0.026\\
+4 & all & -0.047 & 0.881 & 0.747 & -0.024\\
+\addlinespace
+1 & Cases & -0.065 & 0.764 & 0.768 & -0.067\\
+2 & Cases & 0.676 & 0.748 & 1.104 & -0.089\\
+3 & Cases & 1.414 & 0.710 & 1.432 & -0.109\\
+4 & Cases & 2.173 & 0.658 & 1.712 & -0.124\\
+\addlinespace
+1 & Deaths & -0.726 & 0.735 & 0.429 & -0.048\\
+2 & Deaths & -0.553 & 0.766 & 0.490 & -0.045\\
+3 & Deaths & -0.352 & 0.782 & 0.552 & -0.041\\
+4 & Deaths & -0.113 & 0.781 & 0.618 & -0.036\\
+\bottomrule
+\end{tabular}
+\caption{Coefficients of two regressions for the effect of the magnitude of the median forecast on expected scores. The first regression was 
+$\text{CRPS}(F, y) = \exp(\alpha) \cdot \text{median}(F)^{\beta},$ where $F$ is the predictive distribution and $y$ the observed value. The second one was 
+$\text{CRPS}(\log F, \log y) = \alpha^* + \beta^* \cdot \log{(\text{median}(F))}.$
+}
+\label{tab:HUB-regression}
+\end{table}
 

### 2022-12-12 analysis/00 full paper code.R
recreate figures

@@ -125,7 +125,6 @@ nbinom <- filter(nbinom, scale == "natural") |>
 plot_fct <- function(scores, scale_factor, nbinom, filter_scale = "natural") {
   
   scores <- filter(scores, scale == filter_scale)
-  # nbinom <- filter(nbinom, scale == filter_scale)
   
   scores |>
     ggplot(aes(x = id, y = interval_score / scale_factor, color = Forecaster)) + 
@@ -142,7 +141,6 @@ plot_fct <- function(scores, scale_factor, nbinom, filter_scale = "natural") {
                    alpha = 0.2,
                    binwidth = 1) +
     geom_line() +
-    # facet_wrap(~ scale, scales = "free", ncol = 2) + 
     theme_scoringutils() + 
     scale_y_continuous(label = function(x) {paste(scale_factor * x)}) + 
     labs(y = "CRPS / WIS", x = "Observed value") + 
@@ -152,7 +150,7 @@ plot_fct <- function(scores, scale_factor, nbinom, filter_scale = "natural") {
                TeX(r'(B: $\sigma^2 = \mu$)'))
     ) + 
     scale_fill_discrete(
-      labels=c(TeX(r'(A: $\sigma^2 = \mu + \cdot \mu^2$)'), 
+      labels=c(TeX(r'(A: $\sigma^2 = \mu + \mu^2$)'), 
                TeX(r'(B: $\sigma^2 = \mu$)'))
     )
 }
@@ -180,7 +178,7 @@ ggsave(filename = "output/figures/illustration-effect-log-ranking-crps.png",
 ## FIGURE 3
 ## ========================================================================== ##
 if (file.exists("output/data/simulation-negative-binom.Rda")) {
-  scores <- readRDS(file = "output/data/simulation-negative-binom.Rda")
+  scores_fig_3 <- readRDS(file = "output/data/simulation-negative-binom.Rda")
 } else {
   
   time_points <- 1e5
@@ -201,9 +199,9 @@ if (file.exists("output/data/simulation-negative-binom.Rda")) {
     unnest(cols = c("theta", "model")) |>
     mutate(predictive_sample = rnbinom(n = time_points * n_means * length(model_names), 
                                         size = theta,
-                                        mu = mean)) 
+                                        mu = mean) + 1) 
   
-  scores <- df |>
+  scores_fig_3 <- df |>
     group_by(mean, theta, model) |> 
     summarise(
[…]

### 2023-01-09 main.tex
update manuscript, address suggestions, delete paragraph

@@
 Forecast evaluation plays an essential role in improving forecasting models can help inform future decision making.
-In epidemiology, forecasts are often scored using the Weighted Interval Score (WIS) or the Continuous Ranked Probability Score (CRPS), which are a measure of absolute distance between forecast and observation.
-Infectious disease processes are commonly described in terms of the growth rate or the effective reproduction number.
-Scoring based on the absolute distance between forecast and observation may not reflect the true underlying modelling task well (estimating the growth rate or reproduction number) and makes it difficult to compare forecasts across time, space, and different targets.
-As a solution we propose transforming the data before scoring the forecasts.
+In epidemiology, probabilistic forecasts are often scored using the Weighted Interval Score (WIS) or the Continuous Ranked Probability Score (CRPS), which are a measure of absolute distance between forecast and observation.
+Infectious disease transmission is commonly conceptualised as a non-stationary processes in terms of the growth rate or the effective reproduction number.
+Scoring based on the absolute distance between forecast and observation may not reflect this underlying modelling task well and may make it difficult to compare forecasts across time, space, and different forecast targets.
+As a solution, we propose transforming the data before scoring the forecasts to better reflect our understanding of the generative process.
 Depending on the context, many transformations may be appropriate.
-We describe the log-transformation as one specific example in detail and discuss its interpretation as a measure of relative error and as score for an implicit forecast of the exponential growth rate.
-Applying the log-transformation to data from the European COVID-19 Forecast Hub, we found that it changed model rankings, evened scores across time, locations and target types, and emphasised situations in which models missed lower inflection points and consequent rises in numbers, while penalising missed peaks less severely.
+We describe the log-transformation as one specific example in detail and discuss its interpretation as a measure of relative error and as a score for an implicit forecast of the exponential growth rate.
+Applying the log-transformation to data from the European COVID-19 Forecast Hub, we found that it changed model rankings, regardless of stratification by forecast date, location or target types, and emphasised situations in which models missed lower inflection points and consequent rises in numbers, while penalising missed peaks less severely.
 \end{abstract}
@@
 \end{linenomath*}
-where $y$ is the true observed value and $F$ the cumulative distribution function (CDF) of predictive distribution.
-Both the WIS and the CRPS can be understood as a generalisation of the absolute error to predictive distributions.
-The WIS is itself an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in COVID-19 forecast hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
-The WIS is commonly defined as
+where $y$ is the true observed value and $F$ the cumulative distribution function (CDF) of the predictive distribution.
+The CRPS can be understood as a generalisation of the absolute error to predictive distributions.
+The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in COVID-19 forecast hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
+The WIS is defined as
 \begin{linenomath*}
 \begin{equation*}
-    \text{WIS}(F, y) = \frac{1}{K} \times \sum_{k = 1}^{K} \times 2 \left[ 1(y \leq q_{\tau_k}) - \tau_k \right] \times ( q_{\tau_k} - y), 
+    \text{WIS}(F, y) = \frac{1}{K} \times \sum_{k = 1}^{K} 2 \times \left[ \boldsymbol{1}(y \leq q_{\tau_k}) - \tau_k \right] \times ( q_{\tau_k} - y), 
 \end{equation*}
 \end{linenomath*}
-where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, and $K$ is the number of predictive quantiles provided.
+where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, $K$ is the number of predictive quantiles provided and $\boldsymbol{1}$ is the indicator function.
+We show an alternative definition based on central prediction intervals in Section \ref{sec:alternative-wis}.
 
-The dynamics of infectious processes are often described by the complementary concepts of the reproduction number R \citep{gosticPracticalConsiderationsMeasuring2020} and growth rate r \citep{wallingaHowGenerationIntervals2007}, where R describes the strength and r the speed of epidemic growth \citep{dushoffSpeedStrengthEpidemic2021}.
+The dynamics of infectious processes are often described by the complementary concepts of the reproduction number $R$ \citep{gosticPracticalConsiderationsMeasuring2020} and growth rate $r$ \citep{wallingaHowGenerationIntervals2007}, where $R$ describes the strength and $r$ the speed of epidemic growth \citep{dushoffSpeedStrengthEpidemic2021}.
 In the absence of changes in immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
@@
 This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellisChallengesControlCOVID192021}.
-If case numbers are growing exponentially and the true modelling task revolves around estimating the reproduction number or the corresponding growth rate, evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount, as errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
-If one is to assess the ability of forecasts to assess and forecast the underlying infection dynamics, it may be more appropriate to evaluate errors on the growth rate directly.
+If case numbers are evolving based on an exponential process and the  modelling task revolves around estimating the reproduction number or the corresponding growth rate, evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount, as errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
+If one is to measure the ability of forecasts to assess and forecast the underlying infection dynamics, it may thus be more desirable to evaluate errors on the growth rate directly.
 
-Moreover, if forecasts are evaluated based on absolute rather than relative errors, whereas in reality errors scale with the level of observations, scores are strongly influenced by the overall order of magnitude of the quantity to forecast.
-There exist other scores that do not depend on the scale of the forecast target, such as the log score \citep{goodRationalDecisions1952}, which in addition is also invariant under monotonic transformations of the target, or a scaled version of the CRPS \citep{bolinLocalScaleInvariance2021}.
+Evaluating forecasts using the CRPS or WIS means that scores represent a measure of absolute errors.
+However, forecast consumers may find errors on a relative scale easier to interpret and more useful in order to track predictive performance across targets of different orders of magnitude.
+\cite{bolinLocalScaleInvariance2021} have proposed the scaled CRPS (SCRPS) which is locally scale invariant; however, it does not correspond to a relative error measure and lacks a straightforward interpretation as available for the CRPS.
+% is less commonly used and does not have a straightforward equivalent for a setting with predictive quantiles.
+A closely related aspect to relative scores (as opposed to absolute scores) is that in the evaluation one may wish to give similar weight to all considered forecast targets.
+As the CRPS typically scales with the order of magnitude of the quantity to be predicted, this is not the case for the CRPS, which will typically assign higher scores to forecast targets with high expected values (e.g., in large locations or around the peak of an epidemic).
+\cite{bracherEvaluatingEpidemicForecasts2021} have argued that this is a desirable feature, directing attention to situations of particular public health relevance.
+An evaluation based on absolute errors, however, will assign little weight to other potentially important aspects, such as the ability to correctly predict future upswings while observed numbers are still low.
+
+%that do not depend on the scale of the forecast target, such as the log score \citep{goodRationalDecisions1952}, which in addition is also invariant under monotonic transformations of the target, or a scaled version of the CRPS \citep{bolinLocalScaleInvariance2021}.
 These, however, may not have all of the desired properties of the CRPS and in particular cannot easily be transferred to a setting with predictive quantiles.
-With the regular CRPS and WIS the dependence on the scale of the forecast target makes it difficult to compare predictive performance across time and in different phases of an epidemic or to compare performance across targets that generally occur at different orders of magnitudes.
-In the example of COVID-19 where case numbers underwent drastic changes over time following the initial emergence, forecast evaluation based on the absolute numbers would be dominated by performance around the peaks.
-The ability to correctly predict future upswings while case numbers are still low would count little towards overall model scores.
-Scores based on jointly evaluating the ability to predict numbers of cases, hospitalisation and deaths would be dominated by cases as they occur at largest numbers.
+
+%In this paper we argue that these three aspects (evaluating the growth rate rather than absolute values, evaluating relative errors, giving more equal weights to all forecast targets) can all be addressed by simply applying the natural logarithm to forecasts (in our case predictive quantiles) and observations prior to computing the CRPS.
 
@@
 \cite{taylorEvaluatingVolatilityInterval1999} in finance, \cite{mayrLogLevelVAR2015} in macroeconomics, \cite{loweStochasticRainfallrunoffForecasting2014} in hydrology or \cite{fuglstadDoesNonstationarySpatial2015} in meteorology).
-The goal of the transformations is usually to improve accuracy of the predictions.
-In this paper, we discuss how transforming the forecasts and observations prior to applying the WIS (or CRPS) may be an appropriate way to improve the evaluation process by addressing the aforementioned issues.
+While the goal of the transformations is usually to improve accuracy of the predictions, they can also be used to enhance and complement the evaluation process.
+In this paper we argue that the aforementioned issues with evaluating forecasts based on measures of absolute error can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation.
+Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score.
 Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects of the forecasts that forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).
-For conceptual clarity and to allow for a more in-depth discussion, we focus most of our attention on the log-transformation as one example of many possible transformations, chosen for its simplicity and usefulness in an epidemiological context.
-We further focus on the CRPS/WIS as evaluation metric, given its widespread use throughout the COVID-19 pandemic.
 
-We first outline the motivation for transforming forecasts prior to scoring and specifically the motivation for applying a log-transformation (Section \ref{sec:methods}).
-We provide a mathematical intuition of the impact of the log-transformation, its interpretation (Section \ref{sec:methods:relative} and \ref{sec:methods:growthrate}) and effects on forecast rankings (section \ref{sec:methods:rankings} and provide some discuss some practical considerations relevant for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}).
+For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation in the context of epidemic phenomena.
+Instead of a score representing the magnitude of absolute errors, applying a log-transformation prior to the CRPS yields a score which a) measures relative error (see Section \ref{sec:methods:relative}), b) provides a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate}) and c) is less dependent on the expected order of magnitude of the quantity to be predicted (see Section \ref{sec:methods:vst}).
+%The interpretation of the log-transformed score as a score for a forecast of the exponential growth rate of a given quantity aligns the evaluation with our understanding of how epidemiological processes should be modeled and the kinds of errors models would typically make.
+We therefore argue that such evaluations on the logarithmic scale should complement the prevailing evaluations on the natural scale. 
+Other transformations may likewise be of interest, and we briefly explore the square root transformation as an alternative transformation. %[and population standardization?]. 
+Our analysis focuses on the CRPS (or WIS) as evaluation metric, given its widespread use throughout the COVID-19 pandemic.
+
+The remainder of the article is structured as follows.
+In Sections \ref{sec:methods:relative}--\ref{sec:methods:vst} we provide some mathematical intuition on applying the log transformation prior to evaluating the CRPS, highlighting the connections to relative error measures, the epidmemic growth rate and variance stabilizing transformations.
+%We first provide a mathematical intuition of the impact of the log-transformation and its interpretation as relative error (Section \ref{sec:methods:relative}, measure of the accuracy of a forecast of the growth rate (Section \ref{sec:methods:growthrate}) and it's effect as a variance-stabilising transformation that makes can make scores independent of the mean of the forecast target (Section \ref{sec:methods:vst}.
+We then discuss the effect of the transformation on forecast rankings (section \ref{sec:methods:rankings} and discuss some practical considerations relevant for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}).
 We illustrate the effect of the log-transformation using forecasts submitted to the European COVID-19 Forecast Hub  \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022} (Section \ref{sec:results}).
@@
 \label{sec:methods}
-
-Our aim in transforming forecasts and observations is to shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score.
-A particularly relevant transformation in the context of epidemic phenomena is the natural logarithm.
-Instead of a score representing the magnitude of absolute errors this approach makes it possible to obtain a score which a) measures relative error and at the same time b) provides an a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate} and is c) independent from the mean of the observed quantity (see Section \ref{sec:methods:vst}).
-The interpretation of the log-transformed score as a score for a forecast of the exponential growth rate of a given quantity aligns the evaluation with our understanding of how epidemiological processes should be modeled and the kinds of errors models would typically make.
 
@@
 As Figure \ref{fig:SAPE} shows, the alignment with the SAPE is in fact the closest and holds quite well even if predicted and observed value differ by a factor of two or three.
-Moving back to probabilistic forecasts, the CRPS or WIS applied to log-transformed forecasts and outcomes can thus be seen as a probabilistic counterpart to the symmetric absolute percentage error, which offers an appealing intuitive interpretation.
+Moving back to probabilistic forecasts, the CRPS applied to log-transformed forecasts and outcomes can thus be seen as a probabilistic counterpart to the symmetric absolute percentage error, which offers an appealing intuitive interpretation.
 
@@
 \includegraphics[width = 1\textwidth]{output/figures/different-relative-errors.png}
-\caption{Numerical comparison of different measures of relative error/ We consider a prediction of $\hat{y} = 1$ and for different observed values show the  absolute percentage error (APE), relative error (RE), symmetric absolute percentage error (SAPE) and the absolute error applied to log-transformed predictions and observations.
-$\hat{y}$ denotes the predicted value.
-A: Error measures as a function of the observed value with the x-axis shown on a linear scale.
-B: Same, but x-axis shown on a logarithmic scale.}
+\caption{Numerical comparison of different measures of relative error: absolute percentage error (APE), relative error (RE), symmetric absolute percentage error (SAPE) and the absolute error applied to log-transformed predictions and observations.
+We denote the predicted value by $\hat{y}$ and display errors as a function of the ratio of observed and predicted value.
+A: x-axis shown on a linear scale.
+B: x-axis shown on a logarithmic scale.}
 \label{fig:SAPE}
@@
 
-Another interpretation can be based around viewing the process to be forecast as exponential with time-varying growth rate $r(t)$~\citep[see, e.g.,][]{wallingaHowGenerationIntervals2007}, i.e.
+Another interpretation can be based on viewing the process to be forecast as exponential with time-varying growth rate $r(t)$~\citep[see, e.g.,][]{wallingaHowGenerationIntervals2007}, i.e.
 \begin{linenomath*}
@@
 For small $\bar{r}t$, $ y_0 \exp (\bar{r}t)$ is approximately equivalent to $ y_0 (1 + \bar{r}t)$, which is sometimes used as an alternative definition of the growth rate (then called the multiplicative growth rate).
+This means that by applying the CRPS rather than the absolute error we perform a probabilistic evaluation of forecasts of the epidemic growth rate.
 
@@
 
-In epidemiological contexts, the variance of the variable of interest usually scales with its mean.
-A variance-stabilising transformation \citep[VST][]{bartlettSquareRootTransformation1936} is a transformation that serves to make the variance of a variable independent of its mean.
-
-If the predictive distribution for the quantity $Y$ equals the true data-generating process $F$ (an ideal forecast), the expected CRPS is given by \citep{gneitingStrictlyProperScoring2007}
+When evaluating models across sets of forecasting tasks, it may be considered desirable for each target to have a similar impact on the overall results.
+In disease forecasting this is not the case when using the CRPS, as the latter typically scales with the order of magnitude of the quantity to be predicted.
+Average scores are then dominated by the results achieved for targets with high expected outcomes.
+%Bracher et al. mention this as a desirable feature of the CRPS. 
+% In epidemiological contexts, the variance of the variable of interest usually scales with its mean. 
+Specifically, if the predictive distribution for the quantity $Y$ equals the true data-generating process $F$ (an ideal forecast), the expected CRPS is given by \citep{gneitingStrictlyProperScoring2007}
 $$
@@
 $$
-can be used, meaning that the expected CRPS scales roughly with the standard deviation. Denoting by $F^*$ the predictive distribution for $\log(Y)$, the delta method implies that
+can be used.
+This means that the expected CRPS scales roughly with the standard deviation, which in turn typically increases with the mean in epidemiological forecasting.
+In order to make the expected CRPS independent of the expected outcome, a \textit{variance-stabilising transformation} \citep[VST,][]{bartlettSquareRootTransformation1936} can be employed.
+The choice of this transformation depends on the mean-variance relationship of the underlying process.
+
+If the mean-variance relationship is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST \citep{guerreroTimeseriesAnalysisSupported1993}.
+Denoting by $F^*$ the predictive distribution for $\log(Y)$, we can use the delta method to show that
 $$
-\mathbb{E}_F[\text{CRPS}\{F^*, \log(y)\}] \approx \frac{\sigma/\mu}{\sqrt{\pi}}.
+\mathbb{E}_F[\text{CRPS}\{F^*, \log(y)\}] \approx \frac{\sigma/\mu}{\sqrt{\pi}} 
+= \frac{\sqrt{c}}{\sqrt{\pi}}
+.
 $$
-The expected CRPS for the log-transformed outcome thus scales with the coefficient of variation.
-In instances where the variance scales with the square of the mean, the log-transformation serves as a variance-stabilising transformation \citep{guerreroTimeseriesAnalysisSupported1993}, making the expected CRPS independent of the mean of the forecast quantity.
-We illustrate this in Figure \ref{fig:SIM-wis-state-size-mean}, where we show CRPS scores for different ideal forecasters.
-For a negative-binomial distribution with $\sigma^2 = \mu + \mu^2$ and sufficiently large $\mu$ we see that on the log scale scores are independent of the mean of the forecast target.
+The assumption of a quadratic mean-variance relationship is closely linked to the aspects discussed in Sections \ref{sec:methods:relative} and \ref{sec:methods:growthrate}.
+It implies that relative errors have constant variance and can thus be meaningfully compared across different targets.
+Also, it arises naturally if we assume that our capacity to predict the epidemic growth rate does not depend on the expected outcome.
+
+If the variance is linear with $\sigma^2 = c \times \mu$, as with a Poisson-distributed variable, the square root is known to be a VST.
+Denoting by $\tilde{F}$ the predictive distribution for $\sqrt{Y}$, the delta method can again be used to show that
+$$
+\mathbb{E}_F[\text{CRPS}\{\tilde{F}, \sqrt{y}\}] \approx \frac{\sigma/\sqrt{\mu}}{2\sqrt{\pi}} = \frac{\sqrt{c}}{2\sqrt{\pi}}
+.
+$$
+
+% which one is an empirical question - maybe hint that quadratic may result from RW on growth rate?
+
+To strengthen our intuition on how transforming outcomes prior to applying the CRPS shifts the emphasis between targets with high and low expected outcomes, Figure \ref{fig:SIM-wis-state-size-mean} shows the expected CRPS of ideal forecasters under different mean-variance relationships and transformations.
+We consider a Poisson distribution where $\sigma^2 = \mu$, a geometric distribution where $\sigma^2 = \mu + \mu^2$, and a negative binomial distribution where the size parameter $\psi(\mu)$ depends on the mean $\mu$ such that $\sigma^2 = \mu + \mu^{2.5}/1000$.
+We see that when applying the CRPS on the natural scale, the expected CRPS grows in $\mu$ in all cases.
+When applying it after a log-transformation, the expected CRPS is almost independent of $\mu$ for the geometric distribution and large $\mu$, while smaller targets have higher expected CRPS in the Poisson case.
+When applying a square-root-transformation before the CRPS, the expected CRPS is independent of the mean for the Poisson-distribution and grows for the two other distributions.
+%For the negative binomial distribution, targets with very small and very high expectation yield the highest expected CRPS.
+\textbf{Here it would be nice to also add a sentence on what happens under sqrt transformation.} % scores are independent of the mean of the forecast target.
 For a Poisson-distribution with
-$\sigma = \mu$, scores on the log scale decrease as the mean increases.
-The appropriate variance-stabilising transformation here to make scores independent of the mean would be a square-root-transformation \citep{bartlettSquareRootTransformation1936}.
+% $\sigma = \mu$, scores on the log scale decrease as the mean increases.
+The appropriate variance-stabilising transformation here to make scores independent of the mean would be a square-root-transformation .
 
@@
 \label{sec:methods:rankings}
-Rankings between different forecasters based on their scores may change through the log-transformation both when looking at aggregate scores as well as when looking at a single forecast.
+Rankings between different forecasters based on the CRPS may change when making use of a transformation both when looking at aggregate scores as well as when looking at a single forecast.
 We illustrate this in Figure \ref{fig:illustration-ranking} with two forecasters, A and B, issuing two different distributions with different dispersion.
-When showing the obtained score as a function of the observed value, it can be seen that the ranking between the two forecasters may change when scoring the forecast on the logarithmic, rather than the natural scale.
+When showing the obtained CRPS as a function of the observed value, it can be seen that the ranking between the two forecasters may change when scoring the forecast on the logarithmic, rather than the natural scale.
 In particular, on the natural scale, forecaster A, who issues a more uncertain distribution, receives a better score than forecaster B for observed values far away from the centre of the respective predictive distribution.
@@
 The change in rankings of aggregate scores is not mainly driven by the kinds of changes in model rankings for single forecasts just discussed.
-Rather, as observations from exponential processes are likely to be on very different absolute scales, large observations will account for the majority of absolute score but will not dominate relative scores in the same way.
-Evaluating forecasts based on relative errors may help with comparing performance across different targets.
-However, log-transforming forecasts and observations may also result in the opposite extreme because small absolute errors (e.g.
-predicting 9, rather than 3 deaths) may be large in relative terms if the quantity to forecast is very small.
-This can be exacerbated when forecasts are restricted to discrete values, where relative steps between possible values become greater for values closer to zero.
+Rather, average scores will be most influenced by the order of magnitude of scores for different forecast targets across time, location and target type.
+On the natural scale, large observations will account for the majority of the overall score but will not dominate average CRPS values obtained after log transformation in the same way.
+Depending on the relationship between mean and variance of the forecast target, a log-transformation, however, may also lead to systematically larger scores assigned to small forecast targets, as illustrated in Figure \ref{fig:SIM-wis-state-size-mean}.
+A different transformation like the square-root may then be appropriate to stabilise the variance effectively and make scores independent of the mean of the target quantity.
 
@@
 
-Whether forecasts of small or large targets receive higher scores than those for targets with larger quantities, depends on the relationship between the mean and the variability of the quantity of interest (see Section \ref{sec:methods:vst}.
+%Whether forecasts of small or large targets receive higher scores than those for targets with larger quantities, depends on the relationship between the mean and the variability of the quantity of interest (see Section \ref{sec:methods:vst}.
 For the simulated count data, log predictions for small quantities received on average higher scores than forecasts for large quantities if the mean and the variance grew at the same rate $\sigma^2 = \mu$, about equal scores when the variance grew at a rate of $\sigma^2 = \mu + \mu^2$ (which for large $\mu$ is dominated by the quadratic term), and smaller scores when the variance grew faster than that (Figure \ref{fig:SIM-wis-state-size-mean}).
@@
 
-The log-transformation (among many other transformations) is permissible (in the sense that it preserves propriety of the score), because it is a strictly monotonic transformation.
-Even though actual rankings between models may change, forecasters will in expectation still minimise their score if they report a predictive distribution that is equal to the data-generating distribution.
-The order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee propriety.
-In the specific example, taking the logarithm of the scores, rather than scoring the log-transformed forecasts and data, results in an improper score.
+Transformations that are strictly monotonic are permissible in the sense that they maintain the propriety of the score.
+This is because even though rankings of models may change forecasts will in expectation still minimise their score if they report a predictive distribution that is equal to the data-generating distribution.
+This condition holds for both the log and square root transformations, as well as many others.
+However, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee propriety.
+In the case of log transforms, taking the logarithm of the scores, rather than scoring the log-transformed forecasts and data, results in an improper score.
 This is because taking the logarithm of the WIS / CRPS results in a score that does not penalise outliers enough and therefore incentivises overconfident predictions.
@@
 For the log-transformed CRPS, the optimal value is  0.9, i.e.
-there is an incentive to report a too sharp forecast.}
+there is an incentive to report a forecast that is too sharp.}
     \label{fig:log-improper}
@@
 
-In practice, one issue with the log-transformation is that it cannot readily be applied to negative numbers or zero values.
-In addition, it may be difficult to reliably evaluate relative errors when dealing with small discrete values.
-In an epidemiological setting that involves count data, negative values should likely be omitted entirely.
-One solution to deal with zeroes and small values is to also exclude those from ongoing analysis.
-An alternative is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm as is common practice when log-transforming counts.
-This represents a strictly monotonic transformation and therefore preserves propriety of the resulting score.
+In practice, one issue with transformations, such as the log transform, is that they may not be readily applicable to negative numbers or zero values, which need to be removed or handled otherwise.
+Small discrete values can pose an additional problem as it may be difficult to reliably evaluate relative errors.
+One common approach is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm.
+This represents a strictly monotonic transformation and therefore still preserves the propriety of the resulting score.
 The choice of the number to add influences scores and rankings, as measures of relative errors shrink when adding a constant to the forecast and the observation.
-Adding something to the forecasts and observations changes scores for forecasts of small quantities most, as relative errors shrink more the smaller the original value before adding a constant.
+Adding a number to the forecasts and observations changes scores for forecasts of small quantities most, as relative errors shrink more the smaller the original value before adding a constant.
 In principle, it would also be possible to exploit the difference in how scores are affected in order to equalise scores for forecasts of quantities across different scales by adding a larger number (e.g.
@@
 
-As an empirical example for evaluating forecasts on the natural and on the log scale we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
-The European COVID-19 Forecast Hub is one of several COVID-19 Forecast hubs (together with a Hub in the US \citep{cramerEvaluationIndividualEnsemble2021} and one in Germany and Poland \citep{bracherShorttermForecastingCOVID192021}) which have been systematically collecting, aggregating and evaluating forecasts of several COVID-19 targets created by different teams every week.
+As an empirical comparison of evaluating forecasts on the natural and on the log scale, we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
+The European COVID-19 Forecast Hub is one of several COVID-19 Forecast hubs\citep{cramerEvaluationIndividualEnsemble2021, bracherShorttermForecastingCOVID192021} which have been systematically collecting, aggregating and evaluating forecasts of several COVID-19 targets created by different teams every week.
 Forecasts are made one to four weeks ahead into the future and follow a quantile-based format with a set of 22 quantiles plus the median ($0.01, 0.025, 0.05, ..., 0.5, ...
@@
 
-The forecasts used for the purpose of this illustration are forecasts submitted between March 8 2021 and October 18 2021 for reported cases and deaths from COVID-19.
-We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 16 forecast dates (see Figure \ref{fig:HUB-num-avail-models}).
+The forecasts used for the purpose of this illustration are forecasts submitted between March 8 2021 and October 18 2021 CHECK AGAIN for reported cases and deaths from COVID-19.
+We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 16 CHECK AGAIN forecast dates (see Figure \ref{fig:HUB-num-avail-models}).
 We removed all observations marked as data anomalies by the European Forecast Hub as well as all remaining negative observed values.
+In addition, we filtered out erroneous forecasts characterised by a median forecast of 0 and an actual observed value of larger than 150.
 All predictive quantiles were truncated at 0.
@@
 
-As expected on the natural scale, scores correlate strongly with the average number of observed cases and deaths (Spearman rank correlation: 0.96 and 0.97) in a given country (see Figure\ref{fig:HUB-mean-locations}A, B).
-We report the Spearman rank correlation here, rather than the Pearson correlation, so that correlations remain the same regardless of whether we look at observed values on the log or natural scale.
-On the log scale, the correlation is much weaker and slightly negative, at least in terms of rank correlation (Spearman correlation: CHECK AGAIN -0.33 and -0.36), implying a tendency towards larger scores in countries with lower mean incidences.
+As expected on the natural scale, scores correlate strongly with the magnitude of the median forecast (Spearman rank correlation: 0.9 and 0.92) (see Figure\ref{fig:HUB-mean-locations}A, B).
+We report the Spearman rank correlation here, rather than the Pearson correlation, as it tends to be more robust and is invariant to transformations.
+For scores on the log scale, the correlation is weaker and slightly negative (Spearman correlation: -0.35 and -0.55), implying a tendency towards larger scores for forecasts of smaller quantities.
 
@@
 \begin{equation*}
-    \text{CRPS}(\log F, \log y) = \alpha^* + \beta^* \cdot \log{(\text{median}(F))}. 
+    \text{CRPS}(\log F, \log y) = \alpha^* + \beta^* \cdot \log{(\text{median}(F))}
+\end{equation*} 
+and the following regression on scores obtained after applying a square-root-transformation
+
+\begin{equation*}
+    \text{CRPS}(\sqrt{F}, \sqrt{y}) = \alpha^{**} + \beta^{**} \cdot \sqrt{(\text{median}(F))}
 \end{equation*}
@@
 We found that on the natural scale, the order of magnitude of the forecast (represented by the median, rather than the mean, as forecasts were stored as quantiles and the mean was not available) strongly influenced scores, with estimated $\beta$-values close to one.
-On the natural, this influence was small, with estimated $\beta^*$-values close to zero.
+On the log scale, this influence was small, with estimated $\beta^*$-values close to zero.
 Both on the natural and on the log scale we found scores to increase over time (estimated $\alpha$- and $\alpha^*$-values increased over horizons).
@@
 
-In this paper we proposed a log-transformation of forecasts and observed values prior to evaluation with the WIS (or CRPS) in order to address issues that arise when evaluating epidemiological forecasts based on a measure of absolute error.
-We gave a mathematical intuition for the transformation and showed that scores on the log scale can be interpreted as a) a measure of relative prediction errors, as well as b) a score for a forecast of the exponential growth rate of the target quantity.
+In this paper we proposed the use of transformations, with a particular focus on the log transformation, in order to address issues that arise when evaluating epidemiological forecasts based on measures of absolute error and their probabilistic generalisations (i.e CRPS and WIS).
+We gave a mathematical intuition for the log transformation and showed that scores on the log scale can be interpreted as a) a measure of relative prediction errors, as well as b) a score for a forecast of the exponential growth rate of the target quantity and c) as variance stabilising transformation in some settings.
 When applying this approach to forecasts from the European COVID-19 Forecast Hub, we found overall scores on the log scale to be more equal across, time, location and target type (cases, deaths) than scores on the natural scale.
 Scores on the log scale were much less influenced by the overall incidence level in a country, and showed a slight tendency to be higher in locations with very low incidences.
-We found that model rankings changed noticeably, favouring models like epiforecasts-EpiNow2 which explicitly has an exponential error structure, and disfavouring models like MUNI-ARIMA which is a purely statistical model with a linear error structure.
-On the natural scale, missing the peak and overshooting was more severely penalised than missing the nadir and a following upswing in numbers.
+We found that model rankings changed noticeably, favouring models like epiforecasts-EpiNow2 which uses a log scale generative process and a negative binomial observation model, and disfavouring models like MUNI-ARIMA which uses a square root transform to stabilise variance.
+On the natural scale, missing the peak and overshooting was more severely penalised than missing the nadir and the following upswing in numbers.
 Both failure modes tended to be more equally penalised on the log scale (with undershooting receiving slightly higher penalties in our example).
@@
 A second transformation was mentioned above, namely converting forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last observed value.
-One could also think about applying a Box-Cox or a square-root transformation in order to stabilise the variance of the forecasts.
+One could also think about applying a Box-Cox \citep{boxAnalysisTransformations1964} or a square-root transformation in order to stabilise the variance of the forecasts.
 Another promising transformation would be to divide each forecast by the forecast of the previous week (and observations by the observation of the previous week), in order to obtain forecasts for week-to-week multiplicative growth rates, or alternatively, to ask forecasters to provide estimates of the weekly relative change applied to the latest data and subsequent forecast points directly.
-This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluation the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}.
-This, unfortunately is not well defined in the context of forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate, but may be an interesting approach if predictive samples are available.
+This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluate the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}.
+This, unfortunately is not feasible in the context of forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate, but may be an interesting approach if predictive samples are available.
 What makes this approach of transforming forecasts before scoring particularly interesting is the possibility to construct composite scores as a weighted sum of scores based on different transformations.
@@
 
-Exploring these different transformations is a promising avenue for future work that could help bridge the gap between modellers and policy makers by providing scoring rules that better reflect what forecast consumers actually care about.
-We have showed that the log-transformation can lead to significant changes in the relative rankings of models against each other, with potentially important implications for decision makers who rely on the knowledge of past performance to make a decision about which forecasts should inform future decisions.
+Exploring transformations is a promising avenue for future work that could help bridge the gap between modellers and policy makers by providing scoring rules that better reflect what forecast consumers care about.
+Potentially, the variance stabilising time-series forecasting literature may be a useful source of transformations for various forecast settings.
+We have showed that the log-transformation can lead to significant changes in the relative rankings of models against each other, with potentially important implications for decision-makers who rely on the knowledge of past performance to make a decision about which forecasts should inform future decisions.
 While it is commonly accepted that multiple proper scoring rules should usually be considered when comparing forecasts, we think this should be supplemented by considering different transformations of the data to obtain a richer picture of model performance.
-While we have produced this work as an initial case study, more work needs to be done to better understand the effects of applying a logarithmic or other transformations more comprehensively and in different contexts, and how it might affect decision making.
+More work needs to be done to better understand the effects of applying transformations more comprehensively and in different contexts, and how it might affect decision-making.
 
@@
 
+\subsection{Alternative Formulation of the WIS}
+\label{sec:alternative-wis}
+
+Instead of defining the WIS as an average of scores for individual quantiles, we can define it using an average of scores for symmetric predictive intervals.
+For a single prediction interval, the interval scoren (IS) is computed as the sum of three penalty components, dispersion (width of the prediction interval), underprediction and overprediction,
+%
+\begin{linenomath*}
+\begin{align*}
+ IS_\alpha(F,y) &= (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u) \\
+ &= \text{dispersion} + \text{underprediction} + \text{overprediction},    
+\end{align*}
+\end{linenomath*}
+%
+where $1()$ is the indicator function, $y$ is the observed value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution, i.e.
+the lower and upper bound of a single central prediction interval.
+For a set of $K^*$ prediction intervals and the median $m$, the WIS is computed as a weighted sum,
+\begin{linenomath*}
+\begin{equation*}
+\text{WIS} = \frac{1}{K^* + 0.5} \cdot \left(w_0 \cdot |y - m| + \sum_{k = 1}^{K^*} w_k \cdot IS_{\alpha_{k}}(F, y)\right),    
+\end{equation*} 
+\end{linenomath*}
+where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$.
+
 \begin{table}[h!]
@@
 \toprule
-Target & Quantity & Measure & natural & log\\
+target\_type & quantity & measure & natural & log\\
 \midrule
-Cases & Observations & mean & 20013 & 8.33\\
-Cases & Observations & sd & 43256 & 2.07\\
-Cases & Observations & var & 1871039511 & 4.30\\
+Cases & Observations & mean & 62004 & 9.19\\
+Cases & Observations & sd & 171723 & 2.10\\
+Cases & Observations & var & 29488651564 & 4.42\\
 \addlinespace
-Deaths & Observations & mean & 244 & 4.01\\
-Deaths & Observations & sd & 498 & 1.81\\
-Deaths & Observations & var & 248306 & 3.28\\
+Deaths & Observations & mean & 223 & 3.90\\
+Deaths & Observations & sd & 441 & 1.96\\
+Deaths & Observations & var & 194221 & 3.85\\
 \addlinespace
@@
 \addlinespace
-Cases & WIS & mean & 4220 & 0.26\\
-Cases & WIS & sd & 13978 & 0.35\\
+Cases & WIS & mean & 15840 & 0.27\\
+Cases & WIS & sd & 53117 & 0.28\\
 \addlinespace
-Deaths & WIS & mean & 30 & 0.21\\
-Deaths & WIS & sd & 70 & 0.19\\
+Deaths & WIS & mean & 31 & 0.23\\
+Deaths & WIS & sd & 65 & 0.28\\
 \bottomrule
@@
 
-\subsection{Alternative Formulation of the WIS}
-
-Instead of defining the WIS as an average of scores for individual quantiles, we can define it using an average of scores for symmetric predictive intervals.
-For a single prediction interval, the interval scoren (IS) is computed as the sum of three penalty components, dispersion (width of the prediction interval), underprediction and overprediction,
-%
-\begin{linenomath*}
-\begin{align*}
- IS_\alpha(F,y) &= (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u) \\
- &= \text{dispersion} + \text{underprediction} + \text{overprediction},    
-\end{align*}
-\end{linenomath*}
-%
-where $1()$ is the indicator function, $y$ is the observed value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution, i.e.
-the lower and upper bound of a single central prediction interval.
-For a set of $K^*$ prediction intervals and the median $m$, the WIS is computed as a weighted sum,
-\begin{linenomath*}
-\begin{equation*}
-\text{WIS} = \frac{1}{K^* + 0.5} \cdot \left(w_0 \cdot |y - m| + \sum_{k = 1}^{K^*} w_k \cdot IS_{\alpha_{k}}(F, y)\right),    
-\end{equation*} 
-\end{linenomath*}
-where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/SIM-score-approximation.png}
+    \caption{
+    Visualisation of expected CRPS values against approximated scores using the approximation detailed in Section \ref{sec:methods:rankings} (see also Figure \ref{fig:SIM-wis-state-size-mean}).
+Expected CRPS scores are shown for three different distributions once on the natural scale (top row) and once scored on the log scale (bottom row).
+    }
+    \label{fig:HUB-model-comparison-epinow}
+\end{figure}
 
@@
 \centering
-\begin{tabular}{ccrrrr}
+\begin{tabular}{ccrrrrrr}
 \toprule
-Horizon & Target & $\alpha$ & $\beta$ & $\alpha^*$ & $\beta^*$\\
+Horizon & Target & $\alpha$ & $\beta$ & $\alpha^*$ & $\beta^*$ & $\alpha^{**}$ & $\beta^{**}$\\
 \midrule
-all & all & -0.556 & 0.877 & 0.579 & -0.028\\
+all & all & -1.093 & 0.963 & 0.390 & 0.001 & -0.995 & 0.209\\
 \addlinespace
-all & Cases & 1.112 & 0.710 & 1.280 & -0.101\\
-all & Deaths & -0.433 & 0.762 & 0.523 & -0.044\\
+all & Cases & 0.036 & 0.858 & 0.751 & -0.033 & 0.047 & 0.201\\
+all & Deaths & -0.885 & 0.868 & 0.430 & -0.021 & -7.564 & 0.826\\
 \addlinespace
-1 & all & -0.999 & 0.855 & 0.419 & -0.031\\
-2 & all & -0.756 & 0.890 & 0.517 & -0.028\\
-3 & all & -0.430 & 0.895 & 0.628 & -0.026\\
-4 & all & -0.047 & 0.881 & 0.747 & -0.024\\
+1 & all & -1.402 & 0.923 & 0.318 & -0.014 & 0.318 & 0.088\\
+2 & all & -1.221 & 0.967 & 0.364 & -0.003 & 0.111 & 0.164\\
+3 & all & -1.001 & 0.984 & 0.408 & 0.008 & -1.283 & 0.255\\
+4 & all & -0.757 & 0.986 & 0.467 & 0.016 & -1.041 & 0.312\\
 \addlinespace
-1 & Cases & -0.065 & 0.764 & 0.768 & -0.067\\
-2 & Cases & 0.676 & 0.748 & 1.104 & -0.089\\
-3 & Cases & 1.414 & 0.710 & 1.432 & -0.109\\
-4 & Cases & 2.173 & 0.658 & 1.712 & -0.124\\
+1 & Cases & -0.861 & 0.876 & 0.433 & -0.024 & 0.785 & 0.087\\
+2 & Cases & -0.243 & 0.877 & 0.660 & -0.031 & 0.959 & 0.162\\
+3 & Cases & 0.372 & 0.855 & 0.882 & -0.037 & 1.115 & 0.238\\
+4 & Cases & 0.816 & 0.837 & 1.009 & -0.036 & 1.665 & 0.296\\
 \addlinespace
-1 & Deaths & -0.726 & 0.735 & 0.429 & -0.048\\
-2 & Deaths & -0.553 & 0.766 & 0.490 & -0.045\\
-3 & Deaths & -0.352 & 0.782 & 0.552 & -0.041\\
-4 & Deaths & -0.113 & 0.781 & 0.618 & -0.036\\
+1 & Deaths & -1.145 & 0.831 & 0.376 & -0.035 & 0.458 & 0.048\\
+2 & Deaths & -0.980 & 0.866 & 0.415 & -0.027 & 0.445 & 0.084\\
+3 & Deaths & -0.809 & 0.885 & 0.441 & -0.016 & -8.262 & 0.903\\
+4 & Deaths & -0.603 & 0.891 & 0.488 & -0.007 & -7.834 & 0.898\\
 \bottomrule
 \end{tabular}
-\caption{Coefficients of two regressions for the effect of the magnitude of the median forecast on expected scores. The first regression was 
+\caption{Coefficients of three regressions for the effect of the magnitude of the median forecast on expected scores. The first regression was 
 $\text{CRPS}(F, y) = \exp(\alpha) \cdot \text{median}(F)^{\beta},$ where $F$ is the predictive distribution and $y$ the observed value. The second one was 
-$\text{CRPS}(\log F, \log y) = \alpha^* + \beta^* \cdot \log{(\text{median}(F))}.$
+$\text{CRPS}(\log F, \log y) = \alpha^* + \beta^* \cdot \log{(\text{median}(F))}.$ The third one was $\text{CRPS}(\sqrt{F}, \sqrt{y}) = \alpha^{**} + \beta^{**} \cdot \sqrt{(\text{median}(F))}.$
 }

### 2023-01-09 main_backup.tex
update figures, fix code issues

--- 2023-01-09 main.tex
@@
 While the goal of the transformations is usually to improve accuracy of the predictions, they can also be used to enhance and complement the evaluation process.
-In this paper we argue that the aforementioned issues with evaluating forecasts based on measures of absolute error can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation.
-Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score.
+In this paper we argue that the aforementioned issues with evaluating forecasts based on measures of absolute error can be addressed by transforming the forecasts and observations prior to scoring.
+Transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score.
 Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects of the forecasts that forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).
@@
 Rather, average scores will be most influenced by the order of magnitude of scores for different forecast targets across time, location and target type.
-On the natural scale, large observations will account for the majority of the overall score but will not dominate average CRPS values obtained after log transformation in the same way.
+On the natural scale, large observations will account for the majority of the absolute score but will not dominate relative scores on the log scale in the same way.
 Depending on the relationship between mean and variance of the forecast target, a log-transformation, however, may also lead to systematically larger scores assigned to small forecast targets, as illustrated in Figure \ref{fig:SIM-wis-state-size-mean}.
@@
 
-%Whether forecasts of small or large targets receive higher scores than those for targets with larger quantities, depends on the relationship between the mean and the variability of the quantity of interest (see Section \ref{sec:methods:vst}.
+Whether forecasts of small or large targets receive higher scores than those for targets with larger quantities, depends on the relationship between the mean and the variability of the quantity of interest (see Section \ref{sec:methods:vst}.
 For the simulated count data, log predictions for small quantities received on average higher scores than forecasts for large quantities if the mean and the variance grew at the same rate $\sigma^2 = \mu$, about equal scores when the variance grew at a rate of $\sigma^2 = \mu + \mu^2$ (which for large $\mu$ is dominated by the quadratic term), and smaller scores when the variance grew faster than that (Figure \ref{fig:SIM-wis-state-size-mean}).
@@
 
-In practice, one issue with transformations, such as the log transform, is that they may not be readily applicable to negative numbers or zero values, which need to be removed or handled otherwise.
-Small discrete values can pose an additional problem as it may be difficult to reliably evaluate relative errors.
-One common approach is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm.
+In practice, one issue with transformations, such as the log transform, is that they may not be readily applicable to negative numbers or zero values.
+When using a log-transformation, it may also be difficult to reliably evaluate relative errors when dealing with small discrete values.
+In an epidemiological setting that involves count data, negative values should likely be omitted entirely.
+One potential solution to deal with zeroes and small values is to also exclude these from the ongoing analysis.
+An alternative is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm as is common practice when log-transforming counts.
 This represents a strictly monotonic transformation and therefore still preserves the propriety of the resulting score.
@@
 A second transformation was mentioned above, namely converting forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last observed value.
-One could also think about applying a Box-Cox \citep{boxAnalysisTransformations1964} or a square-root transformation in order to stabilise the variance of the forecasts.
+One could also think about applying a Box-Cox or a square-root transformation in order to stabilise the variance of the forecasts.
 Another promising transformation would be to divide each forecast by the forecast of the previous week (and observations by the observation of the previous week), in order to obtain forecasts for week-to-week multiplicative growth rates, or alternatively, to ask forecasters to provide estimates of the weekly relative change applied to the latest data and subsequent forecast points directly.
@@
 Exploring transformations is a promising avenue for future work that could help bridge the gap between modellers and policy makers by providing scoring rules that better reflect what forecast consumers care about.
-Potentially, the variance stabilising time-series forecasting literature may be a useful source of transformations for various forecast settings.
+Potentially the variance stabilising time-series forecasting literature may be a useful source of transformations for various forecast settings.
 We have showed that the log-transformation can lead to significant changes in the relative rankings of models against each other, with potentially important implications for decision-makers who rely on the knowledge of past performance to make a decision about which forecasts should inform future decisions.

### 2023-01-09 R/00 full paper code.R

@@ -174,13 +174,13 @@ plot_fct <- function(scores, scale_factor, nbinom, filter_scale = "natural") {
 }
 
 p1 <- scores |>
-  plot_fct(scale_factor = 240, nbinom = nbinom, filter_scale = "natural")
+  plot_fct(scale_factor = 240, nbinom = nbinom_natural, filter_scale = "natural")
 
 p2 <- scores |>
-  plot_fct(scale_factor = 13, nbinom = nbinom, filter_scale = "log") 
+  plot_fct(scale_factor = 13, nbinom = nbinom_natural, filter_scale = "log") 
 
 p3 <- scores |>
-  plot_fct(scale_factor = 23, nbinom = nbinom, filter_scale = "sqrt") 
+  plot_fct(scale_factor = 23, nbinom = nbinom_natural, filter_scale = "sqrt") 
 
 p1 + p2 +
   plot_layout(guides = "collect") &
@@ -838,17 +838,20 @@ ranking_figure <- function(target = "Cases") {
     facet_wrap(~scale, scales = "free_x")
   
   plot_ranking_change <- summarised_pairwise |>
+    filter(scale %in% c("natural", "log")) |>
     select(model, scale, relative_skill) |>
     group_by(scale) |>
+    # mutate(scale = factor(scale, levels = c("log", "natural"))) |>
     mutate(rank = rank(-relative_skill), 
            x = ifelse(scale == "natural", 0.4, 2.8), 
            x_arrow = ifelse(scale == "natural", 1.7, 2.7)
     ) |>
     group_by(model) |>
-    mutate(color = ifelse(diff(relative_skill) > 0, "deteriorated", "not deteriorated")) |>
-    ggplot(aes(y = rank, x = x, group = model, label = model)) +
+    mutate(custom_color = ifelse(diff(relative_skill) < 0, "deteriorated", "not deteriorated")) |>
+    arrange(desc(scale)) |>
+    ggplot(aes(y = rank, group = model, label = model, color = custom_color)) +
     geom_path(
-      aes(x=x_arrow, color = color), 
+      aes(x=x_arrow), 
       arrow = arrow(length = unit(0.09,"npc")), 
       lineend = "round", linejoin = "mitre",
       size=1, show.legend = FALSE) +

### 2023-01-09 main.tex
update manuscript, remove a few passages

@@
   Edwin van Leeuwen\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom} $^{,}$\thanks{UK Health Security Agency, London, United Kingdom} $^{}$,
-  Johannes Bracher\thanks{Institute of Economic Theory and Statistics, Karlsruhe Institute of Technology, Karlsruhe, Germany} $^{}$, 
+  Johannes Bracher\thanks{Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology, Karlsruhe, Germany} $^{}$, 
   Sebastian Funk\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$}
@@
 \begin{abstract}
-Forecast evaluation plays an essential role in improving forecasting models can help inform future decision making.
+Forecast evaluation plays an essential role in improving forecasting models and can help inform future decision making.
 In epidemiology, probabilistic forecasts are often scored using the Weighted Interval Score (WIS) or the Continuous Ranked Probability Score (CRPS), which are a measure of absolute distance between forecast and observation.
@@
 \end{abstract}
+
+% Suggestion J:
+% Forecast evaluation plays an essential role in the development cycle of pedictive epidemic models and can inform their use for public health decision making.
+Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which are both measures of the absolute distance between the forecast distribution and the observation.
+They are commonly applied directly to predicted and observed incidence counts, but it can be questioned whether this yields the most meaningful results given the exponential nature of epidemic processes and the fact that observed values over space and time can span several orders of magnitude.
+In this paper we argue that applying the CRPS to incidence counts on a logarithmic rather than natural scale effectively mitigates these difficulties, yielding epidemiologically meaningful and easily interpretable results.
+We motivate the procedure threefold: firstly, it can be interpreted as a probabilistic version of a relative error.
+Secondly, it reflects how well models predicted the time-varying epidemic growth rate.
+And lastly, using arguments on variance-stabilizing transformations, it can be shown that under the assumption of a quadratic mean-variance relationship, the logarithmic transformation leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity.
+This implies that different forecasting tasks will will have a similar impact on overall evaluation results.
+Applying the log-transformation to data from the European COVID-19 Forecast Hub, we find that it changes model rankings, regardless of stratification by forecast date, location or target types.
+Situations in which models missed beginning upward swings are more strongly emphasized, while penalising missed peaks are less severely penalized.
+Also, we show that the assumption of a quadratic mean-variance relationship is quite closely aligned with empirical patterns.
 
@@
 Epidemiological modelling in particular has received widespread attention during the COVID-19 pandemic.
-Evaluations of forecasts can provide feedback for forecasters and researchers to improve their models and train ensembles, and help decision makers distinguish good from bad predictions and choosing forecasters and models that should inform future decisions.
+Evaluations of forecasts can provide feedback for forecasters and researchers to improve their models and train ensembles.
+It moreover helps decision makers distinguish good from bad predictions and choose forecasters and models that should inform future decisions.
 
@@
 where $y$ is the true observed value and $F$ the cumulative distribution function (CDF) of the predictive distribution.
-The CRPS can be understood as a generalisation of the absolute error to predictive distributions.
+The CRPS can be understood as a generalisation of the absolute error to predictive distributions, and interpreted n the natural scale of the data.
 The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in COVID-19 forecast hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
@@
 The remainder of the article is structured as follows.
-In Sections \ref{sec:methods:relative}--\ref{sec:methods:vst} we provide some mathematical intuition on applying the log transformation prior to evaluating the CRPS, highlighting the connections to relative error measures, the epidmemic growth rate and variance stabilizing transformations.
+In Sections \ref{sec:methods:relative}--\ref{sec:methods:vst} we provide some mathematical intuition on applying the log transformation prior to evaluating the CRPS, highlighting the connections to relative error measures, the epidemic growth rate and variance stabilizing transformations.
 %We first provide a mathematical intuition of the impact of the log-transformation and its interpretation as relative error (Section \ref{sec:methods:relative}, measure of the accuracy of a forecast of the growth rate (Section \ref{sec:methods:growthrate}) and it's effect as a variance-stabilising transformation that makes can make scores independent of the mean of the forecast target (Section \ref{sec:methods:vst}.
@@
 \label{sec:methods:rankings}
-Rankings between different forecasters based on the CRPS may change when making use of a transformation both when looking at aggregate scores as well as when looking at a single forecast.
+Rankings between different forecasters based on the CRPS may change when making use of a transformation, both in terms of aggregate and individual scores.
 We illustrate this in Figure \ref{fig:illustration-ranking} with two forecasters, A and B, issuing two different distributions with different dispersion.
@@
 Rather, average scores will be most influenced by the order of magnitude of scores for different forecast targets across time, location and target type.
-On the natural scale, large observations will account for the majority of the overall score but will not dominate average CRPS values obtained after log transformation in the same way.
-Depending on the relationship between mean and variance of the forecast target, a log-transformation, however, may also lead to systematically larger scores assigned to small forecast targets, as illustrated in Figure \ref{fig:SIM-wis-state-size-mean}.
-A different transformation like the square-root may then be appropriate to stabilise the variance effectively and make scores independent of the mean of the target quantity.
+Large observations will dominate average CRPS values when evaluation is done on the natural scale, but much less so after log transformation.
+Depending on the relationship between mean and variance of the forecast target, a log-transformation may even lead to systematically larger scores assigned to small forecast targets, as illustrated in Figure \ref{fig:SIM-wis-state-size-mean}.
+A different transformation like the square-root may then be more appropriate to make scores independent of the mean of the target quantity.
 
@@
 
-In practice, one issue with transformations, such as the log transform, is that they may not be readily applicable to negative numbers or zero values, which need to be removed or handled otherwise.
+In practice, one issue with transformations, such as the log transform, is that they are not readily applicable to negative numbers or zero values, which need to be removed or handled otherwise.
 Small discrete values can pose an additional problem as it may be difficult to reliably evaluate relative errors.
@@
 As an empirical comparison of evaluating forecasts on the natural and on the log scale, we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
-The European COVID-19 Forecast Hub is one of several COVID-19 Forecast hubs\citep{cramerEvaluationIndividualEnsemble2021, bracherShorttermForecastingCOVID192021} which have been systematically collecting, aggregating and evaluating forecasts of several COVID-19 targets created by different teams every week.
+The European COVID-19 Forecast Hub is one of several COVID-19 Forecast hubs \citep{cramerEvaluationIndividualEnsemble2021, bracherShorttermForecastingCOVID192021} which have been systematically collecting, aggregating and evaluating forecasts of several COVID-19 targets created by different teams every week.
 Forecasts are made one to four weeks ahead into the future and follow a quantile-based format with a set of 22 quantiles plus the median ($0.01, 0.025, 0.05, ..., 0.5, ...
@@
 
-As expected on the natural scale, scores correlate strongly with the magnitude of the median forecast (Spearman rank correlation: 0.9 and 0.92) (see Figure\ref{fig:HUB-mean-locations}A, B).
+As expected, on the natural scale, scores correlate strongly with the magnitude of the median forecast (Spearman rank correlation: 0.9 and 0.92) (see Figure\ref{fig:HUB-mean-locations}A, B).
 We report the Spearman rank correlation here, rather than the Pearson correlation, as it tends to be more robust and is invariant to transformations.
@@
 
-The rank order of different forecasting models (see Figure \ref{fig:HUB-rank-order} indicates that models that do not make use of a exponential error model and thus would be biased towards under-prediction during periods of exponential growth, such as the MUNI-ARMA model, were ranked lower when using a log-transform.
+The rank order of different forecasting models changes noticeable (see Figure \ref{fig:HUB-rank-order}. 
+% I removed this part as it doesn't seem accurate anymore / we don't even have all the models in there anymore. 
+%indicates that models that do not make use of a exponential error model and thus would be biased towards under-prediction during periods of exponential growth, such as the MUNI-ARMA model, were ranked lower when using a log-transform.
 The epiforecasts-EpiNow2 forecast model, on the other hand~(created by co-authors of this study), which makes forecasts based on the behaviour of the reproduction number and thus implicitly uses an exponential error model, was the most impacted by the change to log-scale scoring moving from last place for reported cases to 4th place.
@@
 Scores on the log scale were much less influenced by the overall incidence level in a country, and showed a slight tendency to be higher in locations with very low incidences.
-We found that model rankings changed noticeably, favouring models like epiforecasts-EpiNow2 which uses a log scale generative process and a negative binomial observation model, and disfavouring models like MUNI-ARIMA which uses a square root transform to stabilise variance.
+We found that model rankings changed noticeably.
+%, favouring models like epiforecasts-EpiNow2 which uses a log scale generative process and a negative binomial observation model, and disfavouring models like MUNI-ARIMA which uses a square root transform to stabilise variance.
 On the natural scale, missing the peak and overshooting was more severely penalised than missing the nadir and the following upswing in numbers.

### 2023-01-09 R/00 full paper code.R
update Figure 8

@@ -410,6 +410,8 @@ hub_data$location |>
 ## ========================================================================== ##
 
 label_fn <- function(x) {
+  x <- ifelse(x%%1 == 0, 
+              as.integer(x), x)
   ifelse(x < 1000, 
          paste(x), 
          ifelse(x < 1e6, 
@@ -776,6 +778,12 @@ scores |>
 ## Figure 8 - Look at pairwise comparisons and changes in rankings (plus Appendix figures)
 ## ========================================================================== ##
 
+simple_labels <- function(x) {
+  ifelse(x%%1 == 0, 
+    paste(as.integer(x)),
+    paste(x))
+}
+
 ranking_figure <- function(target = "Cases") {
   summarised_pairwise <- scores |>
     select(-median_prediction) |>
@@ -795,7 +803,8 @@ ranking_figure <- function(target = "Cases") {
     geom_bar(stat = "identity") +
     theme_scoringutils() +
     facet_wrap(~scale) +
-    labs(y = "Model", x = "Relative skill")
+    labs(y = "Model", x = "Relative skill") +
+    scale_x_continuous(labels = simple_labels, breaks = c(0, 0.5, 1))
   
   plot_rel_skill_log <- summarised_pairwise |>
     filter(scale == "log") |>
@@ -804,7 +813,8 @@ ranking_figure <- function(target = "Cases") {
     theme_scoringutils() +
     facet_wrap(~scale) +
     labs(y = "", x = "Relative skill") +
-    theme(axis.title.y = element_blank())
+    theme(axis.title.y = element_blank()) +
+    scale_x_continuous(labels = simple_labels)
   
   diffs <- summarised_pairwise |>
     select(model, relative_skill, scale) |>
@@ -822,11 +832,14 @@ ranking_figure <- function(target = "Cases") {
           axis.text.y = element_blank(), 
           axis.ticks.y = element_blank(), 
           axis.line.y = element_blank()) +
-    facet_wrap(~title)
+    facet_wrap(~title) +
+    scale_x_continuous(labels = simple_labels)
   
   plot_rel_wis <- scores |>
     filter(horizon == 2, 
-           target_type == "Cases") |>
+           target_type == target, 
+           scale %in% c("log", "natural")) |>
+    mutate(scale = factor(scale, levels = c("natural", "log"))) |>
     mutate(model = factor(model, levels = ranking_log)) |>
[…]

### 2023-01-11 R/00 full paper code.R
update Figure 2 with new simulation

@@ -20,7 +20,8 @@ library(covidHubUtils) # remotes::install_github("https://github.com/reichlab/co
 set.seed(1234)
 
 ## ========================================================================== ##
-## FIGURE 1 + Figure XX
+## FIGURE 1 Numerical comparison of different measures of relative error
+##  + Figure XX
 ## ========================================================================== ##
 
 y_hat <- 1
@@ -91,8 +92,132 @@ p4 <- p3 +
 
 ggsave("output/figures/different-relative-errors-sqrt.png", width = 7, height = 3)
 
+
+
+
 ## ========================================================================== ##
-## FIGURE 2
+## FIGURE 3
+## ========================================================================== ##
+if (file.exists("output/data/simulation-figure-2.Rda")) {
+  scores_fig_2 <- readRDS(file = "output/data/simulation-figure-2.Rda")
+} else {
+  
+  time_points <- 1e5
+  n_means <- 60
+  means = (10^seq(1, 3.3, length.out = n_means))
+  model_names <- c("theta_1", "theta_1e9")
+  
+  df <- data.table(
+    state = 1:length(mean),
+    mean = means
+  ) |>
+    mutate(date = list(1:time_points)) |>
+    unnest(cols = date) |>
+    rowwise() |>
+    mutate(theta = list(c(1, 1e9)), 
+           model = list(model_names)) |>
+    ungroup() |>
+    unnest(cols = c("theta", "model")) |>
+    mutate(predictive_sample = rnbinom(n = time_points * n_means * length(model_names), 
+                                        size = theta,
+                                        mu = mean) + 1) 
+  
+  # add normal
+  df_normal <- filter(df, theta == 1) |>
+    mutate(model = "normal", 
+           theta = 0, 
+           predictive_sample = rnorm(n = time_points * n_means, 
+                                     mean = mean, 
+                                     sd = 1))
+  
+  df <- rbind(df, df_normal)
+  
+  scores_fig_2 <- df |>
+    group_by(mean, theta, model) |> 
+    summarise(
+      crps = mean(abs(predictive_sample[-1] - predictive_sample[-length(predictive_sample)]))/2,
[…]

### 2023-01-11 main.tex
address new comments

@@
 \begin{abstract}
-Forecast evaluation plays an essential role in improving forecasting models and can help inform future decision making.
-In epidemiology, probabilistic forecasts are often scored using the Weighted Interval Score (WIS) or the Continuous Ranked Probability Score (CRPS), which are a measure of absolute distance between forecast and observation.
-Infectious disease transmission is commonly conceptualised as a non-stationary processes in terms of the growth rate or the effective reproduction number.
-Scoring based on the absolute distance between forecast and observation may not reflect this underlying modelling task well and may make it difficult to compare forecasts across time, space, and different forecast targets.
-As a solution, we propose transforming the data before scoring the forecasts to better reflect our understanding of the generative process.
-Depending on the context, many transformations may be appropriate.
-We describe the log-transformation as one specific example in detail and discuss its interpretation as a measure of relative error and as a score for an implicit forecast of the exponential growth rate.
-Applying the log-transformation to data from the European COVID-19 Forecast Hub, we found that it changed model rankings, regardless of stratification by forecast date, location or target types, and emphasised situations in which models missed lower inflection points and consequent rises in numbers, while penalising missed peaks less severely.
+Forecast evaluation plays an essential role in improving forecasting models and can support their role in informing decision-making.
+In epidemiology, probabilistic forecasts are often scored using the Weighted Interval Score (WIS) or the Continuous Ranked Probability Score (CRPS), which are measures of absolute distance between forecast and observation.
+Infectious disease transmission is commonly modelled as an exponential process characterized by the growth rate or the effective reproduction number.
+However, if the true objective of the forecaster is to predict future growth rates or reproduction numbers then scoring based on the absolute distance between forecast and observation may not reflect this underlying task well and thus make it difficult to compare forecasts across time, space, and forecast targets.
+As a solution, we propose transforming the data before scoring the forecasts to better reflect the underlying generative process.
+We further propose the natural logarithm  as a candidate transformation and discuss its interpretation as a measure of relative error and as a score for an implicit forecast of the exponential growth rate.
+Applying the natural logarithm to data from the European COVID-19 Forecast Hub, we found that it changed model rankings, regardless of stratification by forecast date, location or target types, and emphasised situations in which models missed lower inflection points and consequent rises in numbers, while penalising missed peaks less severely.
+We conclude that appropriate transformations, of which the natural logarithm is only one possible option, should be considered when assessing the performance of different models in the context of infectious disease incidence.
 \end{abstract}
@@
 Epidemiological modelling in particular has received widespread attention during the COVID-19 pandemic.
-Evaluations of forecasts can provide feedback for forecasters and researchers to improve their models and train ensembles.
-It moreover helps decision makers distinguish good from bad predictions and choose forecasters and models that should inform future decisions.
+Evaluations of forecasts can provide feedback for researchers to improve their models and train ensembles.
+It moreover helps decision-makers distinguish good from bad predictions and choose forecasters (used here and elsewhere in this manuscript to mean anyone or anything that issues a forecast) and models that should inform future decisions.
 
@@
 The CRPS can be understood as a generalisation of the absolute error to predictive distributions, and interpreted n the natural scale of the data.
-The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in COVID-19 forecast hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
+The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in COVID-19 Forecast Hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
 The WIS is defined as
@@
 where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, $K$ is the number of predictive quantiles provided and $\boldsymbol{1}$ is the indicator function.
-We show an alternative definition based on central prediction intervals in Section \ref{sec:alternative-wis}.
+The WIS can be decomposed into three components, dispersion, overprediction, underprediction, which reflect the width of the forecast and whether it was centred above or below the observed value.
+We show an alternative definition based on central prediction intervals in Section \ref{sec:alternative-wis} which illustrates this decomposition more clearly.
 
@@
 This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellisChallengesControlCOVID192021}.
-If case numbers are evolving based on an exponential process and the  modelling task revolves around estimating the reproduction number or the corresponding growth rate, evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount, as errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
+If case numbers are evolving based on an exponential process and the  modelling task revolves around estimating the reproduction number or the corresponding growth rate, then evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount.
+This is because errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
 If one is to measure the ability of forecasts to assess and forecast the underlying infection dynamics, it may thus be more desirable to evaluate errors on the growth rate directly.
@@
 
-In many fields it is common practice to forecast transformed quantities (see e.g.
+In many fields, it is common practice to forecast transformed quantities (see e.g.
 \cite{taylorEvaluatingVolatilityInterval1999} in finance, \cite{mayrLogLevelVAR2015} in macroeconomics, \cite{loweStochasticRainfallrunoffForecasting2014} in hydrology or \cite{fuglstadDoesNonstationarySpatial2015} in meteorology).
-While the goal of the transformations is usually to improve accuracy of the predictions, they can also be used to enhance and complement the evaluation process.
-In this paper we argue that the aforementioned issues with evaluating forecasts based on measures of absolute error can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation.
+While the goal of the transformations is usually to improve the accuracy of the predictions, they can also be used to enhance and complement the evaluation process.
+In this paper, we argue that the aforementioned issues with evaluating forecasts based on measures of absolute error can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation.
 Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score.
@@
 
-For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation in the context of epidemic phenomena.
+For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation (referred to as the log-transformtation in the remainder of this manuscript) in the context of epidemic phenomena.
 Instead of a score representing the magnitude of absolute errors, applying a log-transformation prior to the CRPS yields a score which a) measures relative error (see Section \ref{sec:methods:relative}), b) provides a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate}) and c) is less dependent on the expected order of magnitude of the quantity to be predicted (see Section \ref{sec:methods:vst}).
@@
 We therefore argue that such evaluations on the logarithmic scale should complement the prevailing evaluations on the natural scale. 
-Other transformations may likewise be of interest, and we briefly explore the square root transformation as an alternative transformation. %[and population standardization?]. 
-Our analysis focuses on the CRPS (or WIS) as evaluation metric, given its widespread use throughout the COVID-19 pandemic.
+Other transformations may likewise be of interest. We briefly explore the square root transformation as an alternative transformation. %[and population standardization?]. 
+Our analysis focuses on the CRPS (or WIS) as an evaluation metric, given its widespread use throughout the COVID-19 pandemic.
 
 The remainder of the article is structured as follows.
-In Sections \ref{sec:methods:relative}--\ref{sec:methods:vst} we provide some mathematical intuition on applying the log transformation prior to evaluating the CRPS, highlighting the connections to relative error measures, the epidemic growth rate and variance stabilizing transformations.
+In Sections \ref{sec:methods:relative}--\ref{sec:methods:vst} we provide some mathematical intuition on applying the log-transformation prior to evaluating the CRPS, highlighting the connections to relative error measures, the epidemic growth rate and variance stabilizing transformations.
 %We first provide a mathematical intuition of the impact of the log-transformation and its interpretation as relative error (Section \ref{sec:methods:relative}, measure of the accuracy of a forecast of the growth rate (Section \ref{sec:methods:growthrate}) and it's effect as a variance-stabilising transformation that makes can make scores independent of the mean of the forecast target (Section \ref{sec:methods:vst}.
-We then discuss the effect of the transformation on forecast rankings (section \ref{sec:methods:rankings} and discuss some practical considerations relevant for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}).
-We illustrate the effect of the log-transformation using forecasts submitted to the European COVID-19 Forecast Hub  \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022} (Section \ref{sec:results}).
-Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues on the use of transformed scores in the evaluation of infectious disease forecasts (Section \ref{sec:discussion}).
+We then discuss practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}) and the effect of the log-transformation on forecast rankings (section \ref{sec:methods:rankings}.
+To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
+Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues (Section \ref{sec:discussion}).
 
@@
 
-To illustrate the effect of applying the logarithm prior to evaluating forecasts we consider the absolute error, which the CRPS and WIS generalize to probabilistic forecasts.
+To illustrate the effect of applying the natural logarithm prior to evaluating forecasts we consider the absolute error, which the CRPS and WIS generalize to probabilistic forecasts.
+We assume strictly positive support (meaning that no specific handling of zero values is needed), a restriction we will address when applying this transformation in practice.
 When considering a point forecast $\hat{y}$ for a quantity of interest $y$, such that
@@
 \end{linenomath*}
-the absolute error is given by $|\varepsilon|$. When taking the logarithm of the forecast and the observation first (assuming that both are strictly positive), thus considering 
+the absolute error is given by $|\varepsilon|$. When taking the logarithm of the forecast and the observation first, thus considering 
 \begin{linenomath*}
@@
 $$
-The absolute error after log transformation is thus an approximation of the \textit{absolute percentage error} \citep[APE,][]{gneitingMakingEvaluatingPoint2011a}.
-As in the approximation we assumed that $\hat{y} \approx y$, we can just as well interpret it as an approximation of the \textit{relative error} (RE)
+The absolute error after log transforming is thus an approximation of the \textit{absolute percentage error} \citep[APE,][]{gneitingMakingEvaluatingPoint2011a}.
+As we assumed that $\hat{y} \approx y$, we can also interpret it as an approximation of the \textit{relative error} (RE)
 $$
@@
 As Figure \ref{fig:SAPE} shows, the alignment with the SAPE is in fact the closest and holds quite well even if predicted and observed value differ by a factor of two or three.
-Moving back to probabilistic forecasts, the CRPS applied to log-transformed forecasts and outcomes can thus be seen as a probabilistic counterpart to the symmetric absolute percentage error, which offers an appealing intuitive interpretation.
+Generalising to probabilistic forecasts, the CRPS applied to log-transformed forecasts and outcomes can thus be seen as a probabilistic counterpart to the symmetric absolute percentage error, which offers an appealing intuitive interpretation.
 
@@
 
-Another interpretation can be based on viewing the process to be forecast as exponential with time-varying growth rate $r(t)$~\citep[see, e.g.,][]{wallingaHowGenerationIntervals2007}, i.e.
+Another interpretation for the log-transform is possible if the generative process is framed as exponential with a time-varying growth rate $r(t)$~\citep[see, e.g.,][]{wallingaHowGenerationIntervals2007}, i.e.
 \begin{linenomath*}
@@
            &= \left| \log \left[ y_0 \exp (\bar{\hat{r}} t ) \right] - \log \left[ y_0 \exp (\bar{r}t) \right] \right|\\
-           &= t \left| \bar{\hat{r}} - \bar{r} \right|.
+           &= t \left| \bar{\hat{r}} - \bar{r} \right|
 \end{align*}
@@
 %
+where $\bar{\hat{r}}$ is the true mean growth rate and $\bar{r}$ is the forecast mean growth rate.
 We thus evaluate the error in the mean exponential growth rate, scaled by the length of the time period considered.
-For small $\bar{r}t$, $ y_0 \exp (\bar{r}t)$ is approximately equivalent to $ y_0 (1 + \bar{r}t)$, which is sometimes used as an alternative definition of the growth rate (then called the multiplicative growth rate).
-This means that by applying the CRPS rather than the absolute error we perform a probabilistic evaluation of forecasts of the epidemic growth rate.
+Again generalising this to the CRPS and WIS implies a probabilistic evaluation of forecasts of the epidemic growth rate.
 
@@
 
-When evaluating models across sets of forecasting tasks, it may be considered desirable for each target to have a similar impact on the overall results.
-In disease forecasting this is not the case when using the CRPS, as the latter typically scales with the order of magnitude of the quantity to be predicted.
+When evaluating models across sets of forecasting tasks, it may be desirable for each target to have a similar impact on the overall results.
+In disease forecasting, this is not the case when using the CRPS, as the latter typically scales with the order of magnitude of the quantity to be predicted.
 Average scores are then dominated by the results achieved for targets with high expected outcomes.
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/SIM-mean-state-size.png}
+    \caption{Expected CRPS scores as a function of the mean and variance of the forecast quantity.
+We generated 10,000 samples for three distributions with differing relationships between mean and variance.
+We then computed expected CRPS values assuming an ideal forecaster with predictive distribution equal to the data-generating distribution.
+The variance of the negative binomial is given as $\sigma^2 = \mu + \mu^2 / \theta$, meaning that for large theta the negative binomial distribution is equal to the Poisson distribution.
+We used values of $\theta = 1$ (red), $\theta = \frac{1000}{\sqrt{\mu}}$ (green) and $\theta = 10^9$ (blue, approximately corresponding to a Poisson distribution).
+To make the scores for the different distributions comparable, scores were normalised to one, meaning that the mean score for every distribution (red, green, blue) is one.
+    A: Normalised expected CRPS for ideal forecasts with increasing means for three distribution with different relationships between mean and variance. B: A but with x axis on the log scale.}
+    \label{fig:SIM-wis-state-size-mean}
+\end{figure}
+
+\subsection{Practical considerations}
+\label{sec:methods:considerations}
+
+Transformations that are strictly monotonic are permissible in the sense that they maintain the propriety of the score.
+This is because even though rankings of models may change forecasts will in expectation still minimise their score if they report a predictive distribution that is equal to the data-generating distribution.
+This condition holds for both the log and square root transformations, as well as many others.
+However, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee propriety.
+In the case of log transforms, taking the logarithm of the scores, rather than scoring the log-transformed forecasts and data, results in an improper score.
+This is because taking the logarithm of the CRPS (or WIS) results in a score that does not penalise outliers enough and therefore incentivises overconfident predictions.
+We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can easily be seen that overconfident models perform best in terms of the log WIS.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/example-log-first.png}
+    \caption{Illustration of impropriety of log-transformed CRPS.
+We assume $Y \sim \text{LogNormal}(0, 1)$ and evaluate the expected CRPS for predictive distributions $\text{LogNormal}(0, \sigma)$ with varying values of $\sigma \in [0.1, 2]$.
+For the regular CRPS (left) and CRPS applied to log-transformed outcomes (middle), the lowest expectation is achieved for the true value $\sigma = 1$.
+For the log-transformed CRPS, the optimal value is  0.9, i.e.
+there is an incentive to report a forecast that is too sharp.}
+    \label{fig:log-improper}
+\end{figure}
+
+In practice, one issue with transformations, such as the log transform, is that they are not readily applicable to negative numbers or zero values, which need to be removed or otherwise handled.
+Small discrete values can pose an additional problem as it may be difficult to reliably evaluate relative errors.
+One common approach is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm CITATION.
+This represents a strictly monotonic transformation and therefore preserves the propriety of the resulting score.
+The choice of the quantity to add does however influences scores and rankings, as measures of relative errors shrink when adding a constant to the forecast and the observation.
+The impact is greatest for forecasts of small quantities, as relative errors shrink more the smaller the original value before adding a constant.
+In principle, it would also be possible to exploit the difference in how scores are affected in order to equalise scores for forecasts of quantities across different scales by adding a larger number (e.g.
+100 or 10,000) if average scores are dominated too strongly by forecasts for small quantities.
+%Add that the dispersion component can blow up
+
 \subsection{Effects on model rankings}
@@
 Overall model rankings would be expected to differ even more when scores are averaged across multiple forecasts or targets.
-The change in rankings of aggregate scores is not mainly driven by the kinds of changes in model rankings for single forecasts just discussed.
-Rather, average scores will be most influenced by the order of magnitude of scores for different forecast targets across time, location and target type.
+The change in rankings of aggregate scores is mainly driven by the order of magnitude of scores for different forecast targets across time, location and target type and less so by the kind of changes in model rankings for single forecasts discussed above.
 Large observations will dominate average CRPS values when evaluation is done on the natural scale, but much less so after log transformation.
-Depending on the relationship between mean and variance of the forecast target, a log-transformation may even lead to systematically larger scores assigned to small forecast targets, as illustrated in Figure \ref{fig:SIM-wis-state-size-mean}.
-A different transformation like the square-root may then be more appropriate to make scores independent of the mean of the target quantity.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/SIM-mean-state-size.png}
-    \caption{Expected CRPS scores as a function of the mean and variance of the forecast quantity.
-We generated 10,000 samples for three distributions with differing relationships between mean and variance.
-We then computed expected CRPS values assuming an ideal forecaster with predictive distribution equal to the data-generating distribution.
-The variance of the negative binomial is given as $\sigma^2 = \mu + \mu^2 / \theta$, meaning that for large theta the negative binomial distribution is equal to the Poisson distribution.
-We used values of $\theta = 1$ (red), $\theta = \frac{1000}{\sqrt{\mu}}$ (green) and $\theta = 10^9$ (blue, approximately corresponding to a Poisson distribution).
-To make the scores for the different distributions comparable, scores were normalised to one, meaning that the mean score for every distribution (red, green, blue) is one.
-    A: Normalised expected CRPS for ideal forecasts with increasing means for three distribution with different relationships between mean and variance. B: A but with x axis on the log scale.}
-    \label{fig:SIM-wis-state-size-mean}
-\end{figure}
+Depending on the relationship between the mean and variance of the forecast target, a log-transformation may even lead to systematically larger scores assigned to small forecast targets, as illustrated in Figure \ref{fig:SIM-wis-state-size-mean}.
+A different transformation like the square-root may then be more appropriate to make scores independent of the mean of the target quantity (as discussed in Section \ref{sec:methods:vst}).
 
@@
 
-\subsection{Practical considerations}
-\label{sec:methods:considerations}
-
-Transformations that are strictly monotonic are permissible in the sense that they maintain the propriety of the score.
-This is because even though rankings of models may change forecasts will in expectation still minimise their score if they report a predictive distribution that is equal to the data-generating distribution.
-This condition holds for both the log and square root transformations, as well as many others.
-However, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee propriety.
-In the case of log transforms, taking the logarithm of the scores, rather than scoring the log-transformed forecasts and data, results in an improper score.
-This is because taking the logarithm of the WIS / CRPS results in a score that does not penalise outliers enough and therefore incentivises overconfident predictions.
-We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can easily be seen that overconfident models perform best in terms of the log WIS.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/example-log-first.png}
-    \caption{Illustration of impropriety of log-transformed CRPS.
-We assume $Y \sim \text{LogNormal}(0, 1)$ and evaluate the expected CRPS for predictive distributions $\text{LogNormal}(0, \sigma)$ with varying values of $\sigma \in [0.1, 2]$.
-For the regular CRPS (left) and CRPS applied to log-transformed outcomes (middle), the lowest expectation is achieved for the true value $\sigma = 1$.
-For the log-transformed CRPS, the optimal value is  0.9, i.e.
-there is an incentive to report a forecast that is too sharp.}
-    \label{fig:log-improper}
-\end{figure}
-
-In practice, one issue with transformations, such as the log transform, is that they are not readily applicable to negative numbers or zero values, which need to be removed or handled otherwise.
-Small discrete values can pose an additional problem as it may be difficult to reliably evaluate relative errors.
-One common approach is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm.
-This represents a strictly monotonic transformation and therefore still preserves the propriety of the resulting score.
-The choice of the number to add influences scores and rankings, as measures of relative errors shrink when adding a constant to the forecast and the observation.
-Adding a number to the forecasts and observations changes scores for forecasts of small quantities most, as relative errors shrink more the smaller the original value before adding a constant.
-In principle, it would also be possible to exploit the difference in how scores are affected in order to equalise scores for forecasts of quantities across different scales by adding a larger number (e.g.
-100 or 10,000) if average scores are dominated too strongly by forecasts for small quantities.
-
 \section{Empirical example: the European Forecast Hub}
@@
 
+\subsection{Setting}
+
 As an empirical comparison of evaluating forecasts on the natural and on the log scale, we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
-The European COVID-19 Forecast Hub is one of several COVID-19 Forecast hubs \citep{cramerEvaluationIndividualEnsemble2021, bracherShorttermForecastingCOVID192021} which have been systematically collecting, aggregating and evaluating forecasts of several COVID-19 targets created by different teams every week.
-Forecasts are made one to four weeks ahead into the future and follow a quantile-based format with a set of 22 quantiles plus the median ($0.01, 0.025, 0.05, ..., 0.5, ...
+The European COVID-19 Forecast Hub is one of several COVID-19 Forecast Hubs \citep{cramerEvaluationIndividualEnsemble2021, bracherShorttermForecastingCOVID192021} which have been systematically collecting, aggregating and evaluating forecasts of several COVID-19 targets created by different teams every week.
+Forecasts are made one to four weeks ahead into the future and follow a quantile-based format with a set of 23 quantiles ($0.01, 0.025, 0.05, ..., 0.5, ...
 0.95, 0.975, 0.99$).
 
-The forecasts used for the purpose of this illustration are forecasts submitted between March 8 2021 and October 18 2021 CHECK AGAIN for reported cases and deaths from COVID-19.
+The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 12th of May 2022 \textbf{CHECK AGAIN} for reported cases and deaths from COVID-19.
 We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 16 CHECK AGAIN forecast dates (see Figure \ref{fig:HUB-num-avail-models}).
-We removed all observations marked as data anomalies by the European Forecast Hub as well as all remaining negative observed values.
+We removed all observations marked as data anomalies by the European Forecast Hub \citep{sherrattPredictivePerformanceMultimodel2022} as well as all remaining negative observed values.
 In addition, we filtered out erroneous forecasts characterised by a median forecast of 0 and an actual observed value of larger than 150.
@@
 
-In addition to the WIS we use pairwise comparisons \citep{cramerEvaluationIndividualEnsemble2021} to evaluate the relative performance of models across countries in the presence of missingness.
-In a first step, score ratios are computed for all pairs of models by taking the set of overlapping forecasts between the two models and dividing the score of one model over the score achieved by the other model.
+In addition to the WIS we use pairwise comparisons \citep{cramerEvaluationIndividualEnsemble2021} to evaluate the relative performance of models across countries in the presence of missing forecasts.
+In the first step, score ratios are computed for all pairs of models by taking the set of overlapping forecasts between the two models and dividing the score of one model by the score achieved by the other model.
 The relative skill for a given model compared to others is then obtained by taking the geometric mean of all score ratios which involve that model.
 Low values are better, and the average model receives a relative skill score of 1.
+
+\subsection{Illustration and qualitative observations}
+
+Show and discuss Fig 7 here? CHECK AGAIN
+
+\subsection{Descriptive analysis of observed values and scores obtained on the natural and logarithmic scales CHECK AGAIN}
 
@@
 Both on the natural scale as well to a slightly lesser extent on the log scale, scores increased considerably with increasing forecast horizon (see Figure \ref{fig:HUB-mean-locations}E).
-This reflects the increasing difficulty of forecasts further into the future.
+This reflects the increasing difficulty of forecasts further into the future and, for the log scale, corresponds with our expectations from Section \ref{sec:methods:growthrate}.
 
@@
 
-As expected, on the natural scale, scores correlate strongly with the magnitude of the median forecast (Spearman rank correlation: 0.9 and 0.92) (see Figure\ref{fig:HUB-mean-locations}A, B).
-We report the Spearman rank correlation here, rather than the Pearson correlation, as it tends to be more robust and is invariant to transformations.
-For scores on the log scale, the correlation is weaker and slightly negative (Spearman correlation: -0.35 and -0.55), implying a tendency towards larger scores for forecasts of smaller quantities.
+\subsection{Regression analysis to determine the variance-stabilizing transformation}
 
-In order to test the relationship between scores and the order of magnitude of the forecast, we ran the following regression on scores on the natural scale
+As argued in Section \ref{sec:methods:vst}, the mean-variance, or mean-CRPS, relationship determines which transformation can serve as a VST. To assess this empirically, we ran the regression
 
 \begin{equation*}
-    \text{CRPS}(F, y) = \exp(\alpha) \cdot \text{median}(F)^{\beta}, 
+    \log[\text{CRPS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], 
 \end{equation*}
-where $F$ is the predictive distribution and $y$ the observed value. In addition, we ran the following regression on scores on the log scale
+where the predictive distribution $F$ and the observation $y$ are on the natural scale. This is equivalent to
+\begin{equation*}
+    \text{CRPS}(F, y) = \exp(\alpha) \times \text{median}(F)^\beta, 
+\end{equation*}
+meaning that we estimate a polynomial relationship between the predictive median and achieved CRPS.
+Note that we are using predictive medians rather than means as only the former are available in the European COVID-19 Forecast Hub.
+We applied the regression to case and death forecasts, pooled across horizons and stratified for one through four-week-ahead forecasts.
+Results are provided in Table \ref{tab:HUB-regression}.
+It can be seen that the estimates of $\beta$ always take a value below 1.
+As the CRPS scales with the standard deviation (see Section \ref{sec:methods:vst}), this can be interpreted as a slightly sub-quadratic mean-variance relationship.
+The logarithmic transformation should thus approximately stabilize the variance (and CRPS), possibly leading to somewhat higher scores for smaller forecast targets.
+The square-root transformation, on the other hand, can be expected to still lead to higher CRPS values for targets of higher orders of magnitude.
 
+To check this, we ran the regressions
 \begin{equation*}
-    \text{CRPS}(\log F, \log y) = \alpha^* + \beta^* \cdot \log{(\text{median}(F))}
+    \text{CRPS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},
 \end{equation*}
-and the following regression on scores obtained after applying a square-root-transformation
+where $F_{\log}$ is the predictive distribution for $\log(y)$, and
+\begin{equation*}
+    \text{CRPS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{\sqrt{\ }} + \beta_{\sqrt{\ }} \cdot \sqrt{\text{median}(F)},
+\end{equation*} 
+where $F_{\sqrt{\ }}$ is the predictive distribution on the square-root scale.
+As can be seen from Table \ref{tab:HUB-regression}, the results indeed indicate small targets lead to larger average CRPS when using the log transform ($\beta_{\log} < 0$), while the opposite is true for the square-root transform ($\beta_{\sqrt{\ }} > 0$).
+The results of the three regressions are also displayed in Figure \ref{fig:HUB-cors}.
+It appears that the log transformation is able to stabilise CRPS value in this empirical example.
 
-\begin{equation*}
-    \text{CRPS}(\sqrt{F}, \sqrt{y}) = \alpha^{**} + \beta^{**} \cdot \sqrt{(\text{median}(F))}
-\end{equation*}
-
-We found that on the natural scale, the order of magnitude of the forecast (represented by the median, rather than the mean, as forecasts were stored as quantiles and the mean was not available) strongly influenced scores, with estimated $\beta$-values close to one.
+% We found that on the natural scale, the order of magnitude of the forecast (represented by the median, rather than the mean, as forecasts were stored as quantiles and the mean was not available) strongly influenced scores, with estimated $\beta$-values close to one.
 On the log scale, this influence was small, with estimated $\beta^*$-values close to zero.
@@
 \end{figure}
+
+\subsection{Impact of logarithmic transformation on model rankings}
 
@@
 
-The rank order of different forecasting models changes noticeable (see Figure \ref{fig:HUB-rank-order}. 
+The rank order of different forecasting models changes noticeably (see Figure \ref{fig:HUB-rank-order}. 
 % I removed this part as it doesn't seem accurate anymore / we don't even have all the models in there anymore. 
@@
 
-In this paper we proposed the use of transformations, with a particular focus on the log transformation, in order to address issues that arise when evaluating epidemiological forecasts based on measures of absolute error and their probabilistic generalisations (i.e CRPS and WIS).
-We gave a mathematical intuition for the log transformation and showed that scores on the log scale can be interpreted as a) a measure of relative prediction errors, as well as b) a score for a forecast of the exponential growth rate of the target quantity and c) as variance stabilising transformation in some settings.
+In this paper we proposed the use of transformations, with a particular focus on the natural logarithmic transformation, when evaluating forecasts in an epidemiological setting.
+These transformations can address issues that arise when evaluating epidemiological forecasts based on measures of absolute error and their probabilistic generalisations (i.e CRPS and WIS).
+We showed that scores obtained after log-transforming both forecasts and observations can be interpreted as a) a measure of relative prediction errors, as well as b) a score for a forecast of the exponential growth rate of the target quantity and c) as variance stabilising transformation in some settings.
 When applying this approach to forecasts from the European COVID-19 Forecast Hub, we found overall scores on the log scale to be more equal across, time, location and target type (cases, deaths) than scores on the natural scale.
@@
 
-The log-transformation is only one of many transformation that may be useful and appropriate in an epidemiological context.
-One easy transformation is to divide all forecasts by the population in a location, in order to obtain forecasts of standardised incidences per e.g.
-100,000 inhabitants.
-A second transformation was mentioned above, namely converting forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last observed value.
-One could also think about applying a Box-Cox \citep{boxAnalysisTransformations1964} or a square-root transformation in order to stabilise the variance of the forecasts.
-Another promising transformation would be to divide each forecast by the forecast of the previous week (and observations by the observation of the previous week), in order to obtain forecasts for week-to-week multiplicative growth rates, or alternatively, to ask forecasters to provide estimates of the weekly relative change applied to the latest data and subsequent forecast points directly.
+The log-transformation is only one of many transformations that may be useful and appropriate in an epidemiological context.
+One obvious option is to apply a population standardization to obtain incidence forecasts e.g., per 100,000 population If one is interested in multiplicative, rather than exponential growth rates, one could convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last observed value.
+We suggested to use the natural logarithm as a variance-stabilising transformation (or alternatively the square-root transformation in the case of a Poisson distributed variable).
+Other VST like the Box-Cox \citep{boxAnalysisTransformations1964} are conceivable as well.
+Another promising transformation would be take differences between forecasts on the log scale, or alternatively
+to divide each forecast by the forecast of the previous week (and analogously for observations), in order to obtain forecasts for week-to-week growth rates.
+One could then also ask forecasters to provide estimates of the weekly relative change applied to the latest data and subsequent forecast points directly.
 This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluate the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}.
@@
 This would allow forecast consumers to assign explicit weights to different qualities of the forecasts they might care about.
+
+In this work we focused on the CRPS and WIS, which are widely used in the evaluation of epidemic forecasts.
+We note that for the logarithmic score, which has also been used e.g., in some editions of the FluSight challenge \cite{reichCollaborativeMultiyearMultimodel2019}, the question of the right scale to evaluate forecasts does not arise.
+It is known that log score differences between different forecasters are invariant to monotonic transformations of the outcome variable (see e.g., \citealt{Diks2011}).
+This is clearly an advantage of the logarithmic score over the CRPS; however, the logarithmic score is known to have other severe downsides, e.g., its low robustness to occasional misguided forecasts; see \cite{bracherEvaluatingEpidemicForecasts2021} for a more detailed discussion.
 

### 2023-01-11 R/00 full paper code.R
udpate Figure 2

@@ -96,7 +96,7 @@ ggsave("output/figures/different-relative-errors-sqrt.png", width = 7, height = 
 
 
 ## ========================================================================== ##
-## FIGURE 3
+## FIGURE 2
 ## ========================================================================== ##
 if (file.exists("output/data/simulation-figure-2.Rda")) {
   scores_fig_2 <- readRDS(file = "output/data/simulation-figure-2.Rda")
@@ -105,7 +105,7 @@ if (file.exists("output/data/simulation-figure-2.Rda")) {
   time_points <- 1e5
   n_means <- 60
   means = (10^seq(1, 3.3, length.out = n_means))
-  model_names <- c("theta_1", "theta_1e9")
+  model_names <- c("theta_10", "theta_1e9")
   
   df <- data.table(
     state = 1:length(mean),
@@ -114,16 +114,16 @@ if (file.exists("output/data/simulation-figure-2.Rda")) {
     mutate(date = list(1:time_points)) |>
     unnest(cols = date) |>
     rowwise() |>
-    mutate(theta = list(c(1, 1e9)), 
+    mutate(theta = list(c(10, 1e9)), 
            model = list(model_names)) |>
     ungroup() |>
     unnest(cols = c("theta", "model")) |>
     mutate(predictive_sample = rnbinom(n = time_points * n_means * length(model_names), 
                                         size = theta,
-                                        mu = mean) + 1) 
+                                        mu = mean)) 
   
   # add normal
-  df_normal <- filter(df, theta == 1) |>
+  df_normal <- filter(df, theta == 10) |>
     mutate(model = "normal", 
            theta = 0, 
            predictive_sample = rnorm(n = time_points * n_means, 
@@ -136,10 +136,11 @@ if (file.exists("output/data/simulation-figure-2.Rda")) {
     group_by(mean, theta, model) |> 
     summarise(
       crps = mean(abs(predictive_sample[-1] - predictive_sample[-length(predictive_sample)]))/2,
-      crps_log = mean(abs(log(predictive_sample[-1]) - log(predictive_sample[-length(predictive_sample)])))/2,
+      crps_log = mean(abs(log(predictive_sample[-1] + 1) - log(predictive_sample[-length(predictive_sample)] + 1)))/2,
[…]

### 2023-01-11 R/00 full paper code.R
update Figures 5 and 6

@@ -217,110 +217,9 @@ scores_fig_2 |>
 ggsave("output/figures/SIM-score-approximation.png", width = 7, height = 8)
 
 
-## ========================================================================== ##
-## FIGURE 4: Illustration of the effect of the log-transformation of 
-##           the ranking for a single forecast
-## ========================================================================== ##
-
-quantile <- seq(0.005, 0.995, 0.005)
-mu <- 10
-theta_1 <- 1
-theta_2 <- 1e9
-
-data <- data.table(
-  quantile = quantile,
-  model_A = qnbinom(p = quantile, size = theta_1, mu = mu), 
-  model_B = qnbinom(p = quantile, size = theta_2, mu = mu)
-) |>
-  pivot_longer(cols = c(model_A, model_B), names_to = "model", values_to = "prediction") |>
-  mutate(true_value = list(1:50), 
-         id = list(1:50)) |>
-  unnest(cols = c(true_value, id))
-
-scores_natural <- data |>
-  score(metrics = "interval_score") |>
-  summarise_scores(by = c("model", "id")) |>
-  mutate(scale = "natural")
-
-scores_log <- data |>
-  mutate(true_value = log(true_value), 
-         prediction = log(prediction)) |>
-  score(metrics = "interval_score") |>
-  summarise_scores(by = c("model", "id"), na.rm = TRUE) |>
-  mutate(scale = "log")
-
-scores_sqrt <- data |>
-  mutate(true_value = sqrt(true_value), 
-         prediction = sqrt(prediction)) |>
-  score(metrics = "interval_score") |>
-  summarise_scores(by = c("model", "id"), na.rm = TRUE) |>
-  mutate(scale = "sqrt")
-
-scores <- rbind(scores_natural, scores_log, scores_sqrt) |>
-  mutate(scale = factor(scale, levels = c("natural", "log", "sqrt")), 
-         Forecaster = ifelse(model == "model_A", "A", "B"))
-
-nbinom_natural <- data.table(
-  A = rnbinom(100000, mu = 10, size = theta_1), 
-  B = rnbinom(100000, mu = 10, size = theta_2)
-) |>
-  pivot_longer(cols = c(A, B), names_to = "Forecaster") 
-
-plot_fct <- function(scores, scale_factor, nbinom, filter_scale = "natural") {
-  
-  scores <- filter(scores, scale == filter_scale)
-  
-  scores |>
-    ggplot(aes(x = id, y = interval_score / scale_factor, color = Forecaster)) + 
[…]

### 2023-01-11 R/Refetch data from Forecast Hub.R

@@ -5,6 +5,8 @@ library(dplyr)
 library(data.table)
 library(purrr)
 library(scoringutils)
+library(here)
+library(stringr)
 
 analysis_date <- "2022-12-12"
   
@@ -185,6 +187,7 @@ hub_data <- filter(hub_data, target_type != "")
 remove <- hub_data |> 
   filter(quantile == 0.5) |>
   filter((true_value > 10) & (prediction > 20*true_value) |
+         (true_value > 0) & (prediction > 100*true_value) |
          (true_value > 50) & (prediction < 1/50 * true_value) |
          (true_value == 0) & (prediction > 100)) |>
   select(-quantile)

### 2023-01-12 main.tex

@@
 % Suggestion J:
-% Forecast evaluation plays an essential role in the development cycle of pedictive epidemic models and can inform their use for public health decision making.
+% Forecast evaluation plays an essential role in the development cycle of predictive epidemic models and can inform their use for public health decision-making.
 Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which are both measures of the absolute distance between the forecast distribution and the observation.
-They are commonly applied directly to predicted and observed incidence counts, but it can be questioned whether this yields the most meaningful results given the exponential nature of epidemic processes and the fact that observed values over space and time can span several orders of magnitude.
-In this paper we argue that applying the CRPS to incidence counts on a logarithmic rather than natural scale effectively mitigates these difficulties, yielding epidemiologically meaningful and easily interpretable results.
-We motivate the procedure threefold: firstly, it can be interpreted as a probabilistic version of a relative error.
+They are commonly applied directly to predicted and observed incidence counts, but it can be questioned whether this yields the most meaningful results given the exponential nature of epidemic processes and the several orders of magnitude that observed values can span over space and time.
+In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results.
+We motivate the procedure threefold using the CRPS on log-transformed counts as an example: firstly, it can be interpreted as a probabilistic version of a relative error.
 Secondly, it reflects how well models predicted the time-varying epidemic growth rate.
 And lastly, using arguments on variance-stabilizing transformations, it can be shown that under the assumption of a quadratic mean-variance relationship, the logarithmic transformation leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity.
-This implies that different forecasting tasks will will have a similar impact on overall evaluation results.
-Applying the log-transformation to data from the European COVID-19 Forecast Hub, we find that it changes model rankings, regardless of stratification by forecast date, location or target types.
-Situations in which models missed beginning upward swings are more strongly emphasized, while penalising missed peaks are less severely penalized.
-Also, we show that the assumption of a quadratic mean-variance relationship is quite closely aligned with empirical patterns.
+This implies that different forecasting tasks will have a similar impact on overall evaluation results.
+Applying the log transformation to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types.
+Situations in which models missed beginning upward swings are more strongly emphasized while failing to predict a downturn following a peak is less severely penalized.
+We conclude that appropriate transformations, of which the natural logarithm is only one particularly attractive option, should be considered when assessing the performance of different models in the context of infectious disease incidence.
 
@@
 We then computed expected CRPS values assuming an ideal forecaster with predictive distribution equal to the data-generating distribution.
-The variance of the negative binomial is given as $\sigma^2 = \mu + \mu^2 / \theta$, meaning that for large theta the negative binomial distribution is equal to the Poisson distribution.
-We used values of $\theta = 1$ (red), $\theta = \frac{1000}{\sqrt{\mu}}$ (green) and $\theta = 10^9$ (blue, approximately corresponding to a Poisson distribution).
+The first distribution (red) is a truncated normal distribution with constant variance (we chose a $\sigma = 1$ in order to only obtain positive samples).
+The second (green) is a geometric distribution (a negative binomial distribution with variance $\sigma^2 = \mu + \mu^2$.
+The third (blue) is a
+    Poisson distribution with $\sigma^2 = \mu + \mu$.
 To make the scores for the different distributions comparable, scores were normalised to one, meaning that the mean score for every distribution (red, green, blue) is one.
@@
 Shown are CRPS (or WIS, respectively) values as a function of the observed value for two forecasters.
-Model A issues a negative binomial forecast with mean = 10 and $\theta = 1$ (resulting in a variance of 110), while Model B issues a Poisson distribution with mean = 10 (resulting in a variance of 10).}
+Model A issues a geometric distribution with mean $\mu = 10$ and variance $\sigma^2 = \mu + \mu^2 = 110$), while Model B issues a Poisson distribution with mean and variance equal to 10.
+Zeroes in this illustrative example were handled by adding one before applying the natural logarithm.}
 \label{fig:illustration-ranking}
@@
 The results of the three regressions are also displayed in Figure \ref{fig:HUB-cors}.
-It appears that the log transformation is able to stabilise CRPS value in this empirical example.
+It appears that the log transformation is most suitable to stabilise the CRPS values in this empirical example.
 
@@
 
-In this paper we proposed the use of transformations, with a particular focus on the natural logarithmic transformation, when evaluating forecasts in an epidemiological setting.
+In this paper, we proposed the use of transformations, with a particular focus on the natural logarithmic transformation, when evaluating forecasts in an epidemiological setting.
 These transformations can address issues that arise when evaluating epidemiological forecasts based on measures of absolute error and their probabilistic generalisations (i.e CRPS and WIS).
@@
 When applying this approach to forecasts from the European COVID-19 Forecast Hub, we found overall scores on the log scale to be more equal across, time, location and target type (cases, deaths) than scores on the natural scale.
-Scores on the log scale were much less influenced by the overall incidence level in a country, and showed a slight tendency to be higher in locations with very low incidences.
+Scores on the log scale were much less influenced by the overall incidence level in a country and showed a slight tendency to be higher in locations with very low incidences.
 We found that model rankings changed noticeably.
@@
 Applying a log-transformation prior to the WIS means that forecasts are evaluated in terms of relative errors and errors on the exponential growth rate, rather than absolute errors.
-The most important strength of this approach is that the evaluation better accommodates the exponential nature of epidemiological process and the types of errors forecasters who accurately model those processes are expected to make.
+The most important strength of this approach is that the evaluation better accommodates the exponential nature of the epidemiological process and the types of errors forecasters who accurately model those processes are expected to make.
 The log-transformation also helps avoid issues with scores being strongly influenced by the order of magnitude of the forecast quantity, which can be an issue when evaluating forecasts on the natural scale.
@@
 One obvious option is to apply a population standardization to obtain incidence forecasts e.g., per 100,000 population If one is interested in multiplicative, rather than exponential growth rates, one could convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last observed value.
-We suggested to use the natural logarithm as a variance-stabilising transformation (or alternatively the square-root transformation in the case of a Poisson distributed variable).
+We suggested using the natural logarithm as a variance-stabilising transformation (VST) or alternatively the square-root transformation in the case of a Poisson distributed variable.
 Other VST like the Box-Cox \citep{boxAnalysisTransformations1964} are conceivable as well.
-Another promising transformation would be take differences between forecasts on the log scale, or alternatively
-to divide each forecast by the forecast of the previous week (and analogously for observations), in order to obtain forecasts for week-to-week growth rates.
+Another promising transformation would be to take differences between forecasts on the log scale, or alternatively to divide each forecast by the forecast of the previous week (and analogously for observations), in order to obtain forecasts for week-to-week growth rates.
 One could then also ask forecasters to provide estimates of the weekly relative change applied to the latest data and subsequent forecast points directly.
-This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluate the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}.
-This, unfortunately is not feasible in the context of forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate, but may be an interesting approach if predictive samples are available.
-What makes this approach of transforming forecasts before scoring particularly interesting is the possibility to construct composite scores as a weighted sum of scores based on different transformations.
+This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluating the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}.
+This, unfortunately, is not feasible in the context of forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate, but may be an interesting approach if predictive samples are available as it enables the possibility of constructing composite scores as a weighted sum of scores based on different transformations.
 This would allow forecast consumers to assign explicit weights to different qualities of the forecasts they might care about.
 
-In this work we focused on the CRPS and WIS, which are widely used in the evaluation of epidemic forecasts.
+In this work, we focused on the CRPS and WIS, which are widely used in the evaluation of epidemic forecasts.
 We note that for the logarithmic score, which has also been used e.g., in some editions of the FluSight challenge \cite{reichCollaborativeMultiyearMultimodel2019}, the question of the right scale to evaluate forecasts does not arise.
 It is known that log score differences between different forecasters are invariant to monotonic transformations of the outcome variable (see e.g., \citealt{Diks2011}).
-This is clearly an advantage of the logarithmic score over the CRPS; however, the logarithmic score is known to have other severe downsides, e.g., its low robustness to occasional misguided forecasts; see \cite{bracherEvaluatingEpidemicForecasts2021} for a more detailed discussion.
+This is clearly an advantage of the logarithmic score over the CRPS; however, the logarithmic score is known to have other severe downsides, e.g., its low robustness to sporadically misguided forecasts; see \cite{bracherEvaluatingEpidemicForecasts2021} for a more detailed discussion.
 
-Exploring transformations is a promising avenue for future work that could help bridge the gap between modellers and policy makers by providing scoring rules that better reflect what forecast consumers care about.
+Exploring transformations is a promising avenue for future work that could help bridge the gap between modellers and policymakers by providing scoring rules that better reflect what forecast consumers care about.
 Potentially, the variance stabilising time-series forecasting literature may be a useful source of transformations for various forecast settings.
-We have showed that the log-transformation can lead to significant changes in the relative rankings of models against each other, with potentially important implications for decision-makers who rely on the knowledge of past performance to make a decision about which forecasts should inform future decisions.
+We have shown that the natural logarithm transformation can lead to significant changes in the relative rankings of models against each other, with potentially important implications for decision-makers who rely on the knowledge of past performance to make a judgement about which forecasts should inform future decisions.
 While it is commonly accepted that multiple proper scoring rules should usually be considered when comparing forecasts, we think this should be supplemented by considering different transformations of the data to obtain a richer picture of model performance.
-More work needs to be done to better understand the effects of applying transformations more comprehensively and in different contexts, and how it might affect decision-making.
+More work needs to be done to better understand the effects of applying transformations in different contexts, and how they may impact decision-making.
 
@@
 \toprule
-Horizon & Target & $\alpha$ & $\beta$ & $\alpha^*$ & $\beta^*$ & $\alpha^{**}$ & $\beta^{**}$\\
+Horizon & Target & $\alpha$ & $\beta$ & $\alpha_{\log}$ & $\beta_{\log}$ & $\alpha_{\sqrt{\ }}$ & $\beta_{\sqrt{\ }}$\\
 \midrule

### 2023-01-12 main.tex

@@
   Edwin van Leeuwen\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom} $^{,}$\thanks{UK Health Security Agency, London, United Kingdom} $^{}$,
-  Johannes Bracher\thanks{Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology, Karlsruhe, Germany} $^{}$, 
+  Johannes Bracher\thanks{Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology, Karlsruhe, Germany}$^{,}$\thanks{Computational Statistics Group, Heidelberg Institute for Theoretical Studies, Heidelberg, Germany} $^{}$,
   Sebastian Funk\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$}
@@
 \begin{abstract}
-Forecast evaluation plays an essential role in improving forecasting models and can support their role in informing decision-making.
+Forecast evaluation plays an essential role in the development cycle of predictive epidemic models and can inform their use for public health decision-making.
+Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which are both measures of the absolute distance between the forecast distribution and the observation.
+They are commonly applied directly to predicted and observed incidence counts, but it can be questioned whether this yields the most meaningful results given the exponential nature of epidemic processes and the several orders of magnitude that observed values can span over space and time.
+In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results.
+We motivate the procedure threefold using the CRPS on log-transformed counts as an example: firstly, it can be interpreted as a probabilistic version of a relative error.
+Secondly, it reflects how well models predicted the time-varying epidemic growth rate.
+And lastly, using arguments on variance-stabilizing transformations, it can be shown that under the assumption of a quadratic mean-variance relationship, the logarithmic transformation leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity.
+Applying the log transformation to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types.
+Situations in which models missed beginning upward swings are more strongly emphasized while failing to predict a downturn following a peak is less severely penalized.
+We conclude that appropriate transformations, of which the natural logarithm is only one particularly attractive option, should be considered when assessing the performance of different models in the context of infectious disease incidence.
+
+%Forecast evaluation plays an essential role in improving forecasting models and can support their role in informing decision-making.
 In epidemiology, probabilistic forecasts are often scored using the Weighted Interval Score (WIS) or the Continuous Ranked Probability Score (CRPS), which are measures of absolute distance between forecast and observation.
@@
 \end{abstract}
-
-% Suggestion J:
-% Forecast evaluation plays an essential role in the development cycle of predictive epidemic models and can inform their use for public health decision-making.
-Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which are both measures of the absolute distance between the forecast distribution and the observation.
-They are commonly applied directly to predicted and observed incidence counts, but it can be questioned whether this yields the most meaningful results given the exponential nature of epidemic processes and the several orders of magnitude that observed values can span over space and time.
-In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results.
-We motivate the procedure threefold using the CRPS on log-transformed counts as an example: firstly, it can be interpreted as a probabilistic version of a relative error.
-Secondly, it reflects how well models predicted the time-varying epidemic growth rate.
-And lastly, using arguments on variance-stabilizing transformations, it can be shown that under the assumption of a quadratic mean-variance relationship, the logarithmic transformation leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity.
-This implies that different forecasting tasks will have a similar impact on overall evaluation results.
-Applying the log transformation to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types.
-Situations in which models missed beginning upward swings are more strongly emphasized while failing to predict a downturn following a peak is less severely penalized.
-We conclude that appropriate transformations, of which the natural logarithm is only one particularly attractive option, should be considered when assessing the performance of different models in the context of infectious disease incidence.
 
@@
 This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellisChallengesControlCOVID192021}.
+
 If case numbers are evolving based on an exponential process and the  modelling task revolves around estimating the reproduction number or the corresponding growth rate, then evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount.
-This is because errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
+This is because for exponential processes errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
 If one is to measure the ability of forecasts to assess and forecast the underlying infection dynamics, it may thus be more desirable to evaluate errors on the growth rate directly.
@@
 \cite{bolinLocalScaleInvariance2021} have proposed the scaled CRPS (SCRPS) which is locally scale invariant; however, it does not correspond to a relative error measure and lacks a straightforward interpretation as available for the CRPS.
-% is less commonly used and does not have a straightforward equivalent for a setting with predictive quantiles.
+
 A closely related aspect to relative scores (as opposed to absolute scores) is that in the evaluation one may wish to give similar weight to all considered forecast targets.
@@
 An evaluation based on absolute errors, however, will assign little weight to other potentially important aspects, such as the ability to correctly predict future upswings while observed numbers are still low.
-
-%that do not depend on the scale of the forecast target, such as the log score \citep{goodRationalDecisions1952}, which in addition is also invariant under monotonic transformations of the target, or a scaled version of the CRPS \citep{bolinLocalScaleInvariance2021}.
-These, however, may not have all of the desired properties of the CRPS and in particular cannot easily be transferred to a setting with predictive quantiles.
-
-%In this paper we argue that these three aspects (evaluating the growth rate rather than absolute values, evaluating relative errors, giving more equal weights to all forecast targets) can all be addressed by simply applying the natural logarithm to forecasts (in our case predictive quantiles) and observations prior to computing the CRPS.
 
@@
 Other transformations may likewise be of interest. We briefly explore the square root transformation as an alternative transformation. %[and population standardization?]. 
-Our analysis focuses on the CRPS (or WIS) as an evaluation metric, given its widespread use throughout the COVID-19 pandemic.
+Our analysis mostly focuses on the CRPS (or WIS) as an evaluation metric for probabilistic forecasts, given its widespread use throughout the COVID-19 pandemic.
 
@@
 If the mean-variance relationship is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST \citep{guerreroTimeseriesAnalysisSupported1993}.
-Denoting by $F^*$ the predictive distribution for $\log(Y)$, we can use the delta method to show that
+Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method to show that
 $$
-\mathbb{E}_F[\text{CRPS}\{F^*, \log(y)\}] \approx \frac{\sigma/\mu}{\sqrt{\pi}} 
+\mathbb{E}_F[\text{CRPS}\{F_{\log}, \log(y)\}] \approx \frac{\sigma/\mu}{\sqrt{\pi}} 
 = \frac{\sqrt{c}}{\sqrt{\pi}}
@@
 If the variance is linear with $\sigma^2 = c \times \mu$, as with a Poisson-distributed variable, the square root is known to be a VST. 
-Denoting by $\tilde{F}$ the predictive distribution for $\sqrt{Y}$, the delta method can again be used to show that
+Denoting by $F_{\sqrt{\ }}$ the predictive distribution for $\sqrt{Y}$, the delta method can again be used to show that
 $$
-\mathbb{E}_F[\text{CRPS}\{\tilde{F}, \sqrt{y}\}] \approx \frac{\sigma/\sqrt{\mu}}{2\sqrt{\pi}} = \frac{\sqrt{c}}{2\sqrt{\pi}}
+\mathbb{E}_F[\text{CRPS}\{F_{\sqrt{\ }}, \sqrt{y}\}] \approx \frac{\sigma/\sqrt{\mu}}{2\sqrt{\pi}} = \frac{\sqrt{c}}{2\sqrt{\pi}}
 .
@@
 In practice, one issue with transformations, such as the log transform, is that they are not readily applicable to negative numbers or zero values, which need to be removed or otherwise handled. 
-Small discrete values can pose an additional problem as it may be difficult to reliably evaluate relative errors.
-One common approach is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm CITATION.
+One common approach to deal with zeros is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm \citep{bellegoDealingLogsZeros2022}.
 This represents a strictly monotonic transformation and therefore preserves the propriety of the resulting score.
@@
 100 or 10,000) if average scores are dominated too strongly by forecasts for small quantities.
-%Add that the dispersion component can blow up
+
+Small discrete observed values can pose an additional problem as it may be difficult to reliably evaluate relative errors.
+A related issues occurs when the predictive distribution has a large probability mass on zero (or on very small values), as this can translate into an excessively wide forecast in relative terms.
+This can be seen in Figure \ref{fig:HUB-model-comparison-baseline}, where the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm where forecasts contained zero in its prediction intervals.
 
@@
 Depending on the relationship between the mean and variance of the forecast target, a log-transformation may even lead to systematically larger scores assigned to small forecast targets, as illustrated in Figure \ref{fig:SIM-wis-state-size-mean}.
-A different transformation like the square-root may then be more appropriate to make scores independent of the mean of the target quantity (as discussed in Section \ref{sec:methods:vst}).
+
+%A different transformation like the square-root may then be more appropriate to make scores independent of the mean of the target quantity (as discussed in Section \ref{sec:methods:vst}).
 
@@
 This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluating the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}.
-This, unfortunately, is not feasible in the context of forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate, but may be an interesting approach if predictive samples are available as it enables the possibility of constructing composite scores as a weighted sum of scores based on different transformations.
-This would allow forecast consumers to assign explicit weights to different qualities of the forecasts they might care about.
+This, unfortunately, is not feasible for forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate, but may be an interesting approach if predictive samples are available.
+It is possible to go beyond choosing a single transformation by constructing composite scores as a weighted sum of scores based on different transformations.
+This would make it possible to create custom scores and allow forecast consumers to assign explicit weights to different qualities of the forecasts they might care about.
 
@@
 \caption{Coefficients of three regressions for the effect of the magnitude of the median forecast on expected scores. The first regression was 
-$\text{CRPS}(F, y) = \exp(\alpha) \cdot \text{median}(F)^{\beta},$ where $F$ is the predictive distribution and $y$ the observed value. The second one was 
-$\text{CRPS}(\log F, \log y) = \alpha^* + \beta^* \cdot \log{(\text{median}(F))}.$ The third one was $\text{CRPS}(\sqrt{F}, \sqrt{y}) = \alpha^{**} + \beta^{**} \cdot \sqrt{(\text{median}(F))}.$
+$\log[\text{CRPS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], $ where $F$ is the predictive distribution and $y$ the observed value. The second one was 
+$\text{CRPS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},$ where $F_{\log}$ is the predictive distribution for $\log y$.
+The third one was $\text{CRPS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{sqrt{\ }} + \beta_{sqrt{\ }} \cdot \sqrt{(\text{median}(F))},$ where $F_{\sqrt{\ }}$ is the predictive distribution for $\sqrt{y}$.
 }

### 2023-01-12 R/Refetch data from Forecast Hub.R
Clean code that refetches data from the Hub

@@ -1,4 +1,6 @@
-# Code to refetch data from the Forecast Hub
+## ========================================================================== ##
+##               Code to refetch data from the Forecast Hub.                  ##
+## ========================================================================== ##
 
 library(covidHubUtils)
 library(dplyr)
@@ -9,12 +11,15 @@ library(here)
 library(stringr)
 
 analysis_date <- "2022-12-12"
-  
-# download folder using SVN
+
+## -------------------------------------------------------------------------- ##  
+##                  Download and process Forecast Hub data                    ##
+## -------------------------------------------------------------------------- ##
+
+# -------------------- Download Forecast Hub folder using SVN ---------------- #
 # svn checkout https://github.com/epiforecasts/covid19-forecast-hub-europe/trunk/data-processed
 
-# load truth data using the covidHubutils package ------------------------------
-
+# ---------------load truth data using the covidHubutils package ------------- #
 if (file.exists("data/weekly-truth-Europe.csv")) {
   truth <- fread("data/weekly-truth-Europe.csv")
 } else {
@@ -29,8 +34,9 @@ if (file.exists("data/weekly-truth-Europe.csv")) {
   fwrite(truth, "data/weekly-truth-Europe.csv")
 }
 
+# ----------------------- Load and handle forecast data ---------------------- #
 
-# get the correct file paths to all forecasts ----------------------------------
+# get the correct file paths to all forecasts 
 folders <- here("data-processed", list.files("data-processed"))
 folders <- folders[
   !(grepl("\\.R", folders) | grepl(".sh", folders) | grepl(".csv", folders))
@@ -44,8 +50,7 @@ file_paths <- purrr::map(folders,
   unlist()
 file_paths <- file_paths[grepl(".csv", file_paths)]
 
-# load all past forecasts ------------------------------------------------------
-# ceate a helper function to get model name from a file path
+# ceate a helper function to obtain the model name from a given file path
 get_model_name <- function(file_path) {
   split <- str_split(file_path, pattern = "/")[[1]]
   model <- split[length(split) - 1]
@@ -85,33 +90,10 @@ hub_data <- merge_pred_and_obs(prediction_data, truth,
 hub_data <- mutate(hub_data, 
                    forecast_date = calc_submission_due_date(forecast_date))
 
[…]

### 2023-01-12 R/00 full paper code.R
split Figure into two, add regressions for sqrt

@@ -409,12 +409,14 @@ hub_data <- rbindlist(list(
   unique() |>
   filter(location != "")
 
-scores <- fread(here("output", "data", "all-scores-european-hub.csv")) |>
+scores <- fread(here("output", "data", "all-scores-european-hub.csv")) 
+
+scores <- scores |>
   mutate(scale = factor(scale, levels = c("natural", "log", "sqrt"))) |>
   mutate(type_and_scale = factor(type_and_scale, 
                                  levels = c("Cases - natural", "Deaths - natural", 
-                                            "Cases - log", "Deaths - log"))) |>
-  filter(horizon <= 4)
+                                            "Cases - log", "Deaths - log", 
+                                            "Cases - sqrt", "Deaths - sqrt")))
 
 
 # dates and number of locations
@@ -554,9 +556,8 @@ box_plot_horizon <- scores |>
   theme_scoringutils() + 
   theme(legend.position = "none") + 
   scale_y_continuous(labels = label_fn, trans = "log10") +
-  labs(y = "Rel. change in WIS", x = "Forecast horizon (weeks)") 
-# + 
-#   coord_flip()
+  labs(y = "Rel. change in WIS", x = "Forecast horizon (weeks)") + 
+  coord_cartesian(ylim = c(0.1, 10))
 
 layout <- "
 AB
@@ -628,27 +629,27 @@ regs_coef <- rbind(
   regression_df(scores, s = "natural", horizons = "2", targets = "Cases"),  
   regression_df(scores, s = "natural", horizons = "2", targets = "Deaths"), 
   regression_df(scores, s = "log", horizons = "2", targets = "Cases"), 
-  regression_df(scores, s = "log", horizons = "2", targets = "Deaths")
+  regression_df(scores, s = "log", horizons = "2", targets = "Deaths"),
+  regression_df(scores, s = "sqrt", horizons = "2", targets = "Cases"), 
+  regression_df(scores, s = "sqrt", horizons = "2", targets = "Deaths")
 ) |>
   mutate(type_and_scale = paste(target_type, scale, sep = " - ")) |>
   mutate(type_and_scale = factor(type_and_scale, 
                                  levels = c("Cases - natural", "Deaths - natural", 
-                                            "Cases - log", "Deaths - log")))
+                                            "Cases - log", "Deaths - log", 
[…]

### 2023-01-12 main.tex
Updates from Overleaf

@@
 \subsection{Setting}
+\label{sec:HUB-setting}
 
@@
 For any given forecast date and location, forecasts were made for four different forecast horizons, resulting in four scores.
-All scores were divided by the score for forecast horizon one.}
+All scores were divided by the score for forecast horizon one.
+To enhance interpretability, the range of visible relative changes was restricted to [0.1, 10].}
     \label{fig:HUB-mean-locations}
@@
 
+\begin{table}
+\centering
+\begin{tabular}{ccrrrrrr}
+\toprule
+Horizon & Target & $\alpha$ & $\beta$ & $\alpha_{\log}$ & $\beta_{\log}$ & $\alpha_{\sqrt{\ }}$ & $\beta_{\sqrt{\ }}$\\
+\midrule
+all & all & -1.093 & 0.963 & 0.390 & 0.001 & -0.995 & 0.209\\
+\addlinespace
+all & Cases & 0.036 & 0.858 & 0.751 & -0.033 & 0.047 & 0.201\\
+all & Deaths & -0.885 & 0.868 & 0.430 & -0.021 & -7.564 & 0.826\\
+\addlinespace
+1 & all & -1.402 & 0.923 & 0.318 & -0.014 & 0.318 & 0.088\\
+2 & all & -1.221 & 0.967 & 0.364 & -0.003 & 0.111 & 0.164\\
+3 & all & -1.001 & 0.984 & 0.408 & 0.008 & -1.283 & 0.255\\
+4 & all & -0.757 & 0.986 & 0.467 & 0.016 & -1.041 & 0.312\\
+\addlinespace
+1 & Cases & -0.861 & 0.876 & 0.433 & -0.024 & 0.785 & 0.087\\
+2 & Cases & -0.243 & 0.877 & 0.660 & -0.031 & 0.959 & 0.162\\
+3 & Cases & 0.372 & 0.855 & 0.882 & -0.037 & 1.115 & 0.238\\
+4 & Cases & 0.816 & 0.837 & 1.009 & -0.036 & 1.665 & 0.296\\
+\addlinespace
+1 & Deaths & -1.145 & 0.831 & 0.376 & -0.035 & 0.458 & 0.048\\
+2 & Deaths & -0.980 & 0.866 & 0.415 & -0.027 & 0.445 & 0.084\\
+3 & Deaths & -0.809 & 0.885 & 0.441 & -0.016 & -8.262 & 0.903\\
+4 & Deaths & -0.603 & 0.891 & 0.488 & -0.007 & -7.834 & 0.898\\
+\bottomrule
+\end{tabular}
+\caption{Coefficients of three regressions for the effect of the magnitude of the median forecast on expected scores. The first regression was 
+$\log[\text{CRPS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], $ where $F$ is the predictive distribution and $y$ the observed value. The second one was 
+$\text{CRPS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},$ where $F_{\log}$ is the predictive distribution for $\log y$.
+The third one was $\text{CRPS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{sqrt{\ }} + \beta_{sqrt{\ }} \cdot \sqrt{(\text{median}(F))},$ where $F_{\sqrt{\ }}$ is the predictive distribution for $\sqrt{y}$.
+}
+\label{tab:HUB-regression}
+\end{table}
+
 As argued in Section \ref{sec:methods:vst}, the mean-variance, or mean-CRPS, relationship determines which transformation can serve as a VST. To assess this empirically, we ran the regression
@@
 The results of the three regressions are also displayed in Figure \ref{fig:HUB-cors}.
-It appears that the log transformation is most suitable to stabilise the CRPS values in this empirical example.
-
-% We found that on the natural scale, the order of magnitude of the forecast (represented by the median, rather than the mean, as forecasts were stored as quantiles and the mean was not available) strongly influenced scores, with estimated $\beta$-values close to one.
-On the log scale, this influence was small, with estimated $\beta^*$-values close to zero.
-Both on the natural and on the log scale we found scores to increase over time (estimated $\alpha$- and $\alpha^*$-values increased over horizons).
-Estimates are summarised in Table \ref{tab:HUB-regression}.
+It appears that the log transformation is most suitable to stabilise the WIS values in this empirical example.
 
@@
     \caption{
-    Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-ensemble made in Germany.
+    Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-ensemble made in Germany. Missing values are due to data anomalies that were removed (see section \ref{sec:HUB-setting}. 
     A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale.
@@
 C: Relative skill scores based on scores on the log scale.
-D: Difference in relative skill scores computed on the natural and on the logarithmic scale.
+D: Difference in relative skill scores computed on the natural and on the logarithmic scale, ordered as in C.
 E: Relative contributions of the different WIS components (overprediction, underprediction, and dispersion) to overall model scores on the natural and the logarithmic scale.
@@
 
-\begin{table}
-\centering
-\begin{tabular}{ccrrrrrr}
-\toprule
-Horizon & Target & $\alpha$ & $\beta$ & $\alpha_{\log}$ & $\beta_{\log}$ & $\alpha_{\sqrt{\ }}$ & $\beta_{\sqrt{\ }}$\\
-\midrule
-all & all & -1.093 & 0.963 & 0.390 & 0.001 & -0.995 & 0.209\\
-\addlinespace
-all & Cases & 0.036 & 0.858 & 0.751 & -0.033 & 0.047 & 0.201\\
-all & Deaths & -0.885 & 0.868 & 0.430 & -0.021 & -7.564 & 0.826\\
-\addlinespace
-1 & all & -1.402 & 0.923 & 0.318 & -0.014 & 0.318 & 0.088\\
-2 & all & -1.221 & 0.967 & 0.364 & -0.003 & 0.111 & 0.164\\
-3 & all & -1.001 & 0.984 & 0.408 & 0.008 & -1.283 & 0.255\\
-4 & all & -0.757 & 0.986 & 0.467 & 0.016 & -1.041 & 0.312\\
-\addlinespace
-1 & Cases & -0.861 & 0.876 & 0.433 & -0.024 & 0.785 & 0.087\\
-2 & Cases & -0.243 & 0.877 & 0.660 & -0.031 & 0.959 & 0.162\\
-3 & Cases & 0.372 & 0.855 & 0.882 & -0.037 & 1.115 & 0.238\\
-4 & Cases & 0.816 & 0.837 & 1.009 & -0.036 & 1.665 & 0.296\\
-\addlinespace
-1 & Deaths & -1.145 & 0.831 & 0.376 & -0.035 & 0.458 & 0.048\\
-2 & Deaths & -0.980 & 0.866 & 0.415 & -0.027 & 0.445 & 0.084\\
-3 & Deaths & -0.809 & 0.885 & 0.441 & -0.016 & -8.262 & 0.903\\
-4 & Deaths & -0.603 & 0.891 & 0.488 & -0.007 & -7.834 & 0.898\\
-\bottomrule
-\end{tabular}
-\caption{Coefficients of three regressions for the effect of the magnitude of the median forecast on expected scores. The first regression was 
-$\log[\text{CRPS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], $ where $F$ is the predictive distribution and $y$ the observed value. The second one was 
-$\text{CRPS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},$ where $F_{\log}$ is the predictive distribution for $\log y$.
-The third one was $\text{CRPS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{sqrt{\ }} + \beta_{sqrt{\ }} \cdot \sqrt{(\text{median}(F))},$ where $F_{\sqrt{\ }}$ is the predictive distribution for $\sqrt{y}$.
-}
-\label{tab:HUB-regression}
-\end{table}
-
 \clearpage

### 2023-01-12 R/00 full paper code.R
update plots

@@ -644,11 +644,11 @@ regs_coef <- regs_coef |>
          beta = signif(beta, 2)) |>
   mutate(label = ifelse(
     scale == "natural",
-    paste0("\\hat{WIS} = e^", alpha, " \\cdot x^", beta), # regression for natural
+    paste0("\\hat{WIS} = e^", alpha, " \\cdot median^", beta), # regression for natural
     
     ifelse(scale == "log", 
-           paste0("\\hat{WIS} = ", alpha, " + ", beta, "\\cdot \\log (x)"), # regression for log
-           paste0("\\hat{WIS} = ", alpha, " + ", beta, "\\cdot \\sqrt{x}")) # regression for sqrt
+           paste0("\\hat{WIS} = ", alpha, " - ", -beta, "\\cdot \\log (median)"), # regression for log
+           paste0("\\hat{WIS} = ", alpha, " + ", beta, "\\cdot \\sqrt{median}")) # regression for sqrt
     ))
 
 regs <- regs_coef |> 

### 2023-01-12 main.tex

@@
 
-The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 12th of May 2022 \textbf{CHECK AGAIN} for reported cases and deaths from COVID-19.
-We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 16 CHECK AGAIN forecast dates (see Figure \ref{fig:HUB-num-avail-models}).
+The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 12th of May 2022 for reported cases and deaths from COVID-19.
+We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 46 forecast dates (see Figure \ref{fig:HUB-num-avail-models}).
 We removed all observations marked as data anomalies by the European Forecast Hub \citep{sherrattPredictivePerformanceMultimodel2022} as well as all remaining negative observed values.
@@
     \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/HUB-transformation-regression.png}
+    \caption{Relationship between median forecasts and scores.
+Black dots represent WIS values for two-week ahead predictions of the EuroCOVIDhub-ensemble.
+Shown in red are the regression lines discussed in \ref{sec:regression} and \ref{tab:HUB-regression}.
+A: WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against median predicted values.
+B: Same as A, with scores obtained after applying a log-transformation to the data.
+C: Same as A, with scores obtained after applying a square-root-transformation to the data.}
+    \label{fig:HUB-regression}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
     \includegraphics[width=0.99\textwidth]{output/figures/HUB-correlations.png}
-    \caption{Correlations for median forecasts, scores, and rankings on the natural and logarithmic scale.
-A: WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against median predicted values.
-B: Average Spearman rank correlation of scores for individual forecasts from all models.
+    \caption{Correlations of rankings on the natural and logarithmic scale.
+A: Average Spearman rank correlation of scores for individual forecasts from all models.
 For every individual target (defined by a combination of forecast date, target type, horizon, location), one score was obtained per model.
@@
 All rank correlations were averaged across locations, and target types and stratified by horizon and target type.
-C: Correlation between relative skill scores.
+B: Correlation between relative skill scores.
 For every forecast horizon and target type, a separate rel.
@@
 4 & Deaths & -0.603 & 0.891 & 0.488 & -0.007 & -7.834 & 0.898\\
+\bottomrule
 \bottomrule
@@
 As can be seen from Table \ref{tab:HUB-regression}, the results indeed indicate small targets lead to larger average CRPS when using the log transform ($\beta_{\log} < 0$), while the opposite is true for the square-root transform ($\beta_{\sqrt{\ }} > 0$).
-The results of the three regressions are also displayed in Figure \ref{fig:HUB-cors}.
+The results of the three regressions are also displayed in Figure \ref{fig:HUB-regression}.
 It appears that the log transformation is most suitable to stabilise the WIS values in this empirical example.
@@
 
-For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}B).
+For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}A).
 When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models change considerably.
-The change is stronger for cases than for deaths (Figure \ref{fig:HUB-cors}C) and possibly tends to increase with increasing forecast horizon.
+The change is stronger for cases than for deaths (Figure \ref{fig:HUB-cors}B) and possibly tends to increase with increasing forecast horizon.
 
@@
 On the log scale, these have less of an influence, as errors are smaller in relative terms (see \ref{fig:HUB-model-comparison-ensemble}C, D).
-We hypothesise that the lower correlation between skill scores for cases (compared to the correlation of scores for deaths) as displayed in Figure \ref{fig:HUB-cors}C is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths, and therefore relative skill scores are more strongly affected for cases.
+We hypothesise that the lower correlation between skill scores for cases (compared to the correlation of scores for deaths) as displayed in Figure \ref{fig:HUB-cors}B is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths, and therefore relative skill scores are more strongly affected for cases.
 Conversely, failure to predict an upswing while numbers are still low, is less severely punished on the natural scale (see forecasts in July in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low.
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width = 1\textwidth]{output/figures/different-relative-errors-sqrt.png}
+\caption{Numerical comparison of different measures of relative error: absolute percentage error (APE), relative error (RE), symmetric absolute percentage error (SAPE) and the absolute error applied to \textit{sqrt}-transformed predictions and observations.
+We denote the predicted value by $\hat{y}$ and display errors as a function of the ratio of observed and predicted value.
+A: x-axis shown on a linear scale.
+B: x-axis shown on a logarithmic scale.}
+\label{fig:SAPE-sqrt}
+\end{figure}
+
 \begin{table}[h!]
     \centering
-
 \begin{tabular}{lllcc}

### 2023-01-12 R/00 full paper code.R
update figures

@@ -184,7 +184,7 @@ p1 / p2 +
   plot_annotation(tag_levels = "A") +   
   plot_layout(guides = "collect") & 
   theme(legend.position = "bottom") &
-  labs(y = TeX("Mean CRPS (normalised)"))
+  labs(y = TeX("CRPS (normalised)"))
 
 ggsave("output/figures/SIM-mean-state-size.png", width = 7, height = 4.1)
 
@@ -707,7 +707,7 @@ p_natural / p_log / p_sqrt +
   plot_annotation(tag_levels = "A") &
   theme(legend.position = "bottom")
 
-ggsave("output/figures/HUB-transformation-regression.png", width = 7, height = 7)
+ggsave("output/figures/HUB-transformation-regression.png", width = 7, height = 6.3)
 
 
 
@@ -765,7 +765,7 @@ p_cor_scores + p_cor_skill +
   plot_annotation(tag_levels = "A") &
   theme(legend.position = "bottom")
 
-ggsave("output/figures/HUB-correlations.png", width = 7, height = 3)
+ggsave("output/figures/HUB-correlations.png", width = 7, height = 2.6)
 
 
 ## ========================================================================== ##
@@ -932,6 +932,7 @@ simple_labels <- function(x) {
 
 ranking_figure <- function(target = "Cases") {
   summarised_pairwise <- scores |>
+    filter(scale %in% c("log", "natural")) |>
     select(-median_prediction) |>
     filter(target_type == target, 
            horizon == 2) |>
@@ -996,18 +997,23 @@ ranking_figure <- function(target = "Cases") {
     scale_x_continuous(labels = label_fn) +
     facet_wrap(~scale, scales = "free_x")
   
-  plot_ranking_change <- summarised_pairwise |>
+  
+  df_ranking_change <- summarised_pairwise |>
     filter(scale %in% c("natural", "log")) |>
     select(model, scale, relative_skill) |>
     group_by(scale) |>
-    # mutate(scale = factor(scale, levels = c("log", "natural"))) |>
+    mutate(scale = factor(scale, levels = c("natural", "log"))) |>
     mutate(rank = rank(-relative_skill), 
            x = ifelse(scale == "natural", 0.4, 2.8), 
            x_arrow = ifelse(scale == "natural", 1.7, 2.7)
     ) |>
     group_by(model) |>
-    mutate(custom_color = ifelse(diff(relative_skill) < 0, "deteriorated", "not deteriorated")) |>
-    arrange(desc(scale)) |>
[…]

### 2023-01-12 main.tex
Updates from Overleaf

@@
 
-% which one is an empirical question - maybe hint that quadratic may result from RW on growth rate?
-
 To strengthen our intuition on how transforming outcomes prior to applying the CRPS shifts the emphasis between targets with high and low expected outcomes, Figure \ref{fig:SIM-wis-state-size-mean} shows the expected CRPS of ideal forecasters under different mean-variance relationships and transformations.
-We consider a Poisson distribution where $\sigma^2 = \mu$, a geometric distribution where $\sigma^2 = \mu + \mu^2$, and a negative binomial distribution where the size parameter $\psi(\mu)$ depends on the mean $\mu$ such that $\sigma^2 = \mu + \mu^{2.5}/1000$.
-We see that when applying the CRPS on the natural scale, the expected CRPS grows in $\mu$ in all cases.
-When applying it after a log-transformation, the expected CRPS is almost independent of $\mu$ for the geometric distribution and large $\mu$, while smaller targets have higher expected CRPS in the Poisson case.
-When applying a square-root-transformation before the CRPS, the expected CRPS is independent of the mean for the Poisson-distribution and grows for the two other distributions.
-%For the negative binomial distribution, targets with very small and very high expectation yield the highest expected CRPS.
-\textbf{Here it would be nice to also add a sentence on what happens under sqrt transformation.} % scores are independent of the mean of the forecast target.
-For a Poisson-distribution with
-% $\sigma = \mu$, scores on the log scale decrease as the mean increases.
-The appropriate variance-stabilising transformation here to make scores independent of the mean would be a square-root-transformation .
+We consider a Poisson distribution where $\sigma^2 = \mu$, a negative binomial distribution with $\theta = 10$ and $\sigma^2 = \mu + 0.1\mu^2$, and a normal distribution with constant variance.
+We see that when applying the CRPS on the natural scale, the expected CRPS grows with the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster).
+The expected CRPS is constant only for the distribution with constant variance, and grows in $\mu$ for the other two.
+When applying a log-transformation first, the expected CRPS is almost independent of $\mu$ for the negative binomial distribution and large $\mu$, while smaller targets have higher expected CRPS in case of the Poisson distribution and the normal distribution with constant variance.
+When applying a square-root-transformation before the CRPS, the expected CRPS is independent of the mean for the Poisson-distribution, but not for the other two.
+As can be seen in Figures \ref{fig:SIM-wis-state-size-mean} and \ref{fig:score-approx}, the approximations presented above work quite well for our simulated example.
 
@@
     \caption{Expected CRPS scores as a function of the mean and variance of the forecast quantity.
-We generated 10,000 samples for three distributions with differing relationships between mean and variance.
-We then computed expected CRPS values assuming an ideal forecaster with predictive distribution equal to the data-generating distribution.
-The first distribution (red) is a truncated normal distribution with constant variance (we chose a $\sigma = 1$ in order to only obtain positive samples).
-The second (green) is a geometric distribution (a negative binomial distribution with variance $\sigma^2 = \mu + \mu^2$.
+We computed expected CRPS values  for three different distributions, assuming an ideal forecaster with predictive distribution equal to the data-generating distribution.
+    These expected CRPS values where computed for different predictive means based on 10,000 samples each and are represented by dots.
+Solid lines show the corresponding approximation of the expected CRPS based on an assumed normal distribution as discussed in section \ref{sec:methods:vst}.
+Figure \ref{fig:XXX} shows the quality of the approximation in more detail.
+    The first distribution (red) is a truncated normal distribution with constant variance (we chose $\sigma = 1$ in order to only obtain positive samples).
+The second (green) is a negative binomial distribution with variance $\theta = 10$ and variance $\sigma^2 = \mu + 0.1\mu^2$.
 The third (blue) is a
-    Poisson distribution with $\sigma^2 = \mu + \mu$.
+    Poisson distribution with $\sigma^2 = \mu$.
 To make the scores for the different distributions comparable, scores were normalised to one, meaning that the mean score for every distribution (red, green, blue) is one.
-    A: Normalised expected CRPS for ideal forecasts with increasing means for three distribution with different relationships between mean and variance. B: A but with x axis on the log scale.}
+    A: Normalised expected CRPS for ideal forecasts with increasing means for three distribution with different relationships between mean and variance.
+Expected CRPS was computed on the natural scale (middle), after adding one and applying a log-transformation to the data (middle) and after applying a square-root transformation (left).
+B: A but with x axis on the log scale.}
     \label{fig:SIM-wis-state-size-mean}
@@
 The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 12th of May 2022 for reported cases and deaths from COVID-19.
+See \cite{sherrattPredictivePerformanceMultimodel2022} for a more thorough description of the data.
 We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 46 forecast dates (see Figure \ref{fig:HUB-num-avail-models}).
 We removed all observations marked as data anomalies by the European Forecast Hub \citep{sherrattPredictivePerformanceMultimodel2022} as well as all remaining negative observed values.
-In addition, we filtered out erroneous forecasts characterised by a median forecast of 0 and an actual observed value of larger than 150.
+In addition, we filtered out erroneous forecasts defined by any of the conditions listed in Table \ref{tab:erroneous}.
+Those forecasts were removed in order to be better able to illustrate the effects of the log-transformation on scores and eliminating distortions caused by outlier forecasters.
 All predictive quantiles were truncated at 0.
@@
 \midrule
-all & all & -1.093 & 0.963 & 0.390 & 0.001 & -0.995 & 0.209\\
+all & all & -1.093 & 0.963 & 0.391 & 0.001 & -0.352 & 0.201\\
 \addlinespace
-all & Cases & 0.036 & 0.858 & 0.751 & -0.033 & 0.047 & 0.201\\
-all & Deaths & -0.885 & 0.868 & 0.430 & -0.021 & -7.564 & 0.826\\
+all & Cases & 0.036 & 0.858 & 0.751 & -0.033 & 0.043 & 0.201\\
+all & Deaths & -0.884 & 0.868 & 0.436 & -0.023 & 0.273 & 0.121\\
 \addlinespace
-1 & all & -1.402 & 0.923 & 0.318 & -0.014 & 0.318 & 0.088\\
-2 & all & -1.221 & 0.967 & 0.364 & -0.003 & 0.111 & 0.164\\
-3 & all & -1.001 & 0.984 & 0.408 & 0.008 & -1.283 & 0.255\\
-4 & all & -0.757 & 0.986 & 0.467 & 0.016 & -1.041 & 0.312\\
+1 & all & -1.402 & 0.923 & 0.318 & -0.014 & 0.320 & 0.088\\
+2 & all & -1.221 & 0.967 & 0.364 & -0.003 & 0.112 & 0.164\\
+3 & all & -1.001 & 0.984 & 0.410 & 0.008 & -0.094 & 0.241\\
+4 & all & -0.757 & 0.986 & 0.469 & 0.015 & 0.000 & 0.299\\
 \addlinespace
-1 & Cases & -0.861 & 0.876 & 0.433 & -0.024 & 0.785 & 0.087\\
+1 & Cases & -0.862 & 0.876 & 0.433 & -0.024 & 0.790 & 0.087\\
 2 & Cases & -0.243 & 0.877 & 0.660 & -0.031 & 0.959 & 0.162\\
-3 & Cases & 0.372 & 0.855 & 0.882 & -0.037 & 1.115 & 0.238\\
-4 & Cases & 0.816 & 0.837 & 1.009 & -0.036 & 1.665 & 0.296\\
+3 & Cases & 0.372 & 0.855 & 0.882 & -0.037 & 1.109 & 0.238\\
+4 & Cases & 0.816 & 0.837 & 1.009 & -0.036 & 1.645 & 0.296\\
 \addlinespace
-1 & Deaths & -1.145 & 0.831 & 0.376 & -0.035 & 0.458 & 0.048\\
-2 & Deaths & -0.980 & 0.866 & 0.415 & -0.027 & 0.445 & 0.084\\
-3 & Deaths & -0.809 & 0.885 & 0.441 & -0.016 & -8.262 & 0.903\\
-4 & Deaths & -0.603 & 0.891 & 0.488 & -0.007 & -7.834 & 0.898\\
-\bottomrule
+1 & Deaths & -1.146 & 0.832 & 0.376 & -0.035 & 0.457 & 0.048\\
+2 & Deaths & -0.981 & 0.867 & 0.416 & -0.028 & 0.443 & 0.084\\
+3 & Deaths & -0.807 & 0.885 & 0.453 & -0.019 & 0.349 & 0.131\\
+4 & Deaths & -0.602 & 0.891 & 0.501 & -0.011 & 0.125 & 0.194\\
 \bottomrule
@@
 We note that for the logarithmic score, which has also been used e.g., in some editions of the FluSight challenge \cite{reichCollaborativeMultiyearMultimodel2019}, the question of the right scale to evaluate forecasts does not arise.
-It is known that log score differences between different forecasters are invariant to monotonic transformations of the outcome variable (see e.g., \citealt{Diks2011}).
+It is known that log score differences between different forecasters are invariant to monotonic transformations of the outcome variable (see e.g., \citealt{diksLikelihoodbasedScoringRules2011}).
 This is clearly an advantage of the logarithmic score over the CRPS; however, the logarithmic score is known to have other severe downsides, e.g., its low robustness to sporadically misguided forecasts; see \cite{bracherEvaluatingEpidemicForecasts2021} for a more detailed discussion.
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width = 1\textwidth]{output/figures/SIM-score-approximation.png}
+\caption{Comparison of expected CRPS values of an ideal forecaster and approximated values.
+Expected scores for three distributions as described in Section \ref{sec:methods:vst} and Figure \ref{fig:SIM-wis-state-size-mean} and approximated as described in \ref{sec:methods:vst}}
+\label{fig:score-approx}
+\end{figure}
+
 \begin{table}[h!]
@@
 \end{figure}
+
+\begin{table}
+\centering
+\begin{tabular}{ccc}
+\toprule
+True value & \& & Median prediction\\
+\midrule
+$>0$ & \ & $>100\times$ true value\\
+$>10$ & \  & $>20\times$ true value\\
+$>50$ & \  & $<1/50\times$ true value\\
+$= 0$ & \  & $>100$\\
+\bottomrule
+\end{tabular}
+\caption{Criteria for removing forecasts.
+Any forecast that met one of the listed criteria (represented by a row in the table), was removed.
+Those forecasts were removed in order to be better able to illustrate the effects of the log-transformation on scores and eliminating distortions caused by outlier forecasters.
+When evaluating models against each other (rather than illustrating the effect of a transformation), one would prefer not to condition on the outcome when deciding whether a forecast should be taken into account.
+}
+\label{tab:erroneous}
+\end{table}
 

### 2023-01-12 R/00 full paper code.R
catch bug associated with computation of relative skill scores for correlation figure

@@ -741,6 +741,8 @@ p_cor_scores <- df |>
 
 # Correlation between pairwise comparisons
 correlation_rel_skill <- scores |>
+  filter(scale %in% c("log", "natural")) |>
+  select(model, location, forecast_date, horizon, target_type, scale, interval_score) |>
   summarise_scores(
     by = c("model", "scale", "horizon", "target_type"), 
     relative_skill = TRUE) |>
@@ -933,7 +935,7 @@ simple_labels <- function(x) {
 ranking_figure <- function(target = "Cases") {
   summarised_pairwise <- scores |>
     filter(scale %in% c("log", "natural")) |>
-    select(-median_prediction) |>
+    select(model, location, forecast_date, horizon, target_type, scale, interval_score) |>
     filter(target_type == target, 
            horizon == 2) |>
     summarise_scores(by = c("model", "scale"), relative_skill = TRUE)

### 2023-01-12 main.tex
Updates from Overleaf

@@
 
-The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 12th of May 2022 for reported cases and deaths from COVID-19.
+The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 5th of December 2022 for reported cases and deaths from COVID-19.
 See \cite{sherrattPredictivePerformanceMultimodel2022} for a more thorough description of the data.
@@
 
-Show and discuss Fig 7 here? CHECK AGAIN
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/HUB-model-comparison-ensemble.png}
+    \caption{
+    Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-ensemble made in Germany. Missing values are due to data anomalies that were removed (see section \ref{sec:HUB-setting}. 
+    A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale.
+B, F: Corresponding scores.
+C, G: Forecasts and observations on the log scale.
+D, H: Corresponding scores.
+    }
+    \label{fig:HUB-model-comparison-ensemble}
+\end{figure}
 
-\subsection{Descriptive analysis of observed values and scores obtained on the natural and logarithmic scales CHECK AGAIN}
+When comparing examples of forecasts on the natural scale with those on the log scale (see Figures \ref{fig:HUB-model-comparison-ensemble}, \ref{fig:HUB-model-comparison-baseline}, \ref{fig:HUB-model-comparison-epinow}) a few interesting patterns emerge.
+Missing the peak, i.e.
+predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts in May in Figure \ref{fig:HUB-model-comparison-ensemble}A, B).
+On the log scale, these have less of an influence, as errors are smaller in relative terms (see \ref{fig:HUB-model-comparison-ensemble}C, D).
+Conversely, failure to predict an upswing while numbers are still low, is less severely punished on the natural scale (see forecasts in July in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low.
+On the log scale, missing lower inflection points tends to lead to more severe penalties (see Figure \ref{fig:HUB-model-comparison-ensemble}C, D)).
+One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure \ref{fig:HUB-model-comparison-ensemble}E, F).
+On the log scale, scores do not exhibit this behaviour and rather increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.
 
@@
 
+\subsection{Regression analysis to determine the variance-stabilizing transformation}
+\label{sec:HUB-regression}
+
+As argued in Section \ref{sec:methods:vst}, the mean-variance, or mean-CRPS, relationship determines which transformation can serve as a VST. To assess this empirically, we ran the regression
+
+\begin{equation*}
+    \log[\text{CRPS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], 
+\end{equation*}
+where the predictive distribution $F$ and the observation $y$ are on the natural scale. This is equivalent to
+\begin{equation*}
+    \text{CRPS}(F, y) = \exp(\alpha) \times \text{median}(F)^\beta, 
+\end{equation*}
+meaning that we estimate a polynomial relationship between the predictive median and achieved CRPS.
+Note that we are using predictive medians rather than means as only the former are available in the European COVID-19 Forecast Hub.
+We applied the regression to case and death forecasts, pooled across horizons and stratified for one through four-week-ahead forecasts.
+Results are provided in Table \ref{tab:HUB-regression}.
+It can be seen that the estimates of $\beta$ always take a value below 1.
+As the CRPS scales with the standard deviation (see Section \ref{sec:methods:vst}), this can be interpreted as a slightly sub-quadratic mean-variance relationship.
+The logarithmic transformation should thus approximately stabilize the variance (and CRPS), possibly leading to somewhat higher scores for smaller forecast targets.
+The square-root transformation, on the other hand, can be expected to still lead to higher CRPS values for targets of higher orders of magnitude.
+
 \begin{figure}[h!]
@@
 Black dots represent WIS values for two-week ahead predictions of the EuroCOVIDhub-ensemble.
-Shown in red are the regression lines discussed in \ref{sec:regression} and \ref{tab:HUB-regression}.
+Shown in red are the regression lines discussed in Section \ref{sec:HUB-regression} shown in Table \ref{tab:HUB-regression}.
 A: WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against median predicted values.
@@
 \end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-correlations.png}
-    \caption{Correlations of rankings on the natural and logarithmic scale.
-A: Average Spearman rank correlation of scores for individual forecasts from all models.
-For every individual target (defined by a combination of forecast date, target type, horizon, location), one score was obtained per model.
-Then, the rank correlation was computed between the scores for all models on the natural scale vs.
-on the log scale.
-All rank correlations were averaged across locations, and target types and stratified by horizon and target type.
-B: Correlation between relative skill scores.
-For every forecast horizon and target type, a separate rel.
-skill score was computed per model using pairwise comparisons.
-The plot shows the correlation between the rel.
-skill scores on the natural vs.
-on the log scale.}
-    \label{fig:HUB-cors}
-\end{figure}
-
-\subsection{Regression analysis to determine the variance-stabilizing transformation}
 
@@
 
-As argued in Section \ref{sec:methods:vst}, the mean-variance, or mean-CRPS, relationship determines which transformation can serve as a VST. To assess this empirically, we ran the regression
-
-\begin{equation*}
-    \log[\text{CRPS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], 
-\end{equation*}
-where the predictive distribution $F$ and the observation $y$ are on the natural scale. This is equivalent to
-\begin{equation*}
-    \text{CRPS}(F, y) = \exp(\alpha) \times \text{median}(F)^\beta, 
-\end{equation*}
-meaning that we estimate a polynomial relationship between the predictive median and achieved CRPS.
-Note that we are using predictive medians rather than means as only the former are available in the European COVID-19 Forecast Hub.
-We applied the regression to case and death forecasts, pooled across horizons and stratified for one through four-week-ahead forecasts.
-Results are provided in Table \ref{tab:HUB-regression}.
-It can be seen that the estimates of $\beta$ always take a value below 1.
-As the CRPS scales with the standard deviation (see Section \ref{sec:methods:vst}), this can be interpreted as a slightly sub-quadratic mean-variance relationship.
-The logarithmic transformation should thus approximately stabilize the variance (and CRPS), possibly leading to somewhat higher scores for smaller forecast targets.
-The square-root transformation, on the other hand, can be expected to still lead to higher CRPS values for targets of higher orders of magnitude.
-
 To check this, we ran the regressions
@@
 The results of the three regressions are also displayed in Figure \ref{fig:HUB-regression}.
-It appears that the log transformation is most suitable to stabilise the WIS values in this empirical example.
+In this empirical example, the log transformation helps (albeit not perfectly), to stabilise WIS values and achieves this better suited than the square-root transformation.
+
+\subsection{Impact of logarithmic transformation on model rankings}
 
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-model-comparison-ensemble.png}
-    \caption{
-    Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-ensemble made in Germany. Missing values are due to data anomalies that were removed (see section \ref{sec:HUB-setting}. 
-    A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale.
-B, F: Corresponding scores.
-C, G: Forecasts and observations on the log scale.
-D, H: Corresponding scores.
-    }
-    \label{fig:HUB-model-comparison-ensemble}
+    \includegraphics[width=0.99\textwidth]{output/figures/HUB-correlations.png}
+    \caption{Correlations of rankings on the natural and logarithmic scale.
+A: Average Spearman rank correlation of scores for individual forecasts from all models.
+For every individual target (defined by a combination of forecast date, target type, horizon, location), one score was obtained per model.
+Then, the rank correlation was computed between the scores for all models on the natural scale vs.
+on the log scale.
+All rank correlations were averaged across locations, and target types and stratified by horizon and target type.
+B: Correlation between relative skill scores.
+For every forecast horizon and target type, a separate rel.
+skill score was computed per model using pairwise comparisons.
+The plot shows the correlation between the rel.
+skill scores on the natural vs.
+on the log scale.}
+    \label{fig:HUB-cors}
 \end{figure}
-
-\subsection{Impact of logarithmic transformation on model rankings}
 
@@
 When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models change considerably.
-The change is stronger for cases than for deaths (Figure \ref{fig:HUB-cors}B) and possibly tends to increase with increasing forecast horizon.
-
-When comparing examples of forecasts on the natural scale with those on the log scale (see Figures \ref{fig:HUB-model-comparison-ensemble}, \ref{fig:HUB-model-comparison-baseline}, \ref{fig:HUB-model-comparison-epinow}) a few interesting patterns emerge.
-Missing the peak, i.e.
-predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts in May in Figure \ref{fig:HUB-model-comparison-ensemble}A, B).
-On the log scale, these have less of an influence, as errors are smaller in relative terms (see \ref{fig:HUB-model-comparison-ensemble}C, D).
-We hypothesise that the lower correlation between skill scores for cases (compared to the correlation of scores for deaths) as displayed in Figure \ref{fig:HUB-cors}B is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths, and therefore relative skill scores are more strongly affected for cases.
-Conversely, failure to predict an upswing while numbers are still low, is less severely punished on the natural scale (see forecasts in July in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low.
-On the log scale, missing lower inflection points tends to lead to more severe penalties (see Figure \ref{fig:HUB-model-comparison-ensemble}C, D)).
-One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure \ref{fig:HUB-model-comparison-ensemble}E, F).
-On the log scale, scores do not exhibit this behaviour and rather increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.
+The change is stronger for cases than for deaths (Figure \ref{fig:HUB-cors}B) and increases noticeably with increasing forecast horizon.
 
@@
 Encouragingly for the European Forecast Hub, the Hub ensemble, which is the forecast the organisers suggest forecast consumers make use of, remains one of the top forecasts across scoring schemes and indeed ranks higher for reported cases when a log-transform is used.
+
+We hypothesise that the lower correlation between skill scores for cases (compared to the correlation of scores for deaths) as displayed in Figure \ref{fig:HUB-cors}B is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths, and therefore relative skill scores are more strongly affected for cases.
 

### 2023-01-13 R/00 full paper code.R
Restructure paper code, update box plot

@@ -20,8 +20,7 @@ library(covidHubUtils) # remotes::install_github("https://github.com/reichlab/co
 set.seed(1234)
 
 ## ========================================================================== ##
-## FIGURE 1 Numerical comparison of different measures of relative error
-##  + Figure XX
+## FIGURE 1: Numerical comparison of different measures of relative error
 ## ========================================================================== ##
 
 y_hat <- 1
@@ -74,6 +73,11 @@ p2 <- p1 +
 ggsave("output/figures/different-relative-errors.png", width = 7, height = 3)
 
 
+## ========================================================================== ##
+## SI Version of Figure 1 (Figure SI.1) 
+## Numerical comparison square-root transformation
+## ========================================================================== ##
+
 p3 <- data |>
   filter(Metric != "AE after log transformation")|>
   plot_fct() +
@@ -96,7 +100,8 @@ ggsave("output/figures/different-relative-errors-sqrt.png", width = 7, height = 
 
 
 ## ========================================================================== ##
-## FIGURE 2
+## FIGURE 2: Expected CRPS scores as a function of the mean 
+## and variance of the forecast quantity.
 ## ========================================================================== ##
 if (file.exists("output/data/simulation-figure-2.Rda")) {
   scores_fig_2 <- readRDS(file = "output/data/simulation-figure-2.Rda")
@@ -190,7 +195,8 @@ ggsave("output/figures/SIM-mean-state-size.png", width = 7, height = 4.1)
 
 
 ## ========================================================================== ##
-## FIGURE 2 - Appendix version - Figure XX
+## SI Version of Figure 2 (Figure SI.2)
+## Illustration of the approximation
 ## ========================================================================== ##
 
 scores_fig_2 |>
@@ -219,7 +225,7 @@ ggsave("output/figures/SIM-score-approximation.png", width = 7, height = 8)
 
 
 ## ========================================================================== ##
-## FIGURE 3
+## FIGURE 3: llustration of impropriety of log-transformed CRPS. 
 ## ========================================================================== ##
 
 # in this plot the CRPS is approximated by the WIS with 99 quantiles
@@ -433,8 +439,172 @@ hub_data$location |>
 
 
 
+
[…]

### 2023-01-13 main.tex
Updates from Overleaf

@@
 Solid lines show the corresponding approximation of the expected CRPS based on an assumed normal distribution as discussed in section \ref{sec:methods:vst}.
-Figure \ref{fig:XXX} shows the quality of the approximation in more detail.
+Figure \ref{fig:score-approx} shows the quality of the approximation in more detail.
     The first distribution (red) is a truncated normal distribution with constant variance (we chose $\sigma = 1$ in order to only obtain positive samples).
@@
 For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}A).
-When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models change considerably.
-The change is stronger for cases than for deaths (Figure \ref{fig:HUB-cors}B) and increases noticeably with increasing forecast horizon.
+When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models change considerably (Figure \ref{fig:HUB-cors}B).
+The correlation between relative skill scores also decreases noticeably with increasing forecast horizon.
 
@@
 
-The rank order of different forecasting models changes noticeably (see Figure \ref{fig:HUB-rank-order}. 
+Figure Figure \ref{fig:HUB-rank-order} shows the changes in the ranking between different forecasting models.
+Encouragingly for the European Forecast Hub, the Hub ensemble, which is the forecast the organisers suggest forecast consumers make use of, remains the top forecasts across scoring schemes.
+For cases, the ILM-EKF model and the Forecast Hub baseline model exhibit the largest change in relative skill scores.
+For the ILM-EKF model the relative proportion of the score that is due to overprediction is reduced when applying a log-transformation before scoring (see Figure \ref{fig:HUB-rank-order}E.
+Instances where the model has overshot are penalised less heavily on the log scale, leading to an overall better score.
+For the Forecast Hub baseline model, the fact that it often puts relevant probability mass on zero (see Figure \ref{fig:HUB-model-comparison-baseline}), leads to worse scores after applying log-transformation due to large dispersion penalties.
+For deaths, the baseline model seems to get similarly penalised for its in relative terms highly forecasts.
+The performance of other models changes as well, but patterns are less discernible on this aggregate level.
+
 % I removed this part as it doesn't seem accurate anymore / we don't even have all the models in there anymore. 
@@
 Other models that are based on modelling the growth rate, such as the LANL-GrowthRate forecaster, do less well after a log-transform, indicating that this potential association between model structure and score under different approximations does not hold strictly.
-Encouragingly for the European Forecast Hub, the Hub ensemble, which is the forecast the organisers suggest forecast consumers make use of, remains one of the top forecasts across scoring schemes and indeed ranks higher for reported cases when a log-transform is used.
 
-We hypothesise that the lower correlation between skill scores for cases (compared to the correlation of scores for deaths) as displayed in Figure \ref{fig:HUB-cors}B is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths, and therefore relative skill scores are more strongly affected for cases.
+% THIS DOESN'T SEEM TRUE ANYMORE
+%We hypothesise that the lower correlation between skill scores for cases (compared to the correlation of scores for deaths) as displayed in Figure \ref{fig:HUB-cors}B is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths, and therefore relative skill scores are more strongly affected for cases.
 
@@
 
+\renewcommand{\thefigure}{SI.\arabic{figure}}
+\setcounter{figure}{0}
+\renewcommand{\thetable}{SI.\arabic{table}} \setcounter{table}{0}
+
 \subsection{Alternative Formulation of the WIS}
@@
 \includegraphics[width = 1\textwidth]{output/figures/SIM-score-approximation.png}
-\caption{Comparison of expected CRPS values of an ideal forecaster and approximated values.
-Expected scores for three distributions as described in Section \ref{sec:methods:vst} and Figure \ref{fig:SIM-wis-state-size-mean} and approximated as described in \ref{sec:methods:vst}}
+\caption{
+Visualisation of expected CRPS values against approximated scores using the approximation detailed in Section \ref{sec:methods:rankings} (see also Figure \ref{fig:SIM-wis-state-size-mean}).
+Expected CRPS scores are shown for three different distributions once on the natural scale (top row) and once scored on the log scale (bottom row).
+}
 \label{fig:score-approx}
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/SIM-score-approximation.png}
-    \caption{
-    Visualisation of expected CRPS values against approximated scores using the approximation detailed in Section \ref{sec:methods:rankings} (see also Figure \ref{fig:SIM-wis-state-size-mean}).
-Expected CRPS scores are shown for three different distributions once on the natural scale (top row) and once scored on the log scale (bottom row).
-    }
-    \label{fig:HUB-model-comparison-epinow}
-\end{figure}
-
 \clearpage

### 2023-01-13 R/00 full paper code.R
update y axis labels for plots

@@ -441,9 +441,23 @@ hub_data$location |>
 
 
 ## ========================================================================== ##
-## Figure 7: Scores for two-week-ahead predictions from the 
+## Figure 5: Scores for two-week-ahead predictions from the 
 ## EuroCOVIDhub-ensemble madein Germany.
 ## ========================================================================== ##
+
+label_fn <- function(x) {
+  x <- ifelse(x%%1 == 0, 
+              as.integer(x), x)
+  ifelse(x < 1000, 
+         paste(x), 
+         ifelse(x < 1e6, 
+                paste0(x / 1000, "k"),
+                ifelse(x < 1e9, 
+                       paste0(x / 1e6, "m"), 
+                       paste0(x / 1e9, "b"))
+         )
+  )
+}
 
 plot_pred_score <- function(hub_data, scores, model,
                             horizon = 2, type = "Cases", 
@@ -507,6 +521,7 @@ plot_pred_score <- function(hub_data, scores, model,
     ) +
     theme_scoringutils() +
     scale_fill_discrete(type = c("#DF536B", "#61D04F", "#2297E6")) +
+    scale_y_continuous(labels = label_fn) + 
     guides(fill = guide_legend(title = "WIS component")) +
     labs(y = "WIS contributions", x = "Target end date") 
   
@@ -571,14 +586,6 @@ DH"
     theme(legend.position = "bottom")
 }
 
-hub_data |> 
-  filter(location == "DE", 
-         horizon == 2,
-         model == "EuroCOVIDhub-ensemble", 
-         target_type == "Cases") |>
-  pull(target_end_date) |>
-  unique() |> sort()
-
 put_plot_together("EuroCOVIDhub-ensemble")
 ggsave(filename = "output/figures/HUB-model-comparison-ensemble.png", width = 10, height = 8.5)
 
@@ -607,20 +614,6 @@ ggsave(filename = "output/figures/HUB-model-comparison-baseline.png", width = 10
 ## Figure 6: Observations and scores across locations and forecast 
 ##           horizons for the European COVID-19 Forecast Hub data
 ## ========================================================================== ##
-
-label_fn <- function(x) {
-  x <- ifelse(x%%1 == 0, 
-              as.integer(x), x)
-  ifelse(x < 1000, 
-         paste(x), 
[…]

### 2023-01-17 R/00 full paper code.R
add illustration for the effect of different offsets when logging

@@ -609,7 +609,6 @@ ggsave(filename = "output/figures/HUB-model-comparison-baseline.png", width = 10
 
 
 
-
 ## ========================================================================== ##
 ## Figure 6: Observations and scores across locations and forecast 
 ##           horizons for the European COVID-19 Forecast Hub data
@@ -653,30 +652,31 @@ box_plot_obs <- hub_data |>
   labs(y = "Observations", x = "")  +
   theme(axis.title.x = element_blank())
 
+mean_scores_plot <- function(scores) {
+  plot_df <- scores |> 
+    filter(model == "EuroCOVIDhub-ensemble", 
+           horizon == 2) |>
+    group_by(target_type, location, scale, type_and_scale) |>
+    summarise(interval_score = mean(interval_score), 
+              .groups = "drop_last") |> 
+    full_join(mean_observations)
+  
+  plot_mean_scores <- plot_df |>
+    ggplot(aes(y = interval_score,
+               x = reorder_within(location, -mean_value, target_type))) + 
+    geom_bar(stat = "identity") + 
+    scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
+    facet_wrap(~ type_and_scale, scale = "free", ncol = 2) + 
+    theme_scoringutils() + 
+    theme(legend.position = "none") + 
+    scale_y_continuous(labels = label_fn) + 
+    scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) +
+    labs(y = "Mean weighted interval score", x = "Loaction") 
+  
+  return(plot_mean_scores)
+}
 
-plot_df <- scores |> 
-  filter(model == "EuroCOVIDhub-ensemble", 
-         horizon == 2, 
-         scale %in% c("log", "natural")) |>
-  group_by(target_type, location, scale, type_and_scale) |>
-  summarise(interval_score = mean(interval_score), 
-            .groups = "drop_last") |> 
-  full_join(mean_observations)
-
-plot_mean_scores <- plot_df |>
-  ggplot(aes(y = interval_score,
-             x = reorder_within(location, -mean_value, target_type))) + 
-  geom_bar(stat = "identity") + 
-  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
-  facet_wrap(~ type_and_scale, scale = "free") + 
-  theme_scoringutils() + 
-  theme(legend.position = "none") + 
-  scale_y_continuous(labels = label_fn) + 
[…]

### 2023-01-17 R/Refetch data from Forecast Hub.R

@@ -220,6 +220,7 @@ for (i in 1:length(indices)) {
 ## Score forecasts, adding versions for the log and sqrt transformations
 scores <- hub_data |>
   mutate(scale = "natural") |>
+  # add log data
   rbind(hub_data |>
           mutate(
             scale = "log", 
@@ -255,6 +256,98 @@ scores <- add_median_forecast(scores, hub_data)
 fwrite(scores, here("output", "data", "all-scores-european-hub.csv"))
 
 
+
+
+
+# create variants of the log score with different offsets
+scores_log_alts <- 
+  rbind(hub_data |>
+          mutate(
+            scale = "log + 0.1", 
+            true_value = log(true_value + 0.1), 
+            prediction = log(pmax(prediction, 0) + 0.1)
+          ), 
+        hub_data |> mutate(
+          scale = "log + 0.001", 
+          true_value = log(true_value + 0.001), 
+          prediction = log(pmax(prediction, 0) + 0.001)
+        ), 
+        hub_data |>
+          mutate(
+            scale = "log + 10x median", 
+            true_value = ifelse(target_type == "Cases", 
+                                log(true_value + 101630), 
+                                log(true_value + 530)), 
+            prediction = ifelse(target_type == "Cases", 
+                                log(pmax(prediction, 0) + 101630), 
+                                log(pmax(prediction, 0) + 530))
+          )
+  ) |>
+  score(metrics = c("interval_score")) |>
+  summarise_scores(by = c("model", "location",
+                          "target_end_date", "forecast_date",
+                          "horizon", "target_type", "scale"), 
+                   na.rm = TRUE)
+
+scores_log_alts <- scores_log_alts |>
+  rbind(scores,
+        fill = TRUE)
+
+scores_log_alts[scale == "log", scale := "log + 1"]
+scores_log_alts[, type_and_scale := paste0(target_type, " - ", scale)]
+
+
+
+# scores_0.001 <- hub_data |>
+#   mutate(scale = "natural") |>
+#   rbind(hub_data |>
+#           mutate(
+#             scale = "log", 
+#             true_value = log(true_value + 0.001), 
+#             prediction = log(pmax(prediction, 0) + 0.001)
+#           )) |>
+#   rbind(hub_data |>
+#           mutate(scale = "sqrt", 
+#                  true_value = sqrt(true_value), 
+#                  prediction = sqrt(prediction))) |>
+#   score(metrics = c("interval_score")) |>
+#   summarise_scores(by = c("model", "location",
[…]

### 2023-01-17 main.tex

@@
 Evaluations of forecasts can provide feedback for researchers to improve their models and train ensembles.
-It moreover helps decision-makers distinguish good from bad predictions and choose forecasters (used here and elsewhere in this manuscript to mean anyone or anything that issues a forecast) and models that should inform future decisions.
+It moreover helps decision-makers distinguish good from bad predictions and choose forecasters and models and models that should inform future decisions.
 
 Probabilistic forecasts are usually evaluated using so-called proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data.
-Proper scoring rules are constructed such that forecasters are incentivised to report their true belief about the future.
+Proper scoring rules are constructed such that forecasters (anyone or anything that issues a forecast) are incentivised to report their true belief about the future.
 Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the Weighted Interval Score \citep{bracherEvaluatingEpidemicForecasts2021}.
@@
 While the goal of the transformations is usually to improve the accuracy of the predictions, they can also be used to enhance and complement the evaluation process.
-In this paper, we argue that the aforementioned issues with evaluating forecasts based on measures of absolute error can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation.
+In this paper, we argue that the aforementioned issues with evaluating forecasts based on measures of absolute error on the natural scale can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation.
 Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score.
@@
 %We first provide a mathematical intuition of the impact of the log-transformation and its interpretation as relative error (Section \ref{sec:methods:relative}, measure of the accuracy of a forecast of the growth rate (Section \ref{sec:methods:growthrate}) and it's effect as a variance-stabilising transformation that makes can make scores independent of the mean of the forecast target (Section \ref{sec:methods:vst}.
-We then discuss practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}) and the effect of the log-transformation on forecast rankings (section \ref{sec:methods:rankings}.
+We then discuss practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}) and the effect of the log-transformation on forecast rankings (section \ref{sec:methods:rankings}).
 To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
@@
 To strengthen our intuition on how transforming outcomes prior to applying the CRPS shifts the emphasis between targets with high and low expected outcomes, Figure \ref{fig:SIM-wis-state-size-mean} shows the expected CRPS of ideal forecasters under different mean-variance relationships and transformations.
-We consider a Poisson distribution where $\sigma^2 = \mu$, a negative binomial distribution with $\theta = 10$ and $\sigma^2 = \mu + 0.1\mu^2$, and a normal distribution with constant variance.
+We consider a Poisson distribution where $\sigma^2 = \mu$, a negative binomial distribution with size parameter $\theta = 10$ and thus $\sigma^2 = \mu + \mu^2/10$, and a normal distribution with constant variance.
 We see that when applying the CRPS on the natural scale, the expected CRPS grows with the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster).
@@
 When applying a log-transformation first, the expected CRPS is almost independent of $\mu$ for the negative binomial distribution and large $\mu$, while smaller targets have higher expected CRPS in case of the Poisson distribution and the normal distribution with constant variance.
-When applying a square-root-transformation before the CRPS, the expected CRPS is independent of the mean for the Poisson-distribution, but not for the other two.
+When applying a square-root-transformation before the CRPS, the expected CRPS is independent of the mean for the Poisson-distribution, but not for the other two (with a positive relationship in the normal case and a negative one for the negative binomial).
 As can be seen in Figures \ref{fig:SIM-wis-state-size-mean} and \ref{fig:score-approx}, the approximations presented above work quite well for our simulated example.
@@
 Small discrete observed values can pose an additional problem as it may be difficult to reliably evaluate relative errors.
-A related issues occurs when the predictive distribution has a large probability mass on zero (or on very small values), as this can translate into an excessively wide forecast in relative terms.
+A related issue occurs when the predictive distribution has a large probability mass on zero (or on very small values), as this can translate into an excessively wide forecast in relative terms.
 This can be seen in Figure \ref{fig:HUB-model-comparison-baseline}, where the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm where forecasts contained zero in its prediction intervals.
@@
 Shown are CRPS (or WIS, respectively) values as a function of the observed value for two forecasters.
-Model A issues a geometric distribution with mean $\mu = 10$ and variance $\sigma^2 = \mu + \mu^2 = 110$), while Model B issues a Poisson distribution with mean and variance equal to 10.
+Model A issues a geometric distribution (a negative binomial distribution with size parameter $\theta = 1$) with mean $\mu = 10$ and variance $\sigma^2 = \mu + \mu^2 = 110$), while Model B issues a Poisson distribution with mean and variance equal to 10.
 Zeroes in this illustrative example were handled by adding one before applying the natural logarithm.}
@@
 We applied the log-transformation after adding 1 to all predictions and observed values.
-Where not otherwise stated, we report results for a two-week-ahead forecast horizon.
+The analysis was conducted in \texttt{R} \citep{R}, using the \texttt{scoringutils} package \citep{bosseEvaluatingForecastsScoringutils2022} for forecast evaluation.
+All code is available on GitHub
+(\url{https://github.com/epiforecasts/transformation-forecast-evaluation}). Where not otherwise stated, we report results for a two-week-ahead forecast horizon.
 
@@
 On the log scale, scores were more evenly distributed between targets (see Figure\ref{fig:HUB-mean-locations}D) and locations (see Figure\ref{fig:HUB-mean-locations}C).
-Both on the natural scale as well to a slightly lesser extent on the log scale, scores increased considerably with increasing forecast horizon (see Figure \ref{fig:HUB-mean-locations}E).
+Both on the natural scale as well on the log scale, scores increased considerably with increasing forecast horizon (see Figure \ref{fig:HUB-mean-locations}E).
 This reflects the increasing difficulty of forecasts further into the future and, for the log scale, corresponds with our expectations from Section \ref{sec:methods:growthrate}.
@@
 
-As argued in Section \ref{sec:methods:vst}, the mean-variance, or mean-CRPS, relationship determines which transformation can serve as a VST. To assess this empirically, we ran the regression
+As argued in Section \ref{sec:methods:vst}, the mean-variance, or mean-CRPS, relationship determines which transformation can serve as a VST.
+We can analyse this relationship empirically by running a regression that explains the CRPS as a function of the central estimate of the predictive distribution.
+We ran the regression
 
@@
 Note that we are using predictive medians rather than means as only the former are available in the European COVID-19 Forecast Hub.
+As the CRPS of an ideal forecaster scales with the standard deviation (see Section \ref{sec:methods:vst}), a value of $\beta = 1$ would imply a quadratic median-variance relationship, whereas a value of $\beta > 1$ ($\beta < 1$) would imply that the variance grows faster (slower) than the median.
 We applied the regression to case and death forecasts, pooled across horizons and stratified for one through four-week-ahead forecasts.
 Results are provided in Table \ref{tab:HUB-regression}.
-It can be seen that the estimates of $\beta$ always take a value below 1.
-As the CRPS scales with the standard deviation (see Section \ref{sec:methods:vst}), this can be interpreted as a slightly sub-quadratic mean-variance relationship.
+It can be seen that the estimates of $\beta$ always take a value below 1, implying a slightly sub-quadratic mean-variance relationship.
 The logarithmic transformation should thus approximately stabilize the variance (and CRPS), possibly leading to somewhat higher scores for smaller forecast targets.
@@
 
-To check this, we ran the regressions
+To check the relationship after the transformation, we ran the regressions
 \begin{equation*}
@@
 where $F_{\sqrt{\ }}$ is the predictive distribution on the square-root scale.
-As can be seen from Table \ref{tab:HUB-regression}, the results indeed indicate small targets lead to larger average CRPS when using the log transform ($\beta_{\log} < 0$), while the opposite is true for the square-root transform ($\beta_{\sqrt{\ }} > 0$).
+A value of $beta_{\log} = 0$ (or $\beta_{\sqrt{\ } = 0}$, respectively, would imply that scores are independent of the median prediction after the transformation.
+A value smaller (larger) than 0 would imply that smaller (larger) targets lead to higher scores.
+As can be seen from Table \ref{tab:HUB-regression}, the results indeed indicate that small targets lead to larger average CRPS when using the log transform ($\beta_{\log} < 0$), while the opposite is true for the square-root transform ($\beta_{\sqrt{\ }} > 0$).
 The results of the three regressions are also displayed in Figure \ref{fig:HUB-regression}.
-In this empirical example, the log transformation helps (albeit not perfectly), to stabilise WIS values and achieves this better suited than the square-root transformation.
+In this empirical example, the log transformation thus helps (albeit not perfectly), to stabilise WIS values, and it does so more successfully than the square-root transformation (with $\beta_{\log}$ closer to zero than $\beta_{\sqrt{\ }}$).
 
@@
 For the Forecast Hub baseline model, the fact that it often puts relevant probability mass on zero (see Figure \ref{fig:HUB-model-comparison-baseline}), leads to worse scores after applying log-transformation due to large dispersion penalties.
-For deaths, the baseline model seems to get similarly penalised for its in relative terms highly forecasts.
+For deaths, the baseline model seems to get similarly penalised for its in relative terms highly dispersed forecasts.
 The performance of other models changes as well, but patterns are less discernible on this aggregate level.
@@
 It also gives higher weight to another type of situation one may care about, namely one in which numbers start to rise from a previously low level.
-If one is interested in scoring a forecast of the multiplicative, rather than the exponential growth rate, it may be better to use a different transformation, e.g.
-dividing every forecast and corresponding observed value by the previously last observed value when the forecast was made.
 
 The log-transformation is only one of many transformations that may be useful and appropriate in an epidemiological context.
-One obvious option is to apply a population standardization to obtain incidence forecasts e.g., per 100,000 population If one is interested in multiplicative, rather than exponential growth rates, one could convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last observed value.
+One obvious option is to apply a population standardization to obtain incidence forecasts e.g., per 100,000 population ( \citep{abbottEvaluatingEpidemiologicallyMotivated2022}.
+If one is interested in multiplicative, rather than exponential growth rates, one could convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last observed value.
 We suggested using the natural logarithm as a variance-stabilising transformation (VST) or alternatively the square-root transformation in the case of a Poisson distributed variable.

### 2023-01-17 R/00 full paper code.R
update figures

@@ -166,6 +166,7 @@ p1 <- scores_fig_2 |>
   filter(mean <= 2000) |>
   mutate(crps = crps / mean(crps), 
          approximation = approximation / mean(approximation)) |>
+  mutate(scale = factor(scale, levels = c("natural", "sqrt", "log"))) |>
   ggplot(aes(y = crps, x = mean, colour = model)) +
   geom_line(aes(y = approximation)) +
   geom_point(size = 0.4) +  
@@ -293,9 +294,49 @@ ggsave("output/figures/example-log-first.png", width = 7, height = 2.1)
 
 
 
+## ========================================================================== ##
+## Figure 4: Illustration of the effect of adding an offset to the logarithm
+## ========================================================================== ##
+
+x <- seq(0.05, 100, 0.05)
+a <- c(0.1, 1, 10)
+
+plot_df <- expand.grid(
+  x = x, 
+  a = c(0, a)
+) |>
+  mutate(y = log(x + a))
+
+cols <- c("0"= "#000000", 
+          "0.1" = "#E69F00", 
+          "1" = "#56B4E9", 
+          "10" = "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+
+plot_df |>
+  ggplot(aes(x = x, y = y, color = factor(a), group = a)) +
+  geom_vline(xintercept = 5 * 0.1, 
+             color = "#E69F00",
+             linetype = "dashed", alpha = 0.4) +
+  geom_vline(xintercept = 5 * 1, 
+             color = "#56B4E9",
+             linetype = "dashed", alpha = 0.4) +
+  geom_vline(xintercept = 5 * 10, 
+             color = "#009E73",
+             linetype = "dashed", alpha = 0.4) +
+  geom_line() + 
+  theme_scoringutils() + 
+  scale_x_continuous(trans = "log10", breaks = c(0.05, 0.5, 5, 50), labels = label_fn) + 
+  scale_y_continuous(labels = label_fn) + 
+  scale_colour_manual(values = cols) + 
+  labs(y = "log (x + a)", color = "a")
+
+ggsave(filename = "output/figures/illustration-effect-offset-log.png", height = 2.5, width = 7)
+
+
 
 ## ========================================================================== ##
-## FIGURE 4: Illustration of the effect of the log-transformation of 
+## FIGURE 5: Illustration of the effect of the log-transformation of 
 ##           the ranking for a single forecast
[…]

### 2023-01-17 R/Refetch data from Forecast Hub.R

@@ -244,7 +244,7 @@ add_median_forecast <- function(scores, hub_data) {
   medians <- hub_data |>
     filter(quantile == 0.5) |> 
     rename(median_prediction = prediction) |>
-    select(-quantile)
+    select(-quantile, -true_value)
   
   out <- scores |>
     inner_join(medians)
@@ -289,68 +289,8 @@ scores_log_alts <-
                           "horizon", "target_type", "scale"), 
                    na.rm = TRUE)
 
-scores_log_alts <- scores_log_alts |>
-  rbind(scores,
-        fill = TRUE)
+scores_log_alts <- add_median_forecast(scores_log_alts, hub_data)
 
-scores_log_alts[scale == "log", scale := "log + 1"]
-scores_log_alts[, type_and_scale := paste0(target_type, " - ", scale)]
-
-
-
-# scores_0.001 <- hub_data |>
-#   mutate(scale = "natural") |>
-#   rbind(hub_data |>
-#           mutate(
-#             scale = "log", 
-#             true_value = log(true_value + 0.001), 
-#             prediction = log(pmax(prediction, 0) + 0.001)
-#           )) |>
-#   rbind(hub_data |>
-#           mutate(scale = "sqrt", 
-#                  true_value = sqrt(true_value), 
-#                  prediction = sqrt(prediction))) |>
-#   score(metrics = c("interval_score")) |>
-#   summarise_scores(by = c("model", "location",
-#                           "target_end_date", "forecast_date",
-#                           "horizon", "target_type", "scale"), 
-#                    na.rm = TRUE)
-# 
-# scores_0.001[, type_and_scale := paste0(target_type, " - ", scale)]
-# 
-# 
-# hub_data |>
-#   group_by(target_type) |>
-#   summarise(max = median(true_value))
-# 
-# scores_10median <- hub_data |>
-#   mutate(scale = "natural") |>
-#   rbind(hub_data |>
-#           mutate(
-#             scale = "log", 
-#             true_value = ifelse(target_type == "Cases", 
-#                                 log(true_value + 101630), 
-#                                 log(true_value + 530)), 
-#             prediction = ifelse(target_type == "Cases", 
-#                                 log(pmax(prediction, 0) + 101630), 
-#                                 log(pmax(prediction, 0) + 530))
-#           )) |>
-#   rbind(hub_data |>
-#           mutate(scale = "sqrt", 
-#                  true_value = sqrt(true_value), 
[…]

### 2023-01-17 R/00 full paper code.R
make points in Figure grey

@@ -910,7 +910,7 @@ scatter_wis_pred <- function(scores, filterscale, a = 1) {
            horizon == 2, 
            scale == filterscale) |>
     ggplot(aes(y = interval_score, x = median_prediction)) +
-    geom_point(size = 0.1, color = "grey20") + 
+    geom_point(size = 0.1, color = "grey20", alpha = 0.4) + 
     geom_line(
       data = filter(regs, scale == filterscale), 
       aes(y = yreg), 

### 2023-01-17 main.tex
Updates from Overleaf

@@
 When evaluating models across sets of forecasting tasks, it may be desirable for each target to have a similar impact on the overall results.
-In disease forecasting, this is not the case when using the CRPS, as the latter typically scales with the order of magnitude of the quantity to be predicted.
+In disease incidence forecasting, this is not the case when using the CRPS on the natural scale, as the latter typically scales with the order of magnitude of the quantity to be predicted.
 Average scores are then dominated by the results achieved for targets with high expected outcomes.
@@
     A: Normalised expected CRPS for ideal forecasts with increasing means for three distribution with different relationships between mean and variance.
-Expected CRPS was computed on the natural scale (middle), after adding one and applying a log-transformation to the data (middle) and after applying a square-root transformation (left).
+Expected CRPS was computed on the natural scale (middle), after adding one and applying a log-transformation to the data (middle) and after applying a square-root transformation (right).
 B: A but with x axis on the log scale.}
@@
 This represents a strictly monotonic transformation and therefore preserves the propriety of the resulting score.
-The choice of the quantity to add does however influences scores and rankings, as measures of relative errors shrink when adding a constant to the forecast and the observation.
-The impact is greatest for forecasts of small quantities, as relative errors shrink more the smaller the original value before adding a constant.
-In principle, it would also be possible to exploit the difference in how scores are affected in order to equalise scores for forecasts of quantities across different scales by adding a larger number (e.g.
-100 or 10,000) if average scores are dominated too strongly by forecasts for small quantities.
+The choice of the quantity to add does however influences scores and rankings, as measures of relative errors shrink when adding a constant $a$ to the forecast and the observation.
+We illustrate this in Figure \ref{fig:illustration-effect-log-offset}.
+As a rule of thumb, if the observation or the forecast is more than 5 times larger than the offset $a$, then deviations from the natural logarithm without offset should be small, and negligible if it is more than 50 times as large as the offset.
+Choosing a suitable offset $a$ balances two competing concerns: on the one hand, choosing a small $a$ makes sure that the transformation is as close to a natural logarithm as possible and scores can be interpreted as outlined above.
+On the other hand, choosing a larger $a$ can help stabilise scores for forecasts and observations close to zero and equalise scores across different orders of magnitude, avoiding giving excessive weight to forecasts for small quantities (see Figure \ref{fig:HUB-log-different-offsets}).
 
-Small discrete observed values can pose an additional problem as it may be difficult to reliably evaluate relative errors.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/illustration-effect-offset-log.png}
+    \caption{Illustration of the effect of adding a small quantity to a value before taking the natural logarithm.
+For increasing x, all lines eventually approach the black line (representing a transformation with no offset applied).
+    For a given solid line, the dashed line of the same colour marks the x-value that  is equal to 5 times the corresponding offset.}
+    \label{fig:illustration-effect-log-offset}
+\end{figure}
+
+Small discrete observed values can pose a problem as it may be difficult to reliably evaluate relative errors.
 A related issue occurs when the predictive distribution has a large probability mass on zero (or on very small values), as this can translate into an excessively wide forecast in relative terms.
@@
 Shown are CRPS (or WIS, respectively) values as a function of the observed value for two forecasters.
-Model A issues a geometric distribution (a negative binomial distribution with size parameter $\theta = 1$) with mean $\mu = 10$ and variance $\sigma^2 = \mu + \mu^2 = 110$), while Model B issues a Poisson distribution with mean and variance equal to 10.
+Model A issues a geometric distribution (a negative binomial distribution with size parameter $\theta = 1$) with mean $\mu = 10$ and variance $\sigma^2 = \mu + \mu^2 = 110$, while Model B issues a Poisson distribution with mean and variance equal to 10.
 Zeroes in this illustrative example were handled by adding one before applying the natural logarithm.}
@@
 All predictive quantiles were truncated at 0.
-We applied the log-transformation after adding 1 to all predictions and observed values.
+We applied the log-transformation after adding a constant $a = 1$ to all predictions and observed values.
+The choice of $a = 1$ in part reflects convention, but also represents a suitable choice as it avoids giving excessive weight to forecasts close to zero, while at the same time ensuring that scores for observations $> 5$ can be interpreted reasonably.
 The analysis was conducted in \texttt{R} \citep{R}, using the \texttt{scoringutils} package \citep{bosseEvaluatingForecastsScoringutils2022} for forecast evaluation.
@@
 \toprule
-Horizon & Target & $\alpha$ & $\beta$ & $\alpha_{\log}$ & $\beta_{\log}$ & $\alpha_{\sqrt{\ }}$ & $\beta_{\sqrt{\ }}$\\
+Horizon & Target & $\alpha$ & $\beta$ & $\alpha_\sqrt{\ }$ & $\beta_\sqrt{\ }$ & $\alpha_\log$ & $\beta_\log$\\
 \midrule
-all & all & -1.093 & 0.963 & 0.391 & 0.001 & -0.352 & 0.201\\
+all & all & -1.093 & 0.963 & -0.352 & 0.201 & 0.391 & 0.001\\
 \addlinespace
-all & Cases & 0.036 & 0.858 & 0.751 & -0.033 & 0.043 & 0.201\\
-all & Deaths & -0.884 & 0.868 & 0.436 & -0.023 & 0.273 & 0.121\\
+all & Cases & 0.036 & 0.858 & 0.043 & 0.201 & 0.751 & -0.033\\
+all & Deaths & -0.884 & 0.868 & 0.273 & 0.121 & 0.436 & -0.023\\
 \addlinespace
-1 & all & -1.402 & 0.923 & 0.318 & -0.014 & 0.320 & 0.088\\
-2 & all & -1.221 & 0.967 & 0.364 & -0.003 & 0.112 & 0.164\\
-3 & all & -1.001 & 0.984 & 0.410 & 0.008 & -0.094 & 0.241\\
-4 & all & -0.757 & 0.986 & 0.469 & 0.015 & 0.000 & 0.299\\
+1 & all & -1.402 & 0.923 & 0.320 & 0.088 & 0.318 & -0.014\\
+2 & all & -1.221 & 0.967 & 0.112 & 0.164 & 0.364 & -0.003\\
+3 & all & -1.001 & 0.984 & -0.094 & 0.241 & 0.410 & 0.008\\
+4 & all & -0.757 & 0.986 & 0.000 & 0.299 & 0.469 & 0.015\\
 \addlinespace
-1 & Cases & -0.862 & 0.876 & 0.433 & -0.024 & 0.790 & 0.087\\
-2 & Cases & -0.243 & 0.877 & 0.660 & -0.031 & 0.959 & 0.162\\
-3 & Cases & 0.372 & 0.855 & 0.882 & -0.037 & 1.109 & 0.238\\
-4 & Cases & 0.816 & 0.837 & 1.009 & -0.036 & 1.645 & 0.296\\
+1 & Cases & -0.862 & 0.876 & 0.790 & 0.087 & 0.433 & -0.024\\
+2 & Cases & -0.243 & 0.877 & 0.959 & 0.162 & 0.660 & -0.031\\
+3 & Cases & 0.372 & 0.855 & 1.109 & 0.238 & 0.882 & -0.037\\
+4 & Cases & 0.816 & 0.837 & 1.645 & 0.296 & 1.009 & -0.036\\
 \addlinespace
-1 & Deaths & -1.146 & 0.832 & 0.376 & -0.035 & 0.457 & 0.048\\
-2 & Deaths & -0.981 & 0.867 & 0.416 & -0.028 & 0.443 & 0.084\\
-3 & Deaths & -0.807 & 0.885 & 0.453 & -0.019 & 0.349 & 0.131\\
-4 & Deaths & -0.602 & 0.891 & 0.501 & -0.011 & 0.125 & 0.194\\
+1 & Deaths & -1.146 & 0.832 & 0.457 & 0.048 & 0.376 & -0.035\\
+2 & Deaths & -0.981 & 0.867 & 0.443 & 0.084 & 0.416 & -0.028\\
+3 & Deaths & -0.807 & 0.885 & 0.349 & 0.131 & 0.453 & -0.019\\
+4 & Deaths & -0.602 & 0.891 & 0.125 & 0.194 & 0.501 & -0.011\\
 \bottomrule
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% HUB plots %
+
 \begin{table}[h!]
@@
 \midrule
-Cases & Observations & mean & 62004 & 9.19\\
-Cases & Observations & sd & 171723 & 2.10\\
-Cases & Observations & var & 29488651564 & 4.42\\
+Cases & Observations & mean & 61979 & 9.19\\
+Cases & Observations & sd & 171916 & 2.10\\
+Cases & Observations & var & 29555122130 & 4.42\\
+Deaths & Observations & mean & 220 & 3.89\\
+Deaths & Observations & sd & 435 & 1.96\\
 \addlinespace
-Deaths & Observations & mean & 223 & 3.90\\
-Deaths & Observations & sd & 441 & 1.96\\
-Deaths & Observations & var & 194221 & 3.85\\
-\addlinespace
-\hline
-\addlinespace
+Deaths & Observations & var & 189051 & 3.83\\
 Cases & WIS & mean & 15840 & 0.27\\
 Cases & WIS & sd & 53117 & 0.28\\
-\addlinespace
 Deaths & WIS & mean & 31 & 0.23\\
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/HUB-scores-locations-log-variants.png}
+    \caption{Mean WIS in different locations for different transformations applied before scoring.
+Shown are scores for two-week-ahead forecasts of the EuroCOVIDhub-ensemble.
+On the natural scale (with no transformation prior applying the WIS), scores correlate strongly with the average number of observed values in a given location.
+The same is true for scores obtained after applying a square-root transformation, or after applying a log-transformation with a large offset $a$.
+For illustrative purposes, $a$ was chosen to be 101630 for cases and 530 for deaths, 10 times the respective median observed value.
+For large values of $a$, $log(x + a)$ grows linearly in $x$, meaning that we expect to observe the same patterns as in the case with no transformation.
+For decreasing values of $a$, we give more relative weight to scores in small locations.}
+    \label{fig:HUB-log-different-offsets}
+\end{figure}
+
 \clearpage

### 2023-01-17 main.tex

@@
     A: Normalised expected CRPS for ideal forecasts with increasing means for three distribution with different relationships between mean and variance.
-Expected CRPS was computed on the natural scale (middle), after adding one and applying a log-transformation to the data (middle) and after applying a square-root transformation (right).
+Expected CRPS was computed on the natural scale (left), after applying a square-root transformation (middle), and after adding one and applying a log-transformation to the data (right).
 B: A but with x axis on the log scale.}
@@
 We illustrate this in Figure \ref{fig:illustration-effect-log-offset}.
-As a rule of thumb, if the observation or the forecast is more than 5 times larger than the offset $a$, then deviations from the natural logarithm without offset should be small, and negligible if it is more than 50 times as large as the offset.
+As a rule of thumb, if if $x > 5a$, the difference between $\log{(x + a)}$ and $\log{(x)}$ is small, and it becomes negligible if $x > 50a$.
 Choosing a suitable offset $a$ balances two competing concerns: on the one hand, choosing a small $a$ makes sure that the transformation is as close to a natural logarithm as possible and scores can be interpreted as outlined above.
-On the other hand, choosing a larger $a$ can help stabilise scores for forecasts and observations close to zero and equalise scores across different orders of magnitude, avoiding giving excessive weight to forecasts for small quantities (see Figure \ref{fig:HUB-log-different-offsets}).
+On the other hand, choosing a larger $a$ can help stabilise scores for forecasts and observations close to zero, avoiding giving excessive weight to forecasts for small quantities (see Figure \ref{fig:HUB-log-different-offsets}).
 
@@
 
-Small discrete observed values can pose a problem as it may be difficult to reliably evaluate relative errors.
 A related issue occurs when the predictive distribution has a large probability mass on zero (or on very small values), as this can translate into an excessively wide forecast in relative terms.
@@
 Shown are CRPS (or WIS, respectively) values as a function of the observed value for two forecasters.
-Model A issues a geometric distribution (a negative binomial distribution with size parameter $\theta = 1$) with mean $\mu = 10$ and variance $\sigma^2 = \mu + \mu^2 = 110$, while Model B issues a Poisson distribution with mean and variance equal to 10.
+Model A issues a geometric distribution (a negative binomial distribution with size parameter $\theta = 1$) with mean $\mu = 10$ and variance $\sigma^2 = \mu + \mu^2 = 110$), while Model B issues a Poisson distribution with mean and variance equal to 10.
 Zeroes in this illustrative example were handled by adding one before applying the natural logarithm.}
@@
 A: WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against median predicted values.
-B: Same as A, with scores obtained after applying a log-transformation to the data.
-C: Same as A, with scores obtained after applying a square-root-transformation to the data.}
+B: Same as A, with scores obtained after applying a square-root-transformation to the data.
+C: Same as A, with scores obtained after applying a log-transformation to the data.}
     \label{fig:HUB-regression}
@@
     \caption{
-    Forecasts and scores for two-week-ahead predictions from the epiforecasts-EpiNow2 model made in Germany.
+    Forecasts and scores for two-week-ahead predictions from the epiforecasts-EpiNow2 model \citep{epinow2} made in Germany.
 A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale.

### 2023-01-20 main.tex
Updates from Overleaf

@@
   Anne Cori\thanks{MRC Centre for Outbreak Analysis and Modelling, Department of Infectious Disease Epidemiology, School of Public Health, Imperial College London, London, United Kingdom} $^{}$, \\
-  Edwin van Leeuwen\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom} $^{,}$\thanks{UK Health Security Agency, London, United Kingdom} $^{}$,
-  Johannes Bracher\thanks{Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology, Karlsruhe, Germany}$^{,}$\thanks{Computational Statistics Group, Heidelberg Institute for Theoretical Studies, Heidelberg, Germany} $^{}$,
-  Sebastian Funk\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$}
+  Edwin van Leeuwen\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom} $^{,}$\thanks{Modelling & Economics Unit and NIHR Health Protection Research Unit in Modelling \& Health Economics, UK Health Security Agency, London, United Kingdom} $^{}$,
+  Johannes Bracher\thanks{Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology, Karlsruhe, Germany } $^{,}$\thanks{Computational Statistics Group, Heidelberg Institute for Theoretical Studies, Heidelberg, Germany } $^{, \dagger}$,
+  Sebastian Funk\footnotemark[1] $^{, }$\footnotemark[2]$ ^{}$ $ ^{, \dagger}$}
 
@@
 \begin{abstract}
+
+%I swapped the commented out version and this one on Wednesday, January 18 2.47pm EST. 
+%Forecast evaluation plays an essential role in improving forecasting models and can support their role in informing decision-making.
+In epidemiology, probabilistic forecasts are often scored using the Weighted Interval Score (WIS) or the Continuous Ranked Probability Score (CRPS), which are measures of absolute distance between forecast and observation.
+Infectious disease transmission is commonly modelled as an exponential process characterized by the growth rate or the effective reproduction number.
+However, if the true objective of the forecaster is to predict future growth rates or reproduction numbers, then scoring based on the absolute distance between forecast and observation may not reflect this underlying task well and thus make it difficult to compare forecasts across time, space, and forecast targets.
+As a solution, we propose transforming the data before scoring the forecasts to better reflect the underlying generative process.
+Specifically, we discuss applying CRPS after log transformation of predictions and observations, which can be interpreted as a measure of relative error and as a score for an implicit forecast of the exponential growth rate.
+Moreover, we discuss its possible role as a variance-stabilizing transformation.
+Applying the natural logarithm to data from the European COVID-19 Forecast Hub, we found that it changed model rankings, regardless of stratification by forecast date, location or target types, and emphasised situations in which models missed lower inflection points and consequent rises in numbers, while penalising missed peaks less severely.
+We conclude that appropriate transformations, of which the natural logarithm is only one possible option, should be considered when assessing the performance of different models in the context of infectious disease incidence.
+
 Forecast evaluation plays an essential role in the development cycle of predictive epidemic models and can inform their use for public health decision-making.
@@
 They are commonly applied directly to predicted and observed incidence counts, but it can be questioned whether this yields the most meaningful results given the exponential nature of epidemic processes and the several orders of magnitude that observed values can span over space and time.
-In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results.
-We motivate the procedure threefold using the CRPS on log-transformed counts as an example: firstly, it can be interpreted as a probabilistic version of a relative error.
+In this paper, we argue that log transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results.
+We motivate the procedure threefold using the CRPS on log-transformed counts as an example: Firstly, it can be interpreted as a probabilistic version of a relative error.
 Secondly, it reflects how well models predicted the time-varying epidemic growth rate.
@@
 Applying the log transformation to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types.
-Situations in which models missed beginning upward swings are more strongly emphasized while failing to predict a downturn following a peak is less severely penalized.
+Situations in which models missed the beginning of upward swings are more strongly emphasized while failing to predict a downturn following a peak is less severely penalized.
 We conclude that appropriate transformations, of which the natural logarithm is only one particularly attractive option, should be considered when assessing the performance of different models in the context of infectious disease incidence.
-
-%Forecast evaluation plays an essential role in improving forecasting models and can support their role in informing decision-making.
-In epidemiology, probabilistic forecasts are often scored using the Weighted Interval Score (WIS) or the Continuous Ranked Probability Score (CRPS), which are measures of absolute distance between forecast and observation.
-Infectious disease transmission is commonly modelled as an exponential process characterized by the growth rate or the effective reproduction number.
-However, if the true objective of the forecaster is to predict future growth rates or reproduction numbers then scoring based on the absolute distance between forecast and observation may not reflect this underlying task well and thus make it difficult to compare forecasts across time, space, and forecast targets.
-As a solution, we propose transforming the data before scoring the forecasts to better reflect the underlying generative process.
-We further propose the natural logarithm  as a candidate transformation and discuss its interpretation as a measure of relative error and as a score for an implicit forecast of the exponential growth rate.
-Applying the natural logarithm to data from the European COVID-19 Forecast Hub, we found that it changed model rankings, regardless of stratification by forecast date, location or target types, and emphasised situations in which models missed lower inflection points and consequent rises in numbers, while penalising missed peaks less severely.
-We conclude that appropriate transformations, of which the natural logarithm is only one possible option, should be considered when assessing the performance of different models in the context of infectious disease incidence.
 \end{abstract}
@@
 
-{\footnotesize $^*$ Correspondence to \url{nikos.bosse@lshtm.ac.uk})}
+{\footnotesize $^*$ Correspondence to \url{nikos.bosse@lshtm.ac.uk}}
+{\footnotesize $^\dagger$ Contributed equally}
 
@@
 
-Probabilistic forecasts \citep{heldProbabilisticForecastingInfectious2017} play an important role in decision-making in epidemiology and public health \citep{reichCollaborativeMultiyearMultimodel2019, funkShorttermForecastsInform2020, cramerEvaluationIndividualEnsemble2021, bracherShorttermForecastingCOVID192021, europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}, as well as other areas as diverse as economics \citep{timmermannForecastingMethodsFinance2018, elliottForecastingEconomicsFinance2016} or meteorology \citep{gneitingWeatherForecastingEnsemble2005, kukkonenReviewOperationalRegionalscale2012}.
-Epidemiological modelling in particular has received widespread attention during the COVID-19 pandemic.
+Probabilistic forecasts \citep{heldProbabilisticForecastingInfectious2017} play an important role in decision-making in epidemiology and public health \citep{doi:10.2105/AJPH.2022.306831}, as well as other areas as diverse as economics \citep{timmermannForecastingMethodsFinance2018} or meteorology \citep{gneitingWeatherForecastingEnsemble2005}.
+Forecasts based on epidemiological modelling in particular has received widespread attention during the COVID-19 pandemic.
 Evaluations of forecasts can provide feedback for researchers to improve their models and train ensembles.
-It moreover helps decision-makers distinguish good from bad predictions and choose forecasters and models and models that should inform future decisions.
+They moreover help decision-makers distinguish good from bad predictions and choose forecasters and models that are best suited to inform future decisions.
 
@@
 where $y$ is the true observed value and $F$ the cumulative distribution function (CDF) of the predictive distribution.
-The CRPS can be understood as a generalisation of the absolute error to predictive distributions, and interpreted n the natural scale of the data.
-The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in COVID-19 Forecast Hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
+The CRPS can be understood as a generalisation of the absolute error to predictive distributions, and interpreted on the natural scale of the data.
+The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in so-called COVID-19 Forecast Hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
 The WIS is defined as
@@
 The WIS can be decomposed into three components, dispersion, overprediction, underprediction, which reflect the width of the forecast and whether it was centred above or below the observed value.
-We show an alternative definition based on central prediction intervals in Section \ref{sec:alternative-wis} which illustrates this decomposition more clearly.
+We show an alternative definition based on central prediction intervals in Section \ref{sec:alternative-wis} which illustrates this decomposition.
 
@@
 
-If case numbers are evolving based on an exponential process and the  modelling task revolves around estimating the reproduction number or the corresponding growth rate, then evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount.
+If case numbers are evolving based on an exponential process and the  modelling task revolves around estimating and forecasting the reproduction number or the corresponding growth rate, then evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount.
 This is because for exponential processes errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
@@
 While the goal of the transformations is usually to improve the accuracy of the predictions, they can also be used to enhance and complement the evaluation process.
-In this paper, we argue that the aforementioned issues with evaluating forecasts based on measures of absolute error on the natural scale can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation.
+In this paper, we argue that the aforementioned issues with evaluating epidemic forecasts based on measures of absolute error on the natural scale can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation.
 Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score.
@@
 %We first provide a mathematical intuition of the impact of the log-transformation and its interpretation as relative error (Section \ref{sec:methods:relative}, measure of the accuracy of a forecast of the growth rate (Section \ref{sec:methods:growthrate}) and it's effect as a variance-stabilising transformation that makes can make scores independent of the mean of the forecast target (Section \ref{sec:methods:vst}.
-We then discuss practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}) and the effect of the log-transformation on forecast rankings (section \ref{sec:methods:rankings}).
-To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
+We then discuss practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}) and the effect of the log-transformation on forecast rankings (Section \ref{sec:methods:rankings}).
+To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  \citep[Section 3]{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
 Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues (Section \ref{sec:discussion}).
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Transforming forecasts and observations}
+\section{Logarithmic transformation of forecasts and observations}
 \label{sec:methods}
@@
 $$
-The absolute error after log transforming is thus an approximation of the \textit{absolute percentage error} \citep[APE,][]{gneitingMakingEvaluatingPoint2011a}.
+The absolute error after log transforming is thus an approximation of the \textit{absolute percentage error} \citep[APE,][]{gneitingMakingEvaluatingPoint2011a} as long as forecast and observation are close.
 As we assumed that $\hat{y} \approx y$, we can also interpret it as an approximation of the \textit{relative error} (RE)
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/example-log-first.png}
-    \caption{Illustration of impropriety of log-transformed CRPS.
-We assume $Y \sim \text{LogNormal}(0, 1)$ and evaluate the expected CRPS for predictive distributions $\text{LogNormal}(0, \sigma)$ with varying values of $\sigma \in [0.1, 2]$.
-For the regular CRPS (left) and CRPS applied to log-transformed outcomes (middle), the lowest expectation is achieved for the true value $\sigma = 1$.
-For the log-transformed CRPS, the optimal value is  0.9, i.e.
-there is an incentive to report a forecast that is too sharp.}
-    \label{fig:log-improper}
-\end{figure}
-
-In practice, one issue with transformations, such as the log transform, is that they are not readily applicable to negative numbers or zero values, which need to be removed or otherwise handled. 
+In practice, one issue with the log transform is that they are not readily applicable to negative numbers or zero values, which need to be removed or otherwise handled. 
 One common approach to deal with zeros is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm \citep{bellegoDealingLogsZeros2022}.
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/illustration-effect-offset-log.png}
-    \caption{Illustration of the effect of adding a small quantity to a value before taking the natural logarithm.
-For increasing x, all lines eventually approach the black line (representing a transformation with no offset applied).
-    For a given solid line, the dashed line of the same colour marks the x-value that  is equal to 5 times the corresponding offset.}
-    \label{fig:illustration-effect-log-offset}
-\end{figure}
-
 A related issue occurs when the predictive distribution has a large probability mass on zero (or on very small values), as this can translate into an excessively wide forecast in relative terms.
-This can be seen in Figure \ref{fig:HUB-model-comparison-baseline}, where the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm where forecasts contained zero in its prediction intervals.
+This can be seen in Figure \ref{fig:HUB-model-comparison-baseline}.
+Here, the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm because forecasts contained zero in its prediction intervals.
 
@@
 Depending on the relationship between the mean and variance of the forecast target, a log-transformation may even lead to systematically larger scores assigned to small forecast targets, as illustrated in Figure \ref{fig:SIM-wis-state-size-mean}.
-
-%A different transformation like the square-root may then be more appropriate to make scores independent of the mean of the target quantity (as discussed in Section \ref{sec:methods:vst}).
-
-%Whether forecasts of small or large targets receive higher scores than those for targets with larger quantities, depends on the relationship between the mean and the variability of the quantity of interest (see Section \ref{sec:methods:vst}.
-For the simulated count data, log predictions for small quantities received on average higher scores than forecasts for large quantities if the mean and the variance grew at the same rate $\sigma^2 = \mu$, about equal scores when the variance grew at a rate of $\sigma^2 = \mu + \mu^2$ (which for large $\mu$ is dominated by the quadratic term), and smaller scores when the variance grew faster than that (Figure \ref{fig:SIM-wis-state-size-mean}).
 
@@
 The relative skill for a given model compared to others is then obtained by taking the geometric mean of all score ratios which involve that model.
-Low values are better, and the average model receives a relative skill score of 1.
+Low values are better, and the "average" model receives a relative skill score of 1.
 
@@
 Across the dataset, the average number of observed cases and deaths varied considerably by location and target type (see Figure \ref{fig:HUB-mean-locations}A and B).
-On the natural scale, scores also varied considerably across targets (see Figure\ref{fig:HUB-mean-locations}D) and locations (see Figure\ref{fig:HUB-mean-locations}C).
+On the natural scale, scores show a pattern quite similar to the observations across targets (see Figure\ref{fig:HUB-mean-locations}D) and locations (see Figure\ref{fig:HUB-mean-locations}C).
 On the log scale, scores were more evenly distributed between targets (see Figure\ref{fig:HUB-mean-locations}D) and locations (see Figure\ref{fig:HUB-mean-locations}C).
@@
 We ran the regression
-
 \begin{equation*}
@@
 Note that we are using predictive medians rather than means as only the former are available in the European COVID-19 Forecast Hub.
-As the CRPS of an ideal forecaster scales with the standard deviation (see Section \ref{sec:methods:vst}), a value of $\beta = 1$ would imply a quadratic median-variance relationship, whereas a value of $\beta > 1$ ($\beta < 1$) would imply that the variance grows faster (slower) than the median.
+As the CRPS of an ideal forecaster scales with the standard deviation (see Section \ref{sec:methods:vst}), a value of $\beta = 1$ would imply a quadratic median-variance relationship; the natural logarithm could then serve as a VST.
+A value of $\beta 0.5$ would imply a linear median-variance relationship, suggesting the square root as a VST.
 We applied the regression to case and death forecasts, pooled across horizons and stratified for one through four-week-ahead forecasts.
 Results are provided in Table \ref{tab:HUB-regression}.
-It can be seen that the estimates of $\beta$ always take a value below 1, implying a slightly sub-quadratic mean-variance relationship.
+It can be seen that the estimates of $\beta$ always take a value somewhat below 1, implying a slightly sub-quadratic mean-variance relationship.
 The logarithmic transformation should thus approximately stabilize the variance (and CRPS), possibly leading to somewhat higher scores for smaller forecast targets.
@@
 \toprule
-Horizon & Target & $\alpha$ & $\beta$ & $\alpha_\sqrt{\ }$ & $\beta_\sqrt{\ }$ & $\alpha_\log$ & $\beta_\log$\\
+Horizon & Target & $\alpha$ & $\beta$ & $\alpha_\sqrt{\ }$ & $\beta_\sqrt{\ }$ & $\alpha_\log$ & $\beta_\log$ \\
 \midrule
@@
 where $F_{\sqrt{\ }}$ is the predictive distribution on the square-root scale.
-A value of $beta_{\log} = 0$ (or $\beta_{\sqrt{\ } = 0}$, respectively, would imply that scores are independent of the median prediction after the transformation.
+A value of $\beta_{\log} = 0$ (or $\beta_{\sqrt{\ } = 0}$, respectively, would imply that scores are independent of the median prediction after the transformation.
 A value smaller (larger) than 0 would imply that smaller (larger) targets lead to higher scores.
@@
 The results of the three regressions are also displayed in Figure \ref{fig:HUB-regression}.
-In this empirical example, the log transformation thus helps (albeit not perfectly), to stabilise WIS values, and it does so more successfully than the square-root transformation (with $\beta_{\log}$ closer to zero than $\beta_{\sqrt{\ }}$).
+In this empirical example, the log transformation thus helps (albeit not perfectly), to stabilise WIS values, and it does so more successfully than the square-root transformation.
+As can be seen from Figure \ref{fig:HUB-regression}, the expected CRPS scores for case targets with medians of 10 and 100,000 differ by more then a factor of ten for the square root transformation, but only a factor of around 2 for the logarithm.
 
@@
 B: Correlation between relative skill scores.
-For every forecast horizon and target type, a separate rel.
-skill score was computed per model using pairwise comparisons.
-The plot shows the correlation between the rel.
-skill scores on the natural vs.
+For every forecast horizon and target type, a separate relative skill score was computed per model using pairwise comparisons.
+The plot shows the correlation between the relative skill scores on the natural vs.
 on the log scale.}
@@
 Figure Figure \ref{fig:HUB-rank-order} shows the changes in the ranking between different forecasting models.
-Encouragingly for the European Forecast Hub, the Hub ensemble, which is the forecast the organisers suggest forecast consumers make use of, remains the top forecasts across scoring schemes.
+Encouragingly for the European Forecast Hub, the Hub ensemble, which is the forecast the organisers suggest forecast consumers make use of, remains the top model across scoring schemes.
 For cases, the ILM-EKF model and the Forecast Hub baseline model exhibit the largest change in relative skill scores.
@@
 The performance of other models changes as well, but patterns are less discernible on this aggregate level.
-
-% I removed this part as it doesn't seem accurate anymore / we don't even have all the models in there anymore. 
-%indicates that models that do not make use of a exponential error model and thus would be biased towards under-prediction during periods of exponential growth, such as the MUNI-ARMA model, were ranked lower when using a log-transform.
-The epiforecasts-EpiNow2 forecast model, on the other hand~(created by co-authors of this study), which makes forecasts based on the behaviour of the reproduction number and thus implicitly uses an exponential error model, was the most impacted by the change to log-scale scoring moving from last place for reported cases to 4th place.
-Other models that are based on modelling the growth rate, such as the LANL-GrowthRate forecaster, do less well after a log-transform, indicating that this potential association between model structure and score under different approximations does not hold strictly.
-
-% THIS DOESN'T SEEM TRUE ANYMORE
-%We hypothesise that the lower correlation between skill scores for cases (compared to the correlation of scores for deaths) as displayed in Figure \ref{fig:HUB-cors}B is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths, and therefore relative skill scores are more strongly affected for cases.
 
@@
 The log-transformation is only one of many transformations that may be useful and appropriate in an epidemiological context.
-One obvious option is to apply a population standardization to obtain incidence forecasts e.g., per 100,000 population ( \citep{abbottEvaluatingEpidemiologicallyMotivated2022}.
+One obvious option is to apply a population standardization to obtain incidence forecasts e.g., per 100,000 population \citep{abbottEvaluatingEpidemiologicallyMotivated2022}.
 If one is interested in multiplicative, rather than exponential growth rates, one could convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last observed value.
@@
 This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluating the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}.
-This, unfortunately, is not feasible for forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate, but may be an interesting approach if predictive samples are available.
+This, unfortunately, is not feasible under the current quantile-based format of the Forecast Hubs, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate.
+However, it may be an interesting approach if predictive samples are available or if quantiles for weekwise growth rates have been collected.
 It is possible to go beyond choosing a single transformation by constructing composite scores as a weighted sum of scores based on different transformations.
@@
     \centering
-\includegraphics[width = 1\textwidth]{output/figures/different-relative-errors-sqrt.png}
-\caption{Numerical comparison of different measures of relative error: absolute percentage error (APE), relative error (RE), symmetric absolute percentage error (SAPE) and the absolute error applied to \textit{sqrt}-transformed predictions and observations.
-We denote the predicted value by $\hat{y}$ and display errors as a function of the ratio of observed and predicted value.
-A: x-axis shown on a linear scale.
-B: x-axis shown on a logarithmic scale.}
-\label{fig:SAPE-sqrt}
+    \includegraphics[width=0.99\textwidth]{output/figures/example-log-first.png}
+    \caption{Illustration of impropriety of log-transformed CRPS.
+We assume $Y \sim \text{LogNormal}(0, 1)$ and evaluate the expected CRPS for predictive distributions $\text{LogNormal}(0, \sigma)$ with varying values of $\sigma \in [0.1, 2]$.
+For the regular CRPS (left) and CRPS applied to log-transformed outcomes (middle), the lowest expectation is achieved for the true value $\sigma = 1$.
+For the log-transformed CRPS, the optimal value is  0.9, i.e.
+there is an incentive to report a forecast that is too sharp.}
+    \label{fig:log-improper}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/illustration-effect-offset-log.png}
+    \caption{Illustration of the effect of adding a small quantity to a value before taking the natural logarithm.
+For increasing x, all lines eventually approach the black line (representing a transformation with no offset applied).
+    For a given solid line, the dashed line of the same colour marks the x-value that  is equal to 5 times the corresponding offset.}
+    \label{fig:illustration-effect-log-offset}
 \end{figure}
@@
 \clearpage
-\bibliography{log-or-not.bib}
+\bibliography{log-or-not.bib, software.bib}
 

### 2023-01-20 main.tex
addressing comments

@@
 We then discuss practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}) and the effect of the log-transformation on forecast rankings (Section \ref{sec:methods:rankings}).
-To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  \citep[Section 3]{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
+To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  \citep[Section \ref{sec:HUB}]{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
 Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues (Section \ref{sec:discussion}).
@@
 \section{Empirical example: the European Forecast Hub}
-\label{sec:results}
+\label{sec:HUB}
 

### 2023-01-23 main.tex

@@
   \author{Nikos I.
-Bosse\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom} $^{,}$\thanks{Centre for the Mathematical Modelling of Infectious Diseases, London, United Kingdom} $^{,*}$,
+Bosse\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom} $^{,}$\thanks{Centre for the Mathematical Modelling of Infectious Diseases, London, United Kingdom} $^{,}$\thanks{NIHR Health Protection Research Unit in Modelling \& Health Economics} $^{,*}$,
   Sam Abbott\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$, 
   Anne Cori\thanks{MRC Centre for Outbreak Analysis and Modelling, Department of Infectious Disease Epidemiology, School of Public Health, Imperial College London, London, United Kingdom} $^{}$, \\
-  Edwin van Leeuwen\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom} $^{,}$\thanks{Modelling & Economics Unit and NIHR Health Protection Research Unit in Modelling \& Health Economics, UK Health Security Agency, London, United Kingdom} $^{}$,
+  Edwin van Leeuwen\footnotemark[1] $^{,}$\thanks{Modelling \& Economics Unit and NIHR Health Protection Research Unit in Modelling \& Health Economics, UK Health Security Agency, London, United Kingdom} $^{}$,
   Johannes Bracher\thanks{Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology, Karlsruhe, Germany } $^{,}$\thanks{Computational Statistics Group, Heidelberg Institute for Theoretical Studies, Heidelberg, Germany } $^{, \dagger}$,
-  Sebastian Funk\footnotemark[1] $^{, }$\footnotemark[2]$ ^{}$ $ ^{, \dagger}$}
+  Sebastian Funk\footnotemark[1] $^{, }$\footnotemark[2] $^{, }$\footnotemark[3] $ ^{, \dagger}$}
 
 \maketitle
-%\tableofcontents
 \begin{abstract}
-
-%I swapped the commented out version and this one on Wednesday, January 18 2.47pm EST. 
-%Forecast evaluation plays an essential role in improving forecasting models and can support their role in informing decision-making.
-In epidemiology, probabilistic forecasts are often scored using the Weighted Interval Score (WIS) or the Continuous Ranked Probability Score (CRPS), which are measures of absolute distance between forecast and observation.
-Infectious disease transmission is commonly modelled as an exponential process characterized by the growth rate or the effective reproduction number.
-However, if the true objective of the forecaster is to predict future growth rates or reproduction numbers, then scoring based on the absolute distance between forecast and observation may not reflect this underlying task well and thus make it difficult to compare forecasts across time, space, and forecast targets.
-As a solution, we propose transforming the data before scoring the forecasts to better reflect the underlying generative process.
-Specifically, we discuss applying CRPS after log transformation of predictions and observations, which can be interpreted as a measure of relative error and as a score for an implicit forecast of the exponential growth rate.
-Moreover, we discuss its possible role as a variance-stabilizing transformation.
-Applying the natural logarithm to data from the European COVID-19 Forecast Hub, we found that it changed model rankings, regardless of stratification by forecast date, location or target types, and emphasised situations in which models missed lower inflection points and consequent rises in numbers, while penalising missed peaks less severely.
-We conclude that appropriate transformations, of which the natural logarithm is only one possible option, should be considered when assessing the performance of different models in the context of infectious disease incidence.
 
@@
 Instead of a score representing the magnitude of absolute errors, applying a log-transformation prior to the CRPS yields a score which a) measures relative error (see Section \ref{sec:methods:relative}), b) provides a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate}) and c) is less dependent on the expected order of magnitude of the quantity to be predicted (see Section \ref{sec:methods:vst}).
-%The interpretation of the log-transformed score as a score for a forecast of the exponential growth rate of a given quantity aligns the evaluation with our understanding of how epidemiological processes should be modeled and the kinds of errors models would typically make.
 We therefore argue that such evaluations on the logarithmic scale should complement the prevailing evaluations on the natural scale. 
-Other transformations may likewise be of interest. We briefly explore the square root transformation as an alternative transformation. %[and population standardization?]. 
+Other transformations may likewise be of interest. We briefly explore the square root transformation as an alternative transformation. 
 Our analysis mostly focuses on the CRPS (or WIS) as an evaluation metric for probabilistic forecasts, given its widespread use throughout the COVID-19 pandemic.
@@
 In Sections \ref{sec:methods:relative}--\ref{sec:methods:vst} we provide some mathematical intuition on applying the log-transformation prior to evaluating the CRPS, highlighting the connections to relative error measures, the epidemic growth rate and variance stabilizing transformations.
-%We first provide a mathematical intuition of the impact of the log-transformation and its interpretation as relative error (Section \ref{sec:methods:relative}, measure of the accuracy of a forecast of the growth rate (Section \ref{sec:methods:growthrate}) and it's effect as a variance-stabilising transformation that makes can make scores independent of the mean of the forecast target (Section \ref{sec:methods:vst}.
 We then discuss practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}) and the effect of the log-transformation on forecast rankings (Section \ref{sec:methods:rankings}).
-To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  \citep[Section \ref{sec:HUB}]{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
+To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  \citep[ Section \ref{sec:HUB}]{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
 Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues (Section \ref{sec:discussion}).
@@
 
-%The analogy between the WIS on the log scale and the absolute error of a point forecast on the log scale is slightly inaccurate: while the absolute error on the log scale is exactly symmetric (e.g., predicting 2 instead of 1 would give the same relative error as predicting 0.5), this is not entirely true for the WIS, which is computed for a set of prediction intervals.
-Due to the fact that upper and lower bounds of individual prediction intervals are affected slightly differently by the log-transformations, small deviations may occur (compare e.g.
-in Figure \ref{fig:change-in-scores}D the small difference between scores for an observed outcome of $\frac{1}{5}$ and 5).
-
 \subsection{Interpretation as scoring the exponential growth rate}
@@
 Average scores are then dominated by the results achieved for targets with high expected outcomes.
-%Bracher et al. mention this as a desirable feature of the CRPS. 
-% In epidemiological contexts, the variance of the variable of interest usually scales with its mean. 
+
 Specifically, if the predictive distribution for the quantity $Y$ equals the true data-generating process $F$ (an ideal forecast), the expected CRPS is given by \citep{gneitingStrictlyProperScoring2007}
@@
 We found that model rankings changed noticeably.
-%, favouring models like epiforecasts-EpiNow2 which uses a log scale generative process and a negative binomial observation model, and disfavouring models like MUNI-ARIMA which uses a square root transform to stabilise variance.
+
 On the natural scale, missing the peak and overshooting was more severely penalised than missing the nadir and the following upswing in numbers.

### 2023-01-24 main.tex
Updates from Overleaf

@@
 As argued in Section \ref{sec:methods:vst}, the mean-variance, or mean-CRPS, relationship determines which transformation can serve as a VST.
-We can analyse this relationship empirically by running a regression that explains the CRPS as a function of the central estimate of the predictive distribution.
+We can analyse this relationship empirically by running a regression that explains the WIS (which approximates the CRPS) as a function of the central estimate of the predictive distribution.
 We ran the regression
 \begin{equation*}
-    \log[\text{CRPS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], 
+    \log[\text{WIS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], 
 \end{equation*}
@@
 \begin{equation*}
-    \text{CRPS}(F, y) = \exp(\alpha) \times \text{median}(F)^\beta, 
+    \text{WIS}(F, y) = \exp(\alpha) \times \text{median}(F)^\beta, 
 \end{equation*}
-meaning that we estimate a polynomial relationship between the predictive median and achieved CRPS.
+meaning that we estimate a polynomial relationship between the predictive median and achieved WIS.
 Note that we are using predictive medians rather than means as only the former are available in the European COVID-19 Forecast Hub.
-As the CRPS of an ideal forecaster scales with the standard deviation (see Section \ref{sec:methods:vst}), a value of $\beta = 1$ would imply a quadratic median-variance relationship; the natural logarithm could then serve as a VST.
-A value of $\beta 0.5$ would imply a linear median-variance relationship, suggesting the square root as a VST.
+As the WIS/CRPS of an ideal forecaster scales with the standard deviation (see Section \ref{sec:methods:vst}), a value of $\beta = 1$ would imply a quadratic median-variance relationship; the natural logarithm could then serve as a VST.
+A value of $\beta = 0.5$ would imply a linear median-variance relationship, suggesting the square root as a VST.
 We applied the regression to case and death forecasts, pooled across horizons and stratified for one through four-week-ahead forecasts.
@@
 It can be seen that the estimates of $\beta$ always take a value somewhat below 1, implying a slightly sub-quadratic mean-variance relationship.
-The logarithmic transformation should thus approximately stabilize the variance (and CRPS), possibly leading to somewhat higher scores for smaller forecast targets.
-The square-root transformation, on the other hand, can be expected to still lead to higher CRPS values for targets of higher orders of magnitude.
+The logarithmic transformation should thus approximately stabilize the variance (and WIS), possibly leading to somewhat higher scores for smaller forecast targets.
+The square-root transformation, on the other hand, can be expected to still lead to higher WIS values for targets of higher orders of magnitude.
 
@@
 \caption{Coefficients of three regressions for the effect of the magnitude of the median forecast on expected scores. The first regression was 
-$\log[\text{CRPS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], $ where $F$ is the predictive distribution and $y$ the observed value. The second one was 
-$\text{CRPS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},$ where $F_{\log}$ is the predictive distribution for $\log y$.
-The third one was $\text{CRPS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{sqrt{\ }} + \beta_{sqrt{\ }} \cdot \sqrt{(\text{median}(F))},$ where $F_{\sqrt{\ }}$ is the predictive distribution for $\sqrt{y}$.
+$\log[\text{WIS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], $ where $F$ is the predictive distribution and $y$ the observed value. The second one was 
+$\text{WIS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},$ where $F_{\log}$ is the predictive distribution for $\log y$.
+The third one was $\text{WIS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{sqrt{\ }} + \beta_{sqrt{\ }} \cdot \sqrt{(\text{median}(F))},$ where $F_{\sqrt{\ }}$ is the predictive distribution for $\sqrt{y}$.
 }
@@
 \begin{equation*}
-    \text{CRPS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},
+    \text{WIS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},
 \end{equation*}
@@
 \begin{equation*}
-    \text{CRPS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{\sqrt{\ }} + \beta_{\sqrt{\ }} \cdot \sqrt{\text{median}(F)},
+    \text{WIS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{\sqrt{\ }} + \beta_{\sqrt{\ }} \cdot \sqrt{\text{median}(F)},
 \end{equation*} 
 where $F_{\sqrt{\ }}$ is the predictive distribution on the square-root scale.
-A value of $\beta_{\log} = 0$ (or $\beta_{\sqrt{\ } = 0}$, respectively, would imply that scores are independent of the median prediction after the transformation.
+A value of $\beta_{\log} = 0$ (or $\beta_{\sqrt{\ }} = 0$, respectively, would imply that scores are independent of the median prediction after the transformation.
 A value smaller (larger) than 0 would imply that smaller (larger) targets lead to higher scores.
-As can be seen from Table \ref{tab:HUB-regression}, the results indeed indicate that small targets lead to larger average CRPS when using the log transform ($\beta_{\log} < 0$), while the opposite is true for the square-root transform ($\beta_{\sqrt{\ }} > 0$).
+As can be seen from Table \ref{tab:HUB-regression}, the results indeed indicate that small targets lead to larger average WIS when using the log transform ($\beta_{\log} < 0$), while the opposite is true for the square-root transform ($\beta_{\sqrt{\ }} > 0$).
 The results of the three regressions are also displayed in Figure \ref{fig:HUB-regression}.
 In this empirical example, the log transformation thus helps (albeit not perfectly), to stabilise WIS values, and it does so more successfully than the square-root transformation.
-As can be seen from Figure \ref{fig:HUB-regression}, the expected CRPS scores for case targets with medians of 10 and 100,000 differ by more then a factor of ten for the square root transformation, but only a factor of around 2 for the logarithm.
+As can be seen from Figure \ref{fig:HUB-regression}, the expected WIS scores for case targets with medians of 10 and 100,000 differ by more then a factor of ten for the square root transformation, but only a factor of around 2 for the logarithm.
 

### 2023-01-24 main.tex
update citation
Updates from Overleaf

@@
 They are commonly applied directly to predicted and observed incidence counts, but it can be questioned whether this yields the most meaningful results given the exponential nature of epidemic processes and the several orders of magnitude that observed values can span over space and time.
-In this paper, we argue that log transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results.
+In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results.
 We motivate the procedure threefold using the CRPS on log-transformed counts as an example: Firstly, it can be interpreted as a probabilistic version of a relative error.

### 2023-01-28 main.tex

@@
 \bibliographystyle{apalike}
+%\bibliographystyle{vancouver}
 
@@
 Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which are both measures of the absolute distance between the forecast distribution and the observation.
-They are commonly applied directly to predicted and observed incidence counts, but it can be questioned whether this yields the most meaningful results given the exponential nature of epidemic processes and the several orders of magnitude that observed values can span over space and time.
+They are commonly applied directly to predicted and observed incidence counts, but it can be questioned whether this is the optimal procedure for comparing models given the exponential nature of epidemic processes and the several orders of magnitude that observed values can span over space and time.
 In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results.
@@
 \clearpage
-\bibliography{log-or-not.bib, software.bib}
+\bibliography{log-or-not-v2.bib, software.bib}
 

### 2023-01-28 main.tex
Update citations, add formula for log-transformed CRPS

@@
 
-{\footnotesize $^*$ Correspondence to \url{nikos.bosse@lshtm.ac.uk}}
+{\footnotesize $^*$ Correspondence to \url{nikos.bosse@lshtm.ac.uk}}, 
 {\footnotesize $^\dagger$ Contributed equally}
@@
 For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation (referred to as the log-transformtation in the remainder of this manuscript) in the context of epidemic phenomena.
+The CRPS after applying a log-transformation can be computed as follows:
+%
+\begin{linenomath*}
+\begin{equation*}
+    \text{CRPS}(F_{\log}, \log y) = \int_{-\infty}^\infty \left( F_{\log}(x) - 1(x \geq \log y) \right)^2 dx,
+\end{equation*}    
+\end{linenomath*}
+%
+where $F_{\log}$ is the log-transformed predictive distribution.
 Instead of a score representing the magnitude of absolute errors, applying a log-transformation prior to the CRPS yields a score which a) measures relative error (see Section \ref{sec:methods:relative}), b) provides a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate}) and c) is less dependent on the expected order of magnitude of the quantity to be predicted (see Section \ref{sec:methods:vst}).

### 2023-04-27 reply_to_reviews.tex
start working on review responses

\documentclass{article}
\usepackage[]{graphicx}
\usepackage[]{xcolor}
\usepackage{alltt}
\usepackage[left=2.3cm,right=2.8cm, top = 2.2cm, bottom = 3cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\PassOptionsToPackage{hyphens}{url}
\usepackage{url} 
\usepackage[disable]{todonotes}
\usepackage{multicol}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage[colorlinks=false]{hyperref} 
\setlength{\parskip}{\baselineskip}%
\setlength{\parindent}{0pt}%

\newcommand{\red}{\color{red}}
\newcommand{\black}{\color{black}}
\newcommand{\blue}{\color{blue}}

\begin{document}

\black \textbf{Black: our comments}

\red \textbf{Reviewer, not addressed}

\blue \textbf{Reviewer, addressed}

\black
We thank the reviewers for their feedback and thoughtful comments which we feel have helped improve the manuscript substantially. Please see our responses to the individual comments as well as explanations on the changes we have made to the manuscript below.

\section{Reviewer \#1}

\blue
Bosse et al. discuss the important problem of forecast evaluation in the context of epidemiological models. They make a clear and compelling argument for the use of different variance-stabilising transformations to reduce the impact of variation in model outputs and data over several orders of magnitude, a consequence of the exponential nature of epidemic growth (with time-varying growth rate).

I only really have minor, hopefully constructive points to raise — overall I think this is a very nice piece of work that deserves to be published in close to its current form. My big picture and lower level comments and suggestions follow.

\black
Thank you for your kind words.

\red
\subsection{Big picture}
\subsubsection{Proper scoring rules}
As noted by the authors, proper scoring rules are a commonly used concept in evaluating probabilistic predictions. While probably familiar to a reasonable fraction of readers, I think there is also likely a reasonable set of readers who are not aware of these or are only vaguely aware of them. For this reason I think a little more general background would help. For example on line 38 after the ‘report their true belief about nature’ another sentence or two along the lines of

“For example, some methods of ‘scoring’ probabilistic predictions can be ‘gamed’ in the sense that forecasters can do better by reporting a probability distribution different to their best estimate. A proper scoring rule ensures that the expected score when reporting a distribution Q, as evaluated under their actual best estimate of the probability distribution P, is best when Q = P. Thus a proper scoring rule encourages a forecaster to provide ‘honest’ predictions."

In addition, I suspect there are a few readers (such as myself!) who are more familiar with more ‘basic’ scores such as e.g. logarithmic score or deviance (i.e. essentially the likelihood or related quantities). I spent a bit of time wondering about why these weren’t considered, especially as these avoid the scale issues at the centre of the present article. Such scores were only briefly considered in the discussion, in the second-to-last paragraph. I think it would be helpful to mention/emphasise up front some of these ‘classic’ scores that don’t have the same issues with scaling and then discuss why they aren’t used, e.g. due to ‘robustness’ or other issues (e.g. lack of full probability distributions from forecasts).

\subsubsection{Log vs other transformations}
The authors take the position that the log transformation is generally preferable for their use case, though they discuss and compare others. Relatedly, their arguments in terms of relative error also involve an essentially log-linear (i.e. log data + additive error) form which may be pragmatic but somewhat inelegant imo.

As the authors note, a particularly controversial issue for the log transformation is the issue of zero values. They take the standard pragmatic stance that we can use log (eps + y) instead of log y in such cases. Many communities e.g. econometrics are vehemently opposed to this, preferring e.g. quasi-Poisson regression and robust standard errors in the context of estimation. While not completely opposed to log (eps + y) myself (I have used it too!) it makes me a bit uncomfortable I still can’t help but feel that something along the lines of e.g. logarithmic/quasi-likelihood or deviance scores could be formulated that would be preferable to the log data transform. This would offer both automatic scaling and handling of zero values, in principle. However, I haven’t thought carefully enough to offer a concrete alternative and the log(eps + y) approach appears to work reasonably here. Instead, and in combination with the previous point, perhaps a bit more discussion of the potential alternatives based on quasi-likelihood-style functions rather than data transformations could be added?

\subsubsection{Observation and process models}
The interpretation in e.g. 2.2 appears to essentially assume a deterministic process model and additive error on the log scale (multiplicative on the natural scale). Although more of a motivating heuristic than strict assumption, a deterministic process model based on a mean will not in general be the same as a stochastic process model, right? Perhaps a further caveat that this is a fairly simplistic motivating tool might be useful?

Furthermore, why not assume that the model mean (say the output of the deterministic model) defines e.g. the mean of something like a Poisson distribution (probably in overdispersed/quasi form)? or negative binomial? The potential use of these distributions is considered in later sections in the context of motivating variance transforms, but then would again seem to motivate a (quasi-)likelihood-style score beyond the approximate transforms (though with pros and cons in terms of robustness, applicability in the presence of partial information).

\subsection{Specific suggestions}

I realised as I was about comment on some equations that none of the equations are numbered. It would probably be good to number them :-)

\black
We added numbers to all equations. 
\textbf{I now numbered everything. Are we fine with that?}

\blue
The first equation (line 43) uses an unbolded indicator function which isn’t defined (this also appears in the same form in the third equation, line 99). The second equation in contrast uses bold and defines the indicator. The definition should be moved forward and a choice of bold or not made.

\black
We added the definition to equation 1 and used bold throughout the manuscript consistently.

\red
I think the term ‘propriety’ (as in having the property of being proper) should either be explicitly defined or re-worded in terms of ‘being proper’.

No reference is provided for the approximate variance-stabilising properties of the square root transformation.

\black

\red
\section{Reviewer \#2}

Summary
This manuscript develops theory, interpretations, and intuition regarding evalua- tion metrics applied to strictly monotonic transformations of observations and forecasts, with focus on log, shifted log, sqrt, and shifted sqrt transformations, and details the application and impact of using a log1p transformation within an evaluation framework for COVID-19 forecasts.
The analysis is extensive and covers many facets of selecting and applying such transformations, providing great value. However, it
\begin{itemize}
    \item focuses on putting evaluations for different locations and times on an even scale, and does not directly answer the question of how to re-emphasize locations and times of import to forecast consumers, and
    \item the discussion of which locations and times forecasts consumers would want to emphasize seems incomplete.
\end{itemize}
The former point could be addressed by adding additional content regarding re-emphasis. The latter point could be addressed by studying what stakeholders want to emphasize and using that to make a recommendation for a particular transformation or transformation selection rule. Alternatively, both could be addressed simply by more explicitly and prominently noting that
\begin{itemize}
    \item when selecting a transformation, one should consider the desired emphasis, which is likely not an even scale for most/all predictions, and
    \item the “natural logarithm” / log1p selection is not meant as a recommendation but to enable demonstration.
\end{itemize}
Additionally, any additional necessary/recommended elements of a scoring proce- dure should be noted more prominently (e.g., dealing with forecast missingness, outlying forecasts, outlying data, selecting a Box-Cox transformation, selecting a shift ($a$ value), etc.).

\subsection{Major comments}
``Natural logarithm'', log1p, and log($a + x$) are conflated in several places, despite the discussion in Section 2.4 noting that this is an important distinction. The terminology should not be overloaded. (E.g., L20 ``Applying the log transformation''.)

L18:
\begin{itemize}
    \item ``relative error'' --- Is this what stakeholders want? Or does it overemphasize low steady periods too much?
    \item ``under the assumption of quadratic mean-variance relationship'' --- And do we expect/observe this sort of relationship in epidemiological surveillance data? This reads more like an arbitrary assumption that will be taken instead of something that is examined in the manuscript. (From Section 3.3 it seems like the answer is ``roughly but not entirely''.)
    \item These points might be more simply resolved simply by revising L15 ``motivate''.
\end{itemize}
\black e.g., \url{https://www.pnas.org/doi/pdf/10.1073/pnas.2103302119} emphasizes that the use of MAPE is motivated by its straightforward interpretation. Can we find another example?

\red
L150: “it may be desirable” — Is it though? This seems unlikely. Stakeholders likely want to emphasize some locations and times over others, and would benefit from some rules about how to do so while maintaining propriety. (E.g., is scaling by some function of recent observations or the population size okay? Is scaling by the forecast or observed value okay? Or are we limited to monotonic transformations?)

\black
On what is okay: scaling by recent observations is okay and in a way that's what we are doing with the log (very similar to scaling with last value).Scaling by population size is okay as mentioned in discussion. Scaling with forecast or observation is not okay (I could probably find examples for that).

\red
L166: “a negative binomial distribution with size parameter $\theta$” — Is this realistic? If so, what’s the interpretation epidemiologically?

\black
This is an illustrative example. But we could argue via Poisson gamma mixture, similar to the model in \url{https://doi.org/10.1111/rssa.12974}

\red
Figure 3: is a geometric distribution realistic for epidemiological surveillance data?

\black
Agree that it is not. Could add something along the lines of ``this can happen in principle, but only in rather constructed cases, such as for the geometric distribution in Fig 3, which is unlikely to occur in an epi setting.

\red
L236: “better able to illustrate the effects” — Why does this removal help illustrate the effects? Is such a removal procedure required to be able to apply this transformation approach at all? And if so, what are the requirements/considerations for such a removal procedure?

L244: “pairwise comparisons” — The necessity of using an approach to handle missingness when combining evaluations should be noted more prominently from the beginning. The same applies to the outlier removal (if necessary) and a value selection.

Figure 4: The data (not) selected seems to obscure the key period of predicting the start of a case wave, as we expect the CRPS of log data to emphasize missed case increase predictions there more heavily than the natural scale. Deaths are not a substitute because they have access to a leading indicator and the same sort of misses may not be observed. The evaluations here may also miss some overshooting of the case peak. Which rules from Table SI.2 are coming into play and why? Is there a more complete alternative example?

Table 1: The beta for cases and beta for deaths are both around 0.86 individually. But fit together they beta is 0.963; this appears due to trying to fit two different (difficulties of) tasks together that follow different trends, evident from the differing alpha values. A joint fit does not seem appropriate without allowing for a separate alpha per task type. The same thing can be said for the different horizons within cases\&deaths. The fits here seem to better support an argument for something between sd and log.

L279: “higher scores for smaller forecast targets” — This seems undesirable, given argument of Bracher et al. 2021a. Perhaps some forecast consumers would be interested in quality of forecasts for locations with lower activity (e.g., policymakers for smaller populations), but very unlikely would they want to prioritize less-active times in general over more-active times.

\subsection{Minor comments}
\begin{itemize}
    \item L12 “over space and time” — Which do stakeholders care most about?
    \item L15 “log-transformed counts” — Immediately sounds problematic due to the possibility of 0 counts, without mention of a shift or threshold. Additionally, there is a truncated normal distribution in the mix; is it rounded to counts, or are these not all count data?
    \item L23–L24 “more strongly emphasized”, “less severely penalized” — Could note that this is relative to evaluations on the natural scale (not relative to each other).
    \item L24 “is only one” — Seems a bit too similar to “the only one”, “the only”, etc.
    \item L46 “WIS is an approximation of the CRPS” — Is it an approximation of CRPS itself or some (2x?) scaling?
    \item L59 “strength” — The meaning of “strength” is nonintuitive without read-
ing the citation and understanding the ties to multiplicative interventions on R and r (the latter of which seems unrealistic to achieve). Consider surrounding with quotes, explaining more/less, relating to growth per generation, etc.
\end{itemize}

L60: “immunity” $\xrightarrow{}$ “population immunity”?

L69: This might benefit from a short justification (e.g.,
$\left| \frac{x_t e^{(r_t + \epsilon) \Delta t} - x_t e^{r_t \Delta t}}{x_t e^{(r_t - \epsilon) \Delta t} - x_t e^{r_t \Delta t}} \right| = e^{\epsilon\Delta t}$)

L70: Clarify: is this suggesting to make the forecasting targets themselves the result of applying some standardized growth rate estimator to current+future observations around the desired time period?

L99: How are atoms at $\exp x = 0$ handled? How are observations of $y = 0$ handled?

L127: “approximation of the absolute percentage error” — Raises a few questions:
\begin{itemize}
    \item If a relative error metric is desired, why not just divide CRPS/WIS by y then? (Propriety?)
    \item How does using a (shifted) log transform differ from dividing by (a constant plus) (a mean of) the most recent observation(s)?
    \item Is this desired? Though already discussed, seems like the major point of doubt here: are we at risk of overemphasizing very-low-activity periods and very-low-population states where standard-deviation-to- mean ratios are much higher?
\end{itemize}

L127: As APE, RE, and SAPE are defined differently in other sources (e.g., with scaling factor in APE and SAPE, and with y in the denominator and no absolute value in RE), it may be helpful to also indicate sources for all three (e.g., that the RE definition is also in Gneiting, 2011).

L138: “$\Bar{r}$ is” — Consider $\Bar{r}_t$ etc.

L146: “$\Bar{\hat{r}}$” — Should be $\hat{\Bar{r}}$, or consider $\hat{\Bar{r}}_t$.

L158 “delta method” — Delta method would seem to also require the distribution to be somewhat tightly distributed, concentrated within a nearly-linear region of the transformation function. But putting significant mass on a wider range of values would also mean having non-negligible mass on negative values if “approximately normal”; it would not make sense to apply the logarithm or sqrt transformations in these cases. So perhaps the “tightly distributed” assumption can be taken. But this seems a bit nonobvious and a justification/application of the delta method would be helpful here.

L159 — Need to discuss / spell out the point of the above calculation (— that it is the same for all $\mu$, when paired with the same $c$?).

L168: “grows with the variance” — May read like this a linear growth. Consider “grows with the standard deviation”.

L180: “data-generating” — Consider “fitted”/“believed”.

L186: “incentivizes overconfident predictions” — Needs justifica- tion/citation/removal. This single example disproves propriety but doesn’t establish overconfidence as being encouraged in general.

Figure 2: Were any zeros encountered for the negative binomial or Poisson distributions? Was a shifted log transformation used rather than an unshifted log transformation?

L215: “is mainly driven” — Please add pointer to justification (e.g., Figure 7 / discussion).

L231: “only include models which” — Consider “only include the 7 models that”

Figure 5 caption:
\begin{itemize}
    \item “logarithmic” — With what a values? (THe first link to Figure 5 is from before this information was stated.)
    \item “D: [. . . ] all individual scores” — of EuroCOVIDhub-ensemble or any of the 7 included systems?
    \item “relative change” — of what change in scores relative to what quantity?
\end{itemize}

L272: “with the standard deviation” — Also note here that this is for approximately normal data distributions?

L282: “independent” — Consider “linearly independent”.

L288–290: What would stakeholders want this difference to be, and in
which direction? I would guess the sqrt would generally be preferable unless applying some extra transformation after the log to re-emphasize locations/times of interest.

Figure 7 caption: “one score” — Consider clarifying that/whether there is potentially missingness here.

Figure 8: What is the meaning of the arrow colors? Initially it seems to match the direction, but there is also blue arrow pointing downward. If it is indicating the direction, please fix this and consider a third color for indicating unchanged ranks.

The WIS contribution decomposition suggests that an alternative baseline should be used based on differences in transformed data rather than differences in natural scale data. Or, alternatively, that the a value is too low to give reasonable comparisons (as data that is hard/impossible to distinguish from 0 on a plot on the natural scale corresponds with very high penalization of forecasts considering 0 as a possibility). This could (also) be noted in Figure SI.5’s caption.

L340: “convert [...] rate” — Please clarify. Does “multiplicative growth rate” refer to a setup with $\mathbb{E}[_{t+k} | X_t] = X_t \theta^k$, or only a single step $\mathbb{E}[_{t+\text{one step}} | X_t] = X_t \theta^k$? Does the proposed conversion apply to both, or only to the latter?

L341–342: “natural logarithm as a [. . . ] (VST)” — for what type of relevant distribution(s)?

L344: “take differences between forecasts on the log scale” — Clarify.

L345: “divide each forecast by the forecast” — Caveat about current forecast formats should be included here / in following sentences, without (one mention of) the idea about directly soliciting forecasts of week-on-week ratios in between.

L366: “[...] settings” — (to standardize between measurements, then
could potentially scale by importance to stakeholders)

Figure SI.2: Would benefit from some more discussion, and perhaps a plot of slightly different quantities. E.g., to select an appropriate value of a, it seems like a common goal would be to deprioritize errors around values of x low enough that we expect there to be primarily noise rather than signal or that are not of concern to policymakers. For count data, this likely includes the values 0 - -10 in any circumstances; how far beyond 10 this range extends depends on the range of realistic growth rates, the forecast horizon, and cleanliness of the data. To assist in this decision, it seems like a similar plot of the evaluations of a particular absolute and/or relative error might be useful, or discussion of how to read it from the current plot. E.g., to illustrate that the same relative error is penalized roughly the same for all the plotted values of a for values of x above 50, but quite differently for $a = 10$ vs. $a = 0.1$ for values of x near 1.

Figure SI.7:
\begin{itemize}
    \item Also specify sorting variable\&direction on x axis. (Would a point plot vs. the sorting variable be more informative?)
    \item Very informative figure, promote to main text? Also would like to see for different “parts” of an epidemic (e.g., steady low periods, increases, peaks/plateaus, decreases). Deaths figures suggest that balancing scores between locations can be tricky, and both cases and deaths suggests that the investigated transformations can’t do this exactly.
\end{itemize}

\subsubsection{Typographical}
\blue
L96: “transformtation” $\xrightarrow{}$ “transformation”

Figure 2 caption: “CRPS values where computed” $\xrightarrow{}$ “[...] were computed”

L192: “does however influences” $\xrightarrow{}$ “does however influence”

L249: use two backticks on the left and two single quotes on the right in LaTeX code for the quoting here

\black
Thank you very much for pointing these errors out. We have corrected them.

\red
Figure 5: “Loaction” → “Location”

Figure SI.7: “Loaction” → “Location”

In several places, e.g., L201, there are links to SI material, when clicked,
that instead redirect to main text figures.

\black
\red
\section{Reviewer \#3}

In this manuscript the authors investigate how common scoring rules for probabilistic forecasts can be adjusted by transforming observed data and forecast predictions prior to scoring. They provide three clear motivations for applying such transformations prior to scoring, assess the implications of different transformations in great detail, and evaluate the effects of these transformations using real-world COVID-19 forecasts obtained from the European Forecast Hub. Approaches that can help us improve how we measure the utility of epidemic forecasts for decision-makers are definitely needed. This is a very nice and timely contribution to epidemic forecast evaluation, and I only have a few comments.

\blue
1. Line 96, typo: "log-transformtation".

\red
2. Figure 3: out of curiosity, do the rankings of forecasts A and B first switch at the same observed value (somewhere around 7?) in both panels? Zooming in, it looks as though maybe the switch occurs very slightly earlier for the log-transformed scoring.

3. Section 3.2, page 9: remarks concerning Figure 4 refer to months (e.g., May, July) but it's a little difficult to relate them to the figure, because the x-axis only has 6-month breaks and spans two years, so it isn't immediately apparent which year (or both?) the reader should focus on. That being said, it's great to see how the log-transformation yields similar scores when the forecasts miss peaks and troughs.

\blue
4. Line 281: missing closing parenthesis, "(or $\beta_{\sqrt{}} = 0$, respectively".

\red
5. Figure 7: I can appreciate that the correlation between the natural and logarithmic scores decreases over the forecast horizon, as the absolute forecast error grows. But I found panel B more challenging to interpret. I gather the key message is that model rankings are quite consistent between the natural and log scales (panel A) even though the absolute scores differ markedly between the two scales — an effect that increases over the forecast horizon (panel B). Perhaps adding a sentence to this effect in the accompanying text (lines 292-295) would help other readers to avoid my confusion.

\blue
6. Line 296, typo: "Figure Figure 8 shows ...".

\red
7. Section 3.4, page 14 and Figure 8: it's a welcome outcome that the Hub ensemble forecast remained the top model, given that the log-transformation clearly affects the individual model rankings. Presumably it would be reasonable to expect that ensemble forecasts for other regions, pathogens, etc, should be the best (or near-best) model when using log-transformed case counts?

8. Lines 327-329: "A potential downside is that forecast evaluation is unreliable in situations where observed values are zero or very small."

I feel there's an argument to be made here that rather than being a potential limitations, it could be considered an accurate reflection of the inherent uncertainty about the future course of an epidemic when case numbers are very small. This transformation also provides a valuable benefit in these situations, as the authors note a few sentences later: "It also gives higher weight to another type of situation one may care about, namely one in which numbers start to rise from a previously low level". This neatly illustrates the importance (as highlighted by the authors in this manuscript) of using multiple scoring rules to evaluate and compare forecasts.

9. Table SI.2, "Criteria for removing forecasts": forecasts were removed if their median prediction differed greatly from the true value. This should probably be mentioned in the manuscript text, somewhere around line 235 (where Table SI.2 is referenced), to notify the reader that the definition of "erroneous forecasts" includes forecasts that are in extremely poor agreement with the ground truth. Otherwise, I feel that "erroneous" may be open to misinterpretation (e.g., only removing forecasts that predicted negative counts, NaN values, etc).

\end{document}

### 2023-04-27 main.tex
start working on review responses

--- 2023-01-28 main.tex
@@
 \begin{linenomath*}
-\begin{equation*}
-    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,
-\end{equation*}    
+\begin{equation}
+    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - \boldsymbol{1}(x \geq y) \right)^2 dx,
+\end{equation}    
 \end{linenomath*}
-where $y$ is the true observed value and $F$ the cumulative distribution function (CDF) of the predictive distribution.
+where $y$ is the true observed value, $F$ is the cumulative distribution function (CDF) of the predictive distribution, and $\boldsymbol{1}()$ is the indicator function.
 The CRPS can be understood as a generalisation of the absolute error to predictive distributions, and interpreted on the natural scale of the data.
@@
 \begin{linenomath*}
-\begin{equation*}
+\begin{equation}
     \text{WIS}(F, y) = \frac{1}{K} \times \sum_{k = 1}^{K} 2 \times \left[ \boldsymbol{1}(y \leq q_{\tau_k}) - \tau_k \right] \times ( q_{\tau_k} - y), 
-\end{equation*}
+\end{equation}
 \end{linenomath*}
-where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, $K$ is the number of predictive quantiles provided and $\boldsymbol{1}$ is the indicator function.
+where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, $K$ is the number of predictive quantiles provided and $\boldsymbol{1}()$ is the indicator function.
 The WIS can be decomposed into three components, dispersion, overprediction, underprediction, which reflect the width of the forecast and whether it was centred above or below the observed value.
@@
 
-For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation (referred to as the log-transformtation in the remainder of this manuscript) in the context of epidemic phenomena.
+For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation (referred to as the log-transformation in the remainder of this manuscript) in the context of epidemic phenomena.
+For a prediction target with strictly positive support, the CRPS after applying a log-transformation is given by
+%
+\begin{linenomath*}
+\begin{equation}
+    \text{CRPS}(F_{\log}, \log y) = \int_{-\infty}^\infty \left( F_{\log}(x) - \boldsymbol{1}(x \geq \log y) \right)^2 dx.
+\end{equation}    
+\end{linenomath*}
+%
+Here, $y$ is again the observed outcome and $F_{\log}$ is the predictive CDF of the log-transformed outcome, i.e.,
+\begin{linenomath*}
+\begin{equation}
+F_{\log}(x) = F(\exp(x)),
+\end{equation}    
+\end{linenomath*}
+with $F$ the CDF on the original scale.
 Instead of a score representing the magnitude of absolute errors, applying a log-transformation prior to the CRPS yields a score which a) measures relative error (see Section \ref{sec:methods:relative}), b) provides a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate}) and c) is less dependent on the expected order of magnitude of the quantity to be predicted (see Section \ref{sec:methods:vst}).
@@
 \begin{linenomath*}
-\begin{equation*}
+\begin{equation}
 y = \hat{y} + \varepsilon,
-\end{equation*}
+\end{equation}
 \end{linenomath*}
@@
 \begin{linenomath*}
-\begin{equation*}
+\begin{equation}
 \log y = \log \hat{y} + \varepsilon^*,
-\end{equation*}
+\end{equation}
 \end{linenomath*}
@@
 Using that $\log(a) \approx a - 1$ if $a$ is close to 1, we get
-$$
+\begin{linenomath*}
+\begin{equation}
 |\varepsilon^*| = |\log \hat{y} - \log y| = \left|\log\left(\frac{\hat{y}}{y}\right) \right| \ \ \stackrel{\text{if } \hat{y} \ \approx \ y}{\approx} \ \ \left| \frac{\hat{y}}{y} - 1 \right| \ \ = \ \ \left| \frac{\hat{y} - y}{y} \right|.
-$$
+\end{equation}
+\end{linenomath*}
 The absolute error after log transforming is thus an approximation of the \textit{absolute percentage error} \citep[APE,][]{gneitingMakingEvaluatingPoint2011a} as long as forecast and observation are close.
 As we assumed that $\hat{y} \approx y$, we can also interpret it as an approximation of the \textit{relative error} (RE)
-$$
+\begin{linenomath*}
+\begin{equation}
 \left| \frac{\hat{y} - y}{\hat{y}} \right|
-$$
+\end{equation}
+\end{linenomath*}
 and the \textit{symmetric absolute percentage error} (SAPE)
-$$
+\begin{linenomath*}
+\begin{equation}
 \left| \ \frac{\hat{y} - y}{y/2 + \hat{y}/2} \ \right|.
-$$
+\end{equation}
+\end{linenomath*}
 As Figure \ref{fig:SAPE} shows, the alignment with the SAPE is in fact the closest and holds quite well even if predicted and observed value differ by a factor of two or three.
@@
 \begin{linenomath*}
-\begin{equation*}
+\begin{equation}
 \frac{d}{dt}y(t) = r(t)y(t)
-\end{equation*}
+\end{equation}
 \end{linenomath*}
@@
 \begin{linenomath*}
-\begin{equation*}
+\begin{equation}
 y(t) = y_0 \exp \left( \int_0^t r(t') dt' \right) = y_0 \exp (\bar{r}t)
-\end{equation*}
+\end{equation}
 \end{linenomath*}
@@
 \begin{linenomath*}
-\begin{align*}
+\begin{align}
+\begin{split}
 \epsilon^* &= \left| \log \left[ \hat{y}( t ) \right] - \log \left[ y ( t ) \right] \right|\\
@@
            &= t \left| \bar{\hat{r}} - \bar{r} \right|
-\end{align*}
+\end{split}
+\end{align}
 \end{linenomath*}
@@
 Specifically, if the predictive distribution for the quantity $Y$ equals the true data-generating process $F$ (an ideal forecast), the expected CRPS is given by \citep{gneitingStrictlyProperScoring2007}
-$$
+\begin{linenomath*}
+\begin{equation}
 \mathbb{E}[\text{CRPS}(F, y)] = 0.5\times\mathbb{E}|Y - Y'|,
-$$
+\end{equation}
+\end{linenomath*}
 where $Y$ and $Y'$ are independent samples from $F$.
@@
 If $F$ is well-approximated by a normal distribution $\text{N}(\mu, \sigma^2)$, the approximation
-$$
+\begin{linenomath*}
+\begin{equation}
 \mathbb{E}_F[\text{CRPS}(F, y)] \approx \frac{\sigma}{\sqrt{\pi}}
-$$
+\end{equation}
+\end{linenomath*}
 can be used.
@@
 Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method to show that
-$$
+\begin{linenomath*}
+\begin{equation}
 \mathbb{E}_F[\text{CRPS}\{F_{\log}, \log(y)\}] \approx \frac{\sigma/\mu}{\sqrt{\pi}} 
@@
 .
-$$
+\end{equation}
+\end{linenomath*}
 The assumption of a quadratic mean-variance relationship is closely linked to the aspects discussed in Sections \ref{sec:methods:relative} and \ref{sec:methods:growthrate}.
@@
 Denoting by $F_{\sqrt{\ }}$ the predictive distribution for $\sqrt{Y}$, the delta method can again be used to show that
-$$
+\begin{linenomath*}
+\begin{equation}
 \mathbb{E}_F[\text{CRPS}\{F_{\sqrt{\ }}, \sqrt{y}\}] \approx \frac{\sigma/\sqrt{\mu}}{2\sqrt{\pi}} = \frac{\sqrt{c}}{2\sqrt{\pi}}
 .
-$$
+\end{equation}
+\end{linenomath*}
 
@@
 We computed expected CRPS values  for three different distributions, assuming an ideal forecaster with predictive distribution equal to the data-generating distribution.
-    These expected CRPS values where computed for different predictive means based on 10,000 samples each and are represented by dots.
+    These expected CRPS values were computed for different predictive means based on 10,000 samples each and are represented by dots.
 Solid lines show the corresponding approximation of the expected CRPS based on an assumed normal distribution as discussed in section \ref{sec:methods:vst}.
@@
 This represents a strictly monotonic transformation and therefore preserves the propriety of the resulting score.
-The choice of the quantity to add does however influences scores and rankings, as measures of relative errors shrink when adding a constant $a$ to the forecast and the observation.
+The choice of the quantity to add does however influence scores and rankings, as measures of relative errors shrink when adding a constant $a$ to the forecast and the observation.
 We illustrate this in Figure \ref{fig:illustration-effect-log-offset}.
@@
 The relative skill for a given model compared to others is then obtained by taking the geometric mean of all score ratios which involve that model.
-Low values are better, and the "average" model receives a relative skill score of 1.
+Low values are better, and the ``average'' model receives a relative skill score of 1.
 
@@
 We ran the regression
-\begin{equation*}
+\begin{linenomath*}
+\begin{equation}
     \log[\text{WIS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], 
-\end{equation*}
+\end{equation}
+\end{linenomath*}
 where the predictive distribution $F$ and the observation $y$ are on the natural scale. This is equivalent to
-\begin{equation*}
+\begin{linenomath*}
+\begin{equation}
     \text{WIS}(F, y) = \exp(\alpha) \times \text{median}(F)^\beta, 
-\end{equation*}
+\end{equation}
+\end{linenomath*}
 meaning that we estimate a polynomial relationship between the predictive median and achieved WIS.
@@
 To check the relationship after the transformation, we ran the regressions
-\begin{equation*}
+\begin{equation}
     \text{WIS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},
-\end{equation*}
+\end{equation}
 where $F_{\log}$ is the predictive distribution for $\log(y)$, and
-\begin{equation*}
+\begin{equation}
     \text{WIS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{\sqrt{\ }} + \beta_{\sqrt{\ }} \cdot \sqrt{\text{median}(F)},
-\end{equation*} 
+\end{equation} 
 where $F_{\sqrt{\ }}$ is the predictive distribution on the square-root scale.
-A value of $\beta_{\log} = 0$ (or $\beta_{\sqrt{\ }} = 0$, respectively, would imply that scores are independent of the median prediction after the transformation.
+A value of $\beta_{\log} = 0$ (or $\beta_{\sqrt{\ }} = 0$, respectively) would imply that scores are independent of the median prediction after the transformation.
 A value smaller (larger) than 0 would imply that smaller (larger) targets lead to higher scores.
@@
 
-Figure Figure \ref{fig:HUB-rank-order} shows the changes in the ranking between different forecasting models.
+Figure \ref{fig:HUB-rank-order} shows the changes in the ranking between different forecasting models.
 Encouragingly for the European Forecast Hub, the Hub ensemble, which is the forecast the organisers suggest forecast consumers make use of, remains the top model across scoring schemes.
@@
 \begin{linenomath*}
-\begin{align*}
- IS_\alpha(F,y) &= (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u) \\
+\begin{align}
+ IS_\alpha(F,y) &= (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot \boldsymbol{1}(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot \boldsymbol{1}(y \geq u) \\
  &= \text{dispersion} + \text{underprediction} + \text{overprediction},    
-\end{align*}
+\end{align}
 \end{linenomath*}
 %
-where $1()$ is the indicator function, $y$ is the observed value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution, i.e.
+where $\boldsymbol{1}()$ is the indicator function, $y$ is the observed value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution, i.e.
 the lower and upper bound of a single central prediction interval.
@@
 \begin{linenomath*}
-\begin{equation*}
+\begin{equation}
 \text{WIS} = \frac{1}{K^* + 0.5} \cdot \left(w_0 \cdot |y - m| + \sum_{k = 1}^{K^*} w_k \cdot IS_{\alpha_{k}}(F, y)\right),    
-\end{equation*} 
+\end{equation} 
 \end{linenomath*}

### 2023-04-27 R/00 full paper code.R
update citations

@@ -264,13 +264,14 @@ summary <- scores |>
   mutate(log_wis = ifelse(scale == "log", log_wis, NA)) |>
   pivot_wider(values_from = wis, names_from = scale) |>
   group_by(sigma) |>
-  summarise(log_wis = sum(log_wis, na.rm = TRUE), 
-            wis_log = sum(log, na.rm = TRUE), 
-            wis = sum(natural, na.rm = TRUE)) |>
-  pivot_longer(cols = c(wis, log_wis, wis_log), values_to = "score", names_to = "type") |>
+  summarise(log_wis = mean(log_wis, na.rm = TRUE), 
+            wis_log = mean(log, na.rm = TRUE), 
+            wis = mean(natural, na.rm = TRUE), 
+            geom_wis = exp(mean(log(natural), na.rm = TRUE))) |>
+  pivot_longer(cols = c(wis, log_wis, wis_log, geom_wis), values_to = "score", names_to = "type") |>
   mutate(type = factor(type, 
-                       levels = c("wis", "wis_log", "log_wis"), 
-                       labels = c("CRPS", "CRPS (log scale)", "log(CRPS)")))
+                       levels = c("wis", "wis_log", "log_wis", "geom_wis"), 
+                       labels = c("CRPS", "CRPS (log scale)", "log(CRPS)", "geom(wis)")))
 
 
 score_plot <- function(summary) {
@@ -572,7 +573,6 @@ plot_pred_score <- function(hub_data, scores, model,
     value.name = "component_value"
     ) |>
     mutate(target_end_date = as.Date(target_end_date)) |>
-    head(2000) |>
     ggplot(aes(x = target_end_date)) +
     geom_col(
       position = "stack",

### 2023-05-06 R/Refetch data from Forecast Hub.R
save data on anomalies and removed forecasts

@@ -2,7 +2,7 @@
 ##               Code to refetch data from the Forecast Hub.                  ##
 ## ========================================================================== ##
 
-library(covidHubUtils)
+library(covidHubUtils) #remotes::install_github("reichlab/covidhubutils")
 library(dplyr)
 library(data.table)
 library(purrr)
@@ -11,6 +11,8 @@ library(here)
 library(stringr)
 
 analysis_date <- "2022-12-12"
+start_date <- "2021-03-08"
+end_date <- "2022-12-05"
 
 ## -------------------------------------------------------------------------- ##  
 ##                  Download and process Forecast Hub data                    ##
@@ -94,30 +96,27 @@ hub_data <- mutate(hub_data,
 # --------------------------- Filter forecast data --------------------------- #
 
 # filter out forecast anomalies based on the Forecast Hub anomalies file
-filter_out_anomalies <- function(hub_data) {
-  anomalies_file <- here::here("data", "anomalies.csv")
-  if (file.exists(anomalies_file)) {
-    anomalies <- data.table::fread(file = anomalies_file)
-  } else {
-    ## get anomalies file at date of last data made
-    owner <- "covid19-forecast-hub-europe"
-    repo <- "covid19-forecast-hub-europe"
-    path <- "data-truth/anomalies/anomalies.csv"
-    commit <- gh::gh(
-      "/repos/{owner}/{repo}/commits?path={path}&until={date}",
-      owner = owner,
-      repo = repo,
-      path = path,
-      until = analysis_date,
-      .limit = 1
-    )
-    anomalies <- data.table::fread(input = URLencode(URL = paste(
-      "https://raw.githubusercontent.com", owner, repo, commit[[1]]$sha, path, sep = "/"
-    )))
-    data.table::fwrite(x = anomalies, file = anomalies_file)
-  }
+anomalies_file <- here::here("data", "anomalies.csv")
+if (file.exists(anomalies_file)) {
+  anomalies <- data.table::fread(file = anomalies_file)
+} else {
+  ## get anomalies file at date of last data made
+  owner <- "covid19-forecast-hub-europe"
+  repo <- "covid19-forecast-hub-europe"
+  path <- "data-truth/anomalies/anomalies.csv"
+  commit <- gh::gh(
+    "/repos/{owner}/{repo}/commits?path={path}&until={date}",
+    owner = owner,
+    repo = repo,
+    path = path,
[…]

### 2023-05-07 R/00 full paper code.R
Add Figures for the number of anomalies and erroneous forecasts removed

@@ -508,8 +508,12 @@ hub_data$location |>
 
 n_total_obs <- scores$target_end_date |> unique() |> length()
 
+scores$target_end_date |> min()
+scores$target_end_date |> max()
+
+
 # Number of anomalies filtered out. 
-anomalies <- anomalies |>
+anomalies <- fread(here::here("data", "anomalies.csv")) |>
   filter(target_type %in% c("Cases", "Deaths")) |> 
   group_by(location, target_type) |>
   summarise(n = n(), 
@@ -519,9 +523,11 @@ complete_anomalies <- expand.grid(
   location = unique(anomalies$location), 
   target_type = unique(anomalies$target_type))
 
-anomalies |>
+anomalies <- anomalies |>
   full_join(complete_anomalies) %>%
-  replace(is.na(.), 0) |>
+  replace(is.na(.), 0) 
+
+anomalies |>
   ggplot(aes(x = location, y = n)) + 
   geom_bar(inherit.aes = FALSE, aes(y = 92, x = location), 
            stat = "identity", fill = "grey95") + 
@@ -532,6 +538,11 @@ anomalies |>
   labs(y = "Number of observations removed as anomaly", x = "Location")
 ggsave(filename = "output/figures/number-anomalies.png", 
        width = 7, height = 4.5)
+
+anomalies |>
+  group_by(target_type) |>
+  summarise(n = mean(n), 
+            n_rel = mean(n) / 92)
 
 
 

### 2023-05-07 R/Refetch data from Forecast Hub.R

@@ -92,6 +92,7 @@ hub_data <- merge_pred_and_obs(prediction_data, truth,
 hub_data <- mutate(hub_data, 
                    forecast_date = calc_submission_due_date(forecast_date))
 
+fwrite(hub_data, "data/raw-hub-data.csv")
 
 # --------------------------- Filter forecast data --------------------------- #
 

### 2023-05-07 main.tex
update manuscript in response to reviewers

@@
 
-Forecast evaluation plays an essential role in the development cycle of predictive epidemic models and can inform their use for public health decision-making.
+Forecast evaluation is essential for the development of predictive epidemic models and can inform their use for public health decision-making.
 Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which are both measures of the absolute distance between the forecast distribution and the observation.
-They are commonly applied directly to predicted and observed incidence counts, but it can be questioned whether this is the optimal procedure for comparing models given the exponential nature of epidemic processes and the several orders of magnitude that observed values can span over space and time.
+However, applying these scores directly to predicted and observed incidence counts may not be optimal due to the exponential nature of epidemic processes and the varying magnitudes of observed values across space and time.
 In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results.
-We motivate the procedure threefold using the CRPS on log-transformed counts as an example: Firstly, it can be interpreted as a probabilistic version of a relative error.
+We motivate the procedure threefold using the CRPS on log-transformed values as an example: Firstly, it can be interpreted as a probabilistic version of a relative error.
 Secondly, it reflects how well models predicted the time-varying epidemic growth rate.
 And lastly, using arguments on variance-stabilizing transformations, it can be shown that under the assumption of a quadratic mean-variance relationship, the logarithmic transformation leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity.
-Applying the log transformation to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types.
-Situations in which models missed the beginning of upward swings are more strongly emphasized while failing to predict a downturn following a peak is less severely penalized.
+Applying a transformation of log(x + 1) to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types.
+Situations in which models missed the beginning of upward swings are more strongly emphasised while failing to predict a downturn following a peak is less severely penalised when scoring transformed forecasts as opposed to untransformed ones.
 We conclude that appropriate transformations, of which the natural logarithm is only one particularly attractive option, should be considered when assessing the performance of different models in the context of infectious disease incidence.
@@
 Probabilistic forecasts are usually evaluated using so-called proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data.
-Proper scoring rules are constructed such that forecasters (anyone or anything that issues a forecast) are incentivised to report their true belief about the future.
-Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the Weighted Interval Score \citep{bracherEvaluatingEpidemicForecasts2021}.
+Proper scoring rules are constructed such that they cannot be gamed, meaning that forecasters (anyone or anything that issues a forecast) are incentivised to report their true belief about the future.
+In expectation, a forecaster receives the best score if their forecast $F$ coincides with the data-generating distribution $G$, that is if $F = G$.
+Common proper scoring rules are the logarithmic score \citep{goodRationalDecisions1952} or the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}.
+The log score is the log-likelihood of the predictive density at the observed value and is scale-invariant under monotonous transformations.
+The log score does not suffer from the issues we address in this paper, but due to a number of reasons is not widely used in fields such as weather forecasting or epidemiology.
+For example, the log score becomes very unstable for outlier forecasts that assign low density to the observed outcome.
+It is also difficult to apply to forecasts of discrete quantities and doesn't work for forecasts in a quantile-based format.
+More common in epidemiology are the CRPS or its discrete equivalent, the ranked probability score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021}.
 The CRPS measures the distance of the predictive distribution to the observed data as 
@@
 
-The dynamics of infectious processes are often described by the complementary concepts of the reproduction number $R$ \citep{gosticPracticalConsiderationsMeasuring2020} and growth rate $r$ \citep{wallingaHowGenerationIntervals2007}, where $R$ describes the strength and $r$ the speed of epidemic growth \citep{dushoffSpeedStrengthEpidemic2021}.
-In the absence of changes in immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
+The dynamics of infectious processes are often described by the complementary concepts of the reproduction number $R$ \citep{gosticPracticalConsiderationsMeasuring2020} and growth rate $r$ \citep{wallingaHowGenerationIntervals2007}.
+The reproduction number $R$ describes the number of people each infected person is expected to infect in turn (irrespective of the time between infections) and therefore is a measure of the strength of epidemic growth.
+The growth rate $r$ describes the speed of epidemic growth \citep{dushoffSpeedStrengthEpidemic2021}.
+In the absence of changes in population immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
 In that case, the number of new infections in the population grows exponentially in time.
@@
 In this paper, we argue that the aforementioned issues with evaluating epidemic forecasts based on measures of absolute error on the natural scale can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation.
-Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score.
+Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while guaranteeing that the score remains proper.
 Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects of the forecasts that forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).
@@
 When evaluating models across sets of forecasting tasks, it may be desirable for each target to have a similar impact on the overall results.
-In disease incidence forecasting, this is not the case when using the CRPS on the natural scale, as the latter typically scales with the order of magnitude of the quantity to be predicted.
-Average scores are then dominated by the results achieved for targets with high expected outcomes.
+This is true if one assumes that forecasts for different targets provide similar amounts of information about how well a forecaster performs.
+Ideally, one would like the magnitude of the score to scale with the difficulty of the prediction tasks and the information that can be obtained about overall performance of the forecaster.
+In disease incidence forecasting, however, CRPS values on the natural scale typically scale with the order of magnitude of the quantity to be predicted.
+Average scores are then dominated by the results achieved for targets with high expected outcomes in a way that does not necessarily reflect the underlying predictive ability well.
 
@@
 We consider a Poisson distribution where $\sigma^2 = \mu$, a negative binomial distribution with size parameter $\theta = 10$ and thus $\sigma^2 = \mu + \mu^2/10$, and a normal distribution with constant variance.
-We see that when applying the CRPS on the natural scale, the expected CRPS grows with the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster).
+We see that when applying the CRPS on the natural scale, the expected CRPS grows as the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster) increases.
 The expected CRPS is constant only for the distribution with constant variance, and grows in $\mu$ for the other two.
@@
     \caption{Expected CRPS scores as a function of the mean and variance of the forecast quantity.
-We computed expected CRPS values  for three different distributions, assuming an ideal forecaster with predictive distribution equal to the data-generating distribution.
+We computed expected CRPS values  for three different distributions, assuming an ideal forecaster with predictive distribution equal to the true underlying (data-generating) distribution.
     These expected CRPS values were computed for different predictive means based on 10,000 samples each and are represented by dots.
@@
 
-Transformations that are strictly monotonic are permissible in the sense that they maintain the propriety of the score.
+Transformations that are strictly monotonic are permissible in the sense that resulting scores remain proper.
 This is because even though rankings of models may change forecasts will in expectation still minimise their score if they report a predictive distribution that is equal to the data-generating distribution.
 This condition holds for both the log and square root transformations, as well as many others.
-However, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee propriety.
+However, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee that the score remains proper.
 In the case of log transforms, taking the logarithm of the scores, rather than scoring the log-transformed forecasts and data, results in an improper score.
@@
 One common approach to deal with zeros is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm \citep{bellegoDealingLogsZeros2022}.
-This represents a strictly monotonic transformation and therefore preserves the propriety of the resulting score.
+This represents a strictly monotonic transformation and therefore guarantees that the resulting score remains proper.
 The choice of the quantity to add does however influence scores and rankings, as measures of relative errors shrink when adding a constant $a$ to the forecast and the observation.
@@
 The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 5th of December 2022 for reported cases and deaths from COVID-19.
+Target dates range from the 13th of March 2021 to the 10th of December 2022, for a total of 92 weeks.
 See \cite{sherrattPredictivePerformanceMultimodel2022} for a more thorough description of the data.
-We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 46 forecast dates (see Figure \ref{fig:HUB-num-avail-models}).
+We filtered all forecasts submitted to the Hub to only include the seven models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 46 forecast dates (see Figure \ref{fig:HUB-num-avail-models}).
 We removed all observations marked as data anomalies by the European Forecast Hub \citep{sherrattPredictivePerformanceMultimodel2022} as well as all remaining negative observed values.
-In addition, we filtered out erroneous forecasts defined by any of the conditions listed in Table \ref{tab:erroneous}.
-Those forecasts were removed in order to be better able to illustrate the effects of the log-transformation on scores and eliminating distortions caused by outlier forecasters.
+These anomalies made up a relevant fraction of all observations.
+On average across locations, 12.1 out of 92 (13.2\%) observations were removed for cases and 12.4 out of 92 (13.5\%) for deaths.
+Figure \ref{fig:number-anomalies} displays the number of anomalies removed for each location.
+In addition, we filtered out erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table \ref{tab:erroneous}.
+Figure \ref{fig:erroneous-forecasts.png} shows the percentage of forecasts removed for each model.
+Those few (less than 0.2\%  of forecasts for each model) erroneous outlier forecasts had excessive influence on average scores and relative skill scores in a way that was not representative of normal model behaviour.
+We removed them here in order to better illustrate the effects of the log-transformation on scores that one would expect in a 'normal' scenario.
+In a regular forecast evaluation such erroneous forecasts should usually not be removed and would count towards overall model scores.
+
 All predictive quantiles were truncated at 0.
@@
 Missing the peak, i.e.
-predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts in May in Figure \ref{fig:HUB-model-comparison-ensemble}A, B).
+predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts during the peak in May 2022 in Figure \ref{fig:HUB-model-comparison-ensemble}A, B).
 On the log scale, these have less of an influence, as errors are smaller in relative terms (see \ref{fig:HUB-model-comparison-ensemble}C, D).
-Conversely, failure to predict an upswing while numbers are still low, is less severely punished on the natural scale (see forecasts in July in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low.
+Conversely, failure to predict an upswing while numbers are still low, is less severely penalised on the natural scale (see forecasts in July 2021 and to a lesser extent in July 2022 in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low.
 On the log scale, missing lower inflection points tends to lead to more severe penalties (see Figure \ref{fig:HUB-model-comparison-ensemble}C, D)).
@@
 B: Corresponding boxplot (y-axis on log-scale) of all cases and deaths.
-C: Scores for two-week-ahead forecasts from the EuroCOVIDhub-ensemble (averaged across all forecast dates) for different locations, evaluated on the natural as well as the logarithmic scale.
-D: Corresponding boxplots of all individual scores for two-week-ahead predictions.
+C: Scores for two-week-ahead forecasts from the EuroCOVIDhub-ensemble (averaged across all forecast dates) for different locations, evaluated on the natural scale as well as after transforming counts by adding one and applying the natural logarithm.
+D: Corresponding boxplots of all individual scores of the EuroCOVIDhub-ensemble for two-week-ahead predictions.
 E: Boxplots for the relative change of scores for the EuroCOVIDhub-ensemble across forecast horizons.
@@
 All scores were divided by the score for forecast horizon one.
-To enhance interpretability, the range of visible relative changes was restricted to [0.1, 10].}
+To enhance interpretability, the range of visible relative changes in scores (relative to horizon = 1) was restricted to [0.1, 10].}
     \label{fig:HUB-mean-locations}
@@
 where $F_{\sqrt{\ }}$ is the predictive distribution on the square-root scale.
-A value of $\beta_{\log} = 0$ (or $\beta_{\sqrt{\ }} = 0$, respectively) would imply that scores are independent of the median prediction after the transformation.
+A value of $\beta_{\log} = 0$ (or $\beta_{\sqrt{\ }} = 0$, respectively) would imply that scores are linearly independent of the median prediction after the transformation.
 A value smaller (larger) than 0 would imply that smaller (larger) targets lead to higher scores.
@@
     \caption{Correlations of rankings on the natural and logarithmic scale.
-A: Average Spearman rank correlation of scores for individual forecasts from all models.
+A: Average Spearman rank correlation of scores for individual forecasts.
 For every individual target (defined by a combination of forecast date, target type, horizon, location), one score was obtained per model.
-Then, the rank correlation was computed between the scores for all models on the natural scale vs.
-on the log scale.
-All rank correlations were averaged across locations, and target types and stratified by horizon and target type.
+Then, for every forecast target, the Spearman rank correlation was computed between scores on the natural scale and on the log scale for all the models that had made a forecast for that specific target.
+These individual rank correlations were then averaged across locations and time and are displayed stratified by horizon and target types, representing average accordance of model ranks for a single forecast target on the natural and on the log scale.
 B: Correlation between relative skill scores.
-For every forecast horizon and target type, a separate relative skill score was computed per model using pairwise comparisons.
+For every forecast horizon and target type, a separate relative skill score was computed per model using pairwise comparisons, which is a measure of performance of a model relative to the others for a given horizon and target type that accounts for missing values.
 The plot shows the correlation between the relative skill scores on the natural vs.
-on the log scale.}
+on the log scale, representing accordance of overall model performance as judged by scores on the natural and on the log scale.}
     \label{fig:HUB-cors}
@@
 For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}A).
-When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models change considerably (Figure \ref{fig:HUB-cors}B).
+While rankings between forecasters remain similar for a single forecast, this is not true anymore when looking at rankings obtained after averaging scores across multiple forecasts made at different times or in different locations.
+As discussed earlier, scores on the natural and on the log scale penalise errors very differently, e.g.
+when looking at performance during peaks or troughs.
+When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models therefore change considerably (Figure \ref{fig:HUB-cors}B).
 The correlation between relative skill scores also decreases noticeably with increasing forecast horizon.
@@
 B: Change in rankings as determined by relative skill scores when moving from an evaluation on the natural scale to one on the logarithmic scale.
+Red arrows indicate that the relative skill score deteriorated when moving from the natural to the log scale, green arrows indicate they improved.
 C: Relative skill scores based on scores on the log scale.
@@
 A potential downside is that forecast evaluation is unreliable in situations where observed values are zero or very small.
+One could argue that this correctly reflect inherent uncertainty about the future course of an epidemic when numbers are small.
+Users nevertheless need to be aware that this can pose issues in practice.
 Including very small values in prediction intervals (see e.g.
@@
 high scores) when evaluating forecasts on the log scale.
-\cite{bracherEvaluatingEpidemicForecasts2021} argue that the large weight given to forecasts for locations with high incidences is a desirable property, as it reflects performance on the targets we should care about most.
+\cite{bracherEvaluatingEpidemicForecasts2021} argue that it is desirable to give large weight to forecasts for locations with high incidences, as this reflects performance on the targets we should care about most.
 On the other hand, scoring forecasts on the log scale may be less influenced by outliers and better reflect consistent performance across time, space, and forecast targets.
@@
 One obvious option is to apply a population standardization to obtain incidence forecasts e.g., per 100,000 population \citep{abbottEvaluatingEpidemiologicallyMotivated2022}.
-If one is interested in multiplicative, rather than exponential growth rates, one could convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last observed value.
-We suggested using the natural logarithm as a variance-stabilising transformation (VST) or alternatively the square-root transformation in the case of a Poisson distributed variable.
+We suggested using the natural logarithm as a variance-stabilising transformation (VST) for variables that are approximately normally distributed and have a quadratic mean-variance relationship with $\sigma^2 = c \times \mu^2$ (this is e.g.
+approximately true for the negative binomoial distribution and large $\mu$, or alternatively the square-root transformation in the case of a Poisson distributed variable.
 Other VST like the Box-Cox \citep{boxAnalysisTransformations1964} are conceivable as well.
-Another promising transformation would be to take differences between forecasts on the log scale, or alternatively to divide each forecast by the forecast of the previous week (and analogously for observations), in order to obtain forecasts for week-to-week growth rates.
-One could then also ask forecasters to provide estimates of the weekly relative change applied to the latest data and subsequent forecast points directly.
-This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluating the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}.
-This, unfortunately, is not feasible under the current quantile-based format of the Forecast Hubs, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate.
+If one is interested in multiplicative, rather than exponential growth rates, one could convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last value that was observed at the time the forecast was made.
+Forecasters would then implicitly predict a separate multiplicative growth rate from today to horizon 1, 2, etc.
+Instead of dividing by the last observed value, another promising transformation would be to divide each forecast by the forecast of the previous week (and analogously for observations), in order to obtain forecasts for week-to-week growth rates.
+Alternatively, one could also take first differences of values on the log scale.
+This approach would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluating the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}.
+Dividing values by the previous value, unfortunately, is not feasible under the current quantile-based format of the Forecast Hubs, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate.
 However, it may be an interesting approach if predictive samples are available or if quantiles for weekwise growth rates have been collected.
-It is possible to go beyond choosing a single transformation by constructing composite scores as a weighted sum of scores based on different transformations.
-This would make it possible to create custom scores and allow forecast consumers to assign explicit weights to different qualities of the forecasts they might care about.
+It is possible to go beyond choosing a single transformation by constructing composite scores as a weighted sum of scores based on different transformations (potentially after standardising different scores to have the same mean and standard deviation).
+This would make it possible to create custom scores and allow forecast consumers to choose and assign explicit weights to different qualities of the forecasts they might care about.
 
@@
     \includegraphics[width=0.99\textwidth]{output/figures/example-log-first.png}
-    \caption{Illustration of impropriety of log-transformed CRPS.
+    \caption{Illustration of the effect of applying a transformation after scoring.
 We assume $Y \sim \text{LogNormal}(0, 1)$ and evaluate the expected CRPS for predictive distributions $\text{LogNormal}(0, \sigma)$ with varying values of $\sigma \in [0.1, 2]$.
@@
 For the log-transformed CRPS, the optimal value is  0.9, i.e.
-there is an incentive to report a forecast that is too sharp.}
+there is an incentive to report a forecast that is too sharp.
+The score is therefore no longer proper.}
     \label{fig:log-improper}
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/number-anomalies.png}
+    \caption{
+    Number of observed values that were removed as anomalous. The values were marked as anomalous by the European Forecast Hub team. 
+    }
+    \label{fig:number-anomalies}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/erroneous-forecasts.png}
+    \caption{
+    Number of forecasts marked as erroneous and removed.
+Forecasts that were in extremely poor agreement with the observed values were removed from the analysis according to the criteria shown in Table \ref{tab:erroneous}.
+    }
+    \label{fig:erroneous-forecasts}
+\end{figure}
+
 \begin{table}
@@
 The model had zero included in some of its 50 percent intervals (e.g.
-for case forecasts in July), leading to excessive dispersion values on the log scale.
+for case forecasts in July 2021), leading to excessive dispersion values on the log scale.
 One could argue that including zero in the prediction intervals constituted an unreasonable forecast that was rightly penalised, but in general care has to be taken with small numbers.
@@
     \caption{Mean WIS in different locations for different transformations applied before scoring.
+Locations are sorted according to the mean observed value in that location.
 Shown are scores for two-week-ahead forecasts of the EuroCOVIDhub-ensemble.
-On the natural scale (with no transformation prior applying the WIS), scores correlate strongly with the average number of observed values in a given location.
+On the natural scale (with no transformation prior to applying the WIS), scores correlate strongly with the average number of observed values in a given location.
 The same is true for scores obtained after applying a square-root transformation, or after applying a log-transformation with a large offset $a$.

### 2023-05-07 reply_to_reviews.tex
Updates from Overleaf

@@
 
-\red
+\blue
 \subsection{Big picture}
@@
 
+\black
+The paragraph now reads:
+
+Proper scoring rules are constructed such that they cannot be gamed, meaning that forecasters (anyone or anything that issues a forecast) are incentivised to report their true belief about the future.
+In expectation, a forecaster receives the best score if their forecast $F$ coincides with the data-generating distribution $G$, that is if $F = G$.
+Common proper scoring rules are the logarithmic score \citep{goodRationalDecisions1952} or the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}.
+The log score is the log-likelihood of the predictive density at the observed value and is scale-invariant under monotonous transformations.
+The log score does not suffer from the issues we address in this paper, but due to a number of reasons is not widely used in fields such as weather forecasting or epidemiology.
+For example, the log score becomes very unstable for outlier forecasts that assign low density to the observed outcome.
+It is also difficult to apply to forecasts of discrete quantities and doesn't work for forecasts in a quantile-based format.
+More common in epidemiology are the CRPS or its discrete equivalent, the ranked probability score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021}.
+
+\red
 \subsubsection{Log vs other transformations}
@@
 
+\black
+Phew hm. Any good ideas?
+
+\red
 \subsubsection{Observation and process models}
@@
 
+\black
+Any ideas?
+
+\blue
 \subsection{Specific suggestions}
@@
 
-\red
+\blue
 I think the term ‘propriety’ (as in having the property of being proper) should either be explicitly defined or re-worded in terms of ‘being proper’.
 
+\black
+We reworded occurences of the term `propriety'.
+
+\red
 No reference is provided for the approximate variance-stabilising properties of the square root transformation.
@@
 \black
+Johannes, do you have one available?
 
@@
 
-Summary
-This manuscript develops theory, interpretations, and intuition regarding evalua- tion metrics applied to strictly monotonic transformations of observations and forecasts, with focus on log, shifted log, sqrt, and shifted sqrt transformations, and details the application and impact of using a log1p transformation within an evaluation framework for COVID-19 forecasts.
+\textbf{Summary}
+
+This manuscript develops theory, interpretations, and intuition regarding evaluation metrics applied to strictly monotonic transformations of observations and forecasts, with focus on log, shifted log, sqrt, and shifted sqrt transformations, and details the application and impact of using a log1p transformation within an evaluation framework for COVID-19 forecasts.
 The analysis is extensive and covers many facets of selecting and applying such transformations, providing great value. However, it
@@
 \end{itemize}
-Additionally, any additional necessary/recommended elements of a scoring proce- dure should be noted more prominently (e.g., dealing with forecast missingness, outlying forecasts, outlying data, selecting a Box-Cox transformation, selecting a shift ($a$ value), etc.).
+
+\black
+It feels to me rather clear in the paper that we're showing people options and try to give them the tools they need to make a decision for themselves.
+Also we do kind of recommend the natural log.
+Any thoughts here?
+
+True we don't really answer how a forecast consumer could re-emphasise locations / times as they want.
+We kind of do with Figure SI.7 where we show what happens for different offsets.
+We're also not saying what emphasis consumers should care about.
+But then again should we aim to do that?
+Alternatively we could argue that the major point about the log is the growth rate and that the more similar weighting is more an icing of the cake and we are not primarily concerned with re-weighing times / locations.
+
+\red
+Additionally, any additional necessary/recommended elements of a scoring procedure should be noted more prominently (e.g., dealing with forecast missingness, outlying forecasts, outlying data, selecting a Box-Cox transformation, selecting a shift ($a$ value), etc.).
+
+\black
+Hm potentially an interesting point.
+At the moment the paper kind of assumes you know what you're doing and suggests to do it differently.
+One could I guess add a paragraph that walks through all the steps?
+Add it where?
+Maybe this should be done in another piece that talks about best practices?
+
+\red
 
@@
 
+\black
+L20 refers to the abstract. I changed ``Applying the log transformation'' to ``Applying a transformation of log($x + 1$)''.
+
+In a more principled sense I'm not sure overloading the terminology is that much of a problem, since we explicitly analyse that adding 1 doesn't make that much of difference.
+We could make it more explicit though.
+
+\red
 L18:
@@
 \end{itemize}
-\black e.g., \url{https://www.pnas.org/doi/pdf/10.1073/pnas.2103302119} emphasizes that the use of MAPE is motivated by its straightforward interpretation. Can we find another example?
+\black
+
+Johannes: e.g., \url{https://www.pnas.org/doi/pdf/10.1073/pnas.2103302119} emphasizes that the use of MAPE is motivated by its straightforward interpretation. Can we find another example?
+
+Nikos: Since this line is just referring to the abstract, I'm not sure what exactly to change there? Should we replace the word `motivate' with something else? With what?
 
@@
 \black
-On what is okay: scaling by recent observations is okay and in a way that's what we are doing with the log (very similar to scaling with last value).Scaling by population size is okay as mentioned in discussion.
+We think there is a case to be made that each forecast target should have a similar impact on the overall result of the evaluation.
+This is true if the goal of the evaluation is to find the best forecasting model for future use and if one assumes that the difficulty of the forecasting task is similar in different locations.
+Take for example two models that make independent forecasts for cases of COVID-19 in Germany, Luxembourg and Liechtenstein.
+CRPS scores on the natural scale would be dominated by the predictive performance in Germany.
+One model could be a lot better in Luxembourg and Liechtenstein and slightly worse in Germany and end up looking worse in terms of overall CRPS.
+However, it is unclear whether predictive performance in Germany really provides e.g.
+100 times more information about how good a model is at predicting disease dynamics.
+If the question is which model to deploy in another location in the future, then it might be desirable to not have average scores be dominated by a single location.
+
+We have updated the paragraph as such:
+
+\textit{When evaluating models across sets of forecasting tasks, it may be desirable for each target to have a similar impact on the overall results.
+This is true if one assumes that forecasts for different targets provide similar amounts of information about how well a forecaster performs.
+Ideally, one would like the magnitude of the score to scale with the difficulty of the prediction tasks and the information that can be obtained about overall performance of the forecaster.
+In disease incidence forecasting, however, CRPS values on the natural scale typically scale with the order of magnitude of the quantity to be predicted.
+Average scores are then dominated by the results achieved for targets with high expected outcomes in a way that does not necessarily reflect the underlying predictive ability well.}
+
+Regarding the question which transformations are permissible:
+
+Johannes: scaling by recent observations is okay and in a way that's what we are doing with the log (very similar to scaling with last value).Scaling by population size is okay as mentioned in discussion.
 Scaling with forecast or observation is not okay (I could probably find examples for that).
+
+Do we need to update anything in the manuscript?
 
@@
 \black
-This is an illustrative example. But we could argue via Poisson gamma mixture, similar to the model in \url{https://doi.org/10.1111/rssa.12974}
+Nikos: This is an illustrative example and is not meant to be completely realistic.
+Rather, the point is to illustrate that for observations which have a standard deviation that grows linearly with the mean, the log transformation can serve as a variance-stabilising transformation.
+However, negative binomial distributions / Poisson gamma mixtures have been used in the past to model epidemiological processes.
+
+Endo et al.
+(https://pubmed.ncbi.nlm.nih.gov/32685698/ ) for example estimated that the distribution of COVID cases could be well described using a negative binomial distribution with size $\Theta = 0.1$.
+This is a much stronger overdispersion then what we use in our example.
+\textbf{I'm actually somewhat confused about this. I remember Johannes rejecting my initial $\Theta = 0.1$ because he thought it was unrealistic. Am I severely misunderstanding Akira's paper?}
+
+Johannes: But we could argue via Poisson gamma mixture, similar to the model in \url{https://doi.org/10.1111/rssa.12974}
 
@@
 \black
-Agree that it is not.
+Nikos: Again refer to Akira's paper?
+
+Johannes: Agree that it is not.
 Could add something along the lines of ``this can happen in principle, but only in rather constructed cases, such as for the geometric distribution in Fig 3, which is unlikely to occur in an epi setting.
@@
 
+\black
+We updated the paragraph as follows:
+
+\textit{In addition, we filtered out erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table SI.2.
+Those few erroneous outlier forecasts had excessive influence on average scores and relative skill scores in a way that was not representative of normal model behaviour.
+We removed them here in order to better illustrate the effects of the log-transformation on scores that one would expect in a 'normal' scenario.
+In a regular forecast evaluation such outlier forecasts should of course not be removed and count towards overall model scores.}
+
 L244: “pairwise comparisons” — The necessity of using an approach to handle missingness when combining evaluations should be noted more prominently from the beginning.
@@
 
+\black
+I'm not sure where we should note this. This is already more or less the beginning of the example part. 
+And for the theoretical part it's not that relevant, because missingness is somewhat independent of whether we transform forecasts.
+
+We could add a sentence like "This transformation approach can be combined with approaches like pairwise comparisons that handle missingness"? earlier in the introduction?
+
+\red
 Figure 4: The data (not) selected seems to obscure the key period of predicting the start of a case wave, as we expect the CRPS of log data to emphasize missed case increase predictions there more heavily than the natural scale.
@@
 
+\black
+The data that was removed here was data revisions. The omissions are not due to the forecasts removed by the filtering. We added two Figures in the SI to make that clearer.
+
+\red
 Table 1: The beta for cases and beta for deaths are both around 0.86 individually.
@@
 
+\black
+Johannes to the rescue: Any ideas for a smart response?
+
+\blue
 L279: “higher scores for smaller forecast targets” — This seems undesirable, given argument of Bracher et al.
@@
 
+\black 
+We generally agree that giving higher weight to targets with lower incidences may be undesirable in many instances (one important exception is instances where times of low incidences are followed by upswings).
+Empirically, in our example, the additional weight given to smaller forecast targets on the log scale was rather small compared to the excessive weight given to larger forecast targets on the natural scale.
+However we agree that this is a trade-off.
+One potential avenue for future exploration might be a composite score that combines scores on the natural and on the log scale.
+
 \subsection{Minor comments}
-\begin{itemize}
-    \item L12 “over space and time” — Which do stakeholders care most about?
-    \item L15 “log-transformed counts” — Immediately sounds problematic due to the possibility of 0 counts, without mention of a shift or threshold.
+
+\blue
+L12 “over space and time” — Which do stakeholders care most about?
+
+\black
+We think that this is entirely up to the stakeholders and may depend on the setting. We hope that the reasoning and trade-offs involved become sufficiently clear throughout the remainder of the paper.
+
+\blue
+L15 “log-transformed counts” — Immediately sounds problematic due to the possibility of 0 counts, without mention of a shift or threshold.
 Additionally, there is a truncated normal distribution in the mix; is it rounded to counts, or are these not all count data?
-    \item L23–L24 “more strongly emphasized”, “less severely penalized” — Could note that this is relative to evaluations on the natural scale (not relative to each other).
-    \item L24 “is only one” — Seems a bit too similar to “the only one”, “the only”, etc.
-    \item L46 “WIS is an approximation of the CRPS” — Is it an approximation of CRPS itself or some (2x?) scaling?
-    \item L59 “strength” — The meaning of “strength” is nonintuitive without read-
-ing the citation and understanding the ties to multiplicative interventions on R and r (the latter of which seems unrealistic to achieve).
+
+\black
+We changed "log-transformed counts" to "log-transformed values" and updated the abstract to point out that we're applying a transformation of log($x + 1$) to our example data:
+
+\textit{Applying a transformation of log(x + 1) to data and forecasts from the European COVID-19 Forecast Hub}
+
+The truncated normal distribution used as an example in Figure 2 is based directly on samples from the normal distribution, so here no rounding was applied.
+
+\blue
+L23–L24 “more strongly emphasized”, “less severely penalized” — Could note that this is relative to evaluations on the natural scale (not relative to each other).
+
+\black
+Thank you for the suggestion. We updated the abstract as such:
+
+\textit{Situations in which models missed the beginning of upward swings are more strongly emphasised while failing to predict a downturn following a peak is less severely penalised when scoring transformed forecasts as opposed to untransformed ones.}
+
+\red
+L24 “is only one” — Seems a bit too similar to “the only one”, “the only”, etc.
+
+\black
+We hope that the remainder makes it sufficiently clear that other transformations are possible.
+
+\blue
+L46 “WIS is an approximation of the CRPS” — Is it an approximation of CRPS itself or some (2x?) scaling?
+
+\black
+The weighted interval score can is a weighted sum of interval scores for the individual prediction intervals.
+When weighing the individual interval scores for the central $(1 − \alpha)$ prediction intervals by
+$w = \alpha / 2$ (which is generally standard), then the WIS converges exactly to the CRPS for an increasing number of equally spaced prediction intervals. 
+\textbf{Proper scoring rule dragon, attack!}
+
+\blue
+L59 “strength” — The meaning of “strength” is nonintuitive without reading the citation and understanding the ties to multiplicative interventions on R and r (the latter of which seems unrealistic to achieve).
 Consider surrounding with quotes, explaining more/less, relating to growth per generation, etc.
-\end{itemize}
 
+\black
+We updated the sentence to make this clearer. The new sentence is now:
+
+\textit{The reproduction number $R$ describes the number of people each infected person is expected to infect in turn (irrespective of the time between infections) and therefore is a measure of the strength of epidemic growth.
+The growth rate $r$ describes the speed of epidemic growth.}
+
+\blue
 L60: “immunity” $\xrightarrow{}$ “population immunity”?
 
+\black
+We changed this to "population immunity" as suggested.
+
+\red
 L69: This might benefit from a short justification (e.g.,
@@
 
+\red
 L70: Clarify: is this suggesting to make the forecasting targets themselves the result of applying some standardized growth rate estimator to current+future observations around the desired time period?
@@
 
+\black
+Johannes to the rescue!
+
+\red
+
 L138: “$\Bar{r}$ is” — Consider $\Bar{r}_t$ etc.
@@
 
+\blue
 L168: “grows with the variance” — May read like this a linear growth. Consider “grows with the standard deviation”.
+
+\black
+We have modified the sentence, which now reads:
+
+\textit{We see that when applying the CRPS on the natural scale, the expected CRPS grows as the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster) increases.}
+
+As the rest of the paragraph talks about the variance, we kept the word variance here as well.
+
+\blue
 
@@
 
+\black
+In this instance "data-generating" is correct, as this is about a general property of proper scoring rules. We changed the sentence slightly to make this clearer:
+
+\textit{We computed expected CRPS values  for three different distributions, assuming an ideal forecaster with predictive distribution equal to the true underlying (data-generating) distribution.}
+
+\red
 L186: “incentivizes overconfident predictions” — Needs justifica- tion/citation/removal. This single example disproves propriety but doesn’t establish overconfidence as being encouraged in general.
 
+\black
+While not proved here this is a statement that holds generally. Johaaannes? Also maybe we need a reference for this...
+
+\red
 Figure 2: Were any zeros encountered for the negative binomial or Poisson distributions? Was a shifted log transformation used rather than an unshifted log transformation?
 
+\black 
+As indicated in the Figure legend ("Expected CRPS was computed on the natural scale (left), after applying a square-root transformation (middle), and after adding one and applying a log-transformation to the data (right)), we have added 1 before applying the log transformation to deal with zero values.
+
+\red
 L215: “is mainly driven” — Please add pointer to justification (e.g., Figure 7 / discussion).
 
+\blue
 L231: “only include models which” — Consider “only include the 7 models that”
 
+\black
+Thank you for the suggestion, we edited the text accordingly.
+
+\blue
 Figure 5 caption:
@@
 
+\black
+We have updated the caption, which now reads:
+
+\textit{Scores for two-week-ahead forecasts from the EuroCOVIDhub-ensemble (averaged across all forecast dates) for different locations, evaluated on the natural scale as well as after transforming counts by adding one and applying the natural logarithm.
+D: Corresponding boxplots of all individual scores of the EuroCOVIDhub-ensemble for two-week-ahead predictions.
+E: Boxplots for the relative change of scores for the EuroCOVIDhub-ensemble across forecast horizons.
+For any given forecast date and location, forecasts were made for four different forecast horizons, resulting in four scores.
+All scores were divided by the score for forecast horizon one.
+To enhance interpretability, the range of visible relative changes in scores (relative to horizon = 1) was restricted to [0.1, 10].}
+
+\red
+
 L272: “with the standard deviation” — Also note here that this is for approximately normal data distributions?
 
+\black 
+The statement that the CRPS/WIS scales with the standard deviation is true in general.
+I think they have a point though that the beta = 1 etc.
+kind of assumes a normal distribution which we say is approximated well?
+
+\blue
 L282: “independent” — Consider “linearly independent”.
 
+\black
+We changed the sentence accordingly. ANY OBJECTIONS?
+
+\black
+
+\red
 L288–290: What would stakeholders want this difference to be, and in
@@
 
+\black
+That likely depends on the exact purpose stakeholders have in mind.
+In many instances it is presumably not desirable to give large weight to smaller forecast targets, and one may therefore e.g.
+choose to use the sqrt transformation or a different one over the log transformation.
+On the other hand, in our example the log transformation only gives a little bit more weight to smaller forecast targets, which may be acceptable given that the log transformation has nice properties in terms of its interpretation in an epidemiological context.
+
+Do we need to make changes to the text?
+
+\blue
 Figure 7 caption: “one score” — Consider clarifying that/whether there is potentially missingness here.
 
+\black
+We have updated the Figure caption to make it more easily understandable. It now reads:
+
+\textit{Correlations of rankings on the natural and logarithmic scale.
+A: Average Spearman rank correlation of scores for individual forecasts.
+For every individual target (defined by a combination of forecast date, target type, horizon, location), one score was obtained per model.
+Then, for every forecast target, the Spearman rank correlation was computed between scores on the natural scale and on the log scale for all the models that had made a forecast for that specific target.
+These individual rank correlations were then averaged across locations and time and are displayed stratified by horizon and target types, representing average accordance of model ranks for a single forecast target on the natural and on the log scale.
+B: Correlation between relative skill scores.
+For every forecast horizon and target type, a separate relative skill score was computed per model using pairwise comparisons, which is a measure of performance of a model relative to the others for a given horizon and target type that accounts for missing values.
+The plot shows the correlation between the relative skill scores on the natural vs.
+on the log scale, representing accordance of overall model performance as judged by scores on the natural and on the log scale.}
+
+\blue
 Figure 8: What is the meaning of the arrow colors?
@@
 
+\black
+The colour of the error indicates whether or not relative skill scores improved or deteriorated.
+We agree this was at all clear and thank you for pointing this out.
+We updated the plot legend and added a more detailed explanation to the Figure legend:
+
+\textit{Red arrows indicate that the relative skill score deteriorated when moving from the natural to the log scale, green arrows indicate they improved.}
+
+\red
 The WIS contribution decomposition suggests that an alternative baseline should be used based on differences in transformed data rather than differences in natural scale data.
@@
 
+\black
+I'm not entirely sure I understand the comment. Ideas, anyone?
+
+\red
 L340: “convert [...] rate” — Please clarify.
-Does “multiplicative growth rate” refer to a setup with $\mathbb{E}[_{t+k} | X_t] = X_t \theta^k$, or only a single step $\mathbb{E}[_{t+\text{one step}} | X_t] = X_t \theta^k$?
+Does “multiplicative growth rate” refer to a setup with $\mathbb{E}[X_{t+k} | X_t] = X_t \theta^k$, or only a single step $\mathbb{E}[X_{t+\text{one step}} | X_t] = X_t \theta^k$?
 Does the proposed conversion apply to both, or only to the latter?
@@
 
+\black
+We updated the paragraph which now reads:
+
+Johannes is that first sentence correct?
+
+\textit{We suggested using the natural logarithm as a variance-stabilising transformation (VST) for variables that are approximately normally distributed and have a quadratic mean-variance relationship with $\sigma^2 = c \times \mu^2$ (this is e.g.
+approximately true for the negative binomoial distribution and large $\mu$, or alternatively the square-root transformation in the case of a Poisson distributed variable.
+Other VST like the Box-Cox (Box and Cox, 1964) are conceivable as well.
+If one is interested in multiplicative, rather than exponential growth rates, one could convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last value that was observed at the time the forecast was made.
+Forecasters would then implicitly predict a separate multiplicative growth rate from today to horizon 1, 2, etc.
+Instead of dividing by the last observed value, another promising transformation would be to divide each forecast by the forecast of the previous week (and analogously for observations), in order to obtain forecasts for week-to-week growth rates.
+Alternatively, one could also take first differences of values on the log scale.
+This approach would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory for a different approach to evaluating the shape of a forecast, see Srivastava et al., 2022).
+Dividing values by the previous value, unfortunately, is not feasible under the current quantile-based format of the Forecast Hubs, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate.
+However, it may be an interesting approach if predictive samples are available or if quantiles for weekwise growth rates have been collected.
+It is possible to go beyond choosing a single transformation by constructing composite scores as a weighted sum of scores based on different transformations.
+This would make it possible to create custom scores and allow forecast consumers to assign explicit weights to different qualities of the forecasts they might care about.
+}
+
+\blue
 L366: “[...] settings” — (to standardize between measurements, then
@@
 
+\black
+We slightly adapted a sentence in the previous paragraph to make this point clearer:
+
+\textit{It is possible to go beyond choosing a single transformation by constructing composite scores as a weighted sum of scores based on different transformations (potentially after standardising different scores to have the same mean and standard deviation).
+This would make it possible to create custom scores and allow forecast consumers to choose and assign explicit weights to different qualities of the forecasts they might care about.}
+
+\textbf{Johannes, the standardising different scores thing should be kosher, right?}
+
+\red
 Figure SI.2: Would benefit from some more discussion, and perhaps a plot of slightly different quantities.
@@
 
+\black
+I'm not sure I can follow.
+In my mind the plot is there to show you that once your values are like 5 times larger than the offset you choose than the offset doesn't have much of an effect.
+So it's designed to show that for most practical purposes, doing log(x + a) should be mostly equivalent to log(x).
+So to me it's less about choosing an appropriate a and more about showing that probably it's just fine to go with some random small a.
+I think Figure SI.7 is more about "choosing an a".
+
+\red
 Figure SI.7:
 \begin{itemize}
+    \blue
     \item Also specify sorting variable\&direction on x axis. (Would a point plot vs. the sorting variable be more informative?)
+    \red
     \item Very informative figure, promote to main text?
@@
 
+\black
+We added a statement to the legend "Locations are sorted according to the mean observed value in that location."
+
+Thoughts on making this a main figure?
+
+I would expect that one could find a value for a (e.g. 10?) that would eventually yield a more or less flat line.
+
+\blue
 \subsubsection{Typographical}
@@
 
+Figure 5: “Loaction” → “Location”
+
+Figure SI.7: “Loaction” → “Location”
+
 \black
@@
 \red
-Figure 5: “Loaction” → “Location”
-
-Figure SI.7: “Loaction” → “Location”
-
-In several places, e.g., L201, there are links to SI material, when clicked,
-that instead redirect to main text figures.
+In several places, e.g., L201, there are links to SI material, when clicked, that instead redirect to main text figures.
 
 \black
+Not sure why this happens - some Latex error? But also not sure worth fixing as they will do that anyway.
+
 \red
@@
 
+\blue
 In this manuscript the authors investigate how common scoring rules for probabilistic forecasts can be adjusted by transforming observed data and forecast predictions prior to scoring.
@@
 
+\black
+Thank you very much
+
 \blue
@@
 
-\red
+\black
+We corrected the typo
+
+\blue
 2.
@@
 
+\black
+On the natural scale, the switch happens between 6.5 and 6.6 and on the log scale between 6.2 and 6.3.
+
+\blue
 3.
@@
 
+\black
+That's a good point, we made the references in the text clearer and added years.
+
 \blue
@@
 
-\red
+\black
+We corrected the typo
+
+\blue
 5.
@@
 
+\black
+We adapted the text and hopefully made it more clear:
+
+For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}A).
+While rankings between forecasters remain similar for a single forecast, this is not true anymore when looking at rankings obtained after averaging scores across multiple forecasts made at different times or in different locations.
+As discussed earlier, scores on the natural and on the log scale penalise errors very differently, e.g.
+when looking at performance during peaks or troughs.
+When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models therefore change considerably.
+
 \blue
@@
 
-\red
+\black
+We corrected the typo
+
+\blue
 7.
@@
 
+\black
+While more research is needed to conclude this with more confidence, we generally agree with the statement.
+We would also expect ensemble forecasts to do well (as they have done in the post) regardless of whether they are evaluated on the log or on the natural scale.
+
+\blue
 8. Lines 327-329: "A potential downside is that forecast evaluation is unreliable in situations where observed values are zero or very small."
@@
 
+\black
+We added a sentence:
+
+One could argue that this correctly reflect inherent uncertainty about the future course of an epidemic when numbers are small.
+Users nevertheless need to be aware that this can pose issues in practice.
+
+\blue
+
 9.
@@
 
+\black
+This is a good point. We changed the sentence to
+
+In addition, we filtered out erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table SI.2.
+
 \end{document}

### 2023-06-06 reply_to_reviews.tex
go through reviewer comments

@@
 \setlength{\parindent}{0pt}%
+
+\bibliographystyle{apalike}
 
@@
 In addition, I suspect there are a few readers (such as myself!) who are more familiar with more ‘basic’ scores such as e.g.
-logarithmic score or deviance (i.e.
+logarithmic score or arise (i.e.
 essentially the likelihood or related quantities).
@@
 In expectation, a forecaster receives the best score if their forecast $F$ coincides with the data-generating distribution $G$, that is if $F = G$.
-Common proper scoring rules are the logarithmic score \citep{goodRationalDecisions1952} or the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}.
+Common proper scoring rules are the logarithmic or log score \citep{goodRationalDecisions1952} or the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}.
 The log score is the log-likelihood of the predictive density at the observed value and is scale-invariant under monotonous transformations.
-The log score does not suffer from the issues we address in this paper, but due to a number of reasons is not widely used in fields such as weather forecasting or epidemiology.
-For example, the log score becomes very unstable for outlier forecasts that assign low density to the observed outcome.
-It is also difficult to apply to forecasts of discrete quantities and doesn't work for forecasts in a quantile-based format.
-More common in epidemiology are the CRPS or its discrete equivalent, the ranked probability score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021}.
+The log score does not suffer from the issues we address in this paper, but due to a number of reasons is not as widely used in fields such as weather forecasting or epidemiology (see e.g., \citealt{Johansson2019} for an exception).
+For example, the log score assigns particularly severe penalties to occasional misguided forecasts.
+This lack of robustness may limit its practical applicability \cite{bracherEvaluatingEpidemicForecasts2021}.
+Moreover, it is not easily applied to forecasts reported as samples or quantiles, as used in many recent disease forecasting efforts.
+For these reasons, the CRPS or its discrete equivalent, the ranked probability score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021} are more commonly used in epidemiology.
 
@@
 
+\black
+ideas? Maybe we can just respond to this and the next one together.
+
+Otherwise suggestion nikos: 
+In the manuscript we simplify the equations a little bit by looking at the case of a single point forecast, rather than a probabilistic forecast that comes in the form of a predictive distribution.
+We feel this makes the discussion more intuitive.
+We want to emphasise that the log-linear error is really the error that is measured by the score, not the error model that would be assumed by any forecaster making a prediction.
+
+\blue
 As the authors note, a particularly controversial issue for the log transformation is the issue of zero values.
@@
 \black
-Phew hm. Any good ideas?
+Thank you for raising this interesting point.
+We have refered to the log score in our manuscript as a score that is scale invariant under monotonic transformations and mention in the discussion that for the log score the question of which scale to evaluate forecasts on does not arise.
+We feel that the ideas based on quasi-likelihood require more investigation and are beyond the scope of this paper.
 
@@
 \black
-Any ideas?
+It is important to note that the equations in 2.2 do not refer to any particular model used by a forecaster or any model that was fit to data.
+Rather, we define the instantaneous growth rate for a single realised epidemic and then show that scoring forecasts for that on the log scale equates to scoring a forecast of the average growth rate according to this definition.
+The forecaster themselves who would make a forecast could assume any error model or model structure they like.
 
@@
 \black
-Johannes, do you have one available?
+We added a reference to the following book:  \url{https://www.google.fr/books/edition/Generalized_Linear_Models_With_Examples/tBh5DwAAQBAJ?hl=en&gbpv=1&dq=variance-stabilizing+transformation+square+root&pg=PA119&printsec=frontcover}, page 118
 
@@
 \black
-It feels to me rather clear in the paper that we're showing people options and try to give them the tools they need to make a decision for themselves.
-Also we do kind of recommend the natural log.
-Any thoughts here?
 
-True we don't really answer how a forecast consumer could re-emphasise locations / times as they want.
-We kind of do with Figure SI.7 where we show what happens for different offsets.
-We're also not saying what emphasis consumers should care about.
-But then again should we aim to do that?
-Alternatively we could argue that the major point about the log is the growth rate and that the more similar weighting is more an icing of the cake and we are not primarily concerned with re-weighing times / locations.
+We added the following to the discussion to make our intentions clearer regarding the preferences of policy makers: 
+\textit{In this paper, we did not make any particular assumptions about policy makers' priorities and preferences.
+Rather, we aimed to enable users to make an informed choice by showing how different transformations lead to different relative weights for the kinds of prediction errors forecast consumers may care about, such as absolute vs.
+relative errors or the size of penalties for over- vs.
+underprediction.
+In practice, engagement with decision makers is important to determine what their priorities are and how different ways to measure predictive importance should be weighed.}
+
+It is important, however, that changing the relative weight for different times and locations is one effect of the scoring forecasts on the log scale, but not necessarily the main point of the transformation.
+Rather, we do recommend to use evaluate forecasts on the log scale (in addition to the natural scale) as we believe the interpretation as a score of the exponential growth rather is meaningful in an epidemiological context.
+
+Users who are explicitly concerned about the relative weights of different times and locations can e.g.
+find information in Figure SI.7 where we show what happens for different offsets.
+A user could e.g.
+choose an offset such that the desired weights would result for different locations / times.
 
@@
 
-Johannes: e.g., \url{https://www.pnas.org/doi/pdf/10.1073/pnas.2103302119} emphasizes that the use of MAPE is motivated by its straightforward interpretation. Can we find another example?
+The sentence in the abstract now reads: 
+\textit{Using the CRPS on log-transformed values as an example, we list three attractive properties:}
 
-Nikos: Since this line is just referring to the abstract, I'm not sure what exactly to change there? Should we replace the word `motivate' with something else? With what?
+relative error is one option that could be interesting. 
+Johannes: e.g., \url{https://www.pnas.org/doi/pdf/10.1073/pnas.2103302119} emphasizes that the use of MAPE is motivated by its
+straightforward interpretation. 
+Ultimately stakeholders need to decide for themselves what they want. 
+Can we find another example?
+
+Quadratic mean-variance relationship means: we can forecast the growth rate equally well regardless of the observed magnitude at the moment. It's a straightforward motivation.
 
@@
 Scaling with forecast or observation is not okay (I could probably find examples for that).
+Paper Sebastian Lerch?
 
@@
 
-\red
+\blue
 L236: “better able to illustrate the effects” — Why does this removal help illustrate the effects?
@@
 
-\textit{In addition, we filtered out erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table SI.2.
-Those few erroneous outlier forecasts had excessive influence on average scores and relative skill scores in a way that was not representative of normal model behaviour.
+\textit{In addition, we filtered out erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table \ref{tab:erroneous}.
+Figure \ref{fig:erroneous-forecasts.png} shows the percentage of forecasts removed for each model.
+Those few (less than 0.2\%  of forecasts for each model) erroneous outlier forecasts had excessive influence on average scores and relative skill scores in a way that was not representative of normal model behaviour.
 We removed them here in order to better illustrate the effects of the log-transformation on scores that one would expect in a 'normal' scenario.
-In a regular forecast evaluation such outlier forecasts should of course not be removed and count towards overall model scores.}
+In a regular forecast evaluation such erroneous forecasts should usually not be removed and would count towards overall model scores.}
 
@@
 \black
-The data that was removed here was data revisions. The omissions are not due to the forecasts removed by the filtering. We added two Figures in the SI to make that clearer.
+The data that was removed here was largely due to data revisions, not due to forecasts removed by the filtering. 
+We added the percentages of observed values in the main text and added the following two Figures in the SI to make that clearer.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.49\textwidth]{output/figures/number-anomalies.png}
+     \includegraphics[width=0.49\textwidth]{output/figures/erroneous-forecasts.png}
+     \caption{}
+    \label{fig:number-anomalies}
+\end{figure}
 
@@
 Johannes to the rescue: Any ideas for a smart response?
+
+So we remoe the joint fit / the first 7 lines of the table.
 
@@
 
-\red
+\blue
 L24 “is only one” — Seems a bit too similar to “the only one”, “the only”, etc.
@@
 \black
-The weighted interval score can is a weighted sum of interval scores for the individual prediction intervals.
-When weighing the individual interval scores for the central $(1 − \alpha)$ prediction intervals by
+The weighted interval score is a weighted sum of interval scores for the individual prediction intervals.
+When weighing the individual interval scores for the central $1 - \alpha$ prediction intervals by
 $w = \alpha / 2$ (which is generally standard), then the WIS converges exactly to the CRPS for an increasing number of equally spaced prediction intervals. 
-\textbf{Proper scoring rule dragon, attack!}
+\textbf{Proper scoring rule dragon, attack!} JB: No scaling. See figure 2 in \cite{bracherEvaluatingEpidemicForecasts2021}.
 
@@
 
+\black
+Do we want to add this?
+
 \red
@@
 
+\black 
+I am confused...
+I think what we want to say is that it would be preferable to evaluate a forecast on the growth rate if one were available. Luckily, we can do that by using a log transformation.
+
+Say that we have something in the discussion that suggests transforming to growth rates directly.
+
+\red
 L99: How are atoms at $\exp x = 0$ handled? How are observations of $y = 0$ handled?
+
+\black
+In this theoretical discussion we defined prediction targets to have strictly positive support, so these cases are excluded.
+In the later parts we follow the common strategy to add an offset before taking the logarithm.
+
+\red
 
@@
 
+\black
+thoughts?
+
 L146: “$\Bar{\hat{r}}$” — Should be $\hat{\Bar{r}}$, or consider $\hat{\Bar{r}}_t$.
+
+\black
+thoughts?
+
+\red
 
@@
 
+\black
+thoughts?
+
+\red
+JB: This argument is a standard justification for the choice of variance-stabilizing transformations, see e.g., the below excerpt from \textit{Generalized Linear Models With Examples in R} by Dunn and Smyth:
+\begin{quote}
+Suppose that $y$ has a mean-variance relationship defined by the function $V(\mu)$, with $\text{var}(y) = \phi V(\mu)$.
+Then, consider a transformation $y^* = h(y) \approx h(\mu) + h'(\mu)(y - \mu)$, from which it can be inferred that
+$$
+\text{var}(y^*) = \text{var}[h(y)] \approx h'(\mu)^2\text{var}(y).
+$$
+Hence, the transformation $y^* = h(y)$ will approximately stabilize the variance if $h'(\mu)$ is proportional to $\text{var}(y)^{-1/2} = V(\mu)^{-1/2}$.
+When $V(\mu) = \mu^2$ [...], the variance-stabilizing transformation is the logarithm, because then $h'(\mu) = 1/\mu$.
+When $V(\mu) = \mu$, the variance-stabilizing transformation is the square root because $h'(\mu) = 1/\mu^{1/2}$.
+\end{quote}
+
+STILL NEED TO ADD THAT REMARK
+We added a reference to this source. We nonetheless agree with the referee that this approximation will work best if the distribution is tight and added a short remark to this regard.
+
+\red
 L159 — Need to discuss / spell out the point of the above calculation (— that it is the same for all $\mu$, when paired with the same $c$?).
+
+\black
+Not sure I understand this.
 
@@
 
-\textit{We see that when applying the CRPS on the natural scale, the expected CRPS grows as the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster) increases.}
+\textit{We see that when applying the CRPS on the natural scale, the expected CRPS grows monotonically as the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster) increases.}
 
@@
 \black
-While not proved here this is a statement that holds generally. Johaaannes? Also maybe we need a reference for this...
+While not proved here this is a statement that holds generally. Johaaannes? Also maybe we need a reference for this... JB: I am not actually sure about this -- may need to remove.
 
@@
 
-\red
+\blue
 L340: “convert [...] rate” — Please clarify.
@@
 
+\red
 L341–342: “natural logarithm as a [. . . ] (VST)” — for what type of relevant distribution(s)?
 
+\blue
 L344: “take differences between forecasts on the log scale” — Clarify.
@@
 
-\textit{We suggested using the natural logarithm as a variance-stabilising transformation (VST) for variables that are approximately normally distributed and have a quadratic mean-variance relationship with $\sigma^2 = c \times \mu^2$ (this is e.g.
-approximately true for the negative binomoial distribution and large $\mu$, or alternatively the square-root transformation in the case of a Poisson distributed variable.
-Other VST like the Box-Cox (Box and Cox, 1964) are conceivable as well.
+\textit{We suggested using the natural logarithm as a variance-stabilising transformation (VST).
+This is appropriate for variables that are approximately normally distributed and have a quadratic mean-variance relationship with $\sigma^2 = c \times \mu^2$ (this is e.g.
+approximately true for the negative binomoial distribution and large $\mu$.
+
+or alternatively the square-root transformation in the case of a Poisson distributed variable. Other VST like the Box-Cox (Box and Cox, 1964) are conceivable as well.
 If one is interested in multiplicative, rather than exponential growth rates, one could convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last value that was observed at the time the forecast was made.
@@
 
-\textbf{Johannes, the standardising different scores thing should be kosher, right?}
+\textbf{Johannes, the standardising different scores thing should be kosher, right?} JB: not sure
 
@@
 
-In addition, we filtered out erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table SI.2.
+In addition, we filtered out a small number of erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table SI.2.
+
+\clearpage
+\bibliography{log-or-not-v2.bib, software.bib}
 

### 2023-06-06 main.tex
go through reviewer comments

@@
 In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results.
-We motivate the procedure threefold using the CRPS on log-transformed values as an example: Firstly, it can be interpreted as a probabilistic version of a relative error.
+Using the CRPS on log-transformed values as an example, we list three attractive properties: Firstly, it can be interpreted as a probabilistic version of a relative error.
 Secondly, it reflects how well models predicted the time-varying epidemic growth rate.
@@
 In expectation, a forecaster receives the best score if their forecast $F$ coincides with the data-generating distribution $G$, that is if $F = G$.
-Common proper scoring rules are the logarithmic score \citep{goodRationalDecisions1952} or the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}.
+Common proper scoring rules are the logarithmic or log score \citep{goodRationalDecisions1952} or the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}.
 The log score is the log-likelihood of the predictive density at the observed value and is scale-invariant under monotonous transformations.
-The log score does not suffer from the issues we address in this paper, but due to a number of reasons is not widely used in fields such as weather forecasting or epidemiology.
-For example, the log score becomes very unstable for outlier forecasts that assign low density to the observed outcome.
-It is also difficult to apply to forecasts of discrete quantities and doesn't work for forecasts in a quantile-based format.
-More common in epidemiology are the CRPS or its discrete equivalent, the ranked probability score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021}.
+The log score does not suffer from the issues we address in this paper, but due to a number of reasons is not as widely used in fields such as weather forecasting or epidemiology (see e.g., \citealt{Johansson2019} for an exception).
+For example, the log score assigns particularly severe penalties to occasional misguided forecasts.
+This lack of robustness may limit its practical applicability \cite{bracherEvaluatingEpidemicForecasts2021}.
+Moreover, it is not easily applied to forecasts reported as samples or quantiles, as used in many recent disease forecasting efforts.
+For these reasons, the CRPS or its discrete equivalent, the ranked probability score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021} are more commonly used in epidemiology.
+
 The CRPS measures the distance of the predictive distribution to the observed data as 
@@
 If the mean-variance relationship is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST \citep{guerreroTimeseriesAnalysisSupported1993}.
-Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method to show that
+Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method (a first-order Taylor approximation, see e.g., \cite{Dunn2018}), to show that
 \begin{linenomath*}
@@
 We consider a Poisson distribution where $\sigma^2 = \mu$, a negative binomial distribution with size parameter $\theta = 10$ and thus $\sigma^2 = \mu + \mu^2/10$, and a normal distribution with constant variance.
-We see that when applying the CRPS on the natural scale, the expected CRPS grows as the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster) increases.
+We see that when applying the CRPS on the natural scale, the expected CRPS grows monotonically as the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster) increases.
 The expected CRPS is constant only for the distribution with constant variance, and grows in $\mu$ for the other two.
@@
 However, it may be an interesting approach if predictive samples are available or if quantiles for weekwise growth rates have been collected.
+Potentially, the variance stabilising time-series forecasting literature may be a useful source of other transformations for various forecast settings.
+
 It is possible to go beyond choosing a single transformation by constructing composite scores as a weighted sum of scores based on different transformations (potentially after standardising different scores to have the same mean and standard deviation).
@@
 Exploring transformations is a promising avenue for future work that could help bridge the gap between modellers and policymakers by providing scoring rules that better reflect what forecast consumers care about.
-Potentially, the variance stabilising time-series forecasting literature may be a useful source of transformations for various forecast settings.
+In this paper, we did not make any particular assumptions about what policy makers care about.
+Rather, we aimed to enable users to make an informed choice by showing how different transformations lead to different relative weights for the kinds of prediction errors forecast consumers may care about, such as absolute vs.
+relative errors or the size of penalties for over- vs.
+underprediction.
+In practice, engagement with decision makers is important to determine what their priorities are and how different ways to measure predictive importance should be weighed.
+
 We have shown that the natural logarithm transformation can lead to significant changes in the relative rankings of models against each other, with potentially important implications for decision-makers who rely on the knowledge of past performance to make a judgement about which forecasts should inform future decisions.

### 2023-06-06 reply_to_reviews.tex
working through review comments

@@
 \black
-We added a reference to the following book:  \url{https://www.google.fr/books/edition/Generalized_Linear_Models_With_Examples/tBh5DwAAQBAJ?hl=en&gbpv=1&dq=variance-stabilizing+transformation+square+root&pg=PA119&printsec=frontcover}, page 118
+We added a reference to the following book:
+Dunn, P.K., Smyth, G.K.(2018). Generalized Linear Models With Examples in R. Germany: Springer New York, page 118.
+
+\url{https://www.google.fr/books/edition/Generalized_Linear_Models_With_Examples/tBh5DwAAQBAJ?hl=en&gbpv=1&dq=variance-stabilizing+transformation+square+root&pg=PA119&printsec=frontcover}, page 118
 
@@
 \subsection{Major comments}
+\blue
 ``Natural logarithm'', log1p, and log($a + x$) are conflated in several places, despite the discussion in Section 2.4 noting that this is an important distinction.
@@
 \black
-L20 refers to the abstract. I changed ``Applying the log transformation'' to ``Applying a transformation of log($x + 1$)''.
+L20 refers to the abstract. We changed ``Applying the log transformation'' to ``Applying a transformation of log($x + 1$)''.
 
-In a more principled sense I'm not sure overloading the terminology is that much of a problem, since we explicitly analyse that adding 1 doesn't make that much of difference.
-We could make it more explicit though.
+We slightly disagree with the notion that the terminology should not be overloaded.
+Not introducing another term for transformations of $\log_{e}(x + \varepsilon)$ makes the manuscript more readable and accessible to readers.
+In addition, our analyses suggest that adding a small context doesn't make much of a difference in most practical applications.
+We nevertheless completely agree that this should be made clear to the reader and have added the following text:
+
+\textit{For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation in the context of epidemic phenomena.
+We refer to this transformation as 'log-transformation' and to scores that have been log-transformed as scores 'on the log scale' (as opposed to scores 'on the natural scale', which have not been transformed).
+In the theoretical discussion in Section \ref{sec:methods}, 'log-transformation' and 'log scale' generally refer to a transformation of $\log_{e}(x)$.
+For practical applications (Section \ref{sec:HUB}) we also use these terms to describe a transformation of $\log_{e}(x + \varepsilon)$ in order to keep the notation simple.}
 
@@
 
-\red
+\blue
 L166: “a negative binomial distribution with size parameter $\theta$” — Is this realistic? If so, what’s the interpretation epidemiologically?
@@
 \black
-Nikos: This is an illustrative example and is not meant to be completely realistic.
+This is an illustrative example and is not meant to be completely realistic.
 Rather, the point is to illustrate that for observations which have a standard deviation that grows linearly with the mean, the log transformation can serve as a variance-stabilising transformation.
-However, negative binomial distributions / Poisson gamma mixtures have been used in the past to model epidemiological processes.
+However, negative binomial distributions / Poisson gamma mixtures have been used in the past (see e.g.
+\url{https://doi.org/10.1111/rssa.12974}) to model epidemiological processes.
+We agree with the reviewer that a size parameter of 0.1 is not realistic, but still think that the example is valuable as an illustration.
 
-Endo et al.
-(https://pubmed.ncbi.nlm.nih.gov/32685698/ ) for example estimated that the distribution of COVID cases could be well described using a negative binomial distribution with size $\Theta = 0.1$.
-This is a much stronger overdispersion then what we use in our example.
-\textbf{I'm actually somewhat confused about this. I remember Johannes rejecting my initial $\Theta = 0.1$ because he thought it was unrealistic. Am I severely misunderstanding Akira's paper?}
-
-Johannes: But we could argue via Poisson gamma mixture, similar to the model in \url{https://doi.org/10.1111/rssa.12974}
-
-\red
+\blue
 Figure 3: is a geometric distribution realistic for epidemiological surveillance data?
@@
 \black
-Nikos: Again refer to Akira's paper?
-
-Johannes: Agree that it is not.
-Could add something along the lines of ``this can happen in principle, but only in rather constructed cases, such as for the geometric distribution in Fig 3, which is unlikely to occur in an epi setting.
+Again, we agree that this is not a typical example we would expect to see in an epidemiological setting.
+It could happen in principle, but only in rather constructed cases.
+We still believe the example is valuable to illustrate the general point that the rankings between two forecasts can change even for a single observation.
 
@@
 
-\red
+\blue
 Table 1: The beta for cases and beta for deaths are both around 0.86 individually.
@@
 \black
-Johannes to the rescue: Any ideas for a smart response?
-
-So we remoe the joint fit / the first 7 lines of the table.
+We agree with the point, thank you for raising it. We removed the joint fit from the table.
 
@@
 \black
-Do we want to add this?
+This is a great suggestion, thank you. We updated the corresponding paragraph, which now reads:
 
-\red
+\textit{If case numbers are evolving based on an exponential process and the  modelling task revolves around estimating and forecasting the reproduction number or the corresponding growth rate, then evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount.
+Consider a forecast that misses the correct growth rate by either $+\epsilon$ or $-\epsilon$:}
+
+\begin{equation}
+\left| \frac{x_t e^{(r_t + \epsilon) \Delta t} - x_t e^{r_t \Delta t}}{x_t e^{(r_t - \epsilon) \Delta t} - x_t e^{r_t \Delta t}} \right| = e^{\epsilon\Delta t}
+\end{equation}
+
+\textit{For exponential processes errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.}
+
+\black
 L70: Clarify: is this suggesting to make the forecasting targets themselves the result of applying some standardized growth rate estimator to current+future observations around the desired time period?
@@
 \black 
-I am confused...
-I think what we want to say is that it would be preferable to evaluate a forecast on the growth rate if one were available. Luckily, we can do that by using a log transformation.
-
-Say that we have something in the discussion that suggests transforming to growth rates directly.
+Here we aren't suggesting a specific transformation yet, but rather just say that it would be desirable to score a forecast of the growth rate (regardless of how that can be achieved).
+Specific suggestions for how to achieve this (either by taking a log transformation or by using another transformation to obtain growth rates directly as mentioned in the Discussion) are discussed in more detail in the remainder of the text.
+We added the word 'conceptually' to the sentence to make this a bit clearer.
 
@@
 
-\black
-thoughts?
-
 L146: “$\Bar{\hat{r}}$” — Should be $\hat{\Bar{r}}$, or consider $\hat{\Bar{r}}_t$.
@@
 \black
-thoughts?
+We applied the suggested changes.
 
 \red
-
 L158 “delta method” — Delta method would seem to also require the distribution to be somewhat tightly distributed, concentrated within a nearly-linear region of the transformation function.
@@
 
-\red
+\blue
 L215: “is mainly driven” — Please add pointer to justification (e.g., Figure 7 / discussion).
+
+\black
+We added a reference to Figure 7.
 
@@
 
-Johannes is that first sentence correct?
-
 \textit{We suggested using the natural logarithm as a variance-stabilising transformation (VST).
 This is appropriate for variables that are approximately normally distributed and have a quadratic mean-variance relationship with $\sigma^2 = c \times \mu^2$ (this is e.g.
-approximately true for the negative binomoial distribution and large $\mu$.
+approximately true for the negative binomoial distribution and large $\mu$).
 

### 2023-06-06 main.tex
Addressing some review comments

@@
 If case numbers are evolving based on an exponential process and the  modelling task revolves around estimating and forecasting the reproduction number or the corresponding growth rate, then evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount.
-This is because for exponential processes errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
-If one is to measure the ability of forecasts to assess and forecast the underlying infection dynamics, it may thus be more desirable to evaluate errors on the growth rate directly.
+Consider a forecast that misses the correct growth rate by either $+\epsilon$ or $-\epsilon$:
+
+\begin{linenomath*}
+\begin{equation}
+\left| \frac{x_t e^{(r_t + \epsilon) \Delta t} - x_t e^{r_t \Delta t}}{x_t e^{(r_t - \epsilon) \Delta t} - x_t e^{r_t \Delta t}} \right| = e^{\epsilon\Delta t}
+\end{equation}
+\end{linenomath*}
+
+For exponential processes errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate. 
+If one is to measure the ability of forecasts to assess and forecast the underlying infection dynamics, it may thus, conceptually, be more desirable to evaluate errors on the growth rate directly.
 
@@
 
-For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation (referred to as the log-transformation in the remainder of this manuscript) in the context of epidemic phenomena.
+For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation in the context of epidemic phenomena.
+We refer to this transformation as 'log-transformation' and to scores that have been log-transformed as scores 'on the log scale' (as opposed to scores 'on the natural scale', which have not been transformed).
+In the theoretical discussion in Section \ref{sec:methods}, 'log-transformation' and 'log scale' generally refer to a transformation of $\log_{e}(x)$.
+For practical applications (Section \ref{sec:HUB}) we also use these terms to describe a transformation of $\log_{e}(x + \varepsilon)$ in order to keep the notation simple.
 For a prediction target with strictly positive support, the CRPS after applying a log-transformation is given by
@@
 \begin{equation}
-y(t) = y_0 \exp \left( \int_0^t r(t') dt' \right) = y_0 \exp (\bar{r}t)
+y(t) = y_0 \exp \left( \int_0^t r(t') dt' \right) = y_0 \exp (\bar{r_t}t)
 \end{equation}
 \end{linenomath*}
-where $y_0$ is an initial data point and $\bar{r}$ is the mean of the growth rate between the initial time point $0$ and time $t$.
+where $y_0$ is an initial data point and $\bar{r_t}$ is the mean of the growth rate between the initial time point $0$ and time $t$.
 
@@
 \epsilon^* &= \left| \log \left[ \hat{y}( t ) \right] - \log \left[ y ( t ) \right] \right|\\
-           &= \left| \log \left[ y_0 \exp (\bar{\hat{r}} t ) \right] - \log \left[ y_0 \exp (\bar{r}t) \right] \right|\\
-           &= t \left| \bar{\hat{r}} - \bar{r} \right|
+           &= \left| \log \left[ y_0 \exp (\hat{\bar{r_t}} t ) \right] - \log \left[ y_0 \exp (\bar{r_t}t) \right] \right|\\
+           &= t \left| \hat{\bar{r_t}} - \bar{r_t} \right|
 \end{split}
@@
 %
-where $\bar{\hat{r}}$ is the true mean growth rate and $\bar{r}$ is the forecast mean growth rate.
+where $\bar{r_t}$ is the true mean growth rate and $\hat{\bar{r_t}}$ is the forecast mean growth rate.
 We thus evaluate the error in the mean exponential growth rate, scaled by the length of the time period considered.
@@
 This means that the expected CRPS scales roughly with the standard deviation, which in turn typically increases with the mean in epidemiological forecasting.
-In order to make the expected CRPS independent of the expected outcome, a \textit{variance-stabilising transformation} \citep[VST,][]{bartlettSquareRootTransformation1936} can be employed.
+In order to make the expected CRPS independent of the expected outcome, a \textit{variance-stabilising transformation} \citep[VST,][]{bartlettSquareRootTransformation1936, dunnGeneralizedLinearModels2018} can be employed.
 The choice of this transformation depends on the mean-variance relationship of the underlying process.
@@
 
-If the variance is linear with $\sigma^2 = c \times \mu$, as with a Poisson-distributed variable, the square root is known to be a VST. 
+If the variance is linear with $\sigma^2 = c \times \mu$, as with a Poisson-distributed variable, the square root is known to be a VST \citep{dunnGeneralizedLinearModels2018}. 
 Denoting by $F_{\sqrt{\ }}$ the predictive distribution for $\sqrt{Y}$, the delta method can again be used to show that
@@
 Overall model rankings would be expected to differ even more when scores are averaged across multiple forecasts or targets.
-The change in rankings of aggregate scores is mainly driven by the order of magnitude of scores for different forecast targets across time, location and target type and less so by the kind of changes in model rankings for single forecasts discussed above.
+The change in rankings of aggregate scores usually is mainly driven by the order of magnitude of scores for different forecast targets across time, location and target type and less so by the kind of changes in model rankings for single forecasts discussed above (see Figure \ref{fig:HUB-cors} for a practical example).
 Large observations will dominate average CRPS values when evaluation is done on the natural scale, but much less so after log transformation.
@@
 \midrule
-all & all & -1.093 & 0.963 & -0.352 & 0.201 & 0.391 & 0.001\\
-\addlinespace
-all & Cases & 0.036 & 0.858 & 0.043 & 0.201 & 0.751 & -0.033\\
-all & Deaths & -0.884 & 0.868 & 0.273 & 0.121 & 0.436 & -0.023\\
-\addlinespace
-1 & all & -1.402 & 0.923 & 0.320 & 0.088 & 0.318 & -0.014\\
-2 & all & -1.221 & 0.967 & 0.112 & 0.164 & 0.364 & -0.003\\
-3 & all & -1.001 & 0.984 & -0.094 & 0.241 & 0.410 & 0.008\\
-4 & all & -0.757 & 0.986 & 0.000 & 0.299 & 0.469 & 0.015\\
+% all & all & -1.093 & 0.963 & -0.352 & 0.201 & 0.391 & 0.001\\
+% \addlinespace
+% all & Cases & 0.036 & 0.858 & 0.043 & 0.201 & 0.751 & -0.033\\
+% all & Deaths & -0.884 & 0.868 & 0.273 & 0.121 & 0.436 & -0.023\\
+% \addlinespace
+% 1 & all & -1.402 & 0.923 & 0.320 & 0.088 & 0.318 & -0.014\\
+% 2 & all & -1.221 & 0.967 & 0.112 & 0.164 & 0.364 & -0.003\\
+% 3 & all & -1.001 & 0.984 & -0.094 & 0.241 & 0.410 & 0.008\\
+% 4 & all & -0.757 & 0.986 & 0.000 & 0.299 & 0.469 & 0.015\\
 \addlinespace
@@
 One obvious option is to apply a population standardization to obtain incidence forecasts e.g., per 100,000 population \citep{abbottEvaluatingEpidemiologicallyMotivated2022}.
-We suggested using the natural logarithm as a variance-stabilising transformation (VST) for variables that are approximately normally distributed and have a quadratic mean-variance relationship with $\sigma^2 = c \times \mu^2$ (this is e.g.
-approximately true for the negative binomoial distribution and large $\mu$, or alternatively the square-root transformation in the case of a Poisson distributed variable.
+We suggested using the natural logarithm as a variance-stabilising transformation (VST).
+This is appropriate for variables that are approximately normally distributed and have a quadratic mean-variance relationship with $\sigma^2 = c \times \mu^2$ (this is e.g.
+approximately true for the negative binomoial distribution and large $\mu$).
+Alternatively, the square-root transformation can be appropriate in the case of a Poisson distributed variable \citep{dunnGeneralizedLinearModels2018}.
 Other VST like the Box-Cox \citep{boxAnalysisTransformations1964} are conceivable as well.
@@
 Exploring transformations is a promising avenue for future work that could help bridge the gap between modellers and policymakers by providing scoring rules that better reflect what forecast consumers care about.
-In this paper, we did not make any particular assumptions about what policy makers care about.
+{In this paper, we did not make any particular assumptions about policy makers' priorities and preferences.
 Rather, we aimed to enable users to make an informed choice by showing how different transformations lead to different relative weights for the kinds of prediction errors forecast consumers may care about, such as absolute vs.

### 2023-06-06 R/00 full paper code.R

@@ -1158,6 +1158,8 @@ df |>
                       "$\\beta_\\log{}$"), escape = FALSE) |>
   kable_styling()
 
+# we removed the first 7 lines from the table following the suggestion from 
+# a reviewer as the regressions may not be particularly meaningful
 
 
 

### 2023-06-20 reply_to_reviews.tex
More comments from Johannes

@@
 \newcommand{\blue}{\color{blue}}
+
+\newcommand{\indented}{\setlength{\leftskip}{1cm}}
+\newcommand{\notindented}{\setlength{\leftskip}{0cm}}
 
@@
 \black
-The paragraph now reads:
+We updated the paragraph, which now reads:
 
-Proper scoring rules are constructed such that they cannot be gamed, meaning that forecasters (anyone or anything that issues a forecast) are incentivised to report their true belief about the future.
-In expectation, a forecaster receives the best score if their forecast $F$ coincides with the data-generating distribution $G$, that is if $F = G$.
+\indented
+
+Proper scoring rules are constructed such that they encourage honest forecasting and cannot be `gamed' or `cheated'.
+% In expectation, a forecaster receives the best score if their forecast $F$ coincides with the data-generating distribution $G$, that is if $F = G$.
+Assuming that the forecaster's actual best judgement corresponds to a predictive distribution $F$, a proper score is constructed such that under $F$, no other distribution $G$ yields a better expected score.
+Forecasters (anyone or anything that issues a forecast) are thus incentivised to report their true belief $F$ about the future.
 Common proper scoring rules are the logarithmic or log score \citep{goodRationalDecisions1952} or the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}.
@@
 
-\red
+\notindented
+
+\blue
 \subsubsection{Log vs other transformations}
@@
 \black
-ideas? Maybe we can just respond to this and the next one together.
-
-Otherwise suggestion nikos: 
-In the manuscript we simplify the equations a little bit by looking at the case of a single point forecast, rather than a probabilistic forecast that comes in the form of a predictive distribution.
+In the theoretical part of the manuscript we simplify the equations a little bit by looking at the case of a single point forecast, rather than a probabilistic forecast that comes in the form of a predictive distribution.
 We feel this makes the discussion more intuitive.
@@
 Thank you for raising this interesting point.
-We have refered to the log score in our manuscript as a score that is scale invariant under monotonic transformations and mention in the discussion that for the log score the question of which scale to evaluate forecasts on does not arise.
-We feel that the ideas based on quasi-likelihood require more investigation and are beyond the scope of this paper.
+The log score is indeed an alternative which is scale invariant under monotonic transformations.
+We mention this in the introduction and the the discussion and state that for the log score the question of which scale to evaluate forecasts on does not arise.
+In our understanding, a quasi-likelihood approach (as opposed to an actual likelihood approach as in the log score) may not actually be necessary as at least in theory full predictive distributions are to be evaluated.
+Developing further ideas based on some form of quasi-likelihood would likely require more in-depth investigation.
+In any case, developing these ideas would require more in-depth investigation and we feel this is beyond the scope of the current paper.
+% and therefore did not include it in this work.
 
-\red
+\blue
 \subsubsection{Observation and process models}
@@
 \black
-We added numbers to all equations. 
-\textbf{I now numbered everything. Are we fine with that?}
+This is a good point, we added numbers to all equations.
 
@@
 \black
-We added the definition to equation 1 and used bold throughout the manuscript consistently.
+We added the definition to equation according to your suggestion 1 and used bold throughout the manuscript consistently.
 
@@
 \black
-We reworded occurences of the term `propriety'.
+We reworded all occurrences of the term `propriety' in the text.
 
-\red
+\blue
 No reference is provided for the approximate variance-stabilising properties of the square root transformation.
@@
 
-\url{https://www.google.fr/books/edition/Generalized_Linear_Models_With_Examples/tBh5DwAAQBAJ?hl=en&gbpv=1&dq=variance-stabilizing+transformation+square+root&pg=PA119&printsec=frontcover}, page 118
-
-\red
+\blue
 \section{Reviewer \#2}
@@
 This manuscript develops theory, interpretations, and intuition regarding evaluation metrics applied to strictly monotonic transformations of observations and forecasts, with focus on log, shifted log, sqrt, and shifted sqrt transformations, and details the application and impact of using a log1p transformation within an evaluation framework for COVID-19 forecasts.
-The analysis is extensive and covers many facets of selecting and applying such transformations, providing great value. However, it
+The analysis is extensive and covers many facets of selecting and applying such transformations, providing great value.
+
+However, it
 \begin{itemize}
@@
 
+Thank you for your thorough and extensive comments.
+
 We added the following to the discussion to make our intentions clearer regarding the preferences of policy makers:
-\textit{In this paper, we did not make any particular assumptions about policy makers' priorities and preferences.
+
+\indented
+In this paper, we did not make any particular assumptions about policy makers' priorities and preferences.
 Rather, we aimed to enable users to make an informed choice by showing how different transformations lead to different relative weights for the kinds of prediction errors forecast consumers may care about, such as absolute vs.
@@
 underprediction.
-In practice, engagement with decision makers is important to determine what their priorities are and how different ways to measure predictive importance should be weighed.}
+In practice, engagement with decision makers is important to determine what their priorities are and how different ways to measure predictive importance should be weighed.
 
-It is important, however, that changing the relative weight for different times and locations is one effect of the scoring forecasts on the log scale, but not necessarily the main point of the transformation.
-Rather, we do recommend to use evaluate forecasts on the log scale (in addition to the natural scale) as we believe the interpretation as a score of the exponential growth rather is meaningful in an epidemiological context.
+\notindented
+It is important to note that changing the relative weight for different times and locations is one effect of the scoring forecasts on the log scale, but not necessarily the main point of the transformation.
+Rather, we do recommend to evaluate forecasts on the log scale as we believe the interpretation as a score of the exponential growth rate is meaningful in an epidemiological context.
+As stated explicitly in the manuscript, we moreover recommend to use this in complement to rather than as a replacement of the common evaluation on the ``natural scale''.
+Evaluators will then becoma aware of whether their results depend strongly on this decision and the implied weighting of different forecasting tasks.
 
 Users who are explicitly concerned about the relative weights of different times and locations can e.g.
-find information in Figure SI.7 where we show what happens for different offsets.
+find information in Figure SI.7 where we show how different forecast targets are emphasized by evaluation on the natural or log-transformed scale.
 A user could e.g.
-choose an offset such that the desired weights would result for different locations / times.
+reproduce such a display with different values of the offset $a$ and choose $a$ such that the relative weighting of times and regions with high and low incidence correspond to their preferences.
 
@@
 \black
+
+To direct the readers' attention to the mentioned challenges already in the conceptual part of the paper we added the following paragraph to Section 2.4, which also contains a discussion of the choice of $a$:
+
+\indented
+We note that in the practical evaluation of operational forecasting systems  several additional challenges arise, which we do not study in detail.
+These concern e.g., the removal of outlying observations and forecasts and the handling of missing forecasts.
+The solutions we employed in practice are provided in Section 3.1.
+
+\notindented
+
+\red not sure what to make of Box-Cox.
+
 Hm potentially an interesting point.
@@
 
-\red
+\blue
+\subsection{Major comments}
 
-\subsection{Major comments}
-\blue
 ``Natural logarithm'', log1p, and log($a + x$) are conflated in several places, despite the discussion in Section 2.4 noting that this is an important distinction.
@@
 \black
-L20 refers to the abstract. We changed ``Applying the log transformation'' to ``Applying a transformation of log($x + 1$)''.
+We changed ``Applying the log transformation'' to ``Applying a transformation of log($x + 1$)'' in L20 (in the abstract).
 
@@
 
-\textit{For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation in the context of epidemic phenomena.
-We refer to this transformation as 'log-transformation' and to scores that have been log-transformed as scores 'on the log scale' (as opposed to scores 'on the natural scale', which have not been transformed).
-In the theoretical discussion in Section \ref{sec:methods}, 'log-transformation' and 'log scale' generally refer to a transformation of $\log_{e}(x)$.
-For practical applications (Section \ref{sec:HUB}) we also use these terms to describe a transformation of $\log_{e}(x + \varepsilon)$ in order to keep the notation simple.}
+\indented
+For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation in the context of epidemic phenomena.
+We refer to this transformation as 'log-transformation' and to scores that have been computed from log-transformed forecasts and observations as scores 'on the log scale' (as opposed to scores 'on the natural scale', which involve no transformation).
+In the theoretical discussion in Section 2, 'log-transformation' and 'log scale' generally refer to a transformation of $\log_{e}(x)$.
+For practical applications (Section 3) we also use these terms to describe a transformation of $\log_{e}(x + \varepsilon)$ in order to keep the terminology and notation simple.
 
-\red
+\notindented
+
+\blue
 L18:
@@
 
-The sentence in the abstract now reads: 
-\textit{Using the CRPS on log-transformed values as an example, we list three attractive properties:}
+We revised the word ``motivate'' in line 15. The sentence in the abstract now reads:
 
-relative error is one option that could be interesting. 
-Johannes: e.g., \url{https://www.pnas.org/doi/pdf/10.1073/pnas.2103302119} emphasizes that the use of MAPE is motivated by its
+\indented
+
+Using the CRPS on log-transformed values as an example, we list three attractive properties: [...].
+
+\notindented
+
+We think the relative error is one option out of many that could be interesting to policy makers. 
+Paireau et al. (2022, \url{https://www.pnas.org/doi/pdf/10.1073/pnas.2103302119}) for example specifically emphasize that in their work the use of MAPE is motivated by its
 straightforward interpretation.
-Ultimately stakeholders need to decide for themselves what they want. 
-Can we find another example?
+But we agree that ultimately stakeholders need to decide for themselves what metrics they are interested in.
+As mentioned above, adjusting the offset before taking the logarithm is one way of dealing with periods of low counts, but this of course moves the interpretation of the score away from a score of the exponential growth rate.
 
-Quadratic mean-variance relationship means: we can forecast the growth rate equally well regardless of the observed magnitude at the moment. It's a straightforward motivation.
+The assumption of a quadratic mean-variance relationship is not entirely arbitrary, but rather has a specific interpretation in the sense that it encodes the assumption that one can forecast the growth rate equally well regardless of the observed magnitude at the moment.
+In the manuscript, we state:
 
-\red
+\indented
+The assumption of a quadratic mean-variance relationship is closely linked to the aspects discussed in Sections 2.1 and 2.2.
+It implies that relative errors have constant variance and can thus be meaningfully compared across different targets.
+Also, it arises naturally if we assume that our capacity to predict the epidemic growth rate does not depend on the expected outcome, i.e.
+does not depend on the current phase of the epidemic or the order of magnitude of current observations.
+
+\notindented
+
+\blue
 L150: “it may be desirable” — Is it though?
@@
 \black
-We think there is a case to be made that each forecast target should have a similar impact on the overall result of the evaluation.
-This is true if the goal of the evaluation is to find the best forecasting model for future use and if one assumes that the difficulty of the forecasting task is similar in different locations.
+We think that while it is certainly not the only relevant approach, there is a case to be made for an evaluation which gives each foreasting target a similar weight.
+This is attractive if the goal of the evaluation is to find the best forecasting model for future use and if one assumes that the difficulty of the forecasting task is similar in different locations.
 Take for example two models that make independent forecasts for cases of COVID-19 in Germany, Luxembourg and Liechtenstein.
@@
 
-\textit{When evaluating models across sets of forecasting tasks, it may be desirable for each target to have a similar impact on the overall results.
-This is true if one assumes that forecasts for different targets provide similar amounts of information about how well a forecaster performs.
-Ideally, one would like the magnitude of the score to scale with the difficulty of the prediction tasks and the information that can be obtained about overall performance of the forecaster.
-In disease incidence forecasting, however, CRPS values on the natural scale typically scale with the order of magnitude of the quantity to be predicted.
-Average scores are then dominated by the results achieved for targets with high expected outcomes in a way that does not necessarily reflect the underlying predictive ability well.}
+\indented
+When evaluating models across sets of forecasting tasks, it may be desirable for each target to have a similar impact on the overall results.
+This could be motivated by the assumption that forecasts from different geographical units and time periods provide similar amounts of information about how well a forecaster performs.
+One would then like the resulting scores to be independent of the order of magnitude of the target to predict.
+CRPS values on the natural scale, however, typically scale with the order of magnitude of the quantity to be predicted.
+Average scores are then dominated by the results achieved for targets with high expected outcomes in a way that does not necessarily reflect the underlying predictive ability well.
+
+\notindented
 
@@
 
-Johannes: scaling by recent observations is okay and in a way that's what we are doing with the log (very similar to scaling with last value).Scaling by population size is okay as mentioned in discussion.
-Scaling with forecast or observation is not okay (I could probably find examples for that).
-Paper Sebastian Lerch?
-
-Do we need to update anything in the manuscript?
+Scaling by past observations is permissible (indeed taking the log is similar to scaling by the last value), as is scaling by the population size as mentioned in the discussion.
+There may of course be exceptions such as dividing by zero, but otherwise this just represents a monotonic transformation, which is unproblematic in terms of propriety.
+Scaling a forecast by the later observed value (as opposed to scaling by past observations) is generally not permissible as it disturbs propriety; see Lerch et al (2017, Statistical Science, preprint: \url{https://arxiv.org/abs/1512.09244}) on the closely related topic of weighting scores with a function of the observed value.
+Similarly, scaling a forecast by predictions made by other forecasters may break propriety, but we are unaware of existing theoretical arguments.
+Scaling scores, e.g.
+dividing a score by the mean score for other forecasters may be acceptable in most applied settings, but we are unsure about exactly when and how this could break propriety (as is the case for example when dividing by zero).
 
@@
 Rather, the point is to illustrate that for observations which have a standard deviation that grows linearly with the mean, the log transformation can serve as a variance-stabilising transformation.
-However, negative binomial distributions / Poisson gamma mixtures have been used in the past (see e.g.
-\url{https://doi.org/10.1111/rssa.12974}) to model epidemiological processes.
-We agree with the reviewer that a size parameter of 0.1 is not realistic, but still think that the example is valuable as an illustration.
+However, negative binomial distributions are commonly used in the epidmeiological literature; see e.g., Section 3 of Meyer and Held (2017, \url{http://dx.doi.org/10.1214/14-AOAS743}).
+An epidemiological interpretation could be the following (see e.g.
+Funk, Abbott and Bracher \url{https://doi.org/10.1111/rssa.12974}):
+\begin{itemize}
+    \item For simplicity assume a generation time of one time interval (this could be relaxed).
+    \item Each of the $Y_{t - 1}$ infected individuals at time $t - 1$  generates new infections at time $t$, with a reproduction number of $R_t$ and independent Poisson offspring distributions.
+Assuming independence across infecting individuals and given $R_t$ and $Y_{t - 1}$, $Y_t$ then follows a Poisson distribution with mean $R_t \times Y_{t - 1}$.
+    \item We assume that $R_t$ is not fixed, but given by
+    $$
+    R_t = R_0 \times \theta_t
+    $$
+    where
+    $$
+    \theta \sim \text{Gamma}(\text{shape} = \psi, \text{rate} = \psi).
+    $$
+    \item This implies
+    $$
+    Y_t \mid Y_{t - 1} \sim \text{NegBin}(R_t \times Y_{t - 1}, \psi).
+    $$
+\end{itemize}
+So a conditional negative binomial distribution with quadratic mean-variance relationship arises if we assume that a suitably distributed environmental factor $\theta_t$ modifies the reproductive number.
+While the assumption of this specific gamma distribution is arbitrary, a quadratic mean-variance relationship will also arise for other distributions of $\theta_t$.
+
+This illustrates that a quadratic mean-variance relationship can arise from a simple yet not completely unreasonable renewal equation model.
+
+% to model epidemiological processes. We agree with the reviewer that a size parameter of 0.1 is not realistic, but still think that the example is valuable as an illustration.
 
@@
 \black
-Again, we agree that this is not a typical example we would expect to see in an epidemiological setting.
+We agree that this is not a typical example we would expect to see in an epidemiological setting.
 It could happen in principle, but only in rather constructed cases.
 We still believe the example is valuable to illustrate the general point that the rankings between two forecasts can change even for a single observation.
+We added the two following sentences:
+
+\indented
+We note that the chosen example involving a geometric forecast distribution is somewhat constructed; for more similarly shaped pairs of forecast distributions this phenomenon will not typically occur.
+
+\notindented
 
@@
 
-\textit{In addition, we filtered out erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table \ref{tab:erroneous}.
+\indented
+In addition, we filtered out erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table \ref{tab:erroneous}.
 Figure \ref{fig:erroneous-forecasts.png} shows the percentage of forecasts removed for each model.
 Those few (less than 0.2\%  of forecasts for each model) erroneous outlier forecasts had excessive influence on average scores and relative skill scores in a way that was not representative of normal model behaviour.
-We removed them here in order to better illustrate the effects of the log-transformation on scores that one would expect in a 'normal' scenario.
-In a regular forecast evaluation such erroneous forecasts should usually not be removed and would count towards overall model scores.}
+We removed them here in order to better illustrate the effects of the log-transformation on scores that one would expect in a well-behaved scenario.
+In a regular forecast evaluation such erroneous forecasts should usually not be removed and would count towards overall model scores.
 
+\notindented
+
+\red
 L244: “pairwise comparisons” — The necessity of using an approach to handle missingness when combining evaluations should be noted more prominently from the beginning.
@@
 \black
-I'm not sure where we should note this. This is already more or less the beginning of the example part. 
-And for the theoretical part it's not that relevant, because missingness is somewhat independent of whether we transform forecasts.
+We added the following short paragraph to Section 2.4 to ensure these aspects are already mentioned in the conceptual part of the paper.
 
-We could add a sentence like "This transformation approach can be combined with approaches like pairwise comparisons that handle missingness"? earlier in the introduction?
+\indented
+We note that in the practical evaluation of operational forecasting systems  several additional challenges arise, which we do not study in detail.
+These concern e.g., the removal of outlying observations and forecasts and the handling of missing forecasts.
+The solutions we employed in practice are provided in Section 3.1.
 
-\red
+\notindented
+% I'm not sure where we should note this. This is already more or less the beginning of the example part. 
+% And for the theoretical part it's not that relevant, because missingness is somewhat independent of whether we transform forecasts.
+
+\blue
 Figure 4: The data (not) selected seems to obscure the key period of predicting the start of a case wave, as we expect the CRPS of log data to emphasize missed case increase predictions there more heavily than the natural scale.
@@
 \black
-The data that was removed here was largely due to data revisions, not due to forecasts removed by the filtering. 
+The data that was removed here was largely due to data revisions, not due to forecasts removed by the filtering on our part. 
 We added the percentages of observed values in the main text and added the following two Figures in the SI to make that clearer.
@@
 $w = \alpha / 2$ (which is generally standard), then the WIS converges exactly to the CRPS for an increasing number of equally spaced prediction intervals.
-\textbf{Proper scoring rule dragon, attack!} JB: No scaling. See figure 2 in \cite{bracherEvaluatingEpidemicForecasts2021}.
+For an illustration see Figure 2 in \cite{bracherEvaluatingEpidemicForecasts2021}.
 
@@
 
-\red
+\blue
 L69: This might benefit from a short justification (e.g.,
@@
 
-\textit{If case numbers are evolving based on an exponential process and the  modelling task revolves around estimating and forecasting the reproduction number or the corresponding growth rate, then evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount.
-Consider a forecast that misses the correct growth rate by either $+\epsilon$ or $-\epsilon$:}
+\indented
+If case numbers are evolving based on an exponential process and the  modelling task revolves around estimating and forecasting the reproduction number or the corresponding growth rate, then evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount.
+Consider a forecast that misses the correct growth rate by either $+\epsilon$ or $-\epsilon$:
 
@@
 
-\textit{For exponential processes errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.}
+For exponential processes errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
 
-\black
+\notindented
+\blue
 L70: Clarify: is this suggesting to make the forecasting targets themselves the result of applying some standardized growth rate estimator to current+future observations around the desired time period?
@@
 
-\red
+\blue
 L99: How are atoms at $\exp x = 0$ handled? How are observations of $y = 0$ handled?
@@
 \black
-In this theoretical discussion we defined prediction targets to have strictly positive support, so these cases are excluded.
+In the theoretical discussion in Section 2 we defined prediction targets to have strictly positive support, so these cases are excluded.
 In the later parts we follow the common strategy to add an offset before taking the logarithm.
 
-\red
+\blue
 
@@
 \begin{itemize}
-    \item If a relative error metric is desired, why not just divide CRPS/WIS by y then? (Propriety?)
-    \item How does using a (shifted) log transform differ from dividing by (a constant plus) (a mean of) the most recent observation(s)?
-    \item Is this desired?
-Though already discussed, seems like the major point of doubt here: are we at risk of overemphasizing very-low-activity periods and very-low-population states where standard-deviation-to- mean ratios are much higher?
+    \item If a relative error metric is desired, why not just divide CRPS/WIS by y then? (Propriety?)\\
+    \black
+Yes, dividing by the observed value generally disturbs propriety.
+This approach would correspond to weighting CRPS scores with weights depending on the observed value, and it has been shown that the resulting average scores are no longer proper; see Lerch et al (2017, Statistical Science, preprint: \url{https://arxiv.org/abs/1512.09244})
+    \item \blue How does using a (shifted) log transform differ from dividing by (a constant plus) (a mean of) the most recent observation(s)?\\
+    \black Like in the related arguments in the paper we address this question for the simpler case of the aboslute error rather than CRPS.
+Using a shifted log transform corresponds to (in lsight variation of equation (8))
+    \begin{equation}
+    |\varepsilon^*_t| = |\log (\hat{y_t} + a) - \log(y_t + a)| = \left|\log\left(\frac{\hat{y}_t + a}{y_t + a}\right) \right| \ \ \stackrel{\text{if } \hat{y}_t \ \approx \ y_t}{\approx} \ \ \left| \frac{\hat{y}_t + a}{y_t + a} - 1 \right| \ \ = \ \ \left| \frac{\hat{y}_t - y_t}{y_t + a} \right|.
+    \end{equation}
+    Dividing by the last observation plus a constant $a$, on the other hand, corresponds to
+    \begin{equation}
+    |\varepsilon^*_t| = \left|\frac{\hat{y}_t + a}{y_{t - 1} + a} - \frac{y_t + a}{y_{t - 1} + a} \right| = \ \ \left| \frac{\hat{y}_{t - 1} - y_t}{y_{t - 1} + a} \right|.
+    \end{equation}
+    If we interpret the denominator as an (inverse) weight for the different absolute errors, log transformation thus corresponds approximately equivalent to weighting with the inverse new observation $y_t$ (though better behaved in terms of the resulting incentives).
+Dividing by the previous value corresponds to weighting with the inverse preceding observation.
+As autocorrelation is typically strong in the time series we consider, this will yield similar, but not identical results.\\
+    In terms of the epidemiological  interpretation, dividing by the last observed value can be interpreted as a score for a multiplicative, rather than an exponential growth rate.
+% We haven't investigated in detail how the two differ in their behaviour, but would expect them to yield mostly similar results.
+    \item \blue Is this desired?
+Though already discussed, seems like the major point of doubt here: are we at risk of overemphasizing very-low-activity periods and very-low-population states where standard-deviation-to- mean ratios are much higher?\\
+    \black We agree that it may be an issue in practice that low-activity-periods could get overemphasised and therefore believe that an evaluation should show both scores on the natural scale as well as on the log scale.
 \end{itemize}
+
+\red
 
@@
 \black
-Johannes to the rescue!
+We added a reference to Gneiting (2011) for the relative error.
+SAPE is not defined in that paper, and we reference the paper by Flores (1986) instead.
+It seems that the earliest occurrence of the SAPE is due to a book by Armstrong (\textit{Long-range Forecasting: From Crystal Ball to Computer}, 1985); however, as we were unable to access the full text of this book we consider Flores (1986) the more helpful reference.
 
-\red
+% Is there a good source for the SAPE?
+It's not in the Gneiting paper.
+I found this, but not entirely convinced: https://bmcinfectdis.biomedcentral.com/articles/10.1186/s12879-017-2365-1.
+Wikipedia (\url{https://en.wikipedia.org/wiki/Symmetric_mean_absolute_percentage_error})also suggests Armstrong and Flores.
+
+\blue
 
@@
 \black
-thoughts?
 
-\red
-JB: This argument is a standard justification for the choice of variance-stabilizing transformations, see e.g., the below excerpt from \textit{Generalized Linear Models With Examples in R} by Dunn and Smyth:
+This argument is a common justification for the choice of variance-stabilizing transformations, see e.g., the below excerpt from \textit{Generalized Linear Models With Examples in R} by Dunn and Smyth:
 \begin{quote}
@@
 
-STILL NEED TO ADD THAT REMARK
 We added a reference to this source. We nonetheless agree with the referee that this approximation will work best if the distribution is tight and added a short remark to this regard.
 
-\red
+The current paragraph reads:
+
+\indented
+
+If the mean-variance relationship is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST \citep{guerreroTimeseriesAnalysisSupported1993}.
+Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method (a first-order Taylor approximation, see e.g., \citealt{Dunn2018}), to show that
+
+\begin{equation}
+\mathbb{E}_F[\text{CRPS}\{F_{\log}, \log(y)\}] \approx \frac{\sigma/\mu}{\sqrt{\pi}} 
+= \frac{\sqrt{c}}{\sqrt{\pi}}
+.
+\end{equation}
+
+The addition on ``tightly distributed'' probability mass reads
+
+\indented
+We note that while standard in the derivation of variance-stabilizing transformations, the application of the delta method in equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear} requires the probability mass of $F$ to be tightly distributed.
+If this is not the case, the approximation may be less accurate.
+
+\notindented
+
+\blue
 L159 — Need to discuss / spell out the point of the above calculation (— that it is the same for all $\mu$, when paired with the same $c$?).
@@
 \black
-Not sure I understand this.
+We added a sentence to make our point clearer:
+
+\indented
+As $\sigma$ and $\mu$ are linked through the quadratic mean-variance relationship ($\sigma = \sqrt{c} \times \mu$), the expected CRPS stays constant regardless of the expeccted value of the data-generating distribution $\mu$.
+
+\notindented
 
@@
 
-\textit{We see that when applying the CRPS on the natural scale, the expected CRPS grows monotonically as the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster) increases.}
+\indented
+We see that when applying the CRPS on the natural scale, the expected CRPS grows monotonically as the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster) increases.
 
+\notindented
 As the rest of the paragraph talks about the variance, we kept the word variance here as well.
@@
 
-\textit{We computed expected CRPS values  for three different distributions, assuming an ideal forecaster with predictive distribution equal to the true underlying (data-generating) distribution.}
+\indented
+We computed expected CRPS values  for three different distributions, assuming an ideal forecaster with predictive distribution equal to the true underlying (data-generating) distribution.
 
-\red
+\notindented
+\blue
 L186: “incentivizes overconfident predictions” — Needs justifica- tion/citation/removal. This single example disproves propriety but doesn’t establish overconfidence as being encouraged in general.
@@
 \black
-While not proved here this is a statement that holds generally. Johaaannes? Also maybe we need a reference for this... JB: I am not actually sure about this -- may need to remove.
+You are right. We removed that sentences as we do not have an analytical result on whether this holds in general.
 
-\red
+\blue
 Figure 2: Were any zeros encountered for the negative binomial or Poisson distributions? Was a shifted log transformation used rather than an unshifted log transformation?
@@
 
-\textit{Scores for two-week-ahead forecasts from the EuroCOVIDhub-ensemble (averaged across all forecast dates) for different locations, evaluated on the natural scale as well as after transforming counts by adding one and applying the natural logarithm.
+\indented
+Scores for two-week-ahead forecasts from the EuroCOVIDhub-ensemble (averaged across all forecast dates) for different locations, evaluated on the natural scale as well as after transforming counts by adding one and applying the natural logarithm.
 D: Corresponding boxplots of all individual scores of the EuroCOVIDhub-ensemble for two-week-ahead predictions.
@@
 All scores were divided by the score for forecast horizon one.
-To enhance interpretability, the range of visible relative changes in scores (relative to horizon = 1) was restricted to [0.1, 10].}
+To enhance interpretability, the range of visible relative changes in scores (relative to horizon = 1) was restricted to [0.1, 10].
 
+\notindented
 \red
@@
 \black 
-The statement that the CRPS/WIS scales with the standard deviation is true in general.
-I think they have a point though that the beta = 1 etc.
-kind of assumes a normal distribution which we say is approximated well?
+We now remind the reader of this assumption by writing
+
+\indented
+
+As (under the simplifying assumption of normality, see Section \ref{sec:methods:vst}) the WIS/CRPS of an ideal forecaster scales with the standard deviation, a value of $\beta = 1$ would imply a quadratic median-variance relationship;
+
+\notindented
 
@@
 \black
-We changed the sentence accordingly. ANY OBJECTIONS?
+We changed the sentence accordingly.
 
@@
 
-\red
+\blue
 L288–290: What would stakeholders want this difference to be, and in
@@
 
-Do we need to make changes to the text?
-
 \blue
@@
 
-\textit{Correlations of rankings on the natural and logarithmic scale.
+\indented
+Correlations of rankings on the natural and logarithmic scale.
 A: Average Spearman rank correlation of scores for individual forecasts.
@@
 The plot shows the correlation between the relative skill scores on the natural vs.
-on the log scale, representing accordance of overall model performance as judged by scores on the natural and on the log scale.}
+on the log scale, representing accordance of overall model performance as judged by scores on the natural and on the log scale.
+
+\notindented
 
@@
 The colour of the error indicates whether or not relative skill scores improved or deteriorated.
-We agree this was at all clear and thank you for pointing this out.
+We agree this was not at all clear and thank you for pointing this out.
 We updated the plot legend and added a more detailed explanation to the Figure legend:
 
-\textit{Red arrows indicate that the relative skill score deteriorated when moving from the natural to the log scale, green arrows indicate they improved.}
+\indented
+Red arrows indicate that the relative skill score deteriorated when moving from the natural to the log scale, green arrows indicate they improved.
 
-\red
+\notindented
+
+\blue
 The WIS contribution decomposition suggests that an alternative baseline should be used based on differences in transformed data rather than differences in natural scale data.
@@
 \black
-I'm not entirely sure I understand the comment. Ideas, anyone?
+We agree that it would be interesting for the Forecast Hubs to explore baseline models based on transformed differences.
+However, in this application we aimed for coherence with previous analyses of the same data and thus included the baseline models created within the Forecast Hub in real time.
+
+We also agree it is useful to flag the issues surrounding the selection of $a$ more prominently. We added the following sentence to the caption of Figure SI.5:
+
+\indented
+One potential way to do deal with this could be to use a higher $a$ value when applying a transformation $\log(x + a)$, for example $a = 10$ instead of $a = 1$.
+
+\notindented
+
+We also added a similar sentence to a paragraph in the main text, which now reads:
+
+\indented
+
+A related issue occurs when the predictive distribution has a large probability mass on zero (or on very small values), as this can translate into an excessively wide forecast in relative terms.
+This can be seen in Figure SI.5.
+Here, the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm because forecasts contained zero in its prediction intervals.
+To deal with this issue one could potentially choose to use a higher $a$ value when applying a transformation $\log(x + a)$, for example $a = 10$ instead of the $a = 1$ that we chose to use.
+
+\notindented
 
@@
 
-\red
+\blue
 L341–342: “natural logarithm as a [. . . ] (VST)” — for what type of relevant distribution(s)?
@@
 \black
-We updated the paragraph which now reads:
+We updated the relevant paragraph which now reads:
 
-\textit{We suggested using the natural logarithm as a variance-stabilising transformation (VST).
+\indented
+
+We suggested using the natural logarithm as a variance-stabilising transformation (VST).
 This is appropriate for variables that are approximately normally distributed and have a quadratic mean-variance relationship with $\sigma^2 = c \times \mu^2$ (this is e.g.
 approximately true for the negative binomoial distribution and large $\mu$).
-
-or alternatively the square-root transformation in the case of a Poisson distributed variable. Other VST like the Box-Cox (Box and Cox, 1964) are conceivable as well.
+Alternatively, the square-root transformation can be appropriate in the case of a Poisson distributed variable \citep{dunnGeneralizedLinearModels2018}.
+Other VST like the Box-Cox \citep{boxAnalysisTransformations1964} are conceivable as well.
 If one is interested in multiplicative, rather than exponential growth rates, one could convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last value that was observed at the time the forecast was made.
@@
 Alternatively, one could also take first differences of values on the log scale.
-This approach would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory for a different approach to evaluating the shape of a forecast, see Srivastava et al., 2022).
+This approach would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluating the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}.
 Dividing values by the previous value, unfortunately, is not feasible under the current quantile-based format of the Forecast Hubs, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate.
 However, it may be an interesting approach if predictive samples are available or if quantiles for weekwise growth rates have been collected.
-It is possible to go beyond choosing a single transformation by constructing composite scores as a weighted sum of scores based on different transformations.
-This would make it possible to create custom scores and allow forecast consumers to assign explicit weights to different qualities of the forecasts they might care about.
-}
+
+\notindented
 
@@
 \black
-We slightly adapted a sentence in the previous paragraph to make this point clearer:
+We slightly adapted the text to make this point clearer:
 
-\textit{It is possible to go beyond choosing a single transformation by constructing composite scores as a weighted sum of scores based on different transformations (potentially after standardising different scores to have the same mean and standard deviation).
-This would make it possible to create custom scores and allow forecast consumers to choose and assign explicit weights to different qualities of the forecasts they might care about.}
+\indented
+It is possible to go beyond choosing a single transformation by constructing composite scores as a weighted sum of scores based on different transformations.
+This would make it possible to create custom scores and allow forecast consumers to choose and assign explicit weights to different qualities of the forecasts they might care about.
 
-\textbf{Johannes, the standardising different scores thing should be kosher, right?} JB: not sure
+\notindented
 
@@
 
-\red
+\blue
 Figure SI.7:
 \begin{itemize}
-    \blue
     \item Also specify sorting variable\&direction on x axis. (Would a point plot vs. the sorting variable be more informative?)
-    \red
     \item Very informative figure, promote to main text?
@@
 \black
-We added a statement to the legend "Locations are sorted according to the mean observed value in that location."
+We added a statement to the legend "Locations are sorted according to the mean observed value in that location." We also followed your suggestion to make this a main figure.
 
-Thoughts on making this a main figure?
-
-I would expect that one could find a value for a (e.g. 10?) that would eventually yield a more or less flat line.
+We would expect that one could find a value for a (e.g. 10?) that would eventually yield a more or less flat line, but we haven't tested this in practice.
 
@@
 
-\red
+\blue
 In several places, e.g., L201, there are links to SI material, when clicked, that instead redirect to main text figures.
@@
 \black
-Not sure why this happens - some Latex error? But also not sure worth fixing as they will do that anyway.
+We're not sure why this happens and apologise for the confusion. We're also admittedly not sure how to fix this, but expect this will be dealt with during final typesetting.
 
-\red
+\blue
 \section{Reviewer \#3}
 
-\blue
 In this manuscript the authors investigate how common scoring rules for probabilistic forecasts can be adjusted by transforming observed data and forecast predictions prior to scoring.
@@
 \black
-Thank you very much
+Thank you for your kind words.
 

### 2023-06-20 main.tex
More comments from Johannes

@@
 For example, the log score assigns particularly severe penalties to occasional misguided forecasts.
-This lack of robustness may limit its practical applicability \cite{bracherEvaluatingEpidemicForecasts2021}.
+This lack of robustness may limit its practical applicability \citep{bracherEvaluatingEpidemicForecasts2021}.
 Moreover, it is not easily applied to forecasts reported as samples or quantiles, as used in many recent disease forecasting efforts.
@@
 The absolute error after log transforming is thus an approximation of the \textit{absolute percentage error} \citep[APE,][]{gneitingMakingEvaluatingPoint2011a} as long as forecast and observation are close.
-As we assumed that $\hat{y} \approx y$, we can also interpret it as an approximation of the \textit{relative error} (RE)
+As we assumed that $\hat{y} \approx y$, we can also interpret it as an approximation of the \textit{relative error} (RE, \citealt{gneitingMakingEvaluatingPoint2011a})
 \begin{linenomath*}
@@
 \end{linenomath*}
-and the \textit{symmetric absolute percentage error} (SAPE)
+and the \textit{symmetric absolute percentage error} (SAPE; see e.g., \citealt{Flores1986})
 \begin{linenomath*}
@@
 If the mean-variance relationship is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST \citep{guerreroTimeseriesAnalysisSupported1993}.
-Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method (a first-order Taylor approximation, see e.g., \cite{Dunn2018}), to show that
+Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method (a first-order Taylor approximation, see e.g., \citealt{Dunn2018}), to show that
 \begin{linenomath*}
@@
 = \frac{\sqrt{c}}{\sqrt{\pi}}
-.
+.\label{eq:taylor_quadratic}
 \end{equation}
@@
 
-If the variance is linear with $\sigma^2 = c \times \mu$, as with a Poisson-distributed variable, the square root is known to be a VST \citep{dunnGeneralizedLinearModels2018}. 
+If the variance is linear with $\sigma^2 = c \times \mu$, as with a Poisson-distributed variable, the square root is known to be a VST \citep{Dunn2018}. 
 Denoting by $F_{\sqrt{\ }}$ the predictive distribution for $\sqrt{Y}$, the delta method can again be used to show that
@@
 \mathbb{E}_F[\text{CRPS}\{F_{\sqrt{\ }}, \sqrt{y}\}] \approx \frac{\sigma/\sqrt{\mu}}{2\sqrt{\pi}} = \frac{\sqrt{c}}{2\sqrt{\pi}}
-.
+.\label{eq:taylor_linear}
 \end{equation}
 \end{linenomath*}
+We note that while standard in the derivation of variance-stabilizing transformations, the application of the delta method in equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear} requires the probability mass of $F$ to be tightly distributed.
+If this is not the case, the approximation may be less accurate.
 
@@
 
+We note that in the practical evaluation of operational forecasting systems  several additional challenges arise, which we do not study in detail.
+These concern e.g., the removal of outlying observations and forecasts and the handling of missing forecasts.
+The solutions we employed in practice are provided in Section \ref{sec:HUB-setting}.
+
 \subsection{Effects on model rankings}
@@
 On the log scale, however, forecaster A receives a lower score for large observed values, being more heavily penalised for assigning large probability to small values (which, in relative terms, are far away from the actual observation).
+We note that the chosen example involving a geometric forecast distribution is somewhat constructed; for more similarly shaped pairs of forecast distributions this phenomenon will not typically occur.
 
@@
 Note that we are using predictive medians rather than means as only the former are available in the European COVID-19 Forecast Hub.
-As the WIS/CRPS of an ideal forecaster scales with the standard deviation (see Section \ref{sec:methods:vst}), a value of $\beta = 1$ would imply a quadratic median-variance relationship; the natural logarithm could then serve as a VST.
+As (under the simplifying assumption of normality; see Section \ref{sec:methods:vst}) the WIS/CRPS of an ideal forecaster scales with the standard deviation, a value of $\beta = 1$ would imply a quadratic median-variance relationship; the natural logarithm could then serve as a VST.
 A value of $\beta = 0.5$ would imply a linear median-variance relationship, suggesting the square root as a VST.

### 2023-06-21 main.tex
fix merge conflicts plus more changes in response to reviewers

@@
 Probabilistic forecasts are usually evaluated using so-called proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data.
-Proper scoring rules are constructed such that they cannot be gamed, meaning that forecasters (anyone or anything that issues a forecast) are incentivised to report their true belief about the future.
-In expectation, a forecaster receives the best score if their forecast $F$ coincides with the data-generating distribution $G$, that is if $F = G$.
+Proper scoring rules are constructed such that they encourage honest forecasting and cannot be `gamed' or `cheated'. 
+Assuming that the forecaster's actual best judgement corresponds to a predictive distribution $F$, a proper score is constructed such that under $F$, no other distribution $G$ yields a better expected score.
+Forecasters (anyone or anything that issues a forecast) are thus incentivised to report their true belief $F$ about the future.
 Common proper scoring rules are the logarithmic or log score \citep{goodRationalDecisions1952} or the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}.
@@
 For example, the log score assigns particularly severe penalties to occasional misguided forecasts.
-This lack of robustness may limit its practical applicability \citep{bracherEvaluatingEpidemicForecasts2021}.
+This lack of robustness may limit its practical applicability \cite{bracherEvaluatingEpidemicForecasts2021}.
 Moreover, it is not easily applied to forecasts reported as samples or quantiles, as used in many recent disease forecasting efforts.
@@
 For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation in the context of epidemic phenomena.
-We refer to this transformation as 'log-transformation' and to scores that have been log-transformed as scores 'on the log scale' (as opposed to scores 'on the natural scale', which have not been transformed).
+We refer to this transformation as 'log-transformation' and to scores that have been computed from log-transformed forecasts and observations as scores 'on the log scale' (as opposed to scores 'on the natural scale', which involve no transformation).
 In the theoretical discussion in Section \ref{sec:methods}, 'log-transformation' and 'log scale' generally refer to a transformation of $\log_{e}(x)$.
-For practical applications (Section \ref{sec:HUB}) we also use these terms to describe a transformation of $\log_{e}(x + \varepsilon)$ in order to keep the notation simple.
+For practical applications (Section \ref{sec:HUB}) we also use these terms to describe a transformation of $\log_{e}(x + \varepsilon)$ in order to keep the terminology and notation simple.
 For a prediction target with strictly positive support, the CRPS after applying a log-transformation is given by
@@
 The absolute error after log transforming is thus an approximation of the \textit{absolute percentage error} \citep[APE,][]{gneitingMakingEvaluatingPoint2011a} as long as forecast and observation are close.
-As we assumed that $\hat{y} \approx y$, we can also interpret it as an approximation of the \textit{relative error} (RE, \citealt{gneitingMakingEvaluatingPoint2011a})
+As we assumed that $\hat{y} \approx y$, we can also interpret it as an approximation of the \textit{relative error} \citep[RE,][]{gneitingMakingEvaluatingPoint2011a}
 \begin{linenomath*}
@@
 When evaluating models across sets of forecasting tasks, it may be desirable for each target to have a similar impact on the overall results.
-This is true if one assumes that forecasts for different targets provide similar amounts of information about how well a forecaster performs.
-Ideally, one would like the magnitude of the score to scale with the difficulty of the prediction tasks and the information that can be obtained about overall performance of the forecaster.
-In disease incidence forecasting, however, CRPS values on the natural scale typically scale with the order of magnitude of the quantity to be predicted.
+This could be motivated by the assumption that forecasts from different geographical units and time periods provide similar amounts of information about how well a forecaster performs.
+One would then like the resulting scores to be independent of the order of magnitude of the target to predict.
+CRPS values on the natural scale, however, typically scale with the order of magnitude of the quantity to be predicted.
 Average scores are then dominated by the results achieved for targets with high expected outcomes in a way that does not necessarily reflect the underlying predictive ability well.
@@
 
-If the mean-variance relationship is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST \citep{guerreroTimeseriesAnalysisSupported1993}.
+If the mean-variance relationship of the data-generating distribution is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST \citep{guerreroTimeseriesAnalysisSupported1993}.
 Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method (a first-order Taylor approximation, see e.g., \citealt{Dunn2018}), to show that
+
 \begin{linenomath*}
@@
 \end{linenomath*}
+As $\sigma$ and $\mu$ are linked through the quadratic mean-variance relationship ($\sigma = \sqrt{c} \times \mu$), the expected CRPS stays constant regardless of the expeccted value of the data-generating distribution $\mu$.
 The assumption of a quadratic mean-variance relationship is closely linked to the aspects discussed in Sections \ref{sec:methods:relative} and \ref{sec:methods:growthrate}.
 It implies that relative errors have constant variance and can thus be meaningfully compared across different targets.
-Also, it arises naturally if we assume that our capacity to predict the epidemic growth rate does not depend on the expected outcome.
+Also, it arises naturally if we assume that our capacity to predict the epidemic growth rate does not depend on the expected outcome, i.e.
+does not depend on the current phase of the epidemic or the order of magnitude of current observations.
 
@@
 In the case of log transforms, taking the logarithm of the scores, rather than scoring the log-transformed forecasts and data, results in an improper score.
-This is because taking the logarithm of the CRPS (or WIS) results in a score that does not penalise outliers enough and therefore incentivises overconfident predictions.
-We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can easily be seen that overconfident models perform best in terms of the log WIS.
+We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can be seen that in the example overconfident models perform best in terms of the log WIS.
 
@@
 On the other hand, choosing a larger $a$ can help stabilise scores for forecasts and observations close to zero, avoiding giving excessive weight to forecasts for small quantities (see Figure \ref{fig:HUB-log-different-offsets}).
+For large values of $a$, $\log(x + a)$ grows linearly in $x$, meaning that we expect to observe the same patterns as in the case with no transformation.
+For decreasing values of $a$, we give more relative weight to scores in small locations.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/HUB-scores-locations-log-variants.png}
+    \caption{Mean WIS in different locations for different transformations applied before scoring.
+Locations are sorted according to the mean observed value in that location.
+Shown are scores for two-week-ahead forecasts of the EuroCOVIDhub-ensemble.
+On the natural scale (with no transformation prior to applying the WIS), scores correlate strongly with the average number of observed values in a given location.
+The same is true for scores obtained after applying a square-root transformation, or after applying a log-transformation with a large offset $a$.
+For illustrative purposes, $a$ was chosen to be 101630 for cases and 530 for deaths, 10 times the respective median observed value.
+For large values of $a$, $log(x + a)$ grows linearly in $x$, meaning that we expect to observe the same patterns as in the case with no transformation.
+For decreasing values of $a$, we give more relative weight to scores in small locations.}
+    \label{fig:HUB-log-different-offsets}
+\end{figure}
 
@@
 Here, the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm because forecasts contained zero in its prediction intervals.
+To deal with this issue one could potentially choose to use a higher $a$ value when applying a transformation $\log(x + a)$, for example $a = 10$ instead of the $a = 1$ that we chose to use.
 
@@
 On the log scale, however, forecaster A receives a lower score for large observed values, being more heavily penalised for assigning large probability to small values (which, in relative terms, are far away from the actual observation).
-We note that the chosen example involving a geometric forecast distribution is somewhat constructed; for more similarly shaped pairs of forecast distributions this phenomenon will not typically occur.
+We note that the chosen example involving a geometric forecast distribution is somewhat constructed; as shown in Section \ref{sec:Hub:cor} and Figure \ref{fig:HUB-cors}A, rankings between models usually stay quite stable for a single forecast.
+
+.
 
@@
 Figure \ref{fig:number-anomalies} displays the number of anomalies removed for each location.
-In addition, we filtered out erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table \ref{tab:erroneous}.
+In addition, we filtered out a small number of erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table \ref{tab:erroneous}.
 Figure \ref{fig:erroneous-forecasts.png} shows the percentage of forecasts removed for each model.
 Those few (less than 0.2\%  of forecasts for each model) erroneous outlier forecasts had excessive influence on average scores and relative skill scores in a way that was not representative of normal model behaviour.
-We removed them here in order to better illustrate the effects of the log-transformation on scores that one would expect in a 'normal' scenario.
+We removed them here in order to better illustrate the effects of the log-transformation on scores that one would expect in a well-behaved scenario.
 In a regular forecast evaluation such erroneous forecasts should usually not be removed and would count towards overall model scores.
@@
 \subsection{Impact of logarithmic transformation on model rankings}
+\label{sec:Hub:cor}
 
@@
 
-It is possible to go beyond choosing a single transformation by constructing composite scores as a weighted sum of scores based on different transformations (potentially after standardising different scores to have the same mean and standard deviation).
+It is possible to go beyond choosing a single transformation by constructing composite scores as a weighted sum of scores based on different transformations.
 This would make it possible to create custom scores and allow forecast consumers to choose and assign explicit weights to different qualities of the forecasts they might care about.
@@
 One could argue that including zero in the prediction intervals constituted an unreasonable forecast that was rightly penalised, but in general care has to be taken with small numbers.
+One potential way to do deal with this could be to use a higher $a$ value when applying a transformation $\log(x + a)$, for example $a = 10$ instead of $a = 1$.
 A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale.
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-scores-locations-log-variants.png}
-    \caption{Mean WIS in different locations for different transformations applied before scoring.
-Locations are sorted according to the mean observed value in that location.
-Shown are scores for two-week-ahead forecasts of the EuroCOVIDhub-ensemble.
-On the natural scale (with no transformation prior to applying the WIS), scores correlate strongly with the average number of observed values in a given location.
-The same is true for scores obtained after applying a square-root transformation, or after applying a log-transformation with a large offset $a$.
-For illustrative purposes, $a$ was chosen to be 101630 for cases and 530 for deaths, 10 times the respective median observed value.
-For large values of $a$, $log(x + a)$ grows linearly in $x$, meaning that we expect to observe the same patterns as in the case with no transformation.
-For decreasing values of $a$, we give more relative weight to scores in small locations.}
-    \label{fig:HUB-log-different-offsets}
-\end{figure}
-
 \clearpage

### 2023-06-21 reply_to_reviews.tex
Updates from Overleaf

@@
 Proper scoring rules are constructed such that they encourage honest forecasting and cannot be `gamed' or `cheated'. 
-% In expectation, a forecaster receives the best score if their forecast $F$ coincides with the data-generating distribution $G$, that is if $F = G$.
 Assuming that the forecaster's actual best judgement corresponds to a predictive distribution $F$, a proper score is constructed such that under $F$, no other distribution $G$ yields a better expected score.
@@
 In any case, developing these ideas would require more in-depth investigation and we feel this is beyond the scope of the current paper.
-% and therefore did not include it in this work.
 
@@
 As stated explicitly in the manuscript, we moreover recommend to use this in complement to rather than as a replacement of the common evaluation on the ``natural scale''.
-Evaluators will then becoma aware of whether their results depend strongly on this decision and the implied weighting of different forecasting tasks.
+Evaluators will then become aware of whether their results depend strongly on this decision and the implied weighting of different forecasting tasks.
 
@@
 
-\red not sure what to make of Box-Cox.
-
-Hm potentially an interesting point.
-At the moment the paper kind of assumes you know what you're doing and suggests to do it differently.
-One could I guess add a paragraph that walks through all the steps?
-Add it where?
-Maybe this should be done in another piece that talks about best practices?
-
 \blue
@@
 Not introducing another term for transformations of $\log_{e}(x + \varepsilon)$ makes the manuscript more readable and accessible to readers.
-In addition, our analyses suggest that adding a small context doesn't make much of a difference in most practical applications.
+In addition, our analyses suggest that adding a small offset doesn't make much of a difference in most practical applications.
 We nevertheless completely agree that this should be made clear to the reader and have added the following text:
@@
 \black
-We think that while it is certainly not the only relevant approach, there is a case to be made for an evaluation which gives each foreasting target a similar weight.
+We think that while it is certainly not the only relevant approach, there is a case to be made for an evaluation which gives each forecasting target a similar weight.
 This is attractive if the goal of the evaluation is to find the best forecasting model for future use and if one assumes that the difficulty of the forecasting task is similar in different locations.
@@
 We still believe the example is valuable to illustrate the general point that the rankings between two forecasts can change even for a single observation.
-We added the two following sentences:
+We added the following sentence:
 
 \indented
-We note that the chosen example involving a geometric forecast distribution is somewhat constructed; for more similarly shaped pairs of forecast distributions this phenomenon will not typically occur.
+We note that the chosen example involving a geometric forecast distribution is somewhat constructed; as shown in Section \ref{sec:Hub:cor} and Figure \ref{fig:HUB-cors}A, rankings between models usually stay quite stable for a single forecast.
+.
 
@@
 \notindented
-% I'm not sure where we should note this. This is already more or less the beginning of the example part. 
-% And for the theoretical part it's not that relevant, because missingness is somewhat independent of whether we transform forecasts.
 
@@
     \item \blue How does using a (shifted) log transform differ from dividing by (a constant plus) (a mean of) the most recent observation(s)?\\
-    \black Like in the related arguments in the paper we address this question for the simpler case of the aboslute error rather than CRPS.
-Using a shifted log transform corresponds to (in lsight variation of equation (8))
+    \black Like in the related arguments in the paper we address this question for the simpler case of the absolute error rather than CRPS.
+Using a shifted log transform corresponds to (in slight variation of equation (8))
     \begin{equation}
@@
 It seems that the earliest occurrence of the SAPE is due to a book by Armstrong (\textit{Long-range Forecasting: From Crystal Ball to Computer}, 1985); however, as we were unable to access the full text of this book we consider Flores (1986) the more helpful reference.
-
-% Is there a good source for the SAPE?
-It's not in the Gneiting paper.
-I found this, but not entirely convinced: https://bmcinfectdis.biomedcentral.com/articles/10.1186/s12879-017-2365-1.
-Wikipedia (\url{https://en.wikipedia.org/wiki/Symmetric_mean_absolute_percentage_error})also suggests Armstrong and Flores.
 
@@
 
-If the mean-variance relationship is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST \citep{guerreroTimeseriesAnalysisSupported1993}.
+If the mean-variance relationship of the data-generating distribution is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST \citep{guerreroTimeseriesAnalysisSupported1993}.
 Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method (a first-order Taylor approximation, see e.g., \citealt{Dunn2018}), to show that
@@
 
-We would expect that one could find a value for a (e.g. 10?) that would eventually yield a more or less flat line, but we haven't tested this in practice.
+We would expect that one could find a value for a that would eventually yield a more or less flat line, but we haven't investigated this in practice.
 

### 2023-06-21 reply_to_reviews.tex
address more reviewer comments

@@
 
-\black \textbf{Black: our comments}
-
-\red \textbf{Reviewer, not addressed}
-
-\blue \textbf{Reviewer, addressed}
+%\black \textbf{Black: our comments}
+%\red \textbf{Reviewer, not addressed}
+%\blue \textbf{Reviewer, addressed}
 
@@
 Users who are explicitly concerned about the relative weights of different times and locations can e.g.
-find information in Figure SI.7 where we show how different forecast targets are emphasized by evaluation on the natural or log-transformed scale.
+find information in Figure 3 where we show how different forecast targets are emphasized by evaluation on the natural or log-transformed scale.
 A user could e.g.
@@
 
-\red
+\blue
 Additionally, any additional necessary/recommended elements of a scoring procedure should be noted more prominently (e.g., dealing with forecast missingness, outlying forecasts, outlying data, selecting a Box-Cox transformation, selecting a shift ($a$ value), etc.).
@@
 The solutions we employed in practice are provided in Section 3.1.
+\notindented
 
-\notindented
+We agree that it would be useful to have a more thorough discussion of all the considerations involved in evaluating forecasts, but feel this is slightly beyond the scope of this paper.
 
@@
 
-% to model epidemiological processes. We agree with the reviewer that a size parameter of 0.1 is not realistic, but still think that the example is valuable as an illustration.
-
 \blue
@@
 \indented
-We note that the chosen example involving a geometric forecast distribution is somewhat constructed; as shown in Section \ref{sec:Hub:cor} and Figure \ref{fig:HUB-cors}A, rankings between models usually stay quite stable for a single forecast.
+We note that the chosen example involving a geometric forecast distribution is somewhat constructed; as shown in Section 3.4 and Figure 8A, rankings between models usually stay quite stable for a single forecast.
 .
@@
 \indented
-In addition, we filtered out erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table \ref{tab:erroneous}.
-Figure \ref{fig:erroneous-forecasts.png} shows the percentage of forecasts removed for each model.
+In addition, we filtered out erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table SI.2.
+Figure SI.6 shows the percentage of forecasts removed for each model.
 Those few (less than 0.2\%  of forecasts for each model) erroneous outlier forecasts had excessive influence on average scores and relative skill scores in a way that was not representative of normal model behaviour.
@@
 
-\red
+\blue
 L244: “pairwise comparisons” — The necessity of using an approach to handle missingness when combining evaluations should be noted more prominently from the beginning.
@@
 
-\textit{Applying a transformation of log(x + 1) to data and forecasts from the European COVID-19 Forecast Hub}
+\indented
+Applying a transformation of log(x + 1) to data and forecasts from the European COVID-19 Forecast Hub
+
+\notindented
 
@@
 
-\textit{Situations in which models missed the beginning of upward swings are more strongly emphasised while failing to predict a downturn following a peak is less severely penalised when scoring transformed forecasts as opposed to untransformed ones.}
+\indented
+Situations in which models missed the beginning of upward swings are more strongly emphasised while failing to predict a downturn following a peak is less severely penalised when scoring transformed forecasts as opposed to untransformed ones.
+
+\notindented
 
@@
 
-\textit{The reproduction number $R$ describes the number of people each infected person is expected to infect in turn (irrespective of the time between infections) and therefore is a measure of the strength of epidemic growth.
-The growth rate $r$ describes the speed of epidemic growth.}
+\indented
+The reproduction number $R$ describes the number of people each infected person is expected to infect in turn (irrespective of the time between infections) and therefore is a measure of the strength of epidemic growth.
+The growth rate $r$ describes the speed of epidemic growth.
+
+\notindented
 
@@
 
-\red
+\blue
 
@@
 
-\red
+\blue
 L158 “delta method” — Delta method would seem to also require the distribution to be somewhat tightly distributed, concentrated within a nearly-linear region of the transformation function.
@@
 
-We added a reference to this source. We nonetheless agree with the referee that this approximation will work best if the distribution is tight and added a short remark to this regard.
+We added a reference to this source. We nonetheless agree with you that this approximation will work best if the distribution is tight and added a short remark to this regard.
 
@@
 \indented
-We note that while standard in the derivation of variance-stabilizing transformations, the application of the delta method in equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear} requires the probability mass of $F$ to be tightly distributed.
+We note that while standard in the derivation of variance-stabilizing transformations, the application of the delta method in equations (16) and (17) requires the probability mass of $F$ to be tightly distributed.
 If this is not the case, the approximation may be less accurate.
@@
 \notindented
-\red
 
+\blue
 L272: “with the standard deviation” — Also note here that this is for approximately normal data distributions?
@@
 
-\red
+\blue
 Figure SI.2: Would benefit from some more discussion, and perhaps a plot of slightly different quantities.
@@
 \black
-I'm not sure I can follow.
-In my mind the plot is there to show you that once your values are like 5 times larger than the offset you choose than the offset doesn't have much of an effect.
-So it's designed to show that for most practical purposes, doing log(x + a) should be mostly equivalent to log(x).
-So to me it's less about choosing an appropriate a and more about showing that probably it's just fine to go with some random small a.
-I think Figure SI.7 is more about "choosing an a".
+
+We added some more explanation to the caption of Figure SI.2. This makes the trade-off involved in choosing a suitable $a$ value clearer, without prescribing a 'correct' solution.
+
+\indented
+For a given solid line, the dashed line of the same colour marks the x-value that  is equal to 5 times the corresponding offset.
+It can be seen that for $a$ values smaller than one fifth of the transformed quantity, the effect of adding an offset is generally small.
+When choosing a suitable $a$, the trade-off is between staying close to the interpretation of a pure log-transformation (choosing a small $a$) and not giving excessive weights to small observations (by choosing a larger $a$, see Figure 3).
+
+\notindented
 

### 2023-06-21 main.tex
address more reviewer comments

@@
 \end{linenomath*}
-As $\sigma$ and $\mu$ are linked through the quadratic mean-variance relationship ($\sigma = \sqrt{c} \times \mu$), the expected CRPS stays constant regardless of the expeccted value of the data-generating distribution $\mu$.
+As $\sigma$ and $\mu$ are linked through the quadratic mean-variance relationship ($\sigma = \sqrt{c} \times \mu$), the expected CRPS stays constant regardless of the expected value of the data-generating distribution $\mu$.
 The assumption of a quadratic mean-variance relationship is closely linked to the aspects discussed in Sections \ref{sec:methods:relative} and \ref{sec:methods:growthrate}.
@@
 This condition holds for both the log and square root transformations, as well as many others.
-However, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee that the score remains proper.
+Some strictly monotonic transformations that can generally be applied are scaling by the population size or scaling by past observations.
+This, as discussed in Section \ref{sec:discussion}, is similar to applying a log-transformation, but corresponds to evaluating a forecast of multiplicative, rather than exponential growth rates.
+There may of course be exceptions such as dividing by zero, but otherwise this just represents a monotonic transformation and the remaining score is still proper.
+Scaling a forecast by the later observed value (as opposed to scaling by past observations) is generally not permissible as it disturbs propriety (see \cite{lerchForecasterDilemmaExtreme2015} on the closely related topic of weighting scores with a function of the observed value).
+Similarly, scaling a forecast by predictions made by other forecasters may result in improper scores, but we are unaware of existing theoretical arguments.
+Scaling scores, e.g.
+dividing a score by the mean score for other forecasters may be acceptable in most applied settings, but we are unaware of existing theoretical arguments that could define exactly when and how this could result in improper scores (as is the case for example when dividing by zero).
+
+When applying a transformation, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee that the score remains proper.
 In the case of log transforms, taking the logarithm of the scores, rather than scoring the log-transformed forecasts and data, results in an improper score.
@@
 For decreasing values of $a$, we give more relative weight to scores in small locations.
+In principle, a user could produce a plot similar to Figure \ref{fig:HUB-log-different-offsets} with different values of the offset $a$ and choose $a$ such that the relative weightings of times and regions with high and low incidence correspond to their preferences.
 
@@
 We note that the chosen example involving a geometric forecast distribution is somewhat constructed; as shown in Section \ref{sec:Hub:cor} and Figure \ref{fig:HUB-cors}A, rankings between models usually stay quite stable for a single forecast.
-
-.
 
@@
 In addition, we filtered out a small number of erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table \ref{tab:erroneous}.
-Figure \ref{fig:erroneous-forecasts.png} shows the percentage of forecasts removed for each model.
+Figure \ref{fig:erroneous-forecasts} shows the percentage of forecasts removed for each model.
 Those few (less than 0.2\%  of forecasts for each model) erroneous outlier forecasts had excessive influence on average scores and relative skill scores in a way that was not representative of normal model behaviour.
@@
 Other VST like the Box-Cox \citep{boxAnalysisTransformations1964} are conceivable as well.
-If one is interested in multiplicative, rather than exponential growth rates, one could convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last value that was observed at the time the forecast was made.
+If one is interested in multiplicative, rather than exponential growth rates, one could, instead of applying a log transformation, convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last value that was observed at the time the forecast was made.
 Forecasters would then implicitly predict a separate multiplicative growth rate from today to horizon 1, 2, etc.
@@
 For increasing x, all lines eventually approach the black line (representing a transformation with no offset applied).
-    For a given solid line, the dashed line of the same colour marks the x-value that  is equal to 5 times the corresponding offset.}
+    For a given solid line, the dashed line of the same colour marks the x-value that  is equal to 5 times the corresponding offset.
+It can be seen that for $a$ values smaller than one fifth of the transformed quantity, the effect of adding an offset is generally small.
+When choosing a suitable $a$, the trade-off is between staying close to the interpretation of a pure log-transformation (choosing a small $a$) and not giving excessive weights to small observations (by choosing a larger $a$, see Figure \ref{fig:HUB-log-different-offsets}).}
     \label{fig:illustration-effect-log-offset}

### 2023-06-22 R/00 full paper code.R
Update Figure to show log(x + 100) instead of log(x + 0.1)

@@ -482,8 +482,8 @@ scores_log_alts <- scores_log_alts |>
                                  levels = c("Cases - natural", "Deaths - natural", 
                                             "Cases - sqrt", "Deaths - sqrt", 
                                             "Cases - log + 10x median", "Deaths - log + 10x median", 
+                                            "Cases - log + 100", "Deaths - log + 100", 
                                             "Cases - log + 1", "Deaths - log + 1", 
-                                            "Cases - log + 0.1", "Deaths - log + 0.1", 
                                             "Cases - log + 0.001", "Deaths - log + 0.001" 
                                  )))
 

### 2023-06-22 R/Refetch data from Forecast Hub.R

@@ -280,9 +280,9 @@ fwrite(scores, here("output", "data", "all-scores-european-hub.csv"))
 scores_log_alts <- 
   rbind(hub_data |>
           mutate(
-            scale = "log + 0.1", 
-            true_value = log(true_value + 0.1), 
-            prediction = log(pmax(prediction, 0) + 0.1)
+            scale = "log + 100", 
+            true_value = log(true_value + 100), 
+            prediction = log(pmax(prediction, 0) + 100)
           ), 
         hub_data |> mutate(
           scale = "log + 0.001", 

### 2023-06-22 reply_to_reviews.tex
Updates from Overleaf

@@
 
+We added a paragraph on which kinds of transformations are permissible in Section 4 (see comments below).
 We agree that it would be useful to have a more thorough discussion of all the considerations involved in evaluating forecasts, but feel this is slightly beyond the scope of this paper.
@@
 dividing a score by the mean score for other forecasters may be acceptable in most applied settings, but we are unsure about exactly when and how this could break propriety (as is the case for example when dividing by zero).
+
+To make this clearer in the text, we added the following paragraphs:
+
+\indented
+A natural question is which other transformations could be applied and whether resulting scores remain (strictly) proper.
+In principle, any transformation function can be applied simultaneously to forecasts and observations as long as the definition of the transformation is independent of the forecasts and any quantities unknown at the time of forecasting, including the observed value.
+This simply corresponds to a re-definition of the forecasting target.
+However, applying non-invertible transformations leads to a loss in information conveyed by forecasts, which we consider undesirable.
+The resulting score will be proper, but it may not be strictly proper anymore (as forecasts differing from the forecaster's true belief on the original scale may be identical on the transformed scale).
+When using the CRPS or the WIS, it seems most appropriate to use only strictly monotonic transformations as otherwise the encoded notion of distance may become meaningless.
+If only strictly monotonic transformations are used, then transforming a forecaster's best belief on the original scale will correspond to their best belief on the transformed scale.
+This holds for both the log and square root transformations, as well as many others.
+
+Some other strictly monotonic transformations that can generally be applied prior to computing  the CRPS or WIS are scaling by the population size or scaling by past observations.
+The latter, as discussed in Section 4, is similar to applying a log-transformation, but corresponds to evaluating a forecast of multiplicative, rather than exponential growth rates.
+The arising issue of dividing by zero can again be solved by adding a small offset.
+This just represents a monotonic transformation and the remaining score is still strictly proper.
+Scaling a forecast by the later observed value (as opposed to scaling by past observations) is generally not permissible as it disturbs propriety (see \cite{lerchForecasterDilemmaExtreme2015} on the closely related topic of weighting scores with a function of the observed value).
+Similarly, scaling a forecast by predictions made by other forecasters may result in improper scores, but we are unaware of existing theoretical arguments.
+
+When applying a transformation, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee that the score remains proper.
+In the case of log transforms, taking the logarithm of the scores, rather than scoring the log-transformed forecasts and data, results in an improper score.
+We illustrate this point using simulated data in Figure SI.1, where it can be seen that in the example overconfident models perform best in terms of the log WIS.
+Scaling scores (rather than forecasts and observations) by the mean score for other forecasters is closely related to the common practice of computing skill scores.
+While these are typically not technically proper \citep{gneitingStrictlyProperScoring2007}, their application is established practice and considered largely unproblematic.
+\notindented
 
@@
 \clearpage
-\bibliography{log-or-not-v2.bib, software.bib}
+\bibliography{log-or-not-v2.bib, log-or-not.bib, software.bib}
 

### 2023-06-22 main.tex

@@
 
-\title{Transformation of forecasts for evaluating predictive performance in an epidemiological context}
+\title{Scoring epidemiological forecasts on transformed scales}
   \author{Nikos I.
@@
 Forecast evaluation is essential for the development of predictive epidemic models and can inform their use for public health decision-making.
-Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which are both measures of the absolute distance between the forecast distribution and the observation.
-However, applying these scores directly to predicted and observed incidence counts may not be optimal due to the exponential nature of epidemic processes and the varying magnitudes of observed values across space and time.
+Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which can be seen as measures of the absolute distance between the forecast distribution and the observation.
+However, applying these scores directly to predicted and observed incidence counts may not be the most appropriate due to the exponential nature of epidemic processes and the varying magnitudes of observed values across space and time.
 In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results.
@@
 
-Probabilistic forecasts are usually evaluated using so-called proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data.
+Probabilistic forecasts are usually evaluated using so-called (strictly) proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data.
 Proper scoring rules are constructed such that they encourage honest forecasting and cannot be `gamed' or `cheated'. 
-Assuming that the forecaster's actual best judgement corresponds to a predictive distribution $F$, a proper score is constructed such that under $F$, no other distribution $G$ yields a better expected score.
+Assuming that the forecaster's actual best judgement corresponds to a predictive distribution $F$, a proper score is constructed such that if $F$ was the data-generating process, no other distribution $G$ would yield a better expected score.
+A scoring rule is called \textit{strictly} proper if there is no other distribution that under $F$ achieves the \textit{same} expected score as $F$, meaning that any deviation from $F$ leads to a worsening of expected scores.
 Forecasters (anyone or anything that issues a forecast) are thus incentivised to report their true belief $F$ about the future.
-Common proper scoring rules are the logarithmic or log score \citep{goodRationalDecisions1952} or the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}.
-The log score is the log-likelihood of the predictive density at the observed value and is scale-invariant under monotonous transformations.
-The log score does not suffer from the issues we address in this paper, but due to a number of reasons is not as widely used in fields such as weather forecasting or epidemiology (see e.g., \citealt{Johansson2019} for an exception).
-For example, the log score assigns particularly severe penalties to occasional misguided forecasts.
-This lack of robustness may limit its practical applicability \cite{bracherEvaluatingEpidemicForecasts2021}.
+Common proper scoring rules are the logarithmic or log score \citep{goodRationalDecisions1952} and the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}.
+The log score is the predictive log density evaluated at the observed value and has many desirable theoretical properties resulting from the likelihood principle; however, the particularly severe penalties it assigns to occasional misguided forecasts make it little robust \citep{bracherEvaluatingEpidemicForecasts2021}.
 Moreover, it is not easily applied to forecasts reported as samples or quantiles, as used in many recent disease forecasting efforts.
-For these reasons, the CRPS or its discrete equivalent, the ranked probability score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021} are more commonly used in epidemiology.
+It is nonetheless occasionally used in epidemiology (see e.g., \citealt{Johansson2019}).
+Especially in recent years, however, the CRPS, its discrete equivalent, the ranked probability score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021} have become popular alternatives.
 
@@
 The CRPS can be understood as a generalisation of the absolute error to predictive distributions, and interpreted on the natural scale of the data.
-The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in so-called COVID-19 Forecast Hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
+The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in the so-called COVID-19 Forecast Hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
 The WIS is defined as
@@
 \end{linenomath*}
-where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, $K$ is the number of predictive quantiles provided and $\boldsymbol{1}()$ is the indicator function.
-The WIS can be decomposed into three components, dispersion, overprediction, underprediction, which reflect the width of the forecast and whether it was centred above or below the observed value.
-We show an alternative definition based on central prediction intervals in Section \ref{sec:alternative-wis} which illustrates this decomposition.
+where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, $K$ is the number of (roughly equally spaced) predictive quantiles provided and $\boldsymbol{1}()$ is the indicator function.
+The WIS can be decomposed into three components, dispersion, overprediction, underprediction, which reflect the spread of the forecast and whether it was centred above or below the observed value.
+We show an alternative definition based on central prediction intervals in Supplement \ref{sec:alternative-wis} which illustrates this decomposition.
 
@@
 In the theoretical discussion in Section \ref{sec:methods}, 'log-transformation' and 'log scale' generally refer to a transformation of $\log_{e}(x)$.
-For practical applications (Section \ref{sec:HUB}) we also use these terms to describe a transformation of $\log_{e}(x + \varepsilon)$ in order to keep the terminology and notation simple.
+For practical applications (Section \ref{sec:HUB}) we also use these terms to describe a transformation of $\log_{e}(x + \varepsilon)$ with a small $\varepsilon > 0$ in order to keep the terminology and notation simple.
 For a prediction target with strictly positive support, the CRPS after applying a log-transformation is given by
@@
 When applying a square-root-transformation before the CRPS, the expected CRPS is independent of the mean for the Poisson-distribution, but not for the other two (with a positive relationship in the normal case and a negative one for the negative binomial).
-As can be seen in Figures \ref{fig:SIM-wis-state-size-mean} and \ref{fig:score-approx}, the approximations presented above work quite well for our simulated example.
+As can be seen in Figures \ref{fig:SIM-wis-state-size-mean} and \ref{fig:score-approx}, the approximations presented in equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear} work quite well for our simulated example.
 
@@
     These expected CRPS values were computed for different predictive means based on 10,000 samples each and are represented by dots.
-Solid lines show the corresponding approximation of the expected CRPS based on an assumed normal distribution as discussed in section \ref{sec:methods:vst}.
+Solid lines show the corresponding approximations of the expected CRPS from equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear}.
 Figure \ref{fig:score-approx} shows the quality of the approximation in more detail.
@@
 
-\subsection{Practical considerations}
+\subsection{Other transformations and practical considerations}
 \label{sec:methods:considerations}
 
-Transformations that are strictly monotonic are permissible in the sense that resulting scores remain proper.
-This is because even though rankings of models may change forecasts will in expectation still minimise their score if they report a predictive distribution that is equal to the data-generating distribution.
-This condition holds for both the log and square root transformations, as well as many others.
-Some strictly monotonic transformations that can generally be applied are scaling by the population size or scaling by past observations.
-This, as discussed in Section \ref{sec:discussion}, is similar to applying a log-transformation, but corresponds to evaluating a forecast of multiplicative, rather than exponential growth rates.
-There may of course be exceptions such as dividing by zero, but otherwise this just represents a monotonic transformation and the remaining score is still proper.
+A natural question is which other transformations could be applied and whether resulting scores remain (strictly) proper.
+In principle, any transformation function can be applied simultaneously to forecasts and observations as long as the definition of the transformation is independent of the forecasts and any quantities unknown at the time of forecasting, including the observed value.
+This simply corresponds to a re-definition of the forecasting target.
+However, applying non-invertible transformations leads to a loss in information conveyed by forecasts, which we consider undesirable.
+The resulting score will be proper, but it may not be strictly proper anymore (as forecasts differing from the forecaster's true belief on the original scale may be identical on the transformed scale).
+When using the CRPS or the WIS, it seems most appropriate to use only strictly monotonic transformations as otherwise the encoded notion of distance may become meaningless.
+If only strictly monotonic transformations are used, then transforming a forecaster's best belief on the original scale will correspond to their best belief on the transformed scale.
+This holds for both the log and square root transformations, as well as many others.
+
+Some other strictly monotonic transformations that can generally be applied prior to computing  the CRPS or WIS are scaling by the population size or scaling by past observations.
+The latter, as discussed in Section \ref{sec:discussion}, is similar to applying a log-transformation, but corresponds to evaluating a forecast of multiplicative, rather than exponential growth rates.
+The arising issue of dividing by zero can again be solved by adding a small offset.
+This just represents a monotonic transformation and the remaining score is still strictly proper.
 Scaling a forecast by the later observed value (as opposed to scaling by past observations) is generally not permissible as it disturbs propriety (see \cite{lerchForecasterDilemmaExtreme2015} on the closely related topic of weighting scores with a function of the observed value).
 Similarly, scaling a forecast by predictions made by other forecasters may result in improper scores, but we are unaware of existing theoretical arguments.
-Scaling scores, e.g.
-dividing a score by the mean score for other forecasters may be acceptable in most applied settings, but we are unaware of existing theoretical arguments that could define exactly when and how this could result in improper scores (as is the case for example when dividing by zero).
 
@@
 We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can be seen that in the example overconfident models perform best in terms of the log WIS.
+Scaling scores (rather than forecasts and observations) by the mean score for other forecasters is closely related to the common practice of computing skill scores.
+While these are typically not technically proper \citep{gneitingStrictlyProperScoring2007}, their application is established practice and considered largely unproblematic.
 
@@
 The choice of the quantity to add does however influence scores and rankings, as measures of relative errors shrink when adding a constant $a$ to the forecast and the observation.
-We illustrate this in Figure \ref{fig:illustration-effect-log-offset}.
+We illustrate this in Figure \ref{fig:illustration-effect-log-offset} and show its effect on relative weights for scores in different locations in Figure \ref{fig:HUB-log-different-offsets}.
 As a rule of thumb, if if $x > 5a$, the difference between $\log{(x + a)}$ and $\log{(x)}$ is small, and it becomes negligible if $x > 50a$.
-Choosing a suitable offset $a$ balances two competing concerns: on the one hand, choosing a small $a$ makes sure that the transformation is as close to a natural logarithm as possible and scores can be interpreted as outlined above.
+Choosing a suitable offset $a$ balances two competing concerns: on the one hand, choosing a small $a$ makes sure that the transformation is as close to a natural logarithm as possible and scores can be interpreted as outlined in the previous sections.
 On the other hand, choosing a larger $a$ can help stabilise scores for forecasts and observations close to zero, avoiding giving excessive weight to forecasts for small quantities (see Figure \ref{fig:HUB-log-different-offsets}).
-For large values of $a$, $\log(x + a)$ grows linearly in $x$, meaning that we expect to observe the same patterns as in the case with no transformation.
+For increasing $a$, less relative weight is given to smaller forecast targets.
+For large values of $a$, $\log(x + a)$ grows again linearly in $x$, meaning that using a large $a$ would be similar to a case with no transformation in terms of the relative weights assigned to different locations.
 For decreasing values of $a$, we give more relative weight to scores in small locations.
@@
 % 4 & all & -0.757 & 0.986 & 0.000 & 0.299 & 0.469 & 0.015\\
-\addlinespace
 1 & Cases & -0.862 & 0.876 & 0.790 & 0.087 & 0.433 & -0.024\\

### 2023-06-24 main.tex
Accept edits made by Johannes

@@
 Probabilistic forecasts \citep{heldProbabilisticForecastingInfectious2017} play an important role in decision-making in epidemiology and public health \citep{doi:10.2105/AJPH.2022.306831}, as well as other areas as diverse as economics \citep{timmermannForecastingMethodsFinance2018} or meteorology \citep{gneitingWeatherForecastingEnsemble2005}.
-Forecasts based on epidemiological modelling in particular has received widespread attention during the COVID-19 pandemic.
+Forecasts based on epidemiological modelling in particular have received widespread attention during the COVID-19 pandemic.
 Evaluations of forecasts can provide feedback for researchers to improve their models and train ensembles.
@@
 Common proper scoring rules are the logarithmic or log score \citep{goodRationalDecisions1952} and the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}.
-The log score is the predictive log density evaluated at the observed value and has many desirable theoretical properties resulting from the likelihood principle; however, the particularly severe penalties it assigns to occasional misguided forecasts make it little robust \citep{bracherEvaluatingEpidemicForecasts2021}.
+The log score is the predictive log density or probability mass evaluated at the observed value.
+It is supported by the likelihood principle \citep{Winkler1996} and has many desirable theoretical properties; however, the particularly severe penalties it assigns to occasional misguided forecasts make it little robust \citep{bracherEvaluatingEpidemicForecasts2021}.
 Moreover, it is not easily applied to forecasts reported as samples or quantiles, as used in many recent disease forecasting efforts.
-It is nonetheless occasionally used in epidemiology (see e.g., \citealt{Johansson2019}).
-Especially in recent years, however, the CRPS, its discrete equivalent, the ranked probability score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021} have become popular alternatives.
+It is nonetheless occasionally used in epidemiology (see e.g., \citealt{heldProbabilisticForecastingInfectious2017, Johansson2019}), but in recent years the CRPS and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021} have become increasingly popular.
 
@@
 The CRPS can be understood as a generalisation of the absolute error to predictive distributions, and interpreted on the natural scale of the data.
-The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in the so-called COVID-19 Forecast Hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
+The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in the so-called COVID-19 Forecast Hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022} and Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US \textit{FluSight} project on influenza forecasting\citep{CdcepiFlusightforecastdata2022}.
 The WIS is defined as
@@
 \end{linenomath*}
-where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, $K$ is the number of (roughly equally spaced) predictive quantiles provided and $\boldsymbol{1}()$ is the indicator function.
-The WIS can be decomposed into three components, dispersion, overprediction, underprediction, which reflect the spread of the forecast and whether it was centred above or below the observed value.
+where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome and $K$ is the number of (roughly equally spaced) predictive quantiles provided.
+The WIS can be decomposed into three components, dispersion, underprediction and overprediction, which reflect the spread of the forecast and whether it was centred above or below the observed value.
 We show an alternative definition based on central prediction intervals in Supplement \ref{sec:alternative-wis} which illustrates this decomposition.
 
-The dynamics of infectious processes are often described by the complementary concepts of the reproduction number $R$ \citep{gosticPracticalConsiderationsMeasuring2020} and growth rate $r$ \citep{wallingaHowGenerationIntervals2007}.
+% The dynamics of infectious processes are often described by the complementary concepts of the reproduction number $R$ \citep{gosticPracticalConsiderationsMeasuring2020} and growth rate $r$ \citep{wallingaHowGenerationIntervals2007}.
 The reproduction number $R$ describes the number of people each infected person is expected to infect in turn (irrespective of the time between infections) and therefore is a measure of the strength of epidemic growth.
 The growth rate $r$ describes the speed of epidemic growth \citep{dushoffSpeedStrengthEpidemic2021}.
-In the absence of changes in population immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
-In that case, the number of new infections in the population grows exponentially in time.
-This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellisChallengesControlCOVID192021}.
+% In the absence of changes in population immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
+% In that case, the number of new infections in the population grows exponentially in time.
+% This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellisChallengesControlCOVID192021}.
 
-If case numbers are evolving based on an exponential process and the  modelling task revolves around estimating and forecasting the reproduction number or the corresponding growth rate, then evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount.
-Consider a forecast that misses the correct growth rate by either $+\epsilon$ or $-\epsilon$:
+The notion of absolute distance encoded by the CRPS and WIS provides a straightforward interpretation, but may not always be the most useful perspective in the context of infectious disease spread.
+Especially in their early phase, outbreaks are best conceived as exponential processes, characterized by potentially time varying reproduction numbers $R_t$ \citep{gosticPracticalConsiderationsMeasuring2020} or epidemic growth rates $r_t$ \citep{dushoffSpeedStrengthEpidemic2021}.
+If the true modelling task revolves around estimating and forecasting these quantities, then evaluating forecasts based on the absolute distance between forecasted and observed incidence values penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount.
+For illustration, consider an incidence forecast issued at time 0 and referring to time $t$ that misses the correct average growth rate $\bar{r}_t$ by either $-\epsilon$ or $+\epsilon$.
+Then the ratio of the resulting absolute errors on the scale of observed incidences $y_{t}$ is
 
@@
 \begin{equation}
-\left| \frac{x_t e^{(r_t + \epsilon) \Delta t} - x_t e^{r_t \Delta t}}{x_t e^{(r_t - \epsilon) \Delta t} - x_t e^{r_t \Delta t}} \right| = e^{\epsilon\Delta t}
+\frac{\left|y_0 \exp[(\bar{r}_t - \epsilon) \times t] - y_0 \exp(\bar{r}_t t)\right|}{\left| y_0 \exp[(\bar{r}_t + \epsilon) \times t] - y_0 \exp(\bar{r}_t t) \right|} = e^{\epsilon t} < 1.
 \end{equation}
@@
 
-For exponential processes errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate. 
-If one is to measure the ability of forecasts to assess and forecast the underlying infection dynamics, it may thus, conceptually, be more desirable to evaluate errors on the growth rate directly.
+% For exponential processes errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate. 
+If one is to measure the ability to forecast the underlying infection dynamics, it may thus be more desirable to evaluate errors on the scale of the growth rate directly.
 
-Evaluating forecasts using the CRPS or WIS means that scores represent a measure of absolute errors.
-However, forecast consumers may find errors on a relative scale easier to interpret and more useful in order to track predictive performance across targets of different orders of magnitude.
+Another argument against using notions of absolute distance between predicted and observed incidence values is that forecast consumers may find errors on a relative scale easier to interpret and more useful in order to track predictive performance across targets of different orders of magnitude.
 \cite{bolinLocalScaleInvariance2021} have proposed the scaled CRPS (SCRPS) which is locally scale invariant; however, it does not correspond to a relative error measure and lacks a straightforward interpretation as available for the CRPS.
 
-A closely related aspect to relative scores (as opposed to absolute scores) is that in the evaluation one may wish to give similar weight to all considered forecast targets.
+% A closely related aspect to relative scores (as opposed to absolute scores) is that in the evaluation one may wish to give similar weight to all 
+Lastly, it may be considered desirable to give all forecast targets similar weight in an overall performance evaluation.
 As the CRPS typically scales with the order of magnitude of the quantity to be predicted, this is not the case for the CRPS, which will typically assign higher scores to forecast targets with high expected values (e.g., in large locations or around the peak of an epidemic).
@@
 \cite{taylorEvaluatingVolatilityInterval1999} in finance, \cite{mayrLogLevelVAR2015} in macroeconomics, \cite{loweStochasticRainfallrunoffForecasting2014} in hydrology or \cite{fuglstadDoesNonstationarySpatial2015} in meteorology).
-While the goal of the transformations is usually to improve the accuracy of the predictions, they can also be used to enhance and complement the evaluation process.
+While the goal of the transformations is often to improve the accuracy of the predictions, they can also be used to enhance and complement the evaluation process.
 In this paper, we argue that the aforementioned issues with evaluating epidemic forecasts based on measures of absolute error on the natural scale can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation.
 Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while guaranteeing that the score remains proper.
-Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects of the forecasts that forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).
+Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).
 
-For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation in the context of epidemic phenomena.
+For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particularly attractive transformation in the context of epidemic phenomena.
 We refer to this transformation as 'log-transformation' and to scores that have been computed from log-transformed forecasts and observations as scores 'on the log scale' (as opposed to scores 'on the natural scale', which involve no transformation).
@@
 Our analysis mostly focuses on the CRPS (or WIS) as an evaluation metric for probabilistic forecasts, given its widespread use throughout the COVID-19 pandemic.
+We note that the logarithmic score has scale invariance properties which imply that score differences between different forecasts are invariant to strictly monotonic transformations (see \citealt{Lehmann1950} on corresponding properties of likelihood ratios).
+The issues discussed here thus do not apply to this score.
 
@@
 
-Specifically, if the predictive distribution for the quantity $Y$ equals the true data-generating process $F$ (an ideal forecast), the expected CRPS is given by \citep{gneitingStrictlyProperScoring2007}
+If the predictive distribution for the quantity $Y$ equals the true data-generating process $F$ (an ideal forecast), the expected CRPS is given by \citep{gneitingStrictlyProperScoring2007}
 \begin{linenomath*}
@@
 
-If the mean-variance relationship of the data-generating distribution is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST \citep{guerreroTimeseriesAnalysisSupported1993}.
-Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method (a first-order Taylor approximation, see e.g., \citealt{Dunn2018}), to show that
+If the mean-variance relationship of the data-generating distribution is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST.
+Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method (a first-order Taylor approximation, see e.g., \citealt{dunnGeneralizedLinearModels2018}), to show that
 
@@
 \end{linenomath*}
-As $\sigma$ and $\mu$ are linked through the quadratic mean-variance relationship ($\sigma = \sqrt{c} \times \mu$), the expected CRPS stays constant regardless of the expected value of the data-generating distribution $\mu$.
+As $\sigma$ and $\mu$ are linked through the quadratic mean-variance relationship ($\sigma = \sqrt{c} \times \mu$), the expected CRPS thus stays constant regardless of the expected value of the data-generating distribution $\mu$.
 The assumption of a quadratic mean-variance relationship is closely linked to the aspects discussed in Sections \ref{sec:methods:relative} and \ref{sec:methods:growthrate}.
@@
 
-If the variance is linear with $\sigma^2 = c \times \mu$, as with a Poisson-distributed variable, the square root is known to be a VST \citep{Dunn2018}. 
+If the mean-variance relationship is linear with $\sigma^2 = c \times \mu$, as with a Poisson-distributed variable, the square root is known to be a VST \citep{dunnGeneralizedLinearModels2018}. 
 Denoting by $F_{\sqrt{\ }}$ the predictive distribution for $\sqrt{Y}$, the delta method can again be used to show that
@@
 We note that while standard in the derivation of variance-stabilizing transformations, the application of the delta method in equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear} requires the probability mass of $F$ to be tightly distributed.
-If this is not the case, the approximation may be less accurate.
+If this is not the case, the approximation and thus the variance stabilization may be less accurate.
 
 To strengthen our intuition on how transforming outcomes prior to applying the CRPS shifts the emphasis between targets with high and low expected outcomes, Figure \ref{fig:SIM-wis-state-size-mean} shows the expected CRPS of ideal forecasters under different mean-variance relationships and transformations.
-We consider a Poisson distribution where $\sigma^2 = \mu$, a negative binomial distribution with size parameter $\theta = 10$ and thus $\sigma^2 = \mu + \mu^2/10$, and a normal distribution with constant variance.
+We consider a Poisson distribution where $\sigma^2 = \mu$, a negative binomial distribution with size parameter $\theta = 10$ and thus $\sigma^2 = \mu + \mu^2/10$, and a truncated normal distribution with practically constant variance.
 We see that when applying the CRPS on the natural scale, the expected CRPS grows monotonically as the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster) increases.
@@
 When applying a log-transformation first, the expected CRPS is almost independent of $\mu$ for the negative binomial distribution and large $\mu$, while smaller targets have higher expected CRPS in case of the Poisson distribution and the normal distribution with constant variance.
-When applying a square-root-transformation before the CRPS, the expected CRPS is independent of the mean for the Poisson-distribution, but not for the other two (with a positive relationship in the normal case and a negative one for the negative binomial).
+When applying a square-root-transformation, the expected CRPS is independent of the mean for the Poisson-distribution, but not for the other two (with a positive relationship in the normal case and a negative one for the negative binomial).
 As can be seen in Figures \ref{fig:SIM-wis-state-size-mean} and \ref{fig:score-approx}, the approximations presented in equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear} work quite well for our simulated example.
@@
 Expected CRPS was computed on the natural scale (left), after applying a square-root transformation (middle), and after adding one and applying a log-transformation to the data (right).
-B: A but with x axis on the log scale.}
+B: A but with x and y axes on the log scale.}
     \label{fig:SIM-wis-state-size-mean}
 \end{figure}
-
-\subsection{Other transformations and practical considerations}
-\label{sec:methods:considerations}
-
-A natural question is which other transformations could be applied and whether resulting scores remain (strictly) proper.
-In principle, any transformation function can be applied simultaneously to forecasts and observations as long as the definition of the transformation is independent of the forecasts and any quantities unknown at the time of forecasting, including the observed value.
-This simply corresponds to a re-definition of the forecasting target.
-However, applying non-invertible transformations leads to a loss in information conveyed by forecasts, which we consider undesirable.
-The resulting score will be proper, but it may not be strictly proper anymore (as forecasts differing from the forecaster's true belief on the original scale may be identical on the transformed scale).
-When using the CRPS or the WIS, it seems most appropriate to use only strictly monotonic transformations as otherwise the encoded notion of distance may become meaningless.
-If only strictly monotonic transformations are used, then transforming a forecaster's best belief on the original scale will correspond to their best belief on the transformed scale.
-This holds for both the log and square root transformations, as well as many others.
-
-Some other strictly monotonic transformations that can generally be applied prior to computing  the CRPS or WIS are scaling by the population size or scaling by past observations.
-The latter, as discussed in Section \ref{sec:discussion}, is similar to applying a log-transformation, but corresponds to evaluating a forecast of multiplicative, rather than exponential growth rates.
-The arising issue of dividing by zero can again be solved by adding a small offset.
-This just represents a monotonic transformation and the remaining score is still strictly proper.
-Scaling a forecast by the later observed value (as opposed to scaling by past observations) is generally not permissible as it disturbs propriety (see \cite{lerchForecasterDilemmaExtreme2015} on the closely related topic of weighting scores with a function of the observed value).
-Similarly, scaling a forecast by predictions made by other forecasters may result in improper scores, but we are unaware of existing theoretical arguments.
-
-When applying a transformation, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee that the score remains proper.
-In the case of log transforms, taking the logarithm of the scores, rather than scoring the log-transformed forecasts and data, results in an improper score.
-We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can be seen that in the example overconfident models perform best in terms of the log WIS.
-Scaling scores (rather than forecasts and observations) by the mean score for other forecasters is closely related to the common practice of computing skill scores.
-While these are typically not technically proper \citep{gneitingStrictlyProperScoring2007}, their application is established practice and considered largely unproblematic.
-
-In practice, one issue with the log transform is that they are not readily applicable to negative numbers or zero values, which need to be removed or otherwise handled. 
-One common approach to deal with zeros is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm \citep{bellegoDealingLogsZeros2022}.
-This represents a strictly monotonic transformation and therefore guarantees that the resulting score remains proper.
-The choice of the quantity to add does however influence scores and rankings, as measures of relative errors shrink when adding a constant $a$ to the forecast and the observation.
-We illustrate this in Figure \ref{fig:illustration-effect-log-offset} and show its effect on relative weights for scores in different locations in Figure \ref{fig:HUB-log-different-offsets}.
-As a rule of thumb, if if $x > 5a$, the difference between $\log{(x + a)}$ and $\log{(x)}$ is small, and it becomes negligible if $x > 50a$.
-Choosing a suitable offset $a$ balances two competing concerns: on the one hand, choosing a small $a$ makes sure that the transformation is as close to a natural logarithm as possible and scores can be interpreted as outlined in the previous sections.
-On the other hand, choosing a larger $a$ can help stabilise scores for forecasts and observations close to zero, avoiding giving excessive weight to forecasts for small quantities (see Figure \ref{fig:HUB-log-different-offsets}).
-For increasing $a$, less relative weight is given to smaller forecast targets.
-For large values of $a$, $\log(x + a)$ grows again linearly in $x$, meaning that using a large $a$ would be similar to a case with no transformation in terms of the relative weights assigned to different locations.
-For decreasing values of $a$, we give more relative weight to scores in small locations.
-In principle, a user could produce a plot similar to Figure \ref{fig:HUB-log-different-offsets} with different values of the offset $a$ and choose $a$ such that the relative weightings of times and regions with high and low incidence correspond to their preferences.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-scores-locations-log-variants.png}
-    \caption{Mean WIS in different locations for different transformations applied before scoring.
-Locations are sorted according to the mean observed value in that location.
-Shown are scores for two-week-ahead forecasts of the EuroCOVIDhub-ensemble.
-On the natural scale (with no transformation prior to applying the WIS), scores correlate strongly with the average number of observed values in a given location.
-The same is true for scores obtained after applying a square-root transformation, or after applying a log-transformation with a large offset $a$.
-For illustrative purposes, $a$ was chosen to be 101630 for cases and 530 for deaths, 10 times the respective median observed value.
-For large values of $a$, $log(x + a)$ grows linearly in $x$, meaning that we expect to observe the same patterns as in the case with no transformation.
-For decreasing values of $a$, we give more relative weight to scores in small locations.}
-    \label{fig:HUB-log-different-offsets}
-\end{figure}
-
-A related issue occurs when the predictive distribution has a large probability mass on zero (or on very small values), as this can translate into an excessively wide forecast in relative terms.
-This can be seen in Figure \ref{fig:HUB-model-comparison-baseline}.
-Here, the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm because forecasts contained zero in its prediction intervals.
-To deal with this issue one could potentially choose to use a higher $a$ value when applying a transformation $\log(x + a)$, for example $a = 10$ instead of the $a = 1$ that we chose to use.
-
-We note that in the practical evaluation of operational forecasting systems  several additional challenges arise, which we do not study in detail.
-These concern e.g., the removal of outlying observations and forecasts and the handling of missing forecasts.
-The solutions we employed in practice are provided in Section \ref{sec:HUB-setting}.
 
@@
 On the log scale, however, forecaster A receives a lower score for large observed values, being more heavily penalised for assigning large probability to small values (which, in relative terms, are far away from the actual observation).
-We note that the chosen example involving a geometric forecast distribution is somewhat constructed; as shown in Section \ref{sec:Hub:cor} and Figure \ref{fig:HUB-cors}A, rankings between models usually stay quite stable for a single forecast.
+We note that the chosen example involving a geometric forecast distribution is somewhat constructed; as shown in Section \ref{sec:Hub:cor} and Figure \ref{fig:HUB-cors}A, rankings between models in practice stay quite stable for a single forecast.
 
@@
 
-Overall model rankings would be expected to differ even more when scores are averaged across multiple forecasts or targets.
+Overall model rankings would be expected to differ more when scores are averaged across multiple forecasts or targets.
 The change in rankings of aggregate scores usually is mainly driven by the order of magnitude of scores for different forecast targets across time, location and target type and less so by the kind of changes in model rankings for single forecasts discussed above (see Figure \ref{fig:HUB-cors} for a practical example).
@@
 Depending on the relationship between the mean and variance of the forecast target, a log-transformation may even lead to systematically larger scores assigned to small forecast targets, as illustrated in Figure \ref{fig:SIM-wis-state-size-mean}.
+
+\subsection{Practical considerations and other transformations}
+\label{sec:methods:considerations}
+
+In practice, one issue with the log transform is that it is not readily applicable to negative or zero values, which need to be removed or otherwise handled. 
+One common approach to this end is to add a small positive quantity, such as $a = 1$, to all observations and predictions before taking the logarithm \citep{bellegoDealingLogsZeros2022}.
+This still represents a strictly monotonic transformation, but the choice of $a$ does influence scores and rankings (measures of relative errors shrink the larger the chosen value $a$).
+As a rule of thumb, if if $x > 5a$, the difference between $\log{(x + a)}$ and $\log{(x)}$ is small, and it becomes negligible if $x > 50a$.
+Choosing a suitable offset $a$ thus balances two competing concerns: on the one hand, choosing a small $a$ makes sure that the transformation is as close to a natural logarithm as possible and scores can be interpreted as outlined in the previous sections.
+On the other hand, choosing a larger $a$ can help stabilise scores for forecasts and observations close to zero, avoiding giving excessive weight to forecasts of small quantities.
+For increasing $a$, less relative weight is given to smaller forecast targets.
+For very large values of $a$, $\log(x + a)$ is roughly linear in $x$, so that using a very large $a$ implies similar relative weighting as applying no transformation at all.
+In practice, a user could explore the effect of different values of $a$ graphically and choose $a$ such that the relative weightings of times and regions with high and low incidence correspond to their preferences (see Figure \ref{fig:HUB-log-different-offsets} in our example application, Section \ref{sec:HUB}).
+
+A related issue occurs when the predictive distribution has a large probability mass on zero (or on very small values), as this can translate into an excessively wide forecast in relative terms.
+In our applied example this is illustrated in Figure \ref{fig:HUB-model-comparison-baseline}.
+Here, the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm because forecasts contained zero in its prediction intervals.
+To deal with this issue one could choose to use a higher $a$ value when applying a transformation $\log(x + a)$, for example $a = 10$ instead of the $a = 1$ that we chose to use.
+
+A natural question is which other transformations could be applied and whether resulting scores remain (strictly) proper.
+In principle, any transformation function can be applied simultaneously to forecasts and observations as long as the definition of the transformation is independent of the forecasts and any quantities unknown at the time of forecasting, including the observed value.
+This simply corresponds to a re-definition of the forecasting target.
+However, applying non-invertible transformations leads to a loss in information conveyed by forecasts, which we consider undesirable.
+The resulting score will be proper, but it may not be strictly proper anymore (as forecasts differing from the forecaster's true belief on the original scale may be identical on the transformed scale).
+When using the CRPS or the WIS, it seems most appropriate to use only strictly monotonic transformations such as the natural logarithm or the square root as otherwise the encoded notion of distance may become meaningless.
+
+Some other strictly monotonic transformations that can be applied are scaling by the population size or scaling by past observations.
+The latter, as discussed in Section \ref{sec:discussion}, is similar to applying a log-transformation, but corresponds to evaluating a forecast of multiplicative, rather than exponential growth rates.
+The arising issue of dividing by zero can again be solved by adding a small offset $a$.
+Scaling a forecast by the later observed value (as opposed to scaling by past observations) is generally not permissible as it disturbs propriety (see \citealt{lerchForecasterDilemmaExtreme2015} on the closely related topic of weighting scores with a function of the observed value).
+Similarly, scaling a forecast by predictions made by a function of the predictive distribution may lead to improper scoes, though we are unaware of existing theoretical arguments.
+
+When applying a transformation, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee that the score remains proper.
+In the case of log transforms, taking the logarithm of the CRPS values, rather than scoring the log-transformed forecasts and data, results in an improper score.
+We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can be seen that in the example overconfident models perform best in terms of the log WIS.
+We note that strictly speaking, re-scaling scores by the score of a baseline model or the average score across different models  to obtain skill scores likewise leads to improper scores\citep{gneitingStrictlyProperScoring2007}.
+The application of such skill scores, however, is established practice and considered largely unproblematic.
+
+We note that in the practical evaluation of operational forecasting systems  several additional challenges arise, which we do not study in detail.
+These concern e.g., the removal of outlying observations and forecasts and the handling of missing forecasts.
+The solutions we employed in practice are provided in Section \ref{sec:HUB-setting}.
 
@@
 This reflects the increasing difficulty of forecasts further into the future and, for the log scale, corresponds with our expectations from Section \ref{sec:methods:growthrate}.
+
+To assess the impact of the choice of offset value $a$ we extend the display from Figure \ref{fig:HUB-mean-locations}C by results obtained under different specifications.
+Results are shown in Figure \ref{fig:HUB-log-different-offsets}, where for completeness we also added the square root transformation.
+As discussed in Section \ref{sec:methods:considerations}, smaller values of $a$ increase the relative weight of smaller locations in the overall evaluation.
+In the most extreme considered case $a = 0.001$, the smallest locations in fact receive the largest weight both for deaths and cases.
+For very large values (see the third row of Figure \ref{fig:HUB-log-different-offsets}), the relative weights strongly resemble those of the evaluation on the natural scale.
+We recommend using displays of this type to get an intuition for the role different locations may play for overall evaluation results.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/HUB-scores-locations-log-variants.png}
+    \caption{Mean WIS in different locations for different transformations applied before scoring.
+Locations are sorted according to the mean observed value in that location.
+Shown are scores for two-week-ahead forecasts of the EuroCOVIDhub-ensemble.
+On the natural scale (with no transformation prior to applying the WIS), scores correlate strongly with the average number of observed values in a given location.
+The same is true for scores obtained after applying a square-root transformation, or after applying a log-transformation with a large offset $a$.
+For illustrative purposes, $a$ was chosen to be 101630 for cases and 530 for deaths, 10 times the respective median observed value.
+For large values of $a$, $log(x + a)$ grows linearly in $x$, meaning that we expect to observe the same patterns as in the case with no transformation.
+For decreasing values of $a$, we give more relative weight to scores in small locations.}
+    \label{fig:HUB-log-different-offsets}
+\end{figure}
 
@@
 
-In this work, we focused on the CRPS and WIS, which are widely used in the evaluation of epidemic forecasts.
-We note that for the logarithmic score, which has also been used e.g., in some editions of the FluSight challenge \cite{reichCollaborativeMultiyearMultimodel2019}, the question of the right scale to evaluate forecasts does not arise.
+%In this work, we focused on the CRPS and WIS, which are widely used in the evaluation of epidemic forecasts.
+As mentioned previously the logarithmic score, the question of the right scale to evaluate forecasts does not arise.
 It is known that log score differences between different forecasters are invariant to monotonic transformations of the outcome variable (see e.g., \citealt{diksLikelihoodbasedScoringRules2011}).
@@
 Instead of defining the WIS as an average of scores for individual quantiles, we can define it using an average of scores for symmetric predictive intervals.
-For a single prediction interval, the interval scoren (IS) is computed as the sum of three penalty components, dispersion (width of the prediction interval), underprediction and overprediction,
+For a single prediction interval, the interval score (IS) is computed as the sum of three penalty components, dispersion (width of the prediction interval), underprediction and overprediction,
 %

### 2023-06-26 main.tex
more changes to the manuscript

@@
 
-% The dynamics of infectious processes are often described by the complementary concepts of the reproduction number $R$ \citep{gosticPracticalConsiderationsMeasuring2020} and growth rate $r$ \citep{wallingaHowGenerationIntervals2007}.
-The reproduction number $R$ describes the number of people each infected person is expected to infect in turn (irrespective of the time between infections) and therefore is a measure of the strength of epidemic growth.
-The growth rate $r$ describes the speed of epidemic growth \citep{dushoffSpeedStrengthEpidemic2021}.
-% In the absence of changes in population immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
-% In that case, the number of new infections in the population grows exponentially in time.
-% This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellisChallengesControlCOVID192021}.
-
 The notion of absolute distance encoded by the CRPS and WIS provides a straightforward interpretation, but may not always be the most useful perspective in the context of infectious disease spread.
@@
 \begin{equation}
-\frac{\left|y_0 \exp[(\bar{r}_t - \epsilon) \times t] - y_0 \exp(\bar{r}_t t)\right|}{\left| y_0 \exp[(\bar{r}_t + \epsilon) \times t] - y_0 \exp(\bar{r}_t t) \right|} = e^{\epsilon t} < 1.
+\frac{\left|y_0 \exp[(\bar{r}_t - \epsilon) \times t] - y_0 \exp(\bar{r}_t t)\right|}{\left| y_0 \exp[(\bar{r}_t + \epsilon) \times t] - y_0 \exp(\bar{r}_t t) \right|} = \exp(-\epsilon t) < 1.
 \end{equation}
@@
 In the theoretical discussion in Section \ref{sec:methods}, 'log-transformation' and 'log scale' generally refer to a transformation of $\log_{e}(x)$.
-For practical applications (Section \ref{sec:HUB}) we also use these terms to describe a transformation of $\log_{e}(x + \varepsilon)$ with a small $\varepsilon > 0$ in order to keep the terminology and notation simple.
+For practical applications (Section \ref{sec:HUB}) we also use these terms to describe a transformation of $\log_{e}(x + a)$ with a small $a > 0$ in order to keep the terminology and notation simple.
 For a prediction target with strictly positive support, the CRPS after applying a log-transformation is given by
@@
 Our analysis mostly focuses on the CRPS (or WIS) as an evaluation metric for probabilistic forecasts, given its widespread use throughout the COVID-19 pandemic.
-We note that the logarithmic score has scale invariance properties which imply that score differences between different forecasts are invariant to strictly monotonic transformations (see \citealt{Lehmann1950} on corresponding properties of likelihood ratios).
-The issues discussed here thus do not apply to this score.
+We note that the logarithmic score has scale invariance properties which imply that score differences between different forecasts are invariant to strictly monotonic transformations (see \citealt{Lehmann1950} on corresponding properties of likelihood ratios and \citealt{diksLikelihoodbasedScoringRules2011}).
+The question of the right scale to evaluate forecasts on does therefore not arise for the log score.
 
@@
 In Sections \ref{sec:methods:relative}--\ref{sec:methods:vst} we provide some mathematical intuition on applying the log-transformation prior to evaluating the CRPS, highlighting the connections to relative error measures, the epidemic growth rate and variance stabilizing transformations.
-We then discuss practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}) and the effect of the log-transformation on forecast rankings (Section \ref{sec:methods:rankings}).
+We then discuss the effect of the log-transformation on forecast rankings (Section \ref{sec:methods:rankings}) as well as practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}).
 To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  \citep[ Section \ref{sec:HUB}]{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
@@
 \end{linenomath*}
-As $\sigma$ and $\mu$ are linked through the quadratic mean-variance relationship ($\sigma = \sqrt{c} \times \mu$), the expected CRPS thus stays constant regardless of the expected value of the data-generating distribution $\mu$.
+As $\sigma$ and $\mu$ are linked through the quadratic mean-variance relationship (or linear mean-standard deviation relationship, $\sigma = \sqrt{c} \times \mu$), the expected CRPS thus stays constant regardless of the expected value of the data-generating distribution $\mu$.
 The assumption of a quadratic mean-variance relationship is closely linked to the aspects discussed in Sections \ref{sec:methods:relative} and \ref{sec:methods:growthrate}.
@@
 Large observations will dominate average CRPS values when evaluation is done on the natural scale, but much less so after log transformation.
-Depending on the relationship between the mean and variance of the forecast target, a log-transformation may even lead to systematically larger scores assigned to small forecast targets, as illustrated in Figure \ref{fig:SIM-wis-state-size-mean}.
+Depending on how different models perform across targets of different orders of magnitude, rankings in terms of average scores may change when applying a transformation.
 
@@
 In our applied example this is illustrated in Figure \ref{fig:HUB-model-comparison-baseline}.
-Here, the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm because forecasts contained zero in its prediction intervals.
+In such cases, the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm because forecasts contained zero in its prediction intervals.
 To deal with this issue one could choose to use a higher $a$ value when applying a transformation $\log(x + a)$, for example $a = 10$ instead of the $a = 1$ that we chose to use.
@@
 The arising issue of dividing by zero can again be solved by adding a small offset $a$.
-Scaling a forecast by the later observed value (as opposed to scaling by past observations) is generally not permissible as it disturbs propriety (see \citealt{lerchForecasterDilemmaExtreme2015} on the closely related topic of weighting scores with a function of the observed value).
-Similarly, scaling a forecast by predictions made by a function of the predictive distribution may lead to improper scoes, though we are unaware of existing theoretical arguments.
+Scaling a forecast by the later observed value (as opposed to scaling by past observations) is generally not permissible as it can result in improper scores (see \citealt{lerchForecasterDilemmaExtreme2015} on the closely related topic of weighting scores with a function of the observed value).
+Similarly, scaling forecasts and observations by a function of the predictive distribution (like the predictive mean) may lead to improper scores; however, we are unaware of existing theoretical arguments on this.
 
@@
 We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can be seen that in the example overconfident models perform best in terms of the log WIS.
-We note that strictly speaking, re-scaling scores by the score of a baseline model or the average score across different models  to obtain skill scores likewise leads to improper scores\citep{gneitingStrictlyProperScoring2007}.
+We note that strictly speaking, re-scaling average scores by the average score of a baseline model or average scores across different models  to obtain skill scores likewise leads to improper scores \citep{gneitingStrictlyProperScoring2007}.
 The application of such skill scores, however, is established practice and considered largely unproblematic.
@@
 Conversely, failure to predict an upswing while numbers are still low, is less severely penalised on the natural scale (see forecasts in July 2021 and to a lesser extent in July 2022 in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low.
-On the log scale, missing lower inflection points tends to lead to more severe penalties (see Figure \ref{fig:HUB-model-comparison-ensemble}C, D)).
+On the log scale, missing lower inflection points tends to lead to more severe penalties (see Figure \ref{fig:HUB-model-comparison-ensemble}C, D).
 One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure \ref{fig:HUB-model-comparison-ensemble}E, F).
@@
 For illustrative purposes, $a$ was chosen to be 101630 for cases and 530 for deaths, 10 times the respective median observed value.
-For large values of $a$, $log(x + a)$ grows linearly in $x$, meaning that we expect to observe the same patterns as in the case with no transformation.
+For large values of $a$, $\log(x + a)$ grows roughly linearly in $x$, meaning that we expect to observe the same patterns as in the case with no transformation.
 For decreasing values of $a$, we give more relative weight to scores in small locations.}
@@
 A value of $\beta = 0.5$ would imply a linear median-variance relationship, suggesting the square root as a VST.
-We applied the regression to case and death forecasts, pooled across horizons and stratified for one through four-week-ahead forecasts.
+We applied the regression to case and death forecasts, stratified for one through four-week-ahead forecasts.
 Results are provided in Table \ref{tab:HUB-regression}.
@@
 
-%In this work, we focused on the CRPS and WIS, which are widely used in the evaluation of epidemic forecasts.
-As mentioned previously the logarithmic score, the question of the right scale to evaluate forecasts does not arise.
-It is known that log score differences between different forecasters are invariant to monotonic transformations of the outcome variable (see e.g., \citealt{diksLikelihoodbasedScoringRules2011}).
-This is clearly an advantage of the logarithmic score over the CRPS; however, the logarithmic score is known to have other severe downsides, e.g., its low robustness to sporadically misguided forecasts; see \cite{bracherEvaluatingEpidemicForecasts2021} for a more detailed discussion.
-
 Exploring transformations is a promising avenue for future work that could help bridge the gap between modellers and policymakers by providing scoring rules that better reflect what forecast consumers care about.
@@
 \toprule
-target\_type & quantity & measure & natural & log\\
+target type & quantity & measure & natural & log\\
 \midrule

### 2023-06-26 R/00 full paper code.R
Update figure legend

@@ -413,12 +413,12 @@ plot_fct <- function(scores, scale_factor, nbinom, filter_scale = "natural") {
     labs(y = "CRPS", x = "Observed value") + 
     coord_cartesian(xlim = c(0, 30)) +
     scale_colour_discrete(
-      labels=c(TeX(r'(A: $\sigma^2 = \mu + \mu^2$)'), 
-               TeX(r'(B: $\sigma^2 = \mu$)'))
+      labels=c("A: Geometric distribution", 
+               "B: Poisson distribution")
     ) + 
     scale_fill_discrete(
-      labels=c(TeX(r'(A: $\sigma^2 = \mu + \mu^2$)'), 
-               TeX(r'(B: $\sigma^2 = \mu$)'))
+      labels=c("A: Geometric distribution", 
+               "B: Poisson distribution")
     ) + 
     facet_wrap(~ scale)
 }

### 2023-06-26 main.tex
Updates from Overleaf

@@
 In our applied example this is illustrated in Figure \ref{fig:HUB-model-comparison-baseline}.
-In such cases, the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm because forecasts contained zero in its prediction intervals.
+In such instances, the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm because forecasts contained zero in its prediction intervals.
 To deal with this issue one could choose to use a higher $a$ value when applying a transformation $\log(x + a)$, for example $a = 10$ instead of the $a = 1$ that we chose to use.
@@
 On the other hand, scoring forecasts on the log scale may be less influenced by outliers and better reflect consistent performance across time, space, and forecast targets.
-It also gives higher weight to another type of situation one may care about, namely one in which numbers start to rise from a previously low level.
+Furthermore, decision makers may specifically care about situations in which numbers start to rise from a previously low level.
 
@@
 Exploring transformations is a promising avenue for future work that could help bridge the gap between modellers and policymakers by providing scoring rules that better reflect what forecast consumers care about.
-{In this paper, we did not make any particular assumptions about policy makers' priorities and preferences.
+In this paper, we did not make any particular assumptions about policy makers' priorities and preferences.
 Rather, we aimed to enable users to make an informed choice by showing how different transformations lead to different relative weights for the kinds of prediction errors forecast consumers may care about, such as absolute vs.

### 2023-06-26 reply_to_reviews.tex
more changes to the manuscript

--- 2023-06-21 reply_to_reviews.tex
@@
 Please see our responses to the individual comments as well as explanations on the changes we have made to the manuscript below.
+In addition to the changes outlined below we some general edits to the paper.
+We changed the title to make it more concise, we altered the order of some sections and shortened the introduction a bit, which became quite lengthy after incorporating all the suggestions.
 
@@
 Proper scoring rules are constructed such that they encourage honest forecasting and cannot be `gamed' or `cheated'. 
-Assuming that the forecaster's actual best judgement corresponds to a predictive distribution $F$, a proper score is constructed such that under $F$, no other distribution $G$ yields a better expected score.
+Assuming that the forecaster's actual best judgement corresponds to a predictive distribution $F$, a proper score is constructed such that if $F$ was the data-generating process, no other distribution $G$ would yield a better expected score.
+A scoring rule is called \textit{strictly} proper if there is no other distribution that under $F$ achieves the \textit{same} expected score as $F$, meaning that any deviation from $F$ leads to a worsening of expected scores.
 Forecasters (anyone or anything that issues a forecast) are thus incentivised to report their true belief $F$ about the future.
-Common proper scoring rules are the logarithmic or log score \citep{goodRationalDecisions1952} or the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}.
-The log score is the log-likelihood of the predictive density at the observed value and is scale-invariant under monotonous transformations.
-The log score does not suffer from the issues we address in this paper, but due to a number of reasons is not as widely used in fields such as weather forecasting or epidemiology (see e.g., \citealt{Johansson2019} for an exception).
-For example, the log score assigns particularly severe penalties to occasional misguided forecasts.
-This lack of robustness may limit its practical applicability \cite{bracherEvaluatingEpidemicForecasts2021}.
+Common proper scoring rules are the logarithmic or log score \citep{goodRationalDecisions1952} and the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}.
+The log score is the predictive log density or probability mass evaluated at the observed value.
+It is supported by the likelihood principle \citep{Winkler1996} and has many desirable theoretical properties; however, the particularly severe penalties it assigns to occasional misguided forecasts make it little robust \citep{bracherEvaluatingEpidemicForecasts2021}.
 Moreover, it is not easily applied to forecasts reported as samples or quantiles, as used in many recent disease forecasting efforts.
-For these reasons, the CRPS or its discrete equivalent, the ranked probability score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021} are more commonly used in epidemiology.
+It is nonetheless occasionally used in epidemiology (see e.g., \citealt{heldProbabilisticForecastingInfectious2017, Johansson2019}), but in recent years the CRPS and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021} have become increasingly popular.
 
@@
 The log score is indeed an alternative which is scale invariant under monotonic transformations.
-We mention this in the introduction and the the discussion and state that for the log score the question of which scale to evaluate forecasts on does not arise.
+We mention this in the introduction and state that for the log score the question of which scale to evaluate forecasts on does not arise.
 In our understanding, a quasi-likelihood approach (as opposed to an actual likelihood approach as in the log score) may not actually be necessary as at least in theory full predictive distributions are to be evaluated.
@@
 \black
-This is a good point, we added numbers to all equations.
+You are completely right, we added numbers to all equations.
 
@@
 
-Thank you for your thorough and extensive comments.
+Thank you for your thorough and helpful comments.
 
@@
 \indented
-For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation in the context of epidemic phenomena.
+For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particularly attractive transformation in the context of epidemic phenomena.
 We refer to this transformation as 'log-transformation' and to scores that have been computed from log-transformed forecasts and observations as scores 'on the log scale' (as opposed to scores 'on the natural scale', which involve no transformation).
 In the theoretical discussion in Section 2, 'log-transformation' and 'log scale' generally refer to a transformation of $\log_{e}(x)$.
-For practical applications (Section 3) we also use these terms to describe a transformation of $\log_{e}(x + \varepsilon)$ in order to keep the terminology and notation simple.
+For practical applications (Section 3) we also use these terms to describe a transformation of $\log_{e}(x + a)$ with a small $a > 0$ in order to keep the terminology and notation simple.
+For a prediction target with strictly positive support, the CRPS after applying a log-transformation is given by
 
@@
 dividing a score by the mean score for other forecasters may be acceptable in most applied settings, but we are unsure about exactly when and how this could break propriety (as is the case for example when dividing by zero).
+
+To make this clearer in the text, we added the following paragraphs:
+
+\indented
+A natural question is which other transformations could be applied and whether resulting scores remain (strictly) proper.
+In principle, any transformation function can be applied simultaneously to forecasts and observations as long as the definition of the transformation is independent of the forecasts and any quantities unknown at the time of forecasting, including the observed value.
+This simply corresponds to a re-definition of the forecasting target.
+However, applying non-invertible transformations leads to a loss in information conveyed by forecasts, which we consider undesirable.
+The resulting score will be proper, but it may not be strictly proper anymore (as forecasts differing from the forecaster's true belief on the original scale may be identical on the transformed scale).
+When using the CRPS or the WIS, it seems most appropriate to use only strictly monotonic transformations such as the natural logarithm or the square root as otherwise the encoded notion of distance may become meaningless.
+
+Some other strictly monotonic transformations that can be applied are scaling by the population size or scaling by past observations.
+The latter, as discussed in Section 4, is similar to applying a log-transformation, but corresponds to evaluating a forecast of multiplicative, rather than exponential growth rates.
+The arising issue of dividing by zero can again be solved by adding a small offset $a$.
+Scaling a forecast by the later observed value (as opposed to scaling by past observations) is generally not permissible as it can result in improper scores (see \citealt{lerchForecasterDilemmaExtreme2015} on the closely related topic of weighting scores with a function of the observed value).
+Similarly, scaling forecasts and observations by a function of the predictive distribution (like the predictive mean) may lead to improper scores; however, we are unaware of existing theoretical arguments on this.
+
+When applying a transformation, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee that the score remains proper.
+In the case of log transforms, taking the logarithm of the CRPS values, rather than scoring the log-transformed forecasts and data, results in an improper score.
+We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can be seen that in the example overconfident models perform best in terms of the log WIS.
+We note that strictly speaking, re-scaling average scores by the average score of a baseline model or average scores across different models  to obtain skill scores likewise leads to improper scores \citep{gneitingStrictlyProperScoring2007}.
+The application of such skill scores, however, is established practice and considered largely unproblematic.
+
+\notindented
 
@@
 \black
-We updated the sentence to make this clearer. The new sentence is now:
-
-\indented
-The reproduction number $R$ describes the number of people each infected person is expected to infect in turn (irrespective of the time between infections) and therefore is a measure of the strength of epidemic growth.
-The growth rate $r$ describes the speed of epidemic growth.
+We revised and shortened the introduction, so this sentence has been removed. Now, it simply reads:
+
+\indented
+
+Especially in their early phase, outbreaks are best conceived as exponential processes, characterized by potentially time varying reproduction numbers $R_t$ \citep{gosticPracticalConsiderationsMeasuring2020} or epidemic growth rates $r_t$ \citep{dushoffSpeedStrengthEpidemic2021}.
 
@@
 \black
-We changed this to "population immunity" as suggested.
+We agree with you, but have simplified the introduction such that this sentence does no longer appear.
 
@@
 \indented
-If case numbers are evolving based on an exponential process and the  modelling task revolves around estimating and forecasting the reproduction number or the corresponding growth rate, then evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount.
-Consider a forecast that misses the correct growth rate by either $+\epsilon$ or $-\epsilon$:
+If the true modelling task revolves around estimating and forecasting these quantities, then evaluating forecasts based on the absolute distance between forecasted and observed incidence values penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount.
+For illustration, consider an incidence forecast issued at time 0 and referring to time $t$ that misses the correct average growth rate $\bar{r}_t$ by either $-\epsilon$ or $+\epsilon$.
+Then the ratio of the resulting absolute errors on the scale of observed incidences $y_{t}$ is
 
 \begin{equation}
-\left| \frac{x_t e^{(r_t + \epsilon) \Delta t} - x_t e^{r_t \Delta t}}{x_t e^{(r_t - \epsilon) \Delta t} - x_t e^{r_t \Delta t}} \right| = e^{\epsilon\Delta t}
+\frac{\left|y_0 \exp[(\bar{r}_t - \epsilon) \times t] - y_0 \exp(\bar{r}_t t)\right|}{\left| y_0 \exp[(\bar{r}_t + \epsilon) \times t] - y_0 \exp(\bar{r}_t t) \right|} = \exp(-\epsilon t) < 1.
 \end{equation}
 
-For exponential processes errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
+If one is to measure the ability to forecast the underlying infection dynamics, it may thus be more desirable to evaluate errors on the scale of the growth rate directly.
 
@@
 Specific suggestions for how to achieve this (either by taking a log transformation or by using another transformation to obtain growth rates directly as mentioned in the Discussion) are discussed in more detail in the remainder of the text.
-We added the word 'conceptually' to the sentence to make this a bit clearer.
 
@@
 
+\notindented
+
 The addition on ``tightly distributed'' probability mass reads
@@
 
-As (under the simplifying assumption of normality, see Section \ref{sec:methods:vst}) the WIS/CRPS of an ideal forecaster scales with the standard deviation, a value of $\beta = 1$ would imply a quadratic median-variance relationship;
+As (under the simplifying assumption of normality, see Section 2) the WIS/CRPS of an ideal forecaster scales with the standard deviation, a value of $\beta = 1$ would imply a quadratic median-variance relationship;
 
@@
 However, it may be an interesting approach if predictive samples are available or if quantiles for weekwise growth rates have been collected.
+Potentially, the variance stabilising time-series forecasting literature may be a useful source of other transformations for various forecast settings.
 
@@
 \black
-We added a statement to the legend "Locations are sorted according to the mean observed value in that location." We also followed your suggestion to make this a main figure.
+We added a statement to the legend "Locations are sorted according to the mean observed value in that location." We also followed your suggestion to make this a main figure and added some explanations in the text.
 
@@
 \black
-On the natural scale, the switch happens between 6.5 and 6.6 and on the log scale between 6.2 and 6.3.
+You're completely right, on the natural scale, the switch happens between 6.5 and 6.6 and on the log scale between 6.2 and 6.3.
 
@@
 \black
-We adapted the text and hopefully made it more clear:
-
-For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}A).
+You'r right, we adapted the text and hopefully made it more clear:
+
+\indented
+
+For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure 8A).
 While rankings between forecasters remain similar for a single forecast, this is not true anymore when looking at rankings obtained after averaging scores across multiple forecasts made at different times or in different locations.
@@
 when looking at performance during peaks or troughs.
-When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models therefore change considerably.
+When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models therefore change considerably (Figure 8B).
+
+\notindented
 
@@
 \black
-We corrected the typo
+Thank you, we corrected the typo.
 
@@
 \black
-We added a sentence:
+We think this is a valid point and added a sentence to that effect:
 

### 2023-06-26 main.tex
fix small latex issue

@@
 \toprule
-Horizon & Target & $\alpha$ & $\beta$ & $\alpha_\sqrt{\ }$ & $\beta_\sqrt{\ }$ & $\alpha_\log$ & $\beta_\log$ \\
+Horizon & Target & $\alpha$ & $\beta$ & $\alpha_{\sqrt{~}}$ & $\beta_{\sqrt{~}}$ & $\alpha_{\log}$ & $\beta_{\log}$ \\
 \midrule
@@
 $\text{WIS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},$ where $F_{\log}$ is the predictive distribution for $\log y$.
-The third one was $\text{WIS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{sqrt{\ }} + \beta_{sqrt{\ }} \cdot \sqrt{(\text{median}(F))},$ where $F_{\sqrt{\ }}$ is the predictive distribution for $\sqrt{y}$.
+The third one was $\text{WIS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{\sqrt{\ }} + \beta_{\sqrt{\ }} \cdot \sqrt{(\text{median}(F))},$ where $F_{\sqrt{\ }}$ is the predictive distribution for $\sqrt{y}$.
 }

### 2023-06-26 second submission after review/reply_to_reviews.tex
Prepare files for second submission

--- 2023-06-26 reply_to_reviews.tex
@@
 In the case of log transforms, taking the logarithm of the CRPS values, rather than scoring the log-transformed forecasts and data, results in an improper score.
-We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can be seen that in the example overconfident models perform best in terms of the log WIS.
+We illustrate this point using simulated data in Figure SI.1, where it can be seen that in the example overconfident models perform best in terms of the log WIS.
 We note that strictly speaking, re-scaling average scores by the average score of a baseline model or average scores across different models  to obtain skill scores likewise leads to improper scores \citep{gneitingStrictlyProperScoring2007}.
@@
     \centering
-    \includegraphics[width=0.49\textwidth]{output/figures/number-anomalies.png}
-     \includegraphics[width=0.49\textwidth]{output/figures/erroneous-forecasts.png}
+    \includegraphics[width=0.49\textwidth]{../output/figures/number-anomalies.png}
+     \includegraphics[width=0.49\textwidth]{../output/figures/erroneous-forecasts.png}
      \caption{}
@@
 \clearpage
-\bibliography{log-or-not-v2.bib, log-or-not.bib, software.bib}
+\bibliography{../bib/log-or-not-v2.bib, ../bib/log-or-not.bib, ../bib/software.bib}
 

### 2023-06-26 first submission/Manuscript_first_submission.tex
Prepare files for second submission

--- 2023-06-26 main.tex
@@
 
-\title{Scoring epidemiological forecasts on transformed scales}
+\title{Transformation of forecasts for evaluating predictive performance in an epidemiological context}
   \author{Nikos I.
@@
 
-Forecast evaluation is essential for the development of predictive epidemic models and can inform their use for public health decision-making.
-Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which can be seen as measures of the absolute distance between the forecast distribution and the observation.
-However, applying these scores directly to predicted and observed incidence counts may not be the most appropriate due to the exponential nature of epidemic processes and the varying magnitudes of observed values across space and time.
+Forecast evaluation plays an essential role in the development cycle of predictive epidemic models and can inform their use for public health decision-making.
+Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which are both measures of the absolute distance between the forecast distribution and the observation.
+They are commonly applied directly to predicted and observed incidence counts, but it can be questioned whether this is the optimal procedure for comparing models given the exponential nature of epidemic processes and the several orders of magnitude that observed values can span over space and time.
 In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results.
-Using the CRPS on log-transformed values as an example, we list three attractive properties: Firstly, it can be interpreted as a probabilistic version of a relative error.
+We motivate the procedure threefold using the CRPS on log-transformed counts as an example: Firstly, it can be interpreted as a probabilistic version of a relative error.
 Secondly, it reflects how well models predicted the time-varying epidemic growth rate.
 And lastly, using arguments on variance-stabilizing transformations, it can be shown that under the assumption of a quadratic mean-variance relationship, the logarithmic transformation leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity.
-Applying a transformation of log(x + 1) to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types.
-Situations in which models missed the beginning of upward swings are more strongly emphasised while failing to predict a downturn following a peak is less severely penalised when scoring transformed forecasts as opposed to untransformed ones.
+Applying the log transformation to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types.
+Situations in which models missed the beginning of upward swings are more strongly emphasized while failing to predict a downturn following a peak is less severely penalized.
 We conclude that appropriate transformations, of which the natural logarithm is only one particularly attractive option, should be considered when assessing the performance of different models in the context of infectious disease incidence.
@@
 Probabilistic forecasts \citep{heldProbabilisticForecastingInfectious2017} play an important role in decision-making in epidemiology and public health \citep{doi:10.2105/AJPH.2022.306831}, as well as other areas as diverse as economics \citep{timmermannForecastingMethodsFinance2018} or meteorology \citep{gneitingWeatherForecastingEnsemble2005}.
-Forecasts based on epidemiological modelling in particular have received widespread attention during the COVID-19 pandemic.
+Forecasts based on epidemiological modelling in particular has received widespread attention during the COVID-19 pandemic.
 Evaluations of forecasts can provide feedback for researchers to improve their models and train ensembles.
@@
 
-Probabilistic forecasts are usually evaluated using so-called (strictly) proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data.
-Proper scoring rules are constructed such that they encourage honest forecasting and cannot be `gamed' or `cheated'. 
-Assuming that the forecaster's actual best judgement corresponds to a predictive distribution $F$, a proper score is constructed such that if $F$ was the data-generating process, no other distribution $G$ would yield a better expected score.
-A scoring rule is called \textit{strictly} proper if there is no other distribution that under $F$ achieves the \textit{same} expected score as $F$, meaning that any deviation from $F$ leads to a worsening of expected scores.
-Forecasters (anyone or anything that issues a forecast) are thus incentivised to report their true belief $F$ about the future.
-Common proper scoring rules are the logarithmic or log score \citep{goodRationalDecisions1952} and the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}.
-The log score is the predictive log density or probability mass evaluated at the observed value.
-It is supported by the likelihood principle \citep{Winkler1996} and has many desirable theoretical properties; however, the particularly severe penalties it assigns to occasional misguided forecasts make it little robust \citep{bracherEvaluatingEpidemicForecasts2021}.
-Moreover, it is not easily applied to forecasts reported as samples or quantiles, as used in many recent disease forecasting efforts.
-It is nonetheless occasionally used in epidemiology (see e.g., \citealt{heldProbabilisticForecastingInfectious2017, Johansson2019}), but in recent years the CRPS and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021} have become increasingly popular.
-
+Probabilistic forecasts are usually evaluated using so-called proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data.
+Proper scoring rules are constructed such that forecasters (anyone or anything that issues a forecast) are incentivised to report their true belief about the future.
+Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the Weighted Interval Score \citep{bracherEvaluatingEpidemicForecasts2021}.
 The CRPS measures the distance of the predictive distribution to the observed data as 
 \begin{linenomath*}
-\begin{equation}
-    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - \boldsymbol{1}(x \geq y) \right)^2 dx,
-\end{equation}    
+\begin{equation*}
+    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,
+\end{equation*}    
 \end{linenomath*}
-where $y$ is the true observed value, $F$ is the cumulative distribution function (CDF) of the predictive distribution, and $\boldsymbol{1}()$ is the indicator function.
+where $y$ is the true observed value and $F$ the cumulative distribution function (CDF) of the predictive distribution.
 The CRPS can be understood as a generalisation of the absolute error to predictive distributions, and interpreted on the natural scale of the data.
-The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in the so-called COVID-19 Forecast Hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022} and Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US \textit{FluSight} project on influenza forecasting\citep{CdcepiFlusightforecastdata2022}.
+The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in so-called COVID-19 Forecast Hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}.
 The WIS is defined as
 \begin{linenomath*}
-\begin{equation}
+\begin{equation*}
     \text{WIS}(F, y) = \frac{1}{K} \times \sum_{k = 1}^{K} 2 \times \left[ \boldsymbol{1}(y \leq q_{\tau_k}) - \tau_k \right] \times ( q_{\tau_k} - y), 
-\end{equation}
+\end{equation*}
 \end{linenomath*}
-where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome and $K$ is the number of (roughly equally spaced) predictive quantiles provided.
-The WIS can be decomposed into three components, dispersion, underprediction and overprediction, which reflect the spread of the forecast and whether it was centred above or below the observed value.
-We show an alternative definition based on central prediction intervals in Supplement \ref{sec:alternative-wis} which illustrates this decomposition.
+where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, $K$ is the number of predictive quantiles provided and $\boldsymbol{1}$ is the indicator function.
+The WIS can be decomposed into three components, dispersion, overprediction, underprediction, which reflect the width of the forecast and whether it was centred above or below the observed value.
+We show an alternative definition based on central prediction intervals in Section \ref{sec:alternative-wis} which illustrates this decomposition.
 
-The notion of absolute distance encoded by the CRPS and WIS provides a straightforward interpretation, but may not always be the most useful perspective in the context of infectious disease spread.
-Especially in their early phase, outbreaks are best conceived as exponential processes, characterized by potentially time varying reproduction numbers $R_t$ \citep{gosticPracticalConsiderationsMeasuring2020} or epidemic growth rates $r_t$ \citep{dushoffSpeedStrengthEpidemic2021}.
-If the true modelling task revolves around estimating and forecasting these quantities, then evaluating forecasts based on the absolute distance between forecasted and observed incidence values penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount.
-For illustration, consider an incidence forecast issued at time 0 and referring to time $t$ that misses the correct average growth rate $\bar{r}_t$ by either $-\epsilon$ or $+\epsilon$.
-Then the ratio of the resulting absolute errors on the scale of observed incidences $y_{t}$ is
+The dynamics of infectious processes are often described by the complementary concepts of the reproduction number $R$ \citep{gosticPracticalConsiderationsMeasuring2020} and growth rate $r$ \citep{wallingaHowGenerationIntervals2007}, where $R$ describes the strength and $r$ the speed of epidemic growth \citep{dushoffSpeedStrengthEpidemic2021}.
+In the absence of changes in immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
+In that case, the number of new infections in the population grows exponentially in time.
+This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellisChallengesControlCOVID192021}.
 
-\begin{linenomath*}
-\begin{equation}
-\frac{\left|y_0 \exp[(\bar{r}_t - \epsilon) \times t] - y_0 \exp(\bar{r}_t t)\right|}{\left| y_0 \exp[(\bar{r}_t + \epsilon) \times t] - y_0 \exp(\bar{r}_t t) \right|} = \exp(-\epsilon t) < 1.
-\end{equation}
-\end{linenomath*}
+If case numbers are evolving based on an exponential process and the  modelling task revolves around estimating and forecasting the reproduction number or the corresponding growth rate, then evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount.
+This is because for exponential processes errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
+If one is to measure the ability of forecasts to assess and forecast the underlying infection dynamics, it may thus be more desirable to evaluate errors on the growth rate directly.
 
-% For exponential processes errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate. 
-If one is to measure the ability to forecast the underlying infection dynamics, it may thus be more desirable to evaluate errors on the scale of the growth rate directly.
-
-Another argument against using notions of absolute distance between predicted and observed incidence values is that forecast consumers may find errors on a relative scale easier to interpret and more useful in order to track predictive performance across targets of different orders of magnitude.
+Evaluating forecasts using the CRPS or WIS means that scores represent a measure of absolute errors.
+However, forecast consumers may find errors on a relative scale easier to interpret and more useful in order to track predictive performance across targets of different orders of magnitude.
 \cite{bolinLocalScaleInvariance2021} have proposed the scaled CRPS (SCRPS) which is locally scale invariant; however, it does not correspond to a relative error measure and lacks a straightforward interpretation as available for the CRPS.
 
-% A closely related aspect to relative scores (as opposed to absolute scores) is that in the evaluation one may wish to give similar weight to all 
-Lastly, it may be considered desirable to give all forecast targets similar weight in an overall performance evaluation.
+A closely related aspect to relative scores (as opposed to absolute scores) is that in the evaluation one may wish to give similar weight to all considered forecast targets.
 As the CRPS typically scales with the order of magnitude of the quantity to be predicted, this is not the case for the CRPS, which will typically assign higher scores to forecast targets with high expected values (e.g., in large locations or around the peak of an epidemic).
@@
 \cite{taylorEvaluatingVolatilityInterval1999} in finance, \cite{mayrLogLevelVAR2015} in macroeconomics, \cite{loweStochasticRainfallrunoffForecasting2014} in hydrology or \cite{fuglstadDoesNonstationarySpatial2015} in meteorology).
-While the goal of the transformations is often to improve the accuracy of the predictions, they can also be used to enhance and complement the evaluation process.
+While the goal of the transformations is usually to improve the accuracy of the predictions, they can also be used to enhance and complement the evaluation process.
 In this paper, we argue that the aforementioned issues with evaluating epidemic forecasts based on measures of absolute error on the natural scale can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation.
-Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while guaranteeing that the score remains proper.
-Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).
+Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score.
+Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects of the forecasts that forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).
 
-For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particularly attractive transformation in the context of epidemic phenomena.
-We refer to this transformation as 'log-transformation' and to scores that have been computed from log-transformed forecasts and observations as scores 'on the log scale' (as opposed to scores 'on the natural scale', which involve no transformation).
-In the theoretical discussion in Section \ref{sec:methods}, 'log-transformation' and 'log scale' generally refer to a transformation of $\log_{e}(x)$.
-For practical applications (Section \ref{sec:HUB}) we also use these terms to describe a transformation of $\log_{e}(x + a)$ with a small $a > 0$ in order to keep the terminology and notation simple.
-For a prediction target with strictly positive support, the CRPS after applying a log-transformation is given by
+For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation (referred to as the log-transformtation in the remainder of this manuscript) in the context of epidemic phenomena.
+The CRPS after applying a log-transformation can be computed as follows:
 %
 \begin{linenomath*}
-\begin{equation}
-    \text{CRPS}(F_{\log}, \log y) = \int_{-\infty}^\infty \left( F_{\log}(x) - \boldsymbol{1}(x \geq \log y) \right)^2 dx.
-\end{equation}    
+\begin{equation*}
+    \text{CRPS}(F_{\log}, \log y) = \int_{-\infty}^\infty \left( F_{\log}(x) - 1(x \geq \log y) \right)^2 dx,
+\end{equation*}    
 \end{linenomath*}
 %
-Here, $y$ is again the observed outcome and $F_{\log}$ is the predictive CDF of the log-transformed outcome, i.e.,
-\begin{linenomath*}
-\begin{equation}
-F_{\log}(x) = F(\exp(x)),
-\end{equation}    
-\end{linenomath*}
-with $F$ the CDF on the original scale.
+where $F_{\log}$ is the log-transformed predictive distribution.
 Instead of a score representing the magnitude of absolute errors, applying a log-transformation prior to the CRPS yields a score which a) measures relative error (see Section \ref{sec:methods:relative}), b) provides a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate}) and c) is less dependent on the expected order of magnitude of the quantity to be predicted (see Section \ref{sec:methods:vst}).
@@
 Our analysis mostly focuses on the CRPS (or WIS) as an evaluation metric for probabilistic forecasts, given its widespread use throughout the COVID-19 pandemic.
-We note that the logarithmic score has scale invariance properties which imply that score differences between different forecasts are invariant to strictly monotonic transformations (see \citealt{Lehmann1950} on corresponding properties of likelihood ratios and \citealt{diksLikelihoodbasedScoringRules2011}).
-The question of the right scale to evaluate forecasts on does therefore not arise for the log score.
 
@@
 In Sections \ref{sec:methods:relative}--\ref{sec:methods:vst} we provide some mathematical intuition on applying the log-transformation prior to evaluating the CRPS, highlighting the connections to relative error measures, the epidemic growth rate and variance stabilizing transformations.
-We then discuss the effect of the log-transformation on forecast rankings (Section \ref{sec:methods:rankings}) as well as practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}).
+We then discuss practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}) and the effect of the log-transformation on forecast rankings (Section \ref{sec:methods:rankings}).
 To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  \citep[ Section \ref{sec:HUB}]{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
@@
 \begin{linenomath*}
-\begin{equation}
+\begin{equation*}
 y = \hat{y} + \varepsilon,
-\end{equation}
+\end{equation*}
 \end{linenomath*}
@@
 \begin{linenomath*}
-\begin{equation}
+\begin{equation*}
 \log y = \log \hat{y} + \varepsilon^*,
-\end{equation}
+\end{equation*}
 \end{linenomath*}
@@
 Using that $\log(a) \approx a - 1$ if $a$ is close to 1, we get
-\begin{linenomath*}
-\begin{equation}
+$$
 |\varepsilon^*| = |\log \hat{y} - \log y| = \left|\log\left(\frac{\hat{y}}{y}\right) \right| \ \ \stackrel{\text{if } \hat{y} \ \approx \ y}{\approx} \ \ \left| \frac{\hat{y}}{y} - 1 \right| \ \ = \ \ \left| \frac{\hat{y} - y}{y} \right|.
-\end{equation}
-\end{linenomath*}
+$$
 The absolute error after log transforming is thus an approximation of the \textit{absolute percentage error} \citep[APE,][]{gneitingMakingEvaluatingPoint2011a} as long as forecast and observation are close.
-As we assumed that $\hat{y} \approx y$, we can also interpret it as an approximation of the \textit{relative error} \citep[RE,][]{gneitingMakingEvaluatingPoint2011a}
-\begin{linenomath*}
-\begin{equation}
+As we assumed that $\hat{y} \approx y$, we can also interpret it as an approximation of the \textit{relative error} (RE)
+$$
 \left| \frac{\hat{y} - y}{\hat{y}} \right|
-\end{equation}
-\end{linenomath*}
-and the \textit{symmetric absolute percentage error} (SAPE; see e.g., \citealt{Flores1986})
-\begin{linenomath*}
-\begin{equation}
+$$
+and the \textit{symmetric absolute percentage error} (SAPE)
+$$
 \left| \ \frac{\hat{y} - y}{y/2 + \hat{y}/2} \ \right|.
-\end{equation}
-\end{linenomath*}
+$$
 As Figure \ref{fig:SAPE} shows, the alignment with the SAPE is in fact the closest and holds quite well even if predicted and observed value differ by a factor of two or three.
@@
 \begin{linenomath*}
-\begin{equation}
+\begin{equation*}
 \frac{d}{dt}y(t) = r(t)y(t)
-\end{equation}
+\end{equation*}
 \end{linenomath*}
@@
 \begin{linenomath*}
-\begin{equation}
-y(t) = y_0 \exp \left( \int_0^t r(t') dt' \right) = y_0 \exp (\bar{r_t}t)
-\end{equation}
+\begin{equation*}
+y(t) = y_0 \exp \left( \int_0^t r(t') dt' \right) = y_0 \exp (\bar{r}t)
+\end{equation*}
 \end{linenomath*}
-where $y_0$ is an initial data point and $\bar{r_t}$ is the mean of the growth rate between the initial time point $0$ and time $t$.
+where $y_0$ is an initial data point and $\bar{r}$ is the mean of the growth rate between the initial time point $0$ and time $t$.
 
@@
 \begin{linenomath*}
-\begin{align}
-\begin{split}
+\begin{align*}
 \epsilon^* &= \left| \log \left[ \hat{y}( t ) \right] - \log \left[ y ( t ) \right] \right|\\
-           &= \left| \log \left[ y_0 \exp (\hat{\bar{r_t}} t ) \right] - \log \left[ y_0 \exp (\bar{r_t}t) \right] \right|\\
-           &= t \left| \hat{\bar{r_t}} - \bar{r_t} \right|
-\end{split}
-\end{align}
+           &= \left| \log \left[ y_0 \exp (\bar{\hat{r}} t ) \right] - \log \left[ y_0 \exp (\bar{r}t) \right] \right|\\
+           &= t \left| \bar{\hat{r}} - \bar{r} \right|
+\end{align*}
 \end{linenomath*}
 %
-where $\bar{r_t}$ is the true mean growth rate and $\hat{\bar{r_t}}$ is the forecast mean growth rate.
+where $\bar{\hat{r}}$ is the true mean growth rate and $\bar{r}$ is the forecast mean growth rate.
 We thus evaluate the error in the mean exponential growth rate, scaled by the length of the time period considered.
@@
 When evaluating models across sets of forecasting tasks, it may be desirable for each target to have a similar impact on the overall results.
-This could be motivated by the assumption that forecasts from different geographical units and time periods provide similar amounts of information about how well a forecaster performs.
-One would then like the resulting scores to be independent of the order of magnitude of the target to predict.
-CRPS values on the natural scale, however, typically scale with the order of magnitude of the quantity to be predicted.
-Average scores are then dominated by the results achieved for targets with high expected outcomes in a way that does not necessarily reflect the underlying predictive ability well.
+In disease incidence forecasting, this is not the case when using the CRPS on the natural scale, as the latter typically scales with the order of magnitude of the quantity to be predicted.
+Average scores are then dominated by the results achieved for targets with high expected outcomes.
 
-If the predictive distribution for the quantity $Y$ equals the true data-generating process $F$ (an ideal forecast), the expected CRPS is given by \citep{gneitingStrictlyProperScoring2007}
-\begin{linenomath*}
-\begin{equation}
+Specifically, if the predictive distribution for the quantity $Y$ equals the true data-generating process $F$ (an ideal forecast), the expected CRPS is given by \citep{gneitingStrictlyProperScoring2007}
+$$
 \mathbb{E}[\text{CRPS}(F, y)] = 0.5\times\mathbb{E}|Y - Y'|,
-\end{equation}
-\end{linenomath*}
+$$
 where $Y$ and $Y'$ are independent samples from $F$.
@@
 If $F$ is well-approximated by a normal distribution $\text{N}(\mu, \sigma^2)$, the approximation
-\begin{linenomath*}
-\begin{equation}
+$$
 \mathbb{E}_F[\text{CRPS}(F, y)] \approx \frac{\sigma}{\sqrt{\pi}}
-\end{equation}
-\end{linenomath*}
+$$
 can be used.
 This means that the expected CRPS scales roughly with the standard deviation, which in turn typically increases with the mean in epidemiological forecasting.
-In order to make the expected CRPS independent of the expected outcome, a \textit{variance-stabilising transformation} \citep[VST,][]{bartlettSquareRootTransformation1936, dunnGeneralizedLinearModels2018} can be employed.
+In order to make the expected CRPS independent of the expected outcome, a \textit{variance-stabilising transformation} \citep[VST,][]{bartlettSquareRootTransformation1936} can be employed.
 The choice of this transformation depends on the mean-variance relationship of the underlying process.
 
-If the mean-variance relationship of the data-generating distribution is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST.
-Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method (a first-order Taylor approximation, see e.g., \citealt{dunnGeneralizedLinearModels2018}), to show that
-
-\begin{linenomath*}
-\begin{equation}
+If the mean-variance relationship is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST \citep{guerreroTimeseriesAnalysisSupported1993}.
+Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method to show that
+$$
 \mathbb{E}_F[\text{CRPS}\{F_{\log}, \log(y)\}] \approx \frac{\sigma/\mu}{\sqrt{\pi}} 
 = \frac{\sqrt{c}}{\sqrt{\pi}}
-.\label{eq:taylor_quadratic}
-\end{equation}
-\end{linenomath*}
-As $\sigma$ and $\mu$ are linked through the quadratic mean-variance relationship (or linear mean-standard deviation relationship, $\sigma = \sqrt{c} \times \mu$), the expected CRPS thus stays constant regardless of the expected value of the data-generating distribution $\mu$.
+.
+$$
 The assumption of a quadratic mean-variance relationship is closely linked to the aspects discussed in Sections \ref{sec:methods:relative} and \ref{sec:methods:growthrate}.
 It implies that relative errors have constant variance and can thus be meaningfully compared across different targets.
-Also, it arises naturally if we assume that our capacity to predict the epidemic growth rate does not depend on the expected outcome, i.e.
-does not depend on the current phase of the epidemic or the order of magnitude of current observations.
+Also, it arises naturally if we assume that our capacity to predict the epidemic growth rate does not depend on the expected outcome.
 
-If the mean-variance relationship is linear with $\sigma^2 = c \times \mu$, as with a Poisson-distributed variable, the square root is known to be a VST \citep{dunnGeneralizedLinearModels2018}. 
+If the variance is linear with $\sigma^2 = c \times \mu$, as with a Poisson-distributed variable, the square root is known to be a VST. 
 Denoting by $F_{\sqrt{\ }}$ the predictive distribution for $\sqrt{Y}$, the delta method can again be used to show that
-\begin{linenomath*}
-\begin{equation}
+$$
 \mathbb{E}_F[\text{CRPS}\{F_{\sqrt{\ }}, \sqrt{y}\}] \approx \frac{\sigma/\sqrt{\mu}}{2\sqrt{\pi}} = \frac{\sqrt{c}}{2\sqrt{\pi}}
-.\label{eq:taylor_linear}
-\end{equation}
-\end{linenomath*}
-We note that while standard in the derivation of variance-stabilizing transformations, the application of the delta method in equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear} requires the probability mass of $F$ to be tightly distributed.
-If this is not the case, the approximation and thus the variance stabilization may be less accurate.
+.
+$$
 
 To strengthen our intuition on how transforming outcomes prior to applying the CRPS shifts the emphasis between targets with high and low expected outcomes, Figure \ref{fig:SIM-wis-state-size-mean} shows the expected CRPS of ideal forecasters under different mean-variance relationships and transformations.
-We consider a Poisson distribution where $\sigma^2 = \mu$, a negative binomial distribution with size parameter $\theta = 10$ and thus $\sigma^2 = \mu + \mu^2/10$, and a truncated normal distribution with practically constant variance.
-We see that when applying the CRPS on the natural scale, the expected CRPS grows monotonically as the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster) increases.
+We consider a Poisson distribution where $\sigma^2 = \mu$, a negative binomial distribution with size parameter $\theta = 10$ and thus $\sigma^2 = \mu + \mu^2/10$, and a normal distribution with constant variance.
+We see that when applying the CRPS on the natural scale, the expected CRPS grows with the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster).
 The expected CRPS is constant only for the distribution with constant variance, and grows in $\mu$ for the other two.
 When applying a log-transformation first, the expected CRPS is almost independent of $\mu$ for the negative binomial distribution and large $\mu$, while smaller targets have higher expected CRPS in case of the Poisson distribution and the normal distribution with constant variance.
-When applying a square-root-transformation, the expected CRPS is independent of the mean for the Poisson-distribution, but not for the other two (with a positive relationship in the normal case and a negative one for the negative binomial).
-As can be seen in Figures \ref{fig:SIM-wis-state-size-mean} and \ref{fig:score-approx}, the approximations presented in equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear} work quite well for our simulated example.
+When applying a square-root-transformation before the CRPS, the expected CRPS is independent of the mean for the Poisson-distribution, but not for the other two (with a positive relationship in the normal case and a negative one for the negative binomial).
+As can be seen in Figures \ref{fig:SIM-wis-state-size-mean} and \ref{fig:score-approx}, the approximations presented above work quite well for our simulated example.
 
@@
     \caption{Expected CRPS scores as a function of the mean and variance of the forecast quantity.
-We computed expected CRPS values  for three different distributions, assuming an ideal forecaster with predictive distribution equal to the true underlying (data-generating) distribution.
-    These expected CRPS values were computed for different predictive means based on 10,000 samples each and are represented by dots.
-Solid lines show the corresponding approximations of the expected CRPS from equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear}.
+We computed expected CRPS values  for three different distributions, assuming an ideal forecaster with predictive distribution equal to the data-generating distribution.
+    These expected CRPS values where computed for different predictive means based on 10,000 samples each and are represented by dots.
+Solid lines show the corresponding approximation of the expected CRPS based on an assumed normal distribution as discussed in section \ref{sec:methods:vst}.
 Figure \ref{fig:score-approx} shows the quality of the approximation in more detail.
@@
 Expected CRPS was computed on the natural scale (left), after applying a square-root transformation (middle), and after adding one and applying a log-transformation to the data (right).
-B: A but with x and y axes on the log scale.}
+B: A but with x axis on the log scale.}
     \label{fig:SIM-wis-state-size-mean}
 \end{figure}
+
+\subsection{Practical considerations}
+\label{sec:methods:considerations}
+
+Transformations that are strictly monotonic are permissible in the sense that they maintain the propriety of the score.
+This is because even though rankings of models may change forecasts will in expectation still minimise their score if they report a predictive distribution that is equal to the data-generating distribution.
+This condition holds for both the log and square root transformations, as well as many others.
+However, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee propriety.
+In the case of log transforms, taking the logarithm of the scores, rather than scoring the log-transformed forecasts and data, results in an improper score.
+This is because taking the logarithm of the CRPS (or WIS) results in a score that does not penalise outliers enough and therefore incentivises overconfident predictions.
+We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can easily be seen that overconfident models perform best in terms of the log WIS.
+
+In practice, one issue with the log transform is that they are not readily applicable to negative numbers or zero values, which need to be removed or otherwise handled. 
+One common approach to deal with zeros is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm \citep{bellegoDealingLogsZeros2022}.
+This represents a strictly monotonic transformation and therefore preserves the propriety of the resulting score.
+The choice of the quantity to add does however influences scores and rankings, as measures of relative errors shrink when adding a constant $a$ to the forecast and the observation.
+We illustrate this in Figure \ref{fig:illustration-effect-log-offset}.
+As a rule of thumb, if if $x > 5a$, the difference between $\log{(x + a)}$ and $\log{(x)}$ is small, and it becomes negligible if $x > 50a$.
+Choosing a suitable offset $a$ balances two competing concerns: on the one hand, choosing a small $a$ makes sure that the transformation is as close to a natural logarithm as possible and scores can be interpreted as outlined above.
+On the other hand, choosing a larger $a$ can help stabilise scores for forecasts and observations close to zero, avoiding giving excessive weight to forecasts for small quantities (see Figure \ref{fig:HUB-log-different-offsets}).
+
+A related issue occurs when the predictive distribution has a large probability mass on zero (or on very small values), as this can translate into an excessively wide forecast in relative terms.
+This can be seen in Figure \ref{fig:HUB-model-comparison-baseline}.
+Here, the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm because forecasts contained zero in its prediction intervals.
 
@@
 On the log scale, however, forecaster A receives a lower score for large observed values, being more heavily penalised for assigning large probability to small values (which, in relative terms, are far away from the actual observation).
-We note that the chosen example involving a geometric forecast distribution is somewhat constructed; as shown in Section \ref{sec:Hub:cor} and Figure \ref{fig:HUB-cors}A, rankings between models in practice stay quite stable for a single forecast.
 
@@
 
-Overall model rankings would be expected to differ more when scores are averaged across multiple forecasts or targets.
-The change in rankings of aggregate scores usually is mainly driven by the order of magnitude of scores for different forecast targets across time, location and target type and less so by the kind of changes in model rankings for single forecasts discussed above (see Figure \ref{fig:HUB-cors} for a practical example).
+Overall model rankings would be expected to differ even more when scores are averaged across multiple forecasts or targets.
+The change in rankings of aggregate scores is mainly driven by the order of magnitude of scores for different forecast targets across time, location and target type and less so by the kind of changes in model rankings for single forecasts discussed above.
 Large observations will dominate average CRPS values when evaluation is done on the natural scale, but much less so after log transformation.
-Depending on how different models perform across targets of different orders of magnitude, rankings in terms of average scores may change when applying a transformation.
-
-\subsection{Practical considerations and other transformations}
-\label{sec:methods:considerations}
-
-In practice, one issue with the log transform is that it is not readily applicable to negative or zero values, which need to be removed or otherwise handled. 
-One common approach to this end is to add a small positive quantity, such as $a = 1$, to all observations and predictions before taking the logarithm \citep{bellegoDealingLogsZeros2022}.
-This still represents a strictly monotonic transformation, but the choice of $a$ does influence scores and rankings (measures of relative errors shrink the larger the chosen value $a$).
-As a rule of thumb, if if $x > 5a$, the difference between $\log{(x + a)}$ and $\log{(x)}$ is small, and it becomes negligible if $x > 50a$.
-Choosing a suitable offset $a$ thus balances two competing concerns: on the one hand, choosing a small $a$ makes sure that the transformation is as close to a natural logarithm as possible and scores can be interpreted as outlined in the previous sections.
-On the other hand, choosing a larger $a$ can help stabilise scores for forecasts and observations close to zero, avoiding giving excessive weight to forecasts of small quantities.
-For increasing $a$, less relative weight is given to smaller forecast targets.
-For very large values of $a$, $\log(x + a)$ is roughly linear in $x$, so that using a very large $a$ implies similar relative weighting as applying no transformation at all.
-In practice, a user could explore the effect of different values of $a$ graphically and choose $a$ such that the relative weightings of times and regions with high and low incidence correspond to their preferences (see Figure \ref{fig:HUB-log-different-offsets} in our example application, Section \ref{sec:HUB}).
-
-A related issue occurs when the predictive distribution has a large probability mass on zero (or on very small values), as this can translate into an excessively wide forecast in relative terms.
-In our applied example this is illustrated in Figure \ref{fig:HUB-model-comparison-baseline}.
-In such cases, the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm because forecasts contained zero in its prediction intervals.
-To deal with this issue one could choose to use a higher $a$ value when applying a transformation $\log(x + a)$, for example $a = 10$ instead of the $a = 1$ that we chose to use.
-
-A natural question is which other transformations could be applied and whether resulting scores remain (strictly) proper.
-In principle, any transformation function can be applied simultaneously to forecasts and observations as long as the definition of the transformation is independent of the forecasts and any quantities unknown at the time of forecasting, including the observed value.
-This simply corresponds to a re-definition of the forecasting target.
-However, applying non-invertible transformations leads to a loss in information conveyed by forecasts, which we consider undesirable.
-The resulting score will be proper, but it may not be strictly proper anymore (as forecasts differing from the forecaster's true belief on the original scale may be identical on the transformed scale).
-When using the CRPS or the WIS, it seems most appropriate to use only strictly monotonic transformations such as the natural logarithm or the square root as otherwise the encoded notion of distance may become meaningless.
-
-Some other strictly monotonic transformations that can be applied are scaling by the population size or scaling by past observations.
-The latter, as discussed in Section \ref{sec:discussion}, is similar to applying a log-transformation, but corresponds to evaluating a forecast of multiplicative, rather than exponential growth rates.
-The arising issue of dividing by zero can again be solved by adding a small offset $a$.
-Scaling a forecast by the later observed value (as opposed to scaling by past observations) is generally not permissible as it can result in improper scores (see \citealt{lerchForecasterDilemmaExtreme2015} on the closely related topic of weighting scores with a function of the observed value).
-Similarly, scaling forecasts and observations by a function of the predictive distribution (like the predictive mean) may lead to improper scores; however, we are unaware of existing theoretical arguments on this.
-
-When applying a transformation, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee that the score remains proper.
-In the case of log transforms, taking the logarithm of the CRPS values, rather than scoring the log-transformed forecasts and data, results in an improper score.
-We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can be seen that in the example overconfident models perform best in terms of the log WIS.
-We note that strictly speaking, re-scaling average scores by the average score of a baseline model or average scores across different models  to obtain skill scores likewise leads to improper scores \citep{gneitingStrictlyProperScoring2007}.
-The application of such skill scores, however, is established practice and considered largely unproblematic.
-
-We note that in the practical evaluation of operational forecasting systems  several additional challenges arise, which we do not study in detail.
-These concern e.g., the removal of outlying observations and forecasts and the handling of missing forecasts.
-The solutions we employed in practice are provided in Section \ref{sec:HUB-setting}.
+Depending on the relationship between the mean and variance of the forecast target, a log-transformation may even lead to systematically larger scores assigned to small forecast targets, as illustrated in Figure \ref{fig:SIM-wis-state-size-mean}.
 
@@
 The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 5th of December 2022 for reported cases and deaths from COVID-19.
-Target dates range from the 13th of March 2021 to the 10th of December 2022, for a total of 92 weeks.
 See \cite{sherrattPredictivePerformanceMultimodel2022} for a more thorough description of the data.
-We filtered all forecasts submitted to the Hub to only include the seven models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 46 forecast dates (see Figure \ref{fig:HUB-num-avail-models}).
+We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 46 forecast dates (see Figure \ref{fig:HUB-num-avail-models}).
 We removed all observations marked as data anomalies by the European Forecast Hub \citep{sherrattPredictivePerformanceMultimodel2022} as well as all remaining negative observed values.
-These anomalies made up a relevant fraction of all observations.
-On average across locations, 12.1 out of 92 (13.2\%) observations were removed for cases and 12.4 out of 92 (13.5\%) for deaths.
-Figure \ref{fig:number-anomalies} displays the number of anomalies removed for each location.
-In addition, we filtered out a small number of erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table \ref{tab:erroneous}.
-Figure \ref{fig:erroneous-forecasts} shows the percentage of forecasts removed for each model.
-Those few (less than 0.2\%  of forecasts for each model) erroneous outlier forecasts had excessive influence on average scores and relative skill scores in a way that was not representative of normal model behaviour.
-We removed them here in order to better illustrate the effects of the log-transformation on scores that one would expect in a well-behaved scenario.
-In a regular forecast evaluation such erroneous forecasts should usually not be removed and would count towards overall model scores.
-
+In addition, we filtered out erroneous forecasts defined by any of the conditions listed in Table \ref{tab:erroneous}.
+Those forecasts were removed in order to be better able to illustrate the effects of the log-transformation on scores and eliminating distortions caused by outlier forecasters.
 All predictive quantiles were truncated at 0.
@@
 The relative skill for a given model compared to others is then obtained by taking the geometric mean of all score ratios which involve that model.
-Low values are better, and the ``average'' model receives a relative skill score of 1.
+Low values are better, and the "average" model receives a relative skill score of 1.
 
@@
 Missing the peak, i.e.
-predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts during the peak in May 2022 in Figure \ref{fig:HUB-model-comparison-ensemble}A, B).
+predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts in May in Figure \ref{fig:HUB-model-comparison-ensemble}A, B).
 On the log scale, these have less of an influence, as errors are smaller in relative terms (see \ref{fig:HUB-model-comparison-ensemble}C, D).
-Conversely, failure to predict an upswing while numbers are still low, is less severely penalised on the natural scale (see forecasts in July 2021 and to a lesser extent in July 2022 in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low.
-On the log scale, missing lower inflection points tends to lead to more severe penalties (see Figure \ref{fig:HUB-model-comparison-ensemble}C, D).
+Conversely, failure to predict an upswing while numbers are still low, is less severely punished on the natural scale (see forecasts in July in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low.
+On the log scale, missing lower inflection points tends to lead to more severe penalties (see Figure \ref{fig:HUB-model-comparison-ensemble}C, D)).
 One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure \ref{fig:HUB-model-comparison-ensemble}E, F).
@@
 B: Corresponding boxplot (y-axis on log-scale) of all cases and deaths.
-C: Scores for two-week-ahead forecasts from the EuroCOVIDhub-ensemble (averaged across all forecast dates) for different locations, evaluated on the natural scale as well as after transforming counts by adding one and applying the natural logarithm.
-D: Corresponding boxplots of all individual scores of the EuroCOVIDhub-ensemble for two-week-ahead predictions.
+C: Scores for two-week-ahead forecasts from the EuroCOVIDhub-ensemble (averaged across all forecast dates) for different locations, evaluated on the natural as well as the logarithmic scale.
+D: Corresponding boxplots of all individual scores for two-week-ahead predictions.
 E: Boxplots for the relative change of scores for the EuroCOVIDhub-ensemble across forecast horizons.
@@
 All scores were divided by the score for forecast horizon one.
-To enhance interpretability, the range of visible relative changes in scores (relative to horizon = 1) was restricted to [0.1, 10].}
+To enhance interpretability, the range of visible relative changes was restricted to [0.1, 10].}
     \label{fig:HUB-mean-locations}
@@
 
-To assess the impact of the choice of offset value $a$ we extend the display from Figure \ref{fig:HUB-mean-locations}C by results obtained under different specifications.
-Results are shown in Figure \ref{fig:HUB-log-different-offsets}, where for completeness we also added the square root transformation.
-As discussed in Section \ref{sec:methods:considerations}, smaller values of $a$ increase the relative weight of smaller locations in the overall evaluation.
-In the most extreme considered case $a = 0.001$, the smallest locations in fact receive the largest weight both for deaths and cases.
-For very large values (see the third row of Figure \ref{fig:HUB-log-different-offsets}), the relative weights strongly resemble those of the evaluation on the natural scale.
-We recommend using displays of this type to get an intuition for the role different locations may play for overall evaluation results.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-scores-locations-log-variants.png}
-    \caption{Mean WIS in different locations for different transformations applied before scoring.
-Locations are sorted according to the mean observed value in that location.
-Shown are scores for two-week-ahead forecasts of the EuroCOVIDhub-ensemble.
-On the natural scale (with no transformation prior to applying the WIS), scores correlate strongly with the average number of observed values in a given location.
-The same is true for scores obtained after applying a square-root transformation, or after applying a log-transformation with a large offset $a$.
-For illustrative purposes, $a$ was chosen to be 101630 for cases and 530 for deaths, 10 times the respective median observed value.
-For large values of $a$, $\log(x + a)$ grows roughly linearly in $x$, meaning that we expect to observe the same patterns as in the case with no transformation.
-For decreasing values of $a$, we give more relative weight to scores in small locations.}
-    \label{fig:HUB-log-different-offsets}
-\end{figure}
-
 \subsection{Regression analysis to determine the variance-stabilizing transformation}
@@
 We ran the regression
-\begin{linenomath*}
-\begin{equation}
+\begin{equation*}
     \log[\text{WIS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], 
-\end{equation}
-\end{linenomath*}
+\end{equation*}
 where the predictive distribution $F$ and the observation $y$ are on the natural scale. This is equivalent to
-\begin{linenomath*}
-\begin{equation}
+\begin{equation*}
     \text{WIS}(F, y) = \exp(\alpha) \times \text{median}(F)^\beta, 
-\end{equation}
-\end{linenomath*}
+\end{equation*}
 meaning that we estimate a polynomial relationship between the predictive median and achieved WIS.
 Note that we are using predictive medians rather than means as only the former are available in the European COVID-19 Forecast Hub.
-As (under the simplifying assumption of normality; see Section \ref{sec:methods:vst}) the WIS/CRPS of an ideal forecaster scales with the standard deviation, a value of $\beta = 1$ would imply a quadratic median-variance relationship; the natural logarithm could then serve as a VST.
+As the WIS/CRPS of an ideal forecaster scales with the standard deviation (see Section \ref{sec:methods:vst}), a value of $\beta = 1$ would imply a quadratic median-variance relationship; the natural logarithm could then serve as a VST.
 A value of $\beta = 0.5$ would imply a linear median-variance relationship, suggesting the square root as a VST.
-We applied the regression to case and death forecasts, stratified for one through four-week-ahead forecasts.
+We applied the regression to case and death forecasts, pooled across horizons and stratified for one through four-week-ahead forecasts.
 Results are provided in Table \ref{tab:HUB-regression}.
@@
 \toprule
-Horizon & Target & $\alpha$ & $\beta$ & $\alpha_\sqrt{\ }$ & $\beta_\sqrt{\ }$ & $\alpha_\log$ & $\beta_\log$ \\
+Horizon & Target & $\alpha$ & $\beta$ & $\alpha_{\sqrt{~}}$ & $\beta_{\sqrt{~}}$ & $\alpha_{\log}$ & $\beta_{\log}$ \\
 \midrule
-% all & all & -1.093 & 0.963 & -0.352 & 0.201 & 0.391 & 0.001\\
-% \addlinespace
-% all & Cases & 0.036 & 0.858 & 0.043 & 0.201 & 0.751 & -0.033\\
-% all & Deaths & -0.884 & 0.868 & 0.273 & 0.121 & 0.436 & -0.023\\
-% \addlinespace
-% 1 & all & -1.402 & 0.923 & 0.320 & 0.088 & 0.318 & -0.014\\
-% 2 & all & -1.221 & 0.967 & 0.112 & 0.164 & 0.364 & -0.003\\
-% 3 & all & -1.001 & 0.984 & -0.094 & 0.241 & 0.410 & 0.008\\
-% 4 & all & -0.757 & 0.986 & 0.000 & 0.299 & 0.469 & 0.015\\
+all & all & -1.093 & 0.963 & -0.352 & 0.201 & 0.391 & 0.001\\
+\addlinespace
+all & Cases & 0.036 & 0.858 & 0.043 & 0.201 & 0.751 & -0.033\\
+all & Deaths & -0.884 & 0.868 & 0.273 & 0.121 & 0.436 & -0.023\\
+\addlinespace
+1 & all & -1.402 & 0.923 & 0.320 & 0.088 & 0.318 & -0.014\\
+2 & all & -1.221 & 0.967 & 0.112 & 0.164 & 0.364 & -0.003\\
+3 & all & -1.001 & 0.984 & -0.094 & 0.241 & 0.410 & 0.008\\
+4 & all & -0.757 & 0.986 & 0.000 & 0.299 & 0.469 & 0.015\\
+\addlinespace
 1 & Cases & -0.862 & 0.876 & 0.790 & 0.087 & 0.433 & -0.024\\
@@
 To check the relationship after the transformation, we ran the regressions
-\begin{equation}
+\begin{equation*}
     \text{WIS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},
-\end{equation}
+\end{equation*}
 where $F_{\log}$ is the predictive distribution for $\log(y)$, and
-\begin{equation}
+\begin{equation*}
     \text{WIS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{\sqrt{\ }} + \beta_{\sqrt{\ }} \cdot \sqrt{\text{median}(F)},
-\end{equation} 
+\end{equation*} 
 where $F_{\sqrt{\ }}$ is the predictive distribution on the square-root scale.
-A value of $\beta_{\log} = 0$ (or $\beta_{\sqrt{\ }} = 0$, respectively) would imply that scores are linearly independent of the median prediction after the transformation.
+A value of $\beta_{\log} = 0$ (or $\beta_{\sqrt{\ }} = 0$, respectively, would imply that scores are independent of the median prediction after the transformation.
 A value smaller (larger) than 0 would imply that smaller (larger) targets lead to higher scores.
@@
 \subsection{Impact of logarithmic transformation on model rankings}
-\label{sec:Hub:cor}
 
@@
     \caption{Correlations of rankings on the natural and logarithmic scale.
-A: Average Spearman rank correlation of scores for individual forecasts.
+A: Average Spearman rank correlation of scores for individual forecasts from all models.
 For every individual target (defined by a combination of forecast date, target type, horizon, location), one score was obtained per model.
-Then, for every forecast target, the Spearman rank correlation was computed between scores on the natural scale and on the log scale for all the models that had made a forecast for that specific target.
-These individual rank correlations were then averaged across locations and time and are displayed stratified by horizon and target types, representing average accordance of model ranks for a single forecast target on the natural and on the log scale.
+Then, the rank correlation was computed between the scores for all models on the natural scale vs.
+on the log scale.
+All rank correlations were averaged across locations, and target types and stratified by horizon and target type.
 B: Correlation between relative skill scores.
-For every forecast horizon and target type, a separate relative skill score was computed per model using pairwise comparisons, which is a measure of performance of a model relative to the others for a given horizon and target type that accounts for missing values.
+For every forecast horizon and target type, a separate relative skill score was computed per model using pairwise comparisons.
 The plot shows the correlation between the relative skill scores on the natural vs.
-on the log scale, representing accordance of overall model performance as judged by scores on the natural and on the log scale.}
+on the log scale.}
     \label{fig:HUB-cors}
@@
 For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}A).
-While rankings between forecasters remain similar for a single forecast, this is not true anymore when looking at rankings obtained after averaging scores across multiple forecasts made at different times or in different locations.
-As discussed earlier, scores on the natural and on the log scale penalise errors very differently, e.g.
-when looking at performance during peaks or troughs.
-When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models therefore change considerably (Figure \ref{fig:HUB-cors}B).
+When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models change considerably (Figure \ref{fig:HUB-cors}B).
 The correlation between relative skill scores also decreases noticeably with increasing forecast horizon.
@@
 B: Change in rankings as determined by relative skill scores when moving from an evaluation on the natural scale to one on the logarithmic scale.
-Red arrows indicate that the relative skill score deteriorated when moving from the natural to the log scale, green arrows indicate they improved.
 C: Relative skill scores based on scores on the log scale.
@@
 
-Figure \ref{fig:HUB-rank-order} shows the changes in the ranking between different forecasting models.
+Figure Figure \ref{fig:HUB-rank-order} shows the changes in the ranking between different forecasting models.
 Encouragingly for the European Forecast Hub, the Hub ensemble, which is the forecast the organisers suggest forecast consumers make use of, remains the top model across scoring schemes.
@@
 A potential downside is that forecast evaluation is unreliable in situations where observed values are zero or very small.
-One could argue that this correctly reflect inherent uncertainty about the future course of an epidemic when numbers are small.
-Users nevertheless need to be aware that this can pose issues in practice.
 Including very small values in prediction intervals (see e.g.
@@
 high scores) when evaluating forecasts on the log scale.
-\cite{bracherEvaluatingEpidemicForecasts2021} argue that it is desirable to give large weight to forecasts for locations with high incidences, as this reflects performance on the targets we should care about most.
+\cite{bracherEvaluatingEpidemicForecasts2021} argue that the large weight given to forecasts for locations with high incidences is a desirable property, as it reflects performance on the targets we should care about most.
 On the other hand, scoring forecasts on the log scale may be less influenced by outliers and better reflect consistent performance across time, space, and forecast targets.
@@
 One obvious option is to apply a population standardization to obtain incidence forecasts e.g., per 100,000 population \citep{abbottEvaluatingEpidemiologicallyMotivated2022}.
-We suggested using the natural logarithm as a variance-stabilising transformation (VST).
-This is appropriate for variables that are approximately normally distributed and have a quadratic mean-variance relationship with $\sigma^2 = c \times \mu^2$ (this is e.g.
-approximately true for the negative binomoial distribution and large $\mu$).
-Alternatively, the square-root transformation can be appropriate in the case of a Poisson distributed variable \citep{dunnGeneralizedLinearModels2018}.
+If one is interested in multiplicative, rather than exponential growth rates, one could convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last observed value.
+We suggested using the natural logarithm as a variance-stabilising transformation (VST) or alternatively the square-root transformation in the case of a Poisson distributed variable.
 Other VST like the Box-Cox \citep{boxAnalysisTransformations1964} are conceivable as well.
-If one is interested in multiplicative, rather than exponential growth rates, one could, instead of applying a log transformation, convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last value that was observed at the time the forecast was made.
-Forecasters would then implicitly predict a separate multiplicative growth rate from today to horizon 1, 2, etc.
-Instead of dividing by the last observed value, another promising transformation would be to divide each forecast by the forecast of the previous week (and analogously for observations), in order to obtain forecasts for week-to-week growth rates.
-Alternatively, one could also take first differences of values on the log scale.
-This approach would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluating the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}.
-Dividing values by the previous value, unfortunately, is not feasible under the current quantile-based format of the Forecast Hubs, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate.
+Another promising transformation would be to take differences between forecasts on the log scale, or alternatively to divide each forecast by the forecast of the previous week (and analogously for observations), in order to obtain forecasts for week-to-week growth rates.
+One could then also ask forecasters to provide estimates of the weekly relative change applied to the latest data and subsequent forecast points directly.
+This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluating the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}.
+This, unfortunately, is not feasible under the current quantile-based format of the Forecast Hubs, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate.
 However, it may be an interesting approach if predictive samples are available or if quantiles for weekwise growth rates have been collected.
-Potentially, the variance stabilising time-series forecasting literature may be a useful source of other transformations for various forecast settings.
+It is possible to go beyond choosing a single transformation by constructing composite scores as a weighted sum of scores based on different transformations.
+This would make it possible to create custom scores and allow forecast consumers to assign explicit weights to different qualities of the forecasts they might care about.
 
-It is possible to go beyond choosing a single transformation by constructing composite scores as a weighted sum of scores based on different transformations.
-This would make it possible to create custom scores and allow forecast consumers to choose and assign explicit weights to different qualities of the forecasts they might care about.
+In this work, we focused on the CRPS and WIS, which are widely used in the evaluation of epidemic forecasts.
+We note that for the logarithmic score, which has also been used e.g., in some editions of the FluSight challenge \cite{reichCollaborativeMultiyearMultimodel2019}, the question of the right scale to evaluate forecasts does not arise.
+It is known that log score differences between different forecasters are invariant to monotonic transformations of the outcome variable (see e.g., \citealt{diksLikelihoodbasedScoringRules2011}).
+This is clearly an advantage of the logarithmic score over the CRPS; however, the logarithmic score is known to have other severe downsides, e.g., its low robustness to sporadically misguided forecasts; see \cite{bracherEvaluatingEpidemicForecasts2021} for a more detailed discussion.
 
 Exploring transformations is a promising avenue for future work that could help bridge the gap between modellers and policymakers by providing scoring rules that better reflect what forecast consumers care about.
-{In this paper, we did not make any particular assumptions about policy makers' priorities and preferences.
-Rather, we aimed to enable users to make an informed choice by showing how different transformations lead to different relative weights for the kinds of prediction errors forecast consumers may care about, such as absolute vs.
-relative errors or the size of penalties for over- vs.
-underprediction.
-In practice, engagement with decision makers is important to determine what their priorities are and how different ways to measure predictive importance should be weighed.
-
+Potentially, the variance stabilising time-series forecasting literature may be a useful source of transformations for various forecast settings.
 We have shown that the natural logarithm transformation can lead to significant changes in the relative rankings of models against each other, with potentially important implications for decision-makers who rely on the knowledge of past performance to make a judgement about which forecasts should inform future decisions.
@@
 Instead of defining the WIS as an average of scores for individual quantiles, we can define it using an average of scores for symmetric predictive intervals.
-For a single prediction interval, the interval score (IS) is computed as the sum of three penalty components, dispersion (width of the prediction interval), underprediction and overprediction,
+For a single prediction interval, the interval scoren (IS) is computed as the sum of three penalty components, dispersion (width of the prediction interval), underprediction and overprediction,
 %
 \begin{linenomath*}
-\begin{align}
- IS_\alpha(F,y) &= (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot \boldsymbol{1}(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot \boldsymbol{1}(y \geq u) \\
+\begin{align*}
+ IS_\alpha(F,y) &= (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u) \\
  &= \text{dispersion} + \text{underprediction} + \text{overprediction},    
-\end{align}
+\end{align*}
 \end{linenomath*}
 %
-where $\boldsymbol{1}()$ is the indicator function, $y$ is the observed value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution, i.e.
+where $1()$ is the indicator function, $y$ is the observed value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution, i.e.
 the lower and upper bound of a single central prediction interval.
@@
 \begin{linenomath*}
-\begin{equation}
+\begin{equation*}
 \text{WIS} = \frac{1}{K^* + 0.5} \cdot \left(w_0 \cdot |y - m| + \sum_{k = 1}^{K^*} w_k \cdot IS_{\alpha_{k}}(F, y)\right),    
-\end{equation} 
+\end{equation*} 
 \end{linenomath*}
@@
     \includegraphics[width=0.99\textwidth]{output/figures/example-log-first.png}
-    \caption{Illustration of the effect of applying a transformation after scoring.
+    \caption{Illustration of impropriety of log-transformed CRPS.
 We assume $Y \sim \text{LogNormal}(0, 1)$ and evaluate the expected CRPS for predictive distributions $\text{LogNormal}(0, \sigma)$ with varying values of $\sigma \in [0.1, 2]$.
@@
 For the log-transformed CRPS, the optimal value is  0.9, i.e.
-there is an incentive to report a forecast that is too sharp.
-The score is therefore no longer proper.}
+there is an incentive to report a forecast that is too sharp.}
     \label{fig:log-improper}
@@
 For increasing x, all lines eventually approach the black line (representing a transformation with no offset applied).
-    For a given solid line, the dashed line of the same colour marks the x-value that  is equal to 5 times the corresponding offset.
-It can be seen that for $a$ values smaller than one fifth of the transformed quantity, the effect of adding an offset is generally small.
-When choosing a suitable $a$, the trade-off is between staying close to the interpretation of a pure log-transformation (choosing a small $a$) and not giving excessive weights to small observations (by choosing a larger $a$, see Figure \ref{fig:HUB-log-different-offsets}).}
+    For a given solid line, the dashed line of the same colour marks the x-value that  is equal to 5 times the corresponding offset.}
     \label{fig:illustration-effect-log-offset}
@@
 \toprule
-target type & quantity & measure & natural & log\\
+target\_type & quantity & measure & natural & log\\
 \midrule
@@
     \label{fig:HUB-num-avail-models}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/number-anomalies.png}
-    \caption{
-    Number of observed values that were removed as anomalous. The values were marked as anomalous by the European Forecast Hub team. 
-    }
-    \label{fig:number-anomalies}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/erroneous-forecasts.png}
-    \caption{
-    Number of forecasts marked as erroneous and removed.
-Forecasts that were in extremely poor agreement with the observed values were removed from the analysis according to the criteria shown in Table \ref{tab:erroneous}.
-    }
-    \label{fig:erroneous-forecasts}
 \end{figure}
@@
 The model had zero included in some of its 50 percent intervals (e.g.
-for case forecasts in July 2021), leading to excessive dispersion values on the log scale.
+for case forecasts in July), leading to excessive dispersion values on the log scale.
 One could argue that including zero in the prediction intervals constituted an unreasonable forecast that was rightly penalised, but in general care has to be taken with small numbers.
-One potential way to do deal with this could be to use a higher $a$ value when applying a transformation $\log(x + a)$, for example $a = 10$ instead of $a = 1$.
 A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale.
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.99\textwidth]{output/figures/HUB-scores-locations-log-variants.png}
+    \caption{Mean WIS in different locations for different transformations applied before scoring.
+Shown are scores for two-week-ahead forecasts of the EuroCOVIDhub-ensemble.
+On the natural scale (with no transformation prior applying the WIS), scores correlate strongly with the average number of observed values in a given location.
+The same is true for scores obtained after applying a square-root transformation, or after applying a log-transformation with a large offset $a$.
+For illustrative purposes, $a$ was chosen to be 101630 for cases and 530 for deaths, 10 times the respective median observed value.
+For large values of $a$, $log(x + a)$ grows linearly in $x$, meaning that we expect to observe the same patterns as in the case with no transformation.
+For decreasing values of $a$, we give more relative weight to scores in small locations.}
+    \label{fig:HUB-log-different-offsets}
+\end{figure}
+
 \clearpage
-\bibliography{log-or-not-v2.bib, log-or-not.bib, software.bib}
+\bibliography{bib/log-or-not-v2.bib, bib/log-or-not.bib, bib/software.bib}
 

### 2023-06-28 second submission after review/reply_to_reviews.tex
Last edits to reviewer reply

@@
 Please see our responses to the individual comments as well as explanations on the changes we have made to the manuscript below.
-In addition to the changes outlined below we some general edits to the paper.
-We changed the title to make it more concise, we altered the order of some sections and shortened the introduction a bit, which became quite lengthy after incorporating all the suggestions.
+In addition to the changes outlined below we made some general edits to the paper.
+We changed the title to make it more concise, we altered the order of some sections and shortened the introduction a bit, which had after the revision had become somewhat long.
 
@@
 \black
-In the theoretical part of the manuscript we simplify the equations a little bit by looking at the case of a single point forecast, rather than a probabilistic forecast that comes in the form of a predictive distribution.
+In the theoretical part of the manuscript we simplify the equations by looking at the case of a single point forecast, rather than a probabilistic forecast that comes in the form of a predictive distribution.
 We feel this makes the discussion more intuitive.
-We want to emphasise that the log-linear error is really the error that is measured by the score, not the error model that would be assumed by any forecaster making a prediction.
+We want to emphasise that the log-linear error $\varepsilon^*$ is really the error that is measured by the score, not the error model that would be assumed by any forecaster making a prediction.
 
@@
 Thank you for raising this interesting point.
-The log score is indeed an alternative which is scale invariant under monotonic transformations.
-We mention this in the introduction and state that for the log score the question of which scale to evaluate forecasts on does not arise.
+The log score is indeed an alternative which is scale-invariant under strictly monotonic transformations.
+We now mention this in the introduction and state that for the log score the question of which scale to evaluate forecasts on does not arise.
 In our understanding, a quasi-likelihood approach (as opposed to an actual likelihood approach as in the log score) may not actually be necessary as at least in theory full predictive distributions are to be evaluated.
-Developing further ideas based on some form of quasi-likelihood would likely require more in-depth investigation.
-In any case, developing these ideas would require more in-depth investigation and we feel this is beyond the scope of the current paper.
+Developing further ideas based on some form of quasi-likelihood would likely require more in-depth investigation, which we feel is beyond the scope of the current paper.
 
@@
 It is important to note that the equations in 2.2 do not refer to any particular model used by a forecaster or any model that was fit to data.
-Rather, we define the instantaneous growth rate for a single realised epidemic and then show that scoring forecasts for that on the log scale equates to scoring a forecast of the average growth rate according to this definition.
+Rather, we define the instantaneous growth rate for a single realised epidemic and then show that scoring forecasts for that on the log scale equates to scoring a forecast of the average growth rate according to this definition (this is essentially the same argument as in the first comment of Section 1.1.2 above).
 The forecaster themselves who would make a forecast could assume any error model or model structure they like.
@@
 The solutions we employed in practice are provided in Section 3.1.
+
 \notindented
@@
 Not introducing another term for transformations of $\log_{e}(x + \varepsilon)$ makes the manuscript more readable and accessible to readers.
-In addition, our analyses suggest that adding a small offset doesn't make much of a difference in most practical applications.
+In addition, our analyses suggest that adding a small offset does not make much of a difference in most practical applications.
 We nevertheless completely agree that this should be made clear to the reader and have added the following text:
@@
 
+For a specific motivation in a simple branching process model please see our reply to comment ``L166'' below.
+
 \blue
@@
 CRPS values on the natural scale, however, typically scale with the order of magnitude of the quantity to be predicted.
-Average scores are then dominated by the results achieved for targets with high expected outcomes in a way that does not necessarily reflect the underlying predictive ability well.
 
@@
     $$
+    $R_t$ thus fluctuates around an average value of $R_0$.
     \item This implies
     $$
-    Y_t \mid Y_{t - 1} \sim \text{NegBin}(R_t \times Y_{t - 1}, \psi).
+    Y_t \mid Y_{t - 1} \sim \text{NegBin}(\text{mean} = R_t \times Y_{t - 1}, \text{size} = \psi).
     $$
@@
 We agree that this is not a typical example we would expect to see in an epidemiological setting.
-It could happen in principle, but only in rather constructed cases.
 We still believe the example is valuable to illustrate the general point that the rankings between two forecasts can change even for a single observation.
@@
 $w = \alpha / 2$ (which is generally standard), then the WIS converges exactly to the CRPS for an increasing number of equally spaced prediction intervals.
-For an illustration see Figure 2 in \cite{bracherEvaluatingEpidemicForecasts2021}.
+For an illustration see Figure 2 in the referenced paper by Bracher, Ray, Gneiting and Reich (2021).
 
@@
 In the theoretical discussion in Section 2 we defined prediction targets to have strictly positive support, so these cases are excluded.
-In the later parts we follow the common strategy to add an offset before taking the logarithm.
+In the later parts we follow the common strategy to add an offset $a$ before taking the logarithm.
 
@@
 \black
-You'r right, we adapted the text and hopefully made it more clear:
+You are right, we adapted the text and hopefully made it more clear:
 

### 2023-07-08 second submission after review/Manuscript-without-SI.tex
add tiff Figures for publication

--- 2023-06-26 main.tex
@@
 \centering
-\includegraphics[width = 1\textwidth]{output/figures/different-relative-errors.png}
+\includegraphics[width = 1\textwidth]{../output/figures/different-relative-errors.png}
 \caption{Numerical comparison of different measures of relative error: absolute percentage error (APE), relative error (RE), symmetric absolute percentage error (SAPE) and the absolute error applied to log-transformed predictions and observations.
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/SIM-mean-state-size.png}
+    \includegraphics[width=0.99\textwidth]{../output/figures/SIM-mean-state-size.png}
     \caption{Expected CRPS scores as a function of the mean and variance of the forecast quantity.
@@
 \centering
-\includegraphics[width = 1\textwidth]{output/figures/illustration-effect-log-ranking-crps.png}
+\includegraphics[width = 1\textwidth]{../output/figures/illustration-effect-log-ranking-crps.png}
 \caption{Illustration of the effect of the log-transformation of the ranking for a single forecast.
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-model-comparison-ensemble.png}
+    \includegraphics[width=0.99\textwidth]{../output/figures/HUB-model-comparison-ensemble.png}
     \caption{
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-mean-obs-location.png}
+    \includegraphics[width=0.99\textwidth]{../output/figures/HUB-mean-obs-location.png}
     \caption{Observations and scores across locations and forecast horizons for the European COVID-19 Forecast Hub data. Locations are sorted according to the mean observed value in that location. 
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-scores-locations-log-variants.png}
+    \includegraphics[width=0.99\textwidth]{../output/figures/HUB-scores-locations-log-variants.png}
     \caption{Mean WIS in different locations for different transformations applied before scoring.
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-transformation-regression.png}
+    \includegraphics[width=0.99\textwidth]{../output/figures/HUB-transformation-regression.png}
     \caption{Relationship between median forecasts and scores.
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-correlations.png}
+    \includegraphics[width=0.99\textwidth]{../output/figures/HUB-correlations.png}
     \caption{Correlations of rankings on the natural and logarithmic scale.
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-pairwise-comparisons.png}
+    \includegraphics[width=0.99\textwidth]{../output/figures/HUB-pairwise-comparisons.png}
     \caption{Changes in model ratings as measured by relative skill for two-week-ahead predictions for cases (top row) and deaths (bottom row).
@@
 
-\subsection{Alternative Formulation of the WIS}
+\paragraph{S1 Text.}
 \label{sec:alternative-wis}
+\textbf{Alternative Formulation of the WIS.}
 
-Instead of defining the WIS as an average of scores for individual quantiles, we can define it using an average of scores for symmetric predictive intervals.
-For a single prediction interval, the interval score (IS) is computed as the sum of three penalty components, dispersion (width of the prediction interval), underprediction and overprediction,
-%
-\begin{linenomath*}
-\begin{align}
- IS_\alpha(F,y) &= (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot \boldsymbol{1}(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot \boldsymbol{1}(y \geq u) \\
- &= \text{dispersion} + \text{underprediction} + \text{overprediction},    
-\end{align}
-\end{linenomath*}
-%
-where $\boldsymbol{1}()$ is the indicator function, $y$ is the observed value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution, i.e.
-the lower and upper bound of a single central prediction interval.
-For a set of $K^*$ prediction intervals and the median $m$, the WIS is computed as a weighted sum,
-\begin{linenomath*}
-\begin{equation}
-\text{WIS} = \frac{1}{K^* + 0.5} \cdot \left(w_0 \cdot |y - m| + \sum_{k = 1}^{K^*} w_k \cdot IS_{\alpha_{k}}(F, y)\right),    
-\end{equation} 
-\end{linenomath*}
-where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/example-log-first.png}
-    \caption{Illustration of the effect of applying a transformation after scoring.
+\paragraph{S1 Figure.}
+\label{fig:log-improper}
+\textbf{Illustration of the effect of applying a transformation after scoring. }
 We assume $Y \sim \text{LogNormal}(0, 1)$ and evaluate the expected CRPS for predictive distributions $\text{LogNormal}(0, \sigma)$ with varying values of $\sigma \in [0.1, 2]$.
@@
 there is an incentive to report a forecast that is too sharp.
-The score is therefore no longer proper.}
-    \label{fig:log-improper}
-\end{figure}
+The score is therefore no longer proper.
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/illustration-effect-offset-log.png}
-    \caption{Illustration of the effect of adding a small quantity to a value before taking the natural logarithm.
+\paragraph{S2 Figure.}
+\label{fig:illustration-effect-log-offset}
+\textbf{Illustration of the effect of adding a small quantity to a value before taking the natural logarithm. }
 For increasing x, all lines eventually approach the black line (representing a transformation with no offset applied).
@@
 It can be seen that for $a$ values smaller than one fifth of the transformed quantity, the effect of adding an offset is generally small.
-When choosing a suitable $a$, the trade-off is between staying close to the interpretation of a pure log-transformation (choosing a small $a$) and not giving excessive weights to small observations (by choosing a larger $a$, see Figure \ref{fig:HUB-log-different-offsets}).}
-    \label{fig:illustration-effect-log-offset}
-\end{figure}
+When choosing a suitable $a$, the trade-off is between staying close to the interpretation of a pure log-transformation (choosing a small $a$) and not giving excessive weights to small observations (by choosing a larger $a$, see Figure \ref{fig:HUB-log-different-offsets}).
 
-\begin{figure}[h!]
-\centering
-\includegraphics[width = 1\textwidth]{output/figures/SIM-score-approximation.png}
-\caption{
-Visualisation of expected CRPS values against approximated scores using the approximation detailed in Section \ref{sec:methods:rankings} (see also Figure \ref{fig:SIM-wis-state-size-mean}).
+\paragraph{S3 Figure.}
+\label{fig:score-approx}
+\textbf{Visualisation of expected CRPS values against approximated scores.} 
+This is using the approximation detailed in Section \ref{sec:methods:rankings} (see also Figure \ref{fig:SIM-wis-state-size-mean}).
 Expected CRPS scores are shown for three different distributions once on the natural scale (top row) and once scored on the log scale (bottom row).
-}
-\label{fig:score-approx}
-\end{figure}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% HUB plots %
+\paragraph{S1 Table.}
+\label{tab:HUB-summary}
+\textbf{Summary statistics for observations and scores for forecasts from the ECDC data set.}
 
-\begin{table}[h!]
-    \centering
-\begin{tabular}{lllcc}
-\toprule
-target type & quantity & measure & natural & log\\
-\midrule
-Cases & Observations & mean & 61979 & 9.19\\
-Cases & Observations & sd & 171916 & 2.10\\
-Cases & Observations & var & 29555122130 & 4.42\\
-Deaths & Observations & mean & 220 & 3.89\\
-Deaths & Observations & sd & 435 & 1.96\\
-\addlinespace
-Deaths & Observations & var & 189051 & 3.83\\
-Cases & WIS & mean & 15840 & 0.27\\
-Cases & WIS & sd & 53117 & 0.28\\
-Deaths & WIS & mean & 31 & 0.23\\
-Deaths & WIS & sd & 65 & 0.28\\
-\bottomrule
-\end{tabular}
-    \caption{Summary statistics for observations and scores for forecasts from the ECDC data set.}
-    \label{tab:HUB-summary}
-\end{table}
+\paragraph{S4 Figure.}
+\label{fig:HUB-num-avail-models}
+\textbf{Number of forecasts available from different models for each forecast date.}
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/number-avail-forecasts.png}
-    \caption{
-    Number of forecasts available from different models for each forecast date. 
-    }
-    \label{fig:HUB-num-avail-models}
-\end{figure}
+\paragraph{S5 Figure.}
+\label{fig:number-anomalies}
+\textbf{Number of observed values that were removed as anomalous. }
+The values were marked as anomalous by the European Forecast Hub team.
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/number-anomalies.png}
-    \caption{
-    Number of observed values that were removed as anomalous. The values were marked as anomalous by the European Forecast Hub team. 
-    }
-    \label{fig:number-anomalies}
-\end{figure}
+\paragraph{S6 Figure.}
+\label{fig:erroneous-forecasts}
+\textbf{Number of forecasts marked as erroneous and removed. } 
+Forecasts that were in extremely poor agreement with the observed values were removed from the analysis according to the criteria shown in Table \ref{tab:erroneous}.
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/erroneous-forecasts.png}
-    \caption{
-    Number of forecasts marked as erroneous and removed.
-Forecasts that were in extremely poor agreement with the observed values were removed from the analysis according to the criteria shown in Table \ref{tab:erroneous}.
-    }
-    \label{fig:erroneous-forecasts}
-\end{figure}
-
-\begin{table}
-\centering
-\begin{tabular}{ccc}
-\toprule
-True value & \& & Median prediction\\
-\midrule
-$>0$ & \ & $>100\times$ true value\\
-$>10$ & \  & $>20\times$ true value\\
-$>50$ & \  & $<1/50\times$ true value\\
-$= 0$ & \  & $>100$\\
-\bottomrule
-\end{tabular}
-\caption{Criteria for removing forecasts.
+\paragraph{S2 Table.}
+\label{tab:erroneous}
+\textbf{Criteria for removing forecasts. }
 Any forecast that met one of the listed criteria (represented by a row in the table), was removed.
@@
 When evaluating models against each other (rather than illustrating the effect of a transformation), one would prefer not to condition on the outcome when deciding whether a forecast should be taken into account.
-}
-\label{tab:erroneous}
-\end{table}
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-model-comparison-baseline.png}
-    \caption{
-    Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-baseline made in Germany.
+\paragraph{S7 Figure.}
+\label{fig:HUB-model-comparison-baseline}
+\textbf{Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-baseline made in Germany. }
 The model had zero included in some of its 50 percent intervals (e.g.
@@
 D, H: Corresponding scores.
-    }
-    \label{fig:HUB-model-comparison-baseline}
-\end{figure}
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-model-comparison-epinow.png}
-    \caption{
-    Forecasts and scores for two-week-ahead predictions from the epiforecasts-EpiNow2 model \citep{epinow2} made in Germany.
-A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale.
+\paragraph{S8 Figure.}
+\label{fig:HUB-model-comparison-epinow}
+\textbf{Forecasts and scores for two-week-ahead predictions from the epiforecasts-EpiNow2 model made in Germany.} 
+A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale from the EpiNow2 model \citep{epinow2}.
 B, F: Corresponding scores.
@@
 D, H: Corresponding scores.
-    }
-    \label{fig:HUB-model-comparison-epinow}
-\end{figure}
 
 \clearpage
-\bibliography{log-or-not-v2.bib, log-or-not.bib, software.bib}
+\bibliography{../bib/log-or-not-v2.bib, ../bib/log-or-not.bib, ../bib/software.bib}
 

### 2023-07-08 R/00 full paper code.R

@@ -48,6 +48,20 @@ plot_fct <- function(data) {
          x = "Observed value")
 }
 
+label_fn <- function(x) {
+  x <- ifelse(x%%1 == 0, 
+              as.integer(x), x)
+  ifelse(x < 1000, 
+         paste(x), 
+         ifelse(x < 1e6, 
+                paste0(x / 1000, "k"),
+                ifelse(x < 1e9, 
+                       paste0(x / 1e6, "m"), 
+                       paste0(x / 1e9, "b"))
+         )
+  )
+}
+
 label_fn_hat <- function(x) {
   ifelse(x%%1 == 0, 
          TeX(paste0(x, "\\hat{y}")), 
@@ -71,10 +85,11 @@ p2 <- p1 +
   theme(legend.position = "bottom")
 
 ggsave("output/figures/different-relative-errors.png", width = 7, height = 3)
+ggsave("second submission after review/figures/Figure1.tiff", width = 7, height = 3, dpi = 300)
 
 
 ## ========================================================================== ##
-## SI Version of Figure 1 (Figure SI.1) 
+## Not used: 
 ## Numerical comparison square-root transformation
 ## ========================================================================== ##
 
@@ -193,10 +208,11 @@ p1 / p2 +
   labs(y = TeX("CRPS (normalised)"))
 
 ggsave("output/figures/SIM-mean-state-size.png", width = 7, height = 4.1)
+ggsave("second submission after review/figures/Figure2.tiff", width = 7, height = 4.1, dpi = 300)
 
 
 ## ========================================================================== ##
-## SI Version of Figure 2 (Figure SI.2)
+## SI Version of Figure 2 (Figure SI.3)
 ## Illustration of the approximation
 ## ========================================================================== ##
 
@@ -222,11 +238,11 @@ scores_fig_2 |>
   expand_limits(y = 0)
 
 ggsave("output/figures/SIM-score-approximation.png", width = 7, height = 8)
-
+ggsave("second submission after review/SI/S3-Figure.tiff", width = 7, height = 4.1, dpi = 300)
 
 
 ## ========================================================================== ##
-## FIGURE 3: llustration of impropriety of log-transformed CRPS. 
+## FIGURE SI.1: llustration of impropriety of log-transformed CRPS. 
 ## ========================================================================== ##
 
 # in this plot the CRPS is approximated by the WIS with 99 quantiles
[…]

### 2023-07-10 second submission after review/Manuscript-without-SI.tex
Add author summary

@@
 \newpage
+
+\section*{Author summary}
+Scores like the Continuous Ranked Probability Score (CRPS) or the Weighted Interval Score (WIS) are commonly used to evaluate epidemiological forecasts and are a measure of absolute distance between forecast and observation.
+Due to the exponential nature of epidemic processes, evaluating the absolute distance between forecast and observation may not be ideal.
+We argue that transforming counts before applying the CRPS or WIS can yield more meaningful results.
+The natural logarithm is a particularly attractive transformation in epidemiological settings.
+Scores computed on log-transformed values can be interpreted as a probabilistic version of a relative error and reflect how well forecasters predict the time-varying epidemic growth rate.
+If the data-generating process has a quadratic mean-variance relationship, the logarithmic transformation also leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity.
+We illustrate these properties using data from the European COVID-19 Forecast Hub and find that scoring transformed counts changes model rankings.
+Stronger emphasis is given to  situations in which forecasters missed the beginning of upward swings, while failing to predict a downturn following a peak is less severely penalised.
+We generally recommend including evaluations of transformed counts when assessing forecaster performance.
 

### 2023-08-04 final submission after reformatting/SI/S1-Text.tex
Adapt formatting for final submission

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\author{}
\date{\vspace{-2.5em}}

\begin{document}

Instead of defining the WIS as an average of scores for individual quantiles, we can define it using an average of scores for symmetric predictive intervals. For a single prediction interval, the interval score (IS) is computed as the sum of three penalty components, dispersion (width of the prediction interval), underprediction and overprediction,  
%
\begin{align}
 IS_\alpha(F,y) &= (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot \boldsymbol{1}(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot \boldsymbol{1}(y \geq u) \\
 &= \text{dispersion} + \text{underprediction} + \text{overprediction},    
\end{align}
%
where $\boldsymbol{1}()$ is the indicator function, $y$ is the observed value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution, i.e. the lower and upper bound of a single central prediction interval. For a set of $K^*$ prediction intervals and the median $m$, the WIS is computed as a weighted sum, 
\begin{equation}
\text{WIS} = \frac{1}{K^* + 0.5} \cdot \left(w_0 \cdot |y - m| + \sum_{k = 1}^{K^*} w_k \cdot IS_{\alpha_{k}}(F, y)\right),    
\end{equation} 
where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$.

\end{document}

### 2023-08-04 final submission after reformatting/final-manuscript.tex
Adapt formatting for final submission

--- 2023-07-10 second submission after review/Manuscript-without-SI.tex
@@
-\documentclass{article}
-\usepackage[]{graphicx}
-\usepackage[]{xcolor}
-\usepackage{alltt}
-\usepackage[left=2.3cm,right=2.8cm, top = 2.2cm, bottom = 3cm]{geometry}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{natbib}
-\PassOptionsToPackage{hyphens}{url}
-\usepackage{url} 
-\usepackage[disable]{todonotes}
-\usepackage{multicol}
-\usepackage{rotating}
+% Template for PLoS
+% Version 3.6 Aug 2022
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% This template contains comments intended 
+% to minimize problems and delays during our production 
+% process. Please follow the template instructions
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Once your paper is accepted for publication, 
+% PLEASE REMOVE ALL TRACKED CHANGES in this file 
+% and leave only the final text of your manuscript. 
+% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
+% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
+%
+%
+% There are no restrictions on package use within the LaTeX files except that no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% For figure citations, please use "Fig" instead of "Figure".
+% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - spacing/line breaks within cells to alter layout or alignment
+% - do not nest tabular environments (no tabular environments within tabular environments)
+% - no graphics or colored text (cell background color/shading OK)
+% See http://journals.plos.org/plosone/s/tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
+%
+% For inline equations, please be sure to include all portions of an equation in the math environment.
+For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
+%
+% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
+%
+% Please add line breaks to long display equations when possible in order to fit size of the column. 
+%
+% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
+%
+% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
+%
+% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
+%
+% % % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
+\usepackage{amsmath,amssymb}
+
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
+
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
+\usepackage[nopatch=eqnum]{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% color can be used to apply background shading to table cells only
+\usepackage[table]{xcolor}
+
+% array package and thick rules for tables
+\usepackage{array}
+
+% create "+" rule type for thick vertical lines
+\newcolumntype{+}{!{\vrule width 2pt}}
+
+% create \thickcline for thick horizontal lines of variable length
+\newlength\savedwidth
+\newcommand\thickcline[1]{%
+  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+  \cline{#1}%
+  \noalign{\vskip\arrayrulewidth}%
+  \noalign{\global\arrayrulewidth\savedwidth}%
+}
+
+% \thickhline command for thick horizontal lines that span the table
+\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+\hline
+\noalign{\global\arrayrulewidth\savedwidth}}
+
+% Remove comment for double spacing
+%\usepackage{setspace} 
+%\doublespacing
+
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in 
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+\renewcommand{\figurename}{Fig}
+
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2015}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+%\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+%\setlength{\headheight}{27.023pt}
+%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\today}
+
+%% Include all macros below
+
+\newcommand{\lorem}{{\bf LOREM}}
+\newcommand{\ipsum}{{\bf IPSUM}}
+
+%% END MACROS SECTION
+
+%% Custom packages
 \usepackage{booktabs}
-\usepackage[colorlinks=false]{hyperref} 
-\setlength{\parskip}{\baselineskip}%
-\setlength{\parindent}{0pt}%
-\urlstyle{same}
-\usepackage{lineno}
-\linenumbers
-\bibliographystyle{apalike}
+% \usepackage[]{xcolor}
+% \usepackage{alltt}
+% \usepackage{amsmath}
+% \usepackage{amssymb}
+% \PassOptionsToPackage{hyphens}{url}
+% \usepackage{url} 
+% \usepackage[disable]{todonotes}
+% \usepackage{multicol}
+% \usepackage{rotating}
+
+% \setlength{\parskip}{\baselineskip}%
+% \setlength{\parindent}{0pt}%
+% \urlstyle{same}
+% \usepackage{lineno}
+% \linenumbers
+
+% \bibliographystyle{apalike}
 %\bibliographystyle{vancouver}
@@
 \begin{document}
+\vspace*{0.2in}
 
-\title{Scoring epidemiological forecasts on transformed scales}
-  \author{Nikos I.
-Bosse\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom} $^{,}$\thanks{Centre for the Mathematical Modelling of Infectious Diseases, London, United Kingdom} $^{,}$\thanks{NIHR Health Protection Research Unit in Modelling \& Health Economics} $^{,*}$,
-  Sam Abbott\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$, 
-  Anne Cori\thanks{MRC Centre for Outbreak Analysis and Modelling, Department of Infectious Disease Epidemiology, School of Public Health, Imperial College London, London, United Kingdom} $^{}$, \\
-  Edwin van Leeuwen\footnotemark[1] $^{,}$\thanks{Modelling \& Economics Unit and NIHR Health Protection Research Unit in Modelling \& Health Economics, UK Health Security Agency, London, United Kingdom} $^{}$,
-  Johannes Bracher\thanks{Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology, Karlsruhe, Germany } $^{,}$\thanks{Computational Statistics Group, Heidelberg Institute for Theoretical Studies, Heidelberg, Germany } $^{, \dagger}$,
-  Sebastian Funk\footnotemark[1] $^{, }$\footnotemark[2] $^{, }$\footnotemark[3] $ ^{, \dagger}$}
+% Title must be 250 characters or less.
+\begin{flushleft}
+{\Large
+\textbf\newline{Scoring epidemiological forecasts on transformed scales} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+}
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+Nikos I. Bosse\textsuperscript{1, 2, 3, *},
+Sam Abbott\textsuperscript{1, 2},
+Anne Cori\textsuperscript{4},
+Edwin van Leeuwen\textsuperscript{1, 3, 5},
+Johannes Bracher\textsuperscript{6, 7, \Yinyang},
+Sebastian Funk\textsuperscript{1, 2, 3, \Yinyang}
+\\
+\bigskip
+\textbf{1} Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom
+\\
+\textbf{2} Centre for the Mathematical Modelling of Infectious Diseases, London, United Kingdom
+\\
+\textbf{3} NIHR Health Protection Research Unit in Modelling \& Health Economics
+\\
+\textbf{4} MRC Centre for Outbreak Analysis and Modelling, Department of Infectious Disease Epidemiology, School of Public Health, Imperial College London, London, United Kingdom
+\\
+\textbf{5} Modelling \& Economics Unit and NIHR Health Protection Research Unit in Modelling \& Health Economics, UK Health Security Agency, London, United Kingdom
+\\
+\textbf{6} Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology, Karlsruhe, Germany
+\\
+\textbf{7} Computational Statistics Group, Heidelberg Institute for Theoretical Studies, Heidelberg, Germany
+\\
+\bigskip
 
-\maketitle
-\begin{abstract}
+% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
+% 
+% Remove or comment out the author notes below if they aren't used.
+%
+% Primary Equal Contribution Note
+\Yinyang These authors contributed equally to this work.
 
+% Use the asterisk to denote corresponding authorship and provide email address in note below.
+* nikos.bosse@lshtm.ac.uk
+
+\end{flushleft}
+% Please keep the abstract below 300 words
+\section*{Abstract}
 Forecast evaluation is essential for the development of predictive epidemic models and can inform their use for public health decision-making.
@@
 We conclude that appropriate transformations, of which the natural logarithm is only one particularly attractive option, should be considered when assessing the performance of different models in the context of infectious disease incidence.
-\end{abstract}
 
-\bigskip
-
-{\footnotesize $^*$ Correspondence to \url{nikos.bosse@lshtm.ac.uk}}, 
-{\footnotesize $^\dagger$ Contributed equally}
-
-\newpage
-
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step. 
+% Author Summary not valid for PLOS ONE submissions.   
 \section*{Author summary}
@@
 
-% ===========================================================
-\section{Introduction}
+\linenumbers
 
-Probabilistic forecasts \citep{heldProbabilisticForecastingInfectious2017} play an important role in decision-making in epidemiology and public health \citep{doi:10.2105/AJPH.2022.306831}, as well as other areas as diverse as economics \citep{timmermannForecastingMethodsFinance2018} or meteorology \citep{gneitingWeatherForecastingEnsemble2005}.
+\section*{Introduction}
+
+Probabilistic forecasts \cite{heldProbabilisticForecastingInfectious2017} play an important role in decision-making in epidemiology and public health \cite{doi:10.2105/AJPH.2022.306831}, as well as other areas as diverse as economics \cite{timmermannForecastingMethodsFinance2018} or meteorology \cite{gneitingWeatherForecastingEnsemble2005}.
 Forecasts based on epidemiological modelling in particular have received widespread attention during the COVID-19 pandemic.
@@
 
-Probabilistic forecasts are usually evaluated using so-called (strictly) proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data.
+Probabilistic forecasts are usually evaluated using so-called (strictly) proper scoring rules \cite{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data.
 Proper scoring rules are constructed such that they encourage honest forecasting and cannot be `gamed' or `cheated'. 
@@
 Forecasters (anyone or anything that issues a forecast) are thus incentivised to report their true belief $F$ about the future.
-Common proper scoring rules are the logarithmic or log score \citep{goodRationalDecisions1952} and the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}.
+Common proper scoring rules are the logarithmic or log score \cite{goodRationalDecisions1952} and the continuous ranked probability score (CRPS,~\cite{gneitingStrictlyProperScoring2007}).
 The log score is the predictive log density or probability mass evaluated at the observed value.
-It is supported by the likelihood principle \citep{Winkler1996} and has many desirable theoretical properties; however, the particularly severe penalties it assigns to occasional misguided forecasts make it little robust \citep{bracherEvaluatingEpidemicForecasts2021}.
+It is supported by the likelihood principle \cite{winklerScoringRulesEvaluation1996} and has many desirable theoretical properties; however, the particularly severe penalties it assigns to occasional misguided forecasts make it little robust \cite{bracherEvaluatingEpidemicForecasts2021}.
 Moreover, it is not easily applied to forecasts reported as samples or quantiles, as used in many recent disease forecasting efforts.
-It is nonetheless occasionally used in epidemiology (see e.g., \citealt{heldProbabilisticForecastingInfectious2017, Johansson2019}), but in recent years the CRPS and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021} have become increasingly popular.
+It is nonetheless occasionally used in epidemiology (see e.g., \cite{heldProbabilisticForecastingInfectious2017, Johansson2019}), but in recent years the CRPS and the weighted interval score (WIS,~\cite{bracherEvaluatingEpidemicForecasts2021}) have become increasingly popular.
 
@@
 \end{linenomath*}
+
 where $y$ is the true observed value, $F$ is the cumulative distribution function (CDF) of the predictive distribution, and $\boldsymbol{1}()$ is the indicator function.
 The CRPS can be understood as a generalisation of the absolute error to predictive distributions, and interpreted on the natural scale of the data.
-The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in the so-called COVID-19 Forecast Hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022} and Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US \textit{FluSight} project on influenza forecasting\citep{CdcepiFlusightforecastdata2022}.
+The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in the so-called COVID-19 Forecast Hubs in the US \cite{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \cite{sherrattPredictivePerformanceMultimodel2022} and Germany and Poland \cite{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US \textit{FluSight} project on influenza forecasting\cite{CdcepiFlusightforecastdata2022}.
 The WIS is defined as
@@
 The notion of absolute distance encoded by the CRPS and WIS provides a straightforward interpretation, but may not always be the most useful perspective in the context of infectious disease spread.
-Especially in their early phase, outbreaks are best conceived as exponential processes, characterized by potentially time varying reproduction numbers $R_t$ \citep{gosticPracticalConsiderationsMeasuring2020} or epidemic growth rates $r_t$ \citep{dushoffSpeedStrengthEpidemic2021}.
+Especially in their early phase, outbreaks are best conceived as exponential processes, characterized by potentially time varying reproduction numbers $R_t$ \cite{gosticPracticalConsiderationsMeasuring2020} or epidemic growth rates $r_t$ \cite{dushoffSpeedStrengthEpidemic2021}.
 If the true modelling task revolves around estimating and forecasting these quantities, then evaluating forecasts based on the absolute distance between forecasted and observed incidence values penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount.
@@
 Our analysis mostly focuses on the CRPS (or WIS) as an evaluation metric for probabilistic forecasts, given its widespread use throughout the COVID-19 pandemic.
-We note that the logarithmic score has scale invariance properties which imply that score differences between different forecasts are invariant to strictly monotonic transformations (see \citealt{Lehmann1950} on corresponding properties of likelihood ratios and \citealt{diksLikelihoodbasedScoringRules2011}).
+We note that the logarithmic score has scale invariance properties which imply that score differences between different forecasts are invariant to strictly monotonic transformations (see \cite{Lehmann1950} on corresponding properties of likelihood ratios and \cite{diksLikelihoodbasedScoringRules2011}).
 The question of the right scale to evaluate forecasts on does therefore not arise for the log score.
@@
 We then discuss the effect of the log-transformation on forecast rankings (Section \ref{sec:methods:rankings}) as well as practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}).
-To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  \citep[ Section \ref{sec:HUB}]{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
+To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  (\cite{europeancovid-19forecasthubEuropeanCovid19Forecast2021,sherrattPredictivePerformanceMultimodel2022}, Section \ref{sec:HUB}).
 Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues (Section \ref{sec:discussion}).
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Logarithmic transformation of forecasts and observations}
+\section*{Logarithmic transformation of forecasts and observations}
 \label{sec:methods}
 
-\subsection{Interpretation as a relative error}
+\subsection*{Interpretation as a relative error}
 \label{sec:methods:relative}
@@
 \end{linenomath*}
-The absolute error after log transforming is thus an approximation of the \textit{absolute percentage error} \citep[APE,][]{gneitingMakingEvaluatingPoint2011a} as long as forecast and observation are close.
-As we assumed that $\hat{y} \approx y$, we can also interpret it as an approximation of the \textit{relative error} \citep[RE,][]{gneitingMakingEvaluatingPoint2011a}
+The absolute error after log transforming is thus an approximation of the \textit{absolute percentage error} (APE,~ \cite{gneitingMakingEvaluatingPoint2011a}) as long as forecast and observation are close.
+As we assumed that $\hat{y} \approx y$, we can also interpret it as an approximation of the \textit{relative error} (RE,~\cite{gneitingMakingEvaluatingPoint2011a})
 \begin{linenomath*}
@@
 \end{linenomath*}
-and the \textit{symmetric absolute percentage error} (SAPE; see e.g., \citealt{Flores1986})
+and the \textit{symmetric absolute percentage error} (SAPE; see e.g., \cite{Flores1986})
 \begin{linenomath*}
@@
 \end{linenomath*}
-As Figure \ref{fig:SAPE} shows, the alignment with the SAPE is in fact the closest and holds quite well even if predicted and observed value differ by a factor of two or three.
+As Fig \ref{fig:SAPE} shows, the alignment with the SAPE is in fact the closest and holds quite well even if predicted and observed value differ by a factor of two or three.
 Generalising to probabilistic forecasts, the CRPS applied to log-transformed forecasts and outcomes can thus be seen as a probabilistic counterpart to the symmetric absolute percentage error, which offers an appealing intuitive interpretation.
@@
 
-\subsection{Interpretation as scoring the exponential growth rate}
+\subsection*{Interpretation as scoring the exponential growth rate}
 \label{sec:methods:growthrate}
 
-Another interpretation for the log-transform is possible if the generative process is framed as exponential with a time-varying growth rate $r(t)$~\citep[see, e.g.,][]{wallingaHowGenerationIntervals2007}, i.e.
+Another interpretation for the log-transform is possible if the generative process is framed as exponential with a time-varying growth rate $r(t)$ (see e.g.
+\cite{wallingaHowGenerationIntervals2007}), i.e.
 \begin{linenomath*}
@@
 
-\subsection{Interpretation as a variance-stabilising transformation}
+\subsection*{Interpretation as a variance-stabilising transformation}
 \label{sec:methods:vst}
@@
 
-If the predictive distribution for the quantity $Y$ equals the true data-generating process $F$ (an ideal forecast), the expected CRPS is given by \citep{gneitingStrictlyProperScoring2007}
+If the predictive distribution for the quantity $Y$ equals the true data-generating process $F$ (an ideal forecast), the expected CRPS is given by \cite{gneitingStrictlyProperScoring2007}
 \begin{linenomath*}
@@
 This means that the expected CRPS scales roughly with the standard deviation, which in turn typically increases with the mean in epidemiological forecasting.
-In order to make the expected CRPS independent of the expected outcome, a \textit{variance-stabilising transformation} \citep[VST,][]{bartlettSquareRootTransformation1936, dunnGeneralizedLinearModels2018} can be employed.
+In order to make the expected CRPS independent of the expected outcome, a \textit{variance-stabilising transformation} (VST, \cite{bartlettSquareRootTransformation1936, dunnGeneralizedLinearModels2018}) can be employed.
 The choice of this transformation depends on the mean-variance relationship of the underlying process.
@@
 If the mean-variance relationship of the data-generating distribution is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST.
-Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method (a first-order Taylor approximation, see e.g., \citealt{dunnGeneralizedLinearModels2018}), to show that
+Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method (a first-order Taylor approximation, see e.g., \cite{dunnGeneralizedLinearModels2018}), to show that
 
@@
 
-If the mean-variance relationship is linear with $\sigma^2 = c \times \mu$, as with a Poisson-distributed variable, the square root is known to be a VST \citep{dunnGeneralizedLinearModels2018}. 
+If the mean-variance relationship is linear with $\sigma^2 = c \times \mu$, as with a Poisson-distributed variable, the square root is known to be a VST \cite{dunnGeneralizedLinearModels2018}. 
 Denoting by $F_{\sqrt{\ }}$ the predictive distribution for $\sqrt{Y}$, the delta method can again be used to show that
@@
 
-To strengthen our intuition on how transforming outcomes prior to applying the CRPS shifts the emphasis between targets with high and low expected outcomes, Figure \ref{fig:SIM-wis-state-size-mean} shows the expected CRPS of ideal forecasters under different mean-variance relationships and transformations.
+To strengthen our intuition on how transforming outcomes prior to applying the CRPS shifts the emphasis between targets with high and low expected outcomes, Fig \ref{fig:SIM-wis-state-size-mean} shows the expected CRPS of ideal forecasters under different mean-variance relationships and transformations.
 We consider a Poisson distribution where $\sigma^2 = \mu$, a negative binomial distribution with size parameter $\theta = 10$ and thus $\sigma^2 = \mu + \mu^2/10$, and a truncated normal distribution with practically constant variance.
@@
 When applying a square-root-transformation, the expected CRPS is independent of the mean for the Poisson-distribution, but not for the other two (with a positive relationship in the normal case and a negative one for the negative binomial).
-As can be seen in Figures \ref{fig:SIM-wis-state-size-mean} and \ref{fig:score-approx}, the approximations presented in equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear} work quite well for our simulated example.
+As can be seen in Fig \ref{fig:SIM-wis-state-size-mean} and \nameref{fig:score-approx}, the approximations presented in equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear} work quite well for our simulated example.
 
@@
 Solid lines show the corresponding approximations of the expected CRPS from equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear}.
-Figure \ref{fig:score-approx} shows the quality of the approximation in more detail.
+\nameref{fig:score-approx} shows the quality of the approximation in more detail.
     The first distribution (red) is a truncated normal distribution with constant variance (we chose $\sigma = 1$ in order to only obtain positive samples).
@@
 
-\subsection{Effects on model rankings}
+\subsection*{Effects on model rankings}
 \label{sec:methods:rankings}
 Rankings between different forecasters based on the CRPS may change when making use of a transformation, both in terms of aggregate and individual scores.
-We illustrate this in Figure \ref{fig:illustration-ranking} with two forecasters, A and B, issuing two different distributions with different dispersion.
+We illustrate this in Fig \ref{fig:illustration-ranking} with two forecasters, A and B, issuing two different distributions with different dispersion.
 When showing the obtained CRPS as a function of the observed value, it can be seen that the ranking between the two forecasters may change when scoring the forecast on the logarithmic, rather than the natural scale.
@@
 On the log scale, however, forecaster A receives a lower score for large observed values, being more heavily penalised for assigning large probability to small values (which, in relative terms, are far away from the actual observation).
-We note that the chosen example involving a geometric forecast distribution is somewhat constructed; as shown in Section \ref{sec:Hub:cor} and Figure \ref{fig:HUB-cors}A, rankings between models in practice stay quite stable for a single forecast.
+We note that the chosen example involving a geometric forecast distribution is somewhat constructed; as shown in Section \ref{sec:Hub:cor} and Fig \ref{fig:HUB-cors}A, rankings between models in practice stay quite stable for a single forecast.
 
@@
 Overall model rankings would be expected to differ more when scores are averaged across multiple forecasts or targets.
-The change in rankings of aggregate scores usually is mainly driven by the order of magnitude of scores for different forecast targets across time, location and target type and less so by the kind of changes in model rankings for single forecasts discussed above (see Figure \ref{fig:HUB-cors} for a practical example).
+The change in rankings of aggregate scores usually is mainly driven by the order of magnitude of scores for different forecast targets across time, location and target type and less so by the kind of changes in model rankings for single forecasts discussed above (see Fig \ref{fig:HUB-cors} for a practical example).
 Large observations will dominate average CRPS values when evaluation is done on the natural scale, but much less so after log transformation.
@@
 
-\subsection{Practical considerations and other transformations}
+\subsection*{Practical considerations and other transformations}
 \label{sec:methods:considerations}
@@
 In practice, one issue with the log transform is that it is not readily applicable to negative or zero values, which need to be removed or otherwise handled. 
-One common approach to this end is to add a small positive quantity, such as $a = 1$, to all observations and predictions before taking the logarithm \citep{bellegoDealingLogsZeros2022}.
+One common approach to this end is to add a small positive quantity, such as $a = 1$, to all observations and predictions before taking the logarithm \cite{bellegoDealingLogsZeros2022}.
 This still represents a strictly monotonic transformation, but the choice of $a$ does influence scores and rankings (measures of relative errors shrink the larger the chosen value $a$).
@@
 For very large values of $a$, $\log(x + a)$ is roughly linear in $x$, so that using a very large $a$ implies similar relative weighting as applying no transformation at all.
-In practice, a user could explore the effect of different values of $a$ graphically and choose $a$ such that the relative weightings of times and regions with high and low incidence correspond to their preferences (see Figure \ref{fig:HUB-log-different-offsets} in our example application, Section \ref{sec:HUB}).
+In practice, a user could explore the effect of different values of $a$ graphically and choose $a$ such that the relative weightings of times and regions with high and low incidence correspond to their preferences (see Fig \ref{fig:HUB-log-different-offsets} in our example application, Section \ref{sec:HUB}).
 
 A related issue occurs when the predictive distribution has a large probability mass on zero (or on very small values), as this can translate into an excessively wide forecast in relative terms.
-In our applied example this is illustrated in Figure \ref{fig:HUB-model-comparison-baseline}.
+In our applied example this is illustrated in \nameref{fig:HUB-model-comparison-baseline}.
 In such instances, the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm because forecasts contained zero in its prediction intervals.
@@
 The arising issue of dividing by zero can again be solved by adding a small offset $a$.
-Scaling a forecast by the later observed value (as opposed to scaling by past observations) is generally not permissible as it can result in improper scores (see \citealt{lerchForecasterDilemmaExtreme2015} on the closely related topic of weighting scores with a function of the observed value).
+Scaling a forecast by the later observed value (as opposed to scaling by past observations) is generally not permissible as it can result in improper scores (see \cite{lerchForecasterDilemmaExtreme2015} on the closely related topic of weighting scores with a function of the observed value).
 Similarly, scaling forecasts and observations by a function of the predictive distribution (like the predictive mean) may lead to improper scores; however, we are unaware of existing theoretical arguments on this.
@@
 In the case of log transforms, taking the logarithm of the CRPS values, rather than scoring the log-transformed forecasts and data, results in an improper score.
-We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can be seen that in the example overconfident models perform best in terms of the log WIS.
-We note that strictly speaking, re-scaling average scores by the average score of a baseline model or average scores across different models  to obtain skill scores likewise leads to improper scores \citep{gneitingStrictlyProperScoring2007}.
+We illustrate this point using simulated data in \nameref{fig:log-improper}, where it can be seen that in the example overconfident models perform best in terms of the log WIS.
+We note that strictly speaking, re-scaling average scores by the average score of a baseline model or average scores across different models  to obtain skill scores likewise leads to improper scores \cite{gneitingStrictlyProperScoring2007}.
 The application of such skill scores, however, is established practice and considered largely unproblematic.
@@
 
-\section{Empirical example: the European Forecast Hub}
+\section*{Empirical example: the European Forecast Hub}
 \label{sec:HUB}
 
-\subsection{Setting}
+\subsection*{Setting}
 \label{sec:HUB-setting}
 
-As an empirical comparison of evaluating forecasts on the natural and on the log scale, we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
-The European COVID-19 Forecast Hub is one of several COVID-19 Forecast Hubs \citep{cramerEvaluationIndividualEnsemble2021, bracherShorttermForecastingCOVID192021} which have been systematically collecting, aggregating and evaluating forecasts of several COVID-19 targets created by different teams every week.
+As an empirical comparison of evaluating forecasts on the natural and on the log scale, we use forecasts from the European Forecast Hub \cite{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
+The European COVID-19 Forecast Hub is one of several COVID-19 Forecast Hubs \cite{cramerEvaluationIndividualEnsemble2021, bracherShorttermForecastingCOVID192021} which have been systematically collecting, aggregating and evaluating forecasts of several COVID-19 targets created by different teams every week.
 Forecasts are made one to four weeks ahead into the future and follow a quantile-based format with a set of 23 quantiles ($0.01, 0.025, 0.05, ..., 0.5, ...
@@
 See \cite{sherrattPredictivePerformanceMultimodel2022} for a more thorough description of the data.
-We filtered all forecasts submitted to the Hub to only include the seven models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 46 forecast dates (see Figure \ref{fig:HUB-num-avail-models}).
-We removed all observations marked as data anomalies by the European Forecast Hub \citep{sherrattPredictivePerformanceMultimodel2022} as well as all remaining negative observed values.
+We filtered all forecasts submitted to the Hub to only include the seven models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 46 forecast dates (see \nameref{fig:HUB-num-avail-models}).
+We removed all observations marked as data anomalies by the European Forecast Hub \cite{sherrattPredictivePerformanceMultimodel2022} as well as all remaining negative observed values.
 These anomalies made up a relevant fraction of all observations.
 On average across locations, 12.1 out of 92 (13.2\%) observations were removed for cases and 12.4 out of 92 (13.5\%) for deaths.
-Figure \ref{fig:number-anomalies} displays the number of anomalies removed for each location.
-In addition, we filtered out a small number of erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table \ref{tab:erroneous}.
-Figure \ref{fig:erroneous-forecasts} shows the percentage of forecasts removed for each model.
+\nameref{fig:number-anomalies} displays the number of anomalies removed for each location.
+In addition, we filtered out a small number of erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in \nameref{tab:erroneous}.
+\nameref{fig:erroneous-forecasts} shows the percentage of forecasts removed for each model.
 Those few (less than 0.2\%  of forecasts for each model) erroneous outlier forecasts had excessive influence on average scores and relative skill scores in a way that was not representative of normal model behaviour.
@@
 The choice of $a = 1$ in part reflects convention, but also represents a suitable choice as it avoids giving excessive weight to forecasts close to zero, while at the same time ensuring that scores for observations $> 5$ can be interpreted reasonably.
-The analysis was conducted in \texttt{R} \citep{R}, using the \texttt{scoringutils} package \citep{bosseEvaluatingForecastsScoringutils2022} for forecast evaluation.
+The analysis was conducted in \texttt{R} \cite{R}, using the \texttt{scoringutils} package \cite{bosseEvaluatingForecastsScoringutils2022} for forecast evaluation.
 All code is available on GitHub
@@
 
-In addition to the WIS we use pairwise comparisons \citep{cramerEvaluationIndividualEnsemble2021} to evaluate the relative performance of models across countries in the presence of missing forecasts.
+In addition to the WIS we use pairwise comparisons \cite{cramerEvaluationIndividualEnsemble2021} to evaluate the relative performance of models across countries in the presence of missing forecasts.
 In the first step, score ratios are computed for all pairs of models by taking the set of overlapping forecasts between the two models and dividing the score of one model by the score achieved by the other model.
@@
 
-\subsection{Illustration and qualitative observations}
+\subsection*{Illustration and qualitative observations}
 
@@
 
-When comparing examples of forecasts on the natural scale with those on the log scale (see Figures \ref{fig:HUB-model-comparison-ensemble}, \ref{fig:HUB-model-comparison-baseline}, \ref{fig:HUB-model-comparison-epinow}) a few interesting patterns emerge.
+When comparing examples of forecasts on the natural scale with those on the log scale (see Fig \ref{fig:HUB-model-comparison-ensemble}, \nameref{fig:HUB-model-comparison-baseline}, \nameref{fig:HUB-model-comparison-epinow}) a few interesting patterns emerge.
 Missing the peak, i.e.
-predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts during the peak in May 2022 in Figure \ref{fig:HUB-model-comparison-ensemble}A, B).
+predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts during the peak in May 2022 in Fig \ref{fig:HUB-model-comparison-ensemble}A, B).
 On the log scale, these have less of an influence, as errors are smaller in relative terms (see \ref{fig:HUB-model-comparison-ensemble}C, D).
-Conversely, failure to predict an upswing while numbers are still low, is less severely penalised on the natural scale (see forecasts in July 2021 and to a lesser extent in July 2022 in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low.
-On the log scale, missing lower inflection points tends to lead to more severe penalties (see Figure \ref{fig:HUB-model-comparison-ensemble}C, D).
-One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure \ref{fig:HUB-model-comparison-ensemble}E, F).
+Conversely, failure to predict an upswing while numbers are still low, is less severely penalised on the natural scale (see forecasts in July 2021 and to a lesser extent in July 2022 in Fig \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low.
+On the log scale, missing lower inflection points tends to lead to more severe penalties (see Fig \ref{fig:HUB-model-comparison-ensemble}C, D).
+One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Fig \ref{fig:HUB-model-comparison-ensemble}E, F).
 On the log scale, scores do not exhibit this behaviour and rather increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.
@@
 
-Across the dataset, the average number of observed cases and deaths varied considerably by location and target type (see Figure \ref{fig:HUB-mean-locations}A and B).
+Across the dataset, the average number of observed cases and deaths varied considerably by location and target type (see Fig \ref{fig:HUB-mean-locations}A and B).
 On the natural scale, scores show a pattern quite similar to the observations across targets (see Figure\ref{fig:HUB-mean-locations}D) and locations (see Figure\ref{fig:HUB-mean-locations}C).
 On the log scale, scores were more evenly distributed between targets (see Figure\ref{fig:HUB-mean-locations}D) and locations (see Figure\ref{fig:HUB-mean-locations}C).
-Both on the natural scale as well on the log scale, scores increased considerably with increasing forecast horizon (see Figure \ref{fig:HUB-mean-locations}E).
+Both on the natural scale as well on the log scale, scores increased considerably with increasing forecast horizon (see Fig \ref{fig:HUB-mean-locations}E).
 This reflects the increasing difficulty of forecasts further into the future and, for the log scale, corresponds with our expectations from Section \ref{sec:methods:growthrate}.
 
-To assess the impact of the choice of offset value $a$ we extend the display from Figure \ref{fig:HUB-mean-locations}C by results obtained under different specifications.
-Results are shown in Figure \ref{fig:HUB-log-different-offsets}, where for completeness we also added the square root transformation.
+To assess the impact of the choice of offset value $a$ we extend the display from Fig \ref{fig:HUB-mean-locations}C by results obtained under different specifications.
+Results are shown in Fig \ref{fig:HUB-log-different-offsets}, where for completeness we also added the square root transformation.
 As discussed in Section \ref{sec:methods:considerations}, smaller values of $a$ increase the relative weight of smaller locations in the overall evaluation.
 In the most extreme considered case $a = 0.001$, the smallest locations in fact receive the largest weight both for deaths and cases.
-For very large values (see the third row of Figure \ref{fig:HUB-log-different-offsets}), the relative weights strongly resemble those of the evaluation on the natural scale.
+For very large values (see the third row of Fig \ref{fig:HUB-log-different-offsets}), the relative weights strongly resemble those of the evaluation on the natural scale.
 We recommend using displays of this type to get an intuition for the role different locations may play for overall evaluation results.
@@
 
-\subsection{Regression analysis to determine the variance-stabilizing transformation}
+\subsection*{Regression analysis to determine the variance-stabilizing transformation}
 \label{sec:HUB-regression}
@@
 As can be seen from Table \ref{tab:HUB-regression}, the results indeed indicate that small targets lead to larger average WIS when using the log transform ($\beta_{\log} < 0$), while the opposite is true for the square-root transform ($\beta_{\sqrt{\ }} > 0$).
-The results of the three regressions are also displayed in Figure \ref{fig:HUB-regression}.
+The results of the three regressions are also displayed in Fig \ref{fig:HUB-regression}.
 In this empirical example, the log transformation thus helps (albeit not perfectly), to stabilise WIS values, and it does so more successfully than the square-root transformation.
-As can be seen from Figure \ref{fig:HUB-regression}, the expected WIS scores for case targets with medians of 10 and 100,000 differ by more then a factor of ten for the square root transformation, but only a factor of around 2 for the logarithm.
+As can be seen from Fig \ref{fig:HUB-regression}, the expected WIS scores for case targets with medians of 10 and 100,000 differ by more then a factor of ten for the square root transformation, but only a factor of around 2 for the logarithm.
 
-\subsection{Impact of logarithmic transformation on model rankings}
+\subsection*{Impact of logarithmic transformation on model rankings}
 \label{sec:Hub:cor}
@@
 
-For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}A).
+For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Fig \ref{fig:HUB-cors}A).
 While rankings between forecasters remain similar for a single forecast, this is not true anymore when looking at rankings obtained after averaging scores across multiple forecasts made at different times or in different locations.
@@
 when looking at performance during peaks or troughs.
-When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models therefore change considerably (Figure \ref{fig:HUB-cors}B).
+When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models therefore change considerably (Fig \ref{fig:HUB-cors}B).
 The correlation between relative skill scores also decreases noticeably with increasing forecast horizon.
@@
 
-Figure \ref{fig:HUB-rank-order} shows the changes in the ranking between different forecasting models.
+Fig \ref{fig:HUB-rank-order} shows the changes in the ranking between different forecasting models.
 Encouragingly for the European Forecast Hub, the Hub ensemble, which is the forecast the organisers suggest forecast consumers make use of, remains the top model across scoring schemes.
 For cases, the ILM-EKF model and the Forecast Hub baseline model exhibit the largest change in relative skill scores.
-For the ILM-EKF model the relative proportion of the score that is due to overprediction is reduced when applying a log-transformation before scoring (see Figure \ref{fig:HUB-rank-order}E.
+For the ILM-EKF model the relative proportion of the score that is due to overprediction is reduced when applying a log-transformation before scoring (see Fig \ref{fig:HUB-rank-order}E.
 Instances where the model has overshot are penalised less heavily on the log scale, leading to an overall better score.
-For the Forecast Hub baseline model, the fact that it often puts relevant probability mass on zero (see Figure \ref{fig:HUB-model-comparison-baseline}), leads to worse scores after applying log-transformation due to large dispersion penalties.
+For the Forecast Hub baseline model, the fact that it often puts relevant probability mass on zero (see \nameref{fig:HUB-model-comparison-baseline}), leads to worse scores after applying log-transformation due to large dispersion penalties.
 For deaths, the baseline model seems to get similarly penalised for its in relative terms highly dispersed forecasts.
@@
 
-\section{Discussion}
+\section*{Discussion}
 \label{sec:discussion}
@@
 Users nevertheless need to be aware that this can pose issues in practice.
-Including very small values in prediction intervals (see e.g.
-Figure \ref{fig:HUB-model-comparison-baseline}) can lead to excessive dispersion values on the log scale.
+Including very small values in prediction intervals (see \nameref{fig:HUB-model-comparison-baseline} for an example) can lead to excessive dispersion values on the log scale.
 Similarly, locations with lower incidences may get disproportionate weight (i.e.
@@
 The log-transformation is only one of many transformations that may be useful and appropriate in an epidemiological context.
-One obvious option is to apply a population standardization to obtain incidence forecasts e.g., per 100,000 population \citep{abbottEvaluatingEpidemiologicallyMotivated2022}.
+One obvious option is to apply a population standardization to obtain incidence forecasts e.g., per 100,000 population \cite{abbottEvaluatingEpidemiologicallyMotivated2022}.
 We suggested using the natural logarithm as a variance-stabilising transformation (VST).
@@
 approximately true for the negative binomoial distribution and large $\mu$).
-Alternatively, the square-root transformation can be appropriate in the case of a Poisson distributed variable \citep{dunnGeneralizedLinearModels2018}.
-Other VST like the Box-Cox \citep{boxAnalysisTransformations1964} are conceivable as well.
+Alternatively, the square-root transformation can be appropriate in the case of a Poisson distributed variable \cite{dunnGeneralizedLinearModels2018}.
+Other VST like the Box-Cox \cite{boxAnalysisTransformations1964} are conceivable as well.
 If one is interested in multiplicative, rather than exponential growth rates, one could, instead of applying a log transformation, convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last value that was observed at the time the forecast was made.
@@
 Alternatively, one could also take first differences of values on the log scale.
-This approach would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluating the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}.
+This approach would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory (for a different approach to evaluating the shape of a forecast, see \cite{srivastavaShapebasedEvaluationEpidemic2022}).
 Dividing values by the previous value, unfortunately, is not feasible under the current quantile-based format of the Forecast Hubs, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate.
@@
 \appendix
-\section{Supplementary information}
+\section*{Supplementary information}
 
@@
 
-\paragraph{S1 Figure.}
+\paragraph{S1 Fig.}
 \label{fig:log-improper}
@@
 
-\paragraph{S2 Figure.}
+\paragraph{S2 Fig.}
 \label{fig:illustration-effect-log-offset}
@@
 It can be seen that for $a$ values smaller than one fifth of the transformed quantity, the effect of adding an offset is generally small.
-When choosing a suitable $a$, the trade-off is between staying close to the interpretation of a pure log-transformation (choosing a small $a$) and not giving excessive weights to small observations (by choosing a larger $a$, see Figure \ref{fig:HUB-log-different-offsets}).
+When choosing a suitable $a$, the trade-off is between staying close to the interpretation of a pure log-transformation (choosing a small $a$) and not giving excessive weights to small observations (by choosing a larger $a$, see Fig \ref{fig:HUB-log-different-offsets}).
 
-\paragraph{S3 Figure.}
+\paragraph{S3 Fig.}
 \label{fig:score-approx}
 \textbf{Visualisation of expected CRPS values against approximated scores.} 
-This is using the approximation detailed in Section \ref{sec:methods:rankings} (see also Figure \ref{fig:SIM-wis-state-size-mean}).
+This is using the approximation detailed in Section \ref{sec:methods:rankings} (see also Fig \ref{fig:SIM-wis-state-size-mean}).
 Expected CRPS scores are shown for three different distributions once on the natural scale (top row) and once scored on the log scale (bottom row).
@@
 
-\paragraph{S4 Figure.}
+\paragraph{S4 Fig.}
 \label{fig:HUB-num-avail-models}
@@
 
-\paragraph{S5 Figure.}
+\paragraph{S5 Fig.}
 \label{fig:number-anomalies}
@@
 
-\paragraph{S6 Figure.}
+\paragraph{S6 Fig.}
 \label{fig:erroneous-forecasts}
 \textbf{Number of forecasts marked as erroneous and removed. } 
-Forecasts that were in extremely poor agreement with the observed values were removed from the analysis according to the criteria shown in Table \ref{tab:erroneous}.
+Forecasts that were in extremely poor agreement with the observed values were removed from the analysis according to the criteria shown in \nameref{tab:erroneous}.
 
@@
 
-\paragraph{S7 Figure.}
+\paragraph{S7 Fig.}
 \label{fig:HUB-model-comparison-baseline}
@@
 
-\paragraph{S8 Figure.}
+\paragraph{S8 Fig.}
 \label{fig:HUB-model-comparison-epinow}
 \textbf{Forecasts and scores for two-week-ahead predictions from the epiforecasts-EpiNow2 model made in Germany.} 
-A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale from the EpiNow2 model \citep{epinow2}.
+A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale from the EpiNow2 model \cite{epinow2}.
 B, F: Corresponding scores.
@@
 \bibliography{../bib/log-or-not-v2.bib, ../bib/log-or-not.bib, ../bib/software.bib}
-
 \end{document}

### 2023-08-07 final submission after reformatting/final-manuscript.tex
remove plots from manuscript

@@
 \centering
-\includegraphics[width = 1\textwidth]{../output/figures/different-relative-errors.png}
 \caption{Numerical comparison of different measures of relative error: absolute percentage error (APE), relative error (RE), symmetric absolute percentage error (SAPE) and the absolute error applied to log-transformed predictions and observations.
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{../output/figures/SIM-mean-state-size.png}
     \caption{Expected CRPS scores as a function of the mean and variance of the forecast quantity.
@@
 \centering
-\includegraphics[width = 1\textwidth]{../output/figures/illustration-effect-log-ranking-crps.png}
 \caption{Illustration of the effect of the log-transformation of the ranking for a single forecast.
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{../output/figures/HUB-model-comparison-ensemble.png}
     \caption{
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{../output/figures/HUB-mean-obs-location.png}
     \caption{Observations and scores across locations and forecast horizons for the European COVID-19 Forecast Hub data. Locations are sorted according to the mean observed value in that location. 
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{../output/figures/HUB-scores-locations-log-variants.png}
     \caption{Mean WIS in different locations for different transformations applied before scoring.
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{../output/figures/HUB-transformation-regression.png}
     \caption{Relationship between median forecasts and scores.
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{../output/figures/HUB-correlations.png}
     \caption{Correlations of rankings on the natural and logarithmic scale.
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{../output/figures/HUB-pairwise-comparisons.png}
     \caption{Changes in model ratings as measured by relative skill for two-week-ahead predictions for cases (top row) and deaths (bottom row).

### 2023-08-15 final submission after reformatting/final-manuscript.tex
change Figure to Fig and add a single bib file

@@
 Across the dataset, the average number of observed cases and deaths varied considerably by location and target type (see Fig \ref{fig:HUB-mean-locations}A and B).
-On the natural scale, scores show a pattern quite similar to the observations across targets (see Figure\ref{fig:HUB-mean-locations}D) and locations (see Figure\ref{fig:HUB-mean-locations}C).
-On the log scale, scores were more evenly distributed between targets (see Figure\ref{fig:HUB-mean-locations}D) and locations (see Figure\ref{fig:HUB-mean-locations}C).
+On the natural scale, scores show a pattern quite similar to the observations across targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}C).
+On the log scale, scores were more evenly distributed between targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}C).
 Both on the natural scale as well on the log scale, scores increased considerably with increasing forecast horizon (see Fig \ref{fig:HUB-mean-locations}E).
@@
 \clearpage
-\bibliography{../bib/log-or-not-v2.bib, ../bib/log-or-not.bib, ../bib/software.bib}
+\bibliography{final-manuscript.bib}
 \end{document}

### 2023-08-15 final submission after reformatting/final-manuscript.tex
Remove cross-references in manuscript

@@
 The WIS can be decomposed into three components, dispersion, underprediction and overprediction, which reflect the spread of the forecast and whether it was centred above or below the observed value.
-We show an alternative definition based on central prediction intervals in Supplement \ref{sec:alternative-wis} which illustrates this decomposition.
+We show an alternative definition based on central prediction intervals in \nameref{sec:alternative-wis} which illustrates this decomposition.
 
@@
 Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while guaranteeing that the score remains proper.
-Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).
+Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects forecast consumers care most about (see Discussion).
 
@@
 We refer to this transformation as 'log-transformation' and to scores that have been computed from log-transformed forecasts and observations as scores 'on the log scale' (as opposed to scores 'on the natural scale', which involve no transformation).
-In the theoretical discussion in Section \ref{sec:methods}, 'log-transformation' and 'log scale' generally refer to a transformation of $\log_{e}(x)$.
-For practical applications (Section \ref{sec:HUB}) we also use these terms to describe a transformation of $\log_{e}(x + a)$ with a small $a > 0$ in order to keep the terminology and notation simple.
+In the theoretical part of the paper, 'log-transformation' and 'log scale' generally refer to a transformation of $\log_{e}(x)$.
+For practical applications in the later sections we also use these terms to describe a transformation of $\log_{e}(x + a)$ with a small $a > 0$ in order to keep the terminology and notation simple.
 For a prediction target with strictly positive support, the CRPS after applying a log-transformation is given by
@@
 with $F$ the CDF on the original scale.
-Instead of a score representing the magnitude of absolute errors, applying a log-transformation prior to the CRPS yields a score which a) measures relative error (see Section \ref{sec:methods:relative}), b) provides a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate}) and c) is less dependent on the expected order of magnitude of the quantity to be predicted (see Section \ref{sec:methods:vst}).
+Instead of a score representing the magnitude of absolute errors, applying a log-transformation prior to the CRPS yields a score which a) measures relative error, b) provides a measure for how well a forecast captures the exponential growth rate of the target quantity and c) is less dependent on the expected order of magnitude of the quantity to be predicted).
 We therefore argue that such evaluations on the logarithmic scale should complement the prevailing evaluations on the natural scale. 
@@
 The remainder of the article is structured as follows.
-In Sections \ref{sec:methods:relative}--\ref{sec:methods:vst} we provide some mathematical intuition on applying the log-transformation prior to evaluating the CRPS, highlighting the connections to relative error measures, the epidemic growth rate and variance stabilizing transformations.
-We then discuss the effect of the log-transformation on forecast rankings (Section \ref{sec:methods:rankings}) as well as practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}).
-To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  (\cite{europeancovid-19forecasthubEuropeanCovid19Forecast2021,sherrattPredictivePerformanceMultimodel2022}, Section \ref{sec:HUB}).
-Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues (Section \ref{sec:discussion}).
+First, we provide some mathematical intuition on applying the log-transformation prior to evaluating the CRPS, highlighting the connections to relative error measures, the epidemic growth rate and variance stabilizing transformations.
+We then discuss the effect of the log-transformation on forecast rankings as well as practical considerations for applying transformations in general and the log-transformation in particular.
+To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  \cite{europeancovid-19forecasthubEuropeanCovid19Forecast2021,sherrattPredictivePerformanceMultimodel2022}.
+Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues).
 
@@
 As $\sigma$ and $\mu$ are linked through the quadratic mean-variance relationship (or linear mean-standard deviation relationship, $\sigma = \sqrt{c} \times \mu$), the expected CRPS thus stays constant regardless of the expected value of the data-generating distribution $\mu$.
-The assumption of a quadratic mean-variance relationship is closely linked to the aspects discussed in Sections \ref{sec:methods:relative} and \ref{sec:methods:growthrate}.
+The assumption of a quadratic mean-variance relationship is closely linked to the aspects discussed earlier.
 It implies that relative errors have constant variance and can thus be meaningfully compared across different targets.
@@
 On the log scale, however, forecaster A receives a lower score for large observed values, being more heavily penalised for assigning large probability to small values (which, in relative terms, are far away from the actual observation).
-We note that the chosen example involving a geometric forecast distribution is somewhat constructed; as shown in Section \ref{sec:Hub:cor} and Fig \ref{fig:HUB-cors}A, rankings between models in practice stay quite stable for a single forecast.
+We note that the chosen example involving a geometric forecast distribution is somewhat constructed; as shown in Fig \ref{fig:HUB-cors}A, rankings between models in practice stay quite stable for a single forecast.
 
@@
 For very large values of $a$, $\log(x + a)$ is roughly linear in $x$, so that using a very large $a$ implies similar relative weighting as applying no transformation at all.
-In practice, a user could explore the effect of different values of $a$ graphically and choose $a$ such that the relative weightings of times and regions with high and low incidence correspond to their preferences (see Fig \ref{fig:HUB-log-different-offsets} in our example application, Section \ref{sec:HUB}).
+In practice, a user could explore the effect of different values of $a$ graphically and choose $a$ such that the relative weightings of times and regions with high and low incidence correspond to their preferences (see Fig \ref{fig:HUB-log-different-offsets} in our example application).
 
@@
 Some other strictly monotonic transformations that can be applied are scaling by the population size or scaling by past observations.
-The latter, as discussed in Section \ref{sec:discussion}, is similar to applying a log-transformation, but corresponds to evaluating a forecast of multiplicative, rather than exponential growth rates.
+The latter is similar to applying a log-transformation, but corresponds to evaluating a forecast of multiplicative, rather than exponential growth rates.
 The arising issue of dividing by zero can again be solved by adding a small offset $a$.
@@
 These concern e.g., the removal of outlying observations and forecasts and the handling of missing forecasts.
-The solutions we employed in practice are provided in Section \ref{sec:HUB-setting}.
+The solutions we employed in practice are detailed below.
 
@@
     \caption{
-    Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-ensemble made in Germany. Missing values are due to data anomalies that were removed (see section \ref{sec:HUB-setting}. 
+    Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-ensemble made in Germany. Missing values are due to data anomalies that were removed. 
     A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale.
@@
 Both on the natural scale as well on the log scale, scores increased considerably with increasing forecast horizon (see Fig \ref{fig:HUB-mean-locations}E).
-This reflects the increasing difficulty of forecasts further into the future and, for the log scale, corresponds with our expectations from Section \ref{sec:methods:growthrate}.
+This reflects the increasing difficulty of forecasts further into the future and, for the log scale, corresponds with our expectations based on the theoretical considerations detailed above.
 
@@
 Results are shown in Fig \ref{fig:HUB-log-different-offsets}, where for completeness we also added the square root transformation.
-As discussed in Section \ref{sec:methods:considerations}, smaller values of $a$ increase the relative weight of smaller locations in the overall evaluation.
+Smaller values of $a$ increase the relative weight of smaller locations in the overall evaluation.
 In the most extreme considered case $a = 0.001$, the smallest locations in fact receive the largest weight both for deaths and cases.
@@
 
-As argued in Section \ref{sec:methods:vst}, the mean-variance, or mean-CRPS, relationship determines which transformation can serve as a VST.
+As argued above, the mean-variance, or mean-CRPS, relationship determines which transformation can serve as a VST.
 We can analyse this relationship empirically by running a regression that explains the WIS (which approximates the CRPS) as a function of the central estimate of the predictive distribution.
@@
 Note that we are using predictive medians rather than means as only the former are available in the European COVID-19 Forecast Hub.
-As (under the simplifying assumption of normality; see Section \ref{sec:methods:vst}) the WIS/CRPS of an ideal forecaster scales with the standard deviation, a value of $\beta = 1$ would imply a quadratic median-variance relationship; the natural logarithm could then serve as a VST.
+As (under the simplifying assumption of normality; see the previous theoretical discussion on the mean-variance relationship) the WIS/CRPS of an ideal forecaster scales with the standard deviation, a value of $\beta = 1$ would imply a quadratic median-variance relationship; the natural logarithm could then serve as a VST.
 A value of $\beta = 0.5$ would imply a linear median-variance relationship, suggesting the square root as a VST.
@@
 Black dots represent WIS values for two-week ahead predictions of the EuroCOVIDhub-ensemble.
-Shown in red are the regression lines discussed in Section \ref{sec:HUB-regression} shown in Table \ref{tab:HUB-regression}.
+Drawn in red are the regression lines as discussed in the main text and shown in Table \ref{tab:HUB-regression}.
 A: WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against median predicted values.
@@
 \textbf{Visualisation of expected CRPS values against approximated scores.} 
-This is using the approximation detailed in Section \ref{sec:methods:rankings} (see also Fig \ref{fig:SIM-wis-state-size-mean}).
+This is using the approximation detailed in theoretical discussion on model rankings (see also Fig \ref{fig:SIM-wis-state-size-mean}).
 Expected CRPS scores are shown for three different distributions once on the natural scale (top row) and once scored on the log scale (bottom row).

### 2023-08-17 final submission after reformatting/final-manuscript.tex
Change order of Figures in tex file
add reference to S2 Fig
Remove references to figures for the practical example so that figures are mentioned in order as required by PLOS

@@
 On the log scale, however, forecaster A receives a lower score for large observed values, being more heavily penalised for assigning large probability to small values (which, in relative terms, are far away from the actual observation).
-We note that the chosen example involving a geometric forecast distribution is somewhat constructed; as shown in Fig \ref{fig:HUB-cors}A, rankings between models in practice stay quite stable for a single forecast.
+We note that the chosen example involving a geometric forecast distribution is somewhat constructed; as illustrated in our practical example, rankings between models in practice stay quite stable for a single forecast.
 
@@
 Overall model rankings would be expected to differ more when scores are averaged across multiple forecasts or targets.
-The change in rankings of aggregate scores usually is mainly driven by the order of magnitude of scores for different forecast targets across time, location and target type and less so by the kind of changes in model rankings for single forecasts discussed above (see Fig \ref{fig:HUB-cors} for a practical example).
+The change in rankings of aggregate scores usually is mainly driven by the order of magnitude of scores for different forecast targets across time, location and target type and less so by the kind of changes in model rankings for single forecasts discussed above.
 Large observations will dominate average CRPS values when evaluation is done on the natural scale, but much less so after log transformation.
@@
 For very large values of $a$, $\log(x + a)$ is roughly linear in $x$, so that using a very large $a$ implies similar relative weighting as applying no transformation at all.
-In practice, a user could explore the effect of different values of $a$ graphically and choose $a$ such that the relative weightings of times and regions with high and low incidence correspond to their preferences (see Fig \ref{fig:HUB-log-different-offsets} in our example application).
+In practice, a user could explore the effect of different values of $a$ graphically and choose $a$ such that the relative weightings of times and regions with high and low incidence correspond to their preferences.
 
@@
 The choice of $a = 1$ in part reflects convention, but also represents a suitable choice as it avoids giving excessive weight to forecasts close to zero, while at the same time ensuring that scores for observations $> 5$ can be interpreted reasonably.
+\nameref{fig:illustration-effect-log-offset} illustrates the effect of adding a small quantity before taking the logarithm.
 The analysis was conducted in \texttt{R} \cite{R}, using the \texttt{scoringutils} package \cite{bosseEvaluatingForecastsScoringutils2022} for forecast evaluation.
@@
 \subsection*{Illustration and qualitative observations}
+
+When comparing examples of forecasts on the natural scale with those on the log scale (see Fig \ref{fig:HUB-model-comparison-ensemble}, \nameref{fig:HUB-model-comparison-baseline}, \nameref{fig:HUB-model-comparison-epinow}) a few interesting patterns emerge.
+Missing the peak, i.e.
+predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts during the peak in May 2022 in Fig \ref{fig:HUB-model-comparison-ensemble}A, B).
+On the log scale, these have less of an influence, as errors are smaller in relative terms (see \ref{fig:HUB-model-comparison-ensemble}C, D).
+Conversely, failure to predict an upswing while numbers are still low, is less severely penalised on the natural scale (see forecasts in July 2021 and to a lesser extent in July 2022 in Fig \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low.
+On the log scale, missing lower inflection points tends to lead to more severe penalties (see Fig \ref{fig:HUB-model-comparison-ensemble}C, D).
+One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Fig \ref{fig:HUB-model-comparison-ensemble}E, F).
+On the log scale, scores do not exhibit this behaviour and rather increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.
 
@@
 
-When comparing examples of forecasts on the natural scale with those on the log scale (see Fig \ref{fig:HUB-model-comparison-ensemble}, \nameref{fig:HUB-model-comparison-baseline}, \nameref{fig:HUB-model-comparison-epinow}) a few interesting patterns emerge.
-Missing the peak, i.e.
-predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts during the peak in May 2022 in Fig \ref{fig:HUB-model-comparison-ensemble}A, B).
-On the log scale, these have less of an influence, as errors are smaller in relative terms (see \ref{fig:HUB-model-comparison-ensemble}C, D).
-Conversely, failure to predict an upswing while numbers are still low, is less severely penalised on the natural scale (see forecasts in July 2021 and to a lesser extent in July 2022 in Fig \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low.
-On the log scale, missing lower inflection points tends to lead to more severe penalties (see Fig \ref{fig:HUB-model-comparison-ensemble}C, D).
-One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Fig \ref{fig:HUB-model-comparison-ensemble}E, F).
-On the log scale, scores do not exhibit this behaviour and rather increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.
+Across the dataset, the average number of observed cases and deaths varied considerably by location and target type (see Fig \ref{fig:HUB-mean-locations}A and B).
+On the natural scale, scores show a pattern quite similar to the observations across targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}C).
+On the log scale, scores were more evenly distributed between targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}C).
+Both on the natural scale as well on the log scale, scores increased considerably with increasing forecast horizon (see Fig \ref{fig:HUB-mean-locations}E).
+This reflects the increasing difficulty of forecasts further into the future and, for the log scale, corresponds with our expectations based on the theoretical considerations detailed above.
 
@@
 \end{figure}
-
-Across the dataset, the average number of observed cases and deaths varied considerably by location and target type (see Fig \ref{fig:HUB-mean-locations}A and B).
-On the natural scale, scores show a pattern quite similar to the observations across targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}C).
-On the log scale, scores were more evenly distributed between targets (see Fig \ref{fig:HUB-mean-locations}D) and locations (see Fig \ref{fig:HUB-mean-locations}C).
-Both on the natural scale as well on the log scale, scores increased considerably with increasing forecast horizon (see Fig \ref{fig:HUB-mean-locations}E).
-This reflects the increasing difficulty of forecasts further into the future and, for the log scale, corresponds with our expectations based on the theoretical considerations detailed above.
 
@@
 
+To check the relationship after the transformation, we ran the regressions
+\begin{equation}
+    \text{WIS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},
+\end{equation}
+where $F_{\log}$ is the predictive distribution for $\log(y)$, and
+\begin{equation}
+    \text{WIS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{\sqrt{\ }} + \beta_{\sqrt{\ }} \cdot \sqrt{\text{median}(F)},
+\end{equation} 
+where $F_{\sqrt{\ }}$ is the predictive distribution on the square-root scale.
+A value of $\beta_{\log} = 0$ (or $\beta_{\sqrt{\ }} = 0$, respectively) would imply that scores are linearly independent of the median prediction after the transformation.
+A value smaller (larger) than 0 would imply that smaller (larger) targets lead to higher scores.
+As can be seen from Table \ref{tab:HUB-regression}, the results indeed indicate that small targets lead to larger average WIS when using the log transform ($\beta_{\log} < 0$), while the opposite is true for the square-root transform ($\beta_{\sqrt{\ }} > 0$).
+The results of the three regressions are also displayed in Fig \ref{fig:HUB-regression}.
+In this empirical example, the log transformation thus helps (albeit not perfectly), to stabilise WIS values, and it does so more successfully than the square-root transformation.
+As can be seen from Fig \ref{fig:HUB-regression}, the expected WIS scores for case targets with medians of 10 and 100,000 differ by more then a factor of ten for the square root transformation, but only a factor of around 2 for the logarithm.
+
 \begin{figure}[h!]
@@
 
-To check the relationship after the transformation, we ran the regressions
-\begin{equation}
-    \text{WIS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},
-\end{equation}
-where $F_{\log}$ is the predictive distribution for $\log(y)$, and
-\begin{equation}
-    \text{WIS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{\sqrt{\ }} + \beta_{\sqrt{\ }} \cdot \sqrt{\text{median}(F)},
-\end{equation} 
-where $F_{\sqrt{\ }}$ is the predictive distribution on the square-root scale.
-A value of $\beta_{\log} = 0$ (or $\beta_{\sqrt{\ }} = 0$, respectively) would imply that scores are linearly independent of the median prediction after the transformation.
-A value smaller (larger) than 0 would imply that smaller (larger) targets lead to higher scores.
-As can be seen from Table \ref{tab:HUB-regression}, the results indeed indicate that small targets lead to larger average WIS when using the log transform ($\beta_{\log} < 0$), while the opposite is true for the square-root transform ($\beta_{\sqrt{\ }} > 0$).
-The results of the three regressions are also displayed in Fig \ref{fig:HUB-regression}.
-In this empirical example, the log transformation thus helps (albeit not perfectly), to stabilise WIS values, and it does so more successfully than the square-root transformation.
-As can be seen from Fig \ref{fig:HUB-regression}, the expected WIS scores for case targets with medians of 10 and 100,000 differ by more then a factor of ten for the square root transformation, but only a factor of around 2 for the logarithm.
-
 \subsection*{Impact of logarithmic transformation on model rankings}
@@
 
+Fig \ref{fig:HUB-rank-order} shows the changes in the ranking between different forecasting models.
+Encouragingly for the European Forecast Hub, the Hub ensemble, which is the forecast the organisers suggest forecast consumers make use of, remains the top model across scoring schemes.
+For cases, the ILM-EKF model and the Forecast Hub baseline model exhibit the largest change in relative skill scores.
+For the ILM-EKF model the relative proportion of the score that is due to overprediction is reduced when applying a log-transformation before scoring (see Fig \ref{fig:HUB-rank-order}E.
+Instances where the model has overshot are penalised less heavily on the log scale, leading to an overall better score.
+For the Forecast Hub baseline model, the fact that it often puts relevant probability mass on zero (see \nameref{fig:HUB-model-comparison-baseline}), leads to worse scores after applying log-transformation due to large dispersion penalties.
+For deaths, the baseline model seems to get similarly penalised for its in relative terms highly dispersed forecasts.
+The performance of other models changes as well, but patterns are less discernible on this aggregate level.
+
 \begin{figure}[h!]
@@
 \end{figure}
-
-Fig \ref{fig:HUB-rank-order} shows the changes in the ranking between different forecasting models.
-Encouragingly for the European Forecast Hub, the Hub ensemble, which is the forecast the organisers suggest forecast consumers make use of, remains the top model across scoring schemes.
-For cases, the ILM-EKF model and the Forecast Hub baseline model exhibit the largest change in relative skill scores.
-For the ILM-EKF model the relative proportion of the score that is due to overprediction is reduced when applying a log-transformation before scoring (see Fig \ref{fig:HUB-rank-order}E.
-Instances where the model has overshot are penalised less heavily on the log scale, leading to an overall better score.
-For the Forecast Hub baseline model, the fact that it often puts relevant probability mass on zero (see \nameref{fig:HUB-model-comparison-baseline}), leads to worse scores after applying log-transformation due to large dispersion penalties.
-For deaths, the baseline model seems to get similarly penalised for its in relative terms highly dispersed forecasts.
-The performance of other models changes as well, but patterns are less discernible on this aggregate level.
 

### 2023-08-24 final submission after reformatting/final-manuscript.tex
Correct typo in caption for Fig 9, add bib entries manually

@@
 E: Relative contributions of the different WIS components (overprediction, underprediction, and dispersion) to overall model scores on the natural and the logarithmic scale.
-F, G, H, I, J, K: Analogously for deaths.}
+F, G, H, I, J: Analogously for deaths.}
     \label{fig:HUB-rank-order}
@@
 \clearpage
-\bibliography{final-manuscript.bib}
+
+\begin{thebibliography}{10}
+
+\bibitem{heldProbabilisticForecastingInfectious2017}
+Held L, Meyer S, Bracher J.
+\newblock Probabilistic Forecasting in Infectious Disease Epidemiology: The
+  13th {{Armitage}} Lecture.
+\newblock Statistics in Medicine. 2017;36(22):3443--3460.
+\newblock doi:{10.1002/sim.7363}.
+
+\bibitem{doi:10.2105/AJPH.2022.306831}
+Reich NG, Lessler J, Funk S, Viboud C, Vespignani A, Tibshirani RJ, et~al.
+\newblock Collaborative Hubs: {{Making}} the Most of Predictive Epidemic
+  Modeling.
+\newblock American Journal of Public Health. 2022;112(6):839--842.
+\newblock doi:{10.2105/AJPH.2022.306831}.
+
+\bibitem{timmermannForecastingMethodsFinance2018}
+Timmermann A.
+\newblock Forecasting {{Methods}} in {{Finance}}.
+\newblock Annual Review of Financial Economics. 2018;10(1):449--479.
+\newblock doi:{10.1146/annurev-financial-110217-022713}.
+
+\bibitem{gneitingWeatherForecastingEnsemble2005}
+Gneiting T, Raftery AE.
+\newblock Weather {{Forecasting}} with {{Ensemble Methods}}.
+\newblock Science. 2005;310(5746):248--249.
+\newblock doi:{10.1126/science.1115255}.
+
+\bibitem{gneitingStrictlyProperScoring2007}
+Gneiting T, Raftery AE.
+\newblock Strictly {{Proper Scoring Rules}}, {{Prediction}}, and
+  {{Estimation}}.
+\newblock Journal of the American Statistical Association.
+  2007;102(477):359--378.
+\newblock doi:{10.1198/016214506000001437}.
+
+\bibitem{goodRationalDecisions1952}
+Good IJ.
+\newblock Rational {{Decisions}}.
+\newblock Journal of the Royal Statistical Society Series B (Methodological).
+  1952;14(1):107--114.
+
+\bibitem{winklerScoringRulesEvaluation1996}
+Winkler RL, Mu{\~n}oz J, Cervera JL, Bernardo JM, Blattenberger G, Kadane JB,
+  et~al.
+\newblock Scoring Rules and the Evaluation of Probabilities.
+\newblock Test. 1996;5(1):1--60.
+\newblock doi:{10.1007/BF02562681}.
+
+\bibitem{bracherEvaluatingEpidemicForecasts2021}
+Bracher J, Ray EL, Gneiting T, Reich NG.
+\newblock Evaluating Epidemic Forecasts in an Interval Format.
+\newblock PLoS computational biology. 2021;17(2):e1008618.
+\newblock doi:{10.1371/journal.pcbi.1008618}.
+
+\bibitem{Johansson2019}
+Johansson MA, Apfeldorf KM, Dobson S, Devita J, Buczak AL, Baugher B, et~al.
+\newblock An Open Challenge to Advance Probabilistic Forecasting for Dengue
+  Epidemics.
+\newblock Proceedings of the National Academy of Sciences.
+  2019;116(48):24268--24274.
+\newblock doi:{10.1073/pnas.1909865116}.
+
+\bibitem{cramerCOVID19ForecastHub2020}
+Cramer E, Reich NG, Wang SY, Niemi J, Hannan A, House K, et~al.. {{COVID-19
+  Forecast Hub}}: 4 {{December}} 2020 Snapshot; 2020.
+
+\bibitem{cramerEvaluationIndividualEnsemble2021}
+Cramer E, Ray EL, Lopez VK, Bracher J, Brennen A, Rivadeneira AJC, et~al.
+\newblock Evaluation of Individual and Ensemble Probabilistic Forecasts of
+  {{COVID-19}} Mortality in the {{US}}.
+\newblock medRxiv. 2021; p. 2021.02.03.21250974.
+\newblock doi:{10.1101/2021.02.03.21250974}.
+
+\bibitem{sherrattPredictivePerformanceMultimodel2022}
+Sherratt K, Gruson H, Grah R, Johnson H, Niehus R, Prasse B, et~al.
+\newblock Predictive Performance of Multi-Model Ensemble Forecasts of
+  {{COVID-19}} across {{European}} Nation.
+\newblock Europe PMC. 2022;doi:{10.1101/2022.06.16.22276024}.
+
+\bibitem{bracherShorttermForecastingCOVID192021}
+Bracher J, Wolffram D, Deuschel J, G{\"o}rgen K, Ketterer JL, Ullrich A, et~al.
+\newblock Short-Term Forecasting of {{COVID-19}} in {{Germany}} and {{Poland}}
+  during the Second Wave \textendash{} a Preregistered Study.
+\newblock medRxiv. 2021; p. 2020.12.24.20248826.
+\newblock doi:{10.1101/2020.12.24.20248826}.
+
+\bibitem{bracherNationalSubnationalShortterm2021}
+Bracher J, Wolffram D, Deuschel J, Goergen K, Ketterer JL, Ullrich A, et~al.
+\newblock National and subnational short-term forecasting of {COVID}-19 in
+  {Germany} and {Poland}, early 2021.
+\newblock {Communications Medicine}. 2022;doi:{10.1038/s43856-022-00191-8}.
+
+\bibitem{CdcepiFlusightforecastdata2022}
+CDC. Cdcepi/{{Flusight-forecast-data}}; 2022.
+\newblock CDC Epidemic Prediction Initiative.
+
+\bibitem{gosticPracticalConsiderationsMeasuring2020}
+Gostic KM, McGough L, Baskerville E, Abbott S, Joshi K, Tedijanto C, et~al.
+\newblock Practical Considerations for Measuring the Effective Reproductive
+  Number, {{Rt}}.
+\newblock medRxiv. 2020;doi:{10.1101/2020.06.18.20134858}.
+
+\bibitem{dushoffSpeedStrengthEpidemic2021}
+Dushoff J, Park SW.
+\newblock Speed and Strength of an Epidemic Intervention.
+\newblock Proceedings of the Royal Society B: Biological Sciences.
+  2021;288(1947):20201556.
+\newblock doi:{10.1098/rspb.2020.1556}.
+
+\bibitem{bolinLocalScaleInvariance2021}
+Bolin D, Wallin J.
+\newblock Local Scale Invariance and Robustness of Proper Scoring Rules.
+\newblock Statistical Science. 2023;38(1):140--159.
+\newblock doi:{10.1214/22-STS864}.
+
+\bibitem{taylorEvaluatingVolatilityInterval1999}
+Taylor JW.
+\newblock Evaluating Volatility and Interval Forecasts.
+\newblock Journal of Forecasting. 1999;18(2):111--128.
+\newblock doi:{10.1002/(SICI)1099-131X(199903)18:2<111::AID-FOR713>3.0.CO;2-C}.
+
+\bibitem{mayrLogLevelVAR2015}
+Mayr J, Ulbricht D.
+\newblock Log versus Level in {{VAR}} Forecasting: 42 Million Empirical
+  Answers\textemdash{{Expect}} the Unexpected.
+\newblock Economics Letters. 2015;126:40--42.
+\newblock doi:{10.1016/j.econlet.2014.11.008}.
+
+\bibitem{loweStochasticRainfallrunoffForecasting2014}
+L{\"o}we R, Mikkelsen PS, Madsen H.
+\newblock Stochastic Rainfall-Runoff Forecasting: Parameter Estimation,
+  Multi-Step Prediction, and Evaluation of Overflow Risk.
+\newblock Stochastic Environmental Research and Risk Assessment.
+  2014;28(3):505--516.
+\newblock doi:{10.1007/s00477-013-0768-0}.
+
+\bibitem{fuglstadDoesNonstationarySpatial2015}
+Fuglstad GA, Simpson D, Lindgren F, Rue H.
+\newblock Does Non-Stationary Spatial Data Always Require Non-Stationary Random
+  Fields?
+\newblock Spatial Statistics. 2015;14:505--531.
+\newblock doi:{10.1016/j.spasta.2015.10.001}.
+
+\bibitem{Lehmann1950}
+Lehmann EL.
+\newblock {Some Principles of the Theory of Testing Hypotheses}.
+\newblock The Annals of Mathematical Statistics. 1950;21(1):1 -- 26.
+\newblock doi:{10.1214/aoms/1177729884}.
+
+\bibitem{diksLikelihoodbasedScoringRules2011}
+Diks C, Panchenko V, {van Dijk} D.
+\newblock Likelihood-Based Scoring Rules for Comparing Density Forecasts in
+  Tails.
+\newblock Journal of Econometrics. 2011;163(2):215--230.
+\newblock doi:{10.1016/j.jeconom.2011.04.001}.
+
+\bibitem{europeancovid-19forecasthubEuropeanCovid19Forecast2021}
+{European Covid-19 Forecast Hub}. European {{Covid-19 Forecast Hub}}; 2021.
+\newblock https://covid19forecasthub.eu/.
+
+\bibitem{gneitingMakingEvaluatingPoint2011a}
+Gneiting T.
+\newblock Making and {{Evaluating Point Forecasts}}.
+\newblock Journal of the American Statistical Association.
+  2011;106(494):746--762.
+\newblock doi:{10.1198/jasa.2011.r10138}.
+
+\bibitem{Flores1986}
+Flores BE.
+\newblock A pragmatic view of accuracy measurement in forecasting.
+\newblock Omega. 1986;14(2):93--98.
+\newblock doi:{https://doi.org/10.1016/0305-0483(86)90013-7}.
+
+\bibitem{wallingaHowGenerationIntervals2007}
+Wallinga J, Lipsitch M.
+\newblock How Generation Intervals Shape the Relationship between Growth Rates
+  and Reproductive Numbers.
+\newblock Proceedings of the Royal Society B: Biological Sciences.
+  2007;274(1609):599--604.
+\newblock doi:{10.1098/rspb.2006.3754}.
+
+\bibitem{bartlettSquareRootTransformation1936}
+Bartlett MS.
+\newblock The {{Square Root Transformation}} in {{Analysis}} of {{Variance}}.
+\newblock Supplement to the Journal of the Royal Statistical Society.
+  1936;3(1):68--78.
+\newblock doi:{10.2307/2983678}.
+
+\bibitem{dunnGeneralizedLinearModels2018}
+Dunn PK, Smyth GK.
+\newblock Generalized {{Linear Models With Examples}} in {{R}}.
+\newblock {Springer}; 2018.
+
+\bibitem{bellegoDealingLogsZeros2022}
+Bell{\'e}go C, Benatia D, Pape L. Dealing with {{Logs}} and {{Zeros}} in
+  {{Regression Models}}; 2022.
+
+\bibitem{lerchForecasterDilemmaExtreme2015}
+Lerch S, Thorarinsdottir TL, Ravazzolo F, Gneiting T. Forecaster's {{Dilemma}}:
+  {{Extreme Events}} and {{Forecast Evaluation}}; 2015.
+
+\bibitem{R}
+{R Core Team}. R: A Language and Environment for Statistical Computing; 2022.
+\newblock Available from: \url{https://www.R-project.org/}.
+
+\bibitem{bosseEvaluatingForecastsScoringutils2022}
+Bosse NI, Gruson H, Cori A, {van Leeuwen} E, Funk S, Abbott S.
+\newblock Evaluating {{Forecasts}} with Scoringutils in {{R}}.
+\newblock arXiv. 2022;doi:{10.48550/arXiv.2205.07090}.
+
+\bibitem{abbottEvaluatingEpidemiologicallyMotivated2022}
+Abbott S, Sherratt K, Bosse N, Gruson H, Bracher J, Funk S. Evaluating an
+  Epidemiologically Motivated Surrogate Model of a Multi-Model Ensemble; 2022.
+
+\bibitem{boxAnalysisTransformations1964}
+Box GEP, Cox DR.
+\newblock An {{Analysis}} of {{Transformations}}.
+\newblock Journal of the Royal Statistical Society Series B (Methodological).
+  1964;26(2):211--252.
+
+\bibitem{srivastavaShapebasedEvaluationEpidemic2022}
+Srivastava A, Singh S, Lee F. Shape-Based {{Evaluation}} of {{Epidemic
+  Forecasts}}; 2022.
+
+\bibitem{epinow2}
+Abbott S, Hellewell J, Sherratt K, Gostic K, Hickson J, Badr HS, et~al..
+  EpiNow2: Estimate Real-Time Case Counts and Time-Varying Epidemiological
+  Parameters; 2020.
+
+\end{thebibliography}
+
 \end{document}

### 2023-09-30 final submission after reformatting/Supplementary-information.tex
Add SI as single PDF

--- 2023-06-26 main.tex
@@
 
-\title{Scoring epidemiological forecasts on transformed scales}
-  \author{Nikos I.
-Bosse\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom} $^{,}$\thanks{Centre for the Mathematical Modelling of Infectious Diseases, London, United Kingdom} $^{,}$\thanks{NIHR Health Protection Research Unit in Modelling \& Health Economics} $^{,*}$,
-  Sam Abbott\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$, 
-  Anne Cori\thanks{MRC Centre for Outbreak Analysis and Modelling, Department of Infectious Disease Epidemiology, School of Public Health, Imperial College London, London, United Kingdom} $^{}$, \\
-  Edwin van Leeuwen\footnotemark[1] $^{,}$\thanks{Modelling \& Economics Unit and NIHR Health Protection Research Unit in Modelling \& Health Economics, UK Health Security Agency, London, United Kingdom} $^{}$,
-  Johannes Bracher\thanks{Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology, Karlsruhe, Germany } $^{,}$\thanks{Computational Statistics Group, Heidelberg Institute for Theoretical Studies, Heidelberg, Germany } $^{, \dagger}$,
-  Sebastian Funk\footnotemark[1] $^{, }$\footnotemark[2] $^{, }$\footnotemark[3] $ ^{, \dagger}$}
-
-\maketitle
-\begin{abstract}
-
-Forecast evaluation is essential for the development of predictive epidemic models and can inform their use for public health decision-making.
-Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which can be seen as measures of the absolute distance between the forecast distribution and the observation.
-However, applying these scores directly to predicted and observed incidence counts may not be the most appropriate due to the exponential nature of epidemic processes and the varying magnitudes of observed values across space and time.
-In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results.
-Using the CRPS on log-transformed values as an example, we list three attractive properties: Firstly, it can be interpreted as a probabilistic version of a relative error.
-Secondly, it reflects how well models predicted the time-varying epidemic growth rate.
-And lastly, using arguments on variance-stabilizing transformations, it can be shown that under the assumption of a quadratic mean-variance relationship, the logarithmic transformation leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity.
-Applying a transformation of log(x + 1) to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types.
-Situations in which models missed the beginning of upward swings are more strongly emphasised while failing to predict a downturn following a peak is less severely penalised when scoring transformed forecasts as opposed to untransformed ones.
-We conclude that appropriate transformations, of which the natural logarithm is only one particularly attractive option, should be considered when assessing the performance of different models in the context of infectious disease incidence.
-\end{abstract}
-
-\bigskip
-
-{\footnotesize $^*$ Correspondence to \url{nikos.bosse@lshtm.ac.uk}}, 
-{\footnotesize $^\dagger$ Contributed equally}
-
-\newpage
-
-% ===========================================================
-\section{Introduction}
-
-Probabilistic forecasts \citep{heldProbabilisticForecastingInfectious2017} play an important role in decision-making in epidemiology and public health \citep{doi:10.2105/AJPH.2022.306831}, as well as other areas as diverse as economics \citep{timmermannForecastingMethodsFinance2018} or meteorology \citep{gneitingWeatherForecastingEnsemble2005}.
-Forecasts based on epidemiological modelling in particular have received widespread attention during the COVID-19 pandemic.
-Evaluations of forecasts can provide feedback for researchers to improve their models and train ensembles.
-They moreover help decision-makers distinguish good from bad predictions and choose forecasters and models that are best suited to inform future decisions.
-
-Probabilistic forecasts are usually evaluated using so-called (strictly) proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data.
-Proper scoring rules are constructed such that they encourage honest forecasting and cannot be `gamed' or `cheated'. 
-Assuming that the forecaster's actual best judgement corresponds to a predictive distribution $F$, a proper score is constructed such that if $F$ was the data-generating process, no other distribution $G$ would yield a better expected score.
-A scoring rule is called \textit{strictly} proper if there is no other distribution that under $F$ achieves the \textit{same} expected score as $F$, meaning that any deviation from $F$ leads to a worsening of expected scores.
-Forecasters (anyone or anything that issues a forecast) are thus incentivised to report their true belief $F$ about the future.
-Common proper scoring rules are the logarithmic or log score \citep{goodRationalDecisions1952} and the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}.
-The log score is the predictive log density or probability mass evaluated at the observed value.
-It is supported by the likelihood principle \citep{Winkler1996} and has many desirable theoretical properties; however, the particularly severe penalties it assigns to occasional misguided forecasts make it little robust \citep{bracherEvaluatingEpidemicForecasts2021}.
-Moreover, it is not easily applied to forecasts reported as samples or quantiles, as used in many recent disease forecasting efforts.
-It is nonetheless occasionally used in epidemiology (see e.g., \citealt{heldProbabilisticForecastingInfectious2017, Johansson2019}), but in recent years the CRPS and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021} have become increasingly popular.
-
-The CRPS measures the distance of the predictive distribution to the observed data as 
-\begin{linenomath*}
-\begin{equation}
-    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - \boldsymbol{1}(x \geq y) \right)^2 dx,
-\end{equation}    
-\end{linenomath*}
-where $y$ is the true observed value, $F$ is the cumulative distribution function (CDF) of the predictive distribution, and $\boldsymbol{1}()$ is the indicator function.
-The CRPS can be understood as a generalisation of the absolute error to predictive distributions, and interpreted on the natural scale of the data.
-The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in the so-called COVID-19 Forecast Hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022} and Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US \textit{FluSight} project on influenza forecasting\citep{CdcepiFlusightforecastdata2022}.
-The WIS is defined as
-\begin{linenomath*}
-\begin{equation}
-    \text{WIS}(F, y) = \frac{1}{K} \times \sum_{k = 1}^{K} 2 \times \left[ \boldsymbol{1}(y \leq q_{\tau_k}) - \tau_k \right] \times ( q_{\tau_k} - y), 
-\end{equation}
-\end{linenomath*}
-where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome and $K$ is the number of (roughly equally spaced) predictive quantiles provided.
-The WIS can be decomposed into three components, dispersion, underprediction and overprediction, which reflect the spread of the forecast and whether it was centred above or below the observed value.
-We show an alternative definition based on central prediction intervals in Supplement \ref{sec:alternative-wis} which illustrates this decomposition.
-
-The notion of absolute distance encoded by the CRPS and WIS provides a straightforward interpretation, but may not always be the most useful perspective in the context of infectious disease spread.
-Especially in their early phase, outbreaks are best conceived as exponential processes, characterized by potentially time varying reproduction numbers $R_t$ \citep{gosticPracticalConsiderationsMeasuring2020} or epidemic growth rates $r_t$ \citep{dushoffSpeedStrengthEpidemic2021}.
-If the true modelling task revolves around estimating and forecasting these quantities, then evaluating forecasts based on the absolute distance between forecasted and observed incidence values penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount.
-For illustration, consider an incidence forecast issued at time 0 and referring to time $t$ that misses the correct average growth rate $\bar{r}_t$ by either $-\epsilon$ or $+\epsilon$.
-Then the ratio of the resulting absolute errors on the scale of observed incidences $y_{t}$ is
-
-\begin{linenomath*}
-\begin{equation}
-\frac{\left|y_0 \exp[(\bar{r}_t - \epsilon) \times t] - y_0 \exp(\bar{r}_t t)\right|}{\left| y_0 \exp[(\bar{r}_t + \epsilon) \times t] - y_0 \exp(\bar{r}_t t) \right|} = \exp(-\epsilon t) < 1.
-\end{equation}
-\end{linenomath*}
-
-% For exponential processes errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate. 
-If one is to measure the ability to forecast the underlying infection dynamics, it may thus be more desirable to evaluate errors on the scale of the growth rate directly.
-
-Another argument against using notions of absolute distance between predicted and observed incidence values is that forecast consumers may find errors on a relative scale easier to interpret and more useful in order to track predictive performance across targets of different orders of magnitude.
-\cite{bolinLocalScaleInvariance2021} have proposed the scaled CRPS (SCRPS) which is locally scale invariant; however, it does not correspond to a relative error measure and lacks a straightforward interpretation as available for the CRPS.
-
-% A closely related aspect to relative scores (as opposed to absolute scores) is that in the evaluation one may wish to give similar weight to all 
-Lastly, it may be considered desirable to give all forecast targets similar weight in an overall performance evaluation.
-As the CRPS typically scales with the order of magnitude of the quantity to be predicted, this is not the case for the CRPS, which will typically assign higher scores to forecast targets with high expected values (e.g., in large locations or around the peak of an epidemic).
-\cite{bracherEvaluatingEpidemicForecasts2021} have argued that this is a desirable feature, directing attention to situations of particular public health relevance.
-An evaluation based on absolute errors, however, will assign little weight to other potentially important aspects, such as the ability to correctly predict future upswings while observed numbers are still low.
-
-In many fields, it is common practice to forecast transformed quantities (see e.g.
-\cite{taylorEvaluatingVolatilityInterval1999} in finance, \cite{mayrLogLevelVAR2015} in macroeconomics, \cite{loweStochasticRainfallrunoffForecasting2014} in hydrology or \cite{fuglstadDoesNonstationarySpatial2015} in meteorology).
-While the goal of the transformations is often to improve the accuracy of the predictions, they can also be used to enhance and complement the evaluation process.
-In this paper, we argue that the aforementioned issues with evaluating epidemic forecasts based on measures of absolute error on the natural scale can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation.
-Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while guaranteeing that the score remains proper.
-Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).
-
-For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particularly attractive transformation in the context of epidemic phenomena.
-We refer to this transformation as 'log-transformation' and to scores that have been computed from log-transformed forecasts and observations as scores 'on the log scale' (as opposed to scores 'on the natural scale', which involve no transformation).
-In the theoretical discussion in Section \ref{sec:methods}, 'log-transformation' and 'log scale' generally refer to a transformation of $\log_{e}(x)$.
-For practical applications (Section \ref{sec:HUB}) we also use these terms to describe a transformation of $\log_{e}(x + a)$ with a small $a > 0$ in order to keep the terminology and notation simple.
-For a prediction target with strictly positive support, the CRPS after applying a log-transformation is given by
-%
-\begin{linenomath*}
-\begin{equation}
-    \text{CRPS}(F_{\log}, \log y) = \int_{-\infty}^\infty \left( F_{\log}(x) - \boldsymbol{1}(x \geq \log y) \right)^2 dx.
-\end{equation}    
-\end{linenomath*}
-%
-Here, $y$ is again the observed outcome and $F_{\log}$ is the predictive CDF of the log-transformed outcome, i.e.,
-\begin{linenomath*}
-\begin{equation}
-F_{\log}(x) = F(\exp(x)),
-\end{equation}    
-\end{linenomath*}
-with $F$ the CDF on the original scale.
-Instead of a score representing the magnitude of absolute errors, applying a log-transformation prior to the CRPS yields a score which a) measures relative error (see Section \ref{sec:methods:relative}), b) provides a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate}) and c) is less dependent on the expected order of magnitude of the quantity to be predicted (see Section \ref{sec:methods:vst}).
-We therefore argue that such evaluations on the logarithmic scale should complement the prevailing evaluations on the natural scale. 
-Other transformations may likewise be of interest. We briefly explore the square root transformation as an alternative transformation. 
-Our analysis mostly focuses on the CRPS (or WIS) as an evaluation metric for probabilistic forecasts, given its widespread use throughout the COVID-19 pandemic.
-We note that the logarithmic score has scale invariance properties which imply that score differences between different forecasts are invariant to strictly monotonic transformations (see \citealt{Lehmann1950} on corresponding properties of likelihood ratios and \citealt{diksLikelihoodbasedScoringRules2011}).
-The question of the right scale to evaluate forecasts on does therefore not arise for the log score.
-
-The remainder of the article is structured as follows.
-In Sections \ref{sec:methods:relative}--\ref{sec:methods:vst} we provide some mathematical intuition on applying the log-transformation prior to evaluating the CRPS, highlighting the connections to relative error measures, the epidemic growth rate and variance stabilizing transformations.
-We then discuss the effect of the log-transformation on forecast rankings (Section \ref{sec:methods:rankings}) as well as practical considerations for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}).
-To analyse the real-world implications of the log-transformation we use forecasts submitted to the European COVID-19 Forecast Hub  \citep[ Section \ref{sec:HUB}]{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
-Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues (Section \ref{sec:discussion}).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Logarithmic transformation of forecasts and observations}
-\label{sec:methods}
-
-\subsection{Interpretation as a relative error}
-\label{sec:methods:relative}
-
-To illustrate the effect of applying the natural logarithm prior to evaluating forecasts we consider the absolute error, which the CRPS and WIS generalize to probabilistic forecasts.
-We assume strictly positive support (meaning that no specific handling of zero values is needed), a restriction we will address when applying this transformation in practice.
-When considering a point forecast $\hat{y}$ for a quantity of interest $y$, such that
-%
-\begin{linenomath*}
-\begin{equation}
-y = \hat{y} + \varepsilon,
-\end{equation}
-\end{linenomath*}
-the absolute error is given by $|\varepsilon|$. When taking the logarithm of the forecast and the observation first, thus considering 
-\begin{linenomath*}
-\begin{equation}
-\log y = \log \hat{y} + \varepsilon^*,
-\end{equation}
-\end{linenomath*}
-the resulting absolute error $\left|\varepsilon^*\right|$ can be interpreted as an approximation of various common relative error measures.
-Using that $\log(a) \approx a - 1$ if $a$ is close to 1, we get
-\begin{linenomath*}
-\begin{equation}
-|\varepsilon^*| = |\log \hat{y} - \log y| = \left|\log\left(\frac{\hat{y}}{y}\right) \right| \ \ \stackrel{\text{if } \hat{y} \ \approx \ y}{\approx} \ \ \left| \frac{\hat{y}}{y} - 1 \right| \ \ = \ \ \left| \frac{\hat{y} - y}{y} \right|.
-\end{equation}
-\end{linenomath*}
-The absolute error after log transforming is thus an approximation of the \textit{absolute percentage error} \citep[APE,][]{gneitingMakingEvaluatingPoint2011a} as long as forecast and observation are close.
-As we assumed that $\hat{y} \approx y$, we can also interpret it as an approximation of the \textit{relative error} \citep[RE,][]{gneitingMakingEvaluatingPoint2011a}
-\begin{linenomath*}
-\begin{equation}
-\left| \frac{\hat{y} - y}{\hat{y}} \right|
-\end{equation}
-\end{linenomath*}
-and the \textit{symmetric absolute percentage error} (SAPE; see e.g., \citealt{Flores1986})
-\begin{linenomath*}
-\begin{equation}
-\left| \ \frac{\hat{y} - y}{y/2 + \hat{y}/2} \ \right|.
-\end{equation}
-\end{linenomath*}
-As Figure \ref{fig:SAPE} shows, the alignment with the SAPE is in fact the closest and holds quite well even if predicted and observed value differ by a factor of two or three.
-Generalising to probabilistic forecasts, the CRPS applied to log-transformed forecasts and outcomes can thus be seen as a probabilistic counterpart to the symmetric absolute percentage error, which offers an appealing intuitive interpretation.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width = 1\textwidth]{output/figures/different-relative-errors.png}
-\caption{Numerical comparison of different measures of relative error: absolute percentage error (APE), relative error (RE), symmetric absolute percentage error (SAPE) and the absolute error applied to log-transformed predictions and observations.
-We denote the predicted value by $\hat{y}$ and display errors as a function of the ratio of observed and predicted value.
-A: x-axis shown on a linear scale.
-B: x-axis shown on a logarithmic scale.}
-\label{fig:SAPE}
-\end{figure}
-
-\subsection{Interpretation as scoring the exponential growth rate}
-\label{sec:methods:growthrate}
-
-Another interpretation for the log-transform is possible if the generative process is framed as exponential with a time-varying growth rate $r(t)$~\citep[see, e.g.,][]{wallingaHowGenerationIntervals2007}, i.e.
-\begin{linenomath*}
-\begin{equation}
-\frac{d}{dt}y(t) = r(t)y(t)
-\end{equation}
-\end{linenomath*}
-%
-which is solved by
-%
-\begin{linenomath*}
-\begin{equation}
-y(t) = y_0 \exp \left( \int_0^t r(t') dt' \right) = y_0 \exp (\bar{r_t}t)
-\end{equation}
-\end{linenomath*}
-where $y_0$ is an initial data point and $\bar{r_t}$ is the mean of the growth rate between the initial time point $0$ and time $t$.
-
-If a forecast $\hat{y}(t)$ for the value of the time series at time $t$ is issued at time 0 based on the data point $y_0$ then the absolute error after log transformation is
-%
-\begin{linenomath*}
-\begin{align}
-\begin{split}
-\epsilon^* &= \left| \log \left[ \hat{y}( t ) \right] - \log \left[ y ( t ) \right] \right|\\
-           &= \left| \log \left[ y_0 \exp (\hat{\bar{r_t}} t ) \right] - \log \left[ y_0 \exp (\bar{r_t}t) \right] \right|\\
-           &= t \left| \hat{\bar{r_t}} - \bar{r_t} \right|
-\end{split}
-\end{align}
-\end{linenomath*}
-%
-where $\bar{r_t}$ is the true mean growth rate and $\hat{\bar{r_t}}$ is the forecast mean growth rate.
-We thus evaluate the error in the mean exponential growth rate, scaled by the length of the time period considered.
-Again generalising this to the CRPS and WIS implies a probabilistic evaluation of forecasts of the epidemic growth rate.
-
-\subsection{Interpretation as a variance-stabilising transformation}
-\label{sec:methods:vst}
-
-When evaluating models across sets of forecasting tasks, it may be desirable for each target to have a similar impact on the overall results.
-This could be motivated by the assumption that forecasts from different geographical units and time periods provide similar amounts of information about how well a forecaster performs.
-One would then like the resulting scores to be independent of the order of magnitude of the target to predict.
-CRPS values on the natural scale, however, typically scale with the order of magnitude of the quantity to be predicted.
-Average scores are then dominated by the results achieved for targets with high expected outcomes in a way that does not necessarily reflect the underlying predictive ability well.
-
-If the predictive distribution for the quantity $Y$ equals the true data-generating process $F$ (an ideal forecast), the expected CRPS is given by \citep{gneitingStrictlyProperScoring2007}
-\begin{linenomath*}
-\begin{equation}
-\mathbb{E}[\text{CRPS}(F, y)] = 0.5\times\mathbb{E}|Y - Y'|,
-\end{equation}
-\end{linenomath*}
-where $Y$ and $Y'$ are independent samples from $F$.
-This corresponds to half the \textit{mean absolute difference}, which is a measure of dispersion.
-If $F$ is well-approximated by a normal distribution $\text{N}(\mu, \sigma^2)$, the approximation
-\begin{linenomath*}
-\begin{equation}
-\mathbb{E}_F[\text{CRPS}(F, y)] \approx \frac{\sigma}{\sqrt{\pi}}
-\end{equation}
-\end{linenomath*}
-can be used.
-This means that the expected CRPS scales roughly with the standard deviation, which in turn typically increases with the mean in epidemiological forecasting.
-In order to make the expected CRPS independent of the expected outcome, a \textit{variance-stabilising transformation} \citep[VST,][]{bartlettSquareRootTransformation1936, dunnGeneralizedLinearModels2018} can be employed.
-The choice of this transformation depends on the mean-variance relationship of the underlying process.
-
-If the mean-variance relationship of the data-generating distribution is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST.
-Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method (a first-order Taylor approximation, see e.g., \citealt{dunnGeneralizedLinearModels2018}), to show that
-
-\begin{linenomath*}
-\begin{equation}
-\mathbb{E}_F[\text{CRPS}\{F_{\log}, \log(y)\}] \approx \frac{\sigma/\mu}{\sqrt{\pi}} 
-= \frac{\sqrt{c}}{\sqrt{\pi}}
-.\label{eq:taylor_quadratic}
-\end{equation}
-\end{linenomath*}
-As $\sigma$ and $\mu$ are linked through the quadratic mean-variance relationship (or linear mean-standard deviation relationship, $\sigma = \sqrt{c} \times \mu$), the expected CRPS thus stays constant regardless of the expected value of the data-generating distribution $\mu$.
-The assumption of a quadratic mean-variance relationship is closely linked to the aspects discussed in Sections \ref{sec:methods:relative} and \ref{sec:methods:growthrate}.
-It implies that relative errors have constant variance and can thus be meaningfully compared across different targets.
-Also, it arises naturally if we assume that our capacity to predict the epidemic growth rate does not depend on the expected outcome, i.e.
-does not depend on the current phase of the epidemic or the order of magnitude of current observations.
-
-If the mean-variance relationship is linear with $\sigma^2 = c \times \mu$, as with a Poisson-distributed variable, the square root is known to be a VST \citep{dunnGeneralizedLinearModels2018}. 
-Denoting by $F_{\sqrt{\ }}$ the predictive distribution for $\sqrt{Y}$, the delta method can again be used to show that
-\begin{linenomath*}
-\begin{equation}
-\mathbb{E}_F[\text{CRPS}\{F_{\sqrt{\ }}, \sqrt{y}\}] \approx \frac{\sigma/\sqrt{\mu}}{2\sqrt{\pi}} = \frac{\sqrt{c}}{2\sqrt{\pi}}
-.\label{eq:taylor_linear}
-\end{equation}
-\end{linenomath*}
-We note that while standard in the derivation of variance-stabilizing transformations, the application of the delta method in equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear} requires the probability mass of $F$ to be tightly distributed.
-If this is not the case, the approximation and thus the variance stabilization may be less accurate.
-
-To strengthen our intuition on how transforming outcomes prior to applying the CRPS shifts the emphasis between targets with high and low expected outcomes, Figure \ref{fig:SIM-wis-state-size-mean} shows the expected CRPS of ideal forecasters under different mean-variance relationships and transformations.
-We consider a Poisson distribution where $\sigma^2 = \mu$, a negative binomial distribution with size parameter $\theta = 10$ and thus $\sigma^2 = \mu + \mu^2/10$, and a truncated normal distribution with practically constant variance.
-We see that when applying the CRPS on the natural scale, the expected CRPS grows monotonically as the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster) increases.
-The expected CRPS is constant only for the distribution with constant variance, and grows in $\mu$ for the other two.
-When applying a log-transformation first, the expected CRPS is almost independent of $\mu$ for the negative binomial distribution and large $\mu$, while smaller targets have higher expected CRPS in case of the Poisson distribution and the normal distribution with constant variance.
-When applying a square-root-transformation, the expected CRPS is independent of the mean for the Poisson-distribution, but not for the other two (with a positive relationship in the normal case and a negative one for the negative binomial).
-As can be seen in Figures \ref{fig:SIM-wis-state-size-mean} and \ref{fig:score-approx}, the approximations presented in equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear} work quite well for our simulated example.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/SIM-mean-state-size.png}
-    \caption{Expected CRPS scores as a function of the mean and variance of the forecast quantity.
-We computed expected CRPS values  for three different distributions, assuming an ideal forecaster with predictive distribution equal to the true underlying (data-generating) distribution.
-    These expected CRPS values were computed for different predictive means based on 10,000 samples each and are represented by dots.
-Solid lines show the corresponding approximations of the expected CRPS from equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear}.
-Figure \ref{fig:score-approx} shows the quality of the approximation in more detail.
-    The first distribution (red) is a truncated normal distribution with constant variance (we chose $\sigma = 1$ in order to only obtain positive samples).
-The second (green) is a negative binomial distribution with variance $\theta = 10$ and variance $\sigma^2 = \mu + 0.1\mu^2$.
-The third (blue) is a
-    Poisson distribution with $\sigma^2 = \mu$.
-To make the scores for the different distributions comparable, scores were normalised to one, meaning that the mean score for every distribution (red, green, blue) is one.
-    A: Normalised expected CRPS for ideal forecasts with increasing means for three distribution with different relationships between mean and variance.
-Expected CRPS was computed on the natural scale (left), after applying a square-root transformation (middle), and after adding one and applying a log-transformation to the data (right).
-B: A but with x and y axes on the log scale.}
-    \label{fig:SIM-wis-state-size-mean}
-\end{figure}
-
-\subsection{Effects on model rankings}
-\label{sec:methods:rankings}
-Rankings between different forecasters based on the CRPS may change when making use of a transformation, both in terms of aggregate and individual scores.
-We illustrate this in Figure \ref{fig:illustration-ranking} with two forecasters, A and B, issuing two different distributions with different dispersion.
-When showing the obtained CRPS as a function of the observed value, it can be seen that the ranking between the two forecasters may change when scoring the forecast on the logarithmic, rather than the natural scale.
-In particular, on the natural scale, forecaster A, who issues a more uncertain distribution, receives a better score than forecaster B for observed values far away from the centre of the respective predictive distribution.
-On the log scale, however, forecaster A receives a lower score for large observed values, being more heavily penalised for assigning large probability to small values (which, in relative terms, are far away from the actual observation).
-We note that the chosen example involving a geometric forecast distribution is somewhat constructed; as shown in Section \ref{sec:Hub:cor} and Figure \ref{fig:HUB-cors}A, rankings between models in practice stay quite stable for a single forecast.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width = 1\textwidth]{output/figures/illustration-effect-log-ranking-crps.png}
-\caption{Illustration of the effect of the log-transformation of the ranking for a single forecast.
-Shown are CRPS (or WIS, respectively) values as a function of the observed value for two forecasters.
-Model A issues a geometric distribution (a negative binomial distribution with size parameter $\theta = 1$) with mean $\mu = 10$ and variance $\sigma^2 = \mu + \mu^2 = 110$), while Model B issues a Poisson distribution with mean and variance equal to 10.
-Zeroes in this illustrative example were handled by adding one before applying the natural logarithm.}
-\label{fig:illustration-ranking}
-\end{figure}
-
-Overall model rankings would be expected to differ more when scores are averaged across multiple forecasts or targets.
-The change in rankings of aggregate scores usually is mainly driven by the order of magnitude of scores for different forecast targets across time, location and target type and less so by the kind of changes in model rankings for single forecasts discussed above (see Figure \ref{fig:HUB-cors} for a practical example).
-Large observations will dominate average CRPS values when evaluation is done on the natural scale, but much less so after log transformation.
-Depending on how different models perform across targets of different orders of magnitude, rankings in terms of average scores may change when applying a transformation.
-
-\subsection{Practical considerations and other transformations}
-\label{sec:methods:considerations}
-
-In practice, one issue with the log transform is that it is not readily applicable to negative or zero values, which need to be removed or otherwise handled. 
-One common approach to this end is to add a small positive quantity, such as $a = 1$, to all observations and predictions before taking the logarithm \citep{bellegoDealingLogsZeros2022}.
-This still represents a strictly monotonic transformation, but the choice of $a$ does influence scores and rankings (measures of relative errors shrink the larger the chosen value $a$).
-As a rule of thumb, if if $x > 5a$, the difference between $\log{(x + a)}$ and $\log{(x)}$ is small, and it becomes negligible if $x > 50a$.
-Choosing a suitable offset $a$ thus balances two competing concerns: on the one hand, choosing a small $a$ makes sure that the transformation is as close to a natural logarithm as possible and scores can be interpreted as outlined in the previous sections.
-On the other hand, choosing a larger $a$ can help stabilise scores for forecasts and observations close to zero, avoiding giving excessive weight to forecasts of small quantities.
-For increasing $a$, less relative weight is given to smaller forecast targets.
-For very large values of $a$, $\log(x + a)$ is roughly linear in $x$, so that using a very large $a$ implies similar relative weighting as applying no transformation at all.
-In practice, a user could explore the effect of different values of $a$ graphically and choose $a$ such that the relative weightings of times and regions with high and low incidence correspond to their preferences (see Figure \ref{fig:HUB-log-different-offsets} in our example application, Section \ref{sec:HUB}).
-
-A related issue occurs when the predictive distribution has a large probability mass on zero (or on very small values), as this can translate into an excessively wide forecast in relative terms.
-In our applied example this is illustrated in Figure \ref{fig:HUB-model-comparison-baseline}.
-In such instances, the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm because forecasts contained zero in its prediction intervals.
-To deal with this issue one could choose to use a higher $a$ value when applying a transformation $\log(x + a)$, for example $a = 10$ instead of the $a = 1$ that we chose to use.
-
-A natural question is which other transformations could be applied and whether resulting scores remain (strictly) proper.
-In principle, any transformation function can be applied simultaneously to forecasts and observations as long as the definition of the transformation is independent of the forecasts and any quantities unknown at the time of forecasting, including the observed value.
-This simply corresponds to a re-definition of the forecasting target.
-However, applying non-invertible transformations leads to a loss in information conveyed by forecasts, which we consider undesirable.
-The resulting score will be proper, but it may not be strictly proper anymore (as forecasts differing from the forecaster's true belief on the original scale may be identical on the transformed scale).
-When using the CRPS or the WIS, it seems most appropriate to use only strictly monotonic transformations such as the natural logarithm or the square root as otherwise the encoded notion of distance may become meaningless.
-
-Some other strictly monotonic transformations that can be applied are scaling by the population size or scaling by past observations.
-The latter, as discussed in Section \ref{sec:discussion}, is similar to applying a log-transformation, but corresponds to evaluating a forecast of multiplicative, rather than exponential growth rates.
-The arising issue of dividing by zero can again be solved by adding a small offset $a$.
-Scaling a forecast by the later observed value (as opposed to scaling by past observations) is generally not permissible as it can result in improper scores (see \citealt{lerchForecasterDilemmaExtreme2015} on the closely related topic of weighting scores with a function of the observed value).
-Similarly, scaling forecasts and observations by a function of the predictive distribution (like the predictive mean) may lead to improper scores; however, we are unaware of existing theoretical arguments on this.
-
-When applying a transformation, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee that the score remains proper.
-In the case of log transforms, taking the logarithm of the CRPS values, rather than scoring the log-transformed forecasts and data, results in an improper score.
-We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can be seen that in the example overconfident models perform best in terms of the log WIS.
-We note that strictly speaking, re-scaling average scores by the average score of a baseline model or average scores across different models  to obtain skill scores likewise leads to improper scores \citep{gneitingStrictlyProperScoring2007}.
-The application of such skill scores, however, is established practice and considered largely unproblematic.
-
-We note that in the practical evaluation of operational forecasting systems  several additional challenges arise, which we do not study in detail.
-These concern e.g., the removal of outlying observations and forecasts and the handling of missing forecasts.
-The solutions we employed in practice are provided in Section \ref{sec:HUB-setting}.
-
-\section{Empirical example: the European Forecast Hub}
-\label{sec:HUB}
-
-\subsection{Setting}
-\label{sec:HUB-setting}
-
-As an empirical comparison of evaluating forecasts on the natural and on the log scale, we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}.
-The European COVID-19 Forecast Hub is one of several COVID-19 Forecast Hubs \citep{cramerEvaluationIndividualEnsemble2021, bracherShorttermForecastingCOVID192021} which have been systematically collecting, aggregating and evaluating forecasts of several COVID-19 targets created by different teams every week.
-Forecasts are made one to four weeks ahead into the future and follow a quantile-based format with a set of 23 quantiles ($0.01, 0.025, 0.05, ..., 0.5, ...
-0.95, 0.975, 0.99$).
-
-The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 5th of December 2022 for reported cases and deaths from COVID-19.
-Target dates range from the 13th of March 2021 to the 10th of December 2022, for a total of 92 weeks.
-See \cite{sherrattPredictivePerformanceMultimodel2022} for a more thorough description of the data.
-We filtered all forecasts submitted to the Hub to only include the seven models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 46 forecast dates (see Figure \ref{fig:HUB-num-avail-models}).
-We removed all observations marked as data anomalies by the European Forecast Hub \citep{sherrattPredictivePerformanceMultimodel2022} as well as all remaining negative observed values.
-These anomalies made up a relevant fraction of all observations.
-On average across locations, 12.1 out of 92 (13.2\%) observations were removed for cases and 12.4 out of 92 (13.5\%) for deaths.
-Figure \ref{fig:number-anomalies} displays the number of anomalies removed for each location.
-In addition, we filtered out a small number of erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table \ref{tab:erroneous}.
-Figure \ref{fig:erroneous-forecasts} shows the percentage of forecasts removed for each model.
-Those few (less than 0.2\%  of forecasts for each model) erroneous outlier forecasts had excessive influence on average scores and relative skill scores in a way that was not representative of normal model behaviour.
-We removed them here in order to better illustrate the effects of the log-transformation on scores that one would expect in a well-behaved scenario.
-In a regular forecast evaluation such erroneous forecasts should usually not be removed and would count towards overall model scores.
-
-All predictive quantiles were truncated at 0.
-We applied the log-transformation after adding a constant $a = 1$ to all predictions and observed values.
-The choice of $a = 1$ in part reflects convention, but also represents a suitable choice as it avoids giving excessive weight to forecasts close to zero, while at the same time ensuring that scores for observations $> 5$ can be interpreted reasonably.
-The analysis was conducted in \texttt{R} \citep{R}, using the \texttt{scoringutils} package \citep{bosseEvaluatingForecastsScoringutils2022} for forecast evaluation.
-All code is available on GitHub
-(\url{https://github.com/epiforecasts/transformation-forecast-evaluation}). Where not otherwise stated, we report results for a two-week-ahead forecast horizon.
-
-In addition to the WIS we use pairwise comparisons \citep{cramerEvaluationIndividualEnsemble2021} to evaluate the relative performance of models across countries in the presence of missing forecasts.
-In the first step, score ratios are computed for all pairs of models by taking the set of overlapping forecasts between the two models and dividing the score of one model by the score achieved by the other model.
-The relative skill for a given model compared to others is then obtained by taking the geometric mean of all score ratios which involve that model.
-Low values are better, and the ``average'' model receives a relative skill score of 1.
-
-\subsection{Illustration and qualitative observations}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-model-comparison-ensemble.png}
-    \caption{
-    Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-ensemble made in Germany. Missing values are due to data anomalies that were removed (see section \ref{sec:HUB-setting}. 
-    A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale.
-B, F: Corresponding scores.
-C, G: Forecasts and observations on the log scale.
-D, H: Corresponding scores.
-    }
-    \label{fig:HUB-model-comparison-ensemble}
-\end{figure}
-
-When comparing examples of forecasts on the natural scale with those on the log scale (see Figures \ref{fig:HUB-model-comparison-ensemble}, \ref{fig:HUB-model-comparison-baseline}, \ref{fig:HUB-model-comparison-epinow}) a few interesting patterns emerge.
-Missing the peak, i.e.
-predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts during the peak in May 2022 in Figure \ref{fig:HUB-model-comparison-ensemble}A, B).
-On the log scale, these have less of an influence, as errors are smaller in relative terms (see \ref{fig:HUB-model-comparison-ensemble}C, D).
-Conversely, failure to predict an upswing while numbers are still low, is less severely penalised on the natural scale (see forecasts in July 2021 and to a lesser extent in July 2022 in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low.
-On the log scale, missing lower inflection points tends to lead to more severe penalties (see Figure \ref{fig:HUB-model-comparison-ensemble}C, D).
-One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure \ref{fig:HUB-model-comparison-ensemble}E, F).
-On the log scale, scores do not exhibit this behaviour and rather increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-mean-obs-location.png}
-    \caption{Observations and scores across locations and forecast horizons for the European COVID-19 Forecast Hub data. Locations are sorted according to the mean observed value in that location. 
-    A: Average (across all time points) of observed cases and deaths for different locations.
-B: Corresponding boxplot (y-axis on log-scale) of all cases and deaths.
-C: Scores for two-week-ahead forecasts from the EuroCOVIDhub-ensemble (averaged across all forecast dates) for different locations, evaluated on the natural scale as well as after transforming counts by adding one and applying the natural logarithm.
-D: Corresponding boxplots of all individual scores of the EuroCOVIDhub-ensemble for two-week-ahead predictions.
-E: Boxplots for the relative change of scores for the EuroCOVIDhub-ensemble across forecast horizons.
-For any given forecast date and location, forecasts were made for four different forecast horizons, resulting in four scores.
-All scores were divided by the score for forecast horizon one.
-To enhance interpretability, the range of visible relative changes in scores (relative to horizon = 1) was restricted to [0.1, 10].}
-    \label{fig:HUB-mean-locations}
-\end{figure}
-
-Across the dataset, the average number of observed cases and deaths varied considerably by location and target type (see Figure \ref{fig:HUB-mean-locations}A and B).
-On the natural scale, scores show a pattern quite similar to the observations across targets (see Figure\ref{fig:HUB-mean-locations}D) and locations (see Figure\ref{fig:HUB-mean-locations}C).
-On the log scale, scores were more evenly distributed between targets (see Figure\ref{fig:HUB-mean-locations}D) and locations (see Figure\ref{fig:HUB-mean-locations}C).
-Both on the natural scale as well on the log scale, scores increased considerably with increasing forecast horizon (see Figure \ref{fig:HUB-mean-locations}E).
-This reflects the increasing difficulty of forecasts further into the future and, for the log scale, corresponds with our expectations from Section \ref{sec:methods:growthrate}.
-
-To assess the impact of the choice of offset value $a$ we extend the display from Figure \ref{fig:HUB-mean-locations}C by results obtained under different specifications.
-Results are shown in Figure \ref{fig:HUB-log-different-offsets}, where for completeness we also added the square root transformation.
-As discussed in Section \ref{sec:methods:considerations}, smaller values of $a$ increase the relative weight of smaller locations in the overall evaluation.
-In the most extreme considered case $a = 0.001$, the smallest locations in fact receive the largest weight both for deaths and cases.
-For very large values (see the third row of Figure \ref{fig:HUB-log-different-offsets}), the relative weights strongly resemble those of the evaluation on the natural scale.
-We recommend using displays of this type to get an intuition for the role different locations may play for overall evaluation results.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-scores-locations-log-variants.png}
-    \caption{Mean WIS in different locations for different transformations applied before scoring.
-Locations are sorted according to the mean observed value in that location.
-Shown are scores for two-week-ahead forecasts of the EuroCOVIDhub-ensemble.
-On the natural scale (with no transformation prior to applying the WIS), scores correlate strongly with the average number of observed values in a given location.
-The same is true for scores obtained after applying a square-root transformation, or after applying a log-transformation with a large offset $a$.
-For illustrative purposes, $a$ was chosen to be 101630 for cases and 530 for deaths, 10 times the respective median observed value.
-For large values of $a$, $\log(x + a)$ grows roughly linearly in $x$, meaning that we expect to observe the same patterns as in the case with no transformation.
-For decreasing values of $a$, we give more relative weight to scores in small locations.}
-    \label{fig:HUB-log-different-offsets}
-\end{figure}
-
-\subsection{Regression analysis to determine the variance-stabilizing transformation}
-\label{sec:HUB-regression}
-
-As argued in Section \ref{sec:methods:vst}, the mean-variance, or mean-CRPS, relationship determines which transformation can serve as a VST.
-We can analyse this relationship empirically by running a regression that explains the WIS (which approximates the CRPS) as a function of the central estimate of the predictive distribution.
-We ran the regression
-\begin{linenomath*}
-\begin{equation}
-    \log[\text{WIS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], 
-\end{equation}
-\end{linenomath*}
-where the predictive distribution $F$ and the observation $y$ are on the natural scale. This is equivalent to
-\begin{linenomath*}
-\begin{equation}
-    \text{WIS}(F, y) = \exp(\alpha) \times \text{median}(F)^\beta, 
-\end{equation}
-\end{linenomath*}
-meaning that we estimate a polynomial relationship between the predictive median and achieved WIS.
-Note that we are using predictive medians rather than means as only the former are available in the European COVID-19 Forecast Hub.
-As (under the simplifying assumption of normality; see Section \ref{sec:methods:vst}) the WIS/CRPS of an ideal forecaster scales with the standard deviation, a value of $\beta = 1$ would imply a quadratic median-variance relationship; the natural logarithm could then serve as a VST.
-A value of $\beta = 0.5$ would imply a linear median-variance relationship, suggesting the square root as a VST.
-We applied the regression to case and death forecasts, stratified for one through four-week-ahead forecasts.
-Results are provided in Table \ref{tab:HUB-regression}.
-It can be seen that the estimates of $\beta$ always take a value somewhat below 1, implying a slightly sub-quadratic mean-variance relationship.
-The logarithmic transformation should thus approximately stabilize the variance (and WIS), possibly leading to somewhat higher scores for smaller forecast targets.
-The square-root transformation, on the other hand, can be expected to still lead to higher WIS values for targets of higher orders of magnitude.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-transformation-regression.png}
-    \caption{Relationship between median forecasts and scores.
-Black dots represent WIS values for two-week ahead predictions of the EuroCOVIDhub-ensemble.
-Shown in red are the regression lines discussed in Section \ref{sec:HUB-regression} shown in Table \ref{tab:HUB-regression}.
-A: WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against median predicted values.
-B: Same as A, with scores obtained after applying a square-root-transformation to the data.
-C: Same as A, with scores obtained after applying a log-transformation to the data.}
-    \label{fig:HUB-regression}
-\end{figure}
-
-\begin{table}
-\centering
-\begin{tabular}{ccrrrrrr}
-\toprule
-Horizon & Target & $\alpha$ & $\beta$ & $\alpha_\sqrt{\ }$ & $\beta_\sqrt{\ }$ & $\alpha_\log$ & $\beta_\log$ \\
-\midrule
-% all & all & -1.093 & 0.963 & -0.352 & 0.201 & 0.391 & 0.001\\
-% \addlinespace
-% all & Cases & 0.036 & 0.858 & 0.043 & 0.201 & 0.751 & -0.033\\
-% all & Deaths & -0.884 & 0.868 & 0.273 & 0.121 & 0.436 & -0.023\\
-% \addlinespace
-% 1 & all & -1.402 & 0.923 & 0.320 & 0.088 & 0.318 & -0.014\\
-% 2 & all & -1.221 & 0.967 & 0.112 & 0.164 & 0.364 & -0.003\\
-% 3 & all & -1.001 & 0.984 & -0.094 & 0.241 & 0.410 & 0.008\\
-% 4 & all & -0.757 & 0.986 & 0.000 & 0.299 & 0.469 & 0.015\\
-1 & Cases & -0.862 & 0.876 & 0.790 & 0.087 & 0.433 & -0.024\\
-2 & Cases & -0.243 & 0.877 & 0.959 & 0.162 & 0.660 & -0.031\\
-3 & Cases & 0.372 & 0.855 & 1.109 & 0.238 & 0.882 & -0.037\\
-4 & Cases & 0.816 & 0.837 & 1.645 & 0.296 & 1.009 & -0.036\\
-\addlinespace
-1 & Deaths & -1.146 & 0.832 & 0.457 & 0.048 & 0.376 & -0.035\\
-2 & Deaths & -0.981 & 0.867 & 0.443 & 0.084 & 0.416 & -0.028\\
-3 & Deaths & -0.807 & 0.885 & 0.349 & 0.131 & 0.453 & -0.019\\
-4 & Deaths & -0.602 & 0.891 & 0.125 & 0.194 & 0.501 & -0.011\\
-\bottomrule
-\end{tabular}
-\caption{Coefficients of three regressions for the effect of the magnitude of the median forecast on expected scores. The first regression was 
-$\log[\text{WIS}(F, y)] = \alpha + \beta \times \log[\text{median}(F)], $ where $F$ is the predictive distribution and $y$ the observed value. The second one was 
-$\text{WIS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},$ where $F_{\log}$ is the predictive distribution for $\log y$.
-The third one was $\text{WIS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{sqrt{\ }} + \beta_{sqrt{\ }} \cdot \sqrt{(\text{median}(F))},$ where $F_{\sqrt{\ }}$ is the predictive distribution for $\sqrt{y}$.
-}
-\label{tab:HUB-regression}
-\end{table}
-
-To check the relationship after the transformation, we ran the regressions
-\begin{equation}
-    \text{WIS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},
-\end{equation}
-where $F_{\log}$ is the predictive distribution for $\log(y)$, and
-\begin{equation}
-    \text{WIS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{\sqrt{\ }} + \beta_{\sqrt{\ }} \cdot \sqrt{\text{median}(F)},
-\end{equation} 
-where $F_{\sqrt{\ }}$ is the predictive distribution on the square-root scale.
-A value of $\beta_{\log} = 0$ (or $\beta_{\sqrt{\ }} = 0$, respectively) would imply that scores are linearly independent of the median prediction after the transformation.
-A value smaller (larger) than 0 would imply that smaller (larger) targets lead to higher scores.
-As can be seen from Table \ref{tab:HUB-regression}, the results indeed indicate that small targets lead to larger average WIS when using the log transform ($\beta_{\log} < 0$), while the opposite is true for the square-root transform ($\beta_{\sqrt{\ }} > 0$).
-The results of the three regressions are also displayed in Figure \ref{fig:HUB-regression}.
-In this empirical example, the log transformation thus helps (albeit not perfectly), to stabilise WIS values, and it does so more successfully than the square-root transformation.
-As can be seen from Figure \ref{fig:HUB-regression}, the expected WIS scores for case targets with medians of 10 and 100,000 differ by more then a factor of ten for the square root transformation, but only a factor of around 2 for the logarithm.
-
-\subsection{Impact of logarithmic transformation on model rankings}
-\label{sec:Hub:cor}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-correlations.png}
-    \caption{Correlations of rankings on the natural and logarithmic scale.
-A: Average Spearman rank correlation of scores for individual forecasts.
-For every individual target (defined by a combination of forecast date, target type, horizon, location), one score was obtained per model.
-Then, for every forecast target, the Spearman rank correlation was computed between scores on the natural scale and on the log scale for all the models that had made a forecast for that specific target.
-These individual rank correlations were then averaged across locations and time and are displayed stratified by horizon and target types, representing average accordance of model ranks for a single forecast target on the natural and on the log scale.
-B: Correlation between relative skill scores.
-For every forecast horizon and target type, a separate relative skill score was computed per model using pairwise comparisons, which is a measure of performance of a model relative to the others for a given horizon and target type that accounts for missing values.
-The plot shows the correlation between the relative skill scores on the natural vs.
-on the log scale, representing accordance of overall model performance as judged by scores on the natural and on the log scale.}
-    \label{fig:HUB-cors}
-\end{figure}
-
-For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}A).
-While rankings between forecasters remain similar for a single forecast, this is not true anymore when looking at rankings obtained after averaging scores across multiple forecasts made at different times or in different locations.
-As discussed earlier, scores on the natural and on the log scale penalise errors very differently, e.g.
-when looking at performance during peaks or troughs.
-When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models therefore change considerably (Figure \ref{fig:HUB-cors}B).
-The correlation between relative skill scores also decreases noticeably with increasing forecast horizon.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-pairwise-comparisons.png}
-    \caption{Changes in model ratings as measured by relative skill for two-week-ahead predictions for cases (top row) and deaths (bottom row).
-A: Relative skill scores for case forecasts from different models submitted to the European COVID-19 Forecast Hub computed on the natural scale.
-B: Change in rankings as determined by relative skill scores when moving from an evaluation on the natural scale to one on the logarithmic scale.
-Red arrows indicate that the relative skill score deteriorated when moving from the natural to the log scale, green arrows indicate they improved.
-C: Relative skill scores based on scores on the log scale.
-D: Difference in relative skill scores computed on the natural and on the logarithmic scale, ordered as in C.
-E: Relative contributions of the different WIS components (overprediction, underprediction, and dispersion) to overall model scores on the natural and the logarithmic scale.
-F, G, H, I, J, K: Analogously for deaths.}
-    \label{fig:HUB-rank-order}
-\end{figure}
-
-Figure \ref{fig:HUB-rank-order} shows the changes in the ranking between different forecasting models.
-Encouragingly for the European Forecast Hub, the Hub ensemble, which is the forecast the organisers suggest forecast consumers make use of, remains the top model across scoring schemes.
-For cases, the ILM-EKF model and the Forecast Hub baseline model exhibit the largest change in relative skill scores.
-For the ILM-EKF model the relative proportion of the score that is due to overprediction is reduced when applying a log-transformation before scoring (see Figure \ref{fig:HUB-rank-order}E.
-Instances where the model has overshot are penalised less heavily on the log scale, leading to an overall better score.
-For the Forecast Hub baseline model, the fact that it often puts relevant probability mass on zero (see Figure \ref{fig:HUB-model-comparison-baseline}), leads to worse scores after applying log-transformation due to large dispersion penalties.
-For deaths, the baseline model seems to get similarly penalised for its in relative terms highly dispersed forecasts.
-The performance of other models changes as well, but patterns are less discernible on this aggregate level.
-
-\section{Discussion}
-\label{sec:discussion}
-
-In this paper, we proposed the use of transformations, with a particular focus on the natural logarithmic transformation, when evaluating forecasts in an epidemiological setting.
-These transformations can address issues that arise when evaluating epidemiological forecasts based on measures of absolute error and their probabilistic generalisations (i.e CRPS and WIS).
-We showed that scores obtained after log-transforming both forecasts and observations can be interpreted as a) a measure of relative prediction errors, as well as b) a score for a forecast of the exponential growth rate of the target quantity and c) as variance stabilising transformation in some settings.
-When applying this approach to forecasts from the European COVID-19 Forecast Hub, we found overall scores on the log scale to be more equal across, time, location and target type (cases, deaths) than scores on the natural scale.
-Scores on the log scale were much less influenced by the overall incidence level in a country and showed a slight tendency to be higher in locations with very low incidences.
-We found that model rankings changed noticeably.
-
-On the natural scale, missing the peak and overshooting was more severely penalised than missing the nadir and the following upswing in numbers.
-Both failure modes tended to be more equally penalised on the log scale (with undershooting receiving slightly higher penalties in our example).
-
-Applying a log-transformation prior to the WIS means that forecasts are evaluated in terms of relative errors and errors on the exponential growth rate, rather than absolute errors.
-The most important strength of this approach is that the evaluation better accommodates the exponential nature of the epidemiological process and the types of errors forecasters who accurately model those processes are expected to make.
-The log-transformation also helps avoid issues with scores being strongly influenced by the order of magnitude of the forecast quantity, which can be an issue when evaluating forecasts on the natural scale.
-A potential downside is that forecast evaluation is unreliable in situations where observed values are zero or very small.
-One could argue that this correctly reflect inherent uncertainty about the future course of an epidemic when numbers are small.
-Users nevertheless need to be aware that this can pose issues in practice.
-Including very small values in prediction intervals (see e.g.
-Figure \ref{fig:HUB-model-comparison-baseline}) can lead to excessive dispersion values on the log scale.
-Similarly, locations with lower incidences may get disproportionate weight (i.e.
-high scores) when evaluating forecasts on the log scale.
-\cite{bracherEvaluatingEpidemicForecasts2021} argue that it is desirable to give large weight to forecasts for locations with high incidences, as this reflects performance on the targets we should care about most.
-On the other hand, scoring forecasts on the log scale may be less influenced by outliers and better reflect consistent performance across time, space, and forecast targets.
-Furthermore, decision makers may specifically care about situations in which numbers start to rise from a previously low level.
-
-The log-transformation is only one of many transformations that may be useful and appropriate in an epidemiological context.
-One obvious option is to apply a population standardization to obtain incidence forecasts e.g., per 100,000 population \citep{abbottEvaluatingEpidemiologicallyMotivated2022}.
-We suggested using the natural logarithm as a variance-stabilising transformation (VST).
-This is appropriate for variables that are approximately normally distributed and have a quadratic mean-variance relationship with $\sigma^2 = c \times \mu^2$ (this is e.g.
-approximately true for the negative binomoial distribution and large $\mu$).
-Alternatively, the square-root transformation can be appropriate in the case of a Poisson distributed variable \citep{dunnGeneralizedLinearModels2018}.
-Other VST like the Box-Cox \citep{boxAnalysisTransformations1964} are conceivable as well.
-If one is interested in multiplicative, rather than exponential growth rates, one could, instead of applying a log transformation, convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last value that was observed at the time the forecast was made.
-Forecasters would then implicitly predict a separate multiplicative growth rate from today to horizon 1, 2, etc.
-Instead of dividing by the last observed value, another promising transformation would be to divide each forecast by the forecast of the previous week (and analogously for observations), in order to obtain forecasts for week-to-week growth rates.
-Alternatively, one could also take first differences of values on the log scale.
-This approach would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluating the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}.
-Dividing values by the previous value, unfortunately, is not feasible under the current quantile-based format of the Forecast Hubs, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate.
-However, it may be an interesting approach if predictive samples are available or if quantiles for weekwise growth rates have been collected.
-Potentially, the variance stabilising time-series forecasting literature may be a useful source of other transformations for various forecast settings.
-
-It is possible to go beyond choosing a single transformation by constructing composite scores as a weighted sum of scores based on different transformations.
-This would make it possible to create custom scores and allow forecast consumers to choose and assign explicit weights to different qualities of the forecasts they might care about.
-
-Exploring transformations is a promising avenue for future work that could help bridge the gap between modellers and policymakers by providing scoring rules that better reflect what forecast consumers care about.
-In this paper, we did not make any particular assumptions about policy makers' priorities and preferences.
-Rather, we aimed to enable users to make an informed choice by showing how different transformations lead to different relative weights for the kinds of prediction errors forecast consumers may care about, such as absolute vs.
-relative errors or the size of penalties for over- vs.
-underprediction.
-In practice, engagement with decision makers is important to determine what their priorities are and how different ways to measure predictive importance should be weighed.
-
-We have shown that the natural logarithm transformation can lead to significant changes in the relative rankings of models against each other, with potentially important implications for decision-makers who rely on the knowledge of past performance to make a judgement about which forecasts should inform future decisions.
-While it is commonly accepted that multiple proper scoring rules should usually be considered when comparing forecasts, we think this should be supplemented by considering different transformations of the data to obtain a richer picture of model performance.
-More work needs to be done to better understand the effects of applying transformations in different contexts, and how they may impact decision-making.
-
-\newpage
-
 \appendix
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/example-log-first.png}
+    \includegraphics[width=0.99\textwidth]{../output/figures/example-log-first.png}
     \caption{Illustration of the effect of applying a transformation after scoring.
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/illustration-effect-offset-log.png}
+    \includegraphics[width=0.99\textwidth]{../output/figures/illustration-effect-offset-log.png}
     \caption{Illustration of the effect of adding a small quantity to a value before taking the natural logarithm.
@@
 \centering
-\includegraphics[width = 1\textwidth]{output/figures/SIM-score-approximation.png}
+\includegraphics[width = 1\textwidth]{../output/figures/SIM-score-approximation.png}
 \caption{
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/number-avail-forecasts.png}
+    \includegraphics[width=0.99\textwidth]{../output/figures/number-avail-forecasts.png}
     \caption{
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/number-anomalies.png}
+    \includegraphics[width=0.99\textwidth]{../output/figures/number-anomalies.png}
     \caption{
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/erroneous-forecasts.png}
+    \includegraphics[width=0.99\textwidth]{../output/figures/erroneous-forecasts.png}
     \caption{
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-model-comparison-baseline.png}
+    \includegraphics[width=0.99\textwidth]{../output/figures/HUB-model-comparison-baseline.png}
     \caption{
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{output/figures/HUB-model-comparison-epinow.png}
+    \includegraphics[width=0.99\textwidth]{../output/figures/HUB-model-comparison-epinow.png}
     \caption{
@@
 
-\clearpage
-\bibliography{log-or-not-v2.bib, log-or-not.bib, software.bib}
-
 \end{document}

### 2024-02-20 final submission after reformatting/Supplementary-information.tex
update SI PDF by removing links to main manuscript

@@
 It can be seen that for $a$ values smaller than one fifth of the transformed quantity, the effect of adding an offset is generally small.
-When choosing a suitable $a$, the trade-off is between staying close to the interpretation of a pure log-transformation (choosing a small $a$) and not giving excessive weights to small observations (by choosing a larger $a$, see Figure \ref{fig:HUB-log-different-offsets}).}
+When choosing a suitable $a$, the trade-off is between staying close to the interpretation of a pure log-transformation (choosing a small $a$) and not giving excessive weights to small observations (by choosing a larger $a$, see Figure 6).}
     \label{fig:illustration-effect-log-offset}
@@
 \caption{
-Visualisation of expected CRPS values against approximated scores using the approximation detailed in Section \ref{sec:methods:rankings} (see also Figure \ref{fig:SIM-wis-state-size-mean}).
+This is using the approximation detailed in the theoretical discussion on model rankings (see also Fig 2).
 Expected CRPS scores are shown for three different distributions once on the natural scale (top row) and once scored on the log scale (bottom row).
